---
title: "Cluster_Waveform_Visualizations"
author: "Sarah Moseley"
date: "2026-05-24"
output: html_document
---

## Setup

```{r setup, message=FALSE, warning=FALSE}

library(dplyr)
library(readr)
library(purrr)
library(stringr)
library(tibble)
library(ggplot2)

if (!requireNamespace("PamBinaries", quietly = TRUE)) {
  install.packages("PamBinaries")
}

library(PamBinaries)

set.seed(123)

# ------------------------------------------------------------
# Main analysis folder
# ------------------------------------------------------------

base_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1"

audit_dir <- file.path(
  base_dir,
  "temporal_visualizations/bearing_source_audit"
)

waveform_dir <- file.path(
  base_dir,
  "plots/waveform_cluster_grids"
)

dir.create(waveform_dir, recursive = TRUE, showWarnings = FALSE)

# ------------------------------------------------------------
# Existing files created during the bearing workflow
# ------------------------------------------------------------
# These files come from the previous workflow where sampled HDBSCAN clicks
# were joined to PAMGuard OfflineClicks metadata and matched to binary files.

sampled_joined_file <- file.path(
  audit_dir,
  "sampled_hdbscan_clicks_joined_to_offline_clicks.csv"
)

manifest_file <- file.path(
  audit_dir,
  "PAMGuard_binary_manifest_for_bearing_export_CLEAN.csv"
)

# ------------------------------------------------------------
# Cluster sets for this visualization
# ------------------------------------------------------------

# Individual 5x5 panels will be saved for all clusters,
# including cluster 0/noise.
individual_grid_clusters <- as.character(0:9)

# Composite 3x3 figure will include only the nine assigned clusters.
composite_grid_clusters <- as.character(1:9)

# ------------------------------------------------------------
# Cluster colors used throughout the clustering workflow
# ------------------------------------------------------------

cluster_colors <- c(
  "0" = "grey80",
  "1" = "#E41A1C",
  "2" = "#FF7F00",
  "3" = "#FFD92F",
  "4" = "#4DAF4A",
  "5" = "#00BFC4",
  "6" = "#377EB8",
  "7" = "#984EA3",
  "8" = "#F781BF",
  "9" = "#A65628"
)

```

## Load and verify input files

```{r load-inputs, message=FALSE, warning=FALSE}

# ------------------------------------------------------------
# Confirm that the required input files exist
# ------------------------------------------------------------

required_files <- c(
  sampled_joined_file,
  manifest_file
)

missing_files <- required_files[!file.exists(required_files)]

if (length(missing_files) > 0) {
  stop(
    "These required files are missing:\n",
    paste(missing_files, collapse = "\n")
  )
}

cat("All required input files found.\n\n")

# ------------------------------------------------------------
# Load sampled HDBSCAN clicks joined to OfflineClicks metadata
# ------------------------------------------------------------

sampled_clicks <- read_csv(
  sampled_joined_file,
  show_col_types = FALSE
) %>%
  mutate(
    UID = as.character(UID),
    BinaryFile = as.character(BinaryFile),
    DriftID_clean = as.character(DriftID_clean),
    HDBSCAN_cluster = as.character(HDBSCAN_cluster)
  )

# ------------------------------------------------------------
# Load clean PAMGuard binary manifest
# ------------------------------------------------------------

manifest_clean <- read_csv(
  manifest_file,
  show_col_types = FALSE
) %>%
  mutate(
    DriftID_clean = as.character(DriftID_clean),
    BinaryFile = as.character(BinaryFile),
    full_path = as.character(full_path),
    pgdx_full_path = as.character(pgdx_full_path)
  )

# ------------------------------------------------------------
# Check required columns
# ------------------------------------------------------------

required_sampled_cols <- c(
  "UID",
  "EventID",
  "DriftID_clean",
  "BinaryFile",
  "HDBSCAN_cluster"
)

required_manifest_cols <- c(
  "DriftID_clean",
  "BinaryFile",
  "full_path",
  "pgdx_full_path"
)

missing_sampled_cols <- setdiff(required_sampled_cols, names(sampled_clicks))
missing_manifest_cols <- setdiff(required_manifest_cols, names(manifest_clean))

if (length(missing_sampled_cols) > 0) {
  stop(
    "sampled_clicks is missing these required columns:\n",
    paste(missing_sampled_cols, collapse = "\n")
  )
}

if (length(missing_manifest_cols) > 0) {
  stop(
    "manifest_clean is missing these required columns:\n",
    paste(missing_manifest_cols, collapse = "\n")
  )
}

cat("All required columns found.\n\n")

# ------------------------------------------------------------
# Basic summaries
# ------------------------------------------------------------

cat("Sampled clicks loaded:", nrow(sampled_clicks), "\n")
cat("Unique events:", n_distinct(sampled_clicks$EventID), "\n")
cat("Unique drifts:", n_distinct(sampled_clicks$DriftID_clean), "\n")
cat("Unique binary files represented in sampled clicks:", n_distinct(sampled_clicks$BinaryFile), "\n\n")

cat("Clean binary manifest rows:", nrow(manifest_clean), "\n")
cat("Unique manifest binary files:", n_distinct(manifest_clean$BinaryFile), "\n\n")

cat("Clicks per HDBSCAN cluster:\n")

sampled_clicks %>%
  count(HDBSCAN_cluster, name = "n_clicks") %>%
  arrange(as.numeric(HDBSCAN_cluster)) %>%
  print(n = Inf)

# ------------------------------------------------------------
# Check that all intended clusters are present
# ------------------------------------------------------------

observed_clusters <- sort(unique(sampled_clicks$HDBSCAN_cluster))
missing_expected_clusters <- setdiff(individual_grid_clusters, observed_clusters)

if (length(missing_expected_clusters) > 0) {
  warning(
    "These expected clusters are missing from sampled_clicks: ",
    paste(missing_expected_clusters, collapse = ", ")
  )
} else {
  cat("\nAll expected clusters 0-9 are present.\n")
}

```

## Smoke test one PAMGuard binary for waveform data

```{r smoke-test-one-binary-waveforms, message=FALSE, warning=FALSE}

# ------------------------------------------------------------
# Pick one binary file that actually contains sampled clicks
# ------------------------------------------------------------

binary_click_counts <- sampled_clicks %>%
  filter(!is.na(BinaryFile)) %>%
  count(DriftID_clean, BinaryFile, name = "n_sampled_clicks") %>%
  left_join(
    manifest_clean,
    by = c("DriftID_clean", "BinaryFile")
  ) %>%
  arrange(desc(n_sampled_clicks))

cat("Binary files available for waveform smoke test:\n")
print(binary_click_counts %>% head(10))

# Use the binary file with the most sampled clicks
test_binary <- binary_click_counts %>%
  filter(!is.na(full_path)) %>%
  slice(1)

test_pgdf <- test_binary$full_path[1]

cat("\nTesting PAMGuard binary file:\n")
cat(test_pgdf, "\n\n")

cat("This file contains", test_binary$n_sampled_clicks[1], "sampled clicks from the current analysis.\n\n")

# ------------------------------------------------------------
# Load the PAMGuard binary WITH large objects included
# ------------------------------------------------------------
# Important:
# skipLarge = FALSE is the key change from the bearing workflow.
# This may take longer than the angle extraction workflow.

pg_test <- PamBinaries::loadPamguardBinaryFile(
  fileName = test_pgdf,
  skipLarge = FALSE,
  convertDate = TRUE
)

cat("Binary loaded successfully.\n\n")

# ------------------------------------------------------------
# Basic object inspection
# ------------------------------------------------------------

cat("Class of loaded object:\n")
print(class(pg_test))

cat("\nTop-level names in loaded object:\n")
print(names(pg_test))

cat("\nStructure preview of loaded object:\n")
str(pg_test, max.level = 2)

# ------------------------------------------------------------
# Try converting to a data frame, as in the bearing workflow
# ------------------------------------------------------------

pg_test_df <- as.data.frame(pg_test) %>%
  as_tibble()

cat("\nRows in converted data frame:", nrow(pg_test_df), "\n")
cat("Columns in converted data frame:", ncol(pg_test_df), "\n\n")

cat("Column names:\n")
print(names(pg_test_df))

# ------------------------------------------------------------
# Identify columns that may contain waveform/sample data
# ------------------------------------------------------------

candidate_waveform_cols <- names(pg_test_df)[
  str_detect(
    names(pg_test_df),
    regex("wave|waveform|raw|sample|samples|signal|data", ignore_case = TRUE)
  )
]

cat("\nCandidate waveform-like columns based on names:\n")
print(candidate_waveform_cols)

# ------------------------------------------------------------
# Identify list columns, because waveform data is often stored as a list column
# ------------------------------------------------------------

list_cols <- names(pg_test_df)[
  map_lgl(pg_test_df, is.list)
]

cat("\nList columns in converted data frame:\n")
print(list_cols)

# ------------------------------------------------------------
# Summarize candidate/list columns so we can see what they contain
# ------------------------------------------------------------

cols_to_inspect <- unique(c(candidate_waveform_cols, list_cols))

if (length(cols_to_inspect) > 0) {
  
  cat("\nPreview of candidate waveform/list columns:\n")
  
  for (col in cols_to_inspect) {
    
    cat("\n----------------------------\n")
    cat("Column:", col, "\n")
    cat("Class:\n")
    print(class(pg_test_df[[col]]))
    
    first_non_null_index <- which(!map_lgl(pg_test_df[[col]], is.null))[1]
    
    if (!is.na(first_non_null_index)) {
      cat("First non-null row index:", first_non_null_index, "\n")
      cat("Structure of first non-null value:\n")
      str(pg_test_df[[col]][[first_non_null_index]], max.level = 2)
    } else {
      cat("No non-null values found in this column.\n")
    }
  }
  
} else {
  
  cat("\nNo obvious waveform/list columns found in the converted data frame.\n")
  cat("We may need to inspect the raw pg_test object directly.\n")
}

# ------------------------------------------------------------
# Inspect a row corresponding to one sampled click in this binary
# ------------------------------------------------------------

sampled_uids_in_test_binary <- sampled_clicks %>%
  filter(
    DriftID_clean == test_binary$DriftID_clean[1],
    BinaryFile == test_binary$BinaryFile[1]
  ) %>%
  distinct(UID, HDBSCAN_cluster, EventID) %>%
  arrange(HDBSCAN_cluster, UID)

cat("\nSampled UIDs from this binary:\n")
print(sampled_uids_in_test_binary %>% head(10))

if ("UID" %in% names(pg_test_df)) {
  
  pg_test_df <- pg_test_df %>%
    mutate(UID = as.character(UID))
  
  matched_rows <- pg_test_df %>%
    semi_join(sampled_uids_in_test_binary, by = "UID")
  
  cat("\nRows in this binary matching sampled clicks:", nrow(matched_rows), "\n")
  
  if (nrow(matched_rows) > 0) {
    cat("\nPreview of matched rows, excluding any very large columns:\n")
    
    matched_rows %>%
      select(-any_of(list_cols)) %>%
      head(10) %>%
      print(width = Inf)
  }
  
} else {
  cat("\nNo UID column found in pg_test_df. We need to inspect the object structure more carefully.\n")
}

```

## Inspect raw PAMGuard click objects for waveform fields

```{r inspect-raw-click-waveform-fields, message=FALSE, warning=FALSE}

# ------------------------------------------------------------
# Check arguments accepted by loadPamguardBinaryFile()
# ------------------------------------------------------------

cat("Arguments for PamBinaries::loadPamguardBinaryFile:\n")
print(args(PamBinaries::loadPamguardBinaryFile))

loader_arg_names <- names(formals(PamBinaries::loadPamguardBinaryFile))

cat("\nFormal argument names:\n")
print(loader_arg_names)

# ------------------------------------------------------------
# Inspect one raw click object from pg_test$data
# ------------------------------------------------------------

cat("\nNumber of raw click objects in pg_test$data:", length(pg_test$data), "\n\n")

cat("Names of first few raw click objects:\n")
print(head(names(pg_test$data), 10))

first_click <- pg_test$data[[1]]

cat("\nClass of first raw click object:\n")
print(class(first_click))

cat("\nNames inside first raw click object:\n")
print(names(first_click))

cat("\nStructure of first raw click object:\n")
str(first_click, max.level = 3)

# ------------------------------------------------------------
# Inspect one raw click that matches our sampled-click UIDs
# ------------------------------------------------------------

matched_uid <- sampled_uids_in_test_binary$UID[1]

cat("\nTrying to inspect sampled UID:", matched_uid, "\n")

if (matched_uid %in% names(pg_test$data)) {
  
  matched_click_raw <- pg_test$data[[matched_uid]]
  
  cat("\nMatched UID found directly in pg_test$data names.\n")
  cat("\nNames inside matched raw click object:\n")
  print(names(matched_click_raw))
  
  cat("\nStructure of matched raw click object:\n")
  str(matched_click_raw, max.level = 3)
  
} else {
  
  cat("\nMatched UID was not found directly in pg_test$data names.\n")
  cat("Trying to locate it by UID field inside raw objects...\n")
  
  raw_uids <- map_chr(
    pg_test$data,
    ~ as.character(.x$UID %||% NA_character_)
  )
  
  matched_index <- which(raw_uids == matched_uid)[1]
  
  if (!is.na(matched_index)) {
    matched_click_raw <- pg_test$data[[matched_index]]
    
    cat("\nMatched UID found at raw object index:", matched_index, "\n")
    cat("\nNames inside matched raw click object:\n")
    print(names(matched_click_raw))
    
    cat("\nStructure of matched raw click object:\n")
    str(matched_click_raw, max.level = 3)
    
  } else {
    cat("\nCould not locate matched UID in raw objects.\n")
  }
}

# ------------------------------------------------------------
# Search raw click fields for likely waveform/sample names
# ------------------------------------------------------------

raw_field_names <- names(first_click)

possible_waveform_fields <- raw_field_names[
  str_detect(
    raw_field_names,
    regex("wave|waveform|raw|sample|samples|signal|data", ignore_case = TRUE)
  )
]

cat("\nPossible waveform-related fields inside raw click object:\n")
print(possible_waveform_fields)

# ------------------------------------------------------------
# If waveform-specific loader arguments exist, try loading with them
# ------------------------------------------------------------

if ("getWave" %in% loader_arg_names) {
  
  cat("\nThe loader supports getWave. Trying getWave = TRUE...\n")
  
  wave_args <- list(
    fileName = test_pgdf,
    skipLarge = FALSE,
    convertDate = TRUE,
    getWave = TRUE
  )
  
  if ("onlyWave" %in% loader_arg_names) {
    wave_args$onlyWave <- FALSE
  }
  
  pg_test_wave <- do.call(
    PamBinaries::loadPamguardBinaryFile,
    wave_args
  )
  
  cat("\nWave-enabled binary loaded successfully.\n")
  
  first_click_wave <- pg_test_wave$data[[1]]
  
  cat("\nNames inside first wave-enabled raw click object:\n")
  print(names(first_click_wave))
  
  cat("\nStructure of first wave-enabled raw click object:\n")
  str(first_click_wave, max.level = 3)
  
  wave_df <- as.data.frame(pg_test_wave) %>%
    as_tibble()
  
  cat("\nColumns after wave-enabled load:\n")
  print(names(wave_df))
  
  wave_list_cols <- names(wave_df)[map_lgl(wave_df, is.list)]
  
  cat("\nList columns after wave-enabled load:\n")
  print(wave_list_cols)
  
} else {
  
  cat("\nThe loader does not appear to expose getWave as a formal argument.\n")
  cat("We will need to inspect the raw object fields or PamBinaries documentation/source behavior.\n")
}

```

## Extract and plot one sampled waveform using channel 1

```{r extract-and-plot-one-waveform-channel-1, message=FALSE, warning=FALSE}

# ------------------------------------------------------------
# Channel choice
# ------------------------------------------------------------
# We will use the first waveform column for every click.
# In the PAMGuard raw click object, wave is samples x channels.

selected_wave_channel <- 1

# ------------------------------------------------------------
# Helper: extract one waveform from a raw PAMGuard click object
# ------------------------------------------------------------

extract_one_waveform <- function(raw_click, channel = selected_wave_channel, normalize = TRUE) {
  
  wave_matrix <- raw_click$wave
  
  if (is.null(wave_matrix)) {
    stop("This raw click does not contain a wave field.")
  }
  
  if (!is.matrix(wave_matrix)) {
    wave_matrix <- as.matrix(wave_matrix)
  }
  
  # PAMGuard wave is samples x channels
  n_samples <- nrow(wave_matrix)
  n_channels <- ncol(wave_matrix)
  
  if (channel > n_channels) {
    stop(
      "Requested channel ",
      channel,
      ", but this click only has ",
      n_channels,
      " channel(s)."
    )
  }
  
  waveform <- wave_matrix[, channel]
  
  # Normalize each click to compare waveform shape rather than amplitude
  if (normalize) {
    max_abs_amp <- max(abs(waveform), na.rm = TRUE)
    
    if (is.finite(max_abs_amp) && max_abs_amp > 0) {
      waveform <- waveform / max_abs_amp
    }
  }
  
  tibble(
    sample_index = seq_len(n_samples),
    amplitude = as.numeric(waveform),
    wave_channel_used = channel,
    n_channels = n_channels,
    n_samples = n_samples,
    UID = as.character(raw_click$UID),
    sampleDuration = as.integer(raw_click$sampleDuration),
    maxAmplitude = as.numeric(raw_click$maxAmplitude)
  )
}

# ------------------------------------------------------------
# Use the matched sampled click we inspected above
# ------------------------------------------------------------

one_waveform <- extract_one_waveform(
  raw_click = matched_click_raw,
  channel = selected_wave_channel,
  normalize = TRUE
)

cat("Extracted waveform summary:\n")

one_waveform %>%
  summarise(
    UID = first(UID),
    n_samples = first(n_samples),
    n_channels = first(n_channels),
    wave_channel_used = first(wave_channel_used),
    min_amplitude = min(amplitude, na.rm = TRUE),
    max_amplitude = max(amplitude, na.rm = TRUE),
    sampleDuration = first(sampleDuration),
    maxAmplitude = first(maxAmplitude)
  ) %>%
  print(width = Inf)

# ------------------------------------------------------------
# Plot one waveform
# ------------------------------------------------------------

p_one_waveform <- ggplot(
  one_waveform,
  aes(x = sample_index, y = amplitude)
) +
  geom_line(linewidth = 0.35) +
  labs(
    title = paste0("Example Waveform: UID ", unique(one_waveform$UID)),
    subtitle = paste0(
      "Waveform channel = ",
      unique(one_waveform$wave_channel_used),
      "; normalized amplitude"
    ),
    x = "Sample index",
    y = "Normalized amplitude"
  ) +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.minor = element_blank()
  )

print(p_one_waveform)

ggsave(
  filename = file.path(
    waveform_dir,
    paste0(
      "example_waveform_UID_",
      unique(one_waveform$UID),
      "_channel_",
      selected_wave_channel,
      ".png"
    )
  ),
  plot = p_one_waveform,
  width = 7,
  height = 4,
  dpi = 300
)

```

## Randomly sample 25 clicks per cluster

```{r sample-clicks-per-cluster, message=FALSE, warning=FALSE}

# ------------------------------------------------------------
# Sampling settings
# ------------------------------------------------------------

waveform_sample_n <- 25

set.seed(123)

# ------------------------------------------------------------
# Check that each cluster has at least 25 clicks available
# ------------------------------------------------------------

cluster_availability <- sampled_clicks %>%
  filter(HDBSCAN_cluster %in% individual_grid_clusters) %>%
  count(HDBSCAN_cluster, name = "n_available_clicks") %>%
  arrange(as.numeric(HDBSCAN_cluster))

cat("Available clicks per cluster:\n")
print(cluster_availability, n = Inf)

clusters_with_too_few_clicks <- cluster_availability %>%
  filter(n_available_clicks < waveform_sample_n)

if (nrow(clusters_with_too_few_clicks) > 0) {
  stop(
    "At least one cluster has fewer than ",
    waveform_sample_n,
    " clicks. Check clusters:\n",
    paste(clusters_with_too_few_clicks$HDBSCAN_cluster, collapse = ", ")
  )
}

cat("\nAll clusters have at least", waveform_sample_n, "clicks available.\n\n")

# ------------------------------------------------------------
# Randomly sample 25 clicks per cluster
# ------------------------------------------------------------

waveform_sample_keys <- sampled_clicks %>%
  filter(
    HDBSCAN_cluster %in% individual_grid_clusters,
    !is.na(UID),
    !is.na(BinaryFile),
    !is.na(DriftID_clean)
  ) %>%
  group_by(HDBSCAN_cluster) %>%
  slice_sample(n = waveform_sample_n) %>%
  ungroup() %>%
  mutate(
    HDBSCAN_cluster = factor(
      HDBSCAN_cluster,
      levels = individual_grid_clusters
    )
  ) %>%
  arrange(HDBSCAN_cluster, DriftID_clean, BinaryFile, UID)

# ------------------------------------------------------------
# Attach PAMGuard binary paths from clean manifest
# ------------------------------------------------------------

waveform_sample_manifest <- waveform_sample_keys %>%
  left_join(
    manifest_clean %>%
      select(
        DriftID_clean,
        BinaryFile,
        full_path,
        pgdx_full_path
      ),
    by = c("DriftID_clean", "BinaryFile")
  )

# ------------------------------------------------------------
# Verify binary paths were found
# ------------------------------------------------------------

missing_binary_paths <- waveform_sample_manifest %>%
  filter(is.na(full_path) | !file.exists(full_path))

if (nrow(missing_binary_paths) > 0) {
  cat("Rows with missing binary paths:\n")
  print(
    missing_binary_paths %>%
      select(HDBSCAN_cluster, EventID, DriftID_clean, BinaryFile, UID, full_path),
    n = Inf
  )
  
  stop("Some sampled clicks do not have valid PAMGuard binary paths.")
}

cat("All sampled clicks have valid PAMGuard binary paths.\n\n")

# ------------------------------------------------------------
# Summaries
# ------------------------------------------------------------

cat("Sampled clicks per cluster:\n")
waveform_sample_manifest %>%
  count(HDBSCAN_cluster, name = "n_sampled_clicks") %>%
  arrange(as.numeric(as.character(HDBSCAN_cluster))) %>%
  print(n = Inf)

cat("\nNumber of unique binary files needed:\n")
waveform_sample_manifest %>%
  summarise(
    n_sampled_clicks = n(),
    n_unique_clusters = n_distinct(HDBSCAN_cluster),
    n_unique_events = n_distinct(EventID),
    n_unique_drifts = n_distinct(DriftID_clean),
    n_unique_binary_files = n_distinct(paste(DriftID_clean, BinaryFile))
  ) %>%
  print(width = Inf)

cat("\nSampled clicks by cluster and binary file:\n")
waveform_sample_manifest %>%
  count(HDBSCAN_cluster, DriftID_clean, BinaryFile, name = "n_clicks") %>%
  arrange(as.numeric(as.character(HDBSCAN_cluster)), DriftID_clean, BinaryFile) %>%
  print(n = Inf)

# ------------------------------------------------------------
# Save sampled click table for reproducibility
# ------------------------------------------------------------

waveform_sample_file <- file.path(
  waveform_dir,
  paste0("waveform_sample_25_clicks_per_cluster_channel_", selected_wave_channel, ".csv")
)

write_csv(
  waveform_sample_manifest,
  waveform_sample_file
)

cat("\nSaved waveform sample table to:\n")
cat(waveform_sample_file, "\n")

```

## Extract waveforms for sampled clicks

```{r extract-sampled-waveforms, message=FALSE, warning=FALSE}

# ------------------------------------------------------------
# Helper: safely read waveforms from one PAMGuard binary file
# ------------------------------------------------------------

read_sampled_waveforms_from_binary <- function(binary_sample_df) {
  
  this_drift <- unique(binary_sample_df$DriftID_clean)
  this_binary <- unique(binary_sample_df$BinaryFile)
  this_path <- unique(binary_sample_df$full_path)
  
  if (length(this_drift) != 1 || length(this_binary) != 1 || length(this_path) != 1) {
    stop("Each binary_sample_df must correspond to exactly one DriftID_clean + BinaryFile + full_path.")
  }
  
  needed_uids_chr <- unique(as.character(binary_sample_df$UID))
  needed_uids_num <- suppressWarnings(as.numeric(needed_uids_chr))
  
  cat("Reading:", this_drift, "|", this_binary, "|", length(needed_uids_chr), "sampled UID(s)\n")
  
  out <- tryCatch({
    
    pg <- PamBinaries::loadPamguardBinaryFile(
      fileName = this_path,
      skipLarge = FALSE,
      skipData = FALSE,
      keepUIDs = needed_uids_num,
      convertDate = TRUE
    )
    
    available_uids <- names(pg$data)
    
    map_dfr(seq_len(nrow(binary_sample_df)), function(i) {
      
      click_meta <- binary_sample_df[i, ]
      uid_chr <- as.character(click_meta$UID)
      
      if (!uid_chr %in% available_uids) {
        warning(
          "UID not found in loaded binary: ",
          uid_chr,
          " | ",
          this_binary
        )
        
        return(tibble())
      }
      
      raw_click <- pg$data[[uid_chr]]
      
      one_wave <- extract_one_waveform(
        raw_click = raw_click,
        channel = selected_wave_channel,
        normalize = TRUE
      )
      
      one_wave %>%
        mutate(
          HDBSCAN_cluster = as.character(click_meta$HDBSCAN_cluster),
          EventID = as.character(click_meta$EventID),
          DriftID_clean = as.character(click_meta$DriftID_clean),
          BinaryFile = as.character(click_meta$BinaryFile),
          waveform_panel_id = paste0(
            "Cluster ",
            as.character(click_meta$HDBSCAN_cluster),
            " | UID ",
            uid_chr
          )
        )
    })
    
  }, error = function(e) {
    
    warning(
      "Failed to read binary: ",
      this_path,
      "\nError: ",
      e$message
    )
    
    tibble()
  })
  
  return(out)
}

# ------------------------------------------------------------
# Split sampled clicks by binary file
# ------------------------------------------------------------

waveform_binary_groups <- waveform_sample_manifest %>%
  mutate(
    UID = as.character(UID),
    HDBSCAN_cluster = as.character(HDBSCAN_cluster)
  ) %>%
  group_by(DriftID_clean, BinaryFile, full_path) %>%
  group_split()

cat("Number of binary files to read:", length(waveform_binary_groups), "\n\n")

# ------------------------------------------------------------
# Extract all sampled waveforms
# ------------------------------------------------------------

sampled_waveforms_long <- map_dfr(
  waveform_binary_groups,
  read_sampled_waveforms_from_binary
)

cat("\nFinished waveform extraction.\n\n")

# ------------------------------------------------------------
# Verify extraction success
# ------------------------------------------------------------

waveform_extraction_summary <- sampled_waveforms_long %>%
  distinct(
    HDBSCAN_cluster,
    EventID,
    DriftID_clean,
    BinaryFile,
    UID,
    wave_channel_used,
    n_channels,
    n_samples
  ) %>%
  summarise(
    n_waveforms_extracted = n(),
    n_unique_clusters = n_distinct(HDBSCAN_cluster),
    n_unique_events = n_distinct(EventID),
    n_unique_drifts = n_distinct(DriftID_clean),
    min_n_samples = min(n_samples, na.rm = TRUE),
    median_n_samples = median(n_samples, na.rm = TRUE),
    max_n_samples = max(n_samples, na.rm = TRUE),
    wave_channel_used = paste(sort(unique(wave_channel_used)), collapse = ", ")
  )

cat("Waveform extraction summary:\n")
print(waveform_extraction_summary, width = Inf)

cat("\nExtracted waveforms per cluster:\n")

sampled_waveforms_long %>%
  distinct(HDBSCAN_cluster, UID) %>%
  count(HDBSCAN_cluster, name = "n_waveforms_extracted") %>%
  arrange(as.numeric(HDBSCAN_cluster)) %>%
  print(n = Inf)

# ------------------------------------------------------------
# Check for missing sampled clicks
# ------------------------------------------------------------

expected_waveform_keys <- waveform_sample_manifest %>%
  transmute(
    HDBSCAN_cluster = as.character(HDBSCAN_cluster),
    DriftID_clean = as.character(DriftID_clean),
    BinaryFile = as.character(BinaryFile),
    UID = as.character(UID)
  ) %>%
  distinct()

extracted_waveform_keys <- sampled_waveforms_long %>%
  transmute(
    HDBSCAN_cluster = as.character(HDBSCAN_cluster),
    DriftID_clean = as.character(DriftID_clean),
    BinaryFile = as.character(BinaryFile),
    UID = as.character(UID)
  ) %>%
  distinct()

missing_waveform_keys <- expected_waveform_keys %>%
  anti_join(
    extracted_waveform_keys,
    by = c("HDBSCAN_cluster", "DriftID_clean", "BinaryFile", "UID")
  )

if (nrow(missing_waveform_keys) > 0) {
  cat("\nWARNING: Some sampled clicks were not extracted:\n")
  print(missing_waveform_keys, n = Inf)
} else {
  cat("\nAll sampled clicks were successfully extracted.\n")
}

# ------------------------------------------------------------
# Save extracted waveform data
# ------------------------------------------------------------

waveform_long_file <- file.path(
  waveform_dir,
  paste0("sampled_waveforms_long_25_per_cluster_channel_", selected_wave_channel, ".rds")
)

saveRDS(
  sampled_waveforms_long,
  waveform_long_file
)

cat("\nSaved extracted waveform data to:\n")
cat(waveform_long_file, "\n")

```

## Save individual 5x5 waveform grids for each cluster

```{r save-individual-cluster-waveform-grids, message=FALSE, warning=FALSE}

# ------------------------------------------------------------
# If needed, reload extracted waveform data
# ------------------------------------------------------------
# This lets the chunk run even if you restart R after extraction.

if (!exists("sampled_waveforms_long")) {
  
  waveform_long_file <- file.path(
    waveform_dir,
    paste0("sampled_waveforms_long_25_per_cluster_channel_", selected_wave_channel, ".rds")
  )
  
  sampled_waveforms_long <- readRDS(waveform_long_file)
}

# ------------------------------------------------------------
# Build stable 1-25 panel IDs within each cluster
# ------------------------------------------------------------

waveform_panel_lookup <- sampled_waveforms_long %>%
  distinct(
    HDBSCAN_cluster,
    EventID,
    DriftID_clean,
    BinaryFile,
    UID
  ) %>%
  arrange(
    as.numeric(HDBSCAN_cluster),
    EventID,
    DriftID_clean,
    BinaryFile,
    UID
  ) %>%
  group_by(HDBSCAN_cluster) %>%
  mutate(
    waveform_number = row_number(),
    waveform_facet_label = paste0("Click ", waveform_number, "\nUID ", UID)
  ) %>%
  ungroup()

sampled_waveforms_for_plot <- sampled_waveforms_long %>%
  left_join(
    waveform_panel_lookup,
    by = c(
      "HDBSCAN_cluster",
      "EventID",
      "DriftID_clean",
      "BinaryFile",
      "UID"
    )
  ) %>%
  mutate(
    HDBSCAN_cluster = as.character(HDBSCAN_cluster),
    waveform_facet_label = factor(
      waveform_facet_label,
      levels = waveform_panel_lookup$waveform_facet_label
    )
  )

# ------------------------------------------------------------
# Helper: make one 5x5 waveform grid for a cluster
# ------------------------------------------------------------

make_cluster_waveform_grid <- function(cluster_id) {
  
  cluster_id <- as.character(cluster_id)
  
  cluster_plot_data <- sampled_waveforms_for_plot %>%
    filter(HDBSCAN_cluster == cluster_id)
  
  n_waveforms <- cluster_plot_data %>%
    distinct(UID, EventID, DriftID_clean, BinaryFile) %>%
    nrow()
  
  if (n_waveforms != 25) {
    warning(
      "Cluster ",
      cluster_id,
      " has ",
      n_waveforms,
      " waveforms instead of 25."
    )
  }
  
  cluster_line_color <- cluster_colors[[cluster_id]]
  
  ggplot(
    cluster_plot_data,
    aes(x = sample_index, y = amplitude)
  ) +
    geom_hline(
      yintercept = 0,
      linewidth = 0.2,
      color = "grey70"
    ) +
    geom_line(
      linewidth = 0.25,
      color = cluster_line_color
    ) +
    facet_wrap(
      ~ waveform_facet_label,
      ncol = 5,
      nrow = 5
    ) +
    coord_cartesian(
      ylim = c(-1, 1)
    ) +
    labs(
      title = paste0("Cluster ", cluster_id, " Waveforms"),
      subtitle = paste0(
        "25 randomly sampled clicks; channel ",
        selected_wave_channel,
        "; normalized amplitude"
      ),
      x = "Sample index",
      y = "Normalized amplitude"
    ) +
    theme_bw(base_size = 11) +
    theme(
      strip.text = element_text(size = 6),
      plot.title = element_text(face = "bold", size = 15),
      plot.subtitle = element_text(size = 10),
      axis.text = element_text(size = 7),
      axis.title = element_text(size = 10),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(linewidth = 0.2),
      panel.spacing = unit(0.35, "lines")
    )
}

# ------------------------------------------------------------
# Save one grid per cluster, including cluster 0/noise
# ------------------------------------------------------------

individual_grid_files <- map_chr(individual_grid_clusters, function(cluster_id) {
  
  p_cluster <- make_cluster_waveform_grid(cluster_id)
  
  out_file <- file.path(
    waveform_dir,
    paste0(
      "cluster_",
      cluster_id,
      "_waveforms_5x5_25_clicks_channel_",
      selected_wave_channel,
      ".png"
    )
  )
  
  ggsave(
    filename = out_file,
    plot = p_cluster,
    width = 10,
    height = 9,
    dpi = 300
  )
  
  cat("Saved:", out_file, "\n")
  
  out_file
})

# ------------------------------------------------------------
# Print one example in the RMarkdown output
# ------------------------------------------------------------

print(make_cluster_waveform_grid("0"))

cat("\nIndividual 5x5 waveform grids saved for clusters:\n")
print(individual_grid_clusters)

```

## Save 3x3 composite waveform grid for clusters 1-9

```{r save-composite-cluster-waveform-grid, message=FALSE, warning=FALSE}

# ------------------------------------------------------------
# Load patchwork for combining ggplots
# ------------------------------------------------------------

if (!requireNamespace("patchwork", quietly = TRUE)) {
  install.packages("patchwork")
}

library(patchwork)

# ------------------------------------------------------------
# Helper: make compact 5x5 cluster panel for composite figure
# ------------------------------------------------------------

make_compact_cluster_waveform_grid <- function(cluster_id) {
  
  cluster_id <- as.character(cluster_id)
  
  cluster_plot_data <- sampled_waveforms_for_plot %>%
    filter(HDBSCAN_cluster == cluster_id) %>%
    mutate(
      waveform_number = factor(
        waveform_number,
        levels = 1:25
      )
    )
  
  n_waveforms <- cluster_plot_data %>%
    distinct(UID, EventID, DriftID_clean, BinaryFile) %>%
    nrow()
  
  if (n_waveforms != 25) {
    warning(
      "Cluster ",
      cluster_id,
      " has ",
      n_waveforms,
      " waveforms instead of 25."
    )
  }
  
  cluster_line_color <- cluster_colors[[cluster_id]]
  
  ggplot(
    cluster_plot_data,
    aes(x = sample_index, y = amplitude)
  ) +
    geom_hline(
      yintercept = 0,
      linewidth = 0.15,
      color = "grey75"
    ) +
    geom_line(
      linewidth = 0.18,
      color = cluster_line_color
    ) +
    facet_wrap(
      ~ waveform_number,
      ncol = 5,
      nrow = 5
    ) +
    coord_cartesian(
      ylim = c(-1, 1)
    ) +
    labs(
      title = paste0("Cluster ", cluster_id),
      x = NULL,
      y = NULL
    ) +
    theme_bw(base_size = 8) +
    theme(
      plot.title = element_text(
        face = "bold",
        size = 11,
        hjust = 0.5
      ),
      strip.text = element_blank(),
      strip.background = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(linewidth = 0.12),
      panel.spacing = unit(0.12, "lines"),
      plot.margin = margin(4, 4, 4, 4)
    )
}

# ------------------------------------------------------------
# Build one compact panel for each assigned cluster
# ------------------------------------------------------------

composite_cluster_plots <- map(
  composite_grid_clusters,
  make_compact_cluster_waveform_grid
)

# ------------------------------------------------------------
# Combine into 3x3 composite
# ------------------------------------------------------------

p_composite_waveforms <- wrap_plots(
  composite_cluster_plots,
  ncol = 3,
  nrow = 3
) +
  plot_annotation(
    title = "Sampled Waveforms by HDBSCAN Cluster",
    subtitle = paste0(
      "Clusters 1-9 only; 25 randomly sampled clicks per cluster; channel ",
      selected_wave_channel,
      "; normalized amplitude"
    ),
    theme = theme(
      plot.title = element_text(
        face = "bold",
        size = 18,
        hjust = 0.5
      ),
      plot.subtitle = element_text(
        size = 12,
        hjust = 0.5
      )
    )
  )

print(p_composite_waveforms)

# ------------------------------------------------------------
# Save composite figure
# ------------------------------------------------------------

composite_file <- file.path(
  waveform_dir,
  paste0(
    "clusters_1_to_9_waveforms_composite_3x3_each_5x5_channel_",
    selected_wave_channel,
    ".png"
  )
)

ggsave(
  filename = composite_file,
  plot = p_composite_waveforms,
  width = 18,
  height = 16,
  dpi = 300
)

cat("Saved composite waveform figure to:\n")
cat(composite_file, "\n")

```

## ============================================================
## NEW OUTPUT SET:
## 25 new random waveforms per HDBSCAN cluster,
## excluding the previous sampled waveform UIDs
## ============================================================

```{r new-25-per-cluster-waveform-output-excluding-prior, message=FALSE, warning=FALSE}

library(dplyr)
library(readr)
library(purrr)
library(stringr)
library(tibble)
library(ggplot2)

if (!requireNamespace("PamBinaries", quietly = TRUE)) {
  install.packages("PamBinaries")
}
library(PamBinaries)

if (!requireNamespace("patchwork", quietly = TRUE)) {
  install.packages("patchwork")
}
library(patchwork)

set.seed(456)

# ------------------------------------------------------------
# Settings
# ------------------------------------------------------------

base_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1"

audit_dir <- file.path(
  base_dir,
  "temporal_visualizations/bearing_source_audit"
)

old_waveform_dir <- file.path(
  base_dir,
  "plots/waveform_cluster_grids"
)

new_waveform_dir <- file.path(
  base_dir,
  "plots/waveform_cluster_grids_NEW_25_per_cluster_excluding_prior"
)

dir.create(new_waveform_dir, recursive = TRUE, showWarnings = FALSE)

sampled_joined_file <- file.path(
  audit_dir,
  "sampled_hdbscan_clicks_joined_to_offline_clicks.csv"
)

manifest_file <- file.path(
  audit_dir,
  "PAMGuard_binary_manifest_for_bearing_export_CLEAN.csv"
)

selected_wave_channel <- 1
waveform_sample_n <- 25

individual_grid_clusters <- as.character(0:9)
composite_grid_clusters <- as.character(1:9)

cluster_colors <- c(
  "0" = "grey80",
  "1" = "#E41A1C",
  "2" = "#FF7F00",
  "3" = "#FFD92F",
  "4" = "#4DAF4A",
  "5" = "#00BFC4",
  "6" = "#377EB8",
  "7" = "#984EA3",
  "8" = "#F781BF",
  "9" = "#A65628"
)

# ------------------------------------------------------------
# Load input files
# ------------------------------------------------------------

required_files <- c(sampled_joined_file, manifest_file)

missing_files <- required_files[!file.exists(required_files)]

if (length(missing_files) > 0) {
  stop(
    "These required files are missing:\n",
    paste(missing_files, collapse = "\n")
  )
}

sampled_clicks <- read_csv(sampled_joined_file, show_col_types = FALSE) %>%
  mutate(
    UID = as.character(UID),
    EventID = as.character(EventID),
    BinaryFile = as.character(BinaryFile),
    DriftID_clean = as.character(DriftID_clean),
    HDBSCAN_cluster = as.character(HDBSCAN_cluster)
  )

manifest_clean <- read_csv(manifest_file, show_col_types = FALSE) %>%
  mutate(
    DriftID_clean = as.character(DriftID_clean),
    BinaryFile = as.character(BinaryFile),
    full_path = as.character(full_path),
    pgdx_full_path = as.character(pgdx_full_path)
  )

# ------------------------------------------------------------
# Load prior waveform sample so we do NOT reuse those clicks
# ------------------------------------------------------------

old_sample_file <- file.path(
  old_waveform_dir,
  paste0("waveform_sample_25_clicks_per_cluster_channel_", selected_wave_channel, ".csv")
)

if (!file.exists(old_sample_file)) {
  stop(
    "Could not find the previous waveform sample table:\n",
    old_sample_file,
    "\n\nRun the original waveform sampling first, or check the old_waveform_dir path."
  )
}

old_sampled_uids <- read_csv(old_sample_file, show_col_types = FALSE) %>%
  transmute(UID = as.character(UID)) %>%
  distinct()

cat("Previous sampled UIDs to exclude:", nrow(old_sampled_uids), "\n\n")

# ------------------------------------------------------------
# Create new sample: 25 unused clicks per cluster
# ------------------------------------------------------------

available_for_new_sample <- sampled_clicks %>%
  filter(
    HDBSCAN_cluster %in% individual_grid_clusters,
    !is.na(UID),
    !is.na(BinaryFile),
    !is.na(DriftID_clean)
  ) %>%
  anti_join(old_sampled_uids, by = "UID")

cluster_availability <- available_for_new_sample %>%
  count(HDBSCAN_cluster, name = "n_available_after_excluding_prior") %>%
  arrange(as.numeric(HDBSCAN_cluster))

cat("Available unused clicks per cluster:\n")
print(cluster_availability, n = Inf)

clusters_with_too_few <- cluster_availability %>%
  filter(n_available_after_excluding_prior < waveform_sample_n)

if (nrow(clusters_with_too_few) > 0) {
  stop(
    "At least one cluster has fewer than ",
    waveform_sample_n,
    " unused clicks after excluding the prior waveform sample:\n",
    paste(clusters_with_too_few$HDBSCAN_cluster, collapse = ", ")
  )
}

waveform_sample_keys <- available_for_new_sample %>%
  group_by(HDBSCAN_cluster) %>%
  slice_sample(n = waveform_sample_n) %>%
  ungroup() %>%
  mutate(
    HDBSCAN_cluster = factor(
      HDBSCAN_cluster,
      levels = individual_grid_clusters
    )
  ) %>%
  arrange(HDBSCAN_cluster, DriftID_clean, BinaryFile, UID)

waveform_sample_manifest <- waveform_sample_keys %>%
  left_join(
    manifest_clean %>%
      select(
        DriftID_clean,
        BinaryFile,
        full_path,
        pgdx_full_path
      ),
    by = c("DriftID_clean", "BinaryFile")
  )

missing_binary_paths <- waveform_sample_manifest %>%
  filter(is.na(full_path) | !file.exists(full_path))

if (nrow(missing_binary_paths) > 0) {
  print(
    missing_binary_paths %>%
      select(HDBSCAN_cluster, EventID, DriftID_clean, BinaryFile, UID, full_path),
    n = Inf
  )
  
  stop("Some newly sampled clicks do not have valid PAMGuard binary paths.")
}

cat("\nNew sampled clicks per cluster:\n")
waveform_sample_manifest %>%
  count(HDBSCAN_cluster, name = "n_sampled_clicks") %>%
  arrange(as.numeric(as.character(HDBSCAN_cluster))) %>%
  print(n = Inf)

waveform_sample_file <- file.path(
  new_waveform_dir,
  paste0(
    "waveform_sample_25_clicks_per_cluster_EXCLUDING_PRIOR_channel_",
    selected_wave_channel,
    ".csv"
  )
)

write_csv(waveform_sample_manifest, waveform_sample_file)

cat("\nSaved new waveform sample table to:\n")
cat(waveform_sample_file, "\n\n")

# ------------------------------------------------------------
# Helper: extract one normalized waveform
# ------------------------------------------------------------

extract_one_waveform <- function(raw_click, channel = selected_wave_channel, normalize = TRUE) {
  
  wave_matrix <- raw_click$wave
  
  if (is.null(wave_matrix)) {
    stop("This raw click does not contain a wave field.")
  }
  
  if (!is.matrix(wave_matrix)) {
    wave_matrix <- as.matrix(wave_matrix)
  }
  
  n_samples <- nrow(wave_matrix)
  n_channels <- ncol(wave_matrix)
  
  if (channel > n_channels) {
    stop(
      "Requested channel ",
      channel,
      ", but this click only has ",
      n_channels,
      " channel(s)."
    )
  }
  
  waveform <- wave_matrix[, channel]
  
  if (normalize) {
    max_abs_amp <- max(abs(waveform), na.rm = TRUE)
    
    if (is.finite(max_abs_amp) && max_abs_amp > 0) {
      waveform <- waveform / max_abs_amp
    }
  }
  
  tibble(
    sample_index = seq_len(n_samples),
    amplitude = as.numeric(waveform),
    wave_channel_used = channel,
    n_channels = n_channels,
    n_samples = n_samples,
    UID = as.character(raw_click$UID),
    sampleDuration = as.integer(raw_click$sampleDuration),
    maxAmplitude = as.numeric(raw_click$maxAmplitude)
  )
}

# ------------------------------------------------------------
# Helper: read selected waveforms from one binary file
# ------------------------------------------------------------

read_sampled_waveforms_from_binary <- function(binary_sample_df) {
  
  this_drift <- unique(binary_sample_df$DriftID_clean)
  this_binary <- unique(binary_sample_df$BinaryFile)
  this_path <- unique(binary_sample_df$full_path)
  
  if (length(this_drift) != 1 || length(this_binary) != 1 || length(this_path) != 1) {
    stop("Each binary_sample_df must correspond to exactly one DriftID_clean + BinaryFile + full_path.")
  }
  
  needed_uids_chr <- unique(as.character(binary_sample_df$UID))
  needed_uids_num <- suppressWarnings(as.numeric(needed_uids_chr))
  
  cat("Reading:", this_drift, "|", this_binary, "|", length(needed_uids_chr), "sampled UID(s)\n")
  
  out <- tryCatch({
    
    pg <- PamBinaries::loadPamguardBinaryFile(
      fileName = this_path,
      skipLarge = FALSE,
      skipData = FALSE,
      keepUIDs = needed_uids_num,
      convertDate = TRUE
    )
    
    available_uids <- names(pg$data)
    
    map_dfr(seq_len(nrow(binary_sample_df)), function(i) {
      
      click_meta <- binary_sample_df[i, ]
      uid_chr <- as.character(click_meta$UID)
      
      if (!uid_chr %in% available_uids) {
        warning(
          "UID not found in loaded binary: ",
          uid_chr,
          " | ",
          this_binary
        )
        
        return(tibble())
      }
      
      raw_click <- pg$data[[uid_chr]]
      
      extract_one_waveform(
        raw_click = raw_click,
        channel = selected_wave_channel,
        normalize = TRUE
      ) %>%
        mutate(
          HDBSCAN_cluster = as.character(click_meta$HDBSCAN_cluster),
          EventID = as.character(click_meta$EventID),
          DriftID_clean = as.character(click_meta$DriftID_clean),
          BinaryFile = as.character(click_meta$BinaryFile),
          waveform_panel_id = paste0(
            "Cluster ",
            as.character(click_meta$HDBSCAN_cluster),
            " | UID ",
            uid_chr
          )
        )
    })
    
  }, error = function(e) {
    
    warning(
      "Failed to read binary: ",
      this_path,
      "\nError: ",
      e$message
    )
    
    tibble()
  })
  
  return(out)
}

# ------------------------------------------------------------
# Extract all newly sampled waveforms
# ------------------------------------------------------------

waveform_binary_groups <- waveform_sample_manifest %>%
  mutate(
    UID = as.character(UID),
    HDBSCAN_cluster = as.character(HDBSCAN_cluster)
  ) %>%
  group_by(DriftID_clean, BinaryFile, full_path) %>%
  group_split()

cat("Number of binary files to read:", length(waveform_binary_groups), "\n\n")

sampled_waveforms_long <- map_dfr(
  waveform_binary_groups,
  read_sampled_waveforms_from_binary
)

cat("\nFinished waveform extraction.\n\n")

if (nrow(sampled_waveforms_long) == 0) {
  stop("No waveforms were extracted. Check binary paths, UIDs, and PamBinaries loading.")
}

# ------------------------------------------------------------
# Verify extraction
# ------------------------------------------------------------

cat("Extracted waveforms per cluster:\n")

sampled_waveforms_long %>%
  distinct(HDBSCAN_cluster, UID) %>%
  count(HDBSCAN_cluster, name = "n_waveforms_extracted") %>%
  arrange(as.numeric(HDBSCAN_cluster)) %>%
  print(n = Inf)

expected_waveform_keys <- waveform_sample_manifest %>%
  transmute(
    HDBSCAN_cluster = as.character(HDBSCAN_cluster),
    DriftID_clean = as.character(DriftID_clean),
    BinaryFile = as.character(BinaryFile),
    UID = as.character(UID)
  ) %>%
  distinct()

extracted_waveform_keys <- sampled_waveforms_long %>%
  transmute(
    HDBSCAN_cluster = as.character(HDBSCAN_cluster),
    DriftID_clean = as.character(DriftID_clean),
    BinaryFile = as.character(BinaryFile),
    UID = as.character(UID)
  ) %>%
  distinct()

missing_waveform_keys <- expected_waveform_keys %>%
  anti_join(
    extracted_waveform_keys,
    by = c("HDBSCAN_cluster", "DriftID_clean", "BinaryFile", "UID")
  )

if (nrow(missing_waveform_keys) > 0) {
  cat("\nWARNING: Some newly sampled clicks were not extracted:\n")
  print(missing_waveform_keys, n = Inf)
} else {
  cat("\nAll newly sampled clicks were successfully extracted.\n")
}

waveform_long_file <- file.path(
  new_waveform_dir,
  paste0(
    "sampled_waveforms_long_25_per_cluster_EXCLUDING_PRIOR_channel_",
    selected_wave_channel,
    ".rds"
  )
)

saveRDS(sampled_waveforms_long, waveform_long_file)

cat("\nSaved extracted waveform data to:\n")
cat(waveform_long_file, "\n\n")

# ------------------------------------------------------------
# Prepare plotting data
# ------------------------------------------------------------

waveform_panel_lookup <- sampled_waveforms_long %>%
  distinct(
    HDBSCAN_cluster,
    EventID,
    DriftID_clean,
    BinaryFile,
    UID
  ) %>%
  arrange(
    as.numeric(HDBSCAN_cluster),
    EventID,
    DriftID_clean,
    BinaryFile,
    UID
  ) %>%
  group_by(HDBSCAN_cluster) %>%
  mutate(
    waveform_number = row_number(),
    waveform_facet_label = paste0("Click ", waveform_number, "\nUID ", UID)
  ) %>%
  ungroup()

sampled_waveforms_for_plot <- sampled_waveforms_long %>%
  left_join(
    waveform_panel_lookup,
    by = c(
      "HDBSCAN_cluster",
      "EventID",
      "DriftID_clean",
      "BinaryFile",
      "UID"
    )
  ) %>%
  mutate(
    HDBSCAN_cluster = as.character(HDBSCAN_cluster),
    waveform_facet_label = factor(
      waveform_facet_label,
      levels = waveform_panel_lookup$waveform_facet_label
    )
  )

# ------------------------------------------------------------
# Save individual 5x5 waveform grids for each cluster
# ------------------------------------------------------------

make_cluster_waveform_grid <- function(cluster_id) {
  
  cluster_id <- as.character(cluster_id)
  
  cluster_plot_data <- sampled_waveforms_for_plot %>%
    filter(HDBSCAN_cluster == cluster_id)
  
  n_waveforms <- cluster_plot_data %>%
    distinct(UID, EventID, DriftID_clean, BinaryFile) %>%
    nrow()
  
  if (n_waveforms != 25) {
    warning(
      "Cluster ",
      cluster_id,
      " has ",
      n_waveforms,
      " waveforms instead of 25."
    )
  }
  
  cluster_line_color <- cluster_colors[[cluster_id]]
  
  ggplot(
    cluster_plot_data,
    aes(x = sample_index, y = amplitude)
  ) +
    geom_hline(
      yintercept = 0,
      linewidth = 0.2,
      color = "grey70"
    ) +
    geom_line(
      linewidth = 0.25,
      color = cluster_line_color
    ) +
    facet_wrap(
      ~ waveform_facet_label,
      ncol = 5,
      nrow = 5
    ) +
    coord_cartesian(
      ylim = c(-1, 1)
    ) +
    labs(
      title = paste0("Cluster ", cluster_id, " Waveforms"),
      subtitle = paste0(
        "25 newly sampled clicks; previous waveform sample excluded; channel ",
        selected_wave_channel,
        "; normalized amplitude"
      ),
      x = "Sample index",
      y = "Normalized amplitude"
    ) +
    theme_bw(base_size = 11) +
    theme(
      strip.text = element_text(size = 6),
      plot.title = element_text(face = "bold", size = 15),
      plot.subtitle = element_text(size = 10),
      axis.text = element_text(size = 7),
      axis.title = element_text(size = 10),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(linewidth = 0.2),
      panel.spacing = unit(0.35, "lines")
    )
}

individual_grid_files <- map_chr(individual_grid_clusters, function(cluster_id) {
  
  p_cluster <- make_cluster_waveform_grid(cluster_id)
  
  out_file <- file.path(
    new_waveform_dir,
    paste0(
      "cluster_",
      cluster_id,
      "_waveforms_5x5_25_clicks_EXCLUDING_PRIOR_channel_",
      selected_wave_channel,
      ".png"
    )
  )
  
  ggsave(
    filename = out_file,
    plot = p_cluster,
    width = 10,
    height = 9,
    dpi = 300
  )
  
  cat("Saved:", out_file, "\n")
  
  out_file
})

print(make_cluster_waveform_grid("0"))

# ------------------------------------------------------------
# Save 3x3 composite waveform grid for clusters 1-9
# ------------------------------------------------------------

make_compact_cluster_waveform_grid <- function(cluster_id) {
  
  cluster_id <- as.character(cluster_id)
  
  cluster_plot_data <- sampled_waveforms_for_plot %>%
    filter(HDBSCAN_cluster == cluster_id) %>%
    mutate(
      waveform_number = factor(
        waveform_number,
        levels = 1:25
      )
    )
  
  n_waveforms <- cluster_plot_data %>%
    distinct(UID, EventID, DriftID_clean, BinaryFile) %>%
    nrow()
  
  if (n_waveforms != 25) {
    warning(
      "Cluster ",
      cluster_id,
      " has ",
      n_waveforms,
      " waveforms instead of 25."
    )
  }
  
  cluster_line_color <- cluster_colors[[cluster_id]]
  
  ggplot(
    cluster_plot_data,
    aes(x = sample_index, y = amplitude)
  ) +
    geom_hline(
      yintercept = 0,
      linewidth = 0.15,
      color = "grey75"
    ) +
    geom_line(
      linewidth = 0.18,
      color = cluster_line_color
    ) +
    facet_wrap(
      ~ waveform_number,
      ncol = 5,
      nrow = 5
    ) +
    coord_cartesian(
      ylim = c(-1, 1)
    ) +
    labs(
      title = paste0("Cluster ", cluster_id),
      x = NULL,
      y = NULL
    ) +
    theme_bw(base_size = 8) +
    theme(
      plot.title = element_text(
        face = "bold",
        size = 11,
        hjust = 0.5
      ),
      strip.text = element_blank(),
      strip.background = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(linewidth = 0.12),
      panel.spacing = unit(0.12, "lines"),
      plot.margin = margin(4, 4, 4, 4)
    )
}

composite_cluster_plots <- map(
  composite_grid_clusters,
  make_compact_cluster_waveform_grid
)

p_composite_waveforms <- wrap_plots(
  composite_cluster_plots,
  ncol = 3,
  nrow = 3
) +
  plot_annotation(
    title = "New Sampled Waveforms by HDBSCAN Cluster",
    subtitle = paste0(
      "Clusters 1-9 only; 25 newly sampled clicks per cluster; previous waveform sample excluded; channel ",
      selected_wave_channel,
      "; normalized amplitude"
    ),
    theme = theme(
      plot.title = element_text(
        face = "bold",
        size = 18,
        hjust = 0.5
      ),
      plot.subtitle = element_text(
        size = 12,
        hjust = 0.5
      )
    )
  )

print(p_composite_waveforms)

composite_file <- file.path(
  new_waveform_dir,
  paste0(
    "clusters_1_to_9_waveforms_composite_3x3_each_5x5_EXCLUDING_PRIOR_channel_",
    selected_wave_channel,
    ".png"
  )
)

ggsave(
  filename = composite_file,
  plot = p_composite_waveforms,
  width = 18,
  height = 16,
  dpi = 300
)

cat("\nSaved composite waveform figure to:\n")
cat(composite_file, "\n")

cat("\nDONE. New outputs saved in:\n")
cat(new_waveform_dir, "\n")

```



































