---
title: "Clustering_HDBSCAN"
author: "Sarah Shiers"
date: "2026-04-07"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

## Load Required Packages

```{r}

required_packages <- c("ggplot2", "dplyr", "cluster", "dbscan")

for (pkg in required_packages) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg, repos = "https://cran.rstudio.com")
    library(pkg, character.only = TRUE)
  }
}

```

## User Settings

```{r}

cluster_method <- "hdbscan"   # options: "kmeans", "kshape", "hdbscan"

output_dir <- "E:/Analysis/Clustering Output"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

freq_min <- 19000
freq_max <- 50000

# HDBSCAN settings
hdbscan_minPts <- 5
min_samples <- NULL   # leave NULL to use default behavior in hdbscan()

# Sampling settings
max_clicks_per_event <- 200

# Optional caps after event-balanced sampling
eval_sample_cap <- NULL

# Whether to run final HDBSCAN on all clicks
run_hdbscan_on_full_dataset <- FALSE

hdbscan_label <- paste0("minPts_", hdbscan_minPts)

hdb <- hdbscan(AllSpec_pca, minPts = hdbscan_minPts)

```

## Load COmbined Tukey Spectra

```{r}

combined_file <- file.choose()
combined_tukey <- readRDS(combined_file)

cat("Object class:", class(combined_tukey), "\n")
cat("Number of events:", length(combined_tukey), "\n")

if (length(combined_tukey) == 0) {
  stop("combined_tukey is empty.")
}

required_fields <- c("freq", "allSpec", "EventID", "UID")

for (i in seq_along(combined_tukey)) {
  
  missing_fields <- setdiff(required_fields, names(combined_tukey[[i]]))
  
  if (length(missing_fields) > 0) {
    stop(
      paste(
        "Event", i, "is missing required fields:",
        paste(missing_fields, collapse = ", ")
      )
    )
  }
}

cat("Required fields confirmed for all events.\n")

```

## Inspect Frequency Grid

```{r}

reference_freq <- combined_tukey[[1]]$freq

freq_check <- data.frame(
  EventID = character(),
  SameAsReference = logical(),
  LengthFreq = integer(),
  FreqStep = numeric(),
  MaxFreq = numeric(),
  stringsAsFactors = FALSE
)

for (i in seq_along(combined_tukey)) {
  
  current_freq <- combined_tukey[[i]]$freq
  
  freq_check <- rbind(
    freq_check,
    data.frame(
      EventID = combined_tukey[[i]]$EventID,
      SameAsReference = identical(current_freq, reference_freq),
      LengthFreq = length(current_freq),
      FreqStep = current_freq[2] - current_freq[1],
      MaxFreq = max(current_freq),
      stringsAsFactors = FALSE
    )
  )
}

freq_check


```

## Choose Reference Frequency Grid

```{r}

reference_freq <- combined_tukey[[1]]$freq
reference_event <- combined_tukey[[1]]$EventID

cat("Reference event:", reference_event, "\n")
cat("Reference frequency step:", reference_freq[2] - reference_freq[1], "Hz\n")
cat("Reference max frequency:", max(reference_freq), "Hz\n")

```

## Interpolate Spectra to Shared Grid

```{r}

interpolate_spectrum <- function(old_freq, old_spec, new_freq) {
  approx(
    x = old_freq,
    y = old_spec,
    xout = new_freq,
    method = "linear",
    rule = 2
  )$y
}

interpolated_tukey <- vector("list", length(combined_tukey))

for (i in seq_along(combined_tukey)) {
  
  tukey_spectra <- combined_tukey[[i]]
  
  old_freq <- tukey_spectra$freq
  old_allSpec <- tukey_spectra$allSpec
  
  interpolated_allSpec <- apply(
    old_allSpec,
    2,
    function(spec) interpolate_spectrum(old_freq, spec, reference_freq)
  )
  
  interpolated_allSpec <- as.matrix(interpolated_allSpec)
  
  tukey_spectra$allSpec <- interpolated_allSpec
  tukey_spectra$freq <- reference_freq
  
  interpolated_tukey[[i]] <- tukey_spectra
}

```

## Confirm Interpolation

```{r}

all_match_after_interp <- all(
  sapply(interpolated_tukey, function(x) identical(x$freq, reference_freq))
)

cat("All frequency grids match reference after interpolation:", all_match_after_interp, "\n")

if (!all_match_after_interp) {
  stop("Interpolation failed: not all events match the reference frequency grid.")
}

```

## Combine Spectra Across All Events

```{r}

all_click_metadata <- list()
all_click_spectra <- list()

for (i in seq_along(interpolated_tukey)) {
  
  tukey_spectra <- interpolated_tukey[[i]]
  
  event_id <- tukey_spectra$EventID
  uid <- tukey_spectra$UID
  
  click_spectra <- t(tukey_spectra$allSpec)
  
  click_meta <- data.frame(
    EventID = event_id,
    UID = uid,
    stringsAsFactors = FALSE
  )
  
  all_click_metadata[[i]] <- click_meta
  all_click_spectra[[i]] <- click_spectra
}

click_metadata <- do.call(rbind, all_click_metadata)
all_spec_matrix <- do.call(rbind, all_click_spectra)

cat("Combined spectra dimensions:", dim(all_spec_matrix)[1], "clicks x", dim(all_spec_matrix)[2], "frequency bins\n")
cat("Combined metadata rows:", nrow(click_metadata), "\n")

if (nrow(all_spec_matrix) != nrow(click_metadata)) {
  stop("Mismatch between number of spectra rows and metadata rows.")
}

```

## Restrict Spectra to 19-50 kHz

```{r}

keepFreq <- which(reference_freq >= freq_min & reference_freq <= freq_max)

AllSpec_Focus <- all_spec_matrix[, keepFreq]
centroid_freq <- reference_freq[keepFreq]

cat("Frequency band retained:", 
    min(centroid_freq), "to", max(centroid_freq), "Hz\n")

cat("Focused spectra dimensions:",
    dim(AllSpec_Focus)[1], "clicks x",
    dim(AllSpec_Focus)[2], "frequency bins\n")

```

## Scale Spectra

```{r}

# Row-wise scaling: emphasizes spectral shape within each click
AllSpec_scale <- t(scale(t(AllSpec_Focus)))
AllSpec_scale[is.na(AllSpec_scale)] <- 0

cat("Scaled spectra dimensions:",
    dim(AllSpec_scale)[1], "clicks x",
    dim(AllSpec_scale)[2], "frequency bins\n")

cat("Number of NA values after replacement:",
    sum(is.na(AllSpec_scale)), "\n")

```

## Helper Function: Sample Up to N Clicks Per Event

```{r}

sample_by_event <- function(metadata_df, max_per_event, seed = NULL, max_total = NULL) {
  
  if (!is.null(seed)) {
    set.seed(seed)
  }
  
  split_indices <- split(seq_len(nrow(metadata_df)), metadata_df$EventID)
  
  sampled_indices <- unlist(
    lapply(split_indices, function(idx) {
      sample(idx, size = min(length(idx), max_per_event))
    }),
    use.names = FALSE
  )
  
  if (!is.null(max_total) && length(sampled_indices) > max_total) {
    sampled_indices <- sample(sampled_indices, size = max_total)
  }
  
  return(sampled_indices)
}

```

## Create Evaluation Sample

This sample is used for silhouette and PCA scatterplot.
It uses up to 200 clicks per event.

```{r}

eval_index <- sample_by_event(
  metadata_df = click_metadata,
  max_per_event = max_clicks_per_event,
  seed = 2,
  max_total = eval_sample_cap
)

AllSpec_eval <- AllSpec_scale[eval_index, ]
click_metadata_eval <- click_metadata[eval_index, ]

cat("Evaluation sample size used:", nrow(AllSpec_eval), "\n")

```

## Sampling Diagnostics

```{r}

event_total <- click_metadata %>%
  count(EventID, name = "total_clicks")

event_sampled <- click_metadata_eval %>%
  count(EventID, name = "sampled_clicks")

sampling_table <- event_total %>%
  left_join(event_sampled, by = "EventID") %>%
  mutate(sampled_clicks = ifelse(is.na(sampled_clicks), 0, sampled_clicks)) %>%
  arrange(EventID)

sampling_table

```

## HDBSCAN Parameter Sweep on Evaluation Sample

```{r}

hdbscan_summary <- data.frame(
  minPts = integer(),
  n_clusters = integer(),
  n_noise = integer(),
  prop_noise = numeric()
)

for (mp in c(5, 10, 15, 20, 30)) {
  
  hdb_tmp <- hdbscan(AllSpec_eval, minPts = mp)
  
  n_clusters <- length(setdiff(unique(hdb_tmp$cluster), 0))
  n_noise <- sum(hdb_tmp$cluster == 0)
  prop_noise <- mean(hdb_tmp$cluster == 0)
  
  hdbscan_summary <- rbind(
    hdbscan_summary,
    data.frame(
      minPts = mp,
      n_clusters = n_clusters,
      n_noise = n_noise,
      prop_noise = prop_noise
    )
  )
  
  cat(
    "minPts =", mp,
    "| clusters =", n_clusters,
    "| noise =", n_noise,
    "| prop_noise =", round(prop_noise, 4), "\n"
  )
}

hdbscan_summary

```

## PCA Reduction for HDBSCAN

```{r}

# Run PCA on the evaluation sample
pca_model <- prcomp(AllSpec_eval, center = FALSE, scale. = FALSE)

# Cumulative proportion of variance explained
var_explained <- cumsum((pca_model$sdev^2) / sum(pca_model$sdev^2))

# Keep the minimum number of PCs explaining at least 95% of the variance
n_pcs <- which(var_explained >= 0.95)[1]

AllSpec_pca <- pca_model$x[, 1:n_pcs, drop = FALSE]

cat("PCs needed for 95% variance:", n_pcs, "\n")
cat("PCA feature matrix dimensions:",
    dim(AllSpec_pca)[1], "clicks x",
    dim(AllSpec_pca)[2], "PCs\n")

```

## HDBSCAN Parameter Sweep on PCA Space

```{r}

hdbscan_pca_summary <- data.frame(
  minPts = integer(),
  n_clusters = integer(),
  n_noise = integer(),
  prop_noise = numeric()
)

for (mp in c(5, 10, 15, 20, 30)) {
  
  hdb_tmp <- hdbscan(AllSpec_pca, minPts = mp)
  
  n_clusters <- length(setdiff(unique(hdb_tmp$cluster), 0))
  n_noise <- sum(hdb_tmp$cluster == 0)
  prop_noise <- mean(hdb_tmp$cluster == 0)
  
  hdbscan_pca_summary <- rbind(
    hdbscan_pca_summary,
    data.frame(
      minPts = mp,
      n_clusters = n_clusters,
      n_noise = n_noise,
      prop_noise = prop_noise
    )
  )
  
  cat(
    "minPts =", mp,
    "| clusters =", n_clusters,
    "| noise =", n_noise,
    "| prop_noise =", round(prop_noise, 4), "\n"
  )
}

hdbscan_pca_summary

```

## Run Final HDBSCAN on PCA Space

```{r}

hdb <- hdbscan(
  AllSpec_pca,
  minPts = 5
)

cluster_labels_eval <- hdb$cluster
click_metadata_eval$cluster <- cluster_labels_eval

n_clusters <- length(setdiff(unique(cluster_labels_eval), 0))
n_noise <- sum(cluster_labels_eval == 0)
prop_noise <- mean(cluster_labels_eval == 0)

cat("Clusters found:", n_clusters, "\n")
cat("Noise points:", n_noise, "\n")
cat("Noise proportion:", round(prop_noise, 4), "\n")

```

## PCA Scatterplot of HDBSCAN Clusters

```{r}

scatter_df <- data.frame(
  PC1 = AllSpec_pca[,1],
  PC2 = AllSpec_pca[,2],
  EventID = click_metadata_eval$EventID,
  UID = click_metadata_eval$UID,
  cluster = as.factor(cluster_labels_eval)
)

scatter_plot <- ggplot(scatter_df, aes(x = PC1, y = PC2, color = cluster)) +
  geom_point(alpha = 0.7, size = 1.2) +
  labs(
    title = "HDBSCAN Clusters (PCA Space)",
    x = "PC1",
    y = "PC2",
    color = "Cluster"
  )

print(scatter_plot)

ggsave(
  filename = file.path(output_dir, paste0("hdbscan_scatterplot_", hdbscan_label, ".png")),
  plot = scatter_plot,
  width = 7,
  height = 5
)

```

## Spectral Density Structure Diagnostic

```{r}

density_plot <- ggplot(scatter_df, aes(x = PC1, y = PC2)) +
  stat_density_2d(
    aes(fill = after_stat(level)),
    geom = "polygon",
    alpha = 0.6
  ) +
  scale_fill_viridis_c() +
  labs(
    title = "Spectral Density Structure (PCA Space)",
    x = "PC1",
    y = "PC2",
    fill = "Density"
  )

print(density_plot)

ggsave(
  filename = file.path(output_dir, paste0("hdbscan_pca_density_structure_", hdbscan_label, ".png")),
  plot = density_plot,
  width = 7,
  height = 5
)
```

## Compute Mean Spectra for HDBSCAN Clusters

```{r}

# Keep only non-noise points for cluster mean spectra
clustered_idx <- which(cluster_labels_eval != 0)

clustered_spectra <- AllSpec_eval[clustered_idx, , drop = FALSE]
clustered_labels <- cluster_labels_eval[clustered_idx]

cluster_mean_spectra <- aggregate(
  clustered_spectra,
  by = list(cluster = clustered_labels),
  FUN = mean
)

cluster_ids <- cluster_mean_spectra$cluster
cluster_mean_matrix <- as.matrix(cluster_mean_spectra[, -1])

cat("Clusters included in mean spectra:", paste(cluster_ids, collapse = ", "), "\n")
cat("Mean spectra matrix dimensions:",
    dim(cluster_mean_matrix)[1], "clusters x",
    dim(cluster_mean_matrix)[2], "frequency bins\n")

```

## Plot Mean Spectra for HDBSCAN Clusters

```{r}

cluster_ids <- as.numeric(cluster_ids)

centroid_file <- file.path(
  output_dir,
  paste0("hdbscan_mean_spectra_", hdbscan_label, ".png")
)

png(
  filename = centroid_file,
  width = 7,
  height = 5,
  units = "in",
  res = 300
)

matplot(
  x = centroid_freq / 1000,
  y = t(cluster_mean_matrix),
  type = "l",
  lty = 1,
  lwd = 2,
  main = "Mean Standardized Spectral Profiles by HDBSCAN Cluster",
  xlab = "Frequency (kHz)",
  ylab = "Standardized Intensity"
)

legend(
  "topright",
  legend = paste("Cluster", cluster_ids),
  col = 1:length(cluster_ids),
  lty = 1,
  lwd = 2
)

dev.off()

```

## Evaluate HDBSCAN with Mean Silhouette Score

```{r}

sil_idx <- which(cluster_labels_eval != 0)

if (length(unique(cluster_labels_eval[sil_idx])) < 2) {
  cat("Silhouette not computed: fewer than 2 non-noise clusters.\n")
} else {
  
  dist_eval_pca <- dist(AllSpec_pca[sil_idx, , drop = FALSE])
  
  sil_hdb <- silhouette(cluster_labels_eval[sil_idx], dist_eval_pca)
  mean_silhouette <- mean(sil_hdb[, "sil_width"])
  
  cat("Mean silhouette (non-noise points only) =", mean_silhouette, "\n")
}

```

## Silhouette Plot for HDBSCAN

```{r}

if (exists("sil_hdb")) {
  
  sil_df <- as.data.frame(sil_hdb) %>%
    mutate(obs = row_number()) %>%
    group_by(cluster) %>%
    arrange(desc(sil_width), .by_group = TRUE) %>%
    mutate(order = row_number()) %>%
    ungroup()
  
  sil_plot <- ggplot(sil_df, aes(x = order, y = sil_width)) +
    geom_col() +
    facet_wrap(~ cluster, scales = "free_x") +
    labs(
      title = "Silhouette Plot for HDBSCAN Clusters",
      x = "Observations",
      y = "Silhouette Width"
    )
  
  print(sil_plot)
  
  ggsave(
    filename = file.path(output_dir, paste0("hdbscan_silhouette_plot_", hdbscan_label, ".png")),
    plot = sil_plot,
    width = 8,
    height = 5
  )
}

```

## Save Standardized Cluster Profiles

```{r}
cluster_profile_df <- data.frame(
  Frequency_kHz = centroid_freq / 1000,
  t(cluster_mean_matrix)
)

colnames(cluster_profile_df) <- c(
  "Frequency_kHz",
  paste0("Cluster_", cluster_ids)
)

write.csv(
  cluster_profile_df,
  file = file.path(output_dir, paste0("hdbscan_scaled_profiles_", hdbscan_label, ".csv")),
  row.names = FALSE
)

```

## Save Summary Table

```{r}

summary_table <- data.frame(
  Method = "hdbscan",
  PCA_Used = TRUE,
  PCA_Variance_Threshold = 0.95,
  PCsRetained = n_pcs,
  MinPts = hdbscan_minPts,
  ClustersFound = n_clusters,
  NoisePoints = n_noise,
  NoiseProportion = prop_noise,
  MaxClicksPerEvent = max_clicks_per_event,
  EvalSampleSizeUsed = nrow(AllSpec_eval),
  TotalClicksInFullDataset = nrow(AllSpec_scale)
)

if (exists("mean_silhouette")) {
  summary_table$MeanSilhouette_NonNoise <- mean_silhouette
} else {
  summary_table$MeanSilhouette_NonNoise <- NA
}

summary_table

write.csv(
  summary_table,
  file = file.path(output_dir, paste0("hdbscan_summary_table_", hdbscan_label, ".csv")),
  row.names = FALSE
)

```



























