---
title: "ClusteringTest"
author: "Sarah Shiers"
date: "2026-03-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This RMD exists to determine best methodology for clustering; extrema versus whole spectrum, kmeans versus k-shape versus hdbscan.

```{r}

library("easypackages")
libraries("ggplot2", "dplyr", "PAMpal", "here","cluster")

source(here('R',"TukeySmoothing.R"))

```

Generate Clean Acoustic Studies

```{r}
# Base directories
baseDir <- here()

# Study identifiers
DriftID <- "ADRIFT_033"
evNum   <- 1
channel <- 1

study <- readRDS(paste0(baseDir,'/AcousticStudies/',DriftID,'_Gg_Clean.rds'))

#Only keep detections from Channel 1
study<-dplyr::filter(study,Channel==channel)

```

Create and Save Spectra

```{r}
#Create output directories
eventFolderName <- sprintf("Event_%03d", evNum)

eventTukeySpectraDir <- file.path(baseDir, 'TukeySpectra', DriftID, eventFolderName)

if (!dir.exists(eventTukeySpectraDir)) {
  dir.create(eventTukeySpectraDir, recursive = TRUE)
}

# Spectral parameters
wl <- 1024

# Frequency band (kHz)
fmin_khz <- 19
fmax_khz <- 50

# Tukey parameters
tukeyMedian      <- 5
tukeyHanningSize <- 5

# Spectra
tukey_spectra <- calculateAverageSpectra_tukey(
  study,
  evNum = evNum,
  channel = channel,
  tukeyMedian = tukeyMedian,
  tukeyHanningSize = tukeyHanningSize,
  plot = FALSE,
  wl = wl
)

# Save RDS
saveRDS(
  tukey_spectra,
  file.path(
    eventTukeySpectraDir,
    sprintf("%s_Event_%03d_Tukey_Spectra.rds", DriftID, evNum)
  )
)

```

To-Do:

-   Need to load in Tukey spectra from multiple events, and add an ID (like "ADRIFT_003_Event_001" to track spectra from each event.

-   Combine all of the spectra into a single matrix and then run the clustering

----

Reduce matrix of spectra to only keep frequencies between 19 - 50 kHz

```{r}
#Index of frequencies between 19 - 50 kHz 
keepFreq<-which(tukey_spectra$freq>=19000 & tukey_spectra$freq<= 50000)

#Matrix with all click spectra 
AllSpec_Focus<-tukey_spectra$allSpec[keepFreq,]
AllSpec$freq<-tukey_spectra$freq[keepFreq]

#Transpose matrix so each row represents one click and each column represents one spectrum
AllSpec_Focus<-t(AllSpec_Focus)

#Scale all spectra
AllSpec_scale<-scale(AllSpec_Focus)
```

Test K-Means clustering on spectra

```{r}
#control randomization
set.seed(1)

#Run kmeans
k=2
km<-kmeans(AllSpec_scale, k, nstart=50)
d<-dist(AllSpec_scale,method="euclidean")

sil <- silhouette(km$cluster, dist(AllSpec_scale))

#mathmatical evaluation of silhouettes (Higher mean silhouette = better separation)
mean(sil[, 3])

sil_df <- as.data.frame(sil) %>%
  mutate(obs = row_number()) %>%
  group_by(cluster) %>%
  arrange(desc(sil_width), .by_group = TRUE) %>%
  mutate(order = row_number()) %>%
  ungroup()


#Plot silhouettes
ggplot(sil_df, aes(x = order, y = sil_width)) +
  geom_col() +
  facet_wrap(~ cluster, scales = "free_x") +
  labs(
    title = paste("Silhouette plot (k =", k, ")"),
    x = "Observations (sorted within cluster)",
    y = "Silhouette width"
  )

#Plot centroids (spectra that are at the center of each cluster)
matplot(
  t(km$centers),
  type = "l",
  lty = 1,
  main = paste("k =", k)
)


```

How to choose number of clusters

```{r}

wss <- numeric(15)

for (k in 1:15) {
  km <- kmeans(AllSpec_scale, centers = k, nstart = 20)
  wss[k] <- km$tot.withinss
}

plot(1:15, wss, type = "b",
     xlab = "Number of Clusters",
     ylab = "Total Within SS")


```
