---
title: "Clustering_kmeans_minmax_Normalized"
author: "Sarah Moseley"
date: "2026-05-03"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

```

This new script runs a simplified k-means clustering workflow on Risso's dolphin click spectra using whole-spectrum min-max normalization.

For each click spectrum, the values across the retained frequency range are rescaled to [0, 1]. This removes most overall amplitude differences while preserving the relative spectral shape of each click.

The workflow is:

1. Load combined Tukey spectra.
2. Check and interpolate spectra to a shared frequency grid.
3. Combine all click spectra into one matrix.
4. Restrict spectra to 19–50 kHz.
5. Apply row-wise min-max normalization.
6. Run k-means clustering.
7. Generate elbow, silhouette, PCA, centroid, variability, and concatenated spectrogram outputs.

## Setup and User Settings

```{r}

library(ggplot2)
library(dplyr)
library(cluster)

# -----------------------------
# Base Output Directory
# -----------------------------
base_output_dir <- "E:/Analysis/Clustering Output/MinMax_Normalized_ADRIFT003_Corrected"
dir.create(base_output_dir, recursive = TRUE, showWarnings = FALSE)

# -----------------------------
# Analysis Label
# -----------------------------
analysis_name <- "Min-Max Normalized"
cluster_method <- "kmeans"

# -----------------------------
# Frequency Settings
# -----------------------------
freq_min <- 19000
freq_max <- 50000

# -----------------------------
# Clustering Settings
# -----------------------------
k_candidates <- 2:15
chosen_k <- NULL
k_label <- "global" #paste0("k_", chosen_k)

max_clicks_per_event <- 200
elbow_sample_cap <- 20000
eval_sample_cap <- NULL

run_kmeans_on_full_dataset <- TRUE

# -----------------------------
# Helper: Create Structured Output Folders
# -----------------------------
create_output_structure <- function(base_dir, analysis_name, method, k_label) {
  
  analysis_dir <- file.path(base_dir, analysis_name)
  method_dir <- file.path(analysis_dir, paste(method, k_label))
  
  dirs <- list(
    root = method_dir,
    plots = file.path(method_dir, "plots"),
    data = file.path(method_dir, "data"),
    diagnostics = file.path(method_dir, "diagnostics"),
    spectrograms = file.path(method_dir, "concatenated_spectrograms")
  )
  
  lapply(dirs, function(d) {
    dir.create(d, recursive = TRUE, showWarnings = FALSE)
  })
  
  return(dirs)
}

dirs <- create_output_structure(
  base_dir = base_output_dir,
  analysis_name = analysis_name,
  method = cluster_method,
  k_label = k_label
)

```

## Load Combined Tukey Spectra

```{r}

combined_file <- file.choose()
combined_tukey <- readRDS(combined_file)

class(combined_tukey)
length(combined_tukey)

```

## Inspect Frequency Grid

```{r}

reference_freq <- combined_tukey[[1]]$freq

freq_check <- data.frame(
  EventID = character(),
  SameAsReference = logical(),
  LengthFreq = integer(),
  FreqStep = numeric(),
  MaxFreq = numeric(),
  stringsAsFactors = FALSE
)

for (i in seq_along(combined_tukey)) {
  
  current_freq <- combined_tukey[[i]]$freq
  
  freq_check <- rbind(
    freq_check,
    data.frame(
      EventID = combined_tukey[[i]]$EventID,
      SameAsReference = identical(current_freq, reference_freq),
      LengthFreq = length(current_freq),
      FreqStep = current_freq[2] - current_freq[1],
      MaxFreq = max(current_freq),
      stringsAsFactors = FALSE
    )
  )
}

freq_check

```

## Interpolate Spectra to Shared Grid

```{r}

interpolate_spectrum <- function(old_freq, old_spec, new_freq) {
  approx(
    x = old_freq,
    y = old_spec,
    xout = new_freq,
    method = "linear",
    rule = 2
  )$y
}

interpolated_tukey <- vector("list", length(combined_tukey))

for (i in seq_along(combined_tukey)) {
  
  tukey_spectra <- combined_tukey[[i]]
  
  old_freq <- tukey_spectra$freq
  old_allSpec <- tukey_spectra$allSpec
  
  interpolated_allSpec <- apply(
    old_allSpec,
    2,
    function(spec) interpolate_spectrum(old_freq, spec, reference_freq)
  )
  
  tukey_spectra$allSpec <- as.matrix(interpolated_allSpec)
  tukey_spectra$freq <- reference_freq
  
  interpolated_tukey[[i]] <- tukey_spectra
}

all_match_after_interp <- all(
  sapply(interpolated_tukey, function(x) identical(x$freq, reference_freq))
)

all_match_after_interp

```

## Combine Spectra Across All Events

```{r}

all_click_metadata <- list()
all_click_spectra <- list()

for (i in seq_along(interpolated_tukey)) {
  
  tukey_spectra <- interpolated_tukey[[i]]
  
  event_id <- tukey_spectra$EventID
  uid <- tukey_spectra$UID
  
  click_spectra <- t(tukey_spectra$allSpec)
  
  click_meta <- data.frame(
    EventID = event_id,
    UID = uid,
    stringsAsFactors = FALSE
  )
  
  all_click_metadata[[i]] <- click_meta
  all_click_spectra[[i]] <- click_spectra
}

click_metadata <- do.call(rbind, all_click_metadata)
all_spec_matrix <- do.call(rbind, all_click_spectra)

dim(all_spec_matrix)
head(click_metadata)

```

## Restrict Spectra to 19–50 kHz

```{r}

keepFreq <- which(reference_freq >= freq_min & reference_freq <= freq_max)

AllSpec_Focus <- all_spec_matrix[, keepFreq]
centroid_freq <- reference_freq[keepFreq]

dim(AllSpec_Focus)
range(centroid_freq)

```

## Apply Row-Wise Min-Max Normalization

```{r}

minmax_normalize_rows <- function(spec_matrix) {
  
  row_min <- apply(spec_matrix, 1, min, na.rm = TRUE)
  row_max <- apply(spec_matrix, 1, max, na.rm = TRUE)
  row_range <- row_max - row_min
  
  normalized_matrix <- sweep(spec_matrix, 1, row_min, FUN = "-")
  normalized_matrix <- sweep(normalized_matrix, 1, row_range, FUN = "/")
  
  # If a click has zero range, all frequency bins are identical.
  # Dividing by zero creates NA/Inf, so set those rows to 0.
  normalized_matrix[!is.finite(normalized_matrix)] <- 0
  
  return(as.matrix(normalized_matrix))
}

AllSpec_minmax <- minmax_normalize_rows(AllSpec_Focus)

cat(
  "Min-max normalized matrix dimensions:",
  dim(AllSpec_minmax)[1],
  "clicks x",
  dim(AllSpec_minmax)[2],
  "frequency bins\n"
)

summary(as.vector(AllSpec_minmax))
range(AllSpec_minmax, na.rm = TRUE)

```

## Quick Sanity Check: Raw vs Min-Max Normalized Example Clicks

```{r}

set.seed(10)
example_clicks <- sample(seq_len(nrow(AllSpec_Focus)), size = min(6, nrow(AllSpec_Focus)))

raw_example_df <- data.frame(
  Frequency_kHz = rep(centroid_freq / 1000, times = length(example_clicks)),
  Intensity = as.vector(t(AllSpec_Focus[example_clicks, , drop = FALSE])),
  Click = factor(rep(example_clicks, each = length(centroid_freq)))
)

minmax_example_df <- data.frame(
  Frequency_kHz = rep(centroid_freq / 1000, times = length(example_clicks)),
  RelativeAmplitude = as.vector(t(AllSpec_minmax[example_clicks, , drop = FALSE])),
  Click = factor(rep(example_clicks, each = length(centroid_freq)))
)

raw_example_plot <- ggplot(
  raw_example_df,
  aes(x = Frequency_kHz, y = Intensity, group = Click, color = Click)
) +
  geom_line(alpha = 0.85, linewidth = 0.8) +
  labs(
    title = "Example Raw Spectra Before Min-Max Normalization",
    x = "Frequency (kHz)",
    y = "Intensity",
    color = "Example Click"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "right"
  )

minmax_example_plot <- ggplot(
  minmax_example_df,
  aes(x = Frequency_kHz, y = RelativeAmplitude, group = Click, color = Click)
) +
  geom_line(alpha = 0.85, linewidth = 0.8) +
  labs(
    title = "Same Example Spectra After Row-Wise Min-Max Normalization",
    x = "Frequency (kHz)",
    y = "Min-max normalized amplitude",
    color = "Example Click"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "right"
  )

print(raw_example_plot)
print(minmax_example_plot)

ggsave(
  file.path(dirs$plots, "sanity_check_raw_example_spectra.png"),
  raw_example_plot,
  width = 7,
  height = 5
)

ggsave(
  file.path(dirs$plots, "sanity_check_minmax_example_spectra.png"),
  minmax_example_plot,
  width = 7,
  height = 5
)

```

## Helper: Sample Up to N Clicks Per Event

```{r}

sample_by_event <- function(metadata_df, max_per_event, seed = NULL, max_total = NULL) {
  
  if (!is.null(seed)) set.seed(seed)
  
  split_indices <- split(seq_len(nrow(metadata_df)), metadata_df$EventID)
  
  sampled_indices <- unlist(
    lapply(split_indices, function(idx) {
      sample(idx, size = min(length(idx), max_per_event))
    }),
    use.names = FALSE
  )
  
  if (!is.null(max_total) && length(sampled_indices) > max_total) {
    sampled_indices <- sample(sampled_indices, size = max_total)
  }
  
  return(sampled_indices)
}

```

## Elbow Plot for Candidate K Values

```{r}

elbow_index <- sample_by_event(
  metadata_df = click_metadata,
  max_per_event = max_clicks_per_event,
  seed = 1,
  max_total = elbow_sample_cap
)

AllSpec_elbow <- AllSpec_minmax[elbow_index, ]

wss <- numeric(length(k_candidates))

for (i in seq_along(k_candidates)) {
  
  k <- k_candidates[i]
  
  km_tmp <- kmeans(
    AllSpec_elbow,
    centers = k,
    nstart = 20,
    algorithm = "Lloyd"
  )
  
  wss[i] <- km_tmp$tot.withinss
}

elbow_df <- data.frame(k = k_candidates, WSS = wss)

elbow_plot <- ggplot(elbow_df, aes(x = k, y = WSS)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Elbow Plot - Min-Max Normalized",
    x = "Number of clusters (k)",
    y = "Total within-cluster sum of squares"
  )

print(elbow_plot)

ggsave(
  file.path(dirs$plots, "elbow_minmax_normalized.png"),
  elbow_plot,
  width = 7,
  height = 5
)

write.csv(
  elbow_df,
  file = file.path(dirs$data, "elbow_wss_values.csv"),
  row.names = FALSE
)

```

## PCA Helper Functions

```{r}

calculate_pca_variance <- function(pca_object, n_components = 2) {
  
  eigenvalues <- pca_object$sdev^2
  
  data.frame(
    RetainedComponents = n_components,
    RetainedEigenvalueSum = sum(eigenvalues[1:n_components]),
    TotalEigenvalueSum = sum(eigenvalues),
    PercentVarianceExplained = 100 * sum(eigenvalues[1:n_components]) / sum(eigenvalues)
  )
}

get_nonzero_variance_columns <- function(mat) {
  col_sd <- apply(mat, 2, sd, na.rm = TRUE)
  is.finite(col_sd) & col_sd > 0
}
```

## Run K-Means Workflow for Multiple K Values

```{r}
k_values_to_run <- 2:12

for (chosen_k in k_values_to_run) {
  
  k_label <- paste0("k_", chosen_k)
  
  dirs <- create_output_structure(
    base_dir = base_output_dir,
    analysis_name = analysis_name,
    method = cluster_method,
    k_label = k_label
  )
  
  cat("\n============================\n")
  cat("Running k =", chosen_k, "\n")
  cat("============================\n")

# ----------------------------------------------------
  
## Evaluation Sample and K-Means Clustering

eval_index <- sample_by_event(
  metadata_df = click_metadata,
  max_per_event = max_clicks_per_event,
  seed = 2
)

AllSpec_eval <- AllSpec_minmax[eval_index, ]
click_metadata_eval <- click_metadata[eval_index, ]

cat("Eval sample size:", nrow(AllSpec_eval), "\n")

# -----------------------------
# KMEANS
# -----------------------------
set.seed(3)

if (run_kmeans_on_full_dataset) {
  
  km <- kmeans(
    AllSpec_minmax,
    centers = chosen_k,
    nstart = 50,
    algorithm = "Lloyd"
  )
  
  cluster_labels_full <- km$cluster
  cluster_labels_eval <- cluster_labels_full[eval_index]
  
} else {
  
  km <- kmeans(
    AllSpec_eval,
    centers = chosen_k,
    nstart = 50
  )
  
  cluster_labels_eval <- km$cluster
}

# ----------------------------------------------------

## Save Cluster Assignments

cluster_assignments_full <- click_metadata %>%
  mutate(cluster = cluster_labels_full)

write.csv(
  cluster_assignments_full,
  file = file.path(dirs$data, "cluster_assignments_full.csv"),
  row.names = FALSE
)

cluster_size_summary <- cluster_assignments_full %>%
  group_by(cluster) %>%
  summarise(
    n_clicks = n(),
    .groups = "drop"
  ) %>%
  mutate(percent_clicks = 100 * n_clicks / sum(n_clicks))

write.csv(
  cluster_size_summary,
  file = file.path(dirs$data, "cluster_size_summary_full.csv"),
  row.names = FALSE
)

print(cluster_size_summary)

# ----------------------------------------------------

## Silhouette Diagnostics

dist_eval <- dist(AllSpec_eval)

sil <- silhouette(cluster_labels_eval, dist_eval)
mean_sil <- mean(sil[, "sil_width"])

cat("Mean silhouette:", mean_sil, "\n")

# Save silhouette summary
write.csv(
  data.frame(
    Analysis = analysis_name,
    K = chosen_k,
    MeanSilhouette = mean_sil
  ),
  file = file.path(dirs$data, "silhouette_summary.csv"),
  row.names = FALSE
)

# Silhouette plot
sil_df <- as.data.frame(sil) %>%
  mutate(obs = row_number()) %>%
  group_by(cluster) %>%
  arrange(desc(sil_width), .by_group = TRUE) %>%
  mutate(order = row_number()) %>%
  ungroup()

sil_plot <- ggplot(sil_df, aes(x = order, y = sil_width)) +
  geom_col() +
  facet_wrap(~ cluster, scales = "free_x") +
  labs(title = "Silhouette - Min-Max Normalized")

print(sil_plot)

ggsave(
  file.path(dirs$plots, "silhouette_plot.png"),
  sil_plot,
  width = 8,
  height = 5
)

# ------------------------------------------------------------

## PCA Diagnostics and Scatterplots

# --------------------------------------------------
# Remove zero-variance columns before PCA
# --------------------------------------------------

pca_keep_cols <- get_nonzero_variance_columns(AllSpec_eval)

AllSpec_eval_pca <- AllSpec_eval[, pca_keep_cols, drop = FALSE]
AllSpec_minmax_pca <- AllSpec_minmax[, pca_keep_cols, drop = FALSE]
centroid_freq_pca <- centroid_freq[pca_keep_cols]

removed_cols <- sum(!pca_keep_cols)

cat(
  "PCA removed",
  removed_cols,
  "zero-variance frequency columns\n"
)

# --------------------------------------------------
# Covariance-style PCA
# --------------------------------------------------

pca_cov_eval <- prcomp(
  AllSpec_eval_pca,
  center = TRUE,
  scale. = FALSE
)

pca_cov_full <- prcomp(
  AllSpec_minmax_pca,
  center = TRUE,
  scale. = FALSE
)

cov_variance_df <- calculate_pca_variance(pca_cov_eval, n_components = 2)
cov_variance_df$Analysis <- analysis_name
cov_variance_df$PCA_Type <- "Covariance-style PCA"
cov_variance_df$ZeroVarianceColumnsRemoved <- removed_cols

# --------------------------------------------------
# Correlation PCA
# --------------------------------------------------

pca_cor_eval <- prcomp(
  AllSpec_eval_pca,
  center = TRUE,
  scale. = TRUE
)

pca_cor_full <- prcomp(
  AllSpec_minmax_pca,
  center = TRUE,
  scale. = TRUE
)

cor_variance_df <- calculate_pca_variance(pca_cor_eval, n_components = 2)
cor_variance_df$Analysis <- analysis_name
cor_variance_df$PCA_Type <- "Correlation PCA"
cor_variance_df$ZeroVarianceColumnsRemoved <- removed_cols

# --------------------------------------------------
# Save PCA variance summary
# --------------------------------------------------

pca_variance_summary <- rbind(cov_variance_df, cor_variance_df)

write.csv(
  pca_variance_summary,
  file = file.path(dirs$diagnostics, "pca_variance_explained_summary.csv"),
  row.names = FALSE
)

print(pca_variance_summary)

# ------------------------------------------------------------

## PCA Scatterplots

# --------------------------------------------------
# Covariance PCA Scatterplot
# --------------------------------------------------

scatter_cov_df <- data.frame(
  PC1 = pca_cov_eval$x[, 1],
  PC2 = pca_cov_eval$x[, 2],
  EventID = click_metadata_eval$EventID,
  UID = click_metadata_eval$UID,
  cluster = as.factor(cluster_labels_eval)
)

scatter_cov_plot <- ggplot(scatter_cov_df, aes(x = PC1, y = PC2, color = cluster)) +
  geom_point(alpha = 0.7, size = 1.2) +
  labs(
    title = "Covariance PCA Projection - Min-Max Normalized",
    subtitle = paste0(
      "PC1 + PC2 variance explained: ",
      round(cov_variance_df$PercentVarianceExplained, 2),
      "%"
    ),
    x = "PC1",
    y = "PC2",
    color = "Cluster"
  )

print(scatter_cov_plot)

ggsave(
  file.path(dirs$plots, "pca_scatter_covariance.png"),
  scatter_cov_plot,
  width = 7,
  height = 5
)

write.csv(
  scatter_cov_df,
  file = file.path(dirs$data, "pca_scatter_covariance_data.csv"),
  row.names = FALSE
)

# --------------------------------------------------
# Correlation PCA Scatterplot
# --------------------------------------------------

scatter_cor_df <- data.frame(
  PC1 = pca_cor_eval$x[, 1],
  PC2 = pca_cor_eval$x[, 2],
  EventID = click_metadata_eval$EventID,
  UID = click_metadata_eval$UID,
  cluster = as.factor(cluster_labels_eval)
)

scatter_cor_plot <- ggplot(scatter_cor_df, aes(x = PC1, y = PC2, color = cluster)) +
  geom_point(alpha = 0.7, size = 1.2) +
  labs(
    title = "Correlation PCA Projection - Min-Max Normalized",
    subtitle = paste0(
      "PC1 + PC2 variance explained: ",
      round(cor_variance_df$PercentVarianceExplained, 2),
      "%"
    ),
    x = "PC1",
    y = "PC2",
    color = "Cluster"
  )

print(scatter_cor_plot)

ggsave(
  file.path(dirs$plots, "pca_scatter_correlation.png"),
  scatter_cor_plot,
  width = 7,
  height = 5
)

write.csv(
  scatter_cor_df,
  file = file.path(dirs$data, "pca_scatter_correlation_data.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------

## Full Dataset PCA Loadings

cov_loadings_df <- data.frame(
  Frequency_kHz = centroid_freq_pca / 1000,
  PC1 = pca_cov_full$rotation[, 1],
  PC2 = pca_cov_full$rotation[, 2],
  PCA_Type = "Covariance-style PCA"
)

cor_loadings_df <- data.frame(
  Frequency_kHz = centroid_freq_pca / 1000,
  PC1 = pca_cor_full$rotation[, 1],
  PC2 = pca_cor_full$rotation[, 2],
  PCA_Type = "Correlation PCA"
)

loadings_df <- rbind(cov_loadings_df, cor_loadings_df)

write.csv(
  loadings_df,
  file = file.path(dirs$data, "full_dataset_pca_loadings.csv"),
  row.names = FALSE
)

cov_loadings_plot <- ggplot(cov_loadings_df, aes(x = Frequency_kHz)) +
  geom_line(aes(y = PC1, color = "PC1"), linewidth = 1) +
  geom_line(aes(y = PC2, color = "PC2"), linewidth = 1) +
  labs(
    title = "Full-Dataset Covariance PCA Loadings - Min-Max Normalized",
    x = "Frequency (kHz)",
    y = "Loading",
    color = "Component"
  )

print(cov_loadings_plot)

ggsave(
  file.path(dirs$plots, "full_dataset_pca_loadings_covariance.png"),
  cov_loadings_plot,
  width = 7,
  height = 5
)

cor_loadings_plot <- ggplot(cor_loadings_df, aes(x = Frequency_kHz)) +
  geom_line(aes(y = PC1, color = "PC1"), linewidth = 1) +
  geom_line(aes(y = PC2, color = "PC2"), linewidth = 1) +
  labs(
    title = "Full-Dataset Correlation PCA Loadings - Min-Max Normalized",
    x = "Frequency (kHz)",
    y = "Loading",
    color = "Component"
  )

print(cor_loadings_plot)

ggsave(
  file.path(dirs$plots, "full_dataset_pca_loadings_correlation.png"),
  cor_loadings_plot,
  width = 7,
  height = 5
)

# ------------------------------------------------------------

## Centroid Spectra

centroid_df <- data.frame(
  Frequency_kHz = centroid_freq / 1000,
  t(km$centers)
)

colnames(centroid_df) <- c(
  "Frequency_kHz",
  paste0("Cluster_", 1:nrow(km$centers))
)

write.csv(
  centroid_df,
  file = file.path(dirs$data, "centroid_spectra.csv"),
  row.names = FALSE
)

centroid_plot_file <- file.path(dirs$plots, "centroid_plot.png")

png(
  filename = centroid_plot_file,
  width = 7,
  height = 5,
  units = "in",
  res = 300
)

matplot(
  x = centroid_freq / 1000,
  y = t(km$centers),
  type = "l",
  lty = 1,
  main = "Cluster Centroids - Min-Max Normalized",
  xlab = "Frequency (kHz)",
  ylab = "Min-max normalized amplitude"
)

dev.off()

# ------------------------------------------------------------

## Cluster Variability


if (run_kmeans_on_full_dataset) {
  
  cluster_labels_full <- km$cluster
  
  dist_to_centroid <- numeric(nrow(AllSpec_minmax))
  
  for (i in seq_len(nrow(AllSpec_minmax))) {
    
    this_cluster <- cluster_labels_full[i]
    this_centroid <- km$centers[this_cluster, ]
    
    dist_to_centroid[i] <- sqrt(
      sum((AllSpec_minmax[i, ] - this_centroid)^2)
    )
  }
  
  variability_df <- data.frame(
    cluster = cluster_labels_full,
    dist = dist_to_centroid
  ) %>%
    group_by(cluster) %>%
    summarise(
      n_clicks = n(),
      mean_dist = mean(dist),
      sd_dist = sd(dist),
      median_dist = median(dist),
      .groups = "drop"
    )
  
  write.csv(
    variability_df,
    file = file.path(dirs$data, "cluster_variability.csv"),
    row.names = FALSE
  )
  
  print(variability_df)
}

# ------------------------------------------------------------

## Concatenated Spectrograms

cluster_ids <- sort(unique(cluster_labels_eval))

spec_cols <- colorRampPalette(
  c("white", "#FFD54F", "#FF9800", "#D84315", "#4E0012")
)(200)

for (cl in cluster_ids) {
  
  cluster_idx <- which(cluster_labels_eval == cl)
  
  if (length(cluster_idx) == 0) next
  
  cluster_spec <- AllSpec_eval[cluster_idx, , drop = FALSE]
  
  out_file <- file.path(
    dirs$spectrograms,
    paste0(
      "concatenated_spectrogram_eval_19_50kHz_",
      k_label,
      "_cluster_",
      cl,
      ".png"
    )
  )
  
  png(out_file, width = 1800, height = 900)
  
  image(
    x = seq_len(nrow(cluster_spec)),
    y = centroid_freq / 1000,
    z = cluster_spec,
    col = spec_cols,
    xlab = "Click Number",
    ylab = "Frequency (kHz)",
    main = paste(
      "Concatenated Click Spectrogram - Min-Max Normalized - Cluster",
      cl
    ),
    useRaster = TRUE,
    axes = FALSE
  )
  
  axis(1)
  axis(2, at = seq(20, 50, by = 10))
  box()
  
  dev.off()
  
  cat("Saved:", out_file, "\n")
}

cat("Finished k =", chosen_k, "\n")
}
```

## Regenerate Labeled Centroid Plots from Existing Environment

```{r}

library(tidyr)

for (chosen_k in 2:12) {
  
  k_label <- paste0("k_", chosen_k)
  
  dirs <- create_output_structure(
    base_dir = base_output_dir,
    analysis_name = analysis_name,
    method = cluster_method,
    k_label = k_label
  )
  
  centroid_csv <- file.path(dirs$data, "centroid_spectra.csv")
  
  if (!file.exists(centroid_csv)) {
    cat("Missing centroid file for", k_label, "\n")
    next
  }
  
  centroid_df <- read.csv(centroid_csv)
  
  centroid_long <- centroid_df %>%
    tidyr::pivot_longer(
      cols = starts_with("Cluster_"),
      names_to = "Cluster",
      values_to = "MinMaxAmplitude"
    )
  
  centroid_labeled_plot <- ggplot(
    centroid_long,
    aes(x = Frequency_kHz, y = MinMaxAmplitude, color = Cluster)
  ) +
    geom_line(linewidth = 1) +
    labs(
      title = paste("Cluster Centroids - Min-Max Normalized -", k_label),
      x = "Frequency (kHz)",
      y = "Min-max normalized amplitude",
      color = "Cluster"
    ) +
    theme_bw()
  
  ggsave(
    file.path(dirs$plots, "centroid_plot_labeled.png"),
    centroid_labeled_plot,
    width = 8,
    height = 5
  )
  
  cat("Saved labeled centroid plot for", k_label, "\n")
}

```

## Generate Offset/Stacked Centroid Plots for k = 2:12

```{r}

library(tidyr)

for (chosen_k in 2:12) {
  
  k_label <- paste0("k_", chosen_k)
  
  dirs <- create_output_structure(
    base_dir = base_output_dir,
    analysis_name = analysis_name,
    method = cluster_method,
    k_label = k_label
  )
  
  centroid_csv <- file.path(dirs$data, "centroid_spectra.csv")
  
  if (!file.exists(centroid_csv)) {
    cat("Missing centroid file for", k_label, "\n")
    next
  }
  
  centroid_df <- read.csv(centroid_csv)
  
  centroid_long <- centroid_df %>%
    pivot_longer(
      cols = starts_with("Cluster_"),
      names_to = "Cluster",
      values_to = "Amplitude"
    ) %>%
    mutate(
      ClusterNumber = as.numeric(gsub("Cluster_", "", Cluster)),
      # Stack clusters vertically, with Cluster 1 at the top
      Offset = chosen_k - ClusterNumber + 1,
      PlotValue = Amplitude + Offset
    )
  
  centroid_offset_plot <- ggplot(
    centroid_long,
    aes(
      x = Frequency_kHz,
      y = PlotValue,
      group = Cluster,
      color = ClusterNumber
    )
  ) +
    geom_line(linewidth = 1.1) +
    scale_y_continuous(
      breaks = chosen_k:1,
      labels = 1:chosen_k
    ) +
    scale_x_continuous(
      limits = c(19, 50),
      breaks = seq(20, 50, by = 10)
    ) +
    labs(
      title = paste("Offset Cluster Centroids -", k_label),
      x = "Frequency (kHz)",
      y = "Cluster",
      color = "Cluster"
    ) +
    theme_classic() +
    theme(
      plot.title = element_text(face = "bold"),
      axis.title = element_text(size = 14),
      axis.text = element_text(size = 12),
      legend.position = "none"
    )
  
  ggsave(
    file.path(dirs$plots, "centroid_plot_offset_labeled.png"),
    centroid_offset_plot,
    width = 4,
    height = max(5, chosen_k * 0.55),
    dpi = 300
  )
  
  cat("Saved offset centroid plot for", k_label, "\n")
}

```







