---
title: "Clustering_kshape_Normalized"
author: "Sarah Moseley"
date: "2026-04-28"
output: html_document
---

```{r}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

```

## Setup and User Settings

```{r}

library(ggplot2)
library(dplyr)
library(cluster)
library(dtwclust)
library(proxy)
library(tidyr)

# -----------------------------
# Base Output Directory
# -----------------------------
base_output_dir <- "E:/Analysis/Clustering Output/Normalized"
dir.create(base_output_dir, recursive = TRUE, showWarnings = FALSE)

# -----------------------------
# Analysis Variants
# -----------------------------
analysis_types <- list(
  "Raw Normalized" = list(type = "raw_norm"),
  "Normalized Whole Spectrum Scaling" = list(type = "row_scale"),
  "Normalized Frequency Scaling" = list(type = "col_scale")
)

# -----------------------------
# Clustering Settings
# -----------------------------
cluster_method <- "kshape"

freq_min <- 19000
freq_max <- 50000

k_candidates <- 2:6
chosen_k <- 2
k_label <- paste0("k_", chosen_k)

max_clicks_per_event <- 200
eval_sample_cap <- NULL

# k-shape settings
kshape_nrep_elbow <- 1L
kshape_nrep_final <- 1L

# Full dataset assignment settings
assignment_batch_size <- 5000

# -----------------------------
# Helper: Create Structured Output Folders
# -----------------------------
create_output_structure <- function(base_dir, analysis_name, method, k_label) {
  
  analysis_dir <- file.path(base_dir, analysis_name)
  method_dir <- file.path(analysis_dir, paste(method, k_label))
  
  dirs <- list(
    root = method_dir,
    plots = file.path(method_dir, "plots"),
    data = file.path(method_dir, "data"),
    diagnostics = file.path(method_dir, "diagnostics"),
    spectrograms = file.path(method_dir, "concatenated_spectrograms")
  )
  
  lapply(dirs, function(d) {
    dir.create(d, recursive = TRUE, showWarnings = FALSE)
  })
  
  return(dirs)
}

```

## Load Combined Tukey Spectra

```{r}

combined_file <- file.choose()
combined_tukey <- readRDS(combined_file)

reference_freq <- combined_tukey[[1]]$freq

```

## Interpolate Spectra to Shared Grid

```{r}

interpolate_spectrum <- function(old_freq, old_spec, new_freq) {
  approx(x = old_freq, y = old_spec, xout = new_freq, rule = 2)$y
}

interpolated_tukey <- vector("list", length(combined_tukey))

for (i in seq_along(combined_tukey)) {
  
  tukey_spectra <- combined_tukey[[i]]
  
  interpolated_allSpec <- apply(
    tukey_spectra$allSpec,
    2,
    function(spec) interpolate_spectrum(
      tukey_spectra$freq,
      spec,
      reference_freq
    )
  )
  
  tukey_spectra$allSpec <- as.matrix(interpolated_allSpec)
  tukey_spectra$freq <- reference_freq
  
  interpolated_tukey[[i]] <- tukey_spectra
}

all_match_after_interp <- all(
  sapply(interpolated_tukey, function(x) identical(x$freq, reference_freq))
)

all_match_after_interp

```

## Combine Spectra

```{r}

all_click_metadata <- list()
all_click_spectra <- list()

for (i in seq_along(interpolated_tukey)) {
  
  tukey_spectra <- interpolated_tukey[[i]]
  
  click_meta <- data.frame(
    EventID = tukey_spectra$EventID,
    UID = tukey_spectra$UID,
    stringsAsFactors = FALSE
  )
  
  all_click_metadata[[i]] <- click_meta
  all_click_spectra[[i]] <- t(tukey_spectra$allSpec)
}

click_metadata <- do.call(rbind, all_click_metadata)
all_spec_matrix <- do.call(rbind, all_click_spectra)

```

## Restrict to 19–50 kHz

```{r}

keepFreq <- which(reference_freq >= freq_min & reference_freq <= freq_max)

AllSpec_Focus <- all_spec_matrix[, keepFreq]
centroid_freq <- reference_freq[keepFreq]

```

## Normalization

```{r}

normalize_to_zero_baseline <- function(spec_matrix) {
  first_val <- spec_matrix[, 1]
  sweep(spec_matrix, 1, first_val, FUN = "-")
}

prepare_analysis_matrix <- function(spec_matrix, analysis_type) {
  
  norm <- normalize_to_zero_baseline(spec_matrix)
  
  if (analysis_type == "raw_norm") {
    
    out <- norm
    
  } else if (analysis_type == "row_scale") {
    
    out <- t(scale(t(norm)))
    out[is.na(out)] <- 0
    
  } else if (analysis_type == "col_scale") {
    
    out <- scale(norm)
    out[is.na(out)] <- 0
    
  } else {
    stop("Unknown analysis_type")
  }
  
  return(as.matrix(out))
}

analysis_matrices <- list()

for (analysis_name in names(analysis_types)) {
  
  type <- analysis_types[[analysis_name]]$type
  
  analysis_matrices[[analysis_name]] <- prepare_analysis_matrix(
    AllSpec_Focus,
    type
  )
  
  cat(
    analysis_name, ":",
    dim(analysis_matrices[[analysis_name]])[1],
    "x",
    dim(analysis_matrices[[analysis_name]])[2], "\n"
  )
}

```

## Helper: Sample by Event (same as kmeans)

```{r}

sample_by_event <- function(metadata_df, max_per_event, seed = NULL) {
  
  if (!is.null(seed)) set.seed(seed)
  
  split_indices <- split(seq_len(nrow(metadata_df)), metadata_df$EventID)
  
  sampled_indices <- unlist(
    lapply(split_indices, function(idx) {
      sample(idx, size = min(length(idx), max_per_event))
    }),
    use.names = FALSE
  )
  
  return(sampled_indices)
}

```

## Run Analysis on Each Normalization Type

```{r}

results_list <- list()

for (analysis_name in names(analysis_matrices)) {
  
  cat("\n============================\n")
  cat("Running k-shape:", analysis_name, "\n")
  cat("============================\n")
  
  AllSpec_current <- analysis_matrices[[analysis_name]]
  
  dirs <- create_output_structure(
    base_dir = base_output_dir,
    analysis_name = analysis_name,
    method = cluster_method,
    k_label = k_label
  )
  
  # --------------------------------------------------
  # Evaluation Sample
  # --------------------------------------------------
  
  eval_index <- sample_by_event(
    metadata_df = click_metadata,
    max_per_event = max_clicks_per_event,
    seed = 2
  )
  
  AllSpec_eval <- AllSpec_current[eval_index, , drop = FALSE]
  click_metadata_eval <- click_metadata[eval_index, ]
  
  cat("Eval sample size:", nrow(AllSpec_eval), "\n")
  
  # --------------------------------------------------
  # Elbow-like Curve Using Total Within-Cluster SBD
  # --------------------------------------------------
  
  elbow_vals <- numeric(length(k_candidates))
  
  for (i in seq_along(k_candidates)) {
    
    k <- k_candidates[i]
    
    cat("Running elbow k =", k, "for", analysis_name, "\n")
    
    ks_tmp <- dtwclust::tsclust(
      series = lapply(seq_len(nrow(AllSpec_eval)), function(j) as.numeric(AllSpec_eval[j, ])),
      type = "partitional",
      k = k,
      preproc = dtwclust::zscore,
      distance = "sbd",
      centroid = "shape",
      seed = 100 + k,
      trace = FALSE,
      control = dtwclust::partitional_control(nrep = kshape_nrep_elbow)
    )
    
    labels_tmp <- as.integer(predict(ks_tmp))
    centroids_tmp <- ks_tmp@centroids
    
    total_within_sbd <- 0
    
    for (j in seq_len(nrow(AllSpec_eval))) {
      
      cl <- labels_tmp[j]
      
      d_j <- as.numeric(proxy::dist(
        x = matrix(as.numeric(AllSpec_eval[j, ]), nrow = 1),
        y = matrix(as.numeric(centroids_tmp[[cl]]), nrow = 1),
        method = "SBD"
      ))
      
      total_within_sbd <- total_within_sbd + d_j
    }
    
    elbow_vals[i] <- total_within_sbd
    
    cat("k =", k, "total within-cluster SBD =", total_within_sbd, "\n")
  }
  
  elbow_df <- data.frame(
    k = k_candidates,
    TotalWithinSBD = elbow_vals
  )
  
  elbow_plot <- ggplot(elbow_df, aes(x = k, y = TotalWithinSBD)) +
    geom_line() +
    geom_point() +
    labs(
      title = paste("Elbow-like Plot for k-shape -", analysis_name),
      x = "Number of Clusters",
      y = "Total Within-Cluster SBD"
    )
  
  ggsave(
    file.path(dirs$plots, "elbow.png"),
    elbow_plot,
    width = 7,
    height = 5
  )
  
  write.csv(
    elbow_df,
    file = file.path(dirs$data, "elbow_data.csv"),
    row.names = FALSE
  )
  
  # --------------------------------------------------
  # Final k-shape Model on Evaluation Sample
  # --------------------------------------------------
  
  cat("Running final k-shape model for", analysis_name, "\n")
  
  kshape_model <- dtwclust::tsclust(
    series = lapply(seq_len(nrow(AllSpec_eval)), function(j) as.numeric(AllSpec_eval[j, ])),
    type = "partitional",
    k = chosen_k,
    preproc = dtwclust::zscore,
    distance = "sbd",
    centroid = "shape",
    seed = 999,
    trace = FALSE,
    control = dtwclust::partitional_control(nrep = kshape_nrep_final)
  )
  
  cluster_labels_eval <- as.integer(predict(kshape_model))
  centroids <- kshape_model@centroids
  
  centroid_matrix <- do.call(
    rbind,
    lapply(centroids, function(x) as.numeric(x))
  )
  
  # --------------------------------------------------
  # Assign Full Dataset to k-shape Centroids in Batches
  # --------------------------------------------------
  
  assign_clusters_sbd_batch <- function(data_matrix, centroid_matrix, batch_size = 5000) {
    
    n <- nrow(data_matrix)
    assignments <- integer(n)
    
    for (start_idx in seq(1, n, by = batch_size)) {
      
      end_idx <- min(start_idx + batch_size - 1, n)
      
      batch <- data_matrix[start_idx:end_idx, , drop = FALSE]
      
      batch_assign <- apply(batch, 1, function(x) {
        
        dists <- sapply(seq_len(nrow(centroid_matrix)), function(cl) {
          
          as.numeric(proxy::dist(
            x = matrix(as.numeric(x), nrow = 1),
            y = matrix(as.numeric(centroid_matrix[cl, ]), nrow = 1),
            method = "SBD"
          ))
        })
        
        which.min(dists)
      })
      
      assignments[start_idx:end_idx] <- batch_assign
      
      cat("Assigned rows", start_idx, "to", end_idx, "\n")
    }
    
    return(assignments)
  }
  
  cluster_labels_full <- assign_clusters_sbd_batch(
    data_matrix = AllSpec_current,
    centroid_matrix = centroid_matrix,
    batch_size = assignment_batch_size
  )
  
  # --------------------------------------------------
  # Silhouette on Evaluation Sample
  # --------------------------------------------------
  
  cat("Computing SBD distance matrix for silhouette...\n")
  
  dist_eval <- proxy::dist(
    x = AllSpec_eval,
    method = "SBD"
  )
  
  sil <- cluster::silhouette(cluster_labels_eval, dist_eval)
  mean_sil <- mean(sil[, "sil_width"])
  
  cat("Mean silhouette:", mean_sil, "\n")
  
  write.csv(
    data.frame(
      Analysis = analysis_name,
      K = chosen_k,
      MeanSilhouette = mean_sil
    ),
    file = file.path(dirs$data, "silhouette_summary.csv"),
    row.names = FALSE
  )
  
  sil_df <- as.data.frame(sil) %>%
    mutate(obs = row_number()) %>%
    group_by(cluster) %>%
    arrange(desc(sil_width), .by_group = TRUE) %>%
    mutate(order = row_number()) %>%
    ungroup()
  
  sil_plot <- ggplot(sil_df, aes(x = order, y = sil_width)) +
    geom_col() +
    facet_wrap(~ cluster, scales = "free_x") +
    labs(
      title = paste("Silhouette -", analysis_name),
      x = "Observations",
      y = "Silhouette Width"
    )
  
  ggsave(
    file.path(dirs$plots, "silhouette_plot.png"),
    sil_plot,
    width = 8,
    height = 5
  )
  
  # --------------------------------------------------
  # Store Results for Later Chunks
  # --------------------------------------------------
  
  results_list[[analysis_name]] <- list(
    dirs = dirs,
    kshape_model = kshape_model,
    centroids = centroids,
    centroid_matrix = centroid_matrix,
    eval_index = eval_index,
    cluster_labels_eval = cluster_labels_eval,
    cluster_labels_full = cluster_labels_full,
    AllSpec_eval = AllSpec_eval,
    AllSpec_current = AllSpec_current,
    mean_silhouette = mean_sil
  )
  
  cat("Finished k-shape core loop for:", analysis_name, "\n")
}

```

## PCA Diagnostics and Plots for Each Analysis Type

```{r}

calculate_pca_variance <- function(pca_object, n_components = 2) {
  
  eigenvalues <- pca_object$sdev^2
  
  data.frame(
    RetainedComponents = n_components,
    RetainedEigenvalueSum = sum(eigenvalues[1:n_components]),
    TotalEigenvalueSum = sum(eigenvalues),
    PercentVarianceExplained = 100 * sum(eigenvalues[1:n_components]) / sum(eigenvalues)
  )
}

get_nonzero_variance_columns <- function(mat) {
  col_sd <- apply(mat, 2, sd, na.rm = TRUE)
  is.finite(col_sd) & col_sd > 0
}

for (analysis_name in names(results_list)) {
  
  cat("\nRunning PCA diagnostics for:", analysis_name, "\n")
  
  result <- results_list[[analysis_name]]
  
  dirs <- result$dirs
  AllSpec_eval <- result$AllSpec_eval
  AllSpec_current <- result$AllSpec_current
  eval_index <- result$eval_index
  cluster_labels_eval <- result$cluster_labels_eval
  
  click_metadata_eval <- click_metadata[eval_index, ]
  
  # --------------------------------------------------
  # Remove zero-variance columns before PCA
  # --------------------------------------------------
  
  pca_keep_cols <- get_nonzero_variance_columns(AllSpec_eval)
  
  AllSpec_eval_pca <- AllSpec_eval[, pca_keep_cols, drop = FALSE]
  AllSpec_current_pca <- AllSpec_current[, pca_keep_cols, drop = FALSE]
  centroid_freq_pca <- centroid_freq[pca_keep_cols]
  
  removed_cols <- sum(!pca_keep_cols)
  
  cat(
    "PCA removed",
    removed_cols,
    "zero-variance frequency columns for",
    analysis_name,
    "\n"
  )
  
  # --------------------------------------------------
  # Covariance-style PCA
  # --------------------------------------------------
  
  pca_cov_eval <- prcomp(
    AllSpec_eval_pca,
    center = FALSE,
    scale. = FALSE
  )
  
  pca_cov_full <- prcomp(
    AllSpec_current_pca,
    center = FALSE,
    scale. = FALSE
  )
  
  cov_variance_df <- calculate_pca_variance(pca_cov_eval, n_components = 2)
  cov_variance_df$Analysis <- analysis_name
  cov_variance_df$PCA_Type <- "Covariance-style PCA"
  cov_variance_df$ZeroVarianceColumnsRemoved <- removed_cols
  
  # --------------------------------------------------
  # Correlation PCA
  # --------------------------------------------------
  
  pca_cor_eval <- prcomp(
    AllSpec_eval_pca,
    center = TRUE,
    scale. = TRUE
  )
  
  pca_cor_full <- prcomp(
    AllSpec_current_pca,
    center = TRUE,
    scale. = TRUE
  )
  
  cor_variance_df <- calculate_pca_variance(pca_cor_eval, n_components = 2)
  cor_variance_df$Analysis <- analysis_name
  cor_variance_df$PCA_Type <- "Correlation PCA"
  cor_variance_df$ZeroVarianceColumnsRemoved <- removed_cols
  
  # --------------------------------------------------
  # Save PCA variance summary
  # --------------------------------------------------
  
  pca_variance_summary <- rbind(cov_variance_df, cor_variance_df)
  
  write.csv(
    pca_variance_summary,
    file = file.path(dirs$diagnostics, "pca_variance_explained_summary.csv"),
    row.names = FALSE
  )
  
  print(pca_variance_summary)
  
  # --------------------------------------------------
  # Covariance PCA Scatterplot
  # --------------------------------------------------
  
  scatter_cov_df <- data.frame(
    PC1 = pca_cov_eval$x[, 1],
    PC2 = pca_cov_eval$x[, 2],
    EventID = click_metadata_eval$EventID,
    UID = click_metadata_eval$UID,
    cluster = as.factor(cluster_labels_eval)
  )
  
  scatter_cov_plot <- ggplot(scatter_cov_df, aes(x = PC1, y = PC2, color = cluster)) +
    geom_point(alpha = 0.7, size = 1.2) +
    labs(
      title = paste("Covariance PCA Projection -", analysis_name),
      subtitle = paste0(
        "PC1 + PC2 variance explained: ",
        round(cov_variance_df$PercentVarianceExplained, 2),
        "%"
      ),
      x = "PC1",
      y = "PC2",
      color = "Cluster"
    )
  
  print(scatter_cov_plot)
  
  ggsave(
    file.path(dirs$plots, "pca_scatter_covariance.png"),
    scatter_cov_plot,
    width = 7,
    height = 5
  )
  
  write.csv(
    scatter_cov_df,
    file = file.path(dirs$data, "pca_scatter_covariance_data.csv"),
    row.names = FALSE
  )
  
  # --------------------------------------------------
  # Correlation PCA Scatterplot
  # --------------------------------------------------
  
  scatter_cor_df <- data.frame(
    PC1 = pca_cor_eval$x[, 1],
    PC2 = pca_cor_eval$x[, 2],
    EventID = click_metadata_eval$EventID,
    UID = click_metadata_eval$UID,
    cluster = as.factor(cluster_labels_eval)
  )
  
  scatter_cor_plot <- ggplot(scatter_cor_df, aes(x = PC1, y = PC2, color = cluster)) +
    geom_point(alpha = 0.7, size = 1.2) +
    labs(
      title = paste("Correlation PCA Projection -", analysis_name),
      subtitle = paste0(
        "PC1 + PC2 variance explained: ",
        round(cor_variance_df$PercentVarianceExplained, 2),
        "%"
      ),
      x = "PC1",
      y = "PC2",
      color = "Cluster"
    )
  
  print(scatter_cor_plot)
  
  ggsave(
    file.path(dirs$plots, "pca_scatter_correlation.png"),
    scatter_cor_plot,
    width = 7,
    height = 5
  )
  
  write.csv(
    scatter_cor_df,
    file = file.path(dirs$data, "pca_scatter_correlation_data.csv"),
    row.names = FALSE
  )
  
  # --------------------------------------------------
  # Full Dataset PCA Loadings
  # --------------------------------------------------
  
  cov_loadings_df <- data.frame(
    Frequency_kHz = centroid_freq_pca / 1000,
    PC1 = pca_cov_full$rotation[, 1],
    PC2 = pca_cov_full$rotation[, 2],
    PCA_Type = "Covariance-style PCA"
  )
  
  cor_loadings_df <- data.frame(
    Frequency_kHz = centroid_freq_pca / 1000,
    PC1 = pca_cor_full$rotation[, 1],
    PC2 = pca_cor_full$rotation[, 2],
    PCA_Type = "Correlation PCA"
  )
  
  loadings_df <- rbind(cov_loadings_df, cor_loadings_df)
  
  write.csv(
    loadings_df,
    file = file.path(dirs$data, "full_dataset_pca_loadings.csv"),
    row.names = FALSE
  )
  
  cov_loadings_plot <- ggplot(cov_loadings_df, aes(x = Frequency_kHz)) +
    geom_line(aes(y = PC1, color = "PC1"), linewidth = 1) +
    geom_line(aes(y = PC2, color = "PC2"), linewidth = 1) +
    labs(
      title = paste("Full-Dataset Covariance PCA Loadings -", analysis_name),
      x = "Frequency (kHz)",
      y = "Loading",
      color = "Component"
    )
  
  print(cov_loadings_plot)
  
  ggsave(
    file.path(dirs$plots, "full_dataset_pca_loadings_covariance.png"),
    cov_loadings_plot,
    width = 7,
    height = 5
  )
  
  cor_loadings_plot <- ggplot(cor_loadings_df, aes(x = Frequency_kHz)) +
    geom_line(aes(y = PC1, color = "PC1"), linewidth = 1) +
    geom_line(aes(y = PC2, color = "PC2"), linewidth = 1) +
    labs(
      title = paste("Full-Dataset Correlation PCA Loadings -", analysis_name),
      x = "Frequency (kHz)",
      y = "Loading",
      color = "Component"
    )
  
  print(cor_loadings_plot)
  
  ggsave(
    file.path(dirs$plots, "full_dataset_pca_loadings_correlation.png"),
    cor_loadings_plot,
    width = 7,
    height = 5
  )
  
  cat("Saved PCA outputs for:", analysis_name, "\n")
}

```

## Centroids, Variability, Summary Tables, and Concatenated Spectrograms

```{r}

## Centroids, Variability, Summary Tables, and Concatenated Spectrograms

for (analysis_name in names(results_list)) {
  
  cat("\nRunning centroid + spectrogram outputs for:", analysis_name, "\n")
  
  result <- results_list[[analysis_name]]
  
  dirs <- result$dirs
  centroid_matrix <- result$centroid_matrix
  cluster_labels_eval <- result$cluster_labels_eval
  cluster_labels_full <- result$cluster_labels_full
  AllSpec_eval <- result$AllSpec_eval
  AllSpec_current <- result$AllSpec_current
  mean_silhouette <- result$mean_silhouette
  
  # --------------------------------------------------
  # Centroid Spectra
  # --------------------------------------------------
  
  centroid_df <- data.frame(
    Frequency_kHz = centroid_freq / 1000,
    t(centroid_matrix)
  )
  
  colnames(centroid_df) <- c(
    "Frequency_kHz",
    paste0("Cluster_", seq_len(nrow(centroid_matrix)))
  )
  
  write.csv(
    centroid_df,
    file = file.path(dirs$data, "centroid_spectra.csv"),
    row.names = FALSE
  )
  
  centroid_plot_file <- file.path(dirs$plots, "centroid_plot.png")
  
  png(
    filename = centroid_plot_file,
    width = 7,
    height = 5,
    units = "in",
    res = 300
  )
  
  matplot(
    x = centroid_freq / 1000,
    y = t(centroid_matrix),
    type = "l",
    lty = 1,
    main = paste("k-shape Cluster Centroids -", analysis_name),
    xlab = "Frequency (kHz)",
    ylab = "Shape-based Intensity"
  )
  
  dev.off()
  
  # --------------------------------------------------
  # Cluster Variability using SBD
  # --------------------------------------------------
  
  dist_to_centroid <- numeric(nrow(AllSpec_current))
  
  for (i in seq_len(nrow(AllSpec_current))) {
    
    this_cluster <- cluster_labels_full[i]
    this_centroid <- centroid_matrix[this_cluster, ]
    
    dist_to_centroid[i] <- as.numeric(proxy::dist(
      x = matrix(as.numeric(AllSpec_current[i, ]), nrow = 1),
      y = matrix(as.numeric(this_centroid), nrow = 1),
      method = "SBD"
    ))
  }
  
  variability_df <- data.frame(
    cluster = cluster_labels_full,
    dist = dist_to_centroid
  ) %>%
    group_by(cluster) %>%
    summarise(
      n_clicks = n(),
      mean_dist = mean(dist),
      sd_dist = sd(dist),
      median_dist = median(dist),
      .groups = "drop"
    )
  
  write.csv(
    variability_df,
    file = file.path(dirs$data, "cluster_variability.csv"),
    row.names = FALSE
  )
  
  print(variability_df)
  
  # --------------------------------------------------
  # Summary Table
  # --------------------------------------------------
  
  summary_table <- data.frame(
    Method = cluster_method,
    Analysis = analysis_name,
    KLabel = k_label,
    ChosenK = chosen_k,
    MeanSilhouette = mean_silhouette,
    MaxClicksPerEvent = max_clicks_per_event,
    EvalSampleSizeUsed = nrow(AllSpec_eval),
    TotalClicksInFullDataset = nrow(AllSpec_current),
    AssignmentBatchSize = assignment_batch_size,
    DistanceMetric = "SBD"
  )
  
  write.csv(
    summary_table,
    file = file.path(dirs$data, "summary_table.csv"),
    row.names = FALSE
  )
  
  print(summary_table)
  
  # --------------------------------------------------
  # Concatenated Spectrograms
  # --------------------------------------------------
  
  cluster_ids <- sort(unique(cluster_labels_eval))
  
  spec_cols <- colorRampPalette(
    c("white", "#FFD54F", "#FF9800", "#D84315", "#4E0012")
  )(200)
  
  for (cl in cluster_ids) {
    
    cluster_idx <- which(cluster_labels_eval == cl)
    
    if (length(cluster_idx) == 0) next
    
    cluster_spec <- AllSpec_eval[cluster_idx, , drop = FALSE]
    
    out_file <- file.path(
      dirs$spectrograms,
      paste0(
        "concatenated_spectrogram_eval_19_50kHz_",
        k_label,
        "_cluster_",
        cl,
        ".png"
      )
    )
    
    png(out_file, width = 1800, height = 900)
    
    image(
      x = seq_len(nrow(cluster_spec)),
      y = centroid_freq / 1000,
      z = cluster_spec,
      col = spec_cols,
      xlab = "Click Number",
      ylab = "Frequency (kHz)",
      main = paste(
        "Concatenated Click Spectrogram -",
        analysis_name,
        "- Cluster",
        cl
      ),
      useRaster = TRUE,
      axes = FALSE
    )
    
    axis(1)
    axis(2, at = seq(20, 50, by = 10))
    box()
    
    dev.off()
    
    cat("Saved:", out_file, "\n")
  }
  
  cat("Finished:", analysis_name, "\n")
}

```

## Helper: Original ggplot-style k-shape centroid plot

```{r}

plot_centroids_kshape_ggplot <- function(centroid_df, k, analysis_name) {
  
  centroid_long <- tidyr::pivot_longer(
    centroid_df,
    cols = starts_with("Cluster_"),
    names_to = "Cluster",
    values_to = "ScaledIntensity"
  )
  
  ggplot(centroid_long, aes(x = Frequency_kHz, y = ScaledIntensity, color = Cluster)) +
    geom_line(linewidth = 0.8) +
    labs(
      title = paste("k-shape Centroids (k =", k, ") -", analysis_name),
      x = "Frequency (kHz)",
      y = "Scaled Intensity",
      color = "Cluster"
    )
}

```






















































