---
title: "Clustering_kmeans_Normalized"
author: "Sarah Moseley"
date: "2026-04-26"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

## Setup and User Settings

```{r}

library(ggplot2)
library(dplyr)
library(cluster)

# -----------------------------
# Base Output Directory
# -----------------------------
base_output_dir <- "E:/Analysis/Clustering Output/Normalized"
dir.create(base_output_dir, recursive = TRUE, showWarnings = FALSE)

# -----------------------------
# Analysis Variants
# -----------------------------
analysis_types <- list(
  "Raw Normalized" = list(type = "raw_norm"),
  "Normalized Whole Spectrum Scaling" = list(type = "row_scale"),
  "Normalized Frequency Scaling" = list(type = "col_scale")
)

# -----------------------------
# Clustering Settings
# -----------------------------
cluster_method <- "kmeans"

freq_min <- 19000
freq_max <- 50000

k_candidates <- 2:15
chosen_k <- 2
k_label <- paste0("k_", chosen_k)

max_clicks_per_event <- 200
elbow_sample_cap <- 20000
eval_sample_cap <- NULL

run_kmeans_on_full_dataset <- TRUE

# -----------------------------
# Helper: Create Structured Output Folders
# -----------------------------
create_output_structure <- function(base_dir, analysis_name, method, k_label) {
  
  analysis_dir <- file.path(base_dir, analysis_name)
  method_dir <- file.path(analysis_dir, paste(method, k_label))
  
  dirs <- list(
    root = method_dir,
    plots = file.path(method_dir, "plots"),
    data = file.path(method_dir, "data"),
    diagnostics = file.path(method_dir, "diagnostics"),
    spectrograms = file.path(method_dir, "concatenated_spectrograms")
  )
  
  lapply(dirs, function(d) {
    dir.create(d, recursive = TRUE, showWarnings = FALSE)
  })
  
  return(dirs)
}

```

## Load Combined Tukey Spectra

```{r}

combined_file <- file.choose()
combined_tukey <- readRDS(combined_file)

class(combined_tukey)
length(combined_tukey)

```

## Inspect Frequency Grid

```{r}

reference_freq <- combined_tukey[[1]]$freq

freq_check <- data.frame(
  EventID = character(),
  SameAsReference = logical(),
  LengthFreq = integer(),
  FreqStep = numeric(),
  MaxFreq = numeric(),
  stringsAsFactors = FALSE
)

for (i in seq_along(combined_tukey)) {
  
  current_freq <- combined_tukey[[i]]$freq
  
  freq_check <- rbind(
    freq_check,
    data.frame(
      EventID = combined_tukey[[i]]$EventID,
      SameAsReference = identical(current_freq, reference_freq),
      LengthFreq = length(current_freq),
      FreqStep = current_freq[2] - current_freq[1],
      MaxFreq = max(current_freq),
      stringsAsFactors = FALSE
    )
  )
}

freq_check

```

## Interpolate Spectra to Shared Grid

```{r}

interpolate_spectrum <- function(old_freq, old_spec, new_freq) {
  approx(
    x = old_freq,
    y = old_spec,
    xout = new_freq,
    method = "linear",
    rule = 2
  )$y
}

interpolated_tukey <- vector("list", length(combined_tukey))

for (i in seq_along(combined_tukey)) {
  
  tukey_spectra <- combined_tukey[[i]]
  
  old_freq <- tukey_spectra$freq
  old_allSpec <- tukey_spectra$allSpec
  
  interpolated_allSpec <- apply(
    old_allSpec,
    2,
    function(spec) interpolate_spectrum(old_freq, spec, reference_freq)
  )
  
  tukey_spectra$allSpec <- as.matrix(interpolated_allSpec)
  tukey_spectra$freq <- reference_freq
  
  interpolated_tukey[[i]] <- tukey_spectra
}

all_match_after_interp <- all(
  sapply(interpolated_tukey, function(x) identical(x$freq, reference_freq))
)

all_match_after_interp

```

## Combine Spectra Across All EVents

```{r}

all_click_metadata <- list()
all_click_spectra <- list()

for (i in seq_along(interpolated_tukey)) {
  
  tukey_spectra <- interpolated_tukey[[i]]
  
  event_id <- tukey_spectra$EventID
  uid <- tukey_spectra$UID
  
  click_spectra <- t(tukey_spectra$allSpec)
  
  click_meta <- data.frame(
    EventID = event_id,
    UID = uid,
    stringsAsFactors = FALSE
  )
  
  all_click_metadata[[i]] <- click_meta
  all_click_spectra[[i]] <- click_spectra
}

click_metadata <- do.call(rbind, all_click_metadata)
all_spec_matrix <- do.call(rbind, all_click_spectra)

dim(all_spec_matrix)

```

## Restrict Spectra to 19–50 kHz

```{r}

keepFreq <- which(reference_freq >= freq_min & reference_freq <= freq_max)

AllSpec_Focus <- all_spec_matrix[, keepFreq]
centroid_freq <- reference_freq[keepFreq]

dim(AllSpec_Focus)
range(centroid_freq)

```

## Build Normalized / Scaled Analysis Matrices

```{r}

normalize_to_zero_baseline <- function(spec_matrix) {
  
  first_freq_value <- spec_matrix[, 1]
  normalized_matrix <- sweep(spec_matrix, 1, first_freq_value, FUN = "-")
  
  return(normalized_matrix)
}

prepare_analysis_matrix <- function(spec_matrix, analysis_type) {
  
  # Step 1: always normalize first
  normalized_matrix <- normalize_to_zero_baseline(spec_matrix)
  
  # Step 2: optionally scale after normalization
  if (analysis_type == "raw_norm") {
    
    output_matrix <- normalized_matrix
    
  } else if (analysis_type == "row_scale") {
    
    output_matrix <- t(scale(t(normalized_matrix)))
    output_matrix[is.na(output_matrix)] <- 0
    
  } else if (analysis_type == "col_scale") {
    
    output_matrix <- scale(normalized_matrix)
    output_matrix[is.na(output_matrix)] <- 0
    
  } else {
    
    stop("Unknown analysis_type: ", analysis_type)
  }
  
  return(as.matrix(output_matrix))
}

analysis_matrices <- list()

for (analysis_name in names(analysis_types)) {
  
  analysis_type <- analysis_types[[analysis_name]]$type
  
  analysis_matrices[[analysis_name]] <- prepare_analysis_matrix(
    spec_matrix = AllSpec_Focus,
    analysis_type = analysis_type
  )
  
  cat(
    analysis_name,
    "matrix dimensions:",
    dim(analysis_matrices[[analysis_name]])[1],
    "clicks x",
    dim(analysis_matrices[[analysis_name]])[2],
    "frequency bins\n"
  )
}

```

## Helper: Sample Up to N Clicks Per Event

```{r}

sample_by_event <- function(metadata_df, max_per_event, seed = NULL, max_total = NULL) {
  
  if (!is.null(seed)) set.seed(seed)
  
  split_indices <- split(seq_len(nrow(metadata_df)), metadata_df$EventID)
  
  sampled_indices <- unlist(
    lapply(split_indices, function(idx) {
      sample(idx, size = min(length(idx), max_per_event))
    }),
    use.names = FALSE
  )
  
  if (!is.null(max_total) && length(sampled_indices) > max_total) {
    sampled_indices <- sample(sampled_indices, size = max_total)
  }
  
  return(sampled_indices)
}

```

## Run Analysis for Each Normalization Type

```{r}

results_list <- list()

for (analysis_name in names(analysis_matrices)) {
  
  cat("\n============================\n")
  cat("Running:", analysis_name, "\n")
  cat("============================\n")
  
  AllSpec_current <- analysis_matrices[[analysis_name]]
  
  # -----------------------------
  # Create Output Structure
  # -----------------------------
  dirs <- create_output_structure(
    base_dir = base_output_dir,
    analysis_name = analysis_name,
    method = cluster_method,
    k_label = k_label
  )
  
  # -----------------------------
  # Elbow Sampling
  # -----------------------------
  elbow_index <- sample_by_event(
    metadata_df = click_metadata,
    max_per_event = max_clicks_per_event,
    seed = 1,
    max_total = elbow_sample_cap
  )
  
  AllSpec_elbow <- AllSpec_current[elbow_index, ]
  
  wss <- numeric(length(k_candidates))
  
  for (i in seq_along(k_candidates)) {
    
    k <- k_candidates[i]
    
    km_tmp <- kmeans(
      AllSpec_elbow,
      centers = k,
      nstart = 20,
      algorithm = "Lloyd"
    )
    
    wss[i] <- km_tmp$tot.withinss
  }
  
  elbow_df <- data.frame(k = k_candidates, WSS = wss)
  
  elbow_plot <- ggplot(elbow_df, aes(x = k, y = WSS)) +
    geom_line() +
    geom_point() +
    labs(title = paste("Elbow Plot -", analysis_name))
  
  ggsave(
    file.path(dirs$plots, paste0("elbow_", analysis_name, ".png")),
    elbow_plot,
    width = 7,
    height = 5
  )
  
  # -----------------------------
  # Evaluation Sample
  # -----------------------------
  eval_index <- sample_by_event(
    metadata_df = click_metadata,
    max_per_event = max_clicks_per_event,
    seed = 2
  )
  
  AllSpec_eval <- AllSpec_current[eval_index, ]
  click_metadata_eval <- click_metadata[eval_index, ]
  
  cat("Eval sample size:", nrow(AllSpec_eval), "\n")
  
  # -----------------------------
  # KMEANS
  # -----------------------------
  set.seed(3)
  
  if (run_kmeans_on_full_dataset) {
    
    km <- kmeans(
      AllSpec_current,
      centers = chosen_k,
      nstart = 50,
      algorithm = "Lloyd"
    )
    
    cluster_labels_full <- km$cluster
    cluster_labels_eval <- cluster_labels_full[eval_index]
    
  } else {
    
    km <- kmeans(
      AllSpec_eval,
      centers = chosen_k,
      nstart = 50
    )
    
    cluster_labels_eval <- km$cluster
  }
  
  # -----------------------------
  # Silhouette
  # -----------------------------
  dist_eval <- dist(AllSpec_eval)
  
  sil <- silhouette(cluster_labels_eval, dist_eval)
  mean_sil <- mean(sil[, "sil_width"])
  
  cat("Mean silhouette:", mean_sil, "\n")
  
  # Save silhouette summary
  write.csv(
    data.frame(
      Analysis = analysis_name,
      K = chosen_k,
      MeanSilhouette = mean_sil
    ),
    file = file.path(dirs$data, "silhouette_summary.csv"),
    row.names = FALSE
  )
  
  # Silhouette plot
  sil_df <- as.data.frame(sil) %>%
    mutate(obs = row_number()) %>%
    group_by(cluster) %>%
    arrange(desc(sil_width), .by_group = TRUE) %>%
    mutate(order = row_number()) %>%
    ungroup()
  
  sil_plot <- ggplot(sil_df, aes(x = order, y = sil_width)) +
    geom_col() +
    facet_wrap(~ cluster, scales = "free_x") +
    labs(title = paste("Silhouette -", analysis_name))
  
  ggsave(
    file.path(dirs$plots, "silhouette_plot.png"),
    sil_plot,
    width = 8,
    height = 5
  )
  
  # Store for later chunks (PCA, centroids, etc.)
  results_list[[analysis_name]] <- list(
    dirs = dirs,
    km = km,
    eval_index = eval_index,
    cluster_labels_eval = cluster_labels_eval,
    AllSpec_eval = AllSpec_eval,
    AllSpec_current = AllSpec_current
  )
}

```

## PCA Diagnostics and Plots for Each Analysis Type

```{r}

calculate_pca_variance <- function(pca_object, n_components = 2) {
  
  eigenvalues <- pca_object$sdev^2
  
  data.frame(
    RetainedComponents = n_components,
    RetainedEigenvalueSum = sum(eigenvalues[1:n_components]),
    TotalEigenvalueSum = sum(eigenvalues),
    PercentVarianceExplained = 100 * sum(eigenvalues[1:n_components]) / sum(eigenvalues)
  )
}

get_nonzero_variance_columns <- function(mat) {
  col_sd <- apply(mat, 2, sd, na.rm = TRUE)
  is.finite(col_sd) & col_sd > 0
}

for (analysis_name in names(results_list)) {
  
  cat("\nRunning PCA diagnostics for:", analysis_name, "\n")
  
  result <- results_list[[analysis_name]]
  
  dirs <- result$dirs
  AllSpec_eval <- result$AllSpec_eval
  AllSpec_current <- result$AllSpec_current
  eval_index <- result$eval_index
  cluster_labels_eval <- result$cluster_labels_eval
  
  click_metadata_eval <- click_metadata[eval_index, ]
  
  # --------------------------------------------------
  # Remove zero-variance columns before PCA
  # --------------------------------------------------
  
  pca_keep_cols <- get_nonzero_variance_columns(AllSpec_eval)
  
  AllSpec_eval_pca <- AllSpec_eval[, pca_keep_cols, drop = FALSE]
  AllSpec_current_pca <- AllSpec_current[, pca_keep_cols, drop = FALSE]
  centroid_freq_pca <- centroid_freq[pca_keep_cols]
  
  removed_cols <- sum(!pca_keep_cols)
  
  cat(
    "PCA removed",
    removed_cols,
    "zero-variance frequency columns for",
    analysis_name,
    "\n"
  )
  
  # --------------------------------------------------
  # Covariance-style PCA
  # --------------------------------------------------
  
  pca_cov_eval <- prcomp(
    AllSpec_eval_pca,
    center = FALSE,
    scale. = FALSE
  )
  
  pca_cov_full <- prcomp(
    AllSpec_current_pca,
    center = FALSE,
    scale. = FALSE
  )
  
  cov_variance_df <- calculate_pca_variance(pca_cov_eval, n_components = 2)
  cov_variance_df$Analysis <- analysis_name
  cov_variance_df$PCA_Type <- "Covariance-style PCA"
  cov_variance_df$ZeroVarianceColumnsRemoved <- removed_cols
  
  # --------------------------------------------------
  # Correlation PCA
  # --------------------------------------------------
  
  pca_cor_eval <- prcomp(
    AllSpec_eval_pca,
    center = TRUE,
    scale. = TRUE
  )
  
  pca_cor_full <- prcomp(
    AllSpec_current_pca,
    center = TRUE,
    scale. = TRUE
  )
  
  cor_variance_df <- calculate_pca_variance(pca_cor_eval, n_components = 2)
  cor_variance_df$Analysis <- analysis_name
  cor_variance_df$PCA_Type <- "Correlation PCA"
  cor_variance_df$ZeroVarianceColumnsRemoved <- removed_cols
  
  # --------------------------------------------------
  # Save PCA variance summary
  # --------------------------------------------------
  
  pca_variance_summary <- rbind(cov_variance_df, cor_variance_df)
  
  write.csv(
    pca_variance_summary,
    file = file.path(dirs$diagnostics, "pca_variance_explained_summary.csv"),
    row.names = FALSE
  )
  
  print(pca_variance_summary)
  
  # --------------------------------------------------
  # Covariance PCA Scatterplot
  # --------------------------------------------------
  
  scatter_cov_df <- data.frame(
    PC1 = pca_cov_eval$x[, 1],
    PC2 = pca_cov_eval$x[, 2],
    EventID = click_metadata_eval$EventID,
    UID = click_metadata_eval$UID,
    cluster = as.factor(cluster_labels_eval)
  )
  
  scatter_cov_plot <- ggplot(scatter_cov_df, aes(x = PC1, y = PC2, color = cluster)) +
    geom_point(alpha = 0.7, size = 1.2) +
    labs(
      title = paste("Covariance PCA Projection -", analysis_name),
      subtitle = paste0(
        "PC1 + PC2 variance explained: ",
        round(cov_variance_df$PercentVarianceExplained, 2),
        "%"
      ),
      x = "PC1",
      y = "PC2",
      color = "Cluster"
    )
  
  print(scatter_cov_plot)
  
  ggsave(
    file.path(dirs$plots, "pca_scatter_covariance.png"),
    scatter_cov_plot,
    width = 7,
    height = 5
  )
  
  write.csv(
    scatter_cov_df,
    file = file.path(dirs$data, "pca_scatter_covariance_data.csv"),
    row.names = FALSE
  )
  
  # --------------------------------------------------
  # Correlation PCA Scatterplot
  # --------------------------------------------------
  
  scatter_cor_df <- data.frame(
    PC1 = pca_cor_eval$x[, 1],
    PC2 = pca_cor_eval$x[, 2],
    EventID = click_metadata_eval$EventID,
    UID = click_metadata_eval$UID,
    cluster = as.factor(cluster_labels_eval)
  )
  
  scatter_cor_plot <- ggplot(scatter_cor_df, aes(x = PC1, y = PC2, color = cluster)) +
    geom_point(alpha = 0.7, size = 1.2) +
    labs(
      title = paste("Correlation PCA Projection -", analysis_name),
      subtitle = paste0(
        "PC1 + PC2 variance explained: ",
        round(cor_variance_df$PercentVarianceExplained, 2),
        "%"
      ),
      x = "PC1",
      y = "PC2",
      color = "Cluster"
    )
  
  print(scatter_cor_plot)
  
  ggsave(
    file.path(dirs$plots, "pca_scatter_correlation.png"),
    scatter_cor_plot,
    width = 7,
    height = 5
  )
  
  write.csv(
    scatter_cor_df,
    file = file.path(dirs$data, "pca_scatter_correlation_data.csv"),
    row.names = FALSE
  )
  
  # --------------------------------------------------
  # Full Dataset PCA Loadings
  # --------------------------------------------------
  
  cov_loadings_df <- data.frame(
    Frequency_kHz = centroid_freq_pca / 1000,
    PC1 = pca_cov_full$rotation[, 1],
    PC2 = pca_cov_full$rotation[, 2],
    PCA_Type = "Covariance-style PCA"
  )
  
  cor_loadings_df <- data.frame(
    Frequency_kHz = centroid_freq_pca / 1000,
    PC1 = pca_cor_full$rotation[, 1],
    PC2 = pca_cor_full$rotation[, 2],
    PCA_Type = "Correlation PCA"
  )
  
  loadings_df <- rbind(cov_loadings_df, cor_loadings_df)
  
  write.csv(
    loadings_df,
    file = file.path(dirs$data, "full_dataset_pca_loadings.csv"),
    row.names = FALSE
  )
  
  cov_loadings_plot <- ggplot(cov_loadings_df, aes(x = Frequency_kHz)) +
    geom_line(aes(y = PC1, color = "PC1"), linewidth = 1) +
    geom_line(aes(y = PC2, color = "PC2"), linewidth = 1) +
    labs(
      title = paste("Full-Dataset Covariance PCA Loadings -", analysis_name),
      x = "Frequency (kHz)",
      y = "Loading",
      color = "Component"
    )
  
  print(cov_loadings_plot)
  
  ggsave(
    file.path(dirs$plots, "full_dataset_pca_loadings_covariance.png"),
    cov_loadings_plot,
    width = 7,
    height = 5
  )
  
  cor_loadings_plot <- ggplot(cor_loadings_df, aes(x = Frequency_kHz)) +
    geom_line(aes(y = PC1, color = "PC1"), linewidth = 1) +
    geom_line(aes(y = PC2, color = "PC2"), linewidth = 1) +
    labs(
      title = paste("Full-Dataset Correlation PCA Loadings -", analysis_name),
      x = "Frequency (kHz)",
      y = "Loading",
      color = "Component"
    )
  
  print(cor_loadings_plot)
  
  ggsave(
    file.path(dirs$plots, "full_dataset_pca_loadings_correlation.png"),
    cor_loadings_plot,
    width = 7,
    height = 5
  )
  
  cat("Saved PCA outputs for:", analysis_name, "\n")
}

```

## Centroids, Variability, and Concatenated Spectrograms

```{r}

for (analysis_name in names(results_list)) {
  
  cat("\nRunning centroid + spectrogram outputs for:", analysis_name, "\n")
  
  result <- results_list[[analysis_name]]
  
  dirs <- result$dirs
  km <- result$km
  eval_index <- result$eval_index
  cluster_labels_eval <- result$cluster_labels_eval
  AllSpec_eval <- result$AllSpec_eval
  AllSpec_current <- result$AllSpec_current
  
  # --------------------------------------------------
  # Centroid Spectra
  # --------------------------------------------------
  
  centroid_df <- data.frame(
    Frequency_kHz = centroid_freq / 1000,
    t(km$centers)
  )
  
  colnames(centroid_df) <- c(
    "Frequency_kHz",
    paste0("Cluster_", 1:nrow(km$centers))
  )
  
  write.csv(
    centroid_df,
    file = file.path(dirs$data, "centroid_spectra.csv"),
    row.names = FALSE
  )
  
  centroid_plot_file <- file.path(dirs$plots, "centroid_plot.png")
  
  png(
    filename = centroid_plot_file,
    width = 7,
    height = 5,
    units = "in",
    res = 300
  )
  
  matplot(
    x = centroid_freq / 1000,
    y = t(km$centers),
    type = "l",
    lty = 1,
    main = paste("Cluster Centroids -", analysis_name),
    xlab = "Frequency (kHz)",
    ylab = "Intensity"
  )
  
  dev.off()
  
  # --------------------------------------------------
  # Cluster Variability
  # --------------------------------------------------
  
  if (run_kmeans_on_full_dataset) {
    
    cluster_labels_full <- km$cluster
    
    dist_to_centroid <- numeric(nrow(AllSpec_current))
    
    for (i in seq_len(nrow(AllSpec_current))) {
      
      this_cluster <- cluster_labels_full[i]
      this_centroid <- km$centers[this_cluster, ]
      
      dist_to_centroid[i] <- sqrt(
        sum((AllSpec_current[i, ] - this_centroid)^2)
      )
    }
    
    variability_df <- data.frame(
      cluster = cluster_labels_full,
      dist = dist_to_centroid
    ) %>%
      group_by(cluster) %>%
      summarise(
        n_clicks = n(),
        mean_dist = mean(dist),
        sd_dist = sd(dist),
        median_dist = median(dist),
        .groups = "drop"
      )
    
    write.csv(
      variability_df,
      file = file.path(dirs$data, "cluster_variability.csv"),
      row.names = FALSE
    )
    
    print(variability_df)
  }
  
  # --------------------------------------------------
  # Concatenated Spectrograms
  # --------------------------------------------------
  
  cluster_ids <- sort(unique(cluster_labels_eval))
  
  spec_cols <- colorRampPalette(
    c("white", "#FFD54F", "#FF9800", "#D84315", "#4E0012")
  )(200)
  
  for (cl in cluster_ids) {
    
    cluster_idx <- which(cluster_labels_eval == cl)
    
    if (length(cluster_idx) == 0) next
    
    cluster_spec <- AllSpec_eval[cluster_idx, , drop = FALSE]
    
    out_file <- file.path(
      dirs$spectrograms,
      paste0(
        "concatenated_spectrogram_eval_19_50kHz_",
        k_label,
        "_cluster_",
        cl,
        ".png"
      )
    )
    
    png(out_file, width = 1800, height = 900)
    
    image(
      x = seq_len(nrow(cluster_spec)),
      y = centroid_freq / 1000,
      z = cluster_spec,
      col = spec_cols,
      xlab = "Click Number",
      ylab = "Frequency (kHz)",
      main = paste(
        "Concatenated Click Spectrogram -",
        analysis_name,
        "- Cluster",
        cl
      ),
      useRaster = TRUE,
      axes = FALSE
    )
    
    axis(1)
    axis(2, at = seq(20, 50, by = 10))
    box()
    
    dev.off()
    
    cat("Saved:", out_file, "\n")
  }
  
  cat("Finished:", analysis_name, "\n")
}

```




























