---
title: "Clustering_kmeans_compare_scaling"
author: "Sarah Shiers"
date: "2026-03-18"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

## Overview

This workflow prepares combined Tukey spectra for clustering by standardizing all events to a shared frequency grid, combining all click spectra into one matrix, and then running k-means clustering under two different scaling approaches:

old_scaling: column-wise scaling across all clicks

new_scaling: row-wise scaling within each click

The goal is to compare how the clustering results differ under these two preprocessing choices while keeping the rest of the workflow identical.

## Load Required Packages

```{r}

library(ggplot2)
library(dplyr)
library(cluster)

```

## User Settings

```{r}

cluster_method <- "kmeans"   # options: "kmeans", "kshape", "hdbscan"

output_dir <- "E:/Analysis/Clustering_Output"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

freq_min <- 19000
freq_max <- 50000

# Candidate cluster numbers to evaluate for elbow plots
k_candidates <- 2:15

# Set after reviewing elbow plots
chosen_k <- 4   # use NULL before choosing

# Sampling settings
max_clicks_per_event <- 200

# Optional cap after event-balanced sampling for elbow
elbow_sample_cap <- 20000

# Evaluation sample uses up to 200 clicks per event with no global cap
eval_sample_cap <- NULL

# Whether to run final kmeans on all clicks
run_kmeans_on_full_dataset <- TRUE

```

## Load Combined Tukey Spectra

```{r}

combined_file <- file.choose()
combined_tukey <- readRDS(combined_file)

class(combined_tukey)
length(combined_tukey)

```

## Inspect Frequency Grids

```{r}

reference_freq <- combined_tukey[[1]]$freq

freq_check <- data.frame(
  EventID = character(),
  SameAsReference = logical(),
  LengthFreq = integer(),
  FreqStep = numeric(),
  MaxFreq = numeric(),
  stringsAsFactors = FALSE
)

for (i in seq_along(combined_tukey)) {
  
  current_freq <- combined_tukey[[i]]$freq
  
  freq_check <- rbind(
    freq_check,
    data.frame(
      EventID = combined_tukey[[i]]$EventID,
      SameAsReference = identical(current_freq, reference_freq),
      LengthFreq = length(current_freq),
      FreqStep = current_freq[2] - current_freq[1],
      MaxFreq = max(current_freq),
      stringsAsFactors = FALSE
    )
  )
}

freq_check

```

## Choose Reference Frequency Grid

```{r}

reference_freq <- combined_tukey[[1]]$freq
reference_event <- combined_tukey[[1]]$EventID

cat("Reference event:", reference_event, "\n")
cat("Reference frequency step:", reference_freq[2] - reference_freq[1], "Hz\n")
cat("Reference max frequency:", max(reference_freq), "Hz\n")

```

## Interpolate Spectra to Shared Grid

```{r}

interpolate_spectrum <- function(old_freq, old_spec, new_freq) {
  approx(
    x = old_freq,
    y = old_spec,
    xout = new_freq,
    method = "linear",
    rule = 2
  )$y
}

interpolated_tukey <- vector("list", length(combined_tukey))

for (i in seq_along(combined_tukey)) {
  
  tukey_spectra <- combined_tukey[[i]]
  
  old_freq <- tukey_spectra$freq
  old_allSpec <- tukey_spectra$allSpec
  
  interpolated_allSpec <- apply(
    old_allSpec,
    2,
    function(spec) interpolate_spectrum(old_freq, spec, reference_freq)
  )
  
  interpolated_allSpec <- as.matrix(interpolated_allSpec)
  
  tukey_spectra$allSpec <- interpolated_allSpec
  tukey_spectra$freq <- reference_freq
  
  interpolated_tukey[[i]] <- tukey_spectra
}

```

## Confirm Interolation

```{r}

all_match_after_interp <- all(
  sapply(interpolated_tukey, function(x) identical(x$freq, reference_freq))
)

all_match_after_interp

```

## Combine Spectra Across Evvents

```{r}

all_click_metadata <- list()
all_click_spectra <- list()

for (i in seq_along(interpolated_tukey)) {
  
  tukey_spectra <- interpolated_tukey[[i]]
  
  event_id <- tukey_spectra$EventID
  uid <- tukey_spectra$UID
  
  click_spectra <- t(tukey_spectra$allSpec)
  
  click_meta <- data.frame(
    EventID = event_id,
    UID = uid,
    stringsAsFactors = FALSE
  )
  
  all_click_metadata[[i]] <- click_meta
  all_click_spectra[[i]] <- click_spectra
}

click_metadata <- do.call(rbind, all_click_metadata)
all_spec_matrix <- do.call(rbind, all_click_spectra)

dim(all_spec_matrix)

```

## Restrict Spectra to 19–50 kHz

```{r}

keepFreq <- which(reference_freq >= freq_min & reference_freq <= freq_max)

AllSpec_Focus <- all_spec_matrix[, keepFreq]
centroid_freq <- reference_freq[keepFreq]

```

## Create Both Scaling Versions

```{r}

AllSpec_scale_old <- scale(AllSpec_Focus)

AllSpec_scale_new <- t(scale(t(AllSpec_Focus)))
AllSpec_scale_new[is.na(AllSpec_scale_new)] <- 0

scaled_data_list <- list(
  old_scaling = AllSpec_scale_old,
  new_scaling = AllSpec_scale_new
)

names(scaled_data_list)

```

## Helper Function: Sample Up to N Clicks Per Event

```{r}

sample_by_event <- function(metadata_df, max_per_event, seed = NULL, max_total = NULL) {
  
  if (!is.null(seed)) {
    set.seed(seed)
  }
  
  split_indices <- split(seq_len(nrow(metadata_df)), metadata_df$EventID)
  
  sampled_indices <- unlist(
    lapply(split_indices, function(idx) {
      sample(idx, size = min(length(idx), max_per_event))
    }),
    use.names = FALSE
  )
  
  if (!is.null(max_total) && length(sampled_indices) > max_total) {
    sampled_indices <- sample(sampled_indices, size = max_total)
  }
  
  return(sampled_indices)
}

```

## Create Shared Elbow and Evaluation Samples

```{r}

elbow_index <- sample_by_event(
  metadata_df = click_metadata,
  max_per_event = max_clicks_per_event,
  seed = 1,
  max_total = elbow_sample_cap
)

eval_index <- sample_by_event(
  metadata_df = click_metadata,
  max_per_event = max_clicks_per_event,
  seed = 2,
  max_total = eval_sample_cap
)

click_metadata_eval <- click_metadata[eval_index, ]

cat("Elbow sample size used:", length(elbow_index), "\n")
cat("Evaluation sample size used:", length(eval_index), "\n")

```

## Sampling Diagnostics

```{r}

event_total <- click_metadata %>%
  count(EventID, name = "total_clicks")

event_sampled <- click_metadata_eval %>%
  count(EventID, name = "sampled_clicks")

sampling_table <- event_total %>%
  left_join(event_sampled, by = "EventID") %>%
  mutate(sampled_clicks = ifelse(is.na(sampled_clicks), 0, sampled_clicks)) %>%
  arrange(EventID)

sampling_table

write.csv(
  sampling_table,
  file = file.path(output_dir, "sampling_table_shared_across_scalings.csv"),
  row.names = FALSE
)

```

##Determine Number of Clusters (Elbow Method) for Both Scaling Approaches

```{r}

elbow_results <- list()

for (scale_name in names(scaled_data_list)) {
  
  cat("\n------------------------------\n")
  cat("Running elbow analysis for:", scale_name, "\n")
  cat("------------------------------\n")
  
  AllSpec_scale_current <- scaled_data_list[[scale_name]]
  AllSpec_elbow <- AllSpec_scale_current[elbow_index, ]
  
  wss <- numeric(length(k_candidates))
  
  for (i in seq_along(k_candidates)) {
    
    k <- k_candidates[i]
    
    km_tmp <- kmeans(
      AllSpec_elbow,
      centers = k,
      nstart = 20,
      algorithm = "Lloyd"
    )
    
    wss[i] <- km_tmp$tot.withinss
  }
  
  elbow_df <- data.frame(
    k = k_candidates,
    WSS = wss,
    ScalingMethod = scale_name
  )
  
  elbow_plot <- ggplot(elbow_df, aes(x = k, y = WSS)) +
    geom_line() +
    geom_point() +
    labs(
      title = paste("Elbow Plot -", scale_name),
      x = "Number of Clusters",
      y = "Total Within SS"
    )
  
  print(elbow_plot)
  
  ggsave(
    filename = file.path(output_dir, paste0("kmeans_elbow_plot_", scale_name, ".png")),
    plot = elbow_plot,
    width = 7,
    height = 5
  )
  
  write.csv(
    elbow_df,
    file = file.path(output_dir, paste0("kmeans_elbow_data_", scale_name, ".csv")),
    row.names = FALSE
  )
  
  elbow_results[[scale_name]] <- elbow_df
}

```

## Confirm Chosen Number of Clusters

```{r}

if (is.null(chosen_k)) {
  stop("Set chosen_k after reviewing BOTH elbow plots before continuing.")
}

chosen_k

```

## Run Full Analysis for Both Scaling Approaches

```{r}

results_list <- list()
summary_rows <- list()
silhouette_comparison_rows <- list()

for (scale_name in names(scaled_data_list)) {
  
  cat("\n====================================\n")
  cat("Running full analysis for:", scale_name, "\n")
  cat("====================================\n")
  
  AllSpec_scale_current <- scaled_data_list[[scale_name]]
  AllSpec_eval <- AllSpec_scale_current[eval_index, ]
  
  click_metadata_eval_current <- click_metadata_eval
  
  if (cluster_method == "kmeans") {
    
    set.seed(3)
    
    if (run_kmeans_on_full_dataset) {
      
      km <- kmeans(
        AllSpec_scale_current,
        centers = chosen_k,
        nstart = 50,
        algorithm = "Lloyd"
      )
      
      cluster_labels_full <- km$cluster
      cluster_labels_eval <- cluster_labels_full[eval_index]
      
      click_metadata_full_current <- click_metadata
      click_metadata_full_current$cluster <- cluster_labels_full
      
    } else {
      
      km <- kmeans(
        AllSpec_eval,
        centers = chosen_k,
        nstart = 50,
        algorithm = "Lloyd"
      )
      
      cluster_labels_eval <- km$cluster
      
      click_metadata_full_current <- NULL
    }
  }
  
  click_metadata_eval_current$cluster <- cluster_labels_eval
  
  current_dist <- dist(AllSpec_eval)
  current_sil <- silhouette(cluster_labels_eval, current_dist)
  mean_silhouette <- mean(current_sil[, 3])
  
  cat("Mean silhouette for", scale_name, "=", mean_silhouette, "\n")
  
  silhouette_k_rows <- list()
  
  for (k_test in 3:6) {
    km_test <- kmeans(AllSpec_eval, centers = k_test, nstart = 20)
    sil_test <- silhouette(km_test$cluster, current_dist)
    silhouette_k_rows[[as.character(k_test)]] <- data.frame(
      ScalingMethod = scale_name,
      TestedK = k_test,
      MeanSilhouette = mean(sil_test[, 3])
    )
    cat("Scaling =", scale_name, "| k =", k_test, "| mean silhouette =", mean(sil_test[, 3]), "\n")
  }
  
  silhouette_k_df <- do.call(rbind, silhouette_k_rows)
  
  sil_df <- as.data.frame(current_sil) %>%
    mutate(obs = row_number()) %>%
    group_by(cluster) %>%
    arrange(desc(sil_width), .by_group = TRUE) %>%
    mutate(order = row_number()) %>%
    ungroup()
  
  sil_plot <- ggplot(sil_df, aes(x = order, y = sil_width)) +
    geom_col() +
    facet_wrap(~ cluster, scales = "free_x") +
    labs(
      title = paste("Silhouette Plot -", scale_name, "(k =", chosen_k, ")"),
      x = "Observations",
      y = "Silhouette Width"
    )
  
  print(sil_plot)
  
  ggsave(
    filename = file.path(output_dir, paste0("kmeans_silhouette_plot_", scale_name, ".png")),
    plot = sil_plot,
    width = 8,
    height = 5
  )
  
  pca_eval <- prcomp(AllSpec_eval)
  
  scatter_df <- data.frame(
    PC1 = pca_eval$x[, 1],
    PC2 = pca_eval$x[, 2],
    EventID = click_metadata_eval_current$EventID,
    UID = click_metadata_eval_current$UID,
    cluster = as.factor(cluster_labels_eval),
    ScalingMethod = scale_name
  )
  
  scatter_plot <- ggplot(scatter_df, aes(x = PC1, y = PC2, color = cluster)) +
    geom_point(alpha = 0.7, size = 1.2) +
    labs(
      title = paste("Clustered Clicks (PCA Projection) -", scale_name, "(k =", chosen_k, ")"),
      x = "PC1",
      y = "PC2",
      color = "Cluster"
    )
  
  print(scatter_plot)
  
  ggsave(
    filename = file.path(output_dir, paste0("kmeans_scatterplot_", scale_name, ".png")),
    plot = scatter_plot,
    width = 7,
    height = 5
  )
  
  centroid_file <- file.path(output_dir, paste0("kmeans_centroid_plot_", scale_name, ".png"))
  
  png(
    filename = centroid_file,
    width = 7,
    height = 5,
    units = "in",
    res = 300
  )
  
  matplot(
    x = centroid_freq / 1000,
    y = t(km$centers),
    type = "l",
    lty = 1,
    main = paste("Cluster Centroids -", scale_name, "(k =", chosen_k, ")"),
    xlab = "Frequency (kHz)",
    ylab = "Scaled Intensity"
  )
  
  dev.off()
  
  centroid_df <- data.frame(
    Frequency_kHz = centroid_freq / 1000,
    t(km$centers)
  )
  
  colnames(centroid_df) <- c(
    "Frequency_kHz",
    paste0("Cluster_", 1:chosen_k)
  )
  
  centroid_distance_df <- as.matrix(dist(km$centers))
  
  write.csv(
    scatter_df,
    file = file.path(output_dir, paste0("kmeans_scatterplot_data_", scale_name, ".csv")),
    row.names = FALSE
  )
  
  write.csv(
    centroid_df,
    file = file.path(output_dir, paste0("kmeans_centroid_spectra_", scale_name, ".csv")),
    row.names = FALSE
  )
  
  write.csv(
    centroid_distance_df,
    file = file.path(output_dir, paste0("kmeans_centroid_distances_", scale_name, ".csv"))
  )
  
  write.csv(
    silhouette_k_df,
    file = file.path(output_dir, paste0("kmeans_silhouette_k_comparison_", scale_name, ".csv")),
    row.names = FALSE
  )
  
  if (run_kmeans_on_full_dataset) {
    write.csv(
      click_metadata_full_current,
      file = file.path(output_dir, paste0("kmeans_full_cluster_assignments_", scale_name, ".csv")),
      row.names = FALSE
    )
  }
  
  results_list[[scale_name]] <- list(
    km = km,
    cluster_labels_eval = cluster_labels_eval,
    click_metadata_eval = click_metadata_eval_current,
    silhouette = current_sil,
    mean_silhouette = mean_silhouette,
    scatter_df = scatter_df,
    centroid_df = centroid_df,
    centroid_distances = centroid_distance_df,
    silhouette_k_df = silhouette_k_df
  )
  
  summary_rows[[scale_name]] <- data.frame(
    ScalingMethod = scale_name,
    Method = cluster_method,
    ChosenK = chosen_k,
    MeanSilhouette = mean_silhouette,
    RunKmeansOnFullDataset = run_kmeans_on_full_dataset,
    MaxClicksPerEvent = max_clicks_per_event,
    ElbowSampleSizeUsed = length(elbow_index),
    EvalSampleSizeUsed = length(eval_index),
    TotalClicksInFullDataset = nrow(AllSpec_scale_current)
  )
  
  silhouette_comparison_rows[[scale_name]] <- silhouette_k_df
}

```

## Combined Summary Table

```{r}

summary_table <- do.call(rbind, summary_rows)

summary_table

write.csv(
  summary_table,
  file = file.path(output_dir, "kmeans_summary_table_all_scalings.csv"),
  row.names = FALSE
)

```

## Combined Silhouette Comparison Table

```{r}

silhouette_comparison_table <- do.call(rbind, silhouette_comparison_rows)

silhouette_comparison_table

write.csv(
  silhouette_comparison_table,
  file = file.path(output_dir, "kmeans_silhouette_k_comparison_all_scalings.csv"),
  row.names = FALSE
)

```

## Quick Direct Comparison of Mean Silhouette by Scaling Method

```{r}

summary_table %>%
  select(ScalingMethod, ChosenK, MeanSilhouette)

```











