---
title: "Clustering_kmeans_kshape_hdbscan"
author: "Sarah Shiers"
date: "2026-03-16"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

## Overview

This workflow prepares combined Tukey spectra for clustering by standardizing all events to a shared frequency grid, combining all click spectra into one matrix, and then running one selected clustering method.

Current methods supported by the workflow structure:

- kmeans
- kshape (placeholder)
- hdbscan (placeholder)

Current method fully implemented:kmeans

## Load Required Packages

```{r}

library(ggplot2)
library(dplyr)
library(cluster)

```

## User Settings

```{r}

cluster_method <- "kmeans"   # options: "kmeans", "kshape", "hdbscan"

output_dir <- "E:/Analysis/Clustering Output"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

freq_min <- 19000
freq_max <- 50000

# Candidate cluster numbers to evaluate for elbow plots
k_candidates <- 2:15

# Set after reviewing elbow plot
chosen_k <- 3 # Use "NULL" prior
k_label <- paste0("k_", chosen_k)

# Sampling settings
max_clicks_per_event <- 200

# Optional caps after event-balanced sampling
elbow_sample_cap <- 20000
eval_sample_cap <- NULL

# Whether to run final kmeans on all clicks
run_kmeans_on_full_dataset <- TRUE

```


## Load Combined Tukey Spectra

```{r}

combined_file <- file.choose()
combined_tukey <- readRDS(combined_file)

class(combined_tukey)
length(combined_tukey)

```

## Inspect Frequency Grids

```{r}

reference_freq <- combined_tukey[[1]]$freq

freq_check <- data.frame(
  EventID = character(),
  SameAsReference = logical(),
  LengthFreq = integer(),
  FreqStep = numeric(),
  MaxFreq = numeric(),
  stringsAsFactors = FALSE
)

for (i in seq_along(combined_tukey)) {
  
  current_freq <- combined_tukey[[i]]$freq
  
  freq_check <- rbind(
    freq_check,
    data.frame(
      EventID = combined_tukey[[i]]$EventID,
      SameAsReference = identical(current_freq, reference_freq),
      LengthFreq = length(current_freq),
      FreqStep = current_freq[2] - current_freq[1],
      MaxFreq = max(current_freq),
      stringsAsFactors = FALSE
    )
  )
}

freq_check

```

## Choose Reference Frequency Grid

```{r}

reference_freq <- combined_tukey[[1]]$freq
reference_event <- combined_tukey[[1]]$EventID

cat("Reference event:", reference_event, "\n")
cat("Reference frequency step:", reference_freq[2] - reference_freq[1], "Hz\n")
cat("Reference max frequency:", max(reference_freq), "Hz\n")

```

## Interpolate Spectra to Shared Grid

```{r}

interpolate_spectrum <- function(old_freq, old_spec, new_freq) {
  
  approx(
    x = old_freq,
    y = old_spec,
    xout = new_freq,
    method = "linear",
    rule = 2
  )$y
  
}

interpolated_tukey <- vector("list", length(combined_tukey))

for (i in seq_along(combined_tukey)) {
  
  tukey_spectra <- combined_tukey[[i]]
  
  old_freq <- tukey_spectra$freq
  old_allSpec <- tukey_spectra$allSpec
  
  interpolated_allSpec <- apply(
    old_allSpec,
    2,
    function(spec) interpolate_spectrum(old_freq, spec, reference_freq)
  )
  
  interpolated_allSpec <- as.matrix(interpolated_allSpec)
  
  tukey_spectra$allSpec <- interpolated_allSpec
  tukey_spectra$freq <- reference_freq
  
  interpolated_tukey[[i]] <- tukey_spectra
  
}

```

## Confirm Interpolation

```{r}

all_match_after_interp <- all(
  sapply(interpolated_tukey, function(x) identical(x$freq, reference_freq))
)

all_match_after_interp

```

## Combine Spectra Across Events

```{r}

all_click_metadata <- list()
all_click_spectra <- list()

for (i in seq_along(interpolated_tukey)) {
  
  tukey_spectra <- interpolated_tukey[[i]]
  
  event_id <- tukey_spectra$EventID
  uid <- tukey_spectra$UID
  
  click_spectra <- t(tukey_spectra$allSpec)
  
  click_meta <- data.frame(
    EventID = event_id,
    UID = uid,
    stringsAsFactors = FALSE
  )
  
  all_click_metadata[[i]] <- click_meta
  all_click_spectra[[i]] <- click_spectra
  
}

click_metadata <- do.call(rbind, all_click_metadata)
all_spec_matrix <- do.call(rbind, all_click_spectra)

dim(all_spec_matrix)

```

## Restrict Spectra to 19–50 kHz

```{r}

keepFreq <- which(reference_freq >= freq_min & reference_freq <= freq_max)

AllSpec_Focus <- all_spec_matrix[, keepFreq]
centroid_freq <- reference_freq[keepFreq]

```

## Scale Spectra

```{r}

# Row-wise scaling: emphasizes spectral shape within each click
AllSpec_scale <- t(scale(t(AllSpec_Focus)))
AllSpec_scale[is.na(AllSpec_scale)] <- 0

```

## Helper Function: Sample Up to N Clicks Per Event

```{r}

sample_by_event <- function(metadata_df, max_per_event, seed = NULL, max_total = NULL) {
  
  if (!is.null(seed)) {
    set.seed(seed)
  }
  
  split_indices <- split(seq_len(nrow(metadata_df)), metadata_df$EventID)
  
  sampled_indices <- unlist(
    lapply(split_indices, function(idx) {
      sample(idx, size = min(length(idx), max_per_event))
    }),
    use.names = FALSE
  )
  
  if (!is.null(max_total) && length(sampled_indices) > max_total) {
    sampled_indices <- sample(sampled_indices, size = max_total)
  }
  
  return(sampled_indices)
}

```

## Determine Number of Clusters (Elbow Method)

```{r}

elbow_index <- sample_by_event(
  metadata_df = click_metadata,
  max_per_event = max_clicks_per_event,
  seed = 1,
  max_total = elbow_sample_cap
)

AllSpec_elbow <- AllSpec_scale[elbow_index, ]

wss <- numeric(length(k_candidates))

for (i in seq_along(k_candidates)) {
  
  k <- k_candidates[i]
  
  km_tmp <- kmeans(
    AllSpec_elbow,
    centers = k,
    nstart = 20,
    algorithm = "Lloyd"
  )
  
  wss[i] <- km_tmp$tot.withinss
}

elbow_df <- data.frame(
  k = k_candidates,
  WSS = wss
)

elbow_plot <- ggplot(elbow_df, aes(x = k, y = WSS)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Elbow Plot",
    x = "Number of Clusters",
    y = "Total Within SS"
  )

print(elbow_plot)

ggsave(
  filename = file.path(output_dir, "kmeans_elbow_plot.png"),
  plot = elbow_plot,
  width = 7,
  height = 5
)

cat("Elbow sample size used:", nrow(AllSpec_elbow), "\n")

```

## Confirm Chosen Number of Clusters

```{r}

if (is.null(chosen_k)) {
  stop("Set chosen_k after reviewing the elbow plot before continuing.")
}

chosen_k

```

## Create Evaluation Sample

This sample is used for silhouette and PCA scatterplot.
It uses up to 200 clicks per event.

```{r}

eval_index <- sample_by_event(
  metadata_df = click_metadata,
  max_per_event = max_clicks_per_event,
  seed = 2,
  max_total = NULL
)

AllSpec_eval <- AllSpec_scale[eval_index, ]
click_metadata_eval <- click_metadata[eval_index, ]

cat("Evaluation sample size used:", nrow(AllSpec_eval), "\n")

```

## Sampling Diagnostics

```{r}

event_total <- click_metadata %>%
  count(EventID, name = "total_clicks")

event_sampled <- click_metadata_eval %>%
  count(EventID, name = "sampled_clicks")

sampling_table <- event_total %>%
  left_join(event_sampled, by = "EventID") %>%
  mutate(sampled_clicks = ifelse(is.na(sampled_clicks), 0, sampled_clicks)) %>%
  arrange(EventID)

sampling_table

```


## Run Final k-means on Evaluation Sample

```{r}

if (cluster_method == "kmeans") {
  
  set.seed(3)
  
  if (run_kmeans_on_full_dataset) {
    
    km <- kmeans(
      AllSpec_scale,
      centers = chosen_k,
      nstart = 50,
      algorithm = "Lloyd"
    )
    
    cluster_labels_full <- km$cluster
    click_metadata$cluster <- cluster_labels_full
    
    # Pull cluster labels for the evaluation subset
    cluster_labels_eval <- cluster_labels_full[eval_index]
    
  } else {
    
    km <- kmeans(
      AllSpec_eval,
      centers = chosen_k,
      nstart = 50,
      algorithm = "Lloyd"
    )
    
    cluster_labels_eval <- km$cluster
    click_metadata_eval$cluster <- cluster_labels_eval
  }
}

```

## Evaluate Clustering with Mean Silhouette Score

```{r}

# Distance matrix for evaluation sample
dist_eval <- dist(AllSpec_eval)

# Silhouette for the chosen k solution only
sil_chosen <- silhouette(cluster_labels_eval, dist_eval)
mean_silhouette <- mean(sil_chosen[, "sil_width"])

cat("Chosen k =", chosen_k, "mean silhouette =", mean_silhouette, "\n")

# Check other candidate k values without overwriting sil_chosen
silhouette_summary <- data.frame(
  k = integer(),
  mean_silhouette = numeric()
)

# Include chosen_k first
silhouette_summary <- rbind(
  silhouette_summary,
  data.frame(
    k = chosen_k,
    mean_silhouette = mean_silhouette
  )
)

# Evaluate other k values
for (k in setdiff(2:6, chosen_k)) {
  
  set.seed(3)
  
  km_test <- kmeans(
    AllSpec_eval,
    centers = k,
    nstart = 20,
    iter.max = 100,
    algorithm = "Lloyd"
  )
  
  sil_test <- silhouette(km_test$cluster, dist_eval)
  mean_sil_test <- mean(sil_test[, "sil_width"])
  
  silhouette_summary <- rbind(
    silhouette_summary,
    data.frame(
      k = k,
      mean_silhouette = mean_sil_test
    )
  )
  
  cat("k =", k, "mean silhouette =", mean_sil_test, "\n")
}

silhouette_summary <- silhouette_summary %>%
  arrange(k)

silhouette_summary

```

## Silhouette Plot

```{r}

sil_df <- as.data.frame(sil_chosen) %>%
  mutate(obs = row_number()) %>%
  group_by(cluster) %>%
  arrange(desc(sil_width), .by_group = TRUE) %>%
  mutate(order = row_number()) %>%
  ungroup()

sil_plot <- ggplot(sil_df, aes(x = order, y = sil_width)) +
  geom_col() +
  facet_wrap(~ cluster, scales = "free_x") +
  labs(
    title = paste("Silhouette Plot (k =", chosen_k, ")"),
    x = "Observations",
    y = "Silhouette Width"
  )

print(sil_plot)

ggsave(
  filename = file.path(output_dir, paste0("kmeans_silhouette_plot_", k_label, ".png")),
  plot = sil_plot,
  width = 8,
  height = 5
)

cat(
  "Silhouette plot saved to:",
  file.path(output_dir, paste0("kmeans_silhouette_plot_", k_label, ".png")),
  "\n"
)

```

## Assign Cluster Labels

```{r}

click_metadata_eval$cluster <- cluster_labels_eval

head(click_metadata_eval)

```

## PCA Scatterplot of Clustered Clicks

```{r}

pca_eval <- prcomp(AllSpec_eval)

scatter_df <- data.frame(
  PC1 = pca_eval$x[, 1],
  PC2 = pca_eval$x[, 2],
  EventID = click_metadata_eval$EventID,
  UID = click_metadata_eval$UID,
  cluster = as.factor(cluster_labels_eval)
)

scatter_plot <- ggplot(scatter_df, aes(x = PC1, y = PC2, color = cluster)) +
  geom_point(alpha = 0.7, size = 1.2) +
  labs(
    title = paste("Clustered Clicks (PCA Projection, k =", chosen_k, ")"),
    x = "PC1",
    y = "PC2",
    color = "Cluster"
  )

print(scatter_plot)

ggsave(
  filename = file.path(output_dir, paste0("kmeans_scatterplot_", k_label, ".png")),
  plot = scatter_plot,
  width = 7,
  height = 5
)

```

## Save Scatterplot Data

```{r}

write.csv(
  scatter_df,
  file = file.path(output_dir, paste0("kmeans_scatterplot_data_", k_label, ".csv")),
  row.names = FALSE
)

```

## Plot Cluster Centroids

```{r}

centroid_file <- file.path(output_dir, paste0("kmeans_centroid_plot_", k_label, ".png"))

png(
  filename = centroid_file,
  width = 7,
  height = 5,
  units = "in",
  res = 300
)

matplot(
  x = centroid_freq / 1000,
  y = t(km$centers),
  type = "l",
  lty = 1,
  main = paste("Cluster Centroids (", k_label, ")", sep = ""),
  xlab = "Frequency (kHz)",
  ylab = "Scaled Intensity"
)

dev.off()

```

## Save Centroid Spectra

```{r}

centroid_df <- data.frame(
  Frequency_kHz = centroid_freq / 1000,
  t(km$centers)
)

colnames(centroid_df) <- c(
  "Frequency_kHz",
  paste0("Cluster_", 1:chosen_k)
)

write.csv(
  centroid_df,
  file = file.path(output_dir, paste0("kmeans_centroid_spectra_", k_label, ".csv")),
  row.names = FALSE
)

```

## Save Summary Table

```{r}

summary_table <- data.frame(
  Method = cluster_method,
  KLabel = k_label,
  ChosenK = chosen_k,
  MeanSilhouette = mean_silhouette,
  RunKmeansOnFullDataset = run_kmeans_on_full_dataset,
  MaxClicksPerEvent = max_clicks_per_event,
  ElbowSampleSizeUsed = nrow(AllSpec_elbow),
  EvalSampleSizeUsed = nrow(AllSpec_eval),
  TotalClicksInFullDataset = nrow(AllSpec_scale)
)

summary_table

write.csv(
  summary_table,
  file = file.path(output_dir, paste0("kmeans_summary_table_", k_label, ".csv")),
  row.names = FALSE
)

dist(km$centers)

```











