---
title: "Clustering_kshape"
author: "Sarah Shiers"
date: "2026-03-24"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

```

This workflow prepares combined Tukey spectra for clustering by standardizing all events to a shared frequency grid, combining all click spectra into one matrix, and then running k-shape clustering in a way that parallels the k-means workflow.

This version:
- interpolates all spectra to a shared frequency grid
- restricts spectra to 19–50 kHz
- applies row-wise scaling to emphasize spectral shape
- uses an event-balanced sample for k-shape fitting and evaluation
- compares candidate k values on the evaluation sample
- fits one final k-shape model using the chosen k
- assigns all clicks in the full dataset to the nearest learned k-shape centroid using SBD in batches
- saves evaluation plots, centroid plots, centroid spectra, evaluation assignments, full-dataset assignments, and summary tables

## Install Packages

```{r}

install.packages(c("ggplot2", "dplyr", "cluster", "dtwclust", "tidyr", "proxy"))

```

## Load Required Packages

```{r}

library(ggplot2)
library(dplyr)
library(cluster)
library(dtwclust)
library(tidyr)
library(proxy)

```

## User Settings

```{r}

cluster_method <- "kshape"

output_dir <- "E:/Analysis/Clustering Output"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

freq_min <- 19000
freq_max <- 50000

# Candidate k values to evaluate on the sample
k_candidates <- 2:6

# Set after reviewing summary metrics
chosen_k <- 2
k_label <- paste0("k_", chosen_k)

# Sampling settings
max_clicks_per_event <- 200
eval_sample_cap <- NULL

# Final k-shape fit settings
kshape_nrep_eval <- 1L
kshape_nrep_final <- 1L

# Batch size for assigning full dataset to learned centroids
assignment_batch_size <- 5000

k_candidates <- c(2)

```

## Load Combined Tukey Spectra

```{r}

combined_file <- file.choose()
combined_tukey <- readRDS(combined_file)

class(combined_tukey)
length(combined_tukey)

```

## Inspect Frequency Grids

```{r}

reference_freq <- combined_tukey[[1]]$freq

freq_check <- data.frame(
  EventID = character(),
  SameAsReference = logical(),
  LengthFreq = integer(),
  FreqStep = numeric(),
  MaxFreq = numeric(),
  stringsAsFactors = FALSE
)

for (i in seq_along(combined_tukey)) {
  
  current_freq <- combined_tukey[[i]]$freq
  
  freq_check <- rbind(
    freq_check,
    data.frame(
      EventID = combined_tukey[[i]]$EventID,
      SameAsReference = identical(current_freq, reference_freq),
      LengthFreq = length(current_freq),
      FreqStep = current_freq[2] - current_freq[1],
      MaxFreq = max(current_freq),
      stringsAsFactors = FALSE
    )
  )
}

freq_check

```

## Choose Reference Frequency Grid

```{r}

reference_freq <- combined_tukey[[1]]$freq
reference_event <- combined_tukey[[1]]$EventID

cat("Reference event:", reference_event, "\n")
cat("Reference frequency step:", reference_freq[2] - reference_freq[1], "Hz\n")
cat("Reference max frequency:", max(reference_freq), "Hz\n")

```

## Interpolate Spectra to Shared Grid

```{r}

interpolate_spectrum <- function(old_freq, old_spec, new_freq) {
  approx(
    x = old_freq,
    y = old_spec,
    xout = new_freq,
    method = "linear",
    rule = 2
  )$y
}

interpolated_tukey <- vector("list", length(combined_tukey))

for (i in seq_along(combined_tukey)) {
  
  current_event <- combined_tukey[[i]]
  old_freq <- current_event$freq
  old_allSpec <- current_event$allSpec
  
  if (length(old_freq) == nrow(old_allSpec)) {
    
    interpolated_allSpec <- apply(
      old_allSpec,
      2,
      function(spec) interpolate_spectrum(old_freq, spec, reference_freq)
    )
    
  } else if (length(old_freq) == ncol(old_allSpec)) {
    
    interpolated_allSpec <- t(apply(
      old_allSpec,
      1,
      function(spec) interpolate_spectrum(old_freq, spec, reference_freq)
    ))
    
  } else {
    
    stop(
      paste0(
        "Frequency length does not match either dimension of allSpec for EventID: ",
        current_event$EventID,
        " | length(freq) = ", length(old_freq),
        " | dim(allSpec) = ", paste(dim(old_allSpec), collapse = " x ")
      )
    )
  }
  
  interpolated_allSpec <- as.matrix(interpolated_allSpec)
  
  interpolated_tukey[[i]] <- list(
    EventID = current_event$EventID,
    UID = current_event$UID,
    freq = reference_freq,
    allSpec = interpolated_allSpec
  )
}

```

## Confrim Interpolation

```{r}

all_match_after_interp <- all(
  sapply(interpolated_tukey, function(x) identical(x$freq, reference_freq))
)

all_match_after_interp

```

## Combine Spectra Across Events

```{r}

all_click_metadata <- list()
all_click_spectra <- list()

for (i in seq_along(interpolated_tukey)) {
  
  tukey_spectra <- interpolated_tukey[[i]]
  
  event_id <- tukey_spectra$EventID
  uid <- tukey_spectra$UID
  
  click_spectra <- t(tukey_spectra$allSpec)
  
  click_meta <- data.frame(
    EventID = event_id,
    UID = uid,
    stringsAsFactors = FALSE
  )
  
  all_click_metadata[[i]] <- click_meta
  all_click_spectra[[i]] <- click_spectra
}

click_metadata <- do.call(rbind, all_click_metadata)
all_spec_matrix <- do.call(rbind, all_click_spectra)

dim(all_spec_matrix)

```

## Restrict Spectra to 19–50 kHz

```{r}

keepFreq <- which(reference_freq >= freq_min & reference_freq <= freq_max)

AllSpec_Focus <- all_spec_matrix[, keepFreq]
centroid_freq <- reference_freq[keepFreq]

```

## Scale Spectra

```{r}

# Row-wise scaling emphasizes spectral shape within each click
AllSpec_scale <- t(scale(t(AllSpec_Focus)))
AllSpec_scale[is.na(AllSpec_scale)] <- 0

```

## Helper Function: Sample Up to N Clicks Per Event

```{r}

sample_by_event <- function(metadata_df, max_per_event, seed = NULL, max_total = NULL) {
  
  if (!is.null(seed)) {
    set.seed(seed)
  }
  
  split_indices <- split(seq_len(nrow(metadata_df)), metadata_df$EventID)
  
  sampled_indices <- unlist(
    lapply(split_indices, function(idx) {
      sample(idx, size = min(length(idx), max_per_event))
    }),
    use.names = FALSE
  )
  
  if (!is.null(max_total) && length(sampled_indices) > max_total) {
    sampled_indices <- sample(sampled_indices, size = max_total)
  }
  
  return(sampled_indices)
}
sample_by_event <- function(metadata_df, max_per_event, seed = NULL, max_total = NULL) {
  
  if (!is.null(seed)) {
    set.seed(seed)
  }
  
  split_indices <- split(seq_len(nrow(metadata_df)), metadata_df$EventID)
  
  sampled_indices <- unlist(
    lapply(split_indices, function(idx) {
      sample(idx, size = min(length(idx), max_per_event))
    }),
    use.names = FALSE
  )
  
  return(sampled_indices)
}
```

## Create Evaluation Sample

```{r}

eval_index <- sample_by_event(
  metadata_df = click_metadata,
  max_per_event = max_clicks_per_event,
  seed = 2,
  max_total = eval_sample_cap
)

AllSpec_eval <- AllSpec_scale[eval_index, ]
click_metadata_eval <- click_metadata[eval_index, ]

cat("Evaluation sample size used:", nrow(AllSpec_eval), "\n")

```
## Compute SBD Distance Matrix for Evaluation Sample

```{r}

cat("Computing SBD distance matrix for evaluation sample...\n")

sbd_dist_eval <- proxy::dist(
  x = AllSpec_eval,
  method = "SBD"
)

```


## Sampling Diagnostics

```{r}

event_total <- click_metadata %>%
  count(EventID, name = "total_clicks")

event_sampled <- click_metadata_eval %>%
  count(EventID, name = "sampled_clicks")

sampling_table <- event_total %>%
  left_join(event_sampled, by = "EventID") %>%
  mutate(sampled_clicks = ifelse(is.na(sampled_clicks), 0, sampled_clicks)) %>%
  arrange(EventID)

sampling_table

```

## Cluster Labelling

```{r}

get_cluster_labels <- function(obj) {
  
  # Case 1: labels stored directly in a list element
  if (is.list(obj) && "cluster" %in% names(obj) && !is.null(obj$cluster)) {
    return(as.integer(unlist(obj$cluster)))
  }
  
  # Case 2: labels stored as clustering
  if (is.list(obj) && "clustering" %in% names(obj) && !is.null(obj$clustering)) {
    return(as.integer(unlist(obj$clustering)))
  }
  
  # Case 3: dtwclust object nested inside a list
  if (is.list(obj) && "obj" %in% names(obj) && !is.null(obj$obj)) {
    
    inner <- obj$obj
    
    if (is.list(inner) && "cluster" %in% names(inner) && !is.null(inner$cluster)) {
      return(as.integer(unlist(inner$cluster)))
    }
    
    if (isS4(inner)) {
      if ("cluster" %in% slotNames(inner)) {
        return(as.integer(methods::slot(inner, "cluster")))
      }
      if ("clustering" %in% slotNames(inner)) {
        return(as.integer(methods::slot(inner, "clustering")))
      }
    }
  }
  
  # Case 4: S4 object directly
  if (isS4(obj)) {
    if ("cluster" %in% slotNames(obj)) {
      return(as.integer(methods::slot(obj, "cluster")))
    }
    if ("clustering" %in% slotNames(obj)) {
      return(as.integer(methods::slot(obj, "clustering")))
    }
  }
  
  stop(
    paste(
      "Could not find cluster labels.",
      "Check names(obj), class(obj), and str(obj, max.level = 2)."
    )
  )
}

```


## Helper Functions for k-shape

```{r}

run_kshape_clustering <- function(data_matrix, k, seed = 3, nrep = 3L) {
  
  set.seed(seed)
  
  data_matrix <- as.matrix(data_matrix)
  storage.mode(data_matrix) <- "double"
  
  if (anyNA(data_matrix)) {
    stop("Input to k-shape contains NA values.")
  }
  
  if (!all(is.finite(data_matrix))) {
    stop("Input to k-shape contains non-finite values.")
  }
  
  series_list <- lapply(seq_len(nrow(data_matrix)), function(i) {
    as.numeric(data_matrix[i, ])
  })
  
  clust <- dtwclust::tsclust(
    series = series_list,
    type = "partitional",
    k = k,
    preproc = dtwclust::zscore,
    distance = "sbd",
    centroid = "shape",
    seed = seed,
    trace = FALSE,
    control = dtwclust::partitional_control(nrep = nrep)
  )
  
  return(clust)
}

extract_kshape_centroids <- function(kshape_obj, freq_vector) {
  
  cents <- kshape_obj@centroids
  
  centroid_mat <- do.call(
    rbind,
    lapply(cents, function(x) as.numeric(x))
  )
  
  centroid_df <- data.frame(
    Frequency_kHz = freq_vector / 1000
  )
  
  for (i in seq_len(nrow(centroid_mat))) {
    centroid_df[[paste0("Cluster_", i)]] <- centroid_mat[i, ]
  }
  
  return(centroid_df)
}

plot_centroids_kshape <- function(centroid_df, k) {
  
  centroid_long <- tidyr::pivot_longer(
    centroid_df,
    cols = starts_with("Cluster_"),
    names_to = "Cluster",
    values_to = "ScaledIntensity"
  )
  
  ggplot(centroid_long, aes(x = Frequency_kHz, y = ScaledIntensity, color = Cluster)) +
    geom_line(linewidth = 0.8) +
    labs(
      title = paste("k-shape Centroids (k =", k, ")"),
      x = "Frequency (kHz)",
      y = "Scaled Intensity",
      color = "Cluster"
    )
}

assign_to_kshape_centroids <- function(data_matrix, centroid_matrix, batch_size = 5000) {
  
  n <- nrow(data_matrix)
  assignments <- integer(n)
  
  for (start_idx in seq(1, n, by = batch_size)) {
    
    end_idx <- min(start_idx + batch_size - 1, n)
    batch <- data_matrix[start_idx:end_idx, , drop = FALSE]
    
    batch_assign <- apply(batch, 1, function(x) {
      dists <- sapply(seq_len(nrow(centroid_matrix)), function(j) {
        as.numeric(proxy::dist(
          x = matrix(as.numeric(x), nrow = 1),
          y = matrix(as.numeric(centroid_matrix[j, ]), nrow = 1),
          method = "SBD"
        ))
      })
      which.min(dists)
    })
    
    assignments[start_idx:end_idx] <- batch_assign
    
    cat("Assigned rows", start_idx, "to", end_idx, "\n")
  }
  
  return(assignments)
}
```

## Evaluate Candidate k Values on the Evaluation Sample

```{r}

results_summary_list <- list()

k <- chosen_k

cat("\n============================\n")
cat("Running k-shape for chosen k =", k, "\n")
cat("============================\n")

ks_eval_tmp <- run_kshape_clustering(
  data_matrix = AllSpec_eval,
  k = k,
  seed = 100 + k,
  nrep = kshape_nrep_eval
)

cluster_labels_tmp <- as.integer(predict(ks_eval_tmp))

sil_tmp <- cluster::silhouette(cluster_labels_tmp, sbd_dist_eval)
mean_silhouette <- mean(sil_tmp[, "sil_width"])

results_summary_list[[paste0("k_", k)]] <- data.frame(
  Method = cluster_method,
  KLabel = paste0("k_", k),
  K = k,
  MeanSilhouette = mean_silhouette,
  EvalSampleSizeUsed = nrow(AllSpec_eval),
  TotalClicksInFullDataset = nrow(AllSpec_scale)
)

cat("k =", k, "mean silhouette =", mean_silhouette, "\n")

summary_table_all_k <- do.call(rbind, results_summary_list)

summary_table_all_k

write.csv(
  summary_table_all_k,
  file = file.path(output_dir, "kshape_summary_table_all_k.csv"),
  row.names = FALSE
)

```

## Confirm Chosen Number of Clusters

```{r}

if (is.null(chosen_k)) {
  stop("Set chosen_k after reviewing the k-shape summary table before continuing.")
}

chosen_k

```

## Run Final k-shape on Evaluation Sample

```{r}

ks_final <- run_kshape_clustering(
  data_matrix = AllSpec_eval,
  k = chosen_k,
  seed = 999,
  nrep = kshape_nrep_final
)

cluster_labels_eval <- as.integer(predict(ks_final))
click_metadata_eval$cluster <- cluster_labels_eval

```

## Concatenated Spectrograms by Cluster (evaluation sample, scaled spectra)

```{r}

# Output one concatenated spectrogram per cluster using the scaled evaluation sample
# This uses the already-restricted 19–50 kHz spectra in AllSpec_eval

cluster_ids <- sort(unique(cluster_labels_eval))

spec_cols <- colorRampPalette(c("white", "#FFD54F", "#FF9800", "#D84315", "#4E0012"))(200)

for (cl in cluster_ids) {
  
  cluster_idx <- which(cluster_labels_eval == cl)
  
  if (length(cluster_idx) == 0) next
  
  cluster_spec <- AllSpec_eval[cluster_idx, , drop = FALSE]
  
  out_file <- paste0(
    "E:/Analysis/Clustering Output/",
    "kshape_concatenated_spectrogram_eval_scaled_19_50kHz_",
    k_label,
    "_cluster_",
    cl,
    ".png"
  )
  
  png(out_file, width = 1800, height = 900)
  
  image(
    x = seq_len(nrow(cluster_spec)),
    y = centroid_freq / 1000,
    z = cluster_spec,
    col = spec_cols,
    xlab = "Click Number",
    ylab = "Frequency (kHz)",
    main = paste("Concatenated Click Spectrogram - Cluster", cl),
    useRaster = TRUE,
    axes = FALSE
  )
  
  axis(1)
  axis(2, at = seq(20, 50, by = 10))
  box()
  
  dev.off()
  
  cat("Saved:", out_file, "\n")
}

```

## Mean Spectrum with 95% CI by Cluster (evaluation sample, scaled spectra)

```{r}

cluster_ids <- sort(unique(cluster_labels_eval))

for (cl in cluster_ids) {
  
  cluster_idx <- which(cluster_labels_eval == cl)
  
  if (length(cluster_idx) < 2) next
  
  cluster_spec <- AllSpec_eval[cluster_idx, , drop = FALSE]
  
  freq_khz <- centroid_freq / 1000
  
  # Mean at each frequency bin
  mean_spec <- colMeans(cluster_spec)
  
  # 95% central interval at each frequency bin
  lower <- apply(cluster_spec, 2, quantile, probs = 0.025, na.rm = TRUE)
  upper <- apply(cluster_spec, 2, quantile, probs = 0.975, na.rm = TRUE)
  
  out_file <- paste0(
    "E:/Analysis/Clustering Output/",
    "kshape_mean_spectrum_quantile_band_eval_scaled_19_50kHz_",
    k_label,
    "_cluster_",
    cl,
    ".png"
  )
  
  y_lim <- range(c(lower, upper), na.rm = TRUE)
  
  png(out_file, width = 1200, height = 700)
  
  plot(
    freq_khz, mean_spec,
    type = "n",
    xlab = "Frequency (kHz)",
    ylab = "Scaled Intensity",
    main = paste("Mean Spectrum with 95% Quantile Band - Cluster", cl),
    ylim = y_lim
  )
  
  polygon(
    c(freq_khz, rev(freq_khz)),
    c(upper, rev(lower)),
    col = rgb(1, 0.5, 0, 0.3),
    border = NA
  )
  
  lines(freq_khz, mean_spec, lwd = 2)
  
  dev.off()
  
  cat("Saved:", out_file, "\n")
}  

```

## Final Mean Silhouette for Chosen k

```{r}

sil_final <- cluster::silhouette(cluster_labels_eval, sbd_dist_eval)
mean_silhouette_final <- mean(sil_final[, "sil_width"])

cat("Chosen k =", chosen_k, "mean silhouette =", mean_silhouette_final, "\n")

```

## PCA Scatterplot of Clustered Evaluation Sample

```{r}

pca_eval <- prcomp(AllSpec_eval)

scatter_df <- data.frame(
  PC1 = pca_eval$x[, 1],
  PC2 = pca_eval$x[, 2],
  EventID = click_metadata_eval$EventID,
  UID = click_metadata_eval$UID,
  cluster = as.factor(cluster_labels_eval)
)

scatter_plot <- ggplot(scatter_df, aes(x = PC1, y = PC2, color = cluster)) +
  geom_point(alpha = 0.7, size = 1.2) +
  labs(
    title = paste("k-shape Clustered Clicks (PCA Projection,", k_label, ")"),
    x = "PC1",
    y = "PC2",
    color = "Cluster"
  )

print(scatter_plot)

ggsave(
  filename = file.path(output_dir, paste0("kshape_pca_scatterplot_", k_label, ".png")),
  plot = scatter_plot,
  width = 7,
  height = 5
)

write.csv(
  scatter_df,
  file = file.path(output_dir, paste0("kshape_pca_scatterplot_data_", k_label, ".csv")),
  row.names = FALSE
)

```

## SBD-based MDS Plot of Clustered Evaluation Sample

```{r}

cat("Running MDS on precomputed SBD distance matrix...\n")

mds_fit <- cmdscale(sbd_dist_eval, k = 3)

mds_df <- data.frame(
  Dim1 = mds_fit[, 1],
  Dim2 = mds_fit[, 2],
  EventID = click_metadata_eval$EventID,
  UID = click_metadata_eval$UID,
  cluster = as.factor(cluster_labels_eval)
)

mds_plot <- ggplot(mds_df, aes(x = Dim1, y = Dim2, color = cluster)) +
  geom_point(alpha = 0.7, size = 1.2) +
  labs(
    title = paste("k-shape Clustered Clicks (SBD-based MDS,", k_label, ")"),
    x = "MDS Dimension 1",
    y = "MDS Dimension 2",
    color = "Cluster"
  )

print(mds_plot)

ggsave(
  filename = file.path(output_dir, paste0("kshape_mds_scatterplot_", k_label, ".png")),
  plot = mds_plot,
  width = 7,
  height = 5
)

write.csv(
  mds_df,
  file = file.path(output_dir, paste0("kshape_mds_scatterplot_data_", k_label, ".csv")),
  row.names = FALSE
)

```


## Plot Cluster Centroids

```{r}

centroid_df <- extract_kshape_centroids(ks_final, centroid_freq)

centroid_plot <- plot_centroids_kshape(centroid_df, chosen_k)
print(centroid_plot)

ggsave(
  filename = file.path(output_dir, paste0("kshape_centroid_plot_", k_label, ".png")),
  plot = centroid_plot,
  width = 7,
  height = 5
)

write.csv(
  centroid_df,
  file = file.path(output_dir, paste0("kshape_centroid_spectra_", k_label, ".csv")),
  row.names = FALSE
)

```

## Save Evaluation Assignments

```{r}

write.csv(
  click_metadata_eval,
  file = file.path(output_dir, paste0("kshape_eval_cluster_assignments_", k_label, ".csv")),
  row.names = FALSE
)

```

## Assign Full Dataset to Learned Centroids in Batches

```{r}

centroid_matrix <- t(as.matrix(centroid_df[, -1]))

cluster_labels_full <- assign_to_kshape_centroids(
  data_matrix = AllSpec_scale,
  centroid_matrix = centroid_matrix,
  batch_size = assignment_batch_size
)

click_metadata_full <- click_metadata
click_metadata_full$cluster <- cluster_labels_full

```

## Save Full-Dataset Assignments

```{r}

write.csv(
  click_metadata_full,
  file = file.path(output_dir, paste0("kshape_full_cluster_assignments_", k_label, ".csv")),
  row.names = FALSE
)

```

## Save Final Summary Table

```{r}

summary_table_final <- data.frame(
  Method = cluster_method,
  KLabel = k_label,
  ChosenK = chosen_k,
  MeanSilhouette = mean_silhouette_final,
  MaxClicksPerEvent = max_clicks_per_event,
  EvalSampleSizeUsed = nrow(AllSpec_eval),
  TotalClicksInFullDataset = nrow(AllSpec_scale),
  AssignmentBatchSize = assignment_batch_size,
  MDSBasedOn = "SBD"
)

summary_table_final

write.csv(
  summary_table_final,
  file = file.path(output_dir, paste0("kshape_summary_table_", k_label, ".csv")),
  row.names = FALSE
)

```

## Solhouette Plot

```{r}

sil_df <- as.data.frame(sil_final) %>%
  mutate(obs = row_number()) %>%
  group_by(cluster) %>%
  arrange(desc(sil_width), .by_group = TRUE) %>%
  mutate(order = row_number()) %>%
  ungroup()

sil_plot <- ggplot(sil_df, aes(x = order, y = sil_width)) +
  geom_col() +
  facet_wrap(~ cluster, scales = "free_x") +
  labs(
    title = paste("Silhouette Plot (k-shape,", k_label, ")"),
    x = "Observations",
    y = "Silhouette Width"
  )

print(sil_plot)

ggsave(
  filename = file.path(output_dir, paste0("kshape_silhouette_plot_", k_label, ".png")),
  plot = sil_plot,
  width = 8,
  height = 5
)

write.csv(
  sil_df,
  file = file.path(output_dir, paste0("kshape_silhouette_plot_data_", k_label, ".csv")),
  row.names = FALSE
)

```

##  Elbow-like Plot for k-shape Using Total Within-Cluster SBD Distance

```{r}

elbow_plot_k_candidates <- 2:6

kshape_within_sbd <- numeric(length(elbow_plot_k_candidates))

for (i in seq_along(elbow_plot_k_candidates)) {
  
  k <- elbow_plot_k_candidates[i]
  
  cat("\n============================\n")
  cat("Running k-shape elbow-like evaluation for k =", k, "\n")
  cat("============================\n")
  
  ks_tmp <- run_kshape_clustering(
    data_matrix = AllSpec_eval,
    k = k,
    seed = 200 + k,
    nrep = 1L
  )
  
  cluster_labels_tmp <- as.integer(predict(ks_tmp))
  centroid_df_tmp <- extract_kshape_centroids(ks_tmp, centroid_freq)
  centroid_matrix_tmp <- t(as.matrix(centroid_df_tmp[, -1]))
  
  total_within_sbd <- 0
  
  for (j in seq_len(nrow(AllSpec_eval))) {
    
    cl <- cluster_labels_tmp[j]
    
    d_j <- as.numeric(proxy::dist(
      x = matrix(AllSpec_eval[j, ], nrow = 1),
      y = matrix(centroid_matrix_tmp[cl, ], nrow = 1),
      method = "SBD"
    ))
    
    total_within_sbd <- total_within_sbd + d_j
  }
  
  kshape_within_sbd[i] <- total_within_sbd
  
  cat("k =", k, "total within-cluster SBD =", total_within_sbd, "\n")
}

kshape_elbow_df <- data.frame(
  k = elbow_plot_k_candidates,
  TotalWithinSBD = kshape_within_sbd
)

kshape_elbow_plot <- ggplot(kshape_elbow_df, aes(x = k, y = TotalWithinSBD)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Elbow-like Plot for k-shape",
    x = "Number of Clusters",
    y = "Total Within-Cluster SBD"
  )

print(kshape_elbow_plot)

ggsave(
  filename = file.path(output_dir, "kshape_elbow_like_plot.png"),
  plot = kshape_elbow_plot,
  width = 7,
  height = 5
)

write.csv(
  kshape_elbow_df,
  file = file.path(output_dir, "kshape_elbow_like_plot_data.csv"),
  row.names = FALSE
)

cat("Evaluation sample size used for k-shape elbow-like plot:", nrow(AllSpec_eval), "\n")


```

## Silhouette Scores for k=2:6 on Eval Sample

```{r}

silhouette_k_candidates <- 2:6

# Silhouette for the chosen k solution only
sil_chosen <- cluster::silhouette(cluster_labels_eval, sbd_dist_eval)
mean_silhouette <- mean(sil_chosen[, "sil_width"])

cat("Chosen k =", chosen_k, "mean silhouette =", mean_silhouette, "\n")

# Check other candidate k values without overwriting sil_chosen
silhouette_summary <- data.frame(
  k = integer(),
  mean_silhouette = numeric()
)

# Include chosen_k first
silhouette_summary <- rbind(
  silhouette_summary,
  data.frame(
    k = chosen_k,
    mean_silhouette = mean_silhouette
  )
)

# Evaluate other candidate k values
for (k in setdiff(silhouette_k_candidates, chosen_k)) {
  
  ks_tmp <- run_kshape_clustering(
    data_matrix = AllSpec_eval,
    k = k,
    seed = 300 + k,
    nrep = 1L
  )
  
  cluster_labels_tmp <- as.integer(predict(ks_tmp))
  
  sil_tmp <- cluster::silhouette(cluster_labels_tmp, sbd_dist_eval)
  mean_sil_tmp <- mean(sil_tmp[, "sil_width"])
  
  silhouette_summary <- rbind(
    silhouette_summary,
    data.frame(
      k = k,
      mean_silhouette = mean_sil_tmp
    )
  )
  
  cat("k =", k, "mean silhouette =", mean_sil_tmp, "\n")
}

silhouette_summary <- silhouette_summary %>%
  arrange(k)

silhouette_summary

```

## Full Dataset Cluster Variability

```{r}

cat("Computing full dataset cluster variability for k-shape...\n")

# Distance of every click to its assigned k-shape centroid using SBD
full_dist_to_centroid <- numeric(nrow(AllSpec_scale))

for (i in seq_len(nrow(AllSpec_scale))) {
  
  this_cluster <- cluster_labels_full[i]
  this_centroid <- centroid_matrix[this_cluster, ]
  
  full_dist_to_centroid[i] <- as.numeric(proxy::dist(
    x = matrix(AllSpec_scale[i, ], nrow = 1),
    y = matrix(this_centroid, nrow = 1),
    method = "SBD"
  ))
}

# Add to full metadata
click_metadata_full$dist_to_centroid <- full_dist_to_centroid

# Summary statistics: within-cluster variability
full_cluster_summary <- click_metadata_full %>%
  group_by(cluster) %>%
  summarise(
    n_clicks = n(),
    mean_dist = mean(dist_to_centroid),
    sd_dist = sd(dist_to_centroid),
    median_dist = median(dist_to_centroid),
    .groups = "drop"
  )

full_cluster_summary

write.csv(
  full_cluster_summary,
  file = file.path(output_dir, paste0("kshape_FULL_cluster_variability_", k_label, ".csv")),
  row.names = FALSE
)

# Between-cluster variability: pairwise distances among centroids
centroid_distances <- as.matrix(proxy::dist(
  x = centroid_matrix,
  method = "SBD"
))

centroid_distances

```

