---
title: "Diel_Plot"
author: "Sarah Moseley"
date: "2026-05-09"
output: html_document
---

## Setup

```{r setup, message=FALSE, warning=FALSE}

library(tidyverse)
library(lubridate)
library(suncalc)
library(scales)

theme_set(theme_bw(base_size = 13))

```


## Chunk 3 — paths

```{r paths}

analysis_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1"

event_tables_dir <- file.path(
  analysis_dir,
  "mapping_outputs/by_event/tables"
)

output_dir <- file.path(
  analysis_dir,
  "diel_event_plots"
)

dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

event_comp_with_noise_file <- file.path(
  event_tables_dir,
  "event_cluster_composition_WITH_noise_18PC_scaledICI_1_minPts50.csv"
)

event_comp_no_noise_file <- file.path(
  event_tables_dir,
  "event_cluster_composition_NO_noise_18PC_scaledICI_1_minPts50.csv"
)

detections_file <- "E:/Logs/ADRIFT_GgDetections.csv"

```


## Chunk 4 — load data

```{r load-data}

event_comp_with_noise <- read_csv(event_comp_with_noise_file, show_col_types = FALSE)
event_comp_no_noise   <- read_csv(event_comp_no_noise_file, show_col_types = FALSE)
detections            <- read_csv(detections_file, show_col_types = FALSE)

glimpse(event_comp_with_noise)
glimpse(event_comp_no_noise)
glimpse(detections)

```

## Clean and Standardize IDs

```{r clean-ids}

clean_event_id <- function(x) {
  x %>%
    as.character() %>%
    str_trim() %>%
    str_replace_all("\\s+", "_")
}

event_comp_with_noise <- event_comp_with_noise %>%
  mutate(
    EventID_clean = clean_event_id(EventID_clean),
    DriftID_clean = as.character(DriftID_clean),
    DriftName = as.character(DriftName),
    HDBSCAN_cluster = as.character(HDBSCAN_cluster)
  )

event_comp_no_noise <- event_comp_no_noise %>%
  mutate(
    EventID_clean = clean_event_id(EventID_clean),
    DriftID_clean = as.character(DriftID_clean),
    DriftName = as.character(DriftName),
    HDBSCAN_cluster = as.character(HDBSCAN_cluster)
  )

detections %>%
  filter(EventID_clean %in% event_comp_with_noise$EventID_clean) %>%
  count(DriftName)

# Quick QA checks
event_comp_with_noise %>% count(HDBSCAN_cluster)
event_comp_no_noise %>% count(HDBSCAN_cluster)

detections %>%
  filter(EventID_clean %in% event_comp_with_noise$EventID_clean) %>%
  count(DriftName)

# Confirm detection table matches the 48 events used in this analysis
n_used_events_in_detections <- detections %>%
  filter(EventID_clean %in% event_comp_with_noise$EventID_clean) %>%
  distinct(EventID_clean) %>%
  nrow()

cat("Number of used analysis events found in detection table:", n_used_events_in_detections, "\n")

stopifnot(n_used_events_in_detections == 48)

```

## Identify Dominant Click Type per Event

```{r}

dominant_with_unassigned <- event_comp_with_noise %>%
  group_by(EventID_clean) %>%
  arrange(desc(n_clicks), HDBSCAN_cluster, .by_group = TRUE) %>%
  slice(1) %>%
  ungroup() %>%
  transmute(
    EventID_clean,
    DriftID_clean,
    DriftName,
    Latitude,
    Longitude,
    dominant_cluster_with_unassigned = HDBSCAN_cluster,
    dominant_cluster_with_unassigned_n = n_clicks,
    dominant_cluster_with_unassigned_percent = percent,
    total_clicks_event
  )

dominant_assigned_only <- event_comp_no_noise %>%
  filter(HDBSCAN_cluster != "0") %>%
  group_by(EventID_clean) %>%
  arrange(desc(n_clicks), HDBSCAN_cluster, .by_group = TRUE) %>%
  slice(1) %>%
  ungroup() %>%
  transmute(
    EventID_clean,
    dominant_cluster_assigned_only = HDBSCAN_cluster,
    dominant_cluster_assigned_only_n = n_clicks,
    dominant_cluster_assigned_only_percent = percent
  )

# QA: should be 48 events
cat("Dominant events including unassigned:", nrow(dominant_with_unassigned), "\n")
cat("Dominant events assigned only:", nrow(dominant_assigned_only), "\n")

stopifnot(nrow(dominant_with_unassigned) == 48)

# Check dominant cluster counts
dominant_with_unassigned %>%
  count(dominant_cluster_with_unassigned, name = "n_events")

dominant_assigned_only %>%
  count(dominant_cluster_assigned_only, name = "n_events")

```

## Join Event Timing from Detection Table

```{r}

event_timing <- detections %>%
  filter(EventID_clean %in% dominant_with_unassigned$EventID_clean) %>%
  mutate(
    UTC_start_raw = UTC,
    UTC_end_raw = end,

    UTC_start = ymd_hms(UTC_start_raw, tz = "UTC"),
    UTC_end = ymd_hms(UTC_end_raw, tz = "UTC"),

    event_duration_min = as.numeric(
      difftime(UTC_end, UTC_start, units = "mins")
    ),

    date_utc = as.Date(UTC_start),

    hour_utc = hour(UTC_start) +
      minute(UTC_start) / 60 +
      second(UTC_start) / 3600,

    end_hour_utc = hour(UTC_end) +
      minute(UTC_end) / 60 +
      second(UTC_end) / 3600
  ) %>%
  select(
    EventID_clean,
    UTC_start_raw,
    UTC_end_raw,
    UTC_start,
    UTC_end,
    event_duration_min,
    date_utc,
    hour_utc,
    end_hour_utc,
    species,
    call,
    DataQuality,
    gpsUncertainty
  ) %>%
  distinct(EventID_clean, .keep_all = TRUE)

diel_events <- dominant_with_unassigned %>%
  left_join(dominant_assigned_only, by = "EventID_clean") %>%
  left_join(event_timing, by = "EventID_clean")

# QA checks
diel_events %>%
  summarise(
    n_events = n(),
    n_with_start_time = sum(!is.na(UTC_start)),
    n_missing_start_time = sum(is.na(UTC_start)),
    n_with_end_time = sum(!is.na(UTC_end)),
    n_missing_end_time = sum(is.na(UTC_end)),
    min_date = min(date_utc, na.rm = TRUE),
    max_date = max(date_utc, na.rm = TRUE),
    median_duration_min = median(event_duration_min, na.rm = TRUE),
    max_duration_min = max(event_duration_min, na.rm = TRUE)
  )

stopifnot(nrow(diel_events) == 48)
stopifnot(sum(is.na(diel_events$UTC_start)) == 0)

```

## Calculate Median Event Location

```{r}

median_lat <- median(diel_events$Latitude, na.rm = TRUE)
median_lon <- median(diel_events$Longitude, na.rm = TRUE)

cat("Median latitude of used events:", median_lat, "\n")
cat("Median longitude of used events:", median_lon, "\n")

# QA: make sure location is present for all 48 analyzed events
diel_events %>%
  summarise(
    n_events = n(),
    n_with_latitude = sum(!is.na(Latitude)),
    n_with_longitude = sum(!is.na(Longitude)),
    min_latitude = min(Latitude, na.rm = TRUE),
    max_latitude = max(Latitude, na.rm = TRUE),
    min_longitude = min(Longitude, na.rm = TRUE),
    max_longitude = max(Longitude, na.rm = TRUE)
  )

stopifnot(nrow(diel_events) == 48)
stopifnot(sum(is.na(diel_events$Latitude)) == 0)
stopifnot(sum(is.na(diel_events$Longitude)) == 0)

```

## Calculate Sunrise and Sunset Times

```{r}

date_sequence <- seq(
  min(diel_events$date_utc, na.rm = TRUE),
  max(diel_events$date_utc, na.rm = TRUE),
  by = "day"
)

sun_times <- getSunlightTimes(
  date = date_sequence,
  lat = median_lat,
  lon = median_lon,
  keep = c("sunrise", "sunset"),
  tz = "UTC"
) %>%
  as_tibble() %>%
  mutate(
    date_utc = as.Date(date),

    sunrise_hour_utc = hour(sunrise) +
      minute(sunrise) / 60 +
      second(sunrise) / 3600,

    sunset_hour_utc = hour(sunset) +
      minute(sunset) / 60 +
      second(sunset) / 3600
  ) %>%
  select(
    date_utc,
    sunrise,
    sunset,
    sunrise_hour_utc,
    sunset_hour_utc
  )

# QA
sun_times %>%
  summarise(
    n_dates = n(),
    min_date = min(date_utc),
    max_date = max(date_utc),
    min_sunrise_hour_utc = min(sunrise_hour_utc, na.rm = TRUE),
    max_sunrise_hour_utc = max(sunrise_hour_utc, na.rm = TRUE),
    min_sunset_hour_utc = min(sunset_hour_utc, na.rm = TRUE),
    max_sunset_hour_utc = max(sunset_hour_utc, na.rm = TRUE)
  )

head(sun_times)

```

## Correct Sunrise and Sunset Hours for UTC Day Wrap

```{r}

sun_times <- sun_times %>%
  mutate(
    sunset_date_utc = as.Date(sunset),
    sunset_hour_utc_raw = sunset_hour_utc,

    sunset_hour_utc = if_else(
      sunset_date_utc > date_utc,
      sunset_hour_utc + 24,
      sunset_hour_utc
    )
  )

# QA: sunset should now be after sunrise
sun_times %>%
  summarise(
    n_dates = n(),
    min_sunrise_hour_utc = min(sunrise_hour_utc, na.rm = TRUE),
    max_sunrise_hour_utc = max(sunrise_hour_utc, na.rm = TRUE),
    min_sunset_hour_utc = min(sunset_hour_utc, na.rm = TRUE),
    max_sunset_hour_utc = max(sunset_hour_utc, na.rm = TRUE),
    any_sunset_before_sunrise = any(sunset_hour_utc < sunrise_hour_utc, na.rm = TRUE)
  )

stopifnot(!any(sun_times$sunset_hour_utc < sun_times$sunrise_hour_utc, na.rm = TRUE))

```

## Join Sunrise and Sunset Times to Events

```{r}

diel_events <- diel_events %>%
  left_join(
    sun_times %>%
      select(
        date_utc,
        sunrise,
        sunset,
        sunrise_hour_utc,
        sunset_hour_utc
      ),
    by = "date_utc"
  ) %>%
  mutate(
    hour_utc_plot = hour_utc,

    period_utc = case_when(
      hour_utc_plot < sunrise_hour_utc ~ "Night",
      hour_utc_plot >= sunrise_hour_utc & hour_utc_plot <= 24 ~ "Day",
      TRUE ~ NA_character_
    )
  )

# QA
diel_events %>%
  count(period_utc)

write_csv(
  diel_events,
  file.path(output_dir, "diel_events_with_dominant_click_type_and_sun_times.csv")
)

```

## Define Cluster Colors

```{r}

rainbow_cluster_colors <- c(
  "1"  = "#E41A1C",  # red
  "2"  = "#FF7F00",  # orange
  "3"  = "#FFD92F",  # yellow
  "4"  = "#4DAF4A",  # green
  "5"  = "#00BFC4",  # cyan
  "6"  = "#377EB8",  # blue
  "7"  = "#984EA3",  # purple
  "8"  = "#F781BF",  # pink
  "9"  = "#A65628",  # brown
  "10" = "#999999",
  "11" = "#66C2A5",
  "12" = "#FC8D62",
  "13" = "#8DA0CB",
  "14" = "#E78AC3",
  "15" = "#A6D854"
)

all_clusters <- sort(unique(as.character(diel_events$dominant_cluster_with_unassigned)))
non_noise <- setdiff(all_clusters, "0")

cluster_colors_with_unassigned <- c(
  "0" = "grey35",
  rainbow_cluster_colors[non_noise]
)

cluster_colors_assigned_only <- rainbow_cluster_colors[
  sort(unique(as.character(diel_events$dominant_cluster_assigned_only)))
]

cluster_colors_with_unassigned
cluster_colors_assigned_only

```

## Plot Diel Detections Including Cluster 0

```{r}

p_diel_with_unassigned <- ggplot(diel_events, aes(x = hour_utc, y = date_utc)) +
  geom_rect(
    data = sun_times,
    aes(
      xmin = 0,
      xmax = sunrise_hour_utc,
      ymin = date_utc - 0.5,
      ymax = date_utc + 0.5
    ),
    inherit.aes = FALSE,
    fill = "grey90",
    alpha = 0.9
  ) +
  geom_rect(
    data = sun_times,
    aes(
      xmin = sunset_hour_utc,
      xmax = 28,
      ymin = date_utc - 0.5,
      ymax = date_utc + 0.5
    ),
    inherit.aes = FALSE,
    fill = "grey90",
    alpha = 0.9
  ) +
  geom_point(
    aes(color = dominant_cluster_with_unassigned),
    size = 3,
    alpha = 0.95
  ) +
  scale_color_manual(values = cluster_colors_with_unassigned, drop = FALSE) +
  scale_x_continuous(
    limits = c(0, 28),
    breaks = seq(0, 24, by = 3),
    labels = function(x) sprintf("%02d:00", x %% 24)
  ) +
  scale_y_date(date_breaks = "6 months", date_labels = "%b %Y") +
  labs(
    title = "Diel Distribution of Risso's Dolphin Events",
    subtitle = paste0(
      "Colored by predominant cluster including cluster 0; sunrise/sunset from median location (",
      round(median_lat, 3), ", ", round(median_lon, 3), ")"
    ),
    x = "Hour of day UTC",
    y = "Date UTC",
    color = "Dominant cluster"
  ) +
  theme_bw(base_size = 13) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "right"
  )

p_diel_with_unassigned

ggsave(
  file.path(output_dir, "diel_plot_WITH_cluster0.png"),
  p_diel_with_unassigned,
  width = 10,
  height = 7,
  dpi = 300
)

```

## Plot Diel Detections Including Cluster 0 — Event-Date Version

```{r}

# ------------------------------------------------------------
# Convert event and sun times to fixed local plotting time: UTC-7
# This matches the example style: h (UTC-7.0)
# ------------------------------------------------------------

utc_offset_hours <- -7

plot_dates <- diel_events %>%
  distinct(date_utc) %>%
  arrange(date_utc) %>%
  mutate(date_index = row_number())

plot_events_with_unassigned <- diel_events %>%
  left_join(plot_dates, by = "date_utc") %>%
  mutate(
    hour_local = (hour_utc + utc_offset_hours) %% 24,
    dominant_cluster_with_unassigned = factor(
      dominant_cluster_with_unassigned,
      levels = names(cluster_colors_with_unassigned)
    )
  )

plot_sun_times <- sun_times %>%
  inner_join(plot_dates, by = "date_utc") %>%
  mutate(
    sunrise_hour_local = (sunrise_hour_utc + utc_offset_hours) %% 24,
    sunset_hour_local = ((sunset_hour_utc %% 24) + utc_offset_hours) %% 24
  )

# ------------------------------------------------------------
# Plot
# ------------------------------------------------------------

p_diel_with_unassigned <- ggplot(
  plot_events_with_unassigned,
  aes(x = hour_local, y = date_index)
) +
  # Night before sunrise
  geom_rect(
    data = plot_sun_times,
    aes(
      xmin = 0,
      xmax = sunrise_hour_local,
      ymin = date_index - 0.45,
      ymax = date_index + 0.45
    ),
    inherit.aes = FALSE,
    fill = "grey85",
    alpha = 0.9
  ) +
  # Night after sunset
  geom_rect(
    data = plot_sun_times,
    aes(
      xmin = sunset_hour_local,
      xmax = 24,
      ymin = date_index - 0.45,
      ymax = date_index + 0.45
    ),
    inherit.aes = FALSE,
    fill = "grey85",
    alpha = 0.9
  ) +
  geom_linerange(
    aes(
      ymin = date_index - 0.28,
      ymax = date_index + 0.28,
      color = dominant_cluster_with_unassigned
    ),
    linewidth = 1.0,
    alpha = 1
  ) +
  scale_color_manual(values = cluster_colors_with_unassigned, drop = FALSE) +
  scale_x_continuous(
    limits = c(0, 24),
    breaks = seq(0, 24, by = 3)
  ) +
  scale_y_continuous(
    breaks = plot_dates$date_index,
    labels = format(plot_dates$date_utc, "%d-%b-%Y"),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  labs(
    x = "h (UTC-7.0)",
    y = NULL,
    color = "Dominant cluster"
  ) +
  theme_bw(base_size = 13) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "right"
  )

p_diel_with_unassigned

ggsave(
  file.path(output_dir, "diel_plot_WITH_cluster0_UTCminus7_event_dates_only.png"),
  p_diel_with_unassigned,
  width = 9,
  height = 7,
  dpi = 300
)

```

## Plot Diel Detections Excluding Cluster 0 — Local Time UTC-7

```{r}

plot_events_assigned_only <- diel_events %>%
  filter(!is.na(dominant_cluster_assigned_only)) %>%
  left_join(plot_dates, by = "date_utc") %>%
  mutate(
    hour_local = (hour_utc + utc_offset_hours) %% 24,
    dominant_cluster_assigned_only = factor(
      dominant_cluster_assigned_only,
      levels = names(cluster_colors_assigned_only)
    )
  )

p_diel_assigned_only <- ggplot(
  plot_events_assigned_only,
  aes(x = hour_local, y = date_index)
) +
  geom_rect(
    data = plot_sun_times,
    aes(
      xmin = 0,
      xmax = sunrise_hour_local,
      ymin = date_index - 0.45,
      ymax = date_index + 0.45
    ),
    inherit.aes = FALSE,
    fill = "grey85",
    alpha = 0.9
  ) +
  geom_rect(
    data = plot_sun_times,
    aes(
      xmin = sunset_hour_local,
      xmax = 24,
      ymin = date_index - 0.45,
      ymax = date_index + 0.45
    ),
    inherit.aes = FALSE,
    fill = "grey85",
    alpha = 0.9
  ) +
  geom_linerange(
    aes(
      ymin = date_index - 0.28,
      ymax = date_index + 0.28,
      color = dominant_cluster_assigned_only
    ),
    linewidth = 1.0,
    alpha = 1
  ) +
  scale_color_manual(values = cluster_colors_assigned_only, drop = FALSE) +
  scale_x_continuous(
    limits = c(0, 24),
    breaks = seq(0, 24, by = 3)
  ) +
  scale_y_continuous(
    breaks = plot_dates$date_index,
    labels = format(plot_dates$date_utc, "%d-%b-%Y"),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  labs(
    x = "h (UTC-7.0)",
    y = NULL,
    color = "Dominant assigned cluster"
  ) +
  theme_bw(base_size = 13) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "right"
  )

p_diel_assigned_only

ggsave(
  file.path(output_dir, "diel_plot_ASSIGNED_only_UTCminus7_event_dates_only.png"),
  p_diel_assigned_only,
  width = 9,
  height = 7,
  dpi = 300
)

```

```{r}

core_diel_events <- diel_events %>%
  select(
    EventID_clean,
    DriftID_clean,
    DriftName,
    Latitude,
    Longitude,

    dominant_cluster_with_unassigned,
    dominant_cluster_with_unassigned_n,
    dominant_cluster_with_unassigned_percent,

    dominant_cluster_assigned_only,
    dominant_cluster_assigned_only_n,
    dominant_cluster_assigned_only_percent,

    total_clicks_event,

    UTC_start_raw,
    UTC_end_raw,
    UTC_start,
    UTC_end,

    event_duration_min,

    date_utc,
    hour_utc,
    end_hour_utc,

    species,
    call,
    DataQuality,
    gpsUncertainty
  )

write_csv(
  core_diel_events,
  file.path(
    output_dir,
    "core_diel_events_with_dominant_clusters.csv"
  )
)

cat(
  "Saved core diel event table:\n",
  file.path(
    output_dir,
    "core_diel_events_with_dominant_clusters.csv"
  ),
  "\n"
)

```






















