---
title: "Event_Level_Peak_Notch_Tables_18PC_scaledICI1_minPts50_Permissive_AllRegions"
author: "Sarah Moseley"
date: "2026-06-08"
output: html_document
---

## Setup

```{r setup, message=FALSE, warning=FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)

required_packages <- c(
  "dplyr",
  "readr",
  "tibble",
  "tidyr",
  "purrr",
  "pracma",
  "gt",
  "stringr",
  "ggplot2"
)

missing_packages <- required_packages[
  !sapply(required_packages, requireNamespace, quietly = TRUE)
]

if (length(missing_packages) > 0) {
  stop(
    "Missing required packages: ",
    paste(missing_packages, collapse = ", "),
    "\nInstall them before running this script."
  )
}

library(dplyr)
library(readr)
library(tibble)
library(tidyr)
library(purrr)
library(pracma)
library(gt)

```

## Load Saved Inputs

```{r load_saved_inputs, message=FALSE, warning=FALSE}

# ------------------------------------------------------------
# Input files
# ------------------------------------------------------------

hdbscan_assignments_file <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/tables/FULL_hdbscan_assignments_18PC_scaledICI_1_minPts50.csv"

sampled_spectra_matrix_file <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/data/sampled_minmax_spectra_matrix_19_50_khz_18PC.rds"

frequency_grid_file <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/tables/frequency_grid_19_50_khz.csv"

# ------------------------------------------------------------
# Output folders
# ------------------------------------------------------------

event_peak_notch_base_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/Event_Level_Peak_Notch_Tables_Permissive_After_Soldevilla_AllRegions"

event_peak_notch_table_dir <- file.path(
  event_peak_notch_base_dir,
  "tables"
)

event_peak_notch_plot_dir <- file.path(
  event_peak_notch_base_dir,
  "plots"
)

dir.create(
  event_peak_notch_table_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

dir.create(
  event_peak_notch_plot_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

# ------------------------------------------------------------
# Load files
# ------------------------------------------------------------

hdbscan_assignments <- readr::read_csv(
  hdbscan_assignments_file,
  show_col_types = FALSE
)

sampled_spectra_matrix <- readRDS(
  sampled_spectra_matrix_file
)

frequency_grid <- readr::read_csv(
  frequency_grid_file,
  show_col_types = FALSE
)

# ------------------------------------------------------------
# Standardize key columns
# ------------------------------------------------------------

hdbscan_assignments <- hdbscan_assignments %>%
  mutate(
    EventID = as.character(EventID),
    HDBSCAN_cluster = as.character(HDBSCAN_cluster),
    RowID = as.integer(RowID)
  )

frequency_grid <- frequency_grid %>%
  mutate(
    Frequency_kHz = as.numeric(Frequency_kHz)
  )

# ------------------------------------------------------------
# Basic previews
# ------------------------------------------------------------

cat("HDBSCAN assignment columns:\n")
print(names(hdbscan_assignments))

cat("\nRows in HDBSCAN assignments:\n")
print(nrow(hdbscan_assignments))

cat("\nDimensions of sampled spectra matrix:\n")
print(dim(sampled_spectra_matrix))

cat("\nRows in frequency grid:\n")
print(nrow(frequency_grid))

cat("\nNumber of events represented:\n")
print(dplyr::n_distinct(hdbscan_assignments$EventID))

cat("\nClusters represented:\n")
print(sort(unique(hdbscan_assignments$HDBSCAN_cluster)))

cat("\nFrequency range:\n")
print(range(frequency_grid$Frequency_kHz, na.rm = TRUE))

# ------------------------------------------------------------
# Critical alignment checks
# ------------------------------------------------------------

if (nrow(hdbscan_assignments) != nrow(sampled_spectra_matrix)) {
  stop(
    "FAIL: HDBSCAN assignment rows do not match sampled spectra matrix rows."
  )
}

if (ncol(sampled_spectra_matrix) != nrow(frequency_grid)) {
  stop(
    "FAIL: Spectra matrix columns do not match frequency grid rows."
  )
}

if (dplyr::n_distinct(hdbscan_assignments$EventID) != 48) {
  warning(
    "Expected 48 events, but found ",
    dplyr::n_distinct(hdbscan_assignments$EventID),
    ". Check this before continuing."
  )
}

# ------------------------------------------------------------
# Save a quick input summary
# ------------------------------------------------------------

input_summary <- tibble::tibble(
  input_name = c(
    "hdbscan_assignments",
    "sampled_spectra_matrix",
    "frequency_grid",
    "n_events",
    "n_clusters"
  ),
  value = c(
    nrow(hdbscan_assignments),
    paste(dim(sampled_spectra_matrix), collapse = " x "),
    nrow(frequency_grid),
    dplyr::n_distinct(hdbscan_assignments$EventID),
    dplyr::n_distinct(hdbscan_assignments$HDBSCAN_cluster)
  )
)

readr::write_csv(
  input_summary,
  file.path(
    event_peak_notch_table_dir,
    "event_peak_notch_input_summary.csv"
  )
)

cat("\nPASS: Saved inputs loaded and aligned.\n")
cat("Output table folder:\n")
cat(event_peak_notch_table_dir, "\n\n")

cat("Output plot folder:\n")
cat(event_peak_notch_plot_dir, "\n")

```

## Build Event-Level Mean Spectra
## With and without noise cluster 0

```{r build_event_level_mean_spectra, message=FALSE, warning=FALSE}

# ------------------------------------------------------------
# Attach row IDs to preserve alignment between metadata and matrix
# ------------------------------------------------------------

hdbscan_assignments_indexed <- hdbscan_assignments %>%
  mutate(
    MatrixRowID = row_number()
  )

sampled_spectra_indexed <- as.data.frame(sampled_spectra_matrix) %>%
  tibble::as_tibble() %>%
  mutate(
    MatrixRowID = row_number()
  )

# ------------------------------------------------------------
# Name spectrum columns by actual frequency
# ------------------------------------------------------------

frequency_labels <- paste0(
  "freq_",
  sprintf("%.3f", frequency_grid$Frequency_kHz),
  "_kHz"
)

spectra_value_cols <- frequency_labels

names(sampled_spectra_indexed)[
  names(sampled_spectra_indexed) != "MatrixRowID"
] <- frequency_labels

# ------------------------------------------------------------
# Combine metadata + spectra
# One row = one sampled click
# ------------------------------------------------------------

click_spectra_with_metadata <- hdbscan_assignments_indexed %>%
  select(
    MatrixRowID,
    EventID,
    DriftID_spectra,
    UID,
    HDBSCAN_cluster
  ) %>%
  left_join(
    sampled_spectra_indexed,
    by = "MatrixRowID"
  )

# ------------------------------------------------------------
# Safety check
# ------------------------------------------------------------

if (nrow(click_spectra_with_metadata) != nrow(hdbscan_assignments)) {
  stop("FAIL: Joining metadata to spectra changed the row count.")
}

cat("Rows in click-level spectra table:\n")
print(nrow(click_spectra_with_metadata))

cat("\nNumber of events:\n")
print(n_distinct(click_spectra_with_metadata$EventID))

cat("\nNumber of frequency columns:\n")
print(length(spectra_value_cols))

# ------------------------------------------------------------
# Convert click-level spectra to long format
# One row = one click at one frequency
# ------------------------------------------------------------

click_spectra_long <- click_spectra_with_metadata %>%
  pivot_longer(
    cols = all_of(spectra_value_cols),
    names_to = "frequency_label",
    values_to = "minmax_spectrum"
  ) %>%
  mutate(
    Frequency_kHz = as.numeric(
      stringr::str_remove_all(
        stringr::str_remove_all(frequency_label, "freq_"),
        "_kHz"
      )
    )
  )

cat("\nRows in long click spectra table:\n")
print(nrow(click_spectra_long))

cat("\nExpected rows:\n")
print(nrow(hdbscan_assignments) * nrow(frequency_grid))

# ------------------------------------------------------------
# Event mean spectra WITH noise cluster 0 included
# ------------------------------------------------------------

event_mean_spectra_with_noise <- click_spectra_long %>%
  group_by(
    EventID,
    DriftID_spectra,
    Frequency_kHz
  ) %>%
  summarise(
    mean_minmax_spectrum = mean(minmax_spectrum, na.rm = TRUE),
    sd_minmax_spectrum = sd(minmax_spectrum, na.rm = TRUE),
    n_clicks_used = n_distinct(MatrixRowID),
    .groups = "drop"
  ) %>%
  mutate(
    output_group = "WITH_noise_cluster0_included"
  ) %>%
  arrange(
    EventID,
    Frequency_kHz
  )

# ------------------------------------------------------------
# Event mean spectra WITHOUT noise cluster 0
# ------------------------------------------------------------

event_mean_spectra_no_noise <- click_spectra_long %>%
  filter(HDBSCAN_cluster != "0") %>%
  group_by(
    EventID,
    DriftID_spectra,
    Frequency_kHz
  ) %>%
  summarise(
    mean_minmax_spectrum = mean(minmax_spectrum, na.rm = TRUE),
    sd_minmax_spectrum = sd(minmax_spectrum, na.rm = TRUE),
    n_clicks_used = n_distinct(MatrixRowID),
    .groups = "drop"
  ) %>%
  mutate(
    output_group = "NO_noise_cluster0_excluded"
  ) %>%
  arrange(
    EventID,
    Frequency_kHz
  )

# ------------------------------------------------------------
# Combine both output groups
# ------------------------------------------------------------

event_mean_spectra_all_groups <- bind_rows(
  event_mean_spectra_with_noise,
  event_mean_spectra_no_noise
)

# ------------------------------------------------------------
# Sanity checks
# ------------------------------------------------------------

cat("\nEvent counts by output group:\n")

event_count_check <- event_mean_spectra_all_groups %>%
  distinct(output_group, EventID) %>%
  count(output_group, name = "n_events")

print(event_count_check)

cat("\nFrequency-bin counts by output group:\n")

frequency_count_check <- event_mean_spectra_all_groups %>%
  group_by(output_group, EventID) %>%
  summarise(
    n_frequency_bins = n_distinct(Frequency_kHz),
    .groups = "drop"
  ) %>%
  count(output_group, n_frequency_bins)

print(frequency_count_check)

cat("\nPreview: event mean spectra WITH noise:\n")
print(head(event_mean_spectra_with_noise, 10))

cat("\nPreview: event mean spectra NO noise:\n")
print(head(event_mean_spectra_no_noise, 10))

# ------------------------------------------------------------
# Save event mean spectra tables
# ------------------------------------------------------------

readr::write_csv(
  event_mean_spectra_with_noise,
  file.path(
    event_peak_notch_table_dir,
    "event_mean_spectra_WITH_noise_cluster0_included.csv"
  )
)

readr::write_csv(
  event_mean_spectra_no_noise,
  file.path(
    event_peak_notch_table_dir,
    "event_mean_spectra_NO_noise_cluster0_excluded.csv"
  )
)

readr::write_csv(
  event_mean_spectra_all_groups,
  file.path(
    event_peak_notch_table_dir,
    "event_mean_spectra_ALL_groups.csv"
  )
)

cat("\nPASS: Event-level mean spectra created and saved.\n")
cat("Saved to:\n")
cat(event_peak_notch_table_dir, "\n")

```

## Pick Peaks and Notches From Event-Level Mean Spectra
## Permissive settings after Soldevilla comparison

```{r pick_event_level_peaks_and_notches, message=FALSE, warning=FALSE}

# ------------------------------------------------------------
# Helper function: pick peaks/notches for one event spectrum
# ------------------------------------------------------------

pick_features_one_event <- function(one_event_spectrum) {
  
  one_event_spectrum <- one_event_spectrum %>%
    arrange(Frequency_kHz)
  
  event_id <- unique(one_event_spectrum$EventID)
  drift_id <- unique(one_event_spectrum$DriftID_spectra)
  output_group_id <- unique(one_event_spectrum$output_group)
  
  if (length(event_id) != 1 || length(output_group_id) != 1) {
    stop("Input to pick_features_one_event should contain exactly one event and one output group.")
  }
  
  # --------------------------
  # Peaks
  # --------------------------
  
  peak_matrix <- pracma::findpeaks(
    one_event_spectrum$mean_minmax_spectrum,
    nups = 1,
    ndowns = 1,
    minpeakdistance = 1,
    minpeakheight = 0.15
  )
  
  if (is.null(peak_matrix)) {
    
    peaks_df <- tibble()
    
  } else {
    
    peaks_df <- tibble(
      output_group = output_group_id,
      EventID = event_id,
      DriftID_spectra = drift_id,
      feature_type = "Peak",
      feature_height = peak_matrix[, 1],
      feature_index = peak_matrix[, 2],
      feature_start_index = peak_matrix[, 3],
      feature_end_index = peak_matrix[, 4],
      Frequency_kHz = one_event_spectrum$Frequency_kHz[feature_index]
    )
  }
  
  # --------------------------
  # Notches
  # Notches are peaks in the inverted spectrum.
  # --------------------------
  
  notch_matrix <- pracma::findpeaks(
    -one_event_spectrum$mean_minmax_spectrum,
    nups = 1,
    ndowns = 1,
    minpeakdistance = 1,
    minpeakheight = -0.85
  )
  
  if (is.null(notch_matrix)) {
    
    notches_df <- tibble()
    
  } else {
    
    notches_df <- tibble(
      output_group = output_group_id,
      EventID = event_id,
      DriftID_spectra = drift_id,
      feature_type = "Notch",
      feature_height = one_event_spectrum$mean_minmax_spectrum[notch_matrix[, 2]],
      feature_index = notch_matrix[, 2],
      feature_start_index = notch_matrix[, 3],
      feature_end_index = notch_matrix[, 4],
      Frequency_kHz = one_event_spectrum$Frequency_kHz[feature_index]
    )
  }
  
  bind_rows(
    peaks_df,
    notches_df
  ) %>%
    arrange(
      feature_type,
      Frequency_kHz
    )
}

# ------------------------------------------------------------
# Run feature picking for every event in both output groups
# ------------------------------------------------------------

event_peak_notch_features_raw <- event_mean_spectra_all_groups %>%
  group_by(output_group, EventID) %>%
  group_split() %>%
  purrr::map_dfr(pick_features_one_event)

# ------------------------------------------------------------
# Rank features within each event from low to high frequency
# Keep first 4 peaks and first 3 notches
# ------------------------------------------------------------

event_peak_notch_features_ranked <- event_peak_notch_features_raw %>%
  group_by(
    output_group,
    EventID,
    feature_type
  ) %>%
  arrange(
    Frequency_kHz,
    .by_group = TRUE
  ) %>%
  mutate(
    feature_rank = row_number(),
    feature_label = paste(feature_type, feature_rank)
  ) %>%
  ungroup() %>%
  filter(
    (feature_type == "Peak" & feature_rank <= 4) |
      (feature_type == "Notch" & feature_rank <= 3)
  ) %>%
  arrange(
    output_group,
    EventID,
    feature_type,
    feature_rank
  )

# ------------------------------------------------------------
# Sanity checks
# ------------------------------------------------------------

cat("Raw detected features:\n")
print(
  event_peak_notch_features_raw %>%
    count(output_group, feature_type, name = "n_features_detected")
)

cat("\nRanked retained features:\n")
print(
  event_peak_notch_features_ranked %>%
    count(output_group, feature_type, feature_label, name = "n_events_with_feature") %>%
    arrange(output_group, feature_type, feature_label)
)

cat("\nEvents missing one or more expected features:\n")

expected_features <- tidyr::expand_grid(
  output_group = unique(event_mean_spectra_all_groups$output_group),
  EventID = unique(event_mean_spectra_all_groups$EventID),
  feature_label = c(
    "Peak 1", "Peak 2", "Peak 3", "Peak 4",
    "Notch 1", "Notch 2", "Notch 3"
  )
)

missing_feature_check <- expected_features %>%
  left_join(
    event_peak_notch_features_ranked %>%
      distinct(output_group, EventID, feature_label) %>%
      mutate(feature_found = TRUE),
    by = c("output_group", "EventID", "feature_label")
  ) %>%
  mutate(
    feature_found = if_else(is.na(feature_found), FALSE, feature_found)
  ) %>%
  filter(!feature_found)

print(missing_feature_check, n = 100)

cat("\nPreview of ranked event-level peak/notch features:\n")
print(event_peak_notch_features_ranked, n = 50)

# ------------------------------------------------------------
# Save feature tables
# ------------------------------------------------------------

readr::write_csv(
  event_peak_notch_features_raw,
  file.path(
    event_peak_notch_table_dir,
    "event_peak_notch_features_RAW_all_detected.csv"
  )
)

readr::write_csv(
  event_peak_notch_features_ranked,
  file.path(
    event_peak_notch_table_dir,
    "event_peak_notch_features_RANKED_first4peaks_first3notches.csv"
  )
)

readr::write_csv(
  missing_feature_check,
  file.path(
    event_peak_notch_table_dir,
    "event_peak_notch_missing_feature_check.csv"
  )
)

cat("\nPASS: Event-level peaks and notches picked, ranked, and saved.\n")
cat("Saved to:\n")
cat(event_peak_notch_table_dir, "\n")

```

## Build Event-Level Peak/Notch Summary Statistics
## Mean ± SD across event-level mean spectra

```{r build_event_peak_notch_summary_statistics, message=FALSE, warning=FALSE}

# ------------------------------------------------------------
# Soldevilla reference values from Table IV
# ------------------------------------------------------------

soldevilla_reference <- tibble::tibble(
  Region = rep(c("PPac", "CC"), each = 7),
  
  feature_label = rep(
    c(
      "Peak 1", "Peak 2", "Peak 3", "Peak 4",
      "Notch 1", "Notch 2", "Notch 3"
    ),
    times = 2
  ),
  
  mean_frequency_khz = c(
    # PPac
    23.5, 26.6, 31.7, 36.0,
    25.4, 28.7, 33.5,
    
    # CC
    21.8, 24.8, 31.2, 37.9,
    23.4, 27.0, 34.5
  ),
  
  sd_frequency_khz = c(
    # PPac
    0.9, 1.5, 0.6, 2.1,
    0.9, 1.2, 0.8,
    
    # CC
    1.0, 1.0, 1.0, 2.3,
    1.2, 0.9, 1.3
  )
) %>%
  mutate(
    summary_value = sprintf(
      "%.1f ± %.1f",
      mean_frequency_khz,
      sd_frequency_khz
    )
  )

# ------------------------------------------------------------
# Our event-level summary statistics
# Mean ± SD across events, not clusters
# ------------------------------------------------------------

event_peak_notch_summary <- event_peak_notch_features_ranked %>%
  group_by(
    output_group,
    feature_type,
    feature_rank,
    feature_label
  ) %>%
  summarise(
    mean_frequency_khz = mean(Frequency_kHz, na.rm = TRUE),
    sd_frequency_khz = sd(Frequency_kHz, na.rm = TRUE),
    n_events_with_feature = n_distinct(EventID),
    .groups = "drop"
  ) %>%
  mutate(
    Region = case_when(
      output_group == "WITH_noise_cluster0_included" ~ "Our Data: events with noise",
      output_group == "NO_noise_cluster0_excluded" ~ "Our Data: events no noise",
      TRUE ~ output_group
    ),
    
    summary_value = sprintf(
      "%.1f ± %.1f",
      mean_frequency_khz,
      sd_frequency_khz
    ),
    
    summary_value_with_n = sprintf(
      "%.1f ± %.1f (n=%s)",
      mean_frequency_khz,
      sd_frequency_khz,
      n_events_with_feature
    )
  ) %>%
  arrange(
    output_group,
    feature_type,
    feature_rank
  )

# ------------------------------------------------------------
# Sanity checks
# ------------------------------------------------------------

cat("Soldevilla reference values:\n")
print(soldevilla_reference)

cat("\nEvent-level peak/notch summary:\n")
print(event_peak_notch_summary, n = Inf)

cat("\nFeature sample sizes:\n")
print(
  event_peak_notch_summary %>%
    select(
      output_group,
      feature_label,
      n_events_with_feature
    ) %>%
    arrange(output_group, feature_label),
  n = Inf
)

# ------------------------------------------------------------
# Save summary statistics
# ------------------------------------------------------------

readr::write_csv(
  soldevilla_reference,
  file.path(
    event_peak_notch_table_dir,
    "soldevilla_reference_values.csv"
  )
)

readr::write_csv(
  event_peak_notch_summary,
  file.path(
    event_peak_notch_table_dir,
    "event_level_peak_notch_summary_statistics.csv"
  )
)

cat("\nPASS: Event-level summary statistics created and saved.\n")
cat("Saved to:\n")
cat(event_peak_notch_table_dir, "\n")

```

## Expanded Soldevilla Reference Table
## Includes SEUS, NEUS, GoMx, IPH, PPac, and CC

```{r expanded_soldevilla_reference_table, message=FALSE, warning=FALSE}

# ------------------------------------------------------------
# Expanded Soldevilla reference values from Table IV
# Regions:
# SEUS = Southeastern Shelf
# NEUS = Northeastern Shelf
# GoMx = Gulf of Mexico
# IPH  = Insular Pacific-Hawaiian
# PPac = Pelagic Pacific
# CC   = California Current
# ------------------------------------------------------------

soldevilla_reference_expanded <- tibble::tibble(
  
  Region = rep(
    c("SEUS", "NEUS", "GoMx", "IPH", "PPac", "CC"),
    each = 7
  ),
  
  N = rep(
    c(15, 8, 7, 8, 6, 25),
    each = 7
  ),
  
  feature_label = rep(
    c(
      "Peak 1", "Peak 2", "Peak 3", "Peak 4",
      "Notch 1", "Notch 2", "Notch 3"
    ),
    times = 6
  ),
  
  mean_frequency_khz = c(
    
    # SEUS
    23.1, 26.8, 33.3, 39.3,
    24.8, 28.8, 36.2,
    
    # NEUS
    22.1, 26.1, 31.5, 37.2,
    23.9, 28.1, 33.8,
    
    # GoMx
    22.4, 25.6, 30.4, 36.5,
    23.8, 27.2, 32.5,
    
    # IPH
    24.5, 27.3, 33.5, 39.1,
    25.9, 29.4, 36.2,
    
    # PPac
    23.5, 26.6, 31.7, 36.0,
    25.4, 28.7, 33.5,
    
    # CC
    21.8, 24.8, 31.2, 37.9,
    23.4, 27.0, 34.5
  ),
  
  sd_frequency_khz = c(
    
    # SEUS
    1.0, 0.7, 0.7, 3.0,
    0.9, 0.8, 1.1,
    
    # NEUS
    1.0, 1.0, 2.1, 4.5,
    0.8, 1.2, 2.7,
    
    # GoMx
    1.0, 1.1, 2.8, 4.3,
    0.5, 0.8, 2.9,
    
    # IPH
    0.5, 0.6, 0.9, 3.4,
    0.8, 0.3, 1.8,
    
    # PPac
    0.9, 1.5, 0.6, 2.1,
    0.9, 1.2, 0.8,
    
    # CC
    1.0, 1.0, 1.0, 2.3,
    1.2, 0.9, 1.3
  )
) %>%
  mutate(
    summary_value = sprintf(
      "%.1f ± %.1f",
      mean_frequency_khz,
      sd_frequency_khz
    ),
    
    Region_with_N = paste0(
      Region,
      "\nN = ",
      N
    )
  )

# ------------------------------------------------------------
# Sanity check
# ------------------------------------------------------------

cat("Expanded Soldevilla reference table:\n")
print(soldevilla_reference_expanded, n = Inf)

cat("\nRegion counts:\n")
print(
  soldevilla_reference_expanded %>%
    count(Region, N)
)

cat("\nFeature counts:\n")
print(
  soldevilla_reference_expanded %>%
    count(feature_label)
)

# ------------------------------------------------------------
# Save expanded reference table
# ------------------------------------------------------------

readr::write_csv(
  soldevilla_reference_expanded,
  file.path(
    event_peak_notch_table_dir,
    "soldevilla_reference_values_EXPANDED_all_regions.csv"
  )
)

cat("\nPASS: Expanded Soldevilla reference table created and saved.\n")

```

## Build Expanded Regional Comparison Tables
## All regions only
## Duplicated for NO noise and WITH noise event-level summaries

```{r build_expanded_regional_comparison_tables_all_regions_only, message=FALSE, warning=FALSE}

# ------------------------------------------------------------
# Region set: all Soldevilla regions only
# ------------------------------------------------------------

all_soldevilla_regions <- c(
  "SEUS",
  "NEUS",
  "GoMx",
  "IPH",
  "PPac",
  "CC"
)

# ------------------------------------------------------------
# Helper: build one comparison table
# ------------------------------------------------------------

build_expanded_comparison_table <- function(
    event_summary,
    soldevilla_expanded,
    output_group_to_use,
    feature_type_to_use,
    soldevilla_regions_to_use
) {
  
  if (feature_type_to_use == "Peak") {
    feature_levels <- c("Peak 1", "Peak 2", "Peak 3", "Peak 4")
  } else if (feature_type_to_use == "Notch") {
    feature_levels <- c("Notch 1", "Notch 2", "Notch 3")
  } else {
    stop("feature_type_to_use must be either 'Peak' or 'Notch'.")
  }
  
  region_levels <- c(
    soldevilla_regions_to_use,
    "Our Data"
  )
  
  # ------------------------------------------------------------
  # Soldevilla values
  # ------------------------------------------------------------
  
  soldevilla_part <- soldevilla_expanded %>%
    filter(
      Region %in% soldevilla_regions_to_use,
      feature_label %in% feature_levels
    ) %>%
    mutate(
      Region = factor(
        Region,
        levels = soldevilla_regions_to_use
      )
    ) %>%
    select(
      feature_label,
      Region,
      summary_value
    )
  
  # ------------------------------------------------------------
  # Our event-level values
  # ------------------------------------------------------------
  
  our_data_part <- event_summary %>%
    filter(
      output_group == output_group_to_use,
      feature_type == feature_type_to_use
    ) %>%
    mutate(
      Region = "Our Data",
      summary_value = sprintf(
        "%.1f ± %.1f",
        mean_frequency_khz,
        sd_frequency_khz
      )
    ) %>%
    select(
      feature_label,
      Region,
      summary_value
    )
  
  # ------------------------------------------------------------
  # Combine and pivot
  # ------------------------------------------------------------
  
  comparison_table <- bind_rows(
    soldevilla_part,
    our_data_part
  ) %>%
    mutate(
      feature_label = factor(
        feature_label,
        levels = feature_levels
      ),
      Region = factor(
        Region,
        levels = region_levels
      )
    ) %>%
    arrange(
      feature_label,
      Region
    ) %>%
    tidyr::pivot_wider(
      names_from = Region,
      values_from = summary_value
    ) %>%
    rename(
      Feature = feature_label
    )
  
  return(comparison_table)
}

# ------------------------------------------------------------
# NO noise tables: all regions only
# ------------------------------------------------------------

peak_all_regions_no_noise <- build_expanded_comparison_table(
  event_summary = event_peak_notch_summary,
  soldevilla_expanded = soldevilla_reference_expanded,
  output_group_to_use = "NO_noise_cluster0_excluded",
  feature_type_to_use = "Peak",
  soldevilla_regions_to_use = all_soldevilla_regions
)

notch_all_regions_no_noise <- build_expanded_comparison_table(
  event_summary = event_peak_notch_summary,
  soldevilla_expanded = soldevilla_reference_expanded,
  output_group_to_use = "NO_noise_cluster0_excluded",
  feature_type_to_use = "Notch",
  soldevilla_regions_to_use = all_soldevilla_regions
)

# ------------------------------------------------------------
# WITH noise tables: all regions only
# ------------------------------------------------------------

peak_all_regions_with_noise <- build_expanded_comparison_table(
  event_summary = event_peak_notch_summary,
  soldevilla_expanded = soldevilla_reference_expanded,
  output_group_to_use = "WITH_noise_cluster0_included",
  feature_type_to_use = "Peak",
  soldevilla_regions_to_use = all_soldevilla_regions
)

notch_all_regions_with_noise <- build_expanded_comparison_table(
  event_summary = event_peak_notch_summary,
  soldevilla_expanded = soldevilla_reference_expanded,
  output_group_to_use = "WITH_noise_cluster0_included",
  feature_type_to_use = "Notch",
  soldevilla_regions_to_use = all_soldevilla_regions
)

# ------------------------------------------------------------
# Preview tables
# ------------------------------------------------------------

cat("Peak table: all Soldevilla regions, NO noise\n")
print(peak_all_regions_no_noise)

cat("\nNotch table: all Soldevilla regions, NO noise\n")
print(notch_all_regions_no_noise)

cat("\nPeak table: all Soldevilla regions, WITH noise\n")
print(peak_all_regions_with_noise)

cat("\nNotch table: all Soldevilla regions, WITH noise\n")
print(notch_all_regions_with_noise)

# ------------------------------------------------------------
# Save CSV outputs: all regions only
# ------------------------------------------------------------

readr::write_csv(
  peak_all_regions_no_noise,
  file.path(
    event_peak_notch_table_dir,
    "expanded_peak_table_ALL_regions_NO_noise_PERMISSIVE.csv"
  )
)

readr::write_csv(
  notch_all_regions_no_noise,
  file.path(
    event_peak_notch_table_dir,
    "expanded_notch_table_ALL_regions_NO_noise_PERMISSIVE.csv"
  )
)

readr::write_csv(
  peak_all_regions_with_noise,
  file.path(
    event_peak_notch_table_dir,
    "expanded_peak_table_ALL_regions_WITH_noise_PERMISSIVE.csv"
  )
)

readr::write_csv(
  notch_all_regions_with_noise,
  file.path(
    event_peak_notch_table_dir,
    "expanded_notch_table_ALL_regions_WITH_noise_PERMISSIVE.csv"
  )
)

# ------------------------------------------------------------
# Save feature-level sample sizes again for reference
# ------------------------------------------------------------

expanded_feature_n_check <- event_peak_notch_summary %>%
  select(
    output_group,
    feature_type,
    feature_label,
    n_events_with_feature
  ) %>%
  arrange(
    output_group,
    feature_type,
    feature_label
  )

readr::write_csv(
  expanded_feature_n_check,
  file.path(
    event_peak_notch_table_dir,
    "expanded_tables_our_data_feature_sample_sizes_PERMISSIVE.csv"
  )
)

cat("\nPASS: Expanded all-region comparison tables created and saved.\n")
cat("Saved to:\n")
cat(event_peak_notch_table_dir, "\n")

```

## Style and Export Expanded Regional Comparison Tables
## All regions only
## Compact Soldevilla-style spacing
## Creates PNG + HTML outputs for the 4 all-region tables

```{r style_and_export_expanded_regional_tables_all_regions_only, message=FALSE, warning=FALSE}

# ------------------------------------------------------------
# Output folder for expanded comparison tables
# ------------------------------------------------------------

expanded_table_output_dir <- file.path(
  event_peak_notch_base_dir,
  "expanded_regional_comparison_tables_ALL_regions_only_PERMISSIVE"
)

dir.create(
  expanded_table_output_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

# ------------------------------------------------------------
# Helper: build dynamic column labels with N values
# ------------------------------------------------------------

make_region_column_labels <- function(input_table) {
  
  region_n_lookup <- c(
    "SEUS" = "N = 15",
    "NEUS" = "N = 8",
    "GoMx" = "N = 7",
    "IPH" = "N = 8",
    "PPac" = "N = 6",
    "CC" = "N = 25",
    "Our Data" = "N = 48"
  )
  
  label_list <- list()
  
  for (col_name in names(input_table)) {
    
    if (col_name == "Feature") {
      
      label_list[[col_name]] <- gt::md("Feature")
      
    } else if (col_name %in% names(region_n_lookup)) {
      
      label_list[[col_name]] <- gt::md(
        paste0(col_name, "<br>", region_n_lookup[[col_name]])
      )
      
    } else {
      
      label_list[[col_name]] <- gt::md(col_name)
    }
  }
  
  return(label_list)
}

# ------------------------------------------------------------
# Helper: make one compact styled gt table
# ------------------------------------------------------------

make_expanded_gt_table <- function(
    input_table,
    title_text,
    include_noise_footnote = FALSE
) {
  
  column_labels <- make_region_column_labels(input_table)
  
  gt_table <- input_table %>%
    gt::gt() %>%
    gt::tab_header(
      title = gt::md(title_text)
    ) %>%
    gt::cols_label(
      .list = column_labels
    ) %>%
    gt::cols_align(
      align = "center",
      columns = everything()
    ) %>%
    gt::tab_style(
      style = gt::cell_text(weight = "bold"),
      locations = gt::cells_column_labels(everything())
    ) %>%
    gt::tab_options(
      table.font.names = "Times New Roman",
      table.font.size = gt::px(13),
      heading.title.font.size = gt::px(13),
      heading.title.font.weight = "normal",
      column_labels.font.size = gt::px(13),
      data_row.padding = gt::px(4),
      column_labels.padding = gt::px(4),
      table.width = gt::px(760),
      table.border.top.width = gt::px(1),
      table.border.bottom.width = gt::px(1),
      column_labels.border.top.width = gt::px(1),
      column_labels.border.bottom.width = gt::px(1),
      table_body.hlines.width = gt::px(0.5)
    ) %>%
    gt::opt_css(
      css = "
      table.gt_table td, 
      table.gt_table th {
        padding-left: 4px !important;
        padding-right: 4px !important;
        white-space: nowrap;
      }
      "
    )
  
  if (include_noise_footnote) {
    
    gt_table <- gt_table %>%
      gt::tab_footnote(
        footnote = gt::md(
          "Feature-level n varies when a peak or notch was not detected in every event; see `expanded_tables_our_data_feature_sample_sizes_PERMISSIVE.csv`."
        ),
        locations = gt::cells_column_labels(
          columns = `Our Data`
        )
      )
  }
  
  return(gt_table)
}

# ------------------------------------------------------------
# Build gt objects: NO noise, all regions only
# ------------------------------------------------------------

gt_peak_all_regions_no_noise <- make_expanded_gt_table(
  input_table = peak_all_regions_no_noise,
  title_text = "Summary statistics (mean ± SD) for spectral peak frequencies (kHz).",
  include_noise_footnote = FALSE
)

gt_notch_all_regions_no_noise <- make_expanded_gt_table(
  input_table = notch_all_regions_no_noise,
  title_text = "Summary statistics (mean ± SD) for spectral notch frequencies (kHz).",
  include_noise_footnote = FALSE
)

# ------------------------------------------------------------
# Build gt objects: WITH noise, all regions only
# ------------------------------------------------------------

gt_peak_all_regions_with_noise <- make_expanded_gt_table(
  input_table = peak_all_regions_with_noise,
  title_text = "Summary statistics (mean ± SD) for spectral peak frequencies (kHz).",
  include_noise_footnote = TRUE
)

gt_notch_all_regions_with_noise <- make_expanded_gt_table(
  input_table = notch_all_regions_with_noise,
  title_text = "Summary statistics (mean ± SD) for spectral notch frequencies (kHz).",
  include_noise_footnote = TRUE
)

# ------------------------------------------------------------
# Display in knitted output
# ------------------------------------------------------------

gt_peak_all_regions_no_noise
gt_notch_all_regions_no_noise

gt_peak_all_regions_with_noise
gt_notch_all_regions_with_noise

# ------------------------------------------------------------
# Helper: save gt table as HTML and PNG
# ------------------------------------------------------------

save_gt_html_and_png <- function(gt_object, base_filename) {
  
  html_file <- file.path(
    expanded_table_output_dir,
    paste0(base_filename, ".html")
  )
  
  png_file <- file.path(
    expanded_table_output_dir,
    paste0(base_filename, ".png")
  )
  
  gt::gtsave(
    gt_object,
    html_file
  )
  
  tryCatch(
    {
      gt::gtsave(
        gt_object,
        png_file,
        vwidth = 900,
        vheight = 500,
        zoom = 2
      )
      
      cat("Saved HTML and PNG:", base_filename, "\n")
    },
    error = function(e) {
      cat("Saved HTML only:", base_filename, "\n")
      cat("PNG export failed with message:\n")
      cat(e$message, "\n\n")
    }
  )
}

# ------------------------------------------------------------
# Save all 4 outputs
# ------------------------------------------------------------

save_gt_html_and_png(
  gt_peak_all_regions_no_noise,
  "expanded_peak_ALL_regions_NO_noise_PERMISSIVE"
)

save_gt_html_and_png(
  gt_notch_all_regions_no_noise,
  "expanded_notch_ALL_regions_NO_noise_PERMISSIVE"
)

save_gt_html_and_png(
  gt_peak_all_regions_with_noise,
  "expanded_peak_ALL_regions_WITH_noise_PERMISSIVE"
)

save_gt_html_and_png(
  gt_notch_all_regions_with_noise,
  "expanded_notch_ALL_regions_WITH_noise_PERMISSIVE"
)

cat("\nPASS: Expanded all-region gt tables exported with compact spacing.\n")
cat("Saved to:\n")
cat(expanded_table_output_dir, "\n")

```

## Regional Peak/Notch Frequency Plots
## Region on x-axis, frequency on y-axis
## Mean point ± SD error bars
## Separate folders for WITH noise and NO noise

```{r regional_peak_notch_frequency_plots, message=FALSE, warning=FALSE}

library(dplyr)
library(ggplot2)
library(stringr)
library(readr)

# ------------------------------------------------------------
# Safety checks
# ------------------------------------------------------------

if (!exists("event_peak_notch_summary")) {
  stop("event_peak_notch_summary does not exist. Run the event summary-statistics chunk first.")
}

if (!exists("soldevilla_reference_expanded")) {
  stop("soldevilla_reference_expanded does not exist. Run the expanded Soldevilla reference chunk first.")
}

if (!exists("event_peak_notch_base_dir")) {
  if (exists("event_peak_notch_table_dir")) {
    event_peak_notch_base_dir <- dirname(event_peak_notch_table_dir)
  } else {
    stop("event_peak_notch_base_dir does not exist.")
  }
}

# ------------------------------------------------------------
# Output folders
# ------------------------------------------------------------

regional_plot_base_dir <- file.path(
  event_peak_notch_base_dir,
  "Regional_Peak_Notch_Plots_ALL_regions_only_PERMISSIVE"
)

regional_plot_no_noise_dir <- file.path(
  regional_plot_base_dir,
  "No Noise"
)

regional_plot_with_noise_dir <- file.path(
  regional_plot_base_dir,
  "With Noise"
)

dir.create(regional_plot_no_noise_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(regional_plot_with_noise_dir, recursive = TRUE, showWarnings = FALSE)

cat("No-noise plots will save to:\n")
cat(regional_plot_no_noise_dir, "\n\n")

cat("With-noise plots will save to:\n")
cat(regional_plot_with_noise_dir, "\n\n")

# ------------------------------------------------------------
# Feature colors
# ------------------------------------------------------------

peak_colors <- c(
  "Peak 1" = "#377EB8",  # blue
  "Peak 2" = "#FF7F00",  # orange
  "Peak 3" = "#4DAF4A",  # green
  "Peak 4" = "#984EA3"   # purple
)

notch_colors <- c(
  "Notch 1" = "#377EB8", # blue
  "Notch 2" = "#FF7F00", # orange
  "Notch 3" = "#4DAF4A"  # green
)

# ------------------------------------------------------------
# Region comparison sets
# ------------------------------------------------------------

comparison_sets <- list(
  
  all_regions = list(
    label = "All Soldevilla Regions + Our Data",
    regions = c("SEUS", "NEUS", "GoMx", "IPH", "PPac", "CC", "Our Data"),
    width = 10.5
  )
)

# ------------------------------------------------------------
# Helper: build plot-ready data for one output group
# ------------------------------------------------------------

build_plot_ready_data <- function(output_group_to_use) {
  
  # Soldevilla values
  soldevilla_plot_data <- soldevilla_reference_expanded %>%
    mutate(
      output_group = output_group_to_use,
      source = "Soldevilla",
      feature_type = if_else(
        str_detect(feature_label, "Peak"),
        "Peak",
        "Notch"
      ),
      n_events_with_feature = N,
      x_label = paste0(Region, "\nN = ", N)
    ) %>%
    select(
      output_group,
      source,
      Region,
      x_label,
      feature_type,
      feature_label,
      mean_frequency_khz,
      sd_frequency_khz,
      n_events_with_feature
    )
  
  # Our event-level values
  our_data_plot_data <- event_peak_notch_summary %>%
    filter(
      output_group == output_group_to_use
    ) %>%
    mutate(
      source = "Our Data",
      Region = "Our Data",
      x_label = if_else(
        output_group_to_use == "WITH_noise_cluster0_included",
        "Our Data\nN = 48 events*",
        "Our Data\nN = 48 events"
      )
    ) %>%
    select(
      output_group,
      source,
      Region,
      x_label,
      feature_type,
      feature_label,
      mean_frequency_khz,
      sd_frequency_khz,
      n_events_with_feature
    )
  
  bind_rows(
    soldevilla_plot_data,
    our_data_plot_data
  )
}

# ------------------------------------------------------------
# Helper: make one plot
# ------------------------------------------------------------

make_regional_frequency_plot <- function(
    plot_data,
    regions_to_use,
    feature_type_to_use,
    comparison_label,
    noise_label,
    output_file,
    plot_width
) {
  
  if (feature_type_to_use == "Peak") {
    
    feature_levels <- c("Peak 1", "Peak 2", "Peak 3", "Peak 4")
    feature_colors <- peak_colors
    y_limits <- c(19, 43)
    title_text <- "Spectral Peak Frequencies by Region"
    
  } else if (feature_type_to_use == "Notch") {
    
    feature_levels <- c("Notch 1", "Notch 2", "Notch 3")
    feature_colors <- notch_colors
    y_limits <- c(21, 39)
    title_text <- "Spectral Notch Frequencies by Region"
    
  } else {
    
    stop("feature_type_to_use must be either 'Peak' or 'Notch'.")
  }
  
  plot_data_filtered <- plot_data %>%
    filter(
      Region %in% regions_to_use,
      feature_type == feature_type_to_use
    ) %>%
    mutate(
      Region = factor(
        Region,
        levels = regions_to_use
      ),
      feature_label = factor(
        feature_label,
        levels = feature_levels
      )
    )
  
  # Preserve ordered x-axis labels with N values
  x_label_levels <- plot_data_filtered %>%
    distinct(Region, x_label) %>%
    arrange(Region) %>%
    pull(x_label)
  
  plot_data_filtered <- plot_data_filtered %>%
    mutate(
      x_label = factor(
        x_label,
        levels = x_label_levels
      )
    )
  
  subtitle_text <- paste0(
    comparison_label,
    "; ",
    noise_label
  )
  
  if (noise_label == "with noise cluster 0 included") {
    subtitle_text <- paste0(
      subtitle_text,
      "\n*Our Data uses a 48-event pool; feature-level n varies when a peak/notch was not detected."
    )
  }
  
  p <- ggplot(
    plot_data_filtered,
    aes(
      x = x_label,
      y = mean_frequency_khz,
      color = feature_label
    )
  ) +
    
    # Error bars and points are NOT dodged.
    # This keeps all features exactly aligned with their region.
    geom_errorbar(
      aes(
        ymin = mean_frequency_khz - sd_frequency_khz,
        ymax = mean_frequency_khz + sd_frequency_khz
      ),
      width = 0.12,
      linewidth = 0.75,
      alpha = 0.9
    ) +
    geom_point(
      size = 3.2,
      alpha = 0.95
    ) +
    scale_color_manual(
      values = feature_colors,
      breaks = feature_levels,
      drop = FALSE
    ) +
    coord_cartesian(
      ylim = y_limits
    ) +
    labs(
      title = title_text,
      subtitle = subtitle_text,
      x = "Region",
      y = "Frequency (kHz)",
      color = "Feature"
    ) +
    theme_bw(base_size = 13) +
    theme(
      plot.title = element_text(face = "bold"),
      plot.subtitle = element_text(size = 10),
      axis.text.x = element_text(size = 11),
      axis.title = element_text(face = "bold"),
      legend.position = "right",
      panel.grid.minor = element_blank()
    )
  
  print(p)
  
  ggsave(
    filename = output_file,
    plot = p,
    width = plot_width,
    height = 5.5,
    dpi = 300
  )
  
  return(p)
}

# ------------------------------------------------------------
# Build plot-ready datasets
# ------------------------------------------------------------

plot_data_no_noise <- build_plot_ready_data(
  output_group_to_use = "NO_noise_cluster0_excluded"
)

plot_data_with_noise <- build_plot_ready_data(
  output_group_to_use = "WITH_noise_cluster0_included"
)

# ------------------------------------------------------------
# Save plot-ready data
# ------------------------------------------------------------

readr::write_csv(
  plot_data_no_noise,
  file.path(
    regional_plot_no_noise_dir,
    "plot_ready_regional_peak_notch_data_NO_noise_PERMISSIVE.csv"
  )
)

readr::write_csv(
  plot_data_with_noise,
  file.path(
    regional_plot_with_noise_dir,
    "plot_ready_regional_peak_notch_data_WITH_noise_PERMISSIVE.csv"
  )
)

# ------------------------------------------------------------
# Generate plots: NO noise
# ------------------------------------------------------------

for (comparison_name in names(comparison_sets)) {
  
  comparison_info <- comparison_sets[[comparison_name]]
  
  make_regional_frequency_plot(
    plot_data = plot_data_no_noise,
    regions_to_use = comparison_info$regions,
    feature_type_to_use = "Peak",
    comparison_label = comparison_info$label,
    noise_label = "noise cluster 0 excluded",
    output_file = file.path(
      regional_plot_no_noise_dir,
      paste0("regional_peak_frequencies_", comparison_name, "_NO_noise_PERMISSIVE.png")
    ),
    plot_width = comparison_info$width
  )
  
  make_regional_frequency_plot(
    plot_data = plot_data_no_noise,
    regions_to_use = comparison_info$regions,
    feature_type_to_use = "Notch",
    comparison_label = comparison_info$label,
    noise_label = "noise cluster 0 excluded",
    output_file = file.path(
      regional_plot_no_noise_dir,
      paste0("regional_notch_frequencies_", comparison_name, "_NO_noise_PERMISSIVE.png")
    ),
    plot_width = comparison_info$width
  )
}

# ------------------------------------------------------------
# Generate plots: WITH noise
# ------------------------------------------------------------

for (comparison_name in names(comparison_sets)) {
  
  comparison_info <- comparison_sets[[comparison_name]]
  
  make_regional_frequency_plot(
    plot_data = plot_data_with_noise,
    regions_to_use = comparison_info$regions,
    feature_type_to_use = "Peak",
    comparison_label = comparison_info$label,
    noise_label = "with noise cluster 0 included",
    output_file = file.path(
      regional_plot_with_noise_dir,
      paste0("regional_peak_frequencies_", comparison_name, "_WITH_noise_PERMISSIVE.png")
    ),
    plot_width = comparison_info$width
  )
  
  make_regional_frequency_plot(
    plot_data = plot_data_with_noise,
    regions_to_use = comparison_info$regions,
    feature_type_to_use = "Notch",
    comparison_label = comparison_info$label,
    noise_label = "with noise cluster 0 included",
    output_file = file.path(
      regional_plot_with_noise_dir,
      paste0("regional_notch_frequencies_", comparison_name, "_WITH_noise_PERMISSIVE.png")
    ),
    plot_width = comparison_info$width
  )
}

cat("\nPASS: Regional peak/notch frequency plots created.\n")
cat("No-noise plots saved to:\n")
cat(regional_plot_no_noise_dir, "\n\n")

cat("With-noise plots saved to:\n")
cat(regional_plot_with_noise_dir, "\n")

```







































