---
title: "Generate_CleanAcousticStudies_TukeySpectra"
author: "Anne Simonis"
date: "2026-03-04"
output: html_document
---

For a new event from the same drift run the following after generating a new clean acoustic study and running the script: 
evNum <- 2
source("Generate_TukeySpectra_only.R")

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

## Load R-packages and needed code

```{r}

library(PAMpal)
library(dplyr)
library(here)

source(here("R", "TukeySmoothing.R"))

```

## User-defined parameters

```{r}

# Base directories
baseDir <- "E:/Analysis"

cleanStudyDir      <- file.path(baseDir, "CleanAcousticStudies")
tukeySpectraBaseDir <- file.path(baseDir, "TukeySpectra")

# Study identifiers
DriftID <- "ADRIFT_003"
evNum   <- 1
channel <- 1

# Spectral parameters
wl <- 1024

# Tukey parameters
tukeyMedian      <- 5
tukeyHanningSize <- 5

```

## Build file and folder paths

```{r}

eventFolderName <- sprintf("Event_%03d", evNum) #run per event

binFolder <- file.path(baseDir, "Binaries", DriftID)

db <- file.path(
  baseDir,
  "Useable Databases",
  paste0(DriftID, " - Clean.sqlite3")
)

# Clean acoustic study path
cleanStudyFile <- file.path(
  cleanStudyDir,
  sprintf("%s_CleanAcousticStudy.rds", DriftID)
)
 
# Tukey spectra path (run per event)
eventTukeySpectraDir <- file.path(
  tukeySpectraBaseDir,
  DriftID,
  eventFolderName
)

tukeySpectraFile <- file.path(
  eventTukeySpectraDir,
  sprintf("%s_Event_%03d_Tukey_Spectra.rds", DriftID, evNum)
)

```

## Create output directories

```{r}

dir.create(cleanStudyDir, recursive = TRUE, showWarnings = FALSE)
dir.create(eventTukeySpectraDir, recursive = TRUE, showWarnings = FALSE)

```

## Generate Clean Acoustic Study

```{r}

pps <- PAMpalSettings(
  db = db,
  binaries = binFolder,
  sr_hz = "auto",
  winLen_sec = 0.0025,
  filterfrom_khz = 10,
  filterto_khz = 80
)

study <- processPgDetections(pps)

study<-dplyr::filter(study, Channel==1)

```

## Save Clean Acoustic Study

```{r}

saveRDS(study, cleanStudyFile)

```

## Generate Tukey Spectra

```{r}

tukey_spectra <- calculateAverageSpectra_tukey(
  study,
  evNum = evNum,
  channel = channel,
  tukeyMedian = tukeyMedian,
  tukeyHanningSize = tukeyHanningSize,
  plot = FALSE,
  wl = wl
)

tukey_spectra$EventID <- sprintf("%s_Event_%03d", DriftID, evNum)

```

## Save Tukey Spectra

```{r}

saveRDS(tukey_spectra, tukeySpectraFile)

```




