---
title: "Generate_CleanAcousticStudies_TukeySpectra"
author: "Anne Simonis"
date: "2026-03-04"
output: html_document
---

For a new event from the same drift run the following after generating a new clean acoustic study and running the script: 
evNum <- 2
source("Generate_TukeySpectra_only.R")

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

## Load R-packages and needed code

```{r}

library(PAMpal)
library(dplyr)
library(here)

source(here("R", "TukeySmoothing.R"))

```

## User-defined parameters

```{r}

# Base directories
baseDir <- "E:/Analysis"

cleanStudyDir      <- file.path(baseDir, "CleanAcousticStudies")
tukeySpectraBaseDir <- file.path(baseDir, "TukeySpectra")

# Study identifiers
DriftID <- "ADRIFT_054"
evNum   <- 1
channel <- 1

# Spectral parameters
wl <- 1024

# Tukey parameters
tukeyMedian      <- 5
tukeyHanningSize <- 5

```

## Build file and folder paths

```{r}

eventFolderName <- sprintf("Event_%03d", evNum) #run per event

binFolder <- file.path(baseDir, "Binaries", DriftID)

db <- file.path(
  baseDir,
  "Useable Databases",
  paste0(DriftID, " - Clean.sqlite3")
)

# Clean acoustic study path
cleanStudyFile <- file.path(
  cleanStudyDir,
  sprintf("%s_CleanAcousticStudy.rds", DriftID)
)
 
# Tukey spectra path (run per event)
eventTukeySpectraDir <- file.path(
  tukeySpectraBaseDir,
  DriftID,
  eventFolderName
)

tukeySpectraFile <- file.path(
  eventTukeySpectraDir,
  sprintf("%s_Event_%03d_Tukey_Spectra.rds", DriftID, evNum)
)

```

## Create output directories

```{r}

dir.create(cleanStudyDir, recursive = TRUE, showWarnings = FALSE)
dir.create(eventTukeySpectraDir, recursive = TRUE, showWarnings = FALSE)

```

## Generate Clean Acoustic Study

```{r}

pps <- PAMpalSettings(
  db = db,
  binaries = binFolder,
  sr_hz = "auto",
  winLen_sec = 0.0025,
  filterfrom_khz = 10,
  filterto_khz = 80
)

study <- processPgDetections(pps)

study<-dplyr::filter(study, Channel==1)

```

## Save Clean Acoustic Study

```{r}

saveRDS(study, cleanStudyFile)

```

## Generate Tukey Spectra

```{r}

tukey_spectra <- calculateAverageSpectra_tukey(
  study,
  evNum = evNum,
  channel = channel,
  tukeyMedian = tukeyMedian,
  tukeyHanningSize = tukeyHanningSize,
  plot = FALSE,
  wl = wl
)

tukey_spectra$EventID <- sprintf("%s_Event_%03d", DriftID, evNum)

```

## Save Tukey Spectra

```{r}

saveRDS(tukey_spectra, tukeySpectraFile)

```

## Combined Tukey Spectra Prep

```{r}

# Recursively find all Tukey spectra RDS files
tukey_files <- list.files(
  path = tukeySpectraBaseDir,
  pattern = "_Tukey_Spectra\\.rds$",
  recursive = TRUE,
  full.names = TRUE
)

# Output file path
combined_output_file <- file.path(tukeySpectraBaseDir, "Combined_Tukey_Spectra.rds")

# Exclude combined output file if it already exists
tukey_files <- tukey_files[basename(tukey_files) != "Combined_Tukey_Spectra.rds"]

if (length(tukey_files) == 0) {
  stop("No Tukey spectra RDS files were found in: ", tukeySpectraBaseDir)
}

# Function to parse EventID from filename
parse_event_id <- function(filepath) {
  
  file_name <- basename(filepath)
  
  # Expected filename format:
  # ADRIFT_003_Event_001_Tukey_Spectra.rds
  match_obj <- regexec("^(.*_Event_\\d+)_Tukey_Spectra\\.rds$", file_name)
  match_parts <- regmatches(file_name, match_obj)[[1]]
  
  if (length(match_parts) == 0) {
    stop("Could not parse EventID from filename: ", filepath)
  }
  
  match_parts[2]
}

cat("Number of Tukey spectra files found:", length(tukey_files), "\n")

```

## Reading, Annotating, and Storing Tukey Spectra

```{r}

tukey_list <- lapply(tukey_files, function(f) {
  
  event_id <- parse_event_id(f)
  obj <- readRDS(f)
  
  if (!is.list(obj)) {
    stop(
      "This Tukey spectra file is not a list: ", f,
      "\nClass found: ", paste(class(obj), collapse = ", ")
    )
  }
  
  # Force EventID to match filename
  obj$EventID <- event_id
  
  # Keep file traceability
  obj$SourceFile <- f
  
  obj
})

# Name each list entry by EventID
names(tukey_list) <- vapply(tukey_list, function(x) x$EventID, character(1))

cat("Successfully read", length(tukey_list), "Tukey spectra objects.\n")

```

## Build Combined Tukey Spectra Object

```{r}

Combined_Tukey_Spectra <- tukey_list

cat("Combined_Tukey_Spectra object created as a named list.\n")

```

## Sanity Check: Summary Table

```{r}

combined_summary <- dplyr::bind_rows(
  lapply(Combined_Tukey_Spectra, function(x) {
    data.frame(
      EventID = x$EventID,
      NumUID = length(x$UID),
      Length_freq = length(x$freq),
      Length_avgSpec = length(x$avgSpec),
      Dim_allSpec_rows = if (is.null(dim(x$allSpec))) NA_integer_ else dim(x$allSpec)[1],
      Dim_allSpec_cols = if (is.null(dim(x$allSpec))) NA_integer_ else dim(x$allSpec)[2],
      Length_avgNoise = length(x$avgNoise),
      Dim_allNoise_rows = if (is.null(dim(x$allNoise))) NA_integer_ else dim(x$allNoise)[1],
      Dim_allNoise_cols = if (is.null(dim(x$allNoise))) NA_integer_ else dim(x$allNoise)[2],
      Length_snrKeep = length(x$snrKeep),
      SourceFile = x$SourceFile,
      stringsAsFactors = FALSE
    )
  })
)

print(combined_summary, n = Inf)

```

## Sanity Check: Consistency

```{r}

# Check that EventIDs are unique
if (anyDuplicated(combined_summary$EventID) > 0) {
  warning("Duplicate EventIDs found in Combined_Tukey_Spectra.")
} else {
  cat("EventID uniqueness check passed.\n")
}

# Check that number of columns in allSpec matches number of UIDs
bad_allSpec <- combined_summary %>%
  dplyr::filter(Dim_allSpec_cols != NumUID)

if (nrow(bad_allSpec) > 0) {
  warning("Some events have ncol(allSpec) != length(UID).")
  print(bad_allSpec, n = Inf)
} else {
  cat("allSpec column count matches UID count for all events.\n")
}

# Check that number of columns in allNoise matches number of UIDs
bad_allNoise <- combined_summary %>%
  dplyr::filter(Dim_allNoise_cols != NumUID)

if (nrow(bad_allNoise) > 0) {
  warning("Some events have ncol(allNoise) != length(UID).")
  print(bad_allNoise, n = Inf)
} else {
  cat("allNoise column count matches UID count for all events.\n")
}

# Check that snrKeep length matches UID count
bad_snrKeep <- combined_summary %>%
  dplyr::filter(Length_snrKeep != NumUID)

if (nrow(bad_snrKeep) > 0) {
  warning("Some events have length(snrKeep) != length(UID).")
  print(bad_snrKeep, n = Inf)
} else {
  cat("snrKeep length matches UID count for all events.\n")
}

# Check that avgSpec and freq lengths match
bad_avgSpec <- combined_summary %>%
  dplyr::filter(Length_avgSpec != Length_freq)

if (nrow(bad_avgSpec) > 0) {
  warning("Some events have length(avgSpec) != length(freq).")
  print(bad_avgSpec, n = Inf)
} else {
  cat("avgSpec length matches freq length for all events.\n")
}

# Check that avgNoise and freq lengths match
bad_avgNoise <- combined_summary %>%
  dplyr::filter(Length_avgNoise != Length_freq)

if (nrow(bad_avgNoise) > 0) {
  warning("Some events have length(avgNoise) != length(freq).")
  print(bad_avgNoise, n = Inf)
} else {
  cat("avgNoise length matches freq length for all events.\n")
}

```

## Save Combined Tukey Spectra

```{r}

saveRDS(Combined_Tukey_Spectra, combined_output_file)

cat("Combined Tukey spectra saved to:\n", combined_output_file, "\n", sep = "")

```
















