---
title: "Generate_CleanAcousticStudies_TukeySpectra"
author: "Anne Simonis"
date: "2026-03-04"
output: html_document
---

For a new event from the same drift run the following after generating a new clean acoustic study and running the script: 
evNum <- 2
source("Generate_TukeySpectra_only.R")

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

## Load R-packages and needed code

```{r}

library(PAMpal)
library(dplyr)
library(here)

source(here("R", "TukeySmoothing.R"))

```

## User-defined parameters

```{r}

# Base directories
baseDir <- "E:/Analysis"

cleanStudyDir      <- file.path(baseDir, "CleanAcousticStudies")
tukeySpectraBaseDir <- file.path(baseDir, "TukeySpectra")

# Study identifiers
DriftID <- "ADRIFT_054"
evNum   <- 1
channel <- 1

# Spectral parameters
wl <- 1024

# Tukey parameters
tukeyMedian      <- 5
tukeyHanningSize <- 5

```

## Build file and folder paths

```{r}

eventFolderName <- sprintf("Event_%03d", evNum) #run per event

binFolder <- file.path(baseDir, "Binaries", DriftID)

db <- file.path(
  baseDir,
  "Useable Databases",
  paste0(DriftID, " - Clean.sqlite3")
)

# Clean acoustic study path
cleanStudyFile <- file.path(
  cleanStudyDir,
  sprintf("%s_CleanAcousticStudy.rds", DriftID)
)
 
# Tukey spectra path (run per event)
eventTukeySpectraDir <- file.path(
  tukeySpectraBaseDir,
  DriftID,
  eventFolderName
)

tukeySpectraFile <- file.path(
  eventTukeySpectraDir,
  sprintf("%s_Event_%03d_Tukey_Spectra.rds", DriftID, evNum)
)

```

## Create output directories

```{r}

dir.create(cleanStudyDir, recursive = TRUE, showWarnings = FALSE)
dir.create(eventTukeySpectraDir, recursive = TRUE, showWarnings = FALSE)

```

## Generate Clean Acoustic Study

```{r}

pps <- PAMpalSettings(
  db = db,
  binaries = binFolder,
  sr_hz = "auto",
  winLen_sec = 0.0025,
  filterfrom_khz = 10,
  filterto_khz = 80
)

study <- processPgDetections(pps)

study<-dplyr::filter(study, Channel==1)

```

## Save Clean Acoustic Study

```{r}

saveRDS(study, cleanStudyFile)

```

## Generate Tukey Spectra

```{r}

tukey_spectra <- calculateAverageSpectra_tukey(
  study,
  evNum = evNum,
  channel = channel,
  tukeyMedian = tukeyMedian,
  tukeyHanningSize = tukeyHanningSize,
  plot = FALSE,
  wl = wl
)

tukey_spectra$EventID <- sprintf("%s_Event_%03d", DriftID, evNum)

```

## Save Tukey Spectra

```{r}

saveRDS(tukey_spectra, tukeySpectraFile)

```

## Combined Tukey Spectra Prep

```{r}

# Recursively find all Tukey spectra RDS files
tukey_files <- list.files(
  path = tukeySpectraBaseDir,
  pattern = "_Tukey_Spectra\\.rds$",
  recursive = TRUE,
  full.names = TRUE
)

# Output file path
combined_output_file <- file.path(tukeySpectraBaseDir, "Combined_Tukey_Spectra.rds")

# Exclude combined output file if it already exists
tukey_files <- tukey_files[basename(tukey_files) != "Combined_Tukey_Spectra.rds"]

if (length(tukey_files) == 0) {
  stop("No Tukey spectra RDS files were found in: ", tukeySpectraBaseDir)
}

# Function to parse EventID from filename
parse_event_id <- function(filepath) {
  
  file_name <- basename(filepath)
  
  # Expected filename format:
  # ADRIFT_003_Event_001_Tukey_Spectra.rds
  match_obj <- regexec("^(.*_Event_\\d+)_Tukey_Spectra\\.rds$", file_name)
  match_parts <- regmatches(file_name, match_obj)[[1]]
  
  if (length(match_parts) == 0) {
    stop("Could not parse EventID from filename: ", filepath)
  }
  
  match_parts[2]
}

cat("Number of Tukey spectra files found:", length(tukey_files), "\n")

```

## Reading, Annotating, and Storing Tukey Spectra

```{r}

tukey_list <- lapply(tukey_files, function(f) {
  
  event_id <- parse_event_id(f)
  obj <- readRDS(f)
  
  if (!is.list(obj)) {
    stop(
      "This Tukey spectra file is not a list: ", f,
      "\nClass found: ", paste(class(obj), collapse = ", ")
    )
  }
  
  # Force EventID to match filename
  obj$EventID <- event_id
  
  # Keep file traceability
  obj$SourceFile <- f
  
  obj
})

# Name each list entry by EventID
names(tukey_list) <- vapply(tukey_list, function(x) x$EventID, character(1))

cat("Successfully read", length(tukey_list), "Tukey spectra objects.\n")

```

## Build Combined Tukey Spectra Object

```{r}

Combined_Tukey_Spectra <- tukey_list

cat("Combined_Tukey_Spectra object created as a named list.\n")

```

## Sanity Check: Summary Table

```{r}

combined_summary <- dplyr::bind_rows(
  lapply(Combined_Tukey_Spectra, function(x) {
    data.frame(
      EventID = x$EventID,
      NumUID = length(x$UID),
      Length_freq = length(x$freq),
      Length_avgSpec = length(x$avgSpec),
      Dim_allSpec_rows = if (is.null(dim(x$allSpec))) NA_integer_ else dim(x$allSpec)[1],
      Dim_allSpec_cols = if (is.null(dim(x$allSpec))) NA_integer_ else dim(x$allSpec)[2],
      Length_avgNoise = length(x$avgNoise),
      Dim_allNoise_rows = if (is.null(dim(x$allNoise))) NA_integer_ else dim(x$allNoise)[1],
      Dim_allNoise_cols = if (is.null(dim(x$allNoise))) NA_integer_ else dim(x$allNoise)[2],
      Length_snrKeep = length(x$snrKeep),
      SourceFile = x$SourceFile,
      stringsAsFactors = FALSE
    )
  })
)

print(combined_summary, n = Inf)

```

## Sanity Check: Consistency

```{r}

# Check that EventIDs are unique
if (anyDuplicated(combined_summary$EventID) > 0) {
  warning("Duplicate EventIDs found in Combined_Tukey_Spectra.")
} else {
  cat("EventID uniqueness check passed.\n")
}

# Check that number of columns in allSpec matches number of UIDs
bad_allSpec <- combined_summary %>%
  dplyr::filter(Dim_allSpec_cols != NumUID)

if (nrow(bad_allSpec) > 0) {
  warning("Some events have ncol(allSpec) != length(UID).")
  print(bad_allSpec, n = Inf)
} else {
  cat("allSpec column count matches UID count for all events.\n")
}

# Check that number of columns in allNoise matches number of UIDs
bad_allNoise <- combined_summary %>%
  dplyr::filter(Dim_allNoise_cols != NumUID)

if (nrow(bad_allNoise) > 0) {
  warning("Some events have ncol(allNoise) != length(UID).")
  print(bad_allNoise, n = Inf)
} else {
  cat("allNoise column count matches UID count for all events.\n")
}

# Check that snrKeep length matches UID count
bad_snrKeep <- combined_summary %>%
  dplyr::filter(Length_snrKeep != NumUID)

if (nrow(bad_snrKeep) > 0) {
  warning("Some events have length(snrKeep) != length(UID).")
  print(bad_snrKeep, n = Inf)
} else {
  cat("snrKeep length matches UID count for all events.\n")
}

# Check that avgSpec and freq lengths match
bad_avgSpec <- combined_summary %>%
  dplyr::filter(Length_avgSpec != Length_freq)

if (nrow(bad_avgSpec) > 0) {
  warning("Some events have length(avgSpec) != length(freq).")
  print(bad_avgSpec, n = Inf)
} else {
  cat("avgSpec length matches freq length for all events.\n")
}

# Check that avgNoise and freq lengths match
bad_avgNoise <- combined_summary %>%
  dplyr::filter(Length_avgNoise != Length_freq)

if (nrow(bad_avgNoise) > 0) {
  warning("Some events have length(avgNoise) != length(freq).")
  print(bad_avgNoise, n = Inf)
} else {
  cat("avgNoise length matches freq length for all events.\n")
}

```

## Save Combined Tukey Spectra

```{r}

saveRDS(Combined_Tukey_Spectra, combined_output_file)

cat("Combined Tukey spectra saved to:\n", combined_output_file, "\n", sep = "")

```

## Deep Dive: ADRIFT_003 Event 001 vs Event 002

## Confirm the two files actually exist as separate files

```{r}

library(dplyr)

baseDir <- "E:/Analysis"

cleanStudyDir <- file.path(baseDir, "CleanAcousticStudies")
tukeySpectraBaseDir <- file.path(baseDir, "TukeySpectra")

target_drift <- "ADRIFT_003"
target_event_nums <- c(1, 2)

target_event_labels <- sprintf(
  "%s_Event_%03d",
  target_drift,
  target_event_nums
)

target_files <- file.path(
  tukeySpectraBaseDir,
  target_drift,
  sprintf("Event_%03d", target_event_nums),
  sprintf(
    "%s_Event_%03d_Tukey_Spectra.rds",
    target_drift,
    target_event_nums
  )
)

target_file_info <- file.info(target_files)

target_file_df <- data.frame(
  EventLabel = target_event_labels,
  File = target_files,
  Exists = file.exists(target_files),
  SizeBytes = target_file_info$size,
  ModifiedTime = target_file_info$mtime,
  stringsAsFactors = FALSE
)

print(target_file_df)

if (!all(target_file_df$Exists)) {
  stop("One or both ADRIFT_003 target Tukey files are missing. Stop here.")
}

cat("\nFile MD5 hashes:\n")
print(tools::md5sum(target_files))

```

## Check whether ADRIFT_003 has more than one offline event

```{r}

library(DBI)
library(RSQLite)
library(dplyr)

target_drift <- "ADRIFT_003"

db_file <- file.path(
  baseDir,
  "Useable Databases",
  paste0(target_drift, " - Clean.sqlite3")
)

cat("Checking database:\n")
print(db_file)

con <- dbConnect(RSQLite::SQLite(), db_file)

offline_events <- dbGetQuery(
  con,
  "SELECT * FROM Click_Detector_OfflineEvents"
)

offline_clicks <- dbGetQuery(
  con,
  "SELECT UID, EventId, UTC, UTCMilliseconds, ClickNo, parentID, parentUID FROM Click_Detector_OfflineClicks"
)

dbDisconnect(con)

cat("\nOffline events table:\n")
print(offline_events)

cat("\nOffline clicks per EventId:\n")
clicks_per_sql_event <- offline_clicks %>%
  count(EventId, name = "n_clicks") %>%
  arrange(EventId)

print(clicks_per_sql_event)

cat("\nUnique EventIds in OfflineClicks:\n")
print(sort(unique(offline_clicks$EventId)))

```

## Compare each Tukey file against SQL EventId 1 and 2

```{r}

library(DBI)
library(RSQLite)
library(dplyr)

baseDir <- "E:/Analysis"
tukeySpectraBaseDir <- file.path(baseDir, "TukeySpectra")
target_drift <- "ADRIFT_003"

event_001_file <- file.path(
  tukeySpectraBaseDir,
  target_drift,
  "Event_001",
  "ADRIFT_003_Event_001_Tukey_Spectra.rds"
)

event_002_file <- file.path(
  tukeySpectraBaseDir,
  target_drift,
  "Event_002",
  "ADRIFT_003_Event_002_Tukey_Spectra.rds"
)

tukey_001 <- readRDS(event_001_file)
tukey_002 <- readRDS(event_002_file)

db_file <- file.path(
  baseDir,
  "Useable Databases",
  paste0(target_drift, " - Clean.sqlite3")
)

con <- dbConnect(RSQLite::SQLite(), db_file)

sql_clicks <- dbGetQuery(
  con,
  "SELECT UID, EventId FROM Click_Detector_OfflineClicks"
)

dbDisconnect(con)

sql_event_uid_sets <- split(
  as.character(sql_clicks$UID),
  sql_clicks$EventId
)

compare_tukey_to_sql_event <- function(tukey_obj, tukey_label, sql_event_uid_sets) {
  
  tukey_uids <- as.character(tukey_obj$UID)
  
  bind_rows(
    lapply(names(sql_event_uid_sets), function(sql_event_id) {
      
      sql_uids <- as.character(sql_event_uid_sets[[sql_event_id]])
      
      data.frame(
        TukeyFileLabel = tukey_label,
        TukeyNumUID = length(tukey_uids),
        TukeyNumUniqueUID = length(unique(tukey_uids)),
        SQLEventId = sql_event_id,
        SQLNumUID = length(sql_uids),
        SQLNumUniqueUID = length(unique(sql_uids)),
        OverlapUIDs = sum(tukey_uids %in% sql_uids),
        PercentTukeyInSQLEvent = 100 * sum(tukey_uids %in% sql_uids) / length(tukey_uids),
        PercentSQLEventInTukey = 100 * sum(sql_uids %in% tukey_uids) / length(sql_uids),
        SameUIDSet = setequal(tukey_uids, sql_uids),
        stringsAsFactors = FALSE
      )
    })
  )
}

uid_match_check <- bind_rows(
  compare_tukey_to_sql_event(
    tukey_001,
    "ADRIFT_003_Event_001_TukeyFile",
    sql_event_uid_sets
  ),
  compare_tukey_to_sql_event(
    tukey_002,
    "ADRIFT_003_Event_002_TukeyFile",
    sql_event_uid_sets
  )
)

print(uid_match_check)

```

## Test which evNum maps to each SQL EventId

```{r}

library(PAMpal)
library(dplyr)
library(DBI)
library(RSQLite)
library(here)

# Make sure Tukey function is available
source(here("R", "TukeySmoothing.R"))

baseDir <- "E:/Analysis"
cleanStudyDir <- file.path(baseDir, "CleanAcousticStudies")
tukeySpectraBaseDir <- file.path(baseDir, "TukeySpectra")

target_drift <- "ADRIFT_003"
channel <- 1

wl <- 1024
tukeyMedian <- 5
tukeyHanningSize <- 5

cleanStudyFile <- file.path(
  cleanStudyDir,
  sprintf("%s_CleanAcousticStudy.rds", target_drift)
)

db_file <- file.path(
  baseDir,
  "Useable Databases",
  paste0(target_drift, " - Clean.sqlite3")
)

cat("Clean study file:\n")
print(cleanStudyFile)

cat("\nDatabase file:\n")
print(db_file)

if (!file.exists(cleanStudyFile)) {
  stop("Clean acoustic study file does not exist: ", cleanStudyFile)
}

if (!file.exists(db_file)) {
  stop("SQLite database file does not exist: ", db_file)
}

# -----------------------------
# Load clean acoustic study
# -----------------------------

study_003 <- readRDS(cleanStudyFile)

cat("\nLoaded clean study class:\n")
print(class(study_003))

# -----------------------------
# Load SQL event UID sets
# -----------------------------

con <- dbConnect(RSQLite::SQLite(), db_file)

sql_clicks_003 <- dbGetQuery(
  con,
  "SELECT UID, EventId FROM Click_Detector_OfflineClicks"
)

dbDisconnect(con)

sql_event_uid_sets <- split(
  as.character(sql_clicks_003$UID),
  sql_clicks_003$EventId
)

cat("\nSQL click counts by EventId:\n")
print(
  sql_clicks_003 %>%
    count(EventId, name = "n_clicks") %>%
    arrange(EventId)
)

# -----------------------------
# Test candidate evNum values
# -----------------------------

test_tukey_evnum <- function(ev_num) {
  
  cat("\n============================\n")
  cat("Testing evNum =", ev_num, "\n")
  cat("============================\n")
  
  result <- tryCatch(
    {
      tukey_obj <- calculateAverageSpectra_tukey(
        study_003,
        evNum = ev_num,
        channel = channel,
        tukeyMedian = tukeyMedian,
        tukeyHanningSize = tukeyHanningSize,
        plot = FALSE,
        wl = wl
      )
      
      tukey_uids <- as.character(tukey_obj$UID)
      
      comparison <- bind_rows(
        lapply(names(sql_event_uid_sets), function(sql_event_id) {
          
          sql_uids <- as.character(sql_event_uid_sets[[sql_event_id]])
          
          data.frame(
            evNum_tested = ev_num,
            TukeyNumUID = length(tukey_uids),
            TukeyNumUniqueUID = length(unique(tukey_uids)),
            SQLEventId = sql_event_id,
            SQLNumUID = length(sql_uids),
            OverlapUIDs = sum(tukey_uids %in% sql_uids),
            PercentTukeyInSQLEvent = 100 * sum(tukey_uids %in% sql_uids) / length(tukey_uids),
            PercentSQLEventInTukey = 100 * sum(sql_uids %in% tukey_uids) / length(sql_uids),
            SameUIDSet = setequal(tukey_uids, sql_uids),
            Error = NA_character_,
            stringsAsFactors = FALSE
          )
        })
      )
      
      comparison
    },
    error = function(e) {
      data.frame(
        evNum_tested = ev_num,
        TukeyNumUID = NA_integer_,
        TukeyNumUniqueUID = NA_integer_,
        SQLEventId = NA_character_,
        SQLNumUID = NA_integer_,
        OverlapUIDs = NA_integer_,
        PercentTukeyInSQLEvent = NA_real_,
        PercentSQLEventInTukey = NA_real_,
        SameUIDSet = NA,
        Error = conditionMessage(e),
        stringsAsFactors = FALSE
      )
    }
  )
  
  return(result)
}

candidate_evnums <- 1:5

evnum_mapping_results <- bind_rows(
  lapply(candidate_evnums, test_tukey_evnum)
)

print(evnum_mapping_results)

```

## Back up the existing bad ADRIFT_003 files

```{r}

backup_dir <- file.path(
  tukeySpectraBaseDir,
  target_drift,
  "BACKUP_bad_duplicate_Event001_Event002"
)

dir.create(backup_dir, recursive = TRUE, showWarnings = FALSE)

backup_files <- file.path(
  backup_dir,
  basename(target_files)
)

backup_result <- file.copy(
  from = target_files,
  to = backup_files,
  overwrite = TRUE
)

backup_df <- data.frame(
  OriginalFile = target_files,
  BackupFile = backup_files,
  BackupSucceeded = backup_result,
  OriginalMD5 = as.character(tools::md5sum(target_files)),
  BackupMD5 = as.character(tools::md5sum(backup_files)),
  stringsAsFactors = FALSE
)

print(backup_df)

if (!all(backup_df$BackupSucceeded)) {
  stop("Backup failed for one or more files. Do not continue.")
}

if (!all(backup_df$OriginalMD5 == backup_df$BackupMD5)) {
  stop("Backup MD5 does not match original. Do not continue.")
}

cat("\nBackup complete. Safe to regenerate files.\n")

```

## Regenerate ADRIFT_003 Event 001 and Event 002

```{r}

library(PAMpal)
library(dplyr)
library(here)

source(here("R", "TukeySmoothing.R"))

target_drift <- "ADRIFT_003"
channel <- 1

wl <- 1024
tukeyMedian <- 5
tukeyHanningSize <- 5

cleanStudyFile <- file.path(
  cleanStudyDir,
  sprintf("%s_CleanAcousticStudy.rds", target_drift)
)

if (!file.exists(cleanStudyFile)) {
  stop("Missing clean study file: ", cleanStudyFile)
}

study_003 <- readRDS(cleanStudyFile)

regenerate_tukey_event <- function(ev_num) {
  
  event_folder_name <- sprintf("Event_%03d", ev_num)
  event_label <- sprintf("%s_Event_%03d", target_drift, ev_num)
  
  event_tukey_dir <- file.path(
    tukeySpectraBaseDir,
    target_drift,
    event_folder_name
  )
  
  dir.create(event_tukey_dir, recursive = TRUE, showWarnings = FALSE)
  
  tukey_file <- file.path(
    event_tukey_dir,
    sprintf("%s_Event_%03d_Tukey_Spectra.rds", target_drift, ev_num)
  )
  
  cat("\n============================\n")
  cat("Regenerating", event_label, "\n")
  cat("Output file:", tukey_file, "\n")
  cat("============================\n")
  
  tukey_spectra <- calculateAverageSpectra_tukey(
    study_003,
    evNum = ev_num,
    channel = channel,
    tukeyMedian = tukeyMedian,
    tukeyHanningSize = tukeyHanningSize,
    plot = FALSE,
    wl = wl
  )
  
  tukey_spectra$EventID <- event_label
  tukey_spectra$SourceFile <- tukey_file
  
  saveRDS(tukey_spectra, tukey_file)
  
  data.frame(
    EventID = event_label,
    File = tukey_file,
    NumUID = length(tukey_spectra$UID),
    NumUniqueUID = length(unique(tukey_spectra$UID)),
    DimAllSpecRows = dim(tukey_spectra$allSpec)[1],
    DimAllSpecCols = dim(tukey_spectra$allSpec)[2],
    MD5 = as.character(tools::md5sum(tukey_file)),
    stringsAsFactors = FALSE
  )
}

regenerated_summary <- dplyr::bind_rows(
  regenerate_tukey_event(1),
  regenerate_tukey_event(2)
)

print(regenerated_summary)

```

## Verify regenerated files against SQL events

```{r}

library(DBI)
library(RSQLite)
library(dplyr)

target_drift <- "ADRIFT_003"

event_001_file <- file.path(
  tukeySpectraBaseDir,
  target_drift,
  "Event_001",
  "ADRIFT_003_Event_001_Tukey_Spectra.rds"
)

event_002_file <- file.path(
  tukeySpectraBaseDir,
  target_drift,
  "Event_002",
  "ADRIFT_003_Event_002_Tukey_Spectra.rds"
)

tukey_001_new <- readRDS(event_001_file)
tukey_002_new <- readRDS(event_002_file)

# Reload SQL UID sets to make this chunk self-contained
db_file <- file.path(
  baseDir,
  "Useable Databases",
  paste0(target_drift, " - Clean.sqlite3")
)

con <- dbConnect(RSQLite::SQLite(), db_file)

sql_clicks_003 <- dbGetQuery(
  con,
  "SELECT UID, EventId FROM Click_Detector_OfflineClicks"
)

dbDisconnect(con)

sql_event_uid_sets <- split(
  as.character(sql_clicks_003$UID),
  sql_clicks_003$EventId
)

compare_tukey_to_sql_event <- function(tukey_obj, tukey_label, sql_event_uid_sets) {
  
  tukey_uids <- as.character(tukey_obj$UID)
  
  bind_rows(
    lapply(names(sql_event_uid_sets), function(sql_event_id) {
      
      sql_uids <- as.character(sql_event_uid_sets[[sql_event_id]])
      
      data.frame(
        TukeyFileLabel = tukey_label,
        TukeyNumUID = length(tukey_uids),
        TukeyNumUniqueUID = length(unique(tukey_uids)),
        SQLEventId = sql_event_id,
        SQLNumUID = length(sql_uids),
        SQLNumUniqueUID = length(unique(sql_uids)),
        OverlapUIDs = sum(tukey_uids %in% sql_uids),
        PercentTukeyInSQLEvent = 100 * sum(tukey_uids %in% sql_uids) / length(tukey_uids),
        PercentSQLEventInTukey = 100 * sum(sql_uids %in% tukey_uids) / length(sql_uids),
        SameUIDSet = setequal(tukey_uids, sql_uids),
        stringsAsFactors = FALSE
      )
    })
  )
}

verification_after_regen <- dplyr::bind_rows(
  compare_tukey_to_sql_event(
    tukey_001_new,
    "ADRIFT_003_Event_001_REGENERATED",
    sql_event_uid_sets
  ),
  compare_tukey_to_sql_event(
    tukey_002_new,
    "ADRIFT_003_Event_002_REGENERATED",
    sql_event_uid_sets
  )
)

print(verification_after_regen)

cat("\nNew MD5 hashes:\n")
print(tools::md5sum(c(event_001_file, event_002_file)))

cat("\nAre regenerated Event_001 and Event_002 byte-identical?\n")
print(identical(
  as.character(tools::md5sum(event_001_file)),
  as.character(tools::md5sum(event_002_file))
))

```

## Back up the existing combined file

```{r}

combined_output_file <- file.path(
  tukeySpectraBaseDir,
  "Combined_Tukey_Spectra.rds"
)

combined_backup_dir <- file.path(
  tukeySpectraBaseDir,
  "BACKUP_combined_before_ADRIFT003_fix"
)

dir.create(combined_backup_dir, recursive = TRUE, showWarnings = FALSE)

timestamp <- format(Sys.time(), "%Y%m%d_%H%M%S")

combined_backup_file <- file.path(
  combined_backup_dir,
  paste0("Combined_Tukey_Spectra_before_ADRIFT003_fix_", timestamp, ".rds")
)

if (file.exists(combined_output_file)) {
  
  backup_ok <- file.copy(
    from = combined_output_file,
    to = combined_backup_file,
    overwrite = FALSE
  )
  
  backup_df <- data.frame(
    OriginalFile = combined_output_file,
    BackupFile = combined_backup_file,
    BackupSucceeded = backup_ok,
    OriginalMD5 = as.character(tools::md5sum(combined_output_file)),
    BackupMD5 = as.character(tools::md5sum(combined_backup_file)),
    stringsAsFactors = FALSE
  )
  
  print(backup_df)
  
  if (!backup_ok) {
    stop("Combined file backup failed. Do not continue.")
  }
  
  if (backup_df$OriginalMD5 != backup_df$BackupMD5) {
    stop("Combined file backup MD5 mismatch. Do not continue.")
  }
  
  cat("\nExisting combined file backed up successfully.\n")
  
} else {
  
  cat("No existing combined file found. A new one will be created.\n")
}

```

## Rebuild combined Tukey object safely

```{r}

# Find Tukey spectra files, but only inside proper Event_### folders.
# This avoids accidentally including backup files.
all_candidate_tukey_files <- list.files(
  path = tukeySpectraBaseDir,
  pattern = "_Tukey_Spectra\\.rds$",
  recursive = TRUE,
  full.names = TRUE
)

tukey_files <- all_candidate_tukey_files[
  grepl("[/\\\\]Event_\\d{3}[/\\\\].*_Tukey_Spectra\\.rds$", all_candidate_tukey_files)
]

# Explicitly exclude any backup folders just in case
tukey_files <- tukey_files[
  !grepl("BACKUP|backup|Backup", tukey_files)
]

# Exclude combined output file if somehow matched
tukey_files <- tukey_files[
  basename(tukey_files) != "Combined_Tukey_Spectra.rds"
]

cat("Total candidate Tukey files found:", length(all_candidate_tukey_files), "\n")
cat("Valid Event_### Tukey files used:", length(tukey_files), "\n\n")

if (length(tukey_files) == 0) {
  stop("No valid Tukey spectra files were found.")
}

parse_event_id <- function(filepath) {
  
  file_name <- basename(filepath)
  
  match_obj <- regexec("^(.*_Event_\\d+)_Tukey_Spectra\\.rds$", file_name)
  match_parts <- regmatches(file_name, match_obj)[[1]]
  
  if (length(match_parts) == 0) {
    stop("Could not parse EventID from filename: ", filepath)
  }
  
  match_parts[2]
}

tukey_list <- lapply(tukey_files, function(f) {
  
  event_id <- parse_event_id(f)
  obj <- readRDS(f)
  
  if (!is.list(obj)) {
    stop(
      "This Tukey spectra file is not a list: ", f,
      "\nClass found: ", paste(class(obj), collapse = ", ")
    )
  }
  
  obj$EventID <- event_id
  obj$SourceFile <- f
  
  obj
})

names(tukey_list) <- vapply(tukey_list, function(x) x$EventID, character(1))

Combined_Tukey_Spectra_NEW <- tukey_list

cat("Combined_Tukey_Spectra_NEW object created.\n")
cat("Number of events:", length(Combined_Tukey_Spectra_NEW), "\n")

```

## Run stronger sanity checks before saving

```{r}

combined_summary_new <- dplyr::bind_rows(
  lapply(Combined_Tukey_Spectra_NEW, function(x) {
    data.frame(
      EventID = x$EventID,
      DriftID = sub("_Event_.*$", "", x$EventID),
      NumUID = length(x$UID),
      NumUniqueUID = length(unique(x$UID)),
      Length_freq = length(x$freq),
      Length_avgSpec = length(x$avgSpec),
      Dim_allSpec_rows = if (is.null(dim(x$allSpec))) NA_integer_ else dim(x$allSpec)[1],
      Dim_allSpec_cols = if (is.null(dim(x$allSpec))) NA_integer_ else dim(x$allSpec)[2],
      Length_avgNoise = length(x$avgNoise),
      Dim_allNoise_rows = if (is.null(dim(x$allNoise))) NA_integer_ else dim(x$allNoise)[1],
      Dim_allNoise_cols = if (is.null(dim(x$allNoise))) NA_integer_ else dim(x$allNoise)[2],
      Length_snrKeep = length(x$snrKeep),
      SourceFile = x$SourceFile,
      FileMD5 = as.character(tools::md5sum(x$SourceFile)),
      stringsAsFactors = FALSE
    )
  })
)

print(combined_summary_new, n = Inf)

cat("\nTotal clicks in rebuilt combined object:\n")
print(sum(combined_summary_new$NumUID))

cat("\nADRIFT_003 rebuilt summary:\n")
print(
  combined_summary_new %>%
    filter(DriftID == "ADRIFT_003")
)

# -----------------------------
# Within-object consistency checks
# -----------------------------

bad_event_id_duplicates <- combined_summary_new %>%
  count(EventID) %>%
  filter(n > 1)

bad_uid_counts <- combined_summary_new %>%
  filter(NumUID != NumUniqueUID)

bad_allSpec <- combined_summary_new %>%
  filter(Dim_allSpec_cols != NumUID)

bad_allNoise <- combined_summary_new %>%
  filter(Dim_allNoise_cols != NumUID)

bad_snrKeep <- combined_summary_new %>%
  filter(Length_snrKeep != NumUID)

bad_avgSpec <- combined_summary_new %>%
  filter(Length_avgSpec != Length_freq)

bad_avgNoise <- combined_summary_new %>%
  filter(Length_avgNoise != Length_freq)

cat("\nDuplicate EventIDs:\n")
print(bad_event_id_duplicates)

cat("\nEvents with duplicated UIDs within the event:\n")
print(bad_uid_counts)

cat("\nEvents with allSpec column mismatch:\n")
print(bad_allSpec)

cat("\nEvents with allNoise column mismatch:\n")
print(bad_allNoise)

cat("\nEvents with snrKeep mismatch:\n")
print(bad_snrKeep)

cat("\nEvents with avgSpec/freq mismatch:\n")
print(bad_avgSpec)

cat("\nEvents with avgNoise/freq mismatch:\n")
print(bad_avgNoise)

# -----------------------------
# Cross-event duplicate UID check within each DriftID
# -----------------------------

combined_click_keys_new <- dplyr::bind_rows(
  lapply(Combined_Tukey_Spectra_NEW, function(x) {
    data.frame(
      EventID = x$EventID,
      DriftID = sub("_Event_.*$", "", x$EventID),
      UID = as.character(x$UID),
      stringsAsFactors = FALSE
    )
  })
)

duplicate_drift_uid_new <- combined_click_keys_new %>%
  count(DriftID, UID, name = "n_rows") %>%
  filter(n_rows > 1) %>%
  arrange(DriftID, UID)

cat("\nDuplicate DriftID + UID combinations across rebuilt combined object:\n")
print(duplicate_drift_uid_new)

cat("\nNumber of duplicate DriftID + UID combinations:\n")
print(nrow(duplicate_drift_uid_new))

# -----------------------------
# Stop if anything serious is wrong
# -----------------------------

if (nrow(bad_event_id_duplicates) > 0) {
  stop("Duplicate EventIDs found. Do not save.")
}

if (nrow(bad_uid_counts) > 0) {
  stop("Duplicated UIDs within at least one event. Do not save.")
}

if (nrow(bad_allSpec) > 0) {
  stop("allSpec column mismatch found. Do not save.")
}

if (nrow(bad_allNoise) > 0) {
  stop("allNoise column mismatch found. Do not save.")
}

if (nrow(bad_snrKeep) > 0) {
  stop("snrKeep length mismatch found. Do not save.")
}

if (nrow(bad_avgSpec) > 0) {
  stop("avgSpec/freq mismatch found. Do not save.")
}

if (nrow(bad_avgNoise) > 0) {
  stop("avgNoise/freq mismatch found. Do not save.")
}

if (nrow(duplicate_drift_uid_new) > 0) {
  stop("Duplicate DriftID + UID combinations found. Do not save.")
}

cat("\nAll pre-save sanity checks passed.\n")

```

## Save the rebuilt combined file

```{r}

saveRDS(
  Combined_Tukey_Spectra_NEW,
  combined_output_file
)

cat("Corrected Combined_Tukey_Spectra saved to:\n")
cat(combined_output_file, "\n")

cat("\nSaved file MD5:\n")
print(tools::md5sum(combined_output_file))

```

## Final Sanity Checks

```{r}

summarize_combined <- function(x, label) {
  dplyr::bind_rows(lapply(x, function(e) {
    data.frame(
      Version = label,
      EventID = e$EventID,
      DriftID = sub("_Event_.*$", "", e$EventID),
      NumUID = length(e$UID),
      NumUniqueUID = length(unique(e$UID)),
      allSpec_cols = ncol(e$allSpec),
      allNoise_cols = ncol(e$allNoise),
      snrKeep_len = length(e$snrKeep),
      stringsAsFactors = FALSE
    )
  }))
}

old_sum <- summarize_combined(combined_tukey, "OLD")
new_sum <- summarize_combined(Combined_Tukey_Spectra_NEW, "NEW")

check_df <- old_sum %>%
  select(
    EventID, DriftID,
    Old_NumUID = NumUID,
    Old_NumUniqueUID = NumUniqueUID,
    Old_allSpec_cols = allSpec_cols,
    Old_allNoise_cols = allNoise_cols,
    Old_snrKeep_len = snrKeep_len
  ) %>%
  full_join(
    new_sum %>%
      select(
        EventID, DriftID,
        New_NumUID = NumUID,
        New_NumUniqueUID = NumUniqueUID,
        New_allSpec_cols = allSpec_cols,
        New_allNoise_cols = allNoise_cols,
        New_snrKeep_len = snrKeep_len
      ),
    by = c("EventID", "DriftID")
  ) %>%
  mutate(
    Delta_NumUID = New_NumUID - Old_NumUID,
    Delta_allSpec_cols = New_allSpec_cols - Old_allSpec_cols,
    Delta_allNoise_cols = New_allNoise_cols - Old_allNoise_cols,
    Delta_snrKeep_len = New_snrKeep_len - Old_snrKeep_len
  ) %>%
  arrange(DriftID, EventID)

changed_events <- check_df %>%
  filter(
    Delta_NumUID != 0 |
      Delta_allSpec_cols != 0 |
      Delta_allNoise_cols != 0 |
      Delta_snrKeep_len != 0 |
      is.na(Delta_NumUID)
  )

old_total <- sum(old_sum$NumUID)
new_total <- sum(new_sum$NumUID)

cat("OLD total clicks:", old_total, "\n")
cat("NEW total clicks:", new_total, "\n")
cat("Difference:", new_total - old_total, "\n\n")

cat("Changed events:\n")
print(changed_events, n = Inf)

# Expected: only ADRIFT_003_Event_001 changed from 13,599 to 3,203
expected_event <- "ADRIFT_003_Event_001"

if (nrow(changed_events) != 1 ||
    changed_events$EventID != expected_event ||
    changed_events$Old_NumUID != 13599 ||
    changed_events$New_NumUID != 3203 ||
    changed_events$Delta_NumUID != -10396 ||
    changed_events$Delta_allSpec_cols != -10396 ||
    changed_events$Delta_allNoise_cols != -10396 ||
    changed_events$Delta_snrKeep_len != -10396) {
  stop("FAIL: Unexpected change detected. Inspect changed_events.")
}

# Check internal consistency of new combined object
new_internal_problems <- new_sum %>%
  filter(
    NumUID != NumUniqueUID |
      NumUID != allSpec_cols |
      NumUID != allNoise_cols |
      NumUID != snrKeep_len
  )

if (nrow(new_internal_problems) > 0) {
  print(new_internal_problems, n = Inf)
  stop("FAIL: Internal UID/allSpec/allNoise/snrKeep mismatch in NEW combined object.")
}

if (new_total != old_total - 10396) {
  stop("FAIL: New total does not equal old total minus 10,396.")
}

cat("\nPASS: Only ADRIFT_003_Event_001 changed, from 13,599 to 3,203 clicks.\n")
cat("PASS: All NEW event UID/allSpec/allNoise/snrKeep counts line up.\n")

```







