---
title: "ICI_diagnostics_minmax_clusters_kmeans"
author: "Sarah Moseley"
date: "2026-05-05"
output: html_document
---

```{r}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
library(readr)

```

## User Settings

```{r}

# Folder from the spectra-only clustering script
base_output_dir <- "E:/Analysis/Clustering Output/MinMax_Normalized"

analysis_name <- "Min-Max Normalized"
cluster_method <- "kmeans"

# Use the k values you actually generated cluster assignments for
k_values_to_check <- 2:12

# We will use this later to remove obviously invalid intervals
# This is just a placeholder for now; don't overinterpret yet.
max_reasonable_ici_ms <- 1000

```

## Load the original combined Tukey spectra object

```{r}

combined_file <- file.choose()
combined_tukey <- readRDS(combined_file)

cat("Number of events:", length(combined_tukey), "\n")
cat("Class of first event object:\n")
print(class(combined_tukey[[1]]))

cat("\nNames in first event object:\n")
print(names(combined_tukey[[1]]))

```

## Inspect object dimensions and UID/source fields

```{r}

first_event <- combined_tukey[[1]]

cat("First EventID:", first_event$EventID, "\n\n")

cat("Length of freq:", length(first_event$freq), "\n")
cat("Length of UID:", length(first_event$UID), "\n")
cat("Dimensions of allSpec:", dim(first_event$allSpec), "\n")
cat("Dimensions of allNoise:", dim(first_event$allNoise), "\n")

cat("\nFirst 20 UID values:\n")
print(head(first_event$UID, 20))

cat("\nSourceFile field:\n")
print(first_event$SourceFile)

cat("\nClass of SourceFile:\n")
print(class(first_event$SourceFile))

cat("\nLength of SourceFile:\n")
print(length(first_event$SourceFile))

```

## Summarize fields across all events

```{r}

event_structure_summary <- data.frame(
  EventIndex = seq_along(combined_tukey),
  EventID = sapply(combined_tukey, function(x) x$EventID),
  n_freq = sapply(combined_tukey, function(x) length(x$freq)),
  n_uid = sapply(combined_tukey, function(x) length(x$UID)),
  allSpec_rows = sapply(combined_tukey, function(x) dim(x$allSpec)[1]),
  allSpec_cols = sapply(combined_tukey, function(x) dim(x$allSpec)[2]),
  sourcefile_length = sapply(combined_tukey, function(x) length(x$SourceFile)),
  stringsAsFactors = FALSE
)

print(event_structure_summary)

cat("\nDo UID counts match allSpec click columns for every event?\n")
print(all(event_structure_summary$n_uid == event_structure_summary$allSpec_cols))

cat("\nTotal clicks represented in combined_tukey:\n")
print(sum(event_structure_summary$allSpec_cols))

```

7. Build click-level metadata table from combined_tukey

```{r}

click_metadata_from_tukey <- list()

for (i in seq_along(combined_tukey)) {
  
  event_obj <- combined_tukey[[i]]
  
  n_clicks <- ncol(event_obj$allSpec)
  
  click_metadata_from_tukey[[i]] <- data.frame(
    EventIndex = i,
    EventID = event_obj$EventID,
    ClickIndexWithinEvent = seq_len(n_clicks),
    UID = as.character(event_obj$UID),
    SourceFile = event_obj$SourceFile,
    stringsAsFactors = FALSE
  )
}

click_metadata_from_tukey <- do.call(rbind, click_metadata_from_tukey)

cat("Total click rows:", nrow(click_metadata_from_tukey), "\n")
cat("Number of events:", length(unique(click_metadata_from_tukey$EventID)), "\n\n")

head(click_metadata_from_tukey)

```

## Check whether UID can be used for ordering

```{r}

uid_diagnostic <- click_metadata_from_tukey %>%
  mutate(
    UID_numeric = suppressWarnings(as.numeric(UID))
  ) %>%
  group_by(EventID) %>%
  summarise(
    n_clicks = n(),
    n_missing_numeric_uid = sum(is.na(UID_numeric)),
    uid_min = min(UID_numeric, na.rm = TRUE),
    uid_max = max(UID_numeric, na.rm = TRUE),
    uid_strictly_increasing_in_current_order = all(diff(UID_numeric) > 0, na.rm = TRUE),
    uid_non_decreasing_in_current_order = all(diff(UID_numeric) >= 0, na.rm = TRUE),
    median_uid_diff = median(diff(UID_numeric), na.rm = TRUE),
    min_uid_diff = min(diff(UID_numeric), na.rm = TRUE),
    max_uid_diff = max(diff(UID_numeric), na.rm = TRUE),
    .groups = "drop"
  )

print(uid_diagnostic)

cat("\nHow many events have UID increasing in the current click order?\n")
print(table(uid_diagnostic$uid_strictly_increasing_in_current_order))

cat("\nSummary of median UID differences across events:\n")
print(summary(uid_diagnostic$median_uid_diff))

```

## Locate possible source databases

```{r}

possible_database_dirs <- c(
  "E:/Analysis/Useable Databases",
  "E:/Analysis/CleanAcousticStudies"
)

existing_database_dirs <- possible_database_dirs[dir.exists(possible_database_dirs)]

cat("Existing search directories:\n")
print(existing_database_dirs)

possible_database_files <- list()

for (search_dir in existing_database_dirs) {
  
  possible_database_files[[search_dir]] <- list.files(
    path = search_dir,
    pattern = "\\.(rds|RDS|RData|rdata|sqlite|sqlite3|db|csv)$",
    recursive = TRUE,
    full.names = TRUE
  )
}

possible_database_files <- unlist(possible_database_files, use.names = FALSE)

cat("\nNumber of possible database/source files found:", length(possible_database_files), "\n\n")

possible_database_df <- data.frame(
  File = possible_database_files,
  FileName = basename(possible_database_files),
  Folder = dirname(possible_database_files),
  stringsAsFactors = FALSE
) %>%
  filter(
    str_detect(
      tolower(FileName),
      "acoustic|study|database|pam|adrift|click|det|event|useable|usable"
    )
  )

print(possible_database_df)

```

## Inspect one source SQLite database

```{r}

if (!requireNamespace("DBI", quietly = TRUE)) {
  install.packages("DBI")
}

if (!requireNamespace("RSQLite", quietly = TRUE)) {
  install.packages("RSQLite")
}

library(DBI)
library(RSQLite)

test_db <- "E:/Analysis/Useable Databases/ADRIFT_003 - Clean.sqlite3"

cat("Testing database:\n")
print(test_db)

con <- dbConnect(RSQLite::SQLite(), test_db)

db_tables <- dbListTables(con)

cat("\nTables in database:\n")
print(db_tables)

# Summarize columns in each table
table_column_summary <- lapply(db_tables, function(tbl) {
  
  cols <- dbListFields(con, tbl)
  
  data.frame(
    Table = tbl,
    Column = cols,
    stringsAsFactors = FALSE
  )
  
}) %>%
  bind_rows()

cat("\nColumn summary:\n")
print(table_column_summary)

# Look for tables/columns likely related to click timing
candidate_columns <- table_column_summary %>%
  filter(
    str_detect(
      tolower(Column),
      "uid|utc|time|sample|start|peak|millis|milli|sec|date"
    )
  )

cat("\nCandidate UID/time columns:\n")
print(candidate_columns)

dbDisconnect(con)

```

## Find which SQLite table contains the Tukey UIDs

```{r}

con <- dbConnect(RSQLite::SQLite(), test_db)

# Use first event's Tukey UIDs as test IDs
test_uids <- head(as.numeric(combined_tukey[[1]]$UID), 20)
uid_list_sql <- paste(test_uids, collapse = ",")

cat("Testing these Tukey UIDs:\n")
print(test_uids)

# Tables that contain a UID column
uid_tables <- table_column_summary %>%
  filter(Column == "UID") %>%
  pull(Table) %>%
  unique()

cat("\nTables with UID column:\n")
print(uid_tables)

uid_match_summary <- list()

for (tbl in uid_tables) {
  
  query <- paste0(
    "SELECT COUNT(*) AS n_matches FROM `", tbl, 
    "` WHERE UID IN (", uid_list_sql, ")"
  )
  
  n_matches <- dbGetQuery(con, query)$n_matches
  
  uid_match_summary[[tbl]] <- data.frame(
    Table = tbl,
    n_matches = n_matches
  )
}

uid_match_summary <- bind_rows(uid_match_summary)

cat("\nUID match summary:\n")
print(uid_match_summary)

dbDisconnect(con)

```

## Preview matching rows from likely click tables

```{r}

con <- dbConnect(RSQLite::SQLite(), test_db)

matching_tables <- uid_match_summary %>%
  filter(n_matches > 0) %>%
  pull(Table)

for (tbl in matching_tables) {
  
  cat("\n============================\n")
  cat("MATCHING TABLE:", tbl, "\n")
  cat("============================\n")
  
  query <- paste0(
    "SELECT * FROM `", tbl, 
    "` WHERE UID IN (", uid_list_sql, ") LIMIT 10"
  )
  
  preview <- dbGetQuery(con, query)
  
  print(preview)
}

dbDisconnect(con)

```

## Inspect timing fields from the matching table

```{r}

con <- dbConnect(RSQLite::SQLite(), test_db)

# If only one table matched, use that.
# If multiple matched, this will show us all of them.
matching_tables <- uid_match_summary %>%
  filter(n_matches > 0) %>%
  pull(Table)

print(matching_tables)

for (tbl in matching_tables) {
  
  cat("\n============================\n")
  cat("TIMING FIELD CHECK:", tbl, "\n")
  cat("============================\n")
  
  table_fields <- dbListFields(con, tbl)
  print(table_fields)
  
  timing_cols <- intersect(
    c("Id", "UID", "UTC", "UTCMilliseconds", "StartSample", "Sample", "ChannelBitmap"),
    table_fields
  )
  
  query <- paste0(
    "SELECT ",
    paste(paste0("`", timing_cols, "`"), collapse = ", "),
    " FROM `", tbl, "` WHERE UID IN (",
    uid_list_sql,
    ") ORDER BY UID LIMIT 20"
  )
  
  timing_preview <- dbGetQuery(con, query)
  
  print(timing_preview)
  
  cat("\nStructure:\n")
  print(str(timing_preview))
}

dbDisconnect(con)

```

## Force-print UTCMilliseconds for matching UIDs

```{r}

con <- dbConnect(RSQLite::SQLite(), test_db)

matching_tables <- uid_match_summary %>%
  filter(n_matches > 0) %>%
  pull(Table)

for (tbl in matching_tables) {
  
  cat("\n============================\n")
  cat("TABLE:", tbl, "\n")
  cat("============================\n")
  
  table_fields <- dbListFields(con, tbl)
  
  cat("Fields in this table:\n")
  print(table_fields)
  
  if ("UTCMilliseconds" %in% table_fields) {
    
    query <- paste0(
      "SELECT UID, UTC, UTCMilliseconds ",
      "FROM `", tbl, "` ",
      "WHERE UID IN (", uid_list_sql, ") ",
      "ORDER BY UTCMilliseconds ",
      "LIMIT 20"
    )
    
    timing_preview <- dbGetQuery(con, query)
    
    cat("\nTiming preview using UTCMilliseconds:\n")
    print(timing_preview)
    
    cat("\nColumn names:\n")
    print(names(timing_preview))
    
    cat("\nStructure:\n")
    str(timing_preview)
    
    cat("\nDifferences between consecutive UTCMilliseconds:\n")
    print(diff(timing_preview$UTCMilliseconds))
    
  } else {
    
    cat("No UTCMilliseconds column found in this table.\n")
    
    query <- paste0(
      "SELECT UID, UTC ",
      "FROM `", tbl, "` ",
      "WHERE UID IN (", uid_list_sql, ") ",
      "ORDER BY UID ",
      "LIMIT 20"
    )
    
    timing_preview <- dbGetQuery(con, query)
    print(timing_preview)
  }
}

dbDisconnect(con)

```

## Test-clean PAMGuard UTC strings

```{r}

clean_pamguard_utc <- function(x) {
  
  # Some UTC strings contain ".character(0)" before the milliseconds.
  # Example:
  # "2021-07-30 12:28:29.character(0)187"
  # should become:
  # "2021-07-30 12:28:29.187"
  
  x_clean <- gsub("character\\(0\\)", "", x)
  
  # If a row has no decimal milliseconds, add .000
  has_decimal <- grepl("\\.\\d+$", x_clean)
  x_clean[!has_decimal] <- paste0(x_clean[!has_decimal], ".000")
  
  as.POSIXct(
    x_clean,
    format = "%Y-%m-%d %H:%M:%OS",
    tz = "UTC"
  )
}

# -----------------------------
# Test parsed UTC ordering
# -----------------------------

con <- dbConnect(RSQLite::SQLite(), test_db)

timing_test <- dbGetQuery(
  con,
  paste0(
    "SELECT Id, UID, UTC, UTCMilliseconds, EventId, ClickNo, parentID, parentUID ",
    "FROM Click_Detector_OfflineClicks ",
    "WHERE UID IN (", uid_list_sql, ") ",
    "ORDER BY Id ",
    "LIMIT 20"
  )
)

dbDisconnect(con)

timing_test <- timing_test %>%
  mutate(
    UTC_clean = gsub("character\\(0\\)", "", UTC),
    UTC_parsed = clean_pamguard_utc(UTC)
  ) %>%
  arrange(UTC_parsed, UID) %>%
  mutate(
    ICI_ms_test = as.numeric(
      difftime(lead(UTC_parsed), UTC_parsed, units = "secs")
    ) * 1000
  )

print(timing_test)

cat("\nStructure of parsed UTC:\n")
str(timing_test$UTC_parsed)

cat("\nTest ICI values in milliseconds:\n")
print(timing_test$ICI_ms_test)

```

## Add DriftID to the Tukey click metadata

```{r}

click_metadata_from_tukey <- click_metadata_from_tukey %>%
  mutate(
    DriftID = sub("_Event_.*$", "", EventID),
    UID = as.character(UID)
  )

cat("Unique DriftIDs in combined_tukey:\n")
print(sort(unique(click_metadata_from_tukey$DriftID)))

cat("\nNumber of clicks by DriftID:\n")
clicks_by_drift <- click_metadata_from_tukey %>%
  count(DriftID, name = "n_tukey_clicks") %>%
  arrange(DriftID)

print(clicks_by_drift)

cat("\nTotal Tukey clicks:\n")
print(sum(clicks_by_drift$n_tukey_clicks))

```

## Extract click timing from all matching SQLite databases

```{r}

sqlite_dir <- "E:/Analysis/Useable Databases"

extract_timing_for_drift <- function(drift_id, uid_vec, sqlite_dir) {
  
  db_file <- file.path(sqlite_dir, paste0(drift_id, " - Clean.sqlite3"))
  
  if (!file.exists(db_file)) {
    warning("Missing SQLite database for ", drift_id, ": ", db_file)
    return(NULL)
  }
  
  cat("\n============================\n")
  cat("Extracting timing for", drift_id, "\n")
  cat("Database:", db_file, "\n")
  cat("Number of Tukey UIDs:", length(uid_vec), "\n")
  cat("============================\n")
  
  con <- dbConnect(RSQLite::SQLite(), db_file)
  
  uid_vec <- unique(as.numeric(uid_vec))
  uid_vec <- uid_vec[!is.na(uid_vec)]
  
  # Query in chunks to avoid giant SQL queries
  chunk_size <- 1000
  uid_chunks <- split(uid_vec, ceiling(seq_along(uid_vec) / chunk_size))
  
  timing_chunks <- list()
  
  for (j in seq_along(uid_chunks)) {
    
    uid_sql <- paste(uid_chunks[[j]], collapse = ",")
    
    query <- paste0(
      "SELECT ",
      "Id, UID, UTC, UTCMilliseconds, EventId, ClickNo, parentID, parentUID, BinaryFile, Channels ",
      "FROM Click_Detector_OfflineClicks ",
      "WHERE UID IN (", uid_sql, ")"
    )
    
    timing_chunks[[j]] <- dbGetQuery(con, query)
  }
  
  dbDisconnect(con)
  
  timing_df <- bind_rows(timing_chunks)
  
  if (nrow(timing_df) == 0) {
    warning("No matching timing rows found for ", drift_id)
    return(NULL)
  }
  
  timing_df <- timing_df %>%
    mutate(
      DriftID = drift_id,
      UID = as.character(UID),
      UTC_clean = gsub("character\\(0\\)", "", UTC),
      UTC_parsed = clean_pamguard_utc(UTC)
    )
  
  cat("Rows extracted:", nrow(timing_df), "\n")
  cat("Rows with parsed UTC missing:", sum(is.na(timing_df$UTC_parsed)), "\n")
  
  return(timing_df)
}

# Split Tukey UIDs by DriftID
uids_by_drift <- split(
  click_metadata_from_tukey$UID,
  click_metadata_from_tukey$DriftID
)

all_timing_list <- list()

for (drift_id in names(uids_by_drift)) {
  
  all_timing_list[[drift_id]] <- extract_timing_for_drift(
    drift_id = drift_id,
    uid_vec = uids_by_drift[[drift_id]],
    sqlite_dir = sqlite_dir
  )
}

click_timing_from_sqlite <- bind_rows(all_timing_list)

cat("\n============================\n")
cat("Timing extraction complete\n")
cat("============================\n")

cat("Total timing rows extracted:", nrow(click_timing_from_sqlite), "\n")
cat("Unique DriftIDs extracted:", length(unique(click_timing_from_sqlite$DriftID)), "\n")
cat("Rows with missing parsed UTC:", sum(is.na(click_timing_from_sqlite$UTC_parsed)), "\n")

timing_by_drift <- click_timing_from_sqlite %>%
  count(DriftID, name = "n_timing_rows") %>%
  arrange(DriftID)

print(timing_by_drift)

```

## Check timing join coverage

```{r}

click_metadata_with_time <- click_metadata_from_tukey %>%
  left_join(
    click_timing_from_sqlite %>%
      select(
        DriftID,
        UID,
        SQL_Id = Id,
        UTC,
        UTC_clean,
        UTC_parsed,
        UTCMilliseconds,
        SQL_EventId = EventId,
        SQL_ClickNo = ClickNo,
        parentID,
        parentUID,
        BinaryFile,
        Channels
      ),
    by = c("DriftID", "UID")
  )

cat("Rows in Tukey click metadata:", nrow(click_metadata_from_tukey), "\n")
cat("Rows after timing join:", nrow(click_metadata_with_time), "\n")
cat("Rows missing UTC_parsed:", sum(is.na(click_metadata_with_time$UTC_parsed)), "\n")

join_coverage_by_drift <- click_metadata_with_time %>%
  group_by(DriftID) %>%
  summarise(
    n_tukey_clicks = n(),
    n_with_time = sum(!is.na(UTC_parsed)),
    n_missing_time = sum(is.na(UTC_parsed)),
    percent_with_time = 100 * n_with_time / n_tukey_clicks,
    .groups = "drop"
  ) %>%
  arrange(DriftID)

print(join_coverage_by_drift)

cat("\nDuplicate DriftID + UID rows after join:\n")
duplicate_join_rows <- click_metadata_with_time %>%
  count(DriftID, UID) %>%
  filter(n > 1)

print(duplicate_join_rows)

cat("\nExample joined rows:\n")
click_metadata_with_time %>%
  select(EventID, DriftID, ClickIndexWithinEvent, UID, UTC_clean, UTC_parsed, SQL_EventId, SQL_ClickNo) %>%
  head(20) %>%
  print()

```

## Diagnose duplicate UIDs across Tukey events

```{r}

duplicate_uid_summary <- click_metadata_from_tukey %>%
  count(DriftID, UID, name = "n_tukey_rows") %>%
  filter(n_tukey_rows > 1) %>%
  arrange(DriftID, UID)

cat("Number of duplicated DriftID + UID combinations:\n")
print(nrow(duplicate_uid_summary))

cat("\nDuplicated UID combinations by DriftID:\n")
duplicate_uid_by_drift <- duplicate_uid_summary %>%
  group_by(DriftID) %>%
  summarise(
    n_duplicated_uids = n(),
    max_duplicate_count = max(n_tukey_rows),
    .groups = "drop"
  )

print(duplicate_uid_by_drift)

cat("\nWhich EventIDs share duplicated UIDs?\n")

duplicate_uid_event_map <- click_metadata_from_tukey %>%
  semi_join(
    duplicate_uid_summary,
    by = c("DriftID", "UID")
  ) %>%
  group_by(DriftID, UID) %>%
  summarise(
    EventIDs = paste(sort(unique(EventID)), collapse = "; "),
    n_event_ids = n_distinct(EventID),
    n_rows = n(),
    .groups = "drop"
  ) %>%
  arrange(DriftID, UID)

print(head(duplicate_uid_event_map, 30))

cat("\nSummary of duplicated UID sharing patterns:\n")
duplicate_sharing_summary <- duplicate_uid_event_map %>%
  count(DriftID, EventIDs, n_event_ids, name = "n_shared_uids") %>%
  arrange(DriftID, desc(n_shared_uids))

print(duplicate_sharing_summary)

# Save diagnostics
ici_output_dir <- file.path(base_output_dir, analysis_name, "ICI_Diagnostics")
dir.create(ici_output_dir, recursive = TRUE, showWarnings = FALSE)

write.csv(
  duplicate_uid_summary,
  file = file.path(ici_output_dir, "duplicate_uid_summary.csv"),
  row.names = FALSE
)

write.csv(
  duplicate_uid_event_map,
  file = file.path(ici_output_dir, "duplicate_uid_event_map.csv"),
  row.names = FALSE
)

write.csv(
  duplicate_sharing_summary,
  file = file.path(ici_output_dir, "duplicate_uid_sharing_summary.csv"),
  row.names = FALSE
)

```







