---
title: "ICI_MinMax_kmeans_Clustering"
author: "Sarah Moseley"
date: "2026-05-05"
output: html_document
---

## Setup

```{r setup}

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)

required_packages <- c(
  "dplyr",
  "ggplot2",
  "tidyr",
  "stringr",
  "readr",
  "DBI",
  "RSQLite"
)

missing_packages <- required_packages[
  !sapply(required_packages, requireNamespace, quietly = TRUE)
]

if (length(missing_packages) > 0) {
  stop(
    "Missing required packages: ",
    paste(missing_packages, collapse = ", "),
    "\nInstall them before running this workflow."
  )
}

library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
library(readr)
library(DBI)
library(RSQLite)

```

## Use settings

```{r user-settings}

# This should point to the FIXED combined Tukey object,
# where ADRIFT_003_Event_001 has already been corrected.
combined_file <- file.choose()

# SQLite databases containing PAMGuard click timing information
sqlite_dir <- "E:/Analysis/Useable Databases"

# Output folder for ICI-specific products
ici_output_dir <- "E:/Analysis/ICI_Output"

dir.create(ici_output_dir, recursive = TRUE, showWarnings = FALSE)

# Expected total clicks after fixing ADRIFT_003_Event_001
expected_total_clicks_new <- 344153

# Used later only for flagging suspicious intervals.
# Do not delete based on this yet.
max_reasonable_ici_ms <- 1000

```

## Load Fixed Combined Tukey

```{r load-fixed-combined-tukey}

combined_tukey <- readRDS(combined_file)

cat("Loaded combined_tukey object.\n")
cat("Number of events:", length(combined_tukey), "\n")

cat("\nClass of combined_tukey:\n")
print(class(combined_tukey))

cat("\nClass of first event object:\n")
print(class(combined_tukey[[1]]))

cat("\nNames in first event object:\n")
print(names(combined_tukey[[1]]))

```

## Santiy Check: Combined Structure

```{r sanity-check-combined-structure}

event_structure_summary <- data.frame(
  EventIndex = seq_along(combined_tukey),
  EventID = sapply(combined_tukey, function(x) x$EventID),
  n_freq = sapply(combined_tukey, function(x) length(x$freq)),
  n_uid = sapply(combined_tukey, function(x) length(x$UID)),
  allSpec_rows = sapply(combined_tukey, function(x) dim(x$allSpec)[1]),
  allSpec_cols = sapply(combined_tukey, function(x) dim(x$allSpec)[2]),
  allNoise_rows = sapply(combined_tukey, function(x) dim(x$allNoise)[1]),
  allNoise_cols = sapply(combined_tukey, function(x) dim(x$allNoise)[2]),
  snrKeep_length = sapply(combined_tukey, function(x) length(x$snrKeep)),
  sourcefile_length = sapply(combined_tukey, function(x) length(x$SourceFile)),
  stringsAsFactors = FALSE
)

print(event_structure_summary)

total_clicks_new <- sum(event_structure_summary$allSpec_cols)

cat("\nTotal clicks in fixed combined_tukey:", total_clicks_new, "\n")
cat("Expected total clicks:", expected_total_clicks_new, "\n")

if (total_clicks_new == expected_total_clicks_new) {
  cat("\nPASS: Total click count matches expected fixed total.\n")
} else {
  cat("\nFAIL: Total click count does not match expected fixed total.\n")
}

uid_matches_allSpec <- all(event_structure_summary$n_uid == event_structure_summary$allSpec_cols)
allNoise_matches_allSpec <- all(event_structure_summary$allNoise_cols == event_structure_summary$allSpec_cols)
snrKeep_matches_allSpec <- all(event_structure_summary$snrKeep_length == event_structure_summary$allSpec_cols)

cat("\nUID count matches allSpec columns:", uid_matches_allSpec, "\n")
cat("allNoise columns match allSpec columns:", allNoise_matches_allSpec, "\n")
cat("snrKeep length matches allSpec columns:", snrKeep_matches_allSpec, "\n")

if (uid_matches_allSpec && allNoise_matches_allSpec && snrKeep_matches_allSpec) {
  cat("\nPASS: Core click-level dimensions line up across all events.\n")
} else {
  cat("\nFAIL: At least one event has mismatched click-level dimensions.\n")
}

```

## Duplicate Events Sanity Check

```{r sanity-check-duplicate-events-and-uids}

# ------------------------------------------------------------
# Build temporary click-level metadata for duplicate checks
# ------------------------------------------------------------

click_metadata_from_tukey <- list()

for (i in seq_along(combined_tukey)) {
  
  event_obj <- combined_tukey[[i]]
  n_clicks <- ncol(event_obj$allSpec)
  
  click_metadata_from_tukey[[i]] <- data.frame(
    EventIndex = i,
    EventID = event_obj$EventID,
    ClickIndexWithinEvent = seq_len(n_clicks),
    UID = as.character(event_obj$UID),
    SourceFile = event_obj$SourceFile,
    stringsAsFactors = FALSE
  )
}

click_metadata_from_tukey <- bind_rows(click_metadata_from_tukey) %>%
  mutate(
    DriftID = sub("_Event_.*$", "", EventID)
  )

cat("Total click metadata rows:", nrow(click_metadata_from_tukey), "\n")
cat("Expected total clicks:", expected_total_clicks_new, "\n\n")

# ------------------------------------------------------------
# 1. Check duplicated EventIDs
# ------------------------------------------------------------

duplicate_event_ids <- click_metadata_from_tukey %>%
  distinct(EventIndex, EventID) %>%
  count(EventID, name = "n_event_indices") %>%
  filter(n_event_indices > 1)

cat("Duplicated EventIDs:\n")
print(duplicate_event_ids)

if (nrow(duplicate_event_ids) == 0) {
  cat("\nPASS: No duplicated EventIDs.\n")
} else {
  cat("\nFAIL: Some EventIDs appear more than once.\n")
}

# ------------------------------------------------------------
# 2. Check duplicated UID within the same EventID
# ------------------------------------------------------------

duplicate_uid_within_event <- click_metadata_from_tukey %>%
  count(EventID, UID, name = "n_rows") %>%
  filter(n_rows > 1) %>%
  arrange(EventID, UID)

cat("\nDuplicated UID values within the same EventID:\n")
print(duplicate_uid_within_event)

if (nrow(duplicate_uid_within_event) == 0) {
  cat("\nPASS: No duplicated UIDs within any EventID.\n")
} else {
  cat("\nFAIL: Some UIDs are duplicated within the same EventID.\n")
}

# ------------------------------------------------------------
# 3. Check duplicated UID across events within the same DriftID
# ------------------------------------------------------------

duplicate_uid_across_events_same_drift <- click_metadata_from_tukey %>%
  group_by(DriftID, UID) %>%
  summarise(
    n_rows = n(),
    n_event_ids = n_distinct(EventID),
    EventIDs = paste(sort(unique(EventID)), collapse = "; "),
    .groups = "drop"
  ) %>%
  filter(n_rows > 1 | n_event_ids > 1) %>%
  arrange(DriftID, UID)

cat("\nDuplicated DriftID + UID combinations across Tukey metadata:\n")
print(duplicate_uid_across_events_same_drift)

if (nrow(duplicate_uid_across_events_same_drift) == 0) {
  cat("\nPASS: No duplicated DriftID + UID combinations across events.\n")
} else {
  cat("\nFAIL: Some DriftID + UID combinations appear more than once.\n")
}

# ------------------------------------------------------------
# 4. Save duplicate diagnostics
# ------------------------------------------------------------

write.csv(
  duplicate_event_ids,
  file = file.path(ici_output_dir, "duplicate_event_ids_NEW.csv"),
  row.names = FALSE
)

write.csv(
  duplicate_uid_within_event,
  file = file.path(ici_output_dir, "duplicate_uid_within_event_NEW.csv"),
  row.names = FALSE
)

write.csv(
  duplicate_uid_across_events_same_drift,
  file = file.path(ici_output_dir, "duplicate_uid_across_events_same_drift_NEW.csv"),
  row.names = FALSE
)

```

## Save Clean Click-Level Tukey Metadata

```{r save-clean-click-metadata}

cat("Clean Tukey click metadata summary:\n\n")

cat("Rows:", nrow(click_metadata_from_tukey), "\n")
cat("Unique EventIDs:", n_distinct(click_metadata_from_tukey$EventID), "\n")
cat("Unique DriftIDs:", n_distinct(click_metadata_from_tukey$DriftID), "\n")
cat("Unique DriftID + UID pairs:",
    nrow(distinct(click_metadata_from_tukey, DriftID, UID)), "\n")

cat("\nClicks by DriftID:\n")

clicks_by_drift <- click_metadata_from_tukey %>%
  count(DriftID, name = "n_tukey_clicks") %>%
  arrange(DriftID)

print(clicks_by_drift)

cat("\nClicks by EventID:\n")

clicks_by_event <- click_metadata_from_tukey %>%
  count(DriftID, EventID, name = "n_tukey_clicks") %>%
  arrange(DriftID, EventID)

print(clicks_by_event)

# Save clean metadata
saveRDS(
  click_metadata_from_tukey,
  file = file.path(ici_output_dir, "click_metadata_from_tukey_NEW.rds")
)

write.csv(
  click_metadata_from_tukey,
  file = file.path(ici_output_dir, "click_metadata_from_tukey_NEW.csv"),
  row.names = FALSE
)

write.csv(
  clicks_by_drift,
  file = file.path(ici_output_dir, "clicks_by_drift_NEW.csv"),
  row.names = FALSE
)

write.csv(
  clicks_by_event,
  file = file.path(ici_output_dir, "clicks_by_event_NEW.csv"),
  row.names = FALSE
)

cat("\nSaved clean Tukey metadata outputs to:\n")
cat(ici_output_dir, "\n")

```

## Define UTC parser and extract click timing from SQLite databases

```{r define-utc-parser}

# ------------------------------------------------------------
# Function to clean and parse PAMGuard UTC strings
# ------------------------------------------------------------

clean_pamguard_utc <- function(x) {
  
  x <- as.character(x)
  
  # Some UTC strings contain ".character(0)" before milliseconds.
  # Example:
  # "2021-07-30 12:28:29.character(0)187"
  # should become:
  # "2021-07-30 12:28:29.187"
  x_clean <- gsub("character\\(0\\)", "", x)
  
  # If no decimal milliseconds are present, add .000
  has_decimal <- grepl("\\.\\d+$", x_clean)
  x_clean[!has_decimal & !is.na(x_clean)] <- paste0(x_clean[!has_decimal & !is.na(x_clean)], ".000")
  
  parsed_time <- as.POSIXct(
    x_clean,
    format = "%Y-%m-%d %H:%M:%OS",
    tz = "UTC"
  )
  
  return(parsed_time)
}

# ------------------------------------------------------------
# Function to extract timing rows for one drift
# ------------------------------------------------------------

extract_timing_for_drift <- function(drift_id, uid_vec, sqlite_dir) {
  
  db_file <- file.path(sqlite_dir, paste0(drift_id, " - Clean.sqlite3"))
  
  if (!file.exists(db_file)) {
    warning("Missing SQLite database for ", drift_id, ": ", db_file)
    return(NULL)
  }
  
  uid_vec <- unique(suppressWarnings(as.numeric(uid_vec)))
  uid_vec <- uid_vec[!is.na(uid_vec)]
  
  if (length(uid_vec) == 0) {
    warning("No numeric UIDs available for ", drift_id)
    return(NULL)
  }
  
  cat("\n============================\n")
  cat("Extracting timing for:", drift_id, "\n")
  cat("Database:", db_file, "\n")
  cat("Number of unique Tukey UIDs:", length(uid_vec), "\n")
  cat("============================\n")
  
  con <- dbConnect(RSQLite::SQLite(), db_file)
  on.exit(dbDisconnect(con), add = TRUE)
  
  # Query in chunks to avoid huge SQL statements
  chunk_size <- 1000
  uid_chunks <- split(uid_vec, ceiling(seq_along(uid_vec) / chunk_size))
  
  timing_chunks <- vector("list", length(uid_chunks))
  
  for (j in seq_along(uid_chunks)) {
    
    uid_sql <- paste(uid_chunks[[j]], collapse = ",")
    
    query <- paste0(
      "SELECT ",
      "Id, UID, UTC, UTCMilliseconds, EventId, ClickNo, parentID, parentUID, BinaryFile, Channels ",
      "FROM Click_Detector_OfflineClicks ",
      "WHERE UID IN (", uid_sql, ")"
    )
    
    timing_chunks[[j]] <- dbGetQuery(con, query)
  }
  
  timing_df <- bind_rows(timing_chunks)
  
  if (nrow(timing_df) == 0) {
    warning("No matching timing rows found for ", drift_id)
    return(NULL)
  }
  
  timing_df <- timing_df %>%
    mutate(
      DriftID = drift_id,
      UID = as.character(UID),
      UTC_clean = gsub("character\\(0\\)", "", as.character(UTC)),
      UTC_parsed = clean_pamguard_utc(UTC)
    )
  
  cat("Rows extracted:", nrow(timing_df), "\n")
  cat("Rows with missing parsed UTC:", sum(is.na(timing_df$UTC_parsed)), "\n")
  
  return(timing_df)
}

```

## Extract timing

```{r extract-timing-from-all-sqlite-databases}

if (!dir.exists(sqlite_dir)) {
  stop("SQLite directory does not exist: ", sqlite_dir)
}

uids_by_drift <- split(
  click_metadata_from_tukey$UID,
  click_metadata_from_tukey$DriftID
)

cat("Number of DriftIDs to extract:", length(uids_by_drift), "\n")
cat("\nDriftIDs:\n")
print(names(uids_by_drift))

all_timing_list <- list()

 for (drift_id in names(uids_by_drift)) {
  
  all_timing_list[[drift_id]] <- extract_timing_for_drift(
    drift_id = drift_id,
    uid_vec = uids_by_drift[[drift_id]],
    sqlite_dir = sqlite_dir
  )
}

click_timing_from_sqlite <- bind_rows(all_timing_list)

cat("\n============================\n")
cat("Timing extraction complete\n")
cat("============================\n")

cat("Total timing rows extracted:", nrow(click_timing_from_sqlite), "\n")
cat("Unique DriftIDs extracted:", n_distinct(click_timing_from_sqlite$DriftID), "\n")
cat("Rows with missing parsed UTC:", sum(is.na(click_timing_from_sqlite$UTC_parsed)), "\n")

timing_by_drift <- click_timing_from_sqlite %>%
  count(DriftID, name = "n_timing_rows") %>%
  arrange(DriftID)

print(timing_by_drift)

# Save raw extracted timing table
saveRDS(
  click_timing_from_sqlite,
  file = file.path(ici_output_dir, "click_timing_from_sqlite_NEW.rds")
)

write.csv(
  timing_by_drift,
  file = file.path(ici_output_dir, "timing_rows_by_drift_NEW.csv"),
  row.names = FALSE
)

cat("\nSaved timing extraction outputs to:\n")
cat(ici_output_dir, "\n")

```

## Diagnose timing uniqueness and join timing to Tukey metadata

```{r diagnose-timing-uniqueness-before-join}

cat("Rows in Tukey click metadata:", nrow(click_metadata_from_tukey), "\n")
cat("Rows in extracted SQLite timing:", nrow(click_timing_from_sqlite), "\n\n")

duplicate_timing_rows <- click_timing_from_sqlite %>%
  count(DriftID, UID, name = "n_sql_rows") %>%
  filter(n_sql_rows > 1) %>%
  arrange(DriftID, UID)

cat("Duplicated DriftID + UID rows in SQLite timing table:\n")
print(duplicate_timing_rows)

if (nrow(duplicate_timing_rows) == 0) {
  cat("\nPASS: SQLite timing has one row per DriftID + UID.\n")
} else {
  cat("\nFAIL: SQLite timing has duplicated DriftID + UID rows.\n")
  cat("Do not join yet. Need to inspect duplicates first.\n")
}

write.csv(
  duplicate_timing_rows,
  file = file.path(ici_output_dir, "duplicate_timing_rows_sqlite_NEW.csv"),
  row.names = FALSE
)

```

##  Join clean Tukey metadata to SQLite timing

```{r join-tukey-metadata-to-sqlite-timing}

if (nrow(duplicate_timing_rows) > 0) {
  stop("Stopping because SQLite timing contains duplicated DriftID + UID rows.")
}

click_metadata_with_time <- click_metadata_from_tukey %>%
  left_join(
    click_timing_from_sqlite %>%
      select(
        DriftID,
        UID,
        SQL_Id = Id,
        UTC,
        UTC_clean,
        UTC_parsed,
        UTCMilliseconds,
        SQL_EventId = EventId,
        SQL_ClickNo = ClickNo,
        parentID,
        parentUID,
        BinaryFile,
        Channels
      ),
    by = c("DriftID", "UID")
  )

cat("Rows before join:", nrow(click_metadata_from_tukey), "\n")
cat("Rows after join:", nrow(click_metadata_with_time), "\n")

if (nrow(click_metadata_with_time) == nrow(click_metadata_from_tukey)) {
  cat("\nPASS: Join did not change row count.\n")
} else {
  cat("\nFAIL: Join changed row count.\n")
}

cat("\nRows missing UTC_parsed:", sum(is.na(click_metadata_with_time$UTC_parsed)), "\n")
cat("Rows with UTC_parsed:", sum(!is.na(click_metadata_with_time$UTC_parsed)), "\n")

join_coverage_by_drift <- click_metadata_with_time %>%
  group_by(DriftID) %>%
  summarise(
    n_tukey_clicks = n(),
    n_with_time = sum(!is.na(UTC_parsed)),
    n_missing_time = sum(is.na(UTC_parsed)),
    percent_with_time = 100 * n_with_time / n_tukey_clicks,
    .groups = "drop"
  ) %>%
  arrange(DriftID)

cat("\nJoin coverage by DriftID:\n")
print(join_coverage_by_drift)

join_coverage_by_event <- click_metadata_with_time %>%
  group_by(DriftID, EventID) %>%
  summarise(
    n_tukey_clicks = n(),
    n_with_time = sum(!is.na(UTC_parsed)),
    n_missing_time = sum(is.na(UTC_parsed)),
    percent_with_time = 100 * n_with_time / n_tukey_clicks,
    .groups = "drop"
  ) %>%
  arrange(DriftID, EventID)

cat("\nJoin coverage by EventID:\n")
print(join_coverage_by_event)

# Save joined timing metadata
saveRDS(
  click_metadata_with_time,
  file = file.path(ici_output_dir, "click_metadata_with_time_NEW.rds")
)

write.csv(
  join_coverage_by_drift,
  file = file.path(ici_output_dir, "join_coverage_by_drift_NEW.csv"),
  row.names = FALSE
)

write.csv(
  join_coverage_by_event,
  file = file.path(ici_output_dir, "join_coverage_by_event_NEW.csv"),
  row.names = FALSE
)

cat("\nSaved joined timing metadata outputs to:\n")
cat(ici_output_dir, "\n")

```

## Check whether multiple clicks share the exact same timestamp within the same Tukey EventID

```{r diagnose-timestamp-duplicates-before-ici}

timestamp_duplicates_within_event <- click_metadata_with_time %>%
  count(EventID, UTC_parsed, name = "n_clicks_same_timestamp") %>%
  filter(n_clicks_same_timestamp > 1) %>%
  arrange(EventID, UTC_parsed)

cat("Number of EventID + timestamp combinations with multiple clicks:\n")
print(nrow(timestamp_duplicates_within_event))

cat("\nPreview of duplicated timestamps within events:\n")
print(head(timestamp_duplicates_within_event, 30))

write.csv(
  timestamp_duplicates_within_event,
  file = file.path(ici_output_dir, "timestamp_duplicates_within_event_NEW.csv"),
  row.names = FALSE
)

if (nrow(timestamp_duplicates_within_event) == 0) {
  cat("\nPASS: No duplicated timestamps within events.\n")
} else {
  cat("\nNOTE: Some clicks share exact timestamps within events.\n")
  cat("This is not automatically wrong, but it may produce zero-ms ICI values.\n")
}

```

## Summarize duplicated timestamps within events

```{r}

timestamp_duplicate_summary <- timestamp_duplicates_within_event %>%
  summarise(
    n_duplicate_timestamp_bins = n(),
    n_clicks_in_duplicate_timestamp_bins = sum(n_clicks_same_timestamp),
    n_extra_clicks_caused_by_timestamp_ties = sum(n_clicks_same_timestamp - 1),
    max_clicks_same_timestamp = max(n_clicks_same_timestamp),
    .groups = "drop"
  ) %>%
  mutate(
    total_clicks = nrow(click_metadata_with_time),
    percent_clicks_in_duplicate_timestamp_bins =
      100 * n_clicks_in_duplicate_timestamp_bins / total_clicks,
    percent_extra_clicks_due_to_ties =
      100 * n_extra_clicks_caused_by_timestamp_ties / total_clicks
  )

cat("Duplicated timestamp summary:\n")
print(timestamp_duplicate_summary)

timestamp_duplicates_by_event <- timestamp_duplicates_within_event %>%
  group_by(EventID) %>%
  summarise(
    n_duplicate_timestamp_bins = n(),
    n_clicks_in_duplicate_timestamp_bins = sum(n_clicks_same_timestamp),
    n_extra_clicks_caused_by_timestamp_ties = sum(n_clicks_same_timestamp - 1),
    max_clicks_same_timestamp = max(n_clicks_same_timestamp),
    .groups = "drop"
  ) %>%
  arrange(desc(n_extra_clicks_caused_by_timestamp_ties))

cat("\nEvents with the most duplicated timestamps:\n")
print(head(timestamp_duplicates_by_event, 20))

write.csv(
  timestamp_duplicate_summary,
  file = file.path(ici_output_dir, "timestamp_duplicate_summary_NEW.csv"),
  row.names = FALSE
)

write.csv(
  timestamp_duplicates_by_event,
  file = file.path(ici_output_dir, "timestamp_duplicates_by_event_NEW.csv"),
  row.names = FALSE
)

```

## Check UTC Clean VS Parsed Duplicates

```{r check-utc-clean-vs-parsed-duplicates}

# Show milliseconds when printing POSIXct
options(digits.secs = 6)

# Compare duplicate counts using UTC_clean string vs UTC_parsed time object

duplicates_by_utc_clean <- click_metadata_with_time %>%
  count(EventID, UTC_clean, name = "n_clicks_same_utc_clean") %>%
  filter(n_clicks_same_utc_clean > 1)

duplicates_by_utc_parsed <- click_metadata_with_time %>%
  count(EventID, UTC_parsed, name = "n_clicks_same_utc_parsed") %>%
  filter(n_clicks_same_utc_parsed > 1)

cat("Duplicate EventID + UTC_clean combinations:\n")
print(nrow(duplicates_by_utc_clean))

cat("\nDuplicate EventID + UTC_parsed combinations:\n")
print(nrow(duplicates_by_utc_parsed))

cat("\nPreview UTC_clean duplicates:\n")
print(head(duplicates_by_utc_clean, 20))

cat("\nPreview UTC_parsed duplicates:\n")
print(head(duplicates_by_utc_parsed, 20))

```

## Robust UTC string cleaner for malformed PAMGuard UTC values

```{r}

clean_pamguard_utc_string <- function(x) {
  
  x <- as.character(x)
  
  out <- vapply(x, function(s) {
    
    if (is.na(s) || s == "") {
      return(NA_character_)
    }
    
    # Remove known malformed character(0) insertion
    s <- gsub("character\\(0\\)", "", s)
    
    # Extract base date-time through whole seconds
    base_match <- regmatches(
      s,
      regexpr("^\\d{4}-\\d{2}-\\d{2} \\d{2}:\\d{2}:\\d{2}", s)
    )
    
    if (length(base_match) == 0 || base_match == "") {
      return(NA_character_)
    }
    
    base_time <- base_match[1]
    
    # Everything after the base timestamp
    remainder <- sub(
      "^\\d{4}-\\d{2}-\\d{2} \\d{2}:\\d{2}:\\d{2}",
      "",
      s
    )
    
    # Default to .000 if no fractional part exists
    frac_digits <- "000"
    
    if (grepl("^\\.", remainder)) {
      
      frac_raw <- sub("^\\.", "", remainder)
      
      # Case: c("0", "0")0 style malformed fraction
      if (grepl("^c\\(", frac_raw)) {
        
        quoted_digits <- regmatches(
          frac_raw,
          gregexpr('"\\d"', frac_raw)
        )[[1]]
        
        quoted_digits <- gsub('"', "", quoted_digits)
        
        after_c <- sub("^c\\([^)]*\\)", "", frac_raw)
        
        trailing_digits <- regmatches(
          after_c,
          gregexpr("\\d", after_c)
        )[[1]]
        
        if (length(trailing_digits) == 1 && trailing_digits[1] == -1) {
          trailing_digits <- character(0)
        }
        
        frac_digits <- paste0(c(quoted_digits, trailing_digits), collapse = "")
        
      } else {
        
        # Normal case: .127
        frac_digits <- sub("^([0-9]+).*$", "\\1", frac_raw)
        
        if (!grepl("^\\d+$", frac_digits)) {
          frac_digits <- "000"
        }
      }
    }
    
    # Force milliseconds: exactly 3 digits
    frac_digits <- substr(paste0(frac_digits, "000"), 1, 3)
    
    paste0(base_time, ".", frac_digits)
    
  }, character(1))
  
  return(out)
}

parse_clean_utc_to_ms <- function(x_clean) {
  
  base_time <- as.POSIXct(
    sub("\\.\\d+$", "", x_clean),
    format = "%Y-%m-%d %H:%M:%S",
    tz = "UTC"
  )
  
  milliseconds <- as.numeric(sub("^.*\\.", "", x_clean))
  
  as.numeric(base_time) * 1000 + milliseconds
}

click_metadata_with_time <- click_metadata_with_time %>%
  mutate(
    UTC_clean_fixed = clean_pamguard_utc_string(UTC),
    UTC_ms_fixed = parse_clean_utc_to_ms(UTC_clean_fixed)
  )

cat("Rows:", nrow(click_metadata_with_time), "\n")
cat("Missing UTC_clean_fixed:", sum(is.na(click_metadata_with_time$UTC_clean_fixed)), "\n")
cat("Missing UTC_ms_fixed:", sum(is.na(click_metadata_with_time$UTC_ms_fixed)), "\n")

cat("\nExamples of malformed c(...) UTC values after cleaning:\n")

click_metadata_with_time %>%
  filter(grepl("c\\(", UTC)) %>%
  select(UID, UTC, UTC_clean_fixed, UTC_ms_fixed) %>%
  head(20) %>%
  print()


```

## Inspect malformed rows

```{r}

c_malformed_rows <- click_metadata_with_time %>%
  filter(grepl("c\\(", UTC)) %>%
  select(
    EventID,
    ClickIndexWithinEvent,
    UID,
    UTC,
    UTC_clean,
    UTC_parsed,
    UTCMilliseconds,
    UTC_clean_fixed,
    UTC_ms_fixed
  )

cat("Number of c(...) malformed UTC rows:", nrow(c_malformed_rows), "\n\n")

print(head(c_malformed_rows, 30))

```

##  Validate that c(...) malformed UTC rows were cleaned correctly

```{r}

c_malformed_check <- click_metadata_with_time %>%
  filter(grepl("c\\(", UTC)) %>%
  mutate(
    fixed_has_normal_timestamp_format = grepl(
      "^\\d{4}-\\d{2}-\\d{2} \\d{2}:\\d{2}:\\d{2}\\.\\d{3}$",
      UTC_clean_fixed
    )
  )

cat("Number of c(...) malformed UTC rows:", nrow(c_malformed_check), "\n")
cat("Rows with missing UTC_clean_fixed:", sum(is.na(c_malformed_check$UTC_clean_fixed)), "\n")
cat("Rows with missing UTC_ms_fixed:", sum(is.na(c_malformed_check$UTC_ms_fixed)), "\n")
cat("Rows with valid fixed timestamp format:", sum(c_malformed_check$fixed_has_normal_timestamp_format), "\n")
cat("Rows with invalid fixed timestamp format:", sum(!c_malformed_check$fixed_has_normal_timestamp_format), "\n\n")

cat("Preview of fixed c(...) rows:\n")

c_malformed_check %>%
  select(
    EventID,
    ClickIndexWithinEvent,
    UID,
    UTC,
    UTC_clean,
    UTC_parsed,
    UTCMilliseconds,
    UTC_clean_fixed,
    UTC_ms_fixed,
    fixed_has_normal_timestamp_format
  ) %>%
  head(30) %>%
  print()

```

## Create final timing fields for ICI workflow

```{r}

# These are the columns we will use from here forward.
#
# Final_UTC_clean = corrected readable timestamp string
# Final_UTC_ms    = corrected numeric absolute timestamp in milliseconds
# ------------------------------------------------------------

click_metadata_with_time <- click_metadata_with_time %>%
  mutate(
    Final_UTC_clean = UTC_clean_fixed,
    Final_UTC_ms = UTC_ms_fixed
  )

cat("Rows:", nrow(click_metadata_with_time), "\n")
cat("Missing Final_UTC_clean:", sum(is.na(click_metadata_with_time$Final_UTC_clean)), "\n")
cat("Missing Final_UTC_ms:", sum(is.na(click_metadata_with_time$Final_UTC_ms)), "\n")

cat("\nPreview final timing fields:\n")

click_metadata_with_time %>%
  select(
    EventID,
    ClickIndexWithinEvent,
    UID,
    UTC,
    UTC_clean,
    UTC_parsed,
    Final_UTC_clean,
    Final_UTC_ms
  ) %>%
  head(20) %>%
  print()

```

## Compute ICI using final cleaned timestamps

```{r}

# ICI is computed within each Tukey EventID.
# This prevents intervals from crossing between separate click events.
#
# Final_UTC_ms is an absolute timestamp in milliseconds.
# ICI = next click timestamp - current click timestamp
# ------------------------------------------------------------

click_metadata_with_ici <- click_metadata_with_time %>%
  arrange(EventID, Final_UTC_ms, SQL_Id, UID) %>%
  group_by(EventID) %>%
  mutate(
    ClickOrderByTime = row_number(),
    n_clicks_in_event = n(),
    
    Previous_UID = lag(UID),
    Next_UID = lead(UID),
    
    Previous_Final_UTC_clean = lag(Final_UTC_clean),
    Next_Final_UTC_clean = lead(Final_UTC_clean),
    
    Previous_Final_UTC_ms = lag(Final_UTC_ms),
    Next_Final_UTC_ms = lead(Final_UTC_ms),
    
    ICI_from_previous_ms = Final_UTC_ms - Previous_Final_UTC_ms,
    ICI_to_next_ms = Next_Final_UTC_ms - Final_UTC_ms,
    
    IsFirstClickInEvent = ClickOrderByTime == 1,
    IsLastClickInEvent = ClickOrderByTime == n_clicks_in_event
  ) %>%
  ungroup() %>%
  mutate(
    ICI_to_next_flag = case_when(
      is.na(ICI_to_next_ms) & IsLastClickInEvent ~ "last_click_no_next_ici",
      is.na(ICI_to_next_ms) ~ "missing_ici",
      ICI_to_next_ms < 0 ~ "negative_ici",
      ICI_to_next_ms == 0 ~ "zero_ici",
      ICI_to_next_ms > max_reasonable_ici_ms ~ "large_ici",
      TRUE ~ "reasonable_ici"
    ),
    
    ICI_from_previous_flag = case_when(
      is.na(ICI_from_previous_ms) & IsFirstClickInEvent ~ "first_click_no_previous_ici",
      is.na(ICI_from_previous_ms) ~ "missing_ici",
      ICI_from_previous_ms < 0 ~ "negative_ici",
      ICI_from_previous_ms == 0 ~ "zero_ici",
      ICI_from_previous_ms > max_reasonable_ici_ms ~ "large_ici",
      TRUE ~ "reasonable_ici"
    )
  )

cat("Rows in click_metadata_with_ici:", nrow(click_metadata_with_ici), "\n")
cat("Expected rows:", expected_total_clicks_new, "\n")

if (nrow(click_metadata_with_ici) == expected_total_clicks_new) {
  cat("\nPASS: ICI table row count matches expected total clicks.\n")
} else {
  cat("\nFAIL: ICI table row count does not match expected total clicks.\n")
}

cat("\nICI_to_next flag counts:\n")
print(table(click_metadata_with_ici$ICI_to_next_flag, useNA = "ifany"))

cat("\nICI_from_previous flag counts:\n")
print(table(click_metadata_with_ici$ICI_from_previous_flag, useNA = "ifany"))

```

## Basic ICI sanity checks

```{r}

cat("Negative ICI_to_next values:\n")
print(sum(click_metadata_with_ici$ICI_to_next_ms < 0, na.rm = TRUE))

cat("\nZero ICI_to_next values:\n")
print(sum(click_metadata_with_ici$ICI_to_next_ms == 0, na.rm = TRUE))

cat("\nLarge ICI_to_next values greater than max_reasonable_ici_ms:\n")
print(sum(click_metadata_with_ici$ICI_to_next_ms > max_reasonable_ici_ms, na.rm = TRUE))

cat("\nSummary of all non-NA ICI_to_next_ms:\n")
print(summary(click_metadata_with_ici$ICI_to_next_ms))

cat("\nSummary of reasonable ICI_to_next_ms only:\n")
print(
  summary(
    click_metadata_with_ici$ICI_to_next_ms[
      click_metadata_with_ici$ICI_to_next_flag == "reasonable_ici"
    ]
  )
)

cat("\nPreview of computed ICI rows:\n")

click_metadata_with_ici %>%
  select(
    EventID,
    ClickOrderByTime,
    UID,
    Final_UTC_clean,
    Previous_Final_UTC_clean,
    Next_Final_UTC_clean,
    ICI_from_previous_ms,
    ICI_to_next_ms,
    ICI_to_next_flag
  ) %>%
  head(30) %>%
  print()

```

## Event-level ICI summary

```{r}

# Core summaries use reasonable ICI values only:
#   ICI_to_next_flag == "reasonable_ici"
# This excludes:
#   - last click of each event
#   - zero-ms timestamp ties
#   - large gaps > max_reasonable_ici_ms
# ------------------------------------------------------------

ici_summary_by_event <- click_metadata_with_ici %>%
  group_by(DriftID, EventID) %>%
  summarise(
    n_clicks = n(),
    
    n_reasonable_ici = sum(ICI_to_next_flag == "reasonable_ici", na.rm = TRUE),
    n_zero_ici = sum(ICI_to_next_flag == "zero_ici", na.rm = TRUE),
    n_large_ici = sum(ICI_to_next_flag == "large_ici", na.rm = TRUE),
    n_last_click_no_next_ici = sum(ICI_to_next_flag == "last_click_no_next_ici", na.rm = TRUE),
    
    percent_reasonable_ici = 100 * n_reasonable_ici / n_clicks,
    percent_zero_ici = 100 * n_zero_ici / n_clicks,
    percent_large_ici = 100 * n_large_ici / n_clicks,
    
    mean_ici_ms = mean(
      ICI_to_next_ms[ICI_to_next_flag == "reasonable_ici"],
      na.rm = TRUE
    ),
    median_ici_ms = median(
      ICI_to_next_ms[ICI_to_next_flag == "reasonable_ici"],
      na.rm = TRUE
    ),
    sd_ici_ms = sd(
      ICI_to_next_ms[ICI_to_next_flag == "reasonable_ici"],
      na.rm = TRUE
    ),
    min_ici_ms = min(
      ICI_to_next_ms[ICI_to_next_flag == "reasonable_ici"],
      na.rm = TRUE
    ),
    max_ici_ms = max(
      ICI_to_next_ms[ICI_to_next_flag == "reasonable_ici"],
      na.rm = TRUE
    ),
    
    .groups = "drop"
  ) %>%
  arrange(DriftID, EventID)

cat("ICI summary by EventID:\n")
print(ici_summary_by_event)

write.csv(
  ici_summary_by_event,
  file = file.path(ici_output_dir, "ici_summary_by_event_NEW.csv"),
  row.names = FALSE
)

```

## Drift-level ICI summary

```{r}

ici_summary_by_drift <- click_metadata_with_ici %>%
  group_by(DriftID) %>%
  summarise(
    n_clicks = n(),
    n_events = n_distinct(EventID),
    
    n_reasonable_ici = sum(ICI_to_next_flag == "reasonable_ici", na.rm = TRUE),
    n_zero_ici = sum(ICI_to_next_flag == "zero_ici", na.rm = TRUE),
    n_large_ici = sum(ICI_to_next_flag == "large_ici", na.rm = TRUE),
    n_last_click_no_next_ici = sum(ICI_to_next_flag == "last_click_no_next_ici", na.rm = TRUE),
    
    percent_reasonable_ici = 100 * n_reasonable_ici / n_clicks,
    percent_zero_ici = 100 * n_zero_ici / n_clicks,
    percent_large_ici = 100 * n_large_ici / n_clicks,
    
    mean_ici_ms = mean(
      ICI_to_next_ms[ICI_to_next_flag == "reasonable_ici"],
      na.rm = TRUE
    ),
    median_ici_ms = median(
      ICI_to_next_ms[ICI_to_next_flag == "reasonable_ici"],
      na.rm = TRUE
    ),
    sd_ici_ms = sd(
      ICI_to_next_ms[ICI_to_next_flag == "reasonable_ici"],
      na.rm = TRUE
    ),
    min_ici_ms = min(
      ICI_to_next_ms[ICI_to_next_flag == "reasonable_ici"],
      na.rm = TRUE
    ),
    max_ici_ms = max(
      ICI_to_next_ms[ICI_to_next_flag == "reasonable_ici"],
      na.rm = TRUE
    ),
    
    .groups = "drop"
  ) %>%
  arrange(DriftID)

cat("ICI summary by DriftID:\n")
print(ici_summary_by_drift)

write.csv(
  ici_summary_by_drift,
  file = file.path(ici_output_dir, "ici_summary_by_drift_NEW.csv"),
  row.names = FALSE
)

```

## Save final click-level ICI table

```{r}

saveRDS(
  click_metadata_with_ici,
  file = file.path(ici_output_dir, "click_metadata_with_ici_NEW.rds")
)

write.csv(
  click_metadata_with_ici,
  file = file.path(ici_output_dir, "click_metadata_with_ici_NEW.csv"),
  row.names = FALSE
)

cat("Saved final ICI table:\n")
cat(file.path(ici_output_dir, "click_metadata_with_ici_NEW.rds"), "\n")
cat(file.path(ici_output_dir, "click_metadata_with_ici_NEW.csv"), "\n")

cat("\nSaved summary tables:\n")
cat(file.path(ici_output_dir, "ici_summary_by_event_NEW.csv"), "\n")
cat(file.path(ici_output_dir, "ici_summary_by_drift_NEW.csv"), "\n")

```

## Locate corrected min-max normalized k-means cluster assignment files

```{r}

cluster_base_dir <- "E:/Analysis/Clustering Output/MinMax_Normalized_ADRIFT003_Corrected/Min-Max Normalized"

if (!dir.exists(cluster_base_dir)) {
  stop("Cluster base directory does not exist: ", cluster_base_dir)
}

all_cluster_files <- list.files(
  path = cluster_base_dir,
  pattern = "\\.(csv|rds|RDS)$",
  recursive = TRUE,
  full.names = TRUE
)

cluster_file_candidates <- data.frame(
  File = all_cluster_files,
  FileName = basename(all_cluster_files),
  Folder = dirname(all_cluster_files),
  stringsAsFactors = FALSE
) %>%
  filter(
    str_detect(tolower(FileName), "cluster|assignment|label|kmeans|click") |
      str_detect(tolower(Folder), "kmeans|cluster|k=")
  ) %>%
  arrange(File)

cat("Cluster base directory:\n")
cat(cluster_base_dir, "\n\n")

cat("Total csv/rds files found:", length(all_cluster_files), "\n")
cat("Number of possible cluster assignment files found:",
    nrow(cluster_file_candidates), "\n\n")

print(cluster_file_candidates)

write.csv(
  cluster_file_candidates,
  file = file.path(ici_output_dir, "corrected_cluster_file_candidates_NEW.csv"),
  row.names = FALSE
)

```

## Preview corrected cluster file candidates

```{r}

preview_cluster_file <- function(file_path) {
  
  cat("\n============================\n")
  cat("FILE:\n")
  cat(file_path, "\n")
  cat("============================\n")
  
  if (str_detect(tolower(file_path), "\\.csv$")) {
    
    preview <- read.csv(file_path, nrows = 10)
    
    cat("CSV columns:\n")
    print(names(preview))
    
    cat("\nPreview:\n")
    print(preview)
    
  } else if (str_detect(tolower(file_path), "\\.rds$")) {
    
    obj <- readRDS(file_path)
    
    cat("RDS class:\n")
    print(class(obj))
    
    if (is.data.frame(obj)) {
      cat("\nRDS dataframe columns:\n")
      print(names(obj))
      
      cat("\nPreview:\n")
      print(head(obj, 10))
    } else if (is.list(obj)) {
      cat("\nRDS list names:\n")
      print(names(obj))
      cat("\nStructure preview:\n")
      print(str(obj, max.level = 2))
    } else {
      cat("\nObject structure:\n")
      print(str(obj))
    }
  }
}

# Preview up to 30 candidates
files_to_preview <- head(cluster_file_candidates$File, 30)

for (f in files_to_preview) {
  preview_cluster_file(f)
}

```

## Merge ICI with corrected min-max k-means cluster assignments

```{r merge-ici-with-corrected-kmeans-clusters}

# This reads the existing clustering outputs but DOES NOT edit them.
# New merged files are written to a new ICI output folder.
# ------------------------------------------------------------

cluster_base_dir <- "E:/Analysis/Clustering Output/MinMax_Normalized_ADRIFT003_Corrected/Min-Max Normalized"

merged_cluster_ici_dir <- file.path(
  ici_output_dir,
  "ICI_Merged_With_Corrected_MinMax_KMeans"
)

dir.create(merged_cluster_ici_dir, recursive = TRUE, showWarnings = FALSE)

k_values_to_merge <- 2:12

# Keep only the ICI/timing fields we need for cluster-level summaries and plotting
ici_for_merge <- click_metadata_with_ici %>%
  select(
    EventID,
    UID,
    ClickIndexWithinEvent,
    DriftID,
    Final_UTC_clean,
    Final_UTC_ms,
    ICI_to_next_ms,
    ICI_to_next_flag,
    ICI_from_previous_ms,
    ICI_from_previous_flag,
    IsFirstClickInEvent,
    IsLastClickInEvent
  ) %>%
  mutate(
    UID = as.character(UID)
  )

# Sanity check: ICI table should have one row per EventID + UID
ici_duplicate_check <- ici_for_merge %>%
  count(EventID, UID, name = "n") %>%
  filter(n > 1)

if (nrow(ici_duplicate_check) > 0) {
  stop("ICI table has duplicated EventID + UID rows. Do not merge yet.")
}

merged_kmeans_ici_list <- list()

for (k in k_values_to_merge) {
  
  cat("\n============================\n")
  cat("Processing k =", k, "\n")
  cat("============================\n")
  
  cluster_file <- file.path(
    cluster_base_dir,
    paste0("kmeans k_", k),
    "data",
    "cluster_assignments_full.csv"
  )
  
  if (!file.exists(cluster_file)) {
    warning("Missing cluster assignment file for k = ", k, ": ", cluster_file)
    next
  }
  
  cluster_assignments <- read.csv(cluster_file, stringsAsFactors = FALSE) %>%
    mutate(
      UID = as.character(UID),
      k = k,
      cluster = as.integer(cluster),
      cluster_label = paste0("k", k, "_cluster_", cluster)
    )
  
  required_cluster_cols <- c("EventID", "UID", "cluster")
  
  missing_cluster_cols <- setdiff(required_cluster_cols, names(cluster_assignments))
  
  if (length(missing_cluster_cols) > 0) {
    stop(
      "Cluster file for k = ", k,
      " is missing required columns: ",
      paste(missing_cluster_cols, collapse = ", ")
    )
  }
  
  cluster_duplicate_check <- cluster_assignments %>%
    count(EventID, UID, name = "n") %>%
    filter(n > 1)
  
  if (nrow(cluster_duplicate_check) > 0) {
    stop("Cluster assignments for k = ", k, " have duplicated EventID + UID rows.")
  }
  
  merged_kmeans_ici <- cluster_assignments %>%
    left_join(
      ici_for_merge,
      by = c("EventID", "UID")
    )
  
  cat("Rows in cluster assignments:", nrow(cluster_assignments), "\n")
  cat("Rows after ICI merge:", nrow(merged_kmeans_ici), "\n")
  cat("Rows missing ICI_to_next_flag:", sum(is.na(merged_kmeans_ici$ICI_to_next_flag)), "\n")
  
  if (nrow(merged_kmeans_ici) != nrow(cluster_assignments)) {
    stop("Merge changed row count for k = ", k)
  }
  
  if (sum(is.na(merged_kmeans_ici$ICI_to_next_flag)) > 0) {
    stop("Some cluster rows did not find matching ICI rows for k = ", k)
  }
  
  # Save one merged file per k
  per_k_output_dir <- file.path(
    merged_cluster_ici_dir,
    paste0("kmeans_k_", k)
  )
  
  dir.create(per_k_output_dir, recursive = TRUE, showWarnings = FALSE)
  
  saveRDS(
    merged_kmeans_ici,
    file = file.path(per_k_output_dir, paste0("kmeans_k_", k, "_with_ici_NEW.rds"))
  )
  
  write.csv(
    merged_kmeans_ici,
    file = file.path(per_k_output_dir, paste0("kmeans_k_", k, "_with_ici_NEW.csv")),
    row.names = FALSE
  )
  
  merged_kmeans_ici_list[[paste0("k_", k)]] <- merged_kmeans_ici
}

# Combine all k values into one long-format object
merged_kmeans_ici_all_k <- bind_rows(merged_kmeans_ici_list)

saveRDS(
  merged_kmeans_ici_all_k,
  file = file.path(merged_cluster_ici_dir, "kmeans_all_k_with_ici_NEW.rds")
)

cat("\n============================\n")
cat("Finished merging ICI with corrected k-means clusters.\n")
cat("============================\n")

cat("Combined all-k rows:", nrow(merged_kmeans_ici_all_k), "\n")
cat("Expected rows if all k = 2:12 are present:",
    expected_total_clicks_new * length(k_values_to_merge), "\n")

cat("\nMerged outputs saved to:\n")
cat(merged_cluster_ici_dir, "\n")

```

## Summarize ICI by k-means cluster

```{r summarize-ici-by-kmeans-cluster}

# Main biological summaries use reasonable ICI only:
#   ICI_to_next_flag == "reasonable_ici"
# ------------------------------------------------------------

ici_summary_by_kmeans_cluster <- merged_kmeans_ici_all_k %>%
  group_by(k, cluster, cluster_label) %>%
  summarise(
    n_clicks = n(),
    
    n_reasonable_ici = sum(ICI_to_next_flag == "reasonable_ici", na.rm = TRUE),
    n_zero_ici = sum(ICI_to_next_flag == "zero_ici", na.rm = TRUE),
    n_large_ici = sum(ICI_to_next_flag == "large_ici", na.rm = TRUE),
    n_last_click_no_next_ici = sum(ICI_to_next_flag == "last_click_no_next_ici", na.rm = TRUE),
    
    percent_reasonable_ici = 100 * n_reasonable_ici / n_clicks,
    percent_zero_ici = 100 * n_zero_ici / n_clicks,
    percent_large_ici = 100 * n_large_ici / n_clicks,
    
    mean_ici_ms = mean(
      ICI_to_next_ms[ICI_to_next_flag == "reasonable_ici"],
      na.rm = TRUE
    ),
    median_ici_ms = median(
      ICI_to_next_ms[ICI_to_next_flag == "reasonable_ici"],
      na.rm = TRUE
    ),
    sd_ici_ms = sd(
      ICI_to_next_ms[ICI_to_next_flag == "reasonable_ici"],
      na.rm = TRUE
    ),
    q25_ici_ms = quantile(
      ICI_to_next_ms[ICI_to_next_flag == "reasonable_ici"],
      probs = 0.25,
      na.rm = TRUE
    ),
    q75_ici_ms = quantile(
      ICI_to_next_ms[ICI_to_next_flag == "reasonable_ici"],
      probs = 0.75,
      na.rm = TRUE
    ),
    min_ici_ms = min(
      ICI_to_next_ms[ICI_to_next_flag == "reasonable_ici"],
      na.rm = TRUE
    ),
    max_ici_ms = max(
      ICI_to_next_ms[ICI_to_next_flag == "reasonable_ici"],
      na.rm = TRUE
    ),
    
    .groups = "drop"
  ) %>%
  arrange(k, cluster)

cat("ICI summary by corrected min-max k-means cluster:\n")
print(ici_summary_by_kmeans_cluster)

write.csv(
  ici_summary_by_kmeans_cluster,
  file = file.path(
    merged_cluster_ici_dir,
    "ici_summary_by_corrected_minmax_kmeans_cluster_NEW.csv"
  ),
  row.names = FALSE
)

```

## Generate ICI boxplots for every k value

```{r}

# For each k:
#   x-axis = spectral cluster
#   y-axis = ICI to next click in ms
#   data = reasonable ICI values only
#
# Output:
#   one PNG per k value
# ------------------------------------------------------------

ici_boxplot_dir <- file.path(
  merged_cluster_ici_dir,
  "plots",
  "boxplots_by_k"
)

dir.create(ici_boxplot_dir, recursive = TRUE, showWarnings = FALSE)

k_values_to_plot <- sort(unique(merged_kmeans_ici_all_k$k))

for (current_k in k_values_to_plot) {
  
  cat("\n============================\n")
  cat("Plotting k =", current_k, "\n")
  cat("============================\n")
  
  plot_data_k <- merged_kmeans_ici_all_k %>%
    filter(
      k == current_k,
      ICI_to_next_flag == "reasonable_ici"
    ) %>%
    mutate(
      cluster = factor(cluster)
    )
  
  cat("Rows used:", nrow(plot_data_k), "\n")
  
  cluster_counts_k <- plot_data_k %>%
    count(cluster, name = "n_reasonable_ici") %>%
    arrange(cluster)
  
  print(cluster_counts_k)
  
  p <- ggplot(
    plot_data_k,
    aes(x = cluster, y = ICI_to_next_ms)
  ) +
    geom_boxplot(
      outlier.alpha = 0.08,
      outlier.size = 0.35
    ) +
    labs(
      title = paste0("ICI Distribution by Spectral Cluster, k = ", current_k),
      subtitle = "Corrected min-max normalized k-means clusters; reasonable ICI values only",
      x = "Spectral cluster",
      y = "ICI to next click (ms)"
    ) +
    theme_bw(base_size = 13)
  
  print(p)
  
  ggsave(
    filename = file.path(
      ici_boxplot_dir,
      paste0("ici_boxplot_by_cluster_k_", current_k, ".png")
    ),
    plot = p,
    width = 8,
    height = 5,
    dpi = 300
  )
}

cat("\nSaved all k-specific boxplots to:\n")
cat(ici_boxplot_dir, "\n")

```

## Generate zoomed ICI boxplots for every k value

```{r}

ici_boxplot_zoom_dir <- file.path(
  merged_cluster_ici_dir,
  "plots",
  "boxplots_by_k_zoomed_0_200ms"
)

dir.create(ici_boxplot_zoom_dir, recursive = TRUE, showWarnings = FALSE)

k_values_to_plot <- sort(unique(merged_kmeans_ici_all_k$k))

for (current_k in k_values_to_plot) {
  
  plot_data_k <- merged_kmeans_ici_all_k %>%
    filter(
      k == current_k,
      ICI_to_next_flag == "reasonable_ici"
    ) %>%
    mutate(
      cluster = factor(cluster)
    )
  
  p <- ggplot(
    plot_data_k,
    aes(x = cluster, y = ICI_to_next_ms)
  ) +
    geom_boxplot(
      outlier.alpha = 0.04,
      outlier.size = 0.25
    ) +
    coord_cartesian(ylim = c(0, 200)) +
    labs(
      title = paste0("ICI Distribution by Spectral Cluster, k = ", current_k),
      subtitle = "Zoomed to 0–200 ms; reasonable ICI values only",
      x = "Spectral cluster",
      y = "ICI to next click (ms)"
    ) +
    theme_bw(base_size = 13)
  
  print(p)
  
  ggsave(
    filename = file.path(
      ici_boxplot_zoom_dir,
      paste0("ici_boxplot_by_cluster_k_", current_k, "_zoom_0_200ms.png")
    ),
    plot = p,
    width = 8,
    height = 5,
    dpi = 300
  )
}

cat("\nSaved zoomed boxplots to:\n")
cat(ici_boxplot_zoom_dir, "\n")

```

## Clean summary table of ICI by k-means cluster

```{r}

ici_cluster_summary_clean <- merged_kmeans_ici_all_k %>%
  filter(ICI_to_next_flag == "reasonable_ici") %>%
  group_by(k, cluster) %>%
  summarise(
    n_clicks = n(),
    median_ici_ms = median(ICI_to_next_ms, na.rm = TRUE),
    mean_ici_ms = mean(ICI_to_next_ms, na.rm = TRUE),
    q25_ici_ms = quantile(ICI_to_next_ms, 0.25, na.rm = TRUE),
    q75_ici_ms = quantile(ICI_to_next_ms, 0.75, na.rm = TRUE),
    iqr_ici_ms = IQR(ICI_to_next_ms, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(k, cluster)

print(ici_cluster_summary_clean)

write.csv(
  ici_cluster_summary_clean,
  file = file.path(
    merged_cluster_ici_dir,
    "ici_cluster_summary_clean_reasonable_only_NEW.csv"
  ),
  row.names = FALSE
)

```



























