---
title: "PCA_ICI_UMAP_HDBSCAN"
author: "Sarah Moseley"
date: "2026-05-07"
output: html_documenthtml_document:
    toc: true
    toc_float: true
    theme: readable
---

## Setup

```{r}

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)

required_packages <- c(
  "dplyr",
  "ggplot2",
  "tidyr",
  "stringr",
  "readr",
  "purrr",
  "tibble",
  "RColorBrewer",
  "uwot",
  "dbscan"
)

missing_packages <- required_packages[
  !sapply(required_packages, requireNamespace, quietly = TRUE)
]

if (length(missing_packages) > 0) {
  stop(
    "Missing required packages: ",
    paste(missing_packages, collapse = ", "),
    "\nInstall them before running this workflow."
  )
}

library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
library(readr)
library(purrr)
library(tibble)
library(RColorBrewer)
library(uwot)
library(dbscan)

```

## User Settings

```{r}

# ------------------------------------------------------------
# Input files
# ------------------------------------------------------------

# Corrected combined Tukey object
# This should be the fixed combined file where ADRIFT_003_Event_001 has already been corrected.
combined_file <- file.choose()

# Previously generated click-level ICI table
# This should come from your ICI workflow:
# click_metadata_with_ici_NEW.rds
ici_file <- file.choose()

# ------------------------------------------------------------
# Output folder
# ------------------------------------------------------------

output_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN"

dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

plot_dir <- file.path(output_dir, "plots")
data_dir <- file.path(output_dir, "data")
table_dir <- file.path(output_dir, "tables")

dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(data_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(table_dir, recursive = TRUE, showWarnings = FALSE)

# ------------------------------------------------------------
# Spectral settings
# ------------------------------------------------------------

freq_min_khz <- 19
freq_max_khz <- 50

# ------------------------------------------------------------
# ICI filtering settings
# ------------------------------------------------------------

ici_min_s <- 0.01
ici_max_s <- 0.4

# ------------------------------------------------------------
# Sampling settings
# ------------------------------------------------------------

clicks_per_event <- 200
sampling_seed <- 123

# ------------------------------------------------------------
# PCA settings
# ------------------------------------------------------------

fixed_n_pcs <- 5
variance_cutoffs <- c(0.90, 0.95)

# ------------------------------------------------------------
# UMAP settings
# ------------------------------------------------------------

umap_neighbors <- 30
umap_min_dist <- 0.1
umap_metric <- "euclidean"
umap_seed <- 123

# ------------------------------------------------------------
# HDBSCAN settings
# ------------------------------------------------------------

hdbscan_minPts <- 30

# ------------------------------------------------------------
# Concatenated spectrogram settings
# ------------------------------------------------------------

max_clicks_per_cluster_spectrogram <- 300
cluster_spectrogram_seed <- 456

cat("Output directory:\n")
cat(output_dir, "\n")

```

## Load Combined Tukey and ICI Tables

```{r}

# ------------------------------------------------------------
# Load corrected combined Tukey object
# ------------------------------------------------------------

combined_tukey <- readRDS(combined_file)

cat("Loaded combined_tukey object.\n")
cat("Number of events:", length(combined_tukey), "\n\n")

# ------------------------------------------------------------
# Load click-level ICI table
# ------------------------------------------------------------

click_metadata_with_ici <- readRDS(ici_file)

cat("Loaded click_metadata_with_ici object.\n")
cat("Rows in ICI table:", nrow(click_metadata_with_ici), "\n\n")

# ------------------------------------------------------------
# Preview structures
# ------------------------------------------------------------

cat("Class of combined_tukey:\n")
print(class(combined_tukey))

cat("\nClass of first event object:\n")
print(class(combined_tukey[[1]]))

cat("\nNames in first event object:\n")
print(names(combined_tukey[[1]]))

cat("\nColumns in click_metadata_with_ici:\n")
print(names(click_metadata_with_ici))

```

## Sanity Checks for Combined Tukey Structure

```{r}

event_structure_summary <- data.frame(
  
  EventIndex = seq_along(combined_tukey),
  
  EventID = sapply(combined_tukey, function(x) x$EventID),
  
  n_freq = sapply(combined_tukey, function(x) length(x$freq)),
  
  n_uid = sapply(combined_tukey, function(x) length(x$UID)),
  
  allSpec_rows = sapply(combined_tukey, function(x) dim(x$allSpec)[1]),
  
  allSpec_cols = sapply(combined_tukey, function(x) dim(x$allSpec)[2]),
  
  stringsAsFactors = FALSE
)

print(event_structure_summary)

# ------------------------------------------------------------
# Total click count
# ------------------------------------------------------------

total_clicks <- sum(event_structure_summary$allSpec_cols)

cat("\nTotal clicks across all events:", total_clicks, "\n")

# ------------------------------------------------------------
# Core dimension checks
# ------------------------------------------------------------

uid_matches_allSpec <- all(
  event_structure_summary$n_uid ==
    event_structure_summary$allSpec_cols
)

freq_matches_allSpec <- all(
  event_structure_summary$n_freq ==
    event_structure_summary$allSpec_rows
)

cat("\nUID count matches allSpec columns:\n")
print(uid_matches_allSpec)

cat("\nFrequency count matches allSpec rows:\n")
print(freq_matches_allSpec)

if (uid_matches_allSpec && freq_matches_allSpec) {
  
  cat("\nPASS: Core spectral dimensions are aligned.\n")
  
} else {
  
  stop("FAIL: Spectral dimensions are misaligned.")
}

```

## Check for Duplicated EventIDs and UID Problems

```{r}

# ------------------------------------------------------------
# Build click-level metadata from Tukey object
# ------------------------------------------------------------

click_metadata_from_tukey <- list()

for (i in seq_along(combined_tukey)) {
  
  event_obj <- combined_tukey[[i]]
  
  n_clicks <- ncol(event_obj$allSpec)
  
  click_metadata_from_tukey[[i]] <- data.frame(
    
    EventIndex = i,
    
    EventID = event_obj$EventID,
    
    UID = as.character(event_obj$UID),
    
    ClickIndexWithinEvent = seq_len(n_clicks),
    
    stringsAsFactors = FALSE
  )
}

click_metadata_from_tukey <- bind_rows(click_metadata_from_tukey)

cat("Rows in Tukey click metadata:",
    nrow(click_metadata_from_tukey), "\n")

# ------------------------------------------------------------
# Duplicate EventID check
# ------------------------------------------------------------

duplicate_event_ids <- click_metadata_from_tukey %>%
  distinct(EventIndex, EventID) %>%
  count(EventID, name = "n_rows") %>%
  filter(n_rows > 1)

cat("\nDuplicated EventIDs:\n")
print(duplicate_event_ids)

# ------------------------------------------------------------
# Duplicate UID within EventID
# ------------------------------------------------------------

duplicate_uid_within_event <- click_metadata_from_tukey %>%
  count(EventID, UID, name = "n_rows") %>%
  filter(n_rows > 1)

cat("\nDuplicated UID values within EventID:\n")
print(duplicate_uid_within_event)

# ------------------------------------------------------------
# PASS / FAIL checks
# ------------------------------------------------------------

if (nrow(duplicate_event_ids) == 0) {
  
  cat("\nPASS: No duplicated EventIDs.\n")
  
} else {
  
  stop("FAIL: Duplicated EventIDs detected.")
}

if (nrow(duplicate_uid_within_event) == 0) {
  
  cat("\nPASS: No duplicated UID values within events.\n")
  
} else {
  
  stop("FAIL: Duplicated UID values detected within events.")
}

```

## Inspect Frequency Grids Before Interpolation

```{r}

freq_grid_summary <- data.frame(
  EventIndex = seq_along(combined_tukey),
  EventID = sapply(combined_tukey, function(x) x$EventID),
  n_freq = sapply(combined_tukey, function(x) length(x$freq)),
  min_freq = sapply(combined_tukey, function(x) min(x$freq, na.rm = TRUE)),
  max_freq = sapply(combined_tukey, function(x) max(x$freq, na.rm = TRUE)),
  median_freq_step = sapply(combined_tukey, function(x) median(diff(x$freq), na.rm = TRUE)),
  stringsAsFactors = FALSE
)

print(freq_grid_summary)

write.csv(
  freq_grid_summary,
  file = file.path(table_dir, "frequency_grid_summary_before_interpolation.csv"),
  row.names = FALSE
)

# Check whether all frequency grids are already identical
reference_freq_original <- combined_tukey[[1]]$freq

all_freq_grids_identical <- all(
  sapply(
    combined_tukey,
    function(x) identical(x$freq, reference_freq_original)
  )
)

cat("\nAre all original frequency grids identical?\n")
print(all_freq_grids_identical)

```

## Interpolate All Events to Shared Frequency Grid

```{r}

# ------------------------------------------------------------
# Use first event frequency grid as reference
# ------------------------------------------------------------

reference_freq <- combined_tukey[[1]]$freq

cat("Reference grid length:", length(reference_freq), "\n")
cat("Reference min frequency:", min(reference_freq), "\n")
cat("Reference max frequency:", max(reference_freq), "\n")
cat("Reference median step:", median(diff(reference_freq)), "\n")

# ------------------------------------------------------------
# Helper function: interpolate one matrix to reference grid
# ------------------------------------------------------------

interpolate_spectra_matrix <- function(original_freq, spectra_matrix, reference_freq) {
  
  interpolated_matrix <- apply(
    spectra_matrix,
    2,
    function(one_spectrum) {
      approx(
        x = original_freq,
        y = one_spectrum,
        xout = reference_freq,
        method = "linear",
        rule = 2
      )$y
    }
  )
  
  return(interpolated_matrix)
}

# ------------------------------------------------------------
# Interpolate each event
# ------------------------------------------------------------

interpolated_tukey <- vector("list", length(combined_tukey))

for (i in seq_along(combined_tukey)) {
  
  event_obj <- combined_tukey[[i]]
  
  cat("\nInterpolating event", i, "of", length(combined_tukey), ":",
      event_obj$EventID, "\n")
  
  interpolated_allSpec <- interpolate_spectra_matrix(
    original_freq = event_obj$freq,
    spectra_matrix = event_obj$allSpec,
    reference_freq = reference_freq
  )
  
  interpolated_tukey[[i]] <- event_obj
  interpolated_tukey[[i]]$freq <- reference_freq
  interpolated_tukey[[i]]$allSpec <- interpolated_allSpec
}

# ------------------------------------------------------------
# Sanity check: all interpolated grids identical
# ------------------------------------------------------------

all_interpolated_grids_identical <- all(
  sapply(
    interpolated_tukey,
    function(x) identical(x$freq, reference_freq)
  )
)

cat("\nAre all interpolated frequency grids identical?\n")
print(all_interpolated_grids_identical)

if (all_interpolated_grids_identical) {
  cat("\nPASS: All events now share the same frequency grid.\n")
} else {
  stop("FAIL: Frequency-grid interpolation did not align all events.")
}

```

## Restrict Spectra to 19-50 kHz

```{r}

# ------------------------------------------------------------
# Convert reference frequency grid to kHz if needed
# ------------------------------------------------------------

if (max(reference_freq, na.rm = TRUE) > 1000) {
  reference_freq_khz <- reference_freq / 1000
  cat("Detected frequency grid in Hz. Converted to kHz.\n")
} else {
  reference_freq_khz <- reference_freq
  cat("Detected frequency grid already in kHz.\n")
}

# ------------------------------------------------------------
# Identify frequency bins inside target range
# ------------------------------------------------------------

freq_keep <- reference_freq_khz >= freq_min_khz & reference_freq_khz <= freq_max_khz

freq_focus <- reference_freq_khz[freq_keep]

cat("Frequency range requested:", freq_min_khz, "to", freq_max_khz, "kHz\n")
cat("Number of frequency bins retained:", length(freq_focus), "\n")

if (length(freq_focus) > 0) {
  cat("Actual minimum retained frequency:", min(freq_focus), "kHz\n")
  cat("Actual maximum retained frequency:", max(freq_focus), "kHz\n")
}

if (length(freq_focus) < 2) {
  stop("Too few frequency bins retained. Check frequency units and freq_min_khz/freq_max_khz.")
}

# ------------------------------------------------------------
# Save retained frequency grid
# ------------------------------------------------------------

freq_focus_table <- data.frame(
  FreqBin = seq_along(freq_focus),
  Frequency_kHz = freq_focus,
  Frequency_OriginalUnits = reference_freq[freq_keep]
)

write.csv(
  freq_focus_table,
  file = file.path(table_dir, "frequency_grid_19_50_khz.csv"),
  row.names = FALSE
)

```

## Build Click-By-Frequency Spectral Matrix

```{r}

# ------------------------------------------------------------
# Build one click-level spectral matrix and matching metadata table
# ------------------------------------------------------------

spectra_list <- vector("list", length(interpolated_tukey))
metadata_list <- vector("list", length(interpolated_tukey))

for (i in seq_along(interpolated_tukey)) {
  
  event_obj <- interpolated_tukey[[i]]
  
  # allSpec is frequency x click
  # First restrict to 19–50 kHz frequency bins, then transpose:
  # rows = clicks, columns = frequency bins
  spectra_focus_event <- t(event_obj$allSpec[freq_keep, , drop = FALSE])
  
  spectra_list[[i]] <- spectra_focus_event
  
  metadata_list[[i]] <- data.frame(
    EventIndex = i,
    EventID = event_obj$EventID,
    DriftID = sub("_Event_.*$", "", event_obj$EventID),
    UID = as.character(event_obj$UID),
    ClickIndexWithinEvent = seq_len(ncol(event_obj$allSpec)),
    stringsAsFactors = FALSE
  )
}

# ------------------------------------------------------------
# Combine all events into one matrix/table
# ------------------------------------------------------------

spectra_matrix <- do.call(rbind, spectra_list)
spectra_metadata <- bind_rows(metadata_list)

# Use kHz labels from corrected Chunk 9
colnames(spectra_matrix) <- paste0(
  "freq_",
  sprintf("%.3f", freq_focus),
  "_kHz"
)

# ------------------------------------------------------------
# Sanity checks
# ------------------------------------------------------------

cat("Rows in spectra matrix:", nrow(spectra_matrix), "\n")
cat("Columns in spectra matrix:", ncol(spectra_matrix), "\n")
cat("Rows in spectra metadata:", nrow(spectra_metadata), "\n")

if (nrow(spectra_matrix) != nrow(spectra_metadata)) {
  stop("FAIL: Spectra matrix rows do not match metadata rows.")
} else {
  cat("\nPASS: Spectra matrix and metadata rows are aligned.\n")
}

if (ncol(spectra_matrix) != length(freq_focus)) {
  stop("FAIL: Spectra matrix columns do not match retained frequency bins.")
} else {
  cat("PASS: Spectra matrix columns match retained frequency bins.\n")
}

cat("\nPreview spectra metadata:\n")
print(head(spectra_metadata))

cat("\nSpectra matrix dimensions:\n")
print(dim(spectra_matrix))

```

## Sanity Check Spectra Values

```{r}


# ------------------------------------------------------------
# Basic matrix checks
# ------------------------------------------------------------

cat("Spectra matrix dimensions:\n")
print(dim(spectra_matrix))

cat("\nNumber of NA values in spectra matrix:\n")
print(sum(is.na(spectra_matrix)))

cat("\nNumber of infinite / non-finite values in spectra matrix:\n")
print(sum(!is.finite(spectra_matrix)))

if (sum(is.na(spectra_matrix)) > 0 || sum(!is.finite(spectra_matrix)) > 0) {
  stop("FAIL: Spectra matrix contains NA, NaN, or infinite values.")
}

# ------------------------------------------------------------
# Check per-click min/max values
# ------------------------------------------------------------

spectra_row_min <- apply(spectra_matrix, 1, min, na.rm = TRUE)
spectra_row_max <- apply(spectra_matrix, 1, max, na.rm = TRUE)
spectra_row_range <- spectra_row_max - spectra_row_min

cat("\nSummary of per-click minimum values:\n")
print(summary(spectra_row_min))

cat("\nSummary of per-click maximum values:\n")
print(summary(spectra_row_max))

cat("\nSummary of per-click spectral ranges:\n")
print(summary(spectra_row_range))

# ------------------------------------------------------------
# Flag flat spectra, if any
# ------------------------------------------------------------

flat_spectra_rows <- which(spectra_row_range == 0)

cat("\nNumber of flat spectra rows:", length(flat_spectra_rows), "\n")

if (length(flat_spectra_rows) > 0) {
  
  cat("\nWARNING: Some spectra have zero range across retained frequencies.\n")
  cat("These may cause issues later if additional scaling is applied.\n")
  
  flat_spectra_table <- spectra_metadata[flat_spectra_rows, ]
  
  write.csv(
    flat_spectra_table,
    file = file.path(table_dir, "flat_spectra_rows_19_50_khz.csv"),
    row.names = FALSE
  )
}

# ------------------------------------------------------------
# Save row-level min/max diagnostics
# ------------------------------------------------------------

spectra_row_diagnostics <- spectra_metadata %>%
  mutate(
    row_min_19_50_khz = spectra_row_min,
    row_max_19_50_khz = spectra_row_max,
    row_range_19_50_khz = spectra_row_range
  )

write.csv(
  spectra_row_diagnostics,
  file = file.path(table_dir, "spectra_row_min_max_diagnostics_19_50_khz.csv"),
  row.names = FALSE
)

cat("\nPASS: Spectra matrix contains finite values only.\n")

```

## Min-Max Normalize Spectra Across 19-50 kHz

```{r}

# ------------------------------------------------------------
# Min-max normalize each click spectrum across 19–50 kHz
#
# rows = clicks
# columns = frequency bins
#
# Formula:
# normalized = (value - row_min) / (row_max - row_min)
# ------------------------------------------------------------

spectra_row_min <- apply(spectra_matrix, 1, min, na.rm = TRUE)
spectra_row_max <- apply(spectra_matrix, 1, max, na.rm = TRUE)
spectra_row_range <- spectra_row_max - spectra_row_min

if (any(spectra_row_range == 0)) {
  stop("Some spectra have zero range and cannot be min-max normalized.")
}

spectra_matrix_minmax <- sweep(
  spectra_matrix,
  1,
  spectra_row_min,
  FUN = "-"
)

spectra_matrix_minmax <- sweep(
  spectra_matrix_minmax,
  1,
  spectra_row_range,
  FUN = "/"
)

# ------------------------------------------------------------
# Sanity checks
# ------------------------------------------------------------

minmax_row_min <- apply(spectra_matrix_minmax, 1, min, na.rm = TRUE)
minmax_row_max <- apply(spectra_matrix_minmax, 1, max, na.rm = TRUE)

cat("Min-max normalized spectra dimensions:\n")
print(dim(spectra_matrix_minmax))

cat("\nSummary of normalized per-click minimum values:\n")
print(summary(minmax_row_min))

cat("\nSummary of normalized per-click maximum values:\n")
print(summary(minmax_row_max))

cat("\nNumber of NA values in normalized spectra:\n")
print(sum(is.na(spectra_matrix_minmax)))

cat("\nNumber of non-finite values in normalized spectra:\n")
print(sum(!is.finite(spectra_matrix_minmax)))

if (sum(is.na(spectra_matrix_minmax)) > 0 || sum(!is.finite(spectra_matrix_minmax)) > 0) {
  stop("FAIL: Min-max normalized spectra contain NA, NaN, or infinite values.")
}

cat("\nPASS: Spectra are now min-max normalized across 19–50 kHz.\n")

# ------------------------------------------------------------
# Save diagnostics
# ------------------------------------------------------------

minmax_diagnostics <- spectra_metadata %>%
  mutate(
    raw_min_19_50_khz = spectra_row_min,
    raw_max_19_50_khz = spectra_row_max,
    raw_range_19_50_khz = spectra_row_range,
    minmax_min_19_50_khz = minmax_row_min,
    minmax_max_19_50_khz = minmax_row_max
  )

write.csv(
  minmax_diagnostics,
  file = file.path(table_dir, "minmax_normalization_diagnostics_19_50_khz.csv"),
  row.names = FALSE
)

```

## Join ICI to Spectra Metadata

```{r}

# ------------------------------------------------------------
# Prepare ICI table for joining
# ------------------------------------------------------------

ici_for_join <- click_metadata_with_ici %>%
  mutate(
    UID = as.character(UID)
  ) %>%
  select(
    EventID,
    UID,
    DriftID,
    Final_UTC_clean,
    Final_UTC_ms,
    ICI_to_next_ms,
    ICI_to_next_flag,
    ICI_from_previous_ms,
    ICI_from_previous_flag,
    IsFirstClickInEvent,
    IsLastClickInEvent
  )

# ------------------------------------------------------------
# Check ICI table uniqueness before join
# ------------------------------------------------------------

ici_duplicate_check <- ici_for_join %>%
  count(EventID, UID, name = "n_rows") %>%
  filter(n_rows > 1)

cat("Duplicated EventID + UID rows in ICI table:\n")
print(ici_duplicate_check)

if (nrow(ici_duplicate_check) > 0) {
  stop("FAIL: ICI table has duplicated EventID + UID rows.")
}

# ------------------------------------------------------------
# Join ICI table to spectra metadata
# ------------------------------------------------------------

spectra_metadata_with_ici <- spectra_metadata %>%
  left_join(
    ici_for_join,
    by = c("EventID", "UID"),
    suffix = c("_spectra", "_ici")
  )

cat("Rows before join:", nrow(spectra_metadata), "\n")
cat("Rows after join:", nrow(spectra_metadata_with_ici), "\n")

if (nrow(spectra_metadata_with_ici) != nrow(spectra_metadata)) {
  stop("FAIL: ICI join changed the number of rows.")
}

# ------------------------------------------------------------
# Create ICI in seconds
# ------------------------------------------------------------

spectra_metadata_with_ici <- spectra_metadata_with_ici %>%
  mutate(
    ICI_to_next_s = ICI_to_next_ms / 1000,
    log10_ICI_to_next_s = log10(ICI_to_next_s)
  )

# ------------------------------------------------------------
# Join coverage checks
# ------------------------------------------------------------

cat("\nRows missing ICI_to_next_ms:\n")
print(sum(is.na(spectra_metadata_with_ici$ICI_to_next_ms)))

cat("\nICI_to_next_flag counts:\n")
print(table(spectra_metadata_with_ici$ICI_to_next_flag, useNA = "ifany"))

cat("\nSummary of ICI_to_next_s:\n")
print(summary(spectra_metadata_with_ici$ICI_to_next_s))

# ------------------------------------------------------------
# Save joined metadata
# ------------------------------------------------------------

saveRDS(
  spectra_metadata_with_ici,
  file = file.path(data_dir, "spectra_metadata_with_ici.rds")
)

write.csv(
  spectra_metadata_with_ici,
  file = file.path(table_dir, "spectra_metadata_with_ici.csv"),
  row.names = FALSE
)

cat("\nPASS: ICI metadata joined to spectra metadata.\n")

```

## Inspect ICI Filtering Effect

```{r}

# ------------------------------------------------------------
# Label clicks by ICI filter category
# ------------------------------------------------------------

spectra_metadata_with_ici <- spectra_metadata_with_ici %>%
  mutate(
    ICI_filter_category = case_when(
      is.na(ICI_to_next_s) ~ "missing_ici",
      ICI_to_next_s == 0 ~ "zero_ici",
      ICI_to_next_s < ici_min_s ~ "below_0.01_s_likely_buzz",
      ICI_to_next_s >= ici_min_s & ICI_to_next_s <= ici_max_s ~ "kept_0.01_to_0.4_s",
      ICI_to_next_s > ici_max_s ~ "above_0.4_s_long_gap",
      TRUE ~ "other"
    )
  )

# ------------------------------------------------------------
# Overall filter summary
# ------------------------------------------------------------

ici_filter_summary <- spectra_metadata_with_ici %>%
  count(ICI_filter_category, name = "n_clicks") %>%
  mutate(
    percent_clicks = 100 * n_clicks / sum(n_clicks)
  ) %>%
  arrange(desc(n_clicks))

cat("Overall ICI filter summary:\n")
print(ici_filter_summary)

write.csv(
  ici_filter_summary,
  file = file.path(table_dir, "ici_filter_summary_overall.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# Filter summary by event
# ------------------------------------------------------------

ici_filter_summary_by_event <- spectra_metadata_with_ici %>%
  group_by(DriftID_spectra, EventID, ICI_filter_category) %>%
  summarise(
    n_clicks = n(),
    .groups = "drop"
  ) %>%
  group_by(DriftID_spectra, EventID) %>%
  mutate(
    percent_within_event = 100 * n_clicks / sum(n_clicks)
  ) %>%
  ungroup() %>%
  arrange(DriftID_spectra, EventID, ICI_filter_category)

cat("\nICI filter summary by event:\n")
print(ici_filter_summary_by_event)

write.csv(
  ici_filter_summary_by_event,
  file = file.path(table_dir, "ici_filter_summary_by_event.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# Kept-click summary by event
# ------------------------------------------------------------

kept_summary_by_event <- spectra_metadata_with_ici %>%
  group_by(DriftID_spectra, EventID) %>%
  summarise(
    n_total_clicks = n(),
    n_kept_0.01_to_0.4_s = sum(ICI_filter_category == "kept_0.01_to_0.4_s"),
    percent_kept = 100 * n_kept_0.01_to_0.4_s / n_total_clicks,
    .groups = "drop"
  ) %>%
  arrange(DriftID_spectra, EventID)

cat("\nKept-click summary by event:\n")
print(kept_summary_by_event)

write.csv(
  kept_summary_by_event,
  file = file.path(table_dir, "kept_click_summary_by_event.csv"),
  row.names = FALSE
)

```

## Filter to 0.01-0.4 s and Sample Up To 200 Clicks Per Event

```{r}

set.seed(sampling_seed)

# ------------------------------------------------------------
# Keep only clicks inside target ICI range
# ------------------------------------------------------------

eligible_clicks <- spectra_metadata_with_ici %>%
  mutate(
    RowID = row_number()
  ) %>%
  filter(
    ICI_filter_category == "kept_0.01_to_0.4_s"
  )

cat("Eligible clicks after ICI filtering:", nrow(eligible_clicks), "\n")

# ------------------------------------------------------------
# Sample up to 200 clicks per event
# ------------------------------------------------------------

sampled_clicks <- eligible_clicks %>%
  group_by(EventID) %>%
  slice_sample(
    n = clicks_per_event,
    replace = FALSE
  ) %>%
  ungroup()

cat("Sampled clicks:", nrow(sampled_clicks), "\n")
cat("Number of events represented:", n_distinct(sampled_clicks$EventID), "\n")

# ------------------------------------------------------------
# Build sampled spectra matrix
# ------------------------------------------------------------

sampled_row_ids <- sampled_clicks$RowID

spectra_matrix_sampled_minmax <- spectra_matrix_minmax[
  sampled_row_ids,
  ,
  drop = FALSE
]

cat("Sampled spectra matrix dimensions:\n")
print(dim(spectra_matrix_sampled_minmax))

if (nrow(spectra_matrix_sampled_minmax) != nrow(sampled_clicks)) {
  stop("FAIL: Sampled spectra matrix rows do not match sampled metadata rows.")
}

# ------------------------------------------------------------
# Summarize sampled clicks by event
# ------------------------------------------------------------

sample_summary_by_event <- sampled_clicks %>%
  count(DriftID_spectra, EventID, name = "n_sampled_clicks") %>%
  arrange(DriftID_spectra, EventID)

cat("\nSampled clicks by event:\n")
print(sample_summary_by_event, n = Inf)

write.csv(
  sample_summary_by_event,
  file = file.path(table_dir, "sampled_clicks_by_event.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# Save sampled metadata and spectra
# ------------------------------------------------------------

saveRDS(
  sampled_clicks,
  file = file.path(data_dir, "sampled_click_metadata_ici_filtered.rds")
)

saveRDS(
  spectra_matrix_sampled_minmax,
  file = file.path(data_dir, "sampled_minmax_spectra_matrix_19_50_khz.rds")
)

cat("\nPASS: ICI filtering and post-filter sampling complete.\n")

```

## Run PCA on Sampled Min-Max Spectra

```{r}

# ------------------------------------------------------------
# PCA on min-max normalized spectra only
#
# rows = sampled clicks
# columns = frequency bins from 19–50 kHz
# ------------------------------------------------------------

pca_spectra <- prcomp(
  spectra_matrix_sampled_minmax,
  center = TRUE,
  scale. = FALSE
)

# ------------------------------------------------------------
# Variance explained
# ------------------------------------------------------------

pca_variance <- pca_spectra$sdev^2
pca_variance_explained <- pca_variance / sum(pca_variance)
pca_cumulative_variance <- cumsum(pca_variance_explained)

pca_variance_table <- data.frame(
  PC = seq_along(pca_variance_explained),
  PC_label = paste0("PC", seq_along(pca_variance_explained)),
  variance_explained = pca_variance_explained,
  cumulative_variance_explained = pca_cumulative_variance
)

cat("PCA complete.\n")
cat("Number of PCs:", ncol(pca_spectra$x), "\n\n")

cat("First 10 PCs variance explained:\n")
print(head(pca_variance_table, 10))

# ------------------------------------------------------------
# PCs needed for variance cutoffs
# ------------------------------------------------------------

n_pcs_90 <- which(pca_cumulative_variance >= 0.90)[1]
n_pcs_95 <- which(pca_cumulative_variance >= 0.95)[1]

cat("\nNumber of PCs needed for 90% variance:", n_pcs_90, "\n")
cat("Number of PCs needed for 95% variance:", n_pcs_95, "\n")
cat("Fixed PC option:", fixed_n_pcs, "\n")

# ------------------------------------------------------------
# Save PCA objects and variance table
# ------------------------------------------------------------

saveRDS(
  pca_spectra,
  file = file.path(data_dir, "pca_spectra_sampled_minmax_19_50_khz.rds")
)

write.csv(
  pca_variance_table,
  file = file.path(table_dir, "pca_variance_explained.csv"),
  row.names = FALSE
)

cat("\nPASS: PCA completed and saved.\n")

```

## Plot PCA Variance Explained

```{r}

# ------------------------------------------------------------
# Scree plot: variance explained by each PC
# ------------------------------------------------------------

p_scree <- ggplot(
  pca_variance_table,
  aes(x = PC, y = variance_explained)
) +
  geom_col() +
  geom_vline(xintercept = fixed_n_pcs, linetype = "dashed") +
  geom_vline(xintercept = n_pcs_90, linetype = "dotted") +
  geom_vline(xintercept = n_pcs_95, linetype = "dotdash") +
  labs(
    title = "PCA Scree Plot",
    subtitle = paste0(
      "Dashed = 5 PCs; dotted = 90% variance (",
      n_pcs_90,
      " PCs); dot-dash = 95% variance (",
      n_pcs_95,
      " PCs)"
    ),
    x = "Principal Component",
    y = "Proportion of variance explained"
  ) +
  theme_bw(base_size = 13)

print(p_scree)

ggsave(
  filename = file.path(plot_dir, "pca_scree_plot.png"),
  plot = p_scree,
  width = 8,
  height = 5,
  dpi = 300
)

# ------------------------------------------------------------
# Cumulative variance plot
# ------------------------------------------------------------

p_cumulative <- ggplot(
  pca_variance_table,
  aes(x = PC, y = cumulative_variance_explained)
) +
  geom_line() +
  geom_point(size = 1) +
  geom_hline(yintercept = 0.90, linetype = "dotted") +
  geom_hline(yintercept = 0.95, linetype = "dotdash") +
  geom_vline(xintercept = fixed_n_pcs, linetype = "dashed") +
  geom_vline(xintercept = n_pcs_90, linetype = "dotted") +
  geom_vline(xintercept = n_pcs_95, linetype = "dotdash") +
  labs(
    title = "PCA Cumulative Variance Explained",
    subtitle = paste0(
      "5 PCs vs. PCs needed for 90% and 95% total spectral variance"
    ),
    x = "Number of principal components",
    y = "Cumulative variance explained"
  ) +
  theme_bw(base_size = 13)

print(p_cumulative)

ggsave(
  filename = file.path(plot_dir, "pca_cumulative_variance_plot.png"),
  plot = p_cumulative,
  width = 8,
  height = 5,
  dpi = 300
)

cat("Saved PCA variance plots to:\n")
cat(plot_dir, "\n")

```

## Build PC + Scaled logICI Feature Matrices

```{r}

# ------------------------------------------------------------
# Add PCA scores to sampled click metadata
# ------------------------------------------------------------

pc_scores <- as.data.frame(pca_spectra$x)

sampled_clicks_with_pcs <- bind_cols(
  sampled_clicks,
  pc_scores
)

# ------------------------------------------------------------
# Prepare log-transformed and scaled ICI
# ------------------------------------------------------------

sampled_clicks_with_pcs <- sampled_clicks_with_pcs %>%
  mutate(
    log10_ICI_to_next_s = log10(ICI_to_next_s),
    scaled_log10_ICI = as.numeric(scale(log10_ICI_to_next_s))
  )

cat("Summary of raw ICI seconds:\n")
print(summary(sampled_clicks_with_pcs$ICI_to_next_s))

cat("\nSummary of log10 ICI seconds:\n")
print(summary(sampled_clicks_with_pcs$log10_ICI_to_next_s))

cat("\nSummary of scaled log10 ICI:\n")
print(summary(sampled_clicks_with_pcs$scaled_log10_ICI))

# ------------------------------------------------------------
# Define PC sets
# ------------------------------------------------------------

pc_set_5 <- paste0("PC", 1:fixed_n_pcs)
pc_set_90 <- paste0("PC", 1:n_pcs_90)
pc_set_95 <- paste0("PC", 1:n_pcs_95)

# ------------------------------------------------------------
# Build feature matrices
# ------------------------------------------------------------

feature_matrix_5pc_ici <- sampled_clicks_with_pcs %>%
  select(all_of(pc_set_5), scaled_log10_ICI) %>%
  as.matrix()

feature_matrix_90pc_ici <- sampled_clicks_with_pcs %>%
  select(all_of(pc_set_90), scaled_log10_ICI) %>%
  as.matrix()

feature_matrix_95pc_ici <- sampled_clicks_with_pcs %>%
  select(all_of(pc_set_95), scaled_log10_ICI) %>%
  as.matrix()

feature_matrix_5pc_spectra_only <- sampled_clicks_with_pcs %>%
  select(all_of(pc_set_5)) %>%
  as.matrix()

feature_matrix_90pc_spectra_only <- sampled_clicks_with_pcs %>%
  select(all_of(pc_set_90)) %>%
  as.matrix()

# ------------------------------------------------------------
# Sanity checks
# ------------------------------------------------------------

cat("\nFeature matrix dimensions:\n")
cat("5 PCs + ICI:", dim(feature_matrix_5pc_ici), "\n")
cat("90% PCs + ICI:", dim(feature_matrix_90pc_ici), "\n")
cat("95% PCs + ICI:", dim(feature_matrix_95pc_ici), "\n")
cat("5 PCs only:", dim(feature_matrix_5pc_spectra_only), "\n")
cat("90% PCs only:", dim(feature_matrix_90pc_spectra_only), "\n")

if (any(!is.finite(feature_matrix_5pc_ici))) {
  stop("FAIL: 5 PC + ICI matrix contains non-finite values.")
}

if (any(!is.finite(feature_matrix_90pc_ici))) {
  stop("FAIL: 90 PC + ICI matrix contains non-finite values.")
}

if (any(!is.finite(feature_matrix_95pc_ici))) {
  stop("FAIL: 95 PC + ICI matrix contains non-finite values.")
}

# ------------------------------------------------------------
# Save feature metadata and matrices
# ------------------------------------------------------------

saveRDS(
  sampled_clicks_with_pcs,
  file = file.path(data_dir, "sampled_clicks_with_pcs_and_scaled_logici.rds")
)

saveRDS(
  feature_matrix_5pc_ici,
  file = file.path(data_dir, "feature_matrix_5pc_plus_scaled_logici.rds")
)

saveRDS(
  feature_matrix_90pc_ici,
  file = file.path(data_dir, "feature_matrix_90pc_plus_scaled_logici.rds")
)

saveRDS(
  feature_matrix_95pc_ici,
  file = file.path(data_dir, "feature_matrix_95pc_plus_scaled_logici.rds")
)

cat("\nPASS: PC + scaled logICI feature matrices created.\n")

```

## Run UMAP on Spectra-Only and Spectra + Scaled logICI Features Matrices

```{r}

set.seed(umap_seed)

# ------------------------------------------------------------
# Helper function to run UMAP cleanly
# ------------------------------------------------------------

run_umap_analysis <- function(feature_matrix, analysis_name) {
  
  cat("\n============================\n")
  cat("Running UMAP:", analysis_name, "\n")
  cat("============================\n")
  
  umap_result <- uwot::umap(
    feature_matrix,
    n_neighbors = umap_neighbors,
    min_dist = umap_min_dist,
    metric = umap_metric,
    n_components = 2,
    ret_model = FALSE,
    verbose = TRUE,
    seed = umap_seed
  )
  
  umap_df <- as.data.frame(umap_result)
  names(umap_df) <- c("UMAP1", "UMAP2")
  
  umap_df$Analysis <- analysis_name
  
  return(umap_df)
}

# ------------------------------------------------------------
# Main UMAP runs
# ------------------------------------------------------------

umap_5pc_spectra_only <- run_umap_analysis(
  feature_matrix = feature_matrix_5pc_spectra_only,
  analysis_name = "5PC_spectra_only"
)

umap_5pc_ici <- run_umap_analysis(
  feature_matrix = feature_matrix_5pc_ici,
  analysis_name = "5PC_plus_scaled_logICI"
)

umap_90pc_spectra_only <- run_umap_analysis(
  feature_matrix = feature_matrix_90pc_spectra_only,
  analysis_name = "90PC_spectra_only"
)

umap_90pc_ici <- run_umap_analysis(
  feature_matrix = feature_matrix_90pc_ici,
  analysis_name = "90PC_plus_scaled_logICI"
)

# ------------------------------------------------------------
# Attach UMAP coordinates to metadata
# ------------------------------------------------------------

umap_results_all <- bind_rows(
  bind_cols(sampled_clicks_with_pcs, umap_5pc_spectra_only),
  bind_cols(sampled_clicks_with_pcs, umap_5pc_ici),
  bind_cols(sampled_clicks_with_pcs, umap_90pc_spectra_only),
  bind_cols(sampled_clicks_with_pcs, umap_90pc_ici)
)

cat("\nRows in combined UMAP results:", nrow(umap_results_all), "\n")
cat("Expected rows:", nrow(sampled_clicks_with_pcs) * 4, "\n")

# ------------------------------------------------------------
# Save UMAP outputs
# ------------------------------------------------------------

saveRDS(
  umap_results_all,
  file = file.path(data_dir, "umap_results_all_main_comparisons.rds")
)

write.csv(
  umap_results_all,
  file = file.path(table_dir, "umap_results_all_main_comparisons.csv"),
  row.names = FALSE
)

cat("\nPASS: UMAP runs complete and saved.\n")

```

## Plot UMAPs Colored by logICI

```{r}

# ------------------------------------------------------------
# Plot helper
# ------------------------------------------------------------

plot_umap_logici <- function(df, analysis_name) {
  
  plot_data <- df %>%
    filter(Analysis == analysis_name)
  
  p <- ggplot(
    plot_data,
    aes(
      x = UMAP1,
      y = UMAP2,
      color = log10_ICI_to_next_s
    )
  ) +
    geom_point(
      size = 0.4,
      alpha = 0.5
    ) +
    scale_color_viridis_c(
      option = "plasma"
    ) +
    labs(
      title = paste0("UMAP: ", analysis_name),
      subtitle = "Colored by log10 ICI (seconds)",
      x = "UMAP1",
      y = "UMAP2",
      color = "log10 ICI"
    ) +
    theme_bw(base_size = 13)
  
  return(p)
}

# ------------------------------------------------------------
# Generate plots
# ------------------------------------------------------------

p_umap_5pc_only <- plot_umap_logici(
  umap_results_all,
  "5PC_spectra_only"
)

p_umap_5pc_ici <- plot_umap_logici(
  umap_results_all,
  "5PC_plus_scaled_logICI"
)

p_umap_90pc_only <- plot_umap_logici(
  umap_results_all,
  "90PC_spectra_only"
)

p_umap_90pc_ici <- plot_umap_logici(
  umap_results_all,
  "90PC_plus_scaled_logICI"
)

# ------------------------------------------------------------
# Print plots
# ------------------------------------------------------------

print(p_umap_5pc_only)
print(p_umap_5pc_ici)
print(p_umap_90pc_only)
print(p_umap_90pc_ici)

# ------------------------------------------------------------
# Save plots
# ------------------------------------------------------------

ggsave(
  file.path(plot_dir, "umap_5pc_spectra_only_logici.png"),
  p_umap_5pc_only,
  width = 7,
  height = 6,
  dpi = 300
)

ggsave(
  file.path(plot_dir, "umap_5pc_plus_logici.png"),
  p_umap_5pc_ici,
  width = 7,
  height = 6,
  dpi = 300
)

ggsave(
  file.path(plot_dir, "umap_90pc_spectra_only_logici.png"),
  p_umap_90pc_only,
  width = 7,
  height = 6,
  dpi = 300
)

ggsave(
  file.path(plot_dir, "umap_90pc_plus_logici.png"),
  p_umap_90pc_ici,
  width = 7,
  height = 6,
  dpi = 300
)

cat("Saved UMAP logICI plots to:\n")
cat(plot_dir, "\n")

```

## Add Raw-Scaled ICI Comparison UMAPS

```{r}

# ------------------------------------------------------------
# Build raw-scaled ICI feature matrices
# ------------------------------------------------------------

sampled_clicks_with_pcs <- sampled_clicks_with_pcs %>%
  mutate(
    scaled_raw_ICI = as.numeric(scale(ICI_to_next_s))
  )

feature_matrix_5pc_rawici <- sampled_clicks_with_pcs %>%
  select(all_of(pc_set_5), scaled_raw_ICI) %>%
  as.matrix()

feature_matrix_90pc_rawici <- sampled_clicks_with_pcs %>%
  select(all_of(pc_set_90), scaled_raw_ICI) %>%
  as.matrix()

cat("Summary of scaled raw ICI:\n")
print(summary(sampled_clicks_with_pcs$scaled_raw_ICI))

# ------------------------------------------------------------
# Run UMAPs
# ------------------------------------------------------------

umap_5pc_rawici <- run_umap_analysis(
  feature_matrix = feature_matrix_5pc_rawici,
  analysis_name = "5PC_plus_scaled_rawICI"
)

umap_90pc_rawici <- run_umap_analysis(
  feature_matrix = feature_matrix_90pc_rawici,
  analysis_name = "90PC_plus_scaled_rawICI"
)

umap_rawici_results <- bind_rows(
  bind_cols(sampled_clicks_with_pcs, umap_5pc_rawici),
  bind_cols(sampled_clicks_with_pcs, umap_90pc_rawici)
)

saveRDS(
  umap_rawici_results,
  file = file.path(data_dir, "umap_results_rawici_comparison.rds")
)

cat("\nPASS: Raw-scaled ICI UMAP comparison complete.\n")

```

## Plot Raw-ICI UMAPs

```{r}

p_umap_5pc_rawici <- plot_umap_logici(
  umap_rawici_results,
  "5PC_plus_scaled_rawICI"
)

p_umap_90pc_rawici <- plot_umap_logici(
  umap_rawici_results,
  "90PC_plus_scaled_rawICI"
)

print(p_umap_5pc_rawici)
print(p_umap_90pc_rawici)

ggsave(
  file.path(plot_dir, "umap_5pc_plus_scaled_rawici_colored_logici.png"),
  p_umap_5pc_rawici,
  width = 7,
  height = 6,
  dpi = 300
)

ggsave(
  file.path(plot_dir, "umap_90pc_plus_scaled_rawici_colored_logici.png"),
  p_umap_90pc_rawici,
  width = 7,
  height = 6,
  dpi = 300
)

cat("Saved raw-ICI comparison UMAP plots.\n")

```

## Plot UMAPs by DriftID

```{r}

plot_umap_drift <- function(df, analysis_name) {
  
  plot_data <- df %>%
    filter(Analysis == analysis_name)
  
  p <- ggplot(
    plot_data,
    aes(
      x = UMAP1,
      y = UMAP2,
      color = DriftID_spectra
    )
  ) +
    geom_point(size = 0.45, alpha = 0.65) +
    labs(
      title = paste0("UMAP: ", analysis_name),
      subtitle = "Colored by DriftID",
      x = "UMAP1",
      y = "UMAP2",
      color = "DriftID"
    ) +
    theme_bw(base_size = 13) +
    theme(
      legend.position = "right"
    )
  
  return(p)
}

drift_umap_analyses <- c(
  "5PC_spectra_only",
  "5PC_plus_scaled_logICI",
  "90PC_spectra_only",
  "90PC_plus_scaled_logICI"
)

for (analysis_name in drift_umap_analyses) {
  
  p <- plot_umap_drift(umap_results_all, analysis_name)
  
  print(p)
  
  ggsave(
    file.path(
      plot_dir,
      paste0("umap_", analysis_name, "_by_drift.png")
    ),
    p,
    width = 8,
    height = 6,
    dpi = 300
  )
}

```

## Load Fixed k-means Cluster Assignments and Merge with UMAP

```{r}

# ------------------------------------------------------------
# Choose which existing k-means result to compare against
# ------------------------------------------------------------

kmeans_k_to_compare <- 5

cluster_base_dir <- "E:/Analysis/Clustering Output/MinMax_Normalized_ADRIFT003_Corrected/Min-Max Normalized"

cluster_file <- file.path(
  cluster_base_dir,
  paste0("kmeans k_", kmeans_k_to_compare),
  "data",
  "cluster_assignments_full.csv"
)

if (!file.exists(cluster_file)) {
  stop("Could not find k-means cluster file: ", cluster_file)
}

kmeans_assignments <- read.csv(cluster_file, stringsAsFactors = FALSE) %>%
  mutate(
    UID = as.character(UID),
    kmeans_cluster = factor(cluster),
    kmeans_k = kmeans_k_to_compare
  ) %>%
  select(EventID, UID, kmeans_cluster, kmeans_k)

# ------------------------------------------------------------
# Check uniqueness
# ------------------------------------------------------------

kmeans_duplicate_check <- kmeans_assignments %>%
  count(EventID, UID, name = "n_rows") %>%
  filter(n_rows > 1)

if (nrow(kmeans_duplicate_check) > 0) {
  stop("K-means assignments contain duplicated EventID + UID rows.")
}

# ------------------------------------------------------------
# Merge with UMAP results
# ------------------------------------------------------------

umap_results_with_kmeans <- umap_results_all %>%
  mutate(UID = as.character(UID)) %>%
  left_join(
    kmeans_assignments,
    by = c("EventID", "UID")
  )

cat("Rows in UMAP results:", nrow(umap_results_all), "\n")
cat("Rows after k-means merge:", nrow(umap_results_with_kmeans), "\n")
cat("Rows missing k-means cluster:", sum(is.na(umap_results_with_kmeans$kmeans_cluster)), "\n")

if (nrow(umap_results_with_kmeans) != nrow(umap_results_all)) {
  stop("K-means merge changed row count.")
}

saveRDS(
  umap_results_with_kmeans,
  file = file.path(data_dir, paste0("umap_results_with_kmeans_k", kmeans_k_to_compare, ".rds"))
)

cat("\nPASS: K-means assignments merged with UMAP results.\n")

```

## Plot UMAPs by Fixed k-means Cluster

```{r}

plot_umap_kmeans <- function(df, analysis_name) {
  
  plot_data <- df %>%
    filter(Analysis == analysis_name)
  
  p <- ggplot(
    plot_data,
    aes(
      x = UMAP1,
      y = UMAP2,
      color = kmeans_cluster
    )
  ) +
    geom_point(size = 0.45, alpha = 0.65) +
    labs(
      title = paste0("UMAP: ", analysis_name),
      subtitle = paste0("Colored by existing fixed min-max k-means cluster, k = ", kmeans_k_to_compare),
      x = "UMAP1",
      y = "UMAP2",
      color = "k-means cluster"
    ) +
    theme_bw(base_size = 13)
  
  return(p)
}

for (analysis_name in drift_umap_analyses) {
  
  p <- plot_umap_kmeans(umap_results_with_kmeans, analysis_name)
  
  print(p)
  
  ggsave(
    file.path(
      plot_dir,
      paste0("umap_", analysis_name, "_by_kmeans_k", kmeans_k_to_compare, ".png")
    ),
    p,
    width = 8,
    height = 6,
    dpi = 300
  )
}

```

## PCA Score Plots Colored by ICI and k-means Cluster

```{r}

# ------------------------------------------------------------
# Merge k-means labels into sampled PCA metadata
# ------------------------------------------------------------

sampled_clicks_pca_kmeans <- sampled_clicks_with_pcs %>%
  mutate(UID = as.character(UID)) %>%
  left_join(
    kmeans_assignments,
    by = c("EventID", "UID")
  )

cat("Rows in sampled PCA metadata:", nrow(sampled_clicks_with_pcs), "\n")
cat("Rows after k-means merge:", nrow(sampled_clicks_pca_kmeans), "\n")
cat("Rows missing k-means cluster:",
    sum(is.na(sampled_clicks_pca_kmeans$kmeans_cluster)), "\n")

if (nrow(sampled_clicks_pca_kmeans) != nrow(sampled_clicks_with_pcs)) {
  stop("FAIL: k-means merge changed row count.")
}

# ------------------------------------------------------------
# PC1 vs PC2 colored by logICI
# ------------------------------------------------------------

p_pc12_logici <- ggplot(
  sampled_clicks_pca_kmeans,
  aes(
    x = PC1,
    y = PC2,
    color = log10_ICI_to_next_s
  )
) +
  geom_point(size = 0.45, alpha = 0.55) +
  scale_color_viridis_c(option = "plasma") +
  labs(
    title = "PCA Scores: PC1 vs PC2",
    subtitle = "Colored by log10 ICI (seconds)",
    x = "PC1",
    y = "PC2",
    color = "log10 ICI"
  ) +
  theme_bw(base_size = 13)

print(p_pc12_logici)

ggsave(
  file.path(plot_dir, "pca_pc1_pc2_colored_by_logici.png"),
  p_pc12_logici,
  width = 7,
  height = 6,
  dpi = 300
)

# ------------------------------------------------------------
# PC1 vs PC2 colored by existing k-means cluster
# ------------------------------------------------------------

p_pc12_kmeans <- ggplot(
  sampled_clicks_pca_kmeans,
  aes(
    x = PC1,
    y = PC2,
    color = kmeans_cluster
  )
) +
  geom_point(size = 0.45, alpha = 0.6) +
  labs(
    title = "PCA Scores: PC1 vs PC2",
    subtitle = paste0(
      "Colored by existing fixed min-max k-means cluster, k = ",
      kmeans_k_to_compare
    ),
    x = "PC1",
    y = "PC2",
    color = "k-means cluster"
  ) +
  theme_bw(base_size = 13)

print(p_pc12_kmeans)

ggsave(
  file.path(plot_dir, paste0("pca_pc1_pc2_colored_by_kmeans_k", kmeans_k_to_compare, ".png")),
  p_pc12_kmeans,
  width = 7,
  height = 6,
  dpi = 300
)

# ------------------------------------------------------------
# PC1 vs PC3 colored by k-means cluster
# ------------------------------------------------------------

p_pc13_kmeans <- ggplot(
  sampled_clicks_pca_kmeans,
  aes(
    x = PC1,
    y = PC3,
    color = kmeans_cluster
  )
) +
  geom_point(size = 0.45, alpha = 0.6) +
  labs(
    title = "PCA Scores: PC1 vs PC3",
    subtitle = paste0(
      "Colored by existing fixed min-max k-means cluster, k = ",
      kmeans_k_to_compare
    ),
    x = "PC1",
    y = "PC3",
    color = "k-means cluster"
  ) +
  theme_bw(base_size = 13)

print(p_pc13_kmeans)

ggsave(
  file.path(plot_dir, paste0("pca_pc1_pc3_colored_by_kmeans_k", kmeans_k_to_compare, ".png")),
  p_pc13_kmeans,
  width = 7,
  height = 6,
  dpi = 300
)

cat("\nPASS: PCA diagnostic plots saved.\n")

```

## Save and Plot PCA Loading Vectors

```{r}

# ------------------------------------------------------------
# Extract PCA loadings
# ------------------------------------------------------------

pca_loadings <- as.data.frame(pca_spectra$rotation)

pca_loadings <- pca_loadings %>%
  mutate(
    Frequency_kHz = freq_focus,
    FreqBin = row_number()
  ) %>%
  relocate(FreqBin, Frequency_kHz)

# ------------------------------------------------------------
# Save full loading table
# ------------------------------------------------------------

write.csv(
  pca_loadings,
  file = file.path(table_dir, "pca_loading_vectors_by_frequency.csv"),
  row.names = FALSE
)

saveRDS(
  pca_loadings,
  file = file.path(data_dir, "pca_loading_vectors_by_frequency.rds")
)

# ------------------------------------------------------------
# Plot first 5 loading vectors
# ------------------------------------------------------------

pca_loadings_long <- pca_loadings %>%
  select(Frequency_kHz, PC1, PC2, PC3, PC4, PC5) %>%
  pivot_longer(
    cols = starts_with("PC"),
    names_to = "PC",
    values_to = "Loading"
  )

p_loadings <- ggplot(
  pca_loadings_long,
  aes(
    x = Frequency_kHz,
    y = Loading
  )
) +
  geom_line(linewidth = 0.8) +
  facet_wrap(~ PC, scales = "free_y", ncol = 1) +
  labs(
    title = "PCA Loading Vectors for First 5 Principal Components",
    subtitle = "Loadings show which frequency bands contribute to each spectral PC",
    x = "Frequency (kHz)",
    y = "Loading"
  ) +
  theme_bw(base_size = 13)

print(p_loadings)

ggsave(
  filename = file.path(plot_dir, "pca_loading_vectors_first_5_pcs.png"),
  plot = p_loadings,
  width = 8,
  height = 10,
  dpi = 300
)

cat("Saved PCA loading vectors and loading plot.\n")

```

## Run HDBSCAN on Primary UMAP Analyses

```{r}

# ------------------------------------------------------------
# Helper function: run HDBSCAN on one UMAP space
# ------------------------------------------------------------

run_hdbscan_analysis <- function(df, analysis_name, minPts = hdbscan_minPts) {
  
  cat("\n============================\n")
  cat("Running HDBSCAN:", analysis_name, "\n")
  cat("============================\n")
  
  umap_subset <- df %>%
    filter(Analysis == analysis_name)
  
  umap_matrix <- umap_subset %>%
    select(UMAP1, UMAP2) %>%
    as.matrix()
  
  hdbscan_result <- dbscan::hdbscan(
    umap_matrix,
    minPts = minPts
  )
  
  output_df <- umap_subset %>%
    mutate(
      HDBSCAN_cluster = factor(hdbscan_result$cluster),
      HDBSCAN_membership_probability = hdbscan_result$membership_prob
    )
  
  cluster_summary <- output_df %>%
    count(HDBSCAN_cluster, name = "n_clicks") %>%
    mutate(
      percent_clicks = 100 * n_clicks / sum(n_clicks),
      cluster_type = if_else(
        HDBSCAN_cluster == "0",
        "noise_unassigned",
        "cluster"
      )
    ) %>%
    arrange(desc(n_clicks))
  
  cat("\nCluster summary:\n")
  print(cluster_summary)
  
  cat("\nNumber of clusters excluding noise:",
      sum(cluster_summary$HDBSCAN_cluster != "0"), "\n")
  
  cat("Percent noise:",
      cluster_summary$percent_clicks[cluster_summary$HDBSCAN_cluster == "0"], "\n")
  
  return(list(
    result_df = output_df,
    hdbscan_object = hdbscan_result,
    cluster_summary = cluster_summary
  ))
}

# ------------------------------------------------------------
# Run primary HDBSCAN analyses
# ------------------------------------------------------------

hdbscan_5pc_only <- run_hdbscan_analysis(
  df = umap_results_with_kmeans,
  analysis_name = "5PC_spectra_only",
  minPts = hdbscan_minPts
)

hdbscan_5pc_ici <- run_hdbscan_analysis(
  df = umap_results_with_kmeans,
  analysis_name = "5PC_plus_scaled_logICI",
  minPts = hdbscan_minPts
)

# ------------------------------------------------------------
# Combine HDBSCAN results
# ------------------------------------------------------------

hdbscan_results_all <- bind_rows(
  hdbscan_5pc_only$result_df,
  hdbscan_5pc_ici$result_df
)

# ------------------------------------------------------------
# Save outputs
# ------------------------------------------------------------

saveRDS(
  hdbscan_results_all,
  file = file.path(data_dir, "hdbscan_results_primary_analyses.rds")
)

write.csv(
  hdbscan_results_all,
  file = file.path(table_dir, "hdbscan_results_primary_analyses.csv"),
  row.names = FALSE
)

write.csv(
  hdbscan_5pc_only$cluster_summary,
  file = file.path(table_dir, "hdbscan_summary_5pc_spectra_only.csv"),
  row.names = FALSE
)

write.csv(
  hdbscan_5pc_ici$cluster_summary,
  file = file.path(table_dir, "hdbscan_summary_5pc_plus_logici.csv"),
  row.names = FALSE
)

cat("\nPASS: HDBSCAN completed for primary analyses.\n")

```

## Plot HDBSCAN Clusters on UMAP

```{r}

# ------------------------------------------------------------
# Plot helper
# ------------------------------------------------------------

plot_hdbscan_umap <- function(df, analysis_name) {
  
  plot_data <- df %>%
    filter(Analysis == analysis_name)
  
  p <- ggplot(
    plot_data,
    aes(
      x = UMAP1,
      y = UMAP2,
      color = HDBSCAN_cluster
    )
  ) +
    geom_point(
      size = 0.5,
      alpha = 0.65
    ) +
    labs(
      title = paste0("HDBSCAN Clusters: ", analysis_name),
      subtitle = "Cluster 0 = noise/unassigned",
      x = "UMAP1",
      y = "UMAP2",
      color = "HDBSCAN\ncluster"
    ) +
    theme_bw(base_size = 13)
  
  return(p)
}

# ------------------------------------------------------------
# Generate plots
# ------------------------------------------------------------

p_hdbscan_5pc_only <- plot_hdbscan_umap(
  hdbscan_results_all,
  "5PC_spectra_only"
)

p_hdbscan_5pc_ici <- plot_hdbscan_umap(
  hdbscan_results_all,
  "5PC_plus_scaled_logICI"
)

# ------------------------------------------------------------
# Print plots
# ------------------------------------------------------------

print(p_hdbscan_5pc_only)
print(p_hdbscan_5pc_ici)

# ------------------------------------------------------------
# Save plots
# ------------------------------------------------------------

ggsave(
  file.path(plot_dir, "hdbscan_umap_5pc_spectra_only.png"),
  p_hdbscan_5pc_only,
  width = 8,
  height = 6,
  dpi = 300
)

ggsave(
  file.path(plot_dir, "hdbscan_umap_5pc_plus_logici.png"),
  p_hdbscan_5pc_ici,
  width = 8,
  height = 6,
  dpi = 300
)

cat("Saved HDBSCAN UMAP plots.\n")

```

## Summarize ICI by HDBSCAN Cluster

```{r}

# ------------------------------------------------------------
# Focus on primary joint spectral-temporal analysis
# ------------------------------------------------------------

joint_clusters <- hdbscan_results_all %>%
  filter(Analysis == "5PC_plus_scaled_logICI")

# ------------------------------------------------------------
# ICI summary by cluster
# ------------------------------------------------------------

ici_summary_by_cluster <- joint_clusters %>%
  group_by(HDBSCAN_cluster) %>%
  summarise(
    
    n_clicks = n(),
    
    median_ici_s = median(ICI_to_next_s, na.rm = TRUE),
    
    mean_ici_s = mean(ICI_to_next_s, na.rm = TRUE),
    
    sd_ici_s = sd(ICI_to_next_s, na.rm = TRUE),
    
    min_ici_s = min(ICI_to_next_s, na.rm = TRUE),
    
    max_ici_s = max(ICI_to_next_s, na.rm = TRUE),
    
    q25_ici_s = quantile(ICI_to_next_s, 0.25, na.rm = TRUE),
    
    q75_ici_s = quantile(ICI_to_next_s, 0.75, na.rm = TRUE),
    
    n_events = n_distinct(EventID),
    
    n_drifts = n_distinct(DriftID_spectra),
    
    .groups = "drop"
    
  ) %>%
  arrange(desc(n_clicks))

print(ici_summary_by_cluster)

write.csv(
  ici_summary_by_cluster,
  file = file.path(table_dir, "ici_summary_by_hdbscan_cluster.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# Boxplot of ICI by cluster
# ------------------------------------------------------------

p_ici_boxplot <- ggplot(
  joint_clusters,
  aes(
    x = HDBSCAN_cluster,
    y = ICI_to_next_s,
    fill = HDBSCAN_cluster
  )
) +
  geom_boxplot(
    outlier.size = 0.3
  ) +
  scale_y_log10() +
  labs(
    title = "ICI Distribution by HDBSCAN Cluster",
    subtitle = "5PC + scaled logICI analysis",
    x = "HDBSCAN cluster",
    y = "ICI (seconds, log scale)"
  ) +
  theme_bw(base_size = 13)

print(p_ici_boxplot)

ggsave(
  file.path(plot_dir, "ici_boxplot_by_hdbscan_cluster.png"),
  p_ici_boxplot,
  width = 8,
  height = 6,
  dpi = 300
)

cat("Saved ICI summaries and boxplot.\n")

```

## Mean Spectra by HDBSCAN Cluster

```{r}

# ------------------------------------------------------------
# Focus on joint spectral-temporal HDBSCAN result
# ------------------------------------------------------------

joint_clusters <- hdbscan_results_all %>%
  filter(Analysis == "5PC_plus_scaled_logICI") %>%
  mutate(
    RowID = as.integer(RowID),
    HDBSCAN_cluster = factor(HDBSCAN_cluster)
  )

# ------------------------------------------------------------
# Pull sampled spectra for these clustered clicks
# ------------------------------------------------------------

joint_spectra <- spectra_matrix_minmax[
  joint_clusters$RowID,
  ,
  drop = FALSE
]

if (nrow(joint_spectra) != nrow(joint_clusters)) {
  stop("FAIL: Joint spectra rows do not match joint cluster metadata.")
}

# ------------------------------------------------------------
# Compute mean and SD spectrum by HDBSCAN cluster
# ------------------------------------------------------------

cluster_spectra_summary <- list()

for (cluster_id in levels(joint_clusters$HDBSCAN_cluster)) {
  
  cluster_rows <- which(joint_clusters$HDBSCAN_cluster == cluster_id)
  
  cluster_matrix <- joint_spectra[
    cluster_rows,
    ,
    drop = FALSE
  ]
  
  cluster_spectra_summary[[cluster_id]] <- data.frame(
    HDBSCAN_cluster = cluster_id,
    Frequency_kHz = freq_focus,
    mean_minmax_spectrum = colMeans(cluster_matrix, na.rm = TRUE),
    sd_minmax_spectrum = apply(cluster_matrix, 2, sd, na.rm = TRUE),
    n_clicks = length(cluster_rows)
  )
}

cluster_spectra_summary <- bind_rows(cluster_spectra_summary)

write.csv(
  cluster_spectra_summary,
  file = file.path(table_dir, "mean_spectra_by_hdbscan_cluster.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# Plot all cluster mean spectra together
# ------------------------------------------------------------

p_cluster_mean_spectra <- ggplot(
  cluster_spectra_summary,
  aes(
    x = Frequency_kHz,
    y = mean_minmax_spectrum,
    color = HDBSCAN_cluster,
    group = HDBSCAN_cluster
  )
) +
  geom_line(linewidth = 1) +
  labs(
    title = "Mean Min-Max Normalized Spectra by HDBSCAN Cluster",
    subtitle = "5PC + scaled logICI analysis; cluster 0 = noise/unassigned",
    x = "Frequency (kHz)",
    y = "Mean min-max normalized amplitude",
    color = "HDBSCAN\ncluster"
  ) +
  theme_bw(base_size = 13)

print(p_cluster_mean_spectra)

ggsave(
  file.path(plot_dir, "mean_spectra_by_hdbscan_cluster_all.png"),
  p_cluster_mean_spectra,
  width = 8,
  height = 6,
  dpi = 300
)

# ------------------------------------------------------------
# Faceted plot by cluster
# ------------------------------------------------------------

p_cluster_mean_spectra_faceted <- ggplot(
  cluster_spectra_summary,
  aes(
    x = Frequency_kHz,
    y = mean_minmax_spectrum
  )
) +
  geom_line(linewidth = 1) +
  facet_wrap(
    ~ HDBSCAN_cluster,
    ncol = 2
  ) +
  labs(
    title = "Mean Min-Max Normalized Spectra by HDBSCAN Cluster",
    subtitle = "Faceted by cluster; cluster 0 = noise/unassigned",
    x = "Frequency (kHz)",
    y = "Mean min-max normalized amplitude"
  ) +
  theme_bw(base_size = 13)

print(p_cluster_mean_spectra_faceted)

ggsave(
  file.path(plot_dir, "mean_spectra_by_hdbscan_cluster_faceted.png"),
  p_cluster_mean_spectra_faceted,
  width = 10,
  height = 8,
  dpi = 300
)

cat("Saved mean spectra by HDBSCAN cluster.\n")

```

## Stacked Mean Spectra by HDBSCAN Cluster

```{r}

# ------------------------------------------------------------
# Create vertically offset cluster mean spectra
# ------------------------------------------------------------

stacked_cluster_spectra <- cluster_spectra_summary %>%
  mutate(
    HDBSCAN_cluster_numeric = as.numeric(as.character(HDBSCAN_cluster)),
    cluster_label = paste0("Cluster ", HDBSCAN_cluster),
    offset = HDBSCAN_cluster_numeric,
    stacked_spectrum = mean_minmax_spectrum + offset
  )

# Optional: remove noise cluster 0 from main stacked figure
stacked_cluster_spectra_no_noise <- stacked_cluster_spectra %>%
  filter(HDBSCAN_cluster != "0")

# ------------------------------------------------------------
# Tall/narrow stacked plot
# ------------------------------------------------------------

p_stacked_spectra <- ggplot(
  stacked_cluster_spectra_no_noise,
  aes(
    x = Frequency_kHz,
    y = stacked_spectrum,
    group = HDBSCAN_cluster,
    color = HDBSCAN_cluster
  )
) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(
    breaks = sort(unique(stacked_cluster_spectra_no_noise$offset)),
    labels = sort(unique(stacked_cluster_spectra_no_noise$HDBSCAN_cluster))
  ) +
  labs(
    title = "Stacked Mean Spectra by HDBSCAN Cluster",
    subtitle = "5PC + scaled logICI analysis; noise cluster removed",
    x = "Frequency (kHz)",
    y = "HDBSCAN cluster"
  ) +
  theme_classic(base_size = 13) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold"),
    axis.title.y = element_text(margin = margin(r = 10))
  )

print(p_stacked_spectra)

ggsave(
  filename = file.path(plot_dir, "stacked_mean_spectra_by_hdbscan_cluster_no_noise.png"),
  plot = p_stacked_spectra,
  width = 4,
  height = 9,
  dpi = 300
)

# ------------------------------------------------------------
# Also save a version including noise cluster 0
# ------------------------------------------------------------

p_stacked_spectra_with_noise <- ggplot(
  stacked_cluster_spectra,
  aes(
    x = Frequency_kHz,
    y = stacked_spectrum,
    group = HDBSCAN_cluster,
    color = HDBSCAN_cluster
  )
) +
  geom_line(linewidth = 1.1) +
  scale_y_continuous(
    breaks = sort(unique(stacked_cluster_spectra$offset)),
    labels = sort(unique(stacked_cluster_spectra$HDBSCAN_cluster))
  ) +
  labs(
    title = "Stacked Mean Spectra by HDBSCAN Cluster",
    subtitle = "5PC + scaled logICI analysis; cluster 0 = noise/unassigned",
    x = "Frequency (kHz)",
    y = "HDBSCAN cluster"
  ) +
  theme_classic(base_size = 13) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold"),
    axis.title.y = element_text(margin = margin(r = 10))
  )

print(p_stacked_spectra_with_noise)

ggsave(
  filename = file.path(plot_dir, "stacked_mean_spectra_by_hdbscan_cluster_with_noise.png"),
  plot = p_stacked_spectra_with_noise,
  width = 4,
  height = 9,
  dpi = 300
)

cat("Saved stacked mean spectra figures.\n")

```

## Concatenated Spectrograms by HDBSCAN Cluster

```{r}

set.seed(cluster_spectrogram_seed)

# ------------------------------------------------------------
# Output folder
# ------------------------------------------------------------

concat_spec_dir <- file.path(
  plot_dir,
  "concatenated_spectrograms_by_hdbscan_cluster_baseR"
)

dir.create(concat_spec_dir, recursive = TRUE, showWarnings = FALSE)

# ------------------------------------------------------------
# Color palette matching previous min-max clustering workflow
# ------------------------------------------------------------

spec_cols <- colorRampPalette(
  c("white", "#FFD54F", "#FF9800", "#D84315", "#4E0012")
)(200)

# ------------------------------------------------------------
# Use joint spectral-temporal HDBSCAN result
# Exclude cluster 0 because it is noise/unassigned
# ------------------------------------------------------------

cluster_ids <- joint_clusters %>%
  filter(HDBSCAN_cluster != "0") %>%
  pull(HDBSCAN_cluster) %>%
  unique() %>%
  as.character() %>%
  sort()

cat("Clusters to plot:\n")
print(cluster_ids)

# ------------------------------------------------------------
# Generate one concatenated spectrogram per HDBSCAN cluster
# ------------------------------------------------------------

for (cl in cluster_ids) {
  
  cluster_metadata <- joint_clusters %>%
    filter(HDBSCAN_cluster == cl)
  
  n_total_cluster <- nrow(cluster_metadata)
  
  if (n_total_cluster == 0) next
  
  # Sample up to max_clicks_per_cluster_spectrogram clicks.
  # This is NOT ordered by ICI; it is random sampled order.
  cluster_metadata_plot <- cluster_metadata %>%
    slice_sample(
      n = min(max_clicks_per_cluster_spectrogram, n_total_cluster),
      replace = FALSE
    )
  
  cluster_spec <- spectra_matrix_minmax[
    cluster_metadata_plot$RowID,
    ,
    drop = FALSE
  ]
  
  median_ici <- median(cluster_metadata$ICI_to_next_s, na.rm = TRUE)
  mean_ici <- mean(cluster_metadata$ICI_to_next_s, na.rm = TRUE)
  n_events <- n_distinct(cluster_metadata$EventID)
  n_drifts <- n_distinct(cluster_metadata$DriftID_spectra)
  
  out_file <- file.path(
    concat_spec_dir,
    paste0(
      "concatenated_spectrogram_hdbscan_cluster_",
      cl,
      "_5PC_plus_scaled_logICI.png"
    )
  )
  
  png(out_file, width = 1800, height = 900)
  
  image(
    x = seq_len(nrow(cluster_spec)),
    y = freq_focus,
    z = cluster_spec,
    col = spec_cols,
    xlab = "Click Number",
    ylab = "Frequency (kHz)",
    main = paste0(
      "Concatenated Click Spectrogram - HDBSCAN Cluster ",
      cl,
      "\n5PC + scaled logICI; n = ", n_total_cluster,
      "; plotted = ", nrow(cluster_metadata_plot),
      "; median ICI = ", round(median_ici, 3), " s",
      "; mean ICI = ", round(mean_ici, 3), " s",
      "; ", n_events, " events; ", n_drifts, " drifts"
    ),
    useRaster = TRUE,
    axes = FALSE
  )
  
  axis(1)
  axis(2, at = seq(20, 50, by = 10))
  box()
  
  dev.off()
  
  cat("Saved:", out_file, "\n")
}

cat("\nFinished concatenated spectrograms for HDBSCAN clusters.\n")
cat("Output folder:\n")
cat(concat_spec_dir, "\n")

```

## Check HDBSCAN Cluster Composition by Event and Drift

```{r}

# ------------------------------------------------------------
# Focus on joint spectral-temporal HDBSCAN result
# ------------------------------------------------------------

joint_clusters <- hdbscan_results_all %>%
  filter(Analysis == "5PC_plus_scaled_logICI") %>%
  mutate(
    HDBSCAN_cluster = factor(HDBSCAN_cluster)
  )

# ------------------------------------------------------------
# Overall cluster coverage summary
# ------------------------------------------------------------

cluster_coverage_summary <- joint_clusters %>%
  group_by(HDBSCAN_cluster) %>%
  summarise(
    n_clicks = n(),
    percent_clicks = 100 * n_clicks / nrow(joint_clusters),
    n_events = n_distinct(EventID),
    n_drifts = n_distinct(DriftID_spectra),
    median_ici_s = median(ICI_to_next_s, na.rm = TRUE),
    mean_ici_s = mean(ICI_to_next_s, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(n_clicks))

cat("Cluster coverage summary:\n")
print(cluster_coverage_summary)

write.csv(
  cluster_coverage_summary,
  file = file.path(table_dir, "hdbscan_cluster_coverage_summary.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# Cluster composition by EventID
# ------------------------------------------------------------

cluster_event_composition <- joint_clusters %>%
  count(HDBSCAN_cluster, EventID, name = "n_clicks") %>%
  group_by(HDBSCAN_cluster) %>%
  mutate(
    cluster_total_clicks = sum(n_clicks),
    percent_of_cluster = 100 * n_clicks / cluster_total_clicks
  ) %>%
  ungroup() %>%
  arrange(HDBSCAN_cluster, desc(n_clicks))

cat("\nCluster composition by EventID:\n")
print(cluster_event_composition, n = 50)

write.csv(
  cluster_event_composition,
  file = file.path(table_dir, "hdbscan_cluster_composition_by_event.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# Top contributing events per cluster
# ------------------------------------------------------------

top_events_per_cluster <- cluster_event_composition %>%
  group_by(HDBSCAN_cluster) %>%
  slice_max(
    order_by = n_clicks,
    n = 5,
    with_ties = FALSE
  ) %>%
  ungroup() %>%
  arrange(HDBSCAN_cluster, desc(n_clicks))

cat("\nTop 5 contributing events per cluster:\n")
print(top_events_per_cluster, n = Inf)

write.csv(
  top_events_per_cluster,
  file = file.path(table_dir, "hdbscan_top_events_per_cluster.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# Cluster composition by DriftID
# ------------------------------------------------------------

cluster_drift_composition <- joint_clusters %>%
  count(HDBSCAN_cluster, DriftID_spectra, name = "n_clicks") %>%
  group_by(HDBSCAN_cluster) %>%
  mutate(
    cluster_total_clicks = sum(n_clicks),
    percent_of_cluster = 100 * n_clicks / cluster_total_clicks
  ) %>%
  ungroup() %>%
  arrange(HDBSCAN_cluster, desc(n_clicks))

cat("\nCluster composition by DriftID:\n")
print(cluster_drift_composition, n = 50)

write.csv(
  cluster_drift_composition,
  file = file.path(table_dir, "hdbscan_cluster_composition_by_drift.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# Top contributing drifts per cluster
# ------------------------------------------------------------

top_drifts_per_cluster <- cluster_drift_composition %>%
  group_by(HDBSCAN_cluster) %>%
  slice_max(
    order_by = n_clicks,
    n = 5,
    with_ties = FALSE
  ) %>%
  ungroup() %>%
  arrange(HDBSCAN_cluster, desc(n_clicks))

cat("\nTop 5 contributing drifts per cluster:\n")
print(top_drifts_per_cluster, n = Inf)

write.csv(
  top_drifts_per_cluster,
  file = file.path(table_dir, "hdbscan_top_drifts_per_cluster.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# Flag potentially event-dominated clusters
# ------------------------------------------------------------

event_dominance_summary <- cluster_event_composition %>%
  group_by(HDBSCAN_cluster) %>%
  summarise(
    top_event = EventID[which.max(n_clicks)],
    top_event_clicks = max(n_clicks),
    top_event_percent = max(percent_of_cluster),
    .groups = "drop"
  ) %>%
  arrange(desc(top_event_percent))

cat("\nEvent dominance summary:\n")
print(event_dominance_summary)

write.csv(
  event_dominance_summary,
  file = file.path(table_dir, "hdbscan_event_dominance_summary.csv"),
  row.names = FALSE
)

drift_dominance_summary <- cluster_drift_composition %>%
  group_by(HDBSCAN_cluster) %>%
  summarise(
    top_drift = DriftID_spectra[which.max(n_clicks)],
    top_drift_clicks = max(n_clicks),
    top_drift_percent = max(percent_of_cluster),
    .groups = "drop"
  ) %>%
  arrange(desc(top_drift_percent))

cat("\nDrift dominance summary:\n")
print(drift_dominance_summary)

write.csv(
  drift_dominance_summary,
  file = file.path(table_dir, "hdbscan_drift_dominance_summary.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# Simple bar plot: top events per cluster
# ------------------------------------------------------------

p_top_events <- ggplot(
  top_events_per_cluster,
  aes(
    x = reorder(EventID, n_clicks),
    y = n_clicks,
    fill = HDBSCAN_cluster
  )
) +
  geom_col() +
  coord_flip() +
  facet_wrap(
    ~ HDBSCAN_cluster,
    scales = "free_y"
  ) +
  labs(
    title = "Top Contributing Events per HDBSCAN Cluster",
    subtitle = "5PC + scaled logICI analysis",
    x = "EventID",
    y = "Number of clicks"
  ) +
  theme_bw(base_size = 12) +
  theme(
    legend.position = "none"
  )

print(p_top_events)

ggsave(
  filename = file.path(plot_dir, "top_events_per_hdbscan_cluster.png"),
  plot = p_top_events,
  width = 12,
  height = 8,
  dpi = 300
)

# ------------------------------------------------------------
# Simple bar plot: top drifts per cluster
# ------------------------------------------------------------

p_top_drifts <- ggplot(
  top_drifts_per_cluster,
  aes(
    x = reorder(DriftID_spectra, n_clicks),
    y = n_clicks,
    fill = HDBSCAN_cluster
  )
) +
  geom_col() +
  coord_flip() +
  facet_wrap(
    ~ HDBSCAN_cluster,
    scales = "free_y"
  ) +
  labs(
    title = "Top Contributing Drifts per HDBSCAN Cluster",
    subtitle = "5PC + scaled logICI analysis",
    x = "DriftID",
    y = "Number of clicks"
  ) +
  theme_bw(base_size = 12) +
  theme(
    legend.position = "none"
  )

print(p_top_drifts)

ggsave(
  filename = file.path(plot_dir, "top_drifts_per_hdbscan_cluster.png"),
  plot = p_top_drifts,
  width = 12,
  height = 8,
  dpi = 300
)

cat("\nPASS: Cluster composition diagnostics completed and saved.\n")

```

## Mean Spectra with Variability Ribbons for All Non-Noise Clusters

```{r}

# ------------------------------------------------------------
# Use non-noise HDBSCAN clusters from joint analysis
# ------------------------------------------------------------

cluster_spectra_ribbon_data <- cluster_spectra_summary %>%
  filter(HDBSCAN_cluster != "0") %>%
  mutate(
    lower_sd = pmax(mean_minmax_spectrum - sd_minmax_spectrum, 0),
    upper_sd = pmin(mean_minmax_spectrum + sd_minmax_spectrum, 1)
  )

# ------------------------------------------------------------
# Faceted mean ± SD ribbons
# ------------------------------------------------------------

p_cluster_spectra_ribbons <- ggplot(
  cluster_spectra_ribbon_data,
  aes(
    x = Frequency_kHz,
    y = mean_minmax_spectrum
  )
) +
  geom_ribbon(
    aes(
      ymin = lower_sd,
      ymax = upper_sd
    ),
    alpha = 0.25
  ) +
  geom_line(linewidth = 1) +
  facet_wrap(
    ~ HDBSCAN_cluster,
    ncol = 2
  ) +
  labs(
    title = "Mean Spectra with Variability by HDBSCAN Cluster",
    subtitle = "Mean ± 1 SD; 5PC + scaled logICI analysis; noise cluster excluded",
    x = "Frequency (kHz)",
    y = "Min-max normalized amplitude"
  ) +
  theme_bw(base_size = 13)

print(p_cluster_spectra_ribbons)

ggsave(
  filename = file.path(plot_dir, "mean_spectra_with_sd_ribbons_hdbscan_clusters_1_to_6.png"),
  plot = p_cluster_spectra_ribbons,
  width = 10,
  height = 8,
  dpi = 300
)

# ------------------------------------------------------------
# Save ribbon data
# ------------------------------------------------------------

write.csv(
  cluster_spectra_ribbon_data,
  file = file.path(table_dir, "mean_spectra_with_sd_ribbons_hdbscan_clusters_1_to_6.csv"),
  row.names = FALSE
)

cat("Saved mean spectra with SD ribbons for all non-noise clusters.\n")

```

----------------------------------------------

## Full Dataset Assignment

## Force-Assign All Eligible Clicks to Nearest HDBSCAN cluster

```{r}

# ------------------------------------------------------------
# Output folder for full-dataset assignment
# ------------------------------------------------------------

full_assignment_dir <- file.path(
  "E:/Analysis/PCA_ICI_UMAP_HDBSCAN",
  "Full_Eligible_Click_Assignment_1x_scaled_logICI"
)

full_assignment_dirs <- make_output_folders(full_assignment_dir)

# ------------------------------------------------------------
# Recreate full eligible click table
# These are all clicks with ICI between 0.01 and 0.4 s
# ------------------------------------------------------------

all_eligible_clicks <- spectra_metadata_with_ici %>%
  mutate(
    RowID = row_number()
  ) %>%
  filter(
    ICI_filter_category == "kept_0.01_to_0.4_s"
  )

cat("All eligible clicks:", nrow(all_eligible_clicks), "\n")

# ------------------------------------------------------------
# Pull min-max spectra for all eligible clicks
# ------------------------------------------------------------

all_eligible_spectra_minmax <- spectra_matrix_minmax[
  all_eligible_clicks$RowID,
  ,
  drop = FALSE
]

cat("All eligible spectra matrix dimensions:\n")
print(dim(all_eligible_spectra_minmax))

if (nrow(all_eligible_spectra_minmax) != nrow(all_eligible_clicks)) {
  stop("FAIL: Eligible spectra rows do not match eligible metadata rows.")
}

# ------------------------------------------------------------
# Project all eligible clicks into the existing PCA space
# IMPORTANT: This uses the PCA fit from the sampled dataset.
# It does NOT refit PCA.
# ------------------------------------------------------------

all_eligible_pc_scores <- predict(
  pca_spectra,
  newdata = all_eligible_spectra_minmax
)

all_eligible_pc_scores <- as.data.frame(all_eligible_pc_scores)

# ------------------------------------------------------------
# Scale logICI using the SAME mean and SD from the sampled analysis
# IMPORTANT: Do not rescale using the full dataset.
# ------------------------------------------------------------

sampled_logici_mean <- mean(sampled_clicks_with_pcs$log10_ICI_to_next_s, na.rm = TRUE)
sampled_logici_sd <- sd(sampled_clicks_with_pcs$log10_ICI_to_next_s, na.rm = TRUE)

all_eligible_with_pcs <- bind_cols(
  all_eligible_clicks,
  all_eligible_pc_scores
) %>%
  mutate(
    log10_ICI_to_next_s = log10(ICI_to_next_s),
    scaled_log10_ICI = as.numeric(
      (log10_ICI_to_next_s - sampled_logici_mean) / sampled_logici_sd
    )
  )

cat("\nSummary of scaled logICI for all eligible clicks:\n")
print(summary(all_eligible_with_pcs$scaled_log10_ICI))

# ------------------------------------------------------------
# Build full feature matrix: 5 PCs + scaled logICI
# ------------------------------------------------------------

all_eligible_feature_matrix <- all_eligible_with_pcs %>%
  select(all_of(pc_set_5), scaled_log10_ICI) %>%
  as.matrix()

if (any(!is.finite(all_eligible_feature_matrix))) {
  stop("FAIL: Full eligible feature matrix contains non-finite values.")
}

cat("\nFull eligible feature matrix dimensions:\n")
print(dim(all_eligible_feature_matrix))

# ------------------------------------------------------------
# Build cluster centroids from original sampled HDBSCAN labels
# Use only non-noise clusters 1–6
# Centroids are computed in 5PC + scaled logICI feature space.
# ------------------------------------------------------------

sampled_joint_clusters <- hdbscan_results_all %>%
  filter(Analysis == "5PC_plus_scaled_logICI") %>%
  mutate(
    HDBSCAN_cluster_original = as.character(HDBSCAN_cluster)
  )

non_noise_sampled <- sampled_joint_clusters %>%
  filter(HDBSCAN_cluster_original != "0")

cluster_centroids <- non_noise_sampled %>%
  group_by(HDBSCAN_cluster_original) %>%
  summarise(
    across(
      all_of(c(pc_set_5, "scaled_log10_ICI")),
      mean,
      na.rm = TRUE
    ),
    n_sampled_cluster_clicks = n(),
    .groups = "drop"
  ) %>%
  arrange(as.numeric(HDBSCAN_cluster_original))

print(cluster_centroids)

write.csv(
  cluster_centroids,
  file = file.path(
    full_assignment_dirs$tables,
    "hdbscan_cluster_centroids_5pc_scaled_logici.csv"
  ),
  row.names = FALSE
)

centroid_matrix <- cluster_centroids %>%
  select(all_of(c(pc_set_5, "scaled_log10_ICI"))) %>%
  as.matrix()

centroid_labels <- cluster_centroids$HDBSCAN_cluster_original

# ------------------------------------------------------------
# Assign each eligible click to nearest non-noise centroid
# ------------------------------------------------------------

distance_to_centroids <- sapply(
  seq_len(nrow(centroid_matrix)),
  function(j) {
    rowSums(
      sweep(
        all_eligible_feature_matrix,
        2,
        centroid_matrix[j, ],
        FUN = "-"
      )^2
    )
  }
)

nearest_centroid_index <- max.col(
  -distance_to_centroids,
  ties.method = "first"
)

nearest_centroid_distance <- sqrt(
  distance_to_centroids[
    cbind(seq_len(nrow(distance_to_centroids)), nearest_centroid_index)
  ]
)

all_eligible_assigned <- all_eligible_with_pcs %>%
  mutate(
    HDBSCAN_cluster_forced = factor(
      centroid_labels[nearest_centroid_index],
      levels = centroid_labels
    ),
    distance_to_forced_cluster_centroid = nearest_centroid_distance
  )

# ------------------------------------------------------------
# Add original HDBSCAN label for sampled clicks only
# For unsampled eligible clicks, this will be NA.
# ------------------------------------------------------------

sampled_original_labels <- sampled_joint_clusters %>%
  select(
    EventID,
    UID,
    HDBSCAN_cluster_original = HDBSCAN_cluster
  ) %>%
  mutate(
    UID = as.character(UID),
    HDBSCAN_cluster_original = as.character(HDBSCAN_cluster_original)
  )

all_eligible_assigned <- all_eligible_assigned %>%
  mutate(UID = as.character(UID)) %>%
  left_join(
    sampled_original_labels,
    by = c("EventID", "UID")
  )

cat("\nRows after adding sampled original labels:", nrow(all_eligible_assigned), "\n")
cat("Sampled clicks with original HDBSCAN labels:",
    sum(!is.na(all_eligible_assigned$HDBSCAN_cluster_original)), "\n")

# ------------------------------------------------------------
# Save assigned full eligible dataset
# ------------------------------------------------------------

saveRDS(
  all_eligible_assigned,
  file = file.path(
    full_assignment_dirs$data,
    "all_eligible_clicks_forced_hdbscan_assignment_1x.rds"
  )
)

write.csv(
  all_eligible_assigned,
  file = file.path(
    full_assignment_dirs$tables,
    "all_eligible_clicks_forced_hdbscan_assignment_1x.csv"
  ),
  row.names = FALSE
)

# ------------------------------------------------------------
# Summarize forced assignments
# ------------------------------------------------------------

forced_assignment_summary <- all_eligible_assigned %>%
  count(HDBSCAN_cluster_forced, name = "n_clicks") %>%
  mutate(
    percent_clicks = 100 * n_clicks / sum(n_clicks)
  ) %>%
  arrange(desc(n_clicks))

cat("\nForced assignment summary for all eligible clicks:\n")
print(forced_assignment_summary)

write.csv(
  forced_assignment_summary,
  file = file.path(
    full_assignment_dirs$tables,
    "forced_assignment_summary_all_eligible_clicks.csv"
  ),
  row.names = FALSE
)

cat("\nPASS: All eligible clicks assigned to nearest non-noise HDBSCAN cluster.\n")

```



































-------------------------------------------------------

#Sensitivy of ICI

## Set Up Sensitivity Output Folders

```{r}

# ------------------------------------------------------------
# Sensitivity analysis output folders
# ------------------------------------------------------------

sensitivity_base_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN"

main_1x_dir <- file.path(sensitivity_base_dir, "Main_1x_scaled_logICI")
weight_05_dir <- file.path(sensitivity_base_dir, "Reduced_0.5x_scaled_logICI")
weight_025_dir <- file.path(sensitivity_base_dir, "Reduced_0.25x_scaled_logICI")

make_output_folders <- function(base_dir) {
  
  dir.create(base_dir, recursive = TRUE, showWarnings = FALSE)
  
  subdirs <- list(
    plots = file.path(base_dir, "plots"),
    data = file.path(base_dir, "data"),
    tables = file.path(base_dir, "tables")
  )
  
  dir.create(subdirs$plots, recursive = TRUE, showWarnings = FALSE)
  dir.create(subdirs$data, recursive = TRUE, showWarnings = FALSE)
  dir.create(subdirs$tables, recursive = TRUE, showWarnings = FALSE)
  
  return(subdirs)
}

main_1x_dirs <- make_output_folders(main_1x_dir)
weight_05_dirs <- make_output_folders(weight_05_dir)
weight_025_dirs <- make_output_folders(weight_025_dir)

cat("Created sensitivity output folders:\n")
cat(main_1x_dir, "\n")
cat(weight_05_dir, "\n")
cat(weight_025_dir, "\n")

```

## Run Reduced ICI Weight Sensitivity Analysis

```{r}

# ------------------------------------------------------------
# Helper function for one weighted ICI sensitivity run
# ------------------------------------------------------------

run_weighted_ici_sensitivity <- function(
    ici_weight,
    analysis_name,
    output_dirs
) {
  
  cat("\n============================\n")
  cat("Running sensitivity analysis:", analysis_name, "\n")
  cat("ICI weight:", ici_weight, "\n")
  cat("============================\n")
  
  # ----------------------------------------------------------
  # Build weighted feature matrix
  # ----------------------------------------------------------
  
  feature_matrix_weighted <- sampled_clicks_with_pcs %>%
    mutate(
      weighted_scaled_log10_ICI = ici_weight * scaled_log10_ICI
    ) %>%
    select(all_of(pc_set_5), weighted_scaled_log10_ICI) %>%
    as.matrix()
  
  cat("Feature matrix dimensions:\n")
  print(dim(feature_matrix_weighted))
  
  if (any(!is.finite(feature_matrix_weighted))) {
    stop("Weighted feature matrix contains non-finite values.")
  }
  
  # ----------------------------------------------------------
  # Run UMAP
  # ----------------------------------------------------------
  
  umap_weighted <- run_umap_analysis(
    feature_matrix = feature_matrix_weighted,
    analysis_name = analysis_name
  )
  
  umap_weighted_df <- bind_cols(
    sampled_clicks_with_pcs,
    umap_weighted
  ) %>%
    mutate(
      ICI_weight = ici_weight
    )
  
  # ----------------------------------------------------------
  # Run HDBSCAN
  # ----------------------------------------------------------
  
  hdbscan_result <- dbscan::hdbscan(
    umap_weighted_df %>%
      select(UMAP1, UMAP2) %>%
      as.matrix(),
    minPts = hdbscan_minPts
  )
  
  hdbscan_weighted_df <- umap_weighted_df %>%
    mutate(
      HDBSCAN_cluster = factor(hdbscan_result$cluster),
      HDBSCAN_membership_probability = hdbscan_result$membership_prob
    )
  
  # ----------------------------------------------------------
  # Cluster summary
  # ----------------------------------------------------------
  
  cluster_summary <- hdbscan_weighted_df %>%
    count(HDBSCAN_cluster, name = "n_clicks") %>%
    mutate(
      percent_clicks = 100 * n_clicks / sum(n_clicks),
      cluster_type = if_else(
        HDBSCAN_cluster == "0",
        "noise_unassigned",
        "cluster"
      )
    ) %>%
    arrange(desc(n_clicks))
  
  print(cluster_summary)
  
  # ----------------------------------------------------------
  # ICI summary by cluster
  # ----------------------------------------------------------
  
  ici_summary <- hdbscan_weighted_df %>%
    group_by(HDBSCAN_cluster) %>%
    summarise(
      n_clicks = n(),
      percent_clicks = 100 * n_clicks / nrow(hdbscan_weighted_df),
      median_ici_s = median(ICI_to_next_s, na.rm = TRUE),
      mean_ici_s = mean(ICI_to_next_s, na.rm = TRUE),
      q25_ici_s = quantile(ICI_to_next_s, 0.25, na.rm = TRUE),
      q75_ici_s = quantile(ICI_to_next_s, 0.75, na.rm = TRUE),
      n_events = n_distinct(EventID),
      n_drifts = n_distinct(DriftID_spectra),
      .groups = "drop"
    ) %>%
    arrange(desc(n_clicks))
  
  # ----------------------------------------------------------
  # Save data/tables
  # ----------------------------------------------------------
  
  saveRDS(
    feature_matrix_weighted,
    file = file.path(output_dirs$data, paste0("feature_matrix_", analysis_name, ".rds"))
  )
  
  saveRDS(
    umap_weighted_df,
    file = file.path(output_dirs$data, paste0("umap_results_", analysis_name, ".rds"))
  )
  
  saveRDS(
    hdbscan_weighted_df,
    file = file.path(output_dirs$data, paste0("hdbscan_results_", analysis_name, ".rds"))
  )
  
  write.csv(
    cluster_summary,
    file = file.path(output_dirs$tables, paste0("hdbscan_cluster_summary_", analysis_name, ".csv")),
    row.names = FALSE
  )
  
  write.csv(
    ici_summary,
    file = file.path(output_dirs$tables, paste0("ici_summary_by_hdbscan_cluster_", analysis_name, ".csv")),
    row.names = FALSE
  )
  
  # ----------------------------------------------------------
  # Plot UMAP colored by logICI
  # ----------------------------------------------------------
  
  p_logici <- ggplot(
    hdbscan_weighted_df,
    aes(
      x = UMAP1,
      y = UMAP2,
      color = log10_ICI_to_next_s
    )
  ) +
    geom_point(size = 0.45, alpha = 0.55) +
    scale_color_viridis_c(option = "plasma") +
    labs(
      title = paste0("UMAP: ", analysis_name),
      subtitle = paste0("5 PCs + ", ici_weight, " × scaled log10 ICI; colored by log10 ICI"),
      x = "UMAP1",
      y = "UMAP2",
      color = "log10 ICI"
    ) +
    theme_bw(base_size = 13)
  
  print(p_logici)
  
  ggsave(
    filename = file.path(output_dirs$plots, paste0("umap_", analysis_name, "_colored_by_logici.png")),
    plot = p_logici,
    width = 7,
    height = 6,
    dpi = 300
  )
  
  # ----------------------------------------------------------
  # Plot UMAP colored by HDBSCAN cluster
  # ----------------------------------------------------------
  
  p_hdbscan <- ggplot(
    hdbscan_weighted_df,
    aes(
      x = UMAP1,
      y = UMAP2,
      color = HDBSCAN_cluster
    )
  ) +
    geom_point(size = 0.5, alpha = 0.65) +
    labs(
      title = paste0("HDBSCAN Clusters: ", analysis_name),
      subtitle = paste0("5 PCs + ", ici_weight, " × scaled log10 ICI; cluster 0 = noise/unassigned"),
      x = "UMAP1",
      y = "UMAP2",
      color = "HDBSCAN\ncluster"
    ) +
    theme_bw(base_size = 13)
  
  print(p_hdbscan)
  
  ggsave(
    filename = file.path(output_dirs$plots, paste0("hdbscan_umap_", analysis_name, ".png")),
    plot = p_hdbscan,
    width = 8,
    height = 6,
    dpi = 300
  )
  
  # ----------------------------------------------------------
  # ICI boxplot by HDBSCAN cluster
  # ----------------------------------------------------------
  
  p_ici_boxplot <- ggplot(
    hdbscan_weighted_df,
    aes(
      x = HDBSCAN_cluster,
      y = ICI_to_next_s,
      fill = HDBSCAN_cluster
    )
  ) +
    geom_boxplot(outlier.size = 0.3) +
    scale_y_log10() +
    labs(
      title = paste0("ICI Distribution by HDBSCAN Cluster: ", analysis_name),
      subtitle = paste0("5 PCs + ", ici_weight, " × scaled log10 ICI"),
      x = "HDBSCAN cluster",
      y = "ICI (seconds, log scale)"
    ) +
    theme_bw(base_size = 13) +
    theme(legend.position = "none")
  
  print(p_ici_boxplot)
  
  ggsave(
    filename = file.path(output_dirs$plots, paste0("ici_boxplot_by_hdbscan_cluster_", analysis_name, ".png")),
    plot = p_ici_boxplot,
    width = 8,
    height = 6,
    dpi = 300
  )
  
  return(list(
    feature_matrix = feature_matrix_weighted,
    umap_df = umap_weighted_df,
    hdbscan_df = hdbscan_weighted_df,
    hdbscan_object = hdbscan_result,
    cluster_summary = cluster_summary,
    ici_summary = ici_summary
  ))
}

# ------------------------------------------------------------
# Run reduced-weight sensitivity analyses
# ------------------------------------------------------------

sensitivity_05x <- run_weighted_ici_sensitivity(
  ici_weight = 0.5,
  analysis_name = "5PC_plus_0.5x_scaled_logICI",
  output_dirs = weight_05_dirs
)

sensitivity_025x <- run_weighted_ici_sensitivity(
  ici_weight = 0.25,
  analysis_name = "5PC_plus_0.25x_scaled_logICI",
  output_dirs = weight_025_dirs
)

cat("\nPASS: Reduced ICI weight sensitivity analyses complete.\n")

```

## Mean Spectra for Reduced ICI-Weight Clusters

```{r}

make_weighted_cluster_spectra <- function(
    sensitivity_result,
    analysis_name,
    output_dirs
) {
  
  hdbscan_df <- sensitivity_result$hdbscan_df %>%
    mutate(
      RowID = as.integer(RowID),
      HDBSCAN_cluster = factor(HDBSCAN_cluster)
    )
  
  spectra_for_clusters <- spectra_matrix_minmax[
    hdbscan_df$RowID,
    ,
    drop = FALSE
  ]
  
  cluster_ids <- levels(hdbscan_df$HDBSCAN_cluster)
  
  spectra_summary <- list()
  
  for (cl in cluster_ids) {
    
    cluster_rows <- which(hdbscan_df$HDBSCAN_cluster == cl)
    
    cluster_matrix <- spectra_for_clusters[
      cluster_rows,
      ,
      drop = FALSE
    ]
    
    spectra_summary[[cl]] <- data.frame(
      HDBSCAN_cluster = cl,
      Frequency_kHz = freq_focus,
      mean_minmax_spectrum = colMeans(cluster_matrix, na.rm = TRUE),
      sd_minmax_spectrum = apply(cluster_matrix, 2, sd, na.rm = TRUE),
      n_clicks = length(cluster_rows)
    )
  }
  
  spectra_summary <- bind_rows(spectra_summary) %>%
    mutate(
      lower_sd = pmax(mean_minmax_spectrum - sd_minmax_spectrum, 0),
      upper_sd = pmin(mean_minmax_spectrum + sd_minmax_spectrum, 1)
    )
  
  write.csv(
    spectra_summary,
    file = file.path(
      output_dirs$tables,
      paste0("mean_spectra_by_hdbscan_cluster_", analysis_name, ".csv")
    ),
    row.names = FALSE
  )
  
  # ------------------------------------------------------------
  # Overlay plot
  # ------------------------------------------------------------
  
  p_overlay <- ggplot(
    spectra_summary,
    aes(
      x = Frequency_kHz,
      y = mean_minmax_spectrum,
      color = HDBSCAN_cluster,
      group = HDBSCAN_cluster
    )
  ) +
    geom_line(linewidth = 1) +
    labs(
      title = paste0("Mean Spectra by HDBSCAN Cluster: ", analysis_name),
      subtitle = "Cluster 0 = noise/unassigned",
      x = "Frequency (kHz)",
      y = "Mean min-max normalized amplitude",
      color = "HDBSCAN\ncluster"
    ) +
    theme_bw(base_size = 13)
  
  print(p_overlay)
  
  ggsave(
    filename = file.path(
      output_dirs$plots,
      paste0("mean_spectra_overlay_", analysis_name, ".png")
    ),
    plot = p_overlay,
    width = 8,
    height = 6,
    dpi = 300
  )
  
  # ------------------------------------------------------------
  # Faceted mean ± SD plot
  # ------------------------------------------------------------
  
  p_ribbons <- ggplot(
    spectra_summary,
    aes(
      x = Frequency_kHz,
      y = mean_minmax_spectrum
    )
  ) +
    geom_ribbon(
      aes(
        ymin = lower_sd,
        ymax = upper_sd
      ),
      alpha = 0.25
    ) +
    geom_line(linewidth = 1) +
    facet_wrap(
      ~ HDBSCAN_cluster,
      ncol = 2
    ) +
    labs(
      title = paste0("Mean Spectra with Variability: ", analysis_name),
      subtitle = "Mean ± 1 SD; cluster 0 = noise/unassigned",
      x = "Frequency (kHz)",
      y = "Min-max normalized amplitude"
    ) +
    theme_bw(base_size = 13)
  
  print(p_ribbons)
  
  ggsave(
    filename = file.path(
      output_dirs$plots,
      paste0("mean_spectra_with_sd_ribbons_", analysis_name, ".png")
    ),
    plot = p_ribbons,
    width = 10,
    height = 8,
    dpi = 300
  )
  
  return(spectra_summary)
}

spectra_summary_05x <- make_weighted_cluster_spectra(
  sensitivity_result = sensitivity_05x,
  analysis_name = "5PC_plus_0.5x_scaled_logICI",
  output_dirs = weight_05_dirs
)

spectra_summary_025x <- make_weighted_cluster_spectra(
  sensitivity_result = sensitivity_025x,
  analysis_name = "5PC_plus_0.25x_scaled_logICI",
  output_dirs = weight_025_dirs
)

cat("Saved spectra plots for reduced ICI-weight clusters.\n")

```

-------------------------------------------------------

## Reload Existing PCA/UMAP/HDBSCAN Objects

```{r}

## Reload Existing PCA/ICI/UMAP/HDBSCAN Outputs

library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)
library(purrr)
library(tibble)
library(stringr)
library(dbscan)

output_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN"

# Find all RDS files under the workflow folder
all_rds_files <- list.files(
  output_dir,
  pattern = "\\.rds$",
  recursive = TRUE,
  full.names = TRUE
)

cat("RDS files found:\n")
print(all_rds_files)

find_rds <- function(pattern) {
  matches <- all_rds_files[str_detect(basename(all_rds_files), pattern)]
  
  if (length(matches) == 0) {
    stop("No RDS file found matching: ", pattern)
  }
  
  if (length(matches) > 1) {
    cat("Multiple matches found for:", pattern, "\n")
    print(matches)
    cat("Using first match:\n", matches[1], "\n")
  }
  
  matches[1]
}

sampled_clicks_with_pcs <- readRDS(
  find_rds("sampled_clicks_with_pcs_and_scaled_logici")
)

umap_results_all <- readRDS(
  find_rds("umap_results_all_main_comparisons")
)

hdbscan_results_all <- readRDS(
  find_rds("hdbscan_results_primary_analyses")
)

spectra_matrix_sampled_minmax <- readRDS(
  find_rds("sampled_minmax_spectra_matrix_19_50_khz")
)

pca_spectra <- readRDS(
  find_rds("pca_spectra_sampled_minmax_19_50_khz")
)

pca_loadings <- readRDS(
  find_rds("pca_loading_vectors_by_frequency")
)

freq_focus <- pca_loadings$Frequency_kHz

#Sanity check

cat("\nReload check:\n")
cat("sampled_clicks_with_pcs rows:", nrow(sampled_clicks_with_pcs), "\n")
cat("umap_results_all rows:", nrow(umap_results_all), "\n")
cat("hdbscan_results_all rows:", nrow(hdbscan_results_all), "\n")
cat("spectra_matrix_sampled_minmax dim:", dim(spectra_matrix_sampled_minmax), "\n")
cat("freq_focus length:", length(freq_focus), "\n")

cat("\nUMAP analyses available:\n")
print(unique(umap_results_all$Analysis))

cat("\nHDBSCAN analyses available:\n")
print(unique(hdbscan_results_all$Analysis))

```

------------------------------------------------------

## HDBSCAN Tuning

## HDBSCAN Tuning Output Folder

```{r}

tuning_dir <- file.path(
  output_dir,
  "HDBSCAN_Tuning_Safe"
)

tuning_plot_dir <- file.path(
  tuning_dir,
  "plots"
)

tuning_table_dir <- file.path(
  tuning_dir,
  "tables"
)

tuning_data_dir <- file.path(
  tuning_dir,
  "data"
)

dir.create(tuning_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(tuning_plot_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(tuning_table_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(tuning_data_dir, recursive = TRUE, showWarnings = FALSE)

cat("Created safe tuning folder:\n")
cat(tuning_dir)

```

## HDBSCAN Tuning — 5PC + scaled logICI

```{r}

# ------------------------------------------------------------
# minPts values to test
# ------------------------------------------------------------

minPts_values <- c(10, 30, 50)

# ------------------------------------------------------------
# Pull only the primary joint spectral-temporal analysis
# ------------------------------------------------------------

umap_5pc_logici <- umap_results_all %>%
  filter(Analysis == "5PC_plus_scaled_logICI")

cat("Rows in 5PC + scaled logICI UMAP:\n")
print(nrow(umap_5pc_logici))

# ------------------------------------------------------------
# Extract UMAP coordinates
# ------------------------------------------------------------

umap_matrix <- umap_5pc_logici %>%
  select(UMAP1, UMAP2) %>%
  as.matrix()

# ------------------------------------------------------------
# Run HDBSCAN for each minPts value
# ------------------------------------------------------------

hdbscan_tuning_results <- list()

for (m in minPts_values) {
  
  cat("\n============================\n")
  cat("Running HDBSCAN with minPts =", m, "\n")
  cat("============================\n")
  
  hdb <- dbscan::hdbscan(
    umap_matrix,
    minPts = m
  )
  
  result_df <- umap_5pc_logici %>%
    mutate(
      HDBSCAN_cluster = factor(hdb$cluster),
      HDBSCAN_membership_probability = hdb$membership_prob,
      minPts = m
    )
  
  cluster_summary <- result_df %>%
    count(HDBSCAN_cluster, name = "n_clicks") %>%
    mutate(
      percent_clicks = 100 * n_clicks / sum(n_clicks),
      cluster_type = if_else(
        HDBSCAN_cluster == "0",
        "noise_unassigned",
        "cluster"
      )
    ) %>%
    arrange(desc(n_clicks))
  
  cat("\nCluster summary:\n")
  print(cluster_summary)
  
  hdbscan_tuning_results[[paste0("minPts_", m)]] <- result_df
  
  # --------------------------------------------------------
  # Save summary table
  # --------------------------------------------------------
  
  write.csv(
    cluster_summary,
    file.path(
      tuning_table_dir,
      paste0("hdbscan_summary_minPts_", m, ".csv")
    ),
    row.names = FALSE
  )
}

cat("\nPASS: HDBSCAN tuning runs complete.\n")

```

## Compare HDBSCAN UMAP Clusters Across minPts

```{r}

# ------------------------------------------------------------
# Combine tuning outputs into one dataframe
# ------------------------------------------------------------

hdbscan_tuning_combined <- bind_rows(hdbscan_tuning_results)

cat("Rows in combined tuning dataframe:\n")
print(nrow(hdbscan_tuning_combined))

cat("\nminPts values present:\n")
print(unique(hdbscan_tuning_combined$minPts))

# ------------------------------------------------------------
# Plot
# ------------------------------------------------------------

p_hdbscan_tuning_compare <- ggplot(
  hdbscan_tuning_combined,
  aes(
    x = UMAP1,
    y = UMAP2,
    color = HDBSCAN_cluster
  )
) +
  geom_point(
    size = 0.4,
    alpha = 0.7
  ) +
  facet_wrap(
    ~ minPts,
    nrow = 1
  ) +
  labs(
    title = "HDBSCAN Cluster Comparison Across minPts Values",
    subtitle = "5PC + scaled logICI analysis",
    x = "UMAP1",
    y = "UMAP2",
    color = "HDBSCAN\ncluster"
  ) +
  theme_bw(base_size = 13)

print(p_hdbscan_tuning_compare)

# ------------------------------------------------------------
# Save
# ------------------------------------------------------------

ggsave(
  filename = file.path(
    tuning_plot_dir,
    "hdbscan_umap_compare_minPts_10_30_50.png"
  ),
  plot = p_hdbscan_tuning_compare,
  width = 18,
  height = 6,
  dpi = 300
)

cat("\nSaved HDBSCAN comparison plot.\n")

```

## Clean Comparison — minPts 30 vs 50

```{r}

# ------------------------------------------------------------
# Prepare plotting data
# ------------------------------------------------------------

plot_30_50 <- hdbscan_tuning_combined %>%
  filter(minPts %in% c(30, 50)) %>%
  mutate(
    HDBSCAN_cluster = as.character(HDBSCAN_cluster)
  )

# ------------------------------------------------------------
# Plot
# ------------------------------------------------------------

p_30_50_compare <- ggplot(
  plot_30_50,
  aes(
    x = UMAP1,
    y = UMAP2,
    color = HDBSCAN_cluster
  )
) +
  geom_point(
    size = 0.4,
    alpha = 0.75
  ) +
  facet_wrap(
    ~ minPts,
    nrow = 1
  ) +
  
  # --------------------------------------------------------
  # Manual colors preserving cluster labels
  # --------------------------------------------------------
  
  scale_color_manual(
    values = c(
      
      # shared
      "0" = "grey80",
      
      # minPts = 30 major clusters
      "4" = "#17A8E3",  # blue/right
      "5" = "#8E63CE",  # purple/upper
      "6" = "#F15BB5",  # pink/lower
      
      # minPts = 30 minor clusters
      "1" = "#C49A00",
      "2" = "#4CAF50",
      "3" = "#2EC4B6",
      
      # minPts = 50 major clusters
      "2" = "#17A8E3",  # blue/right
      "3" = "#8E63CE",  # purple/upper
      "4" = "#F15BB5",  # pink/lower
      
      # minPts = 50 minor
      "1" = "#2EC4B6"
    )
  ) +
  
  labs(
    title = "HDBSCAN Cluster Stability Across Density Thresholds",
    subtitle = "5PC + scaled logICI analysis",
    x = "UMAP1",
    y = "UMAP2",
    color = "HDBSCAN\ncluster"
  ) +
  
  theme_bw(base_size = 13)

print(p_30_50_compare)

# ------------------------------------------------------------
# Save
# ------------------------------------------------------------

ggsave(
  filename = file.path(
    tuning_plot_dir,
    "hdbscan_compare_minPts30_vs_50_clusterIDs.png"
  ),
  plot = p_30_50_compare,
  width = 12,
  height = 6,
  dpi = 300
)

```

## Compare ICI Distributions — minPts 30 vs 50

```{r}

# ------------------------------------------------------------
# Build stable region labels
# ------------------------------------------------------------

plot_30_50 <- hdbscan_tuning_combined %>%
  filter(minPts %in% c(30, 50)) %>%
  mutate(
    
    PlotCluster = case_when(
      
      # minPts = 30
      minPts == 30 & HDBSCAN_cluster == "5" ~ "Upper region",
      minPts == 30 & HDBSCAN_cluster == "4" ~ "Right region",
      minPts == 30 & HDBSCAN_cluster == "6" ~ "Lower region",
      
      # minPts = 50
      minPts == 50 & HDBSCAN_cluster == "3" ~ "Upper region",
      minPts == 50 & HDBSCAN_cluster == "2" ~ "Right region",
      minPts == 50 & HDBSCAN_cluster == "4" ~ "Lower region",
      
      # noise
      HDBSCAN_cluster == "0" ~ "Noise / unassigned",
      
      # everything else
      TRUE ~ "Minor clusters"
    )
  )

# ------------------------------------------------------------
# Keep only major stable regions
# ------------------------------------------------------------

ici_compare_df <- plot_30_50 %>%
  filter(
    PlotCluster %in% c(
      "Upper region",
      "Right region",
      "Lower region"
    )
  )

cat("Rows in ICI comparison dataframe:\n")
print(nrow(ici_compare_df))

# ------------------------------------------------------------
# Plot
# ------------------------------------------------------------

p_ici_compare <- ggplot(
  ici_compare_df,
  aes(
    x = PlotCluster,
    y = ICI_to_next_s,
    fill = PlotCluster
  )
) +
  
  geom_boxplot(
    outlier.size = 0.4,
    alpha = 0.8
  ) +
  
  facet_wrap(
    ~ minPts,
    nrow = 1
  ) +
  
  scale_fill_manual(
    values = c(
      "Upper region" = "#8E63CE",
      "Right region" = "#17A8E3",
      "Lower region" = "#F15BB5"
    )
  ) +
  
  scale_y_log10() +
  
  labs(
    title = "ICI Distributions Across HDBSCAN Density Thresholds",
    subtitle = "Major spectral-temporal regions remain temporally distinct",
    x = NULL,
    y = "ICI (seconds, log scale)",
    fill = NULL
  ) +
  
  theme_bw(base_size = 13)

print(p_ici_compare)

# ------------------------------------------------------------
# Save
# ------------------------------------------------------------

ggsave(
  filename = file.path(
    tuning_plot_dir,
    "ici_boxplot_compare_minPts30_vs_50.png"
  ),
  plot = p_ici_compare,
  width = 12,
  height = 6,
  dpi = 300
)

```

## Compare Mean Spectra — minPts 30 vs 50

```{r}

# ------------------------------------------------------------
# Keep non-noise clusters only
# ------------------------------------------------------------

spectra_compare_clusters <- hdbscan_tuning_combined %>%
  filter(
    minPts %in% c(30, 50),
    HDBSCAN_cluster != "0"
  )

# ------------------------------------------------------------
# Match spectra rows
# ------------------------------------------------------------

spectra_matrix_compare <- spectra_matrix_sampled_minmax[
  match(
    spectra_compare_clusters$UID,
    sampled_clicks_with_pcs$UID
  ),
]

# ------------------------------------------------------------
# Convert spectra to long format
# ------------------------------------------------------------

spectra_long_compare <- as.data.frame(
  spectra_matrix_compare
)

colnames(spectra_long_compare) <- paste0(
  "Freq_",
  seq_len(ncol(spectra_long_compare))
)

spectra_long_compare$UID <- spectra_compare_clusters$UID
spectra_long_compare$HDBSCAN_cluster <- spectra_compare_clusters$HDBSCAN_cluster
spectra_long_compare$minPts <- spectra_compare_clusters$minPts

spectra_long_compare <- spectra_long_compare %>%
  pivot_longer(
    cols = starts_with("Freq_"),
    names_to = "FreqIndex",
    values_to = "Amplitude"
  ) %>%
  mutate(
    FreqIndex = as.numeric(
      gsub("Freq_", "", FreqIndex)
    ),
    Frequency_kHz = freq_focus[FreqIndex] / 1000
  )

# ------------------------------------------------------------
# Mean spectra
# ------------------------------------------------------------

mean_spectra_compare <- spectra_long_compare %>%
  group_by(
    minPts,
    HDBSCAN_cluster,
    Frequency_kHz
  ) %>%
  summarise(
    MeanAmplitude = mean(
      Amplitude,
      na.rm = TRUE
    ),
    SDAmplitude = sd(
      Amplitude,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

# ------------------------------------------------------------
# Plot
# ------------------------------------------------------------

p_spectra_cluster_compare <- ggplot(
  mean_spectra_compare,
  aes(
    x = Frequency_kHz,
    y = MeanAmplitude
  )
) +
  
  geom_ribbon(
    aes(
      ymin = MeanAmplitude - SDAmplitude,
      ymax = MeanAmplitude + SDAmplitude
    ),
    fill = "grey70",
    alpha = 0.5
  ) +
  
  geom_line(
    linewidth = 1.2,
    color = "black"
  ) +
  
  facet_grid(
    minPts ~ HDBSCAN_cluster
  ) +
  
  labs(
    title = "Mean Spectra with Variability by HDBSCAN Cluster",
    subtitle = "Comparison across minPts = 30 and 50",
    x = "Frequency (kHz)",
    y = "Min-max normalized amplitude"
  ) +
  
  theme_bw(base_size = 13)

print(p_spectra_cluster_compare)

# ------------------------------------------------------------
# Save
# ------------------------------------------------------------

ggsave(
  filename = file.path(
    tuning_plot_dir,
    "mean_spectra_compare_clusters_minPts30_vs_50.png"
  ),
  plot = p_spectra_cluster_compare,
  width = 16,
  height = 8,
  dpi = 300
)

```

------------------------------------------------------
# minPts = 50 Outputs

## Export Main HDBSCAN Result — minPts = 50

```{r}

# ------------------------------------------------------------
# Output folders for selected minPts = 50 result
# ------------------------------------------------------------

minpts50_dir <- file.path(output_dir, "Main_minPts50_5PC_scaled_logICI")
minpts50_data_dir <- file.path(minpts50_dir, "data")
minpts50_table_dir <- file.path(minpts50_dir, "tables")
minpts50_plot_dir <- file.path(minpts50_dir, "plots")

dir.create(minpts50_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(minpts50_data_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(minpts50_table_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(minpts50_plot_dir, recursive = TRUE, showWarnings = FALSE)

# ------------------------------------------------------------
# Pull minPts = 50 cluster assignments
# ------------------------------------------------------------

hdbscan_minpts50_assignments <- hdbscan_tuning_combined %>%
  filter(minPts == 50) %>%
  mutate(
    HDBSCAN_cluster = as.character(HDBSCAN_cluster),
    HDBSCAN_cluster_type = if_else(
      HDBSCAN_cluster == "0",
      "noise_unassigned",
      "cluster"
    )
  ) %>%
  select(
    EventID,
    DriftID_spectra,
    UID,
    RowID,
    UMAP1,
    UMAP2,
    minPts,
    HDBSCAN_cluster,
    HDBSCAN_cluster_type,
    HDBSCAN_membership_probability,
    ICI_to_next_s,
    log10_ICI_to_next_s,
    scaled_log10_ICI,
    starts_with("PC")
  )

# ------------------------------------------------------------
# Sanity checks
# ------------------------------------------------------------

cat("Rows in minPts = 50 assignment table:\n")
print(nrow(hdbscan_minpts50_assignments))

cat("\nUnique clusters:\n")
print(sort(unique(hdbscan_minpts50_assignments$HDBSCAN_cluster)))

cat("\nCluster counts:\n")
print(table(hdbscan_minpts50_assignments$HDBSCAN_cluster))

# ------------------------------------------------------------
# Save assignment table
# ------------------------------------------------------------

saveRDS(
  hdbscan_minpts50_assignments,
  file.path(minpts50_data_dir, "hdbscan_minPts50_click_assignments.rds")
)

write.csv(
  hdbscan_minpts50_assignments,
  file.path(minpts50_table_dir, "hdbscan_minPts50_click_assignments.csv"),
  row.names = FALSE
)

cat("\nSaved minPts = 50 click-level assignments to:\n")
cat(minpts50_dir, "\n")

```

## Cluster Size and ICI Summary — minPts = 50

```{r}

# ------------------------------------------------------------
# Cluster size / percent summary
# ------------------------------------------------------------

hdbscan_minpts50_cluster_summary <- hdbscan_minpts50_assignments %>%
  count(
    HDBSCAN_cluster,
    HDBSCAN_cluster_type,
    name = "n_clicks"
  ) %>%
  mutate(
    percent_clicks = 100 * n_clicks / sum(n_clicks)
  ) %>%
  arrange(
    HDBSCAN_cluster_type,
    desc(n_clicks)
  )

print(hdbscan_minpts50_cluster_summary)

write.csv(
  hdbscan_minpts50_cluster_summary,
  file.path(minpts50_table_dir, "hdbscan_minPts50_cluster_summary.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# ICI summary by cluster
# ------------------------------------------------------------

hdbscan_minpts50_ici_summary <- hdbscan_minpts50_assignments %>%
  group_by(
    HDBSCAN_cluster,
    HDBSCAN_cluster_type
  ) %>%
  summarise(
    n_clicks = n(),
    median_ici_s = median(ICI_to_next_s, na.rm = TRUE),
    mean_ici_s = mean(ICI_to_next_s, na.rm = TRUE),
    sd_ici_s = sd(ICI_to_next_s, na.rm = TRUE),
    q25_ici_s = quantile(ICI_to_next_s, 0.25, na.rm = TRUE),
    q75_ici_s = quantile(ICI_to_next_s, 0.75, na.rm = TRUE),
    min_ici_s = min(ICI_to_next_s, na.rm = TRUE),
    max_ici_s = max(ICI_to_next_s, na.rm = TRUE),
    n_events = n_distinct(EventID),
    n_drifts = n_distinct(DriftID_spectra),
    .groups = "drop"
  ) %>%
  arrange(
    HDBSCAN_cluster_type,
    desc(n_clicks)
  )

print(hdbscan_minpts50_ici_summary)

write.csv(
  hdbscan_minpts50_ici_summary,
  file.path(minpts50_table_dir, "hdbscan_minPts50_ici_summary_by_cluster.csv"),
  row.names = FALSE
)

cat("\nSaved cluster and ICI summary tables.\n")

```

## Generate Main minPts = 50 Tables and Plots

```{r}

# ------------------------------------------------------------
# Frequency grid safety
# ------------------------------------------------------------

freq_plot_khz <- if (max(freq_focus, na.rm = TRUE) > 1000) {
  freq_focus / 1000
} else {
  freq_focus
}

# ------------------------------------------------------------
# Colors for minPts = 50 clusters
# 0 = noise/unassigned
# ------------------------------------------------------------

cluster_colors_50 <- c(
  "0" = "#F8766D",
  "1" = "#C49A00",
  "2" = "#53B400",
  "3" = "#00C1A2",
  "4" = "#00A9E0"
)

# ------------------------------------------------------------
# Build long spectra table for minPts = 50 assignments
# ------------------------------------------------------------

sampled_key <- sampled_clicks_with_pcs %>%
  mutate(
    spectrum_row = row_number(),
    UID = as.character(UID)
  ) %>%
  select(EventID, UID, spectrum_row)

assignments_with_spectrum_row <- hdbscan_minpts50_assignments %>%
  mutate(UID = as.character(UID)) %>%
  left_join(sampled_key, by = c("EventID", "UID"))

if (any(is.na(assignments_with_spectrum_row$spectrum_row))) {
  stop("Some minPts = 50 assignments could not be matched to spectra rows.")
}

spectra_minpts50_matrix <- spectra_matrix_sampled_minmax[
  assignments_with_spectrum_row$spectrum_row,
  ,
  drop = FALSE
]

spectra_minpts50_long <- as.data.frame(spectra_minpts50_matrix)

colnames(spectra_minpts50_long) <- paste0("Freq_", seq_len(ncol(spectra_minpts50_long)))

spectra_minpts50_long <- spectra_minpts50_long %>%
  mutate(
    EventID = assignments_with_spectrum_row$EventID,
    DriftID_spectra = assignments_with_spectrum_row$DriftID_spectra,
    UID = assignments_with_spectrum_row$UID,
    HDBSCAN_cluster = assignments_with_spectrum_row$HDBSCAN_cluster,
    HDBSCAN_cluster_type = assignments_with_spectrum_row$HDBSCAN_cluster_type
  ) %>%
  pivot_longer(
    cols = starts_with("Freq_"),
    names_to = "FreqIndex",
    values_to = "Amplitude"
  ) %>%
  mutate(
    FreqIndex = as.integer(gsub("Freq_", "", FreqIndex)),
    Frequency_kHz = freq_plot_khz[FreqIndex]
  )

mean_spectra_minpts50 <- spectra_minpts50_long %>%
  group_by(HDBSCAN_cluster, HDBSCAN_cluster_type, Frequency_kHz) %>%
  summarise(
    mean_minmax_spectrum = mean(Amplitude, na.rm = TRUE),
    sd_minmax_spectrum = sd(Amplitude, na.rm = TRUE),
    n_clicks = n_distinct(paste(EventID, UID)),
    .groups = "drop"
  )

write.csv(
  mean_spectra_minpts50,
  file.path(minpts50_table_dir, "hdbscan_minPts50_mean_spectra_by_cluster.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# 1. UMAP cluster plot
# ------------------------------------------------------------

p_umap_minpts50 <- ggplot(
  hdbscan_minpts50_assignments,
  aes(x = UMAP1, y = UMAP2, color = HDBSCAN_cluster)
) +
  geom_point(size = 0.5, alpha = 0.65) +
  scale_color_manual(values = cluster_colors_50) +
  labs(
    title = "HDBSCAN Clusters: 5PC + 1x scaled logICI",
    subtitle = "minPts = 50; Cluster 0 = noise/unassigned",
    x = "UMAP1",
    y = "UMAP2",
    color = "HDBSCAN\ncluster"
  ) +
  theme_bw(base_size = 13)

print(p_umap_minpts50)

ggsave(
  file.path(minpts50_plot_dir, "hdbscan_minPts50_umap_clusters.png"),
  p_umap_minpts50,
  width = 8,
  height = 6,
  dpi = 300
)

# ------------------------------------------------------------
# 2. ICI boxplot
# ------------------------------------------------------------

p_ici_minpts50 <- ggplot(
  hdbscan_minpts50_assignments,
  aes(x = HDBSCAN_cluster, y = ICI_to_next_s, fill = HDBSCAN_cluster)
) +
  geom_boxplot(outlier.size = 0.3) +
  scale_fill_manual(values = cluster_colors_50) +
  scale_y_log10() +
  labs(
    title = "ICI Distribution by HDBSCAN Cluster",
    subtitle = "5PC + 1x scaled logICI; minPts = 50",
    x = "HDBSCAN cluster",
    y = "ICI (seconds, log scale)",
    fill = "HDBSCAN\ncluster"
  ) +
  theme_bw(base_size = 13)

print(p_ici_minpts50)

ggsave(
  file.path(minpts50_plot_dir, "hdbscan_minPts50_ici_boxplot_by_cluster.png"),
  p_ici_minpts50,
  width = 8,
  height = 6,
  dpi = 300
)

# ------------------------------------------------------------
# 3. Mean spectra, all clusters together
# ------------------------------------------------------------

p_mean_spectra_all_minpts50 <- ggplot(
  mean_spectra_minpts50,
  aes(
    x = Frequency_kHz,
    y = mean_minmax_spectrum,
    color = HDBSCAN_cluster,
    group = HDBSCAN_cluster
  )
) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = cluster_colors_50) +
  labs(
    title = "Mean Min-Max Normalized Spectra by HDBSCAN Cluster",
    subtitle = "5PC + 1x scaled logICI; minPts = 50; cluster 0 = noise/unassigned",
    x = "Frequency (kHz)",
    y = "Mean min-max normalized amplitude",
    color = "HDBSCAN\ncluster"
  ) +
  theme_bw(base_size = 13)

print(p_mean_spectra_all_minpts50)

ggsave(
  file.path(minpts50_plot_dir, "hdbscan_minPts50_mean_spectra_all_clusters.png"),
  p_mean_spectra_all_minpts50,
  width = 8,
  height = 6,
  dpi = 300
)

# ------------------------------------------------------------
# 4. Mean spectra faceted
# ------------------------------------------------------------

p_mean_spectra_faceted_minpts50 <- ggplot(
  mean_spectra_minpts50,
  aes(x = Frequency_kHz, y = mean_minmax_spectrum)
) +
  geom_line(linewidth = 1.1, color = "black") +
  facet_wrap(~ HDBSCAN_cluster, ncol = 2) +
  labs(
    title = "Mean Min-Max Normalized Spectra by HDBSCAN Cluster",
    subtitle = "Faceted by cluster; minPts = 50; cluster 0 = noise/unassigned",
    x = "Frequency (kHz)",
    y = "Mean min-max normalized amplitude"
  ) +
  theme_bw(base_size = 13)

print(p_mean_spectra_faceted_minpts50)

ggsave(
  file.path(minpts50_plot_dir, "hdbscan_minPts50_mean_spectra_faceted.png"),
  p_mean_spectra_faceted_minpts50,
  width = 10,
  height = 8,
  dpi = 300
)

# ------------------------------------------------------------
# 5. Mean spectra with SD ribbons, noise excluded
# ------------------------------------------------------------

p_mean_spectra_sd_minpts50 <- mean_spectra_minpts50 %>%
  filter(HDBSCAN_cluster != "0") %>%
  ggplot(aes(x = Frequency_kHz, y = mean_minmax_spectrum)) +
  geom_ribbon(
    aes(
      ymin = mean_minmax_spectrum - sd_minmax_spectrum,
      ymax = mean_minmax_spectrum + sd_minmax_spectrum
    ),
    fill = "grey70",
    alpha = 0.55
  ) +
  geom_line(linewidth = 1.1, color = "black") +
  facet_wrap(~ HDBSCAN_cluster, ncol = 2) +
  labs(
    title = "Mean Spectra with Variability by HDBSCAN Cluster",
    subtitle = "Mean ± 1 SD; 5PC + 1x scaled logICI; minPts = 50; noise excluded",
    x = "Frequency (kHz)",
    y = "Min-max normalized amplitude"
  ) +
  theme_bw(base_size = 13)

print(p_mean_spectra_sd_minpts50)

ggsave(
  file.path(minpts50_plot_dir, "hdbscan_minPts50_mean_spectra_with_sd_ribbons_noise_excluded.png"),
  p_mean_spectra_sd_minpts50,
  width = 10,
  height = 8,
  dpi = 300
)

# ------------------------------------------------------------
# 6. Stacked mean spectra, noise excluded
# ------------------------------------------------------------

stacked_spectra_minpts50 <- mean_spectra_minpts50 %>%
  mutate(
    cluster_numeric = as.numeric(HDBSCAN_cluster),
    stacked_spectrum = mean_minmax_spectrum + cluster_numeric
  )

p_stacked_no_noise_minpts50 <- stacked_spectra_minpts50 %>%
  filter(HDBSCAN_cluster != "0") %>%
  ggplot(
    aes(
      x = Frequency_kHz,
      y = stacked_spectrum,
      color = HDBSCAN_cluster,
      group = HDBSCAN_cluster
    )
  ) +
  geom_line(linewidth = 1.2) +
  scale_color_manual(values = cluster_colors_50) +
  scale_y_continuous(
    breaks = sort(unique(stacked_spectra_minpts50$cluster_numeric)),
    labels = sort(unique(stacked_spectra_minpts50$HDBSCAN_cluster))
  ) +
  labs(
    title = "Stacked Mean Spectra by HDBSCAN Cluster",
    subtitle = "5PC + 1x scaled logICI; minPts = 50; noise cluster removed",
    x = "Frequency (kHz)",
    y = "HDBSCAN cluster"
  ) +
  theme_classic(base_size = 13) +
  theme(legend.position = "none")

print(p_stacked_no_noise_minpts50)

ggsave(
  file.path(minpts50_plot_dir, "hdbscan_minPts50_stacked_mean_spectra_no_noise.png"),
  p_stacked_no_noise_minpts50,
  width = 4,
  height = 7,
  dpi = 300
)

# ------------------------------------------------------------
# 7. Drift composition table and plot
# ------------------------------------------------------------

drift_composition_minpts50 <- hdbscan_minpts50_assignments %>%
  count(HDBSCAN_cluster, DriftID_spectra, name = "n_clicks") %>%
  group_by(HDBSCAN_cluster) %>%
  mutate(percent_within_cluster = 100 * n_clicks / sum(n_clicks)) %>%
  ungroup() %>%
  arrange(HDBSCAN_cluster, desc(n_clicks))

write.csv(
  drift_composition_minpts50,
  file.path(minpts50_table_dir, "hdbscan_minPts50_drift_composition_by_cluster.csv"),
  row.names = FALSE
)

top_drifts_minpts50 <- drift_composition_minpts50 %>%
  group_by(HDBSCAN_cluster) %>%
  slice_max(n_clicks, n = 5, with_ties = FALSE) %>%
  ungroup()

p_top_drifts_minpts50 <- ggplot(
  top_drifts_minpts50,
  aes(
    x = n_clicks,
    y = reorder(DriftID_spectra, n_clicks),
    fill = HDBSCAN_cluster
  )
) +
  geom_col() +
  facet_wrap(~ HDBSCAN_cluster, scales = "free", ncol = 3) +
  scale_fill_manual(values = cluster_colors_50) +
  labs(
    title = "Top Contributing Drifts per HDBSCAN Cluster",
    subtitle = "5PC + 1x scaled logICI; minPts = 50",
    x = "Number of clicks",
    y = "DriftID"
  ) +
  theme_bw(base_size = 13) +
  theme(legend.position = "none")

print(p_top_drifts_minpts50)

ggsave(
  file.path(minpts50_plot_dir, "hdbscan_minPts50_top_drifts_per_cluster.png"),
  p_top_drifts_minpts50,
  width = 12,
  height = 8,
  dpi = 300
)

# ------------------------------------------------------------
# 8. Event composition table and plot
# ------------------------------------------------------------

event_composition_minpts50 <- hdbscan_minpts50_assignments %>%
  count(HDBSCAN_cluster, EventID, name = "n_clicks") %>%
  group_by(HDBSCAN_cluster) %>%
  mutate(percent_within_cluster = 100 * n_clicks / sum(n_clicks)) %>%
  ungroup() %>%
  arrange(HDBSCAN_cluster, desc(n_clicks))

write.csv(
  event_composition_minpts50,
  file.path(minpts50_table_dir, "hdbscan_minPts50_event_composition_by_cluster.csv"),
  row.names = FALSE
)

top_events_minpts50 <- event_composition_minpts50 %>%
  group_by(HDBSCAN_cluster) %>%
  slice_max(n_clicks, n = 5, with_ties = FALSE) %>%
  ungroup()

p_top_events_minpts50 <- ggplot(
  top_events_minpts50,
  aes(
    x = n_clicks,
    y = reorder(EventID, n_clicks),
    fill = HDBSCAN_cluster
  )
) +
  geom_col() +
  facet_wrap(~ HDBSCAN_cluster, scales = "free", ncol = 3) +
  scale_fill_manual(values = cluster_colors_50) +
  labs(
    title = "Top Contributing Events per HDBSCAN Cluster",
    subtitle = "5PC + 1x scaled logICI; minPts = 50",
    x = "Number of clicks",
    y = "EventID"
  ) +
  theme_bw(base_size = 13) +
  theme(legend.position = "none")

print(p_top_events_minpts50)

ggsave(
  file.path(minpts50_plot_dir, "hdbscan_minPts50_top_events_per_cluster.png"),
  p_top_events_minpts50,
  width = 12,
  height = 8,
  dpi = 300
)

cat("\nSaved all minPts = 50 plots and tables to:\n")
cat(minpts50_dir, "\n")

```

--------------------------------------------------

# Map HDBSCAN cluster composition by event

## Setup

```{r}

library(dplyr)
library(readr)
library(ggplot2)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(scatterpie)
library(tidyr)
library(stringr)

# ------------------------------------------------------------
# Input: drift/event latitude-longitude table
# ------------------------------------------------------------

adrift_detection_file <- "E:/Logs/ADRIFT_GgDetections.csv"

adrift_detections <- read_csv(adrift_detection_file, show_col_types = FALSE)

cat("Columns in ADRIFT_GgDetections:\n")
print(names(adrift_detections))

```

## Inspect location file contents

```{r}

cat("Rows in location file:", nrow(adrift_detections), "\n\n")

cat("Unique DriftName prefixes:\n")
print(
  adrift_detections %>%
    mutate(DriftPrefix = stringr::str_extract(DriftName, "^[^_]+")) %>%
    count(DriftPrefix, name = "n_rows") %>%
    arrange(DriftPrefix)
)

cat("\nUnique EventID prefixes:\n")
print(
  adrift_detections %>%
    mutate(EventPrefix = stringr::str_extract(EventID, "^[^_]+")) %>%
    count(EventPrefix, name = "n_rows") %>%
    arrange(EventPrefix)
)

cat("\nPreview across all drift types:\n")
print(
  adrift_detections %>%
    select(DriftName, EventID, Latitude, Longitude) %>%
    arrange(DriftName, EventID) %>%
    group_by(stringr::str_extract(DriftName, "^[^_]+")) %>%
    slice_head(n = 5) %>%
    ungroup()
)

```

## Sanity check: events and click counts used in final HDBSCAN

```{r}

hdbscan_event_click_counts <- hdbscan_minpts50_assignments %>%
  mutate(
    EventID = as.character(EventID),
    DriftID = stringr::str_remove(EventID, "_[0-9]+$"),
    EventPrefix = stringr::str_extract(EventID, "^[^_]+")
  ) %>%
  count(EventPrefix, DriftID, EventID, name = "n_clicks_used") %>%
  arrange(EventPrefix, DriftID, EventID)

cat("Total clicks in final HDBSCAN object:", sum(hdbscan_event_click_counts$n_clicks_used), "\n")
cat("Unique events used:", nrow(hdbscan_event_click_counts), "\n")
cat("Unique drifts used:", n_distinct(hdbscan_event_click_counts$DriftID), "\n\n")

cat("Events by prefix:\n")
print(
  hdbscan_event_click_counts %>%
    count(EventPrefix, name = "n_events")
)

cat("\nClick-count summary per event:\n")
print(summary(hdbscan_event_click_counts$n_clicks_used))

cat("\nFull event click-count table:\n")
print(hdbscan_event_click_counts, n = Inf)

```

##  Match final HDBSCAN events to location file using cleaned EventID

```{r}

hdbscan_event_click_counts_clean <- hdbscan_event_click_counts %>%
  mutate(
    EventID_clean = stringr::str_replace(EventID, "_Event_", "_"),
    DriftID_clean = stringr::str_remove(EventID_clean, "_[0-9]+$")
  )

event_locations <- adrift_detections %>%
  mutate(
    EventID_clean = as.character(EventID),
    Latitude = as.numeric(Latitude),
    Longitude = as.numeric(Longitude)
  ) %>%
  filter(
    !is.na(EventID_clean),
    !is.na(Latitude),
    !is.na(Longitude)
  ) %>%
  select(EventID_clean, DriftName, Latitude, Longitude)

hdbscan_events_with_locations <- hdbscan_event_click_counts_clean %>%
  left_join(event_locations, by = "EventID_clean")

cat("HDBSCAN events:", nrow(hdbscan_event_click_counts_clean), "\n")
cat("Matched location rows:", sum(!is.na(hdbscan_events_with_locations$Latitude)), "\n")
cat("Missing location rows:", sum(is.na(hdbscan_events_with_locations$Latitude)), "\n\n")

cat("Missing events, if any:\n")
print(
  hdbscan_events_with_locations %>%
    filter(is.na(Latitude) | is.na(Longitude))
)

cat("\nMatched event coordinate preview:\n")
print(hdbscan_events_with_locations, n = Inf)

```

## Build event-level cluster composition tables

```{r}

event_cluster_counts <- hdbscan_minpts50_assignments %>%
  mutate(
    EventID = as.character(EventID),
    EventID_clean = stringr::str_replace(EventID, "_Event_", "_"),
    HDBSCAN_cluster = as.character(HDBSCAN_cluster)
  ) %>%
  count(EventID, EventID_clean, HDBSCAN_cluster, name = "n_clicks") %>%
  left_join(
    hdbscan_events_with_locations %>%
      select(EventID, EventID_clean, DriftName, Latitude, Longitude),
    by = c("EventID", "EventID_clean")
  )

cat("Rows in event-cluster count table:", nrow(event_cluster_counts), "\n")
cat("Unique mapped events:", n_distinct(event_cluster_counts$EventID), "\n")
cat("Total clicks represented:", sum(event_cluster_counts$n_clicks), "\n\n")

cat("Cluster composition preview:\n")
print(head(event_cluster_counts, 30))

```

## Create wide pie-map tables

```{r}

event_pies_including_noise <- event_cluster_counts %>%
  select(EventID, EventID_clean, DriftName, Latitude, Longitude, HDBSCAN_cluster, n_clicks) %>%
  tidyr::pivot_wider(
    names_from = HDBSCAN_cluster,
    values_from = n_clicks,
    values_fill = 0,
    names_prefix = "cluster_"
  )

event_pies_excluding_noise <- event_cluster_counts %>%
  filter(HDBSCAN_cluster != "0") %>%
  select(EventID, EventID_clean, DriftName, Latitude, Longitude, HDBSCAN_cluster, n_clicks) %>%
  tidyr::pivot_wider(
    names_from = HDBSCAN_cluster,
    values_from = n_clicks,
    values_fill = 0,
    names_prefix = "cluster_"
  )

cat("Including noise: events =", nrow(event_pies_including_noise), "\n")
cat("Excluding noise: events =", nrow(event_pies_excluding_noise), "\n\n")

cat("Including noise columns:\n")
print(names(event_pies_including_noise))

cat("\nExcluding noise columns:\n")
print(names(event_pies_excluding_noise))

cat("\nPreview including noise:\n")
print(head(event_pies_including_noise, 10))

cat("\nPreview excluding noise:\n")
print(head(event_pies_excluding_noise, 10))

```

## Add pie radius and save map-ready tables

```{r}

# ------------------------------------------------------------
# Output folders
# ------------------------------------------------------------

map_output_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/minPts50_event_cluster_maps"

map_plot_dir <- file.path(map_output_dir, "plots")
map_table_dir <- file.path(map_output_dir, "tables")

dir.create(map_plot_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(map_table_dir, recursive = TRUE, showWarnings = FALSE)

# ------------------------------------------------------------
# Add pie radius
# ------------------------------------------------------------

add_pie_radius <- function(df) {
  
  cluster_cols <- names(df)[stringr::str_detect(names(df), "^cluster_")]
  
  df %>%
    mutate(
      total_clicks_mapped = rowSums(across(all_of(cluster_cols))),
      pie_radius = sqrt(total_clicks_mapped) /
        max(sqrt(total_clicks_mapped), na.rm = TRUE) * 0.55
    )
}

event_pies_including_noise <- add_pie_radius(event_pies_including_noise)

event_pies_excluding_noise <- add_pie_radius(event_pies_excluding_noise)

# ------------------------------------------------------------
# Save tables
# ------------------------------------------------------------

write.csv(
  event_pies_including_noise,
  file.path(
    map_table_dir,
    "event_pie_map_data_including_noise.csv"
  ),
  row.names = FALSE
)

write.csv(
  event_pies_excluding_noise,
  file.path(
    map_table_dir,
    "event_pie_map_data_excluding_noise.csv"
  ),
  row.names = FALSE
)

# ------------------------------------------------------------
# Sanity checks
# ------------------------------------------------------------

cat("Including-noise total clicks:",
    sum(event_pies_including_noise$total_clicks_mapped), "\n")

cat("Excluding-noise total clicks:",
    sum(event_pies_excluding_noise$total_clicks_mapped), "\n\n")

cat("Pie radius summary including noise:\n")
print(summary(event_pies_including_noise$pie_radius))

cat("\nPie radius summary excluding noise:\n")
print(summary(event_pies_excluding_noise$pie_radius))

cat("\nSaved map-ready tables to:\n")
cat(map_table_dir, "\n")

```

## Build California Current basemap and pie-map function

```{r}

world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")

make_event_pie_map <- function(pie_data, include_noise = TRUE, output_name) {
  
  cluster_cols <- names(pie_data)[stringr::str_detect(names(pie_data), "^cluster_")]
  
  cluster_cols <- cluster_cols[
    order(as.numeric(stringr::str_remove(cluster_cols, "cluster_")))
  ]
  
  xlim <- range(pie_data$Longitude, na.rm = TRUE) + c(-1.5, 1.5)
  ylim <- range(pie_data$Latitude, na.rm = TRUE) + c(-1.5, 1.5)
  
  p <- ggplot() +
    geom_sf(
      data = world,
      fill = "gray92",
      color = "gray60",
      linewidth = 0.25
    ) +
    coord_sf(
      xlim = xlim,
      ylim = ylim,
      expand = FALSE
    ) +
    scatterpie::geom_scatterpie(
      data = pie_data,
      aes(x = Longitude, y = Latitude, r = pie_radius),
      cols = cluster_cols,
      color = "black",
      linewidth = 0.15,
      alpha = 0.95
    ) +
    labs(
      title = ifelse(
        include_noise,
        "HDBSCAN Cluster Composition by Event: Including Noise",
        "HDBSCAN Cluster Composition by Event: Excluding Noise"
      ),
      subtitle = "Pies show proportional click membership per analyzed event",
      x = "Longitude",
      y = "Latitude",
      fill = "HDBSCAN cluster"
    ) +
    theme_bw(base_size = 13) +
    theme(
      plot.title = element_text(face = "bold"),
      panel.grid.major = element_line(color = "gray85", linewidth = 0.25),
      legend.position = "right"
    )
  
  ggsave(
    filename = file.path(map_plot_dir, output_name),
    plot = p,
    width = 8.5,
    height = 7,
    dpi = 300
  )
  
  return(p)
}

```

## Generate both event pie maps

```{r}

make_event_pie_map <- function(pie_data, include_noise = TRUE, output_name) {
  
  cluster_cols <- names(pie_data)[stringr::str_detect(names(pie_data), "^cluster_")]
  cluster_cols <- cluster_cols[
    order(as.numeric(stringr::str_remove(cluster_cols, "cluster_")))
  ]
  
  xlim <- range(pie_data$Longitude, na.rm = TRUE) + c(-1.5, 1.5)
  ylim <- range(pie_data$Latitude, na.rm = TRUE) + c(-1.5, 1.5)
  
  p <- ggplot() +
    geom_sf(data = world, fill = "gray92", color = "gray60") +
    coord_sf(xlim = xlim, ylim = ylim, expand = FALSE) +
    scatterpie::geom_scatterpie(
      data = pie_data,
      aes(x = Longitude, y = Latitude, r = pie_radius),
      cols = cluster_cols,
      color = "black",
      alpha = 0.95
    ) +
    labs(
      title = ifelse(
        include_noise,
        "HDBSCAN Cluster Composition by Event: Including Noise",
        "HDBSCAN Cluster Composition by Event: Excluding Noise"
      ),
      subtitle = "Pies show proportional click membership per analyzed event",
      x = "Longitude",
      y = "Latitude",
      fill = "HDBSCAN cluster"
    ) +
    theme_bw(base_size = 13) +
    theme(
      plot.title = element_text(face = "bold"),
      legend.position = "right"
    )
  
  ggsave(
    filename = file.path(map_plot_dir, output_name),
    plot = p,
    width = 8.5,
    height = 7,
    dpi = 300
  )
  
  return(p)
}

```

## Troubleshooting

```{r}

## Check map coordinate ranges

cat("Including noise longitude range:\n")
print(range(event_pies_including_noise$Longitude, na.rm = TRUE))

cat("Including noise latitude range:\n")
print(range(event_pies_including_noise$Latitude, na.rm = TRUE))

cat("\nRows with missing coordinates:\n")
print(
  event_pies_including_noise %>%
    filter(is.na(Longitude) | is.na(Latitude))
)

cat("\nCluster columns:\n")
print(names(event_pies_including_noise)[stringr::str_detect(names(event_pies_including_noise), "^cluster_")])

cat("\nFirst few map rows:\n")
print(head(event_pies_including_noise))

# Simple point map test

p_test_points <- ggplot() +
  geom_sf(data = world, fill = "gray92", color = "gray60") +
  geom_point(
    data = event_pies_including_noise,
    aes(x = Longitude, y = Latitude),
    size = 2
  ) +
  coord_sf(
    xlim = range(event_pies_including_noise$Longitude, na.rm = TRUE) + c(-2, 2),
    ylim = range(event_pies_including_noise$Latitude, na.rm = TRUE) + c(-2, 2),
    expand = FALSE
  ) +
  theme_bw()

print(p_test_points)

```

## Manual pie-map function, no scatterpie

```{r}

make_pie_slice_data <- function(pie_data, cluster_cols, radius_col = "pie_radius", n_points = 80) {
  
  out <- list()
  row_i <- 1
  
  for (i in seq_len(nrow(pie_data))) {
    
    vals <- as.numeric(pie_data[i, cluster_cols])
    total <- sum(vals, na.rm = TRUE)
    
    if (total <= 0) next
    
    props <- vals / total
    angle_starts <- c(0, cumsum(props)[-length(props)]) * 2 * pi
    angle_ends <- cumsum(props) * 2 * pi
    
    for (j in seq_along(cluster_cols)) {
      
      if (vals[j] <= 0) next
      
      angles <- seq(angle_starts[j], angle_ends[j], length.out = n_points)
      
      out[[row_i]] <- data.frame(
        EventID = pie_data$EventID[i],
        cluster = stringr::str_remove(cluster_cols[j], "cluster_"),
        Longitude = c(pie_data$Longitude[i], pie_data$Longitude[i] + pie_data[[radius_col]][i] * cos(angles)),
        Latitude  = c(pie_data$Latitude[i],  pie_data$Latitude[i]  + pie_data[[radius_col]][i] * sin(angles)),
        piece_id = paste0(pie_data$EventID[i], "_", cluster_cols[j])
      )
      
      row_i <- row_i + 1
    }
  }
  
  bind_rows(out)
}

make_event_pie_map <- function(pie_data, include_noise = TRUE, output_name) {
  
  cluster_cols <- names(pie_data)[stringr::str_detect(names(pie_data), "^cluster_")]
  cluster_cols <- cluster_cols[
    order(as.numeric(stringr::str_remove(cluster_cols, "cluster_")))
  ]
  
  pie_slices <- make_pie_slice_data(
    pie_data = pie_data,
    cluster_cols = cluster_cols,
    radius_col = "pie_radius"
  )
  
  xlim <- range(pie_data$Longitude, na.rm = TRUE) + c(-1.8, 1.8)
  ylim <- range(pie_data$Latitude, na.rm = TRUE) + c(-1.8, 1.8)
  
  p <- ggplot() +
    geom_sf(data = world, fill = "gray92", color = "gray60") +
    geom_polygon(
      data = pie_slices,
      aes(
        x = Longitude,
        y = Latitude,
        group = piece_id,
        fill = cluster
      ),
      color = "black",
      linewidth = 0.12,
      alpha = 0.95
    ) +
    coord_sf(
      xlim = xlim,
      ylim = ylim,
      expand = FALSE
    ) +
    labs(
      title = ifelse(
        include_noise,
        "HDBSCAN Cluster Composition by Event: Including Noise",
        "HDBSCAN Cluster Composition by Event: Excluding Noise"
      ),
      subtitle = "Pies show proportional click membership per analyzed event",
      x = "Longitude",
      y = "Latitude",
      fill = "HDBSCAN cluster"
    ) +
    theme_bw(base_size = 13) +
    theme(
      plot.title = element_text(face = "bold"),
      legend.position = "right"
    )
  
  ggsave(
    filename = file.path(map_plot_dir, output_name),
    plot = p,
    width = 8.5,
    height = 7,
    dpi = 300
  )
  
  return(p)
}

```

## Generate both manual event pie maps

```{r}

p_event_map_including_noise <- make_event_pie_map(
  pie_data = event_pies_including_noise,
  include_noise = TRUE,
  output_name = "hdbscan_event_pie_map_including_noise_manual.png"
)

p_event_map_excluding_noise <- make_event_pie_map(
  pie_data = event_pies_excluding_noise,
  include_noise = FALSE,
  output_name = "hdbscan_event_pie_map_excluding_noise_manual.png"
)

print(p_event_map_including_noise)
print(p_event_map_excluding_noise)

cat("Saved maps to:\n")
cat(map_plot_dir, "\n")

```

## Smaller circular pies

```{r}

make_pie_slice_data <- function(pie_data, cluster_cols, radius_col = "pie_radius", n_points = 80) {
  
  out <- list()
  row_i <- 1
  
  for (i in seq_len(nrow(pie_data))) {
    
    vals <- as.numeric(pie_data[i, cluster_cols])
    total <- sum(vals, na.rm = TRUE)
    
    if (total <= 0) next
    
    props <- vals / total
    angle_starts <- c(0, cumsum(props)[-length(props)]) * 2 * pi
    angle_ends <- cumsum(props) * 2 * pi
    
    lat_center <- pie_data$Latitude[i]
    lon_center <- pie_data$Longitude[i]
    
    # Correct longitude radius so pies look circular on lon/lat map
    lon_correction <- 1 / cos(lat_center * pi / 180)
    
    r_lat <- pie_data[[radius_col]][i]
    r_lon <- r_lat * lon_correction
    
    for (j in seq_along(cluster_cols)) {
      
      if (vals[j] <= 0) next
      
      angles <- seq(angle_starts[j], angle_ends[j], length.out = n_points)
      
      out[[row_i]] <- data.frame(
        EventID = pie_data$EventID[i],
        cluster = stringr::str_remove(cluster_cols[j], "cluster_"),
        Longitude = c(lon_center, lon_center + r_lon * cos(angles)),
        Latitude  = c(lat_center, lat_center + r_lat * sin(angles)),
        piece_id = paste0(pie_data$EventID[i], "_", cluster_cols[j])
      )
      
      row_i <- row_i + 1
    }
  }
  
  bind_rows(out)
}

# Shrink pies
event_pies_including_noise <- event_pies_including_noise %>%
  mutate(pie_radius = pie_radius * 0.45)

event_pies_excluding_noise <- event_pies_excluding_noise %>%
  mutate(pie_radius = pie_radius * 0.45)

p_event_map_including_noise <- make_event_pie_map(
  pie_data = event_pies_including_noise,
  include_noise = TRUE,
  output_name = "hdbscan_event_pie_map_including_noise_manual_smaller_circular.png"
)

p_event_map_excluding_noise <- make_event_pie_map(
  pie_data = event_pies_excluding_noise,
  include_noise = FALSE,
  output_name = "hdbscan_event_pie_map_excluding_noise_manual_smaller_circular.png"
)

print(p_event_map_including_noise)
print(p_event_map_excluding_noise)

```

## Stacked bar charts of event-level cluster proportions by longitude

```{r}

make_event_stacked_barplot <- function(pie_data, include_noise = TRUE, output_name) {
  
  cluster_cols <- names(pie_data)[stringr::str_detect(names(pie_data), "^cluster_")]
  cluster_cols <- cluster_cols[
    order(as.numeric(stringr::str_remove(cluster_cols, "cluster_")))
  ]
  
  bar_data <- pie_data %>%
    mutate(
      Event_label = paste0(EventID_clean, "\n", round(abs(Longitude), 1), "°W")
    ) %>%
    arrange(Longitude) %>%
    mutate(
      Event_label = factor(Event_label, levels = Event_label)
    ) %>%
    select(EventID, EventID_clean, Event_label, Longitude, Latitude, all_of(cluster_cols)) %>%
    pivot_longer(
      cols = all_of(cluster_cols),
      names_to = "cluster",
      values_to = "n_clicks"
    ) %>%
    mutate(
      cluster = stringr::str_remove(cluster, "cluster_")
    ) %>%
    group_by(EventID, EventID_clean, Event_label, Longitude, Latitude) %>%
    mutate(
      prop_clicks = n_clicks / sum(n_clicks)
    ) %>%
    ungroup()
  
  p <- ggplot(
    bar_data,
    aes(
      x = Event_label,
      y = prop_clicks,
      fill = cluster
    )
  ) +
    geom_col(width = 0.9, color = "black", linewidth = 0.1) +
    scale_y_continuous(
      labels = scales::percent_format(accuracy = 1),
      expand = c(0, 0)
    ) +
    labs(
      title = ifelse(
        include_noise,
        "Event-Level HDBSCAN Cluster Proportions: Including Noise",
        "Event-Level HDBSCAN Cluster Proportions: Excluding Noise"
      ),
      subtitle = "Events sorted west-to-east by longitude",
      x = "Event / Longitude",
      y = "Clicks per cluster (%)",
      fill = "HDBSCAN cluster"
    ) +
    theme_bw(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold"),
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 6),
      panel.grid.major.x = element_blank(),
      panel.grid.minor = element_blank(),
      legend.position = "right"
    )
  
  ggsave(
    filename = file.path(map_plot_dir, output_name),
    plot = p,
    width = 14,
    height = 6,
    dpi = 300
  )
  
  return(p)
}

p_bar_including_noise <- make_event_stacked_barplot(
  pie_data = event_pies_including_noise,
  include_noise = TRUE,
  output_name = "hdbscan_event_stacked_barplot_including_noise.png"
)

p_bar_excluding_noise <- make_event_stacked_barplot(
  pie_data = event_pies_excluding_noise,
  include_noise = FALSE,
  output_name = "hdbscan_event_stacked_barplot_excluding_noise.png"
)

print(p_bar_including_noise)
print(p_bar_excluding_noise)

```

## Stacked mean spectra with map/bar colors

```{r}

cluster_colors_consistent <- c(
  "0" = "#F8766D",
  "1" = "#A3A500",
  "2" = "#00BF7D",
  "3" = "#00B0F6",
  "4" = "#E76BF3"
)

# Use whichever spectra-summary object exists
spectra_summary_for_plot <- if (exists("mean_spectra_minpts50")) {
  mean_spectra_minpts50
} else {
  cluster_spectra_summary
}

spectra_summary_for_plot <- spectra_summary_for_plot %>%
  mutate(
    HDBSCAN_cluster = as.character(HDBSCAN_cluster)
  )

make_stacked_spectra_plot <- function(include_noise = TRUE, output_name) {
  
  plot_data <- spectra_summary_for_plot
  
  if (!include_noise) {
    plot_data <- plot_data %>%
      filter(HDBSCAN_cluster != "0")
  }
  
  cluster_order <- sort(as.numeric(unique(plot_data$HDBSCAN_cluster)))
  
  plot_data <- plot_data %>%
    mutate(
      HDBSCAN_cluster = factor(HDBSCAN_cluster, levels = as.character(cluster_order)),
      offset = as.numeric(HDBSCAN_cluster),
      
      # Vertical emphasis: increases peak/notch visibility without changing cluster ordering
      stacked_spectrum = offset + mean_minmax_spectrum * 1.25
    )
  
  p <- ggplot(
    plot_data,
    aes(
      x = Frequency_kHz,
      y = stacked_spectrum,
      color = HDBSCAN_cluster,
      group = HDBSCAN_cluster
    )
  ) +
    geom_line(linewidth = 1.5) +
    scale_color_manual(values = cluster_colors_consistent) +
    scale_y_continuous(
      breaks = cluster_order,
      labels = cluster_order
    ) +
    labs(
      title = ifelse(
        include_noise,
        "Stacked Mean Spectra by HDBSCAN Cluster: Including Noise",
        "Stacked Mean Spectra by HDBSCAN Cluster: Excluding Noise"
      ),
      subtitle = "5PC + scaled logICI; minPts = 50; vertical emphasis applied",
      x = "Frequency (kHz)",
      y = "HDBSCAN cluster",
      color = "HDBSCAN cluster"
    ) +
    theme_classic(base_size = 16) +
    theme(
      plot.title = element_text(face = "bold"),
      legend.position = "right"
    )
  
  ggsave(
    filename = file.path(map_plot_dir, output_name),
    plot = p,
    width = 4.2,
    height = 9.5,
    dpi = 300
  )
  
  return(p)
}

p_spectra_including_noise <- make_stacked_spectra_plot(
  include_noise = TRUE,
  output_name = "hdbscan_minPts50_stacked_mean_spectra_including_noise_matching_colors.png"
)

p_spectra_excluding_noise <- make_stacked_spectra_plot(
  include_noise = FALSE,
  output_name = "hdbscan_minPts50_stacked_mean_spectra_excluding_noise_matching_colors.png"
)

print(p_spectra_including_noise)
print(p_spectra_excluding_noise)

```

## Fix cut-off title/subtitle on stacked spectra

```{r}

make_stacked_spectra_plot <- function(include_noise = TRUE, output_name) {
  
  plot_data <- spectra_summary_for_plot
  
  if (!include_noise) {
    plot_data <- plot_data %>%
      filter(HDBSCAN_cluster != "0")
  }
  
  cluster_order <- sort(as.numeric(unique(plot_data$HDBSCAN_cluster)))
  
  plot_data <- plot_data %>%
    mutate(
      HDBSCAN_cluster = factor(HDBSCAN_cluster, levels = as.character(cluster_order)),
      offset = as.numeric(as.character(HDBSCAN_cluster)),
      stacked_spectrum = offset + mean_minmax_spectrum * 1.25
    )
  
  p <- ggplot(
    plot_data,
    aes(
      x = Frequency_kHz,
      y = stacked_spectrum,
      color = HDBSCAN_cluster,
      group = HDBSCAN_cluster
    )
  ) +
    geom_line(linewidth = 1.5) +
    scale_color_manual(values = cluster_colors_consistent) +
    scale_y_continuous(
      breaks = cluster_order,
      labels = cluster_order
    ) +
    labs(
      title = ifelse(
        include_noise,
        "Mean Spectra: Including Noise",
        "Mean Spectra: Excluding Noise"
      ),
      subtitle = "5PC + scaled logICI; minPts = 50",
      x = "Frequency (kHz)",
      y = "HDBSCAN cluster",
      color = "HDBSCAN cluster"
    ) +
    theme_classic(base_size = 16) +
    theme(
      plot.title = element_text(face = "bold"),
      plot.margin = margin(15, 20, 15, 15),
      legend.position = "right"
    )
  
  ggsave(
    filename = file.path(map_plot_dir, output_name),
    plot = p,
    width = 6.2,
    height = 9.5,
    dpi = 300
  )
  
  return(p)
}

p_spectra_including_noise <- make_stacked_spectra_plot(
  include_noise = TRUE,
  output_name = "hdbscan_minPts50_stacked_mean_spectra_including_noise_matching_colors_fixed.png"
)

p_spectra_excluding_noise <- make_stacked_spectra_plot(
  include_noise = FALSE,
  output_name = "hdbscan_minPts50_stacked_mean_spectra_excluding_noise_matching_colors_fixed.png"
)

print(p_spectra_including_noise)
print(p_spectra_excluding_noise)

```

## Re-generate maps, stacked barplots, and spectra with consistent colors

```{r}

cluster_colors_consistent <- c(
  "0" = "gray60",   # noise/unassigned
  "1" = "#F8766D", # dark blue
  "2" = "#7CAE00", # green
  "3" = "#1FBFCF", # cyan
  "4" = "#C77CFF"  # purple
)

# ============================================================
# 1. MAPS
# ============================================================

make_event_pie_map <- function(pie_data, include_noise = TRUE, output_name) {
  
  cluster_cols <- names(pie_data)[stringr::str_detect(names(pie_data), "^cluster_")]
  cluster_cols <- cluster_cols[
    order(as.numeric(stringr::str_remove(cluster_cols, "cluster_")))
  ]
  
  pie_slices <- make_pie_slice_data(
    pie_data = pie_data,
    cluster_cols = cluster_cols,
    radius_col = "pie_radius"
  )
  
  xlim <- range(pie_data$Longitude, na.rm = TRUE) + c(-1.8, 1.8)
  ylim <- range(pie_data$Latitude, na.rm = TRUE) + c(-1.8, 1.8)
  
  p <- ggplot() +
    geom_sf(data = world, fill = "gray92", color = "gray60") +
    geom_polygon(
      data = pie_slices,
      aes(
        x = Longitude,
        y = Latitude,
        group = piece_id,
        fill = cluster
      ),
      color = "black",
      linewidth = 0.12,
      alpha = 0.95
    ) +
    scale_fill_manual(values = cluster_colors_consistent) +
    coord_sf(xlim = xlim, ylim = ylim, expand = FALSE) +
    labs(
      title = ifelse(include_noise, "Event Pies: Including Noise", "Event Pies: Excluding Noise"),
      subtitle = "Proportional HDBSCAN cluster membership per analyzed event",
      x = "Longitude",
      y = "Latitude",
      fill = "HDBSCAN cluster"
    ) +
    theme_bw(base_size = 13) +
    theme(plot.title = element_text(face = "bold"))
  
  ggsave(file.path(map_plot_dir, output_name), p, width = 8.5, height = 7, dpi = 300)
  return(p)
}

p_map_including_noise <- make_event_pie_map(
  event_pies_including_noise,
  TRUE,
  "hdbscan_event_pie_map_including_noise_final_colors.png"
)

p_map_excluding_noise <- make_event_pie_map(
  event_pies_excluding_noise,
  FALSE,
  "hdbscan_event_pie_map_excluding_noise_final_colors.png"
)

# ============================================================
# 2. STACKED BARPLOTS
# ============================================================

make_event_stacked_barplot <- function(pie_data, include_noise = TRUE, output_name) {
  
  cluster_cols <- names(pie_data)[stringr::str_detect(names(pie_data), "^cluster_")]
  cluster_cols <- cluster_cols[
    order(as.numeric(stringr::str_remove(cluster_cols, "cluster_")))
  ]
  
  bar_data <- pie_data %>%
    mutate(
      Event_label = paste0(EventID_clean, "\n", round(abs(Longitude), 1), "°W")
    ) %>%
    arrange(Longitude) %>%
    mutate(Event_label = factor(Event_label, levels = Event_label)) %>%
    select(EventID, EventID_clean, Event_label, Longitude, Latitude, all_of(cluster_cols)) %>%
    pivot_longer(
      cols = all_of(cluster_cols),
      names_to = "cluster",
      values_to = "n_clicks"
    ) %>%
    mutate(cluster = stringr::str_remove(cluster, "cluster_")) %>%
    group_by(EventID, EventID_clean, Event_label, Longitude, Latitude) %>%
    mutate(prop_clicks = n_clicks / sum(n_clicks)) %>%
    ungroup()
  
  p <- ggplot(bar_data, aes(x = Event_label, y = prop_clicks, fill = cluster)) +
    geom_col(width = 0.9, color = "black", linewidth = 0.1) +
    scale_fill_manual(values = cluster_colors_consistent) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = c(0, 0)) +
    labs(
      title = ifelse(include_noise, "Cluster Proportions by Event: Including Noise", "Cluster Proportions by Event: Excluding Noise"),
      subtitle = "Events sorted west-to-east by longitude",
      x = "Event / Longitude",
      y = "Clicks per cluster (%)",
      fill = "HDBSCAN cluster"
    ) +
    theme_bw(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold"),
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 6),
      panel.grid.major.x = element_blank(),
      panel.grid.minor = element_blank()
    )
  
  ggsave(file.path(map_plot_dir, output_name), p, width = 14, height = 6, dpi = 300)
  return(p)
}

p_bar_including_noise <- make_event_stacked_barplot(
  event_pies_including_noise,
  TRUE,
  "hdbscan_event_stacked_barplot_including_noise_final_colors.png"
)

p_bar_excluding_noise <- make_event_stacked_barplot(
  event_pies_excluding_noise,
  FALSE,
  "hdbscan_event_stacked_barplot_excluding_noise_final_colors.png"
)

# ============================================================
# 3. STACKED MEAN SPECTRA
# ============================================================

spectra_summary_for_plot <- if (exists("mean_spectra_minpts50")) {
  mean_spectra_minpts50
} else {
  cluster_spectra_summary
}

spectra_summary_for_plot <- spectra_summary_for_plot %>%
  mutate(HDBSCAN_cluster = as.character(HDBSCAN_cluster))

make_stacked_spectra_plot <- function(include_noise = TRUE, output_name) {
  
  plot_data <- spectra_summary_for_plot
  
  if (!include_noise) {
    plot_data <- plot_data %>% filter(HDBSCAN_cluster != "0")
  }
  
  cluster_order <- sort(as.numeric(unique(plot_data$HDBSCAN_cluster)))
  
  plot_data <- plot_data %>%
    mutate(
      HDBSCAN_cluster = factor(HDBSCAN_cluster, levels = as.character(cluster_order)),
      offset = as.numeric(as.character(HDBSCAN_cluster)),
      stacked_spectrum = offset + mean_minmax_spectrum * 1.25
    )
  
  p <- ggplot(
    plot_data,
    aes(
      x = Frequency_kHz,
      y = stacked_spectrum,
      color = HDBSCAN_cluster,
      group = HDBSCAN_cluster
    )
  ) +
    geom_line(linewidth = 1.5) +
    scale_color_manual(values = cluster_colors_consistent) +
    scale_y_continuous(breaks = cluster_order, labels = cluster_order) +
    labs(
      title = ifelse(include_noise, "Mean Spectra: Including Noise", "Mean Spectra: Excluding Noise"),
      subtitle = "5PC + scaled logICI; minPts = 50",
      x = "Frequency (kHz)",
      y = "HDBSCAN cluster",
      color = "HDBSCAN cluster"
    ) +
    theme_classic(base_size = 16) +
    theme(
      plot.title = element_text(face = "bold"),
      plot.margin = margin(15, 20, 15, 15)
    )
  
  ggsave(file.path(map_plot_dir, output_name), p, width = 6.2, height = 9.5, dpi = 300)
  return(p)
}

p_spectra_including_noise <- make_stacked_spectra_plot(
  TRUE,
  "hdbscan_minPts50_stacked_mean_spectra_including_noise_final_colors.png"
)

p_spectra_excluding_noise <- make_stacked_spectra_plot(
  FALSE,
  "hdbscan_minPts50_stacked_mean_spectra_excluding_noise_final_colors.png"
)

print(p_map_including_noise)
print(p_map_excluding_noise)
print(p_bar_including_noise)
print(p_bar_excluding_noise)
print(p_spectra_including_noise)
print(p_spectra_excluding_noise)

```

## Big event pie maps with gray noise + matching cluster colors

```{r}

cluster_colors_consistent <- c(
  "0" = "gray60",   # noise/unassigned
  "1" = "#F8766D", # red
  "2" = "#7CAE00", # green
  "3" = "#1FBFCF", # cyan
  "4" = "#C77CFF"  # purple
)

# Rebuild radii fresh so pies are big again
add_pie_radius_custom <- function(df, max_radius = 0.50) {
  
  cluster_cols <- names(df)[stringr::str_detect(names(df), "^cluster_")]
  
  df %>%
    mutate(
      total_clicks_mapped = rowSums(across(all_of(cluster_cols))),
      pie_radius = sqrt(total_clicks_mapped) /
        max(sqrt(total_clicks_mapped), na.rm = TRUE) * max_radius
    )
}

event_pies_including_noise_big <- add_pie_radius_custom(
  event_pies_including_noise,
  max_radius = 0.50
)

event_pies_excluding_noise_big <- add_pie_radius_custom(
  event_pies_excluding_noise,
  max_radius = 0.42
)

make_event_pie_map_big <- function(pie_data, include_noise = TRUE, output_name) {
  
  cluster_cols <- names(pie_data)[stringr::str_detect(names(pie_data), "^cluster_")]
  cluster_cols <- cluster_cols[
    order(as.numeric(stringr::str_remove(cluster_cols, "cluster_")))
  ]
  
  pie_slices <- make_pie_slice_data(
    pie_data = pie_data,
    cluster_cols = cluster_cols,
    radius_col = "pie_radius"
  )
  
  xlim <- range(pie_data$Longitude, na.rm = TRUE) + c(-1.8, 1.8)
  ylim <- range(pie_data$Latitude, na.rm = TRUE) + c(-1.8, 1.8)
  
  p <- ggplot() +
    geom_sf(data = world, fill = "gray92", color = "gray60") +
    geom_polygon(
      data = pie_slices,
      aes(
        x = Longitude,
        y = Latitude,
        group = piece_id,
        fill = cluster
      ),
      color = "black",
      linewidth = 0.12,
      alpha = 0.95
    ) +
    scale_fill_manual(values = cluster_colors_consistent) +
    coord_sf(xlim = xlim, ylim = ylim, expand = FALSE) +
    labs(
      title = ifelse(
        include_noise,
        "Event Pies: Including Noise",
        "Event Pies: Excluding Noise"
      ),
      subtitle = "Proportional HDBSCAN cluster membership per analyzed event",
      x = "Longitude",
      y = "Latitude",
      fill = "HDBSCAN cluster"
    ) +
    theme_bw(base_size = 13) +
    theme(
      plot.title = element_text(face = "bold"),
      legend.position = "right"
    )
  
  ggsave(
    file.path(map_plot_dir, output_name),
    p,
    width = 8.5,
    height = 7,
    dpi = 300
  )
  
  return(p)
}

p_map_including_noise_big <- make_event_pie_map_big(
  event_pies_including_noise_big,
  TRUE,
  "hdbscan_event_pie_map_including_noise_big_matching_colors.png"
)

p_map_excluding_noise_big <- make_event_pie_map_big(
  event_pies_excluding_noise_big,
  FALSE,
  "hdbscan_event_pie_map_excluding_noise_big_matching_colors.png"
)

print(p_map_including_noise_big)
print(p_map_excluding_noise_big)

```

## Drift-Level plotting

```{r}



```

## Drift-level cluster composition maps and stacked plots

```{r}

# ============================================================
# DRIFT-LEVEL PIE MAPS + STACKED COMPOSITION PLOTS
# Same style as original event-level workflow
# ============================================================

# ------------------------------------------------------------
# Output folders: same parent folder as event-level maps
# ------------------------------------------------------------

map_output_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/minPts50_event_cluster_maps"
map_plot_dir  <- file.path(map_output_dir, "plots")
map_table_dir <- file.path(map_output_dir, "tables")

dir.create(map_plot_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(map_table_dir, recursive = TRUE, showWarnings = FALSE)

# ------------------------------------------------------------
# Build one clean drift coordinate table
# IMPORTANT: one row per drift only
# ------------------------------------------------------------

drift_coords <- event_cluster_counts %>%
  mutate(
    DriftID_clean = stringr::str_remove(EventID_clean, "_[0-9]+$")
  ) %>%
  distinct(EventID_clean, DriftID_clean, Latitude, Longitude) %>%
  group_by(DriftID_clean) %>%
  summarise(
    Latitude = mean(Latitude, na.rm = TRUE),
    Longitude = mean(Longitude, na.rm = TRUE),
    n_events = n_distinct(EventID_clean),
    .groups = "drop"
  )

# ------------------------------------------------------------
# Build drift-level cluster counts
# This is the only real analytical change:
# event-level counts are summed up to drift-level counts
# ------------------------------------------------------------

drift_cluster_counts <- event_cluster_counts %>%
  mutate(
    HDBSCAN_cluster = as.character(HDBSCAN_cluster),
    DriftID_clean = stringr::str_remove(EventID_clean, "_[0-9]+$")
  ) %>%
  group_by(DriftID_clean, HDBSCAN_cluster) %>%
  summarise(
    n_clicks = sum(n_clicks, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(
    drift_coords,
    by = "DriftID_clean"
  )

cat("Rows in drift_cluster_counts:", nrow(drift_cluster_counts), "\n")
cat("Unique drifts:", dplyr::n_distinct(drift_cluster_counts$DriftID_clean), "\n")
cat("Total clicks represented:", sum(drift_cluster_counts$n_clicks), "\n\n")

print(drift_cluster_counts, n = Inf)

# ------------------------------------------------------------
# Make wide drift pie tables
# These mimic the original event pie tables
# ------------------------------------------------------------

drift_pies_including_noise <- drift_cluster_counts %>%
  select(
    DriftID_clean,
    Latitude,
    Longitude,
    n_events,
    HDBSCAN_cluster,
    n_clicks
  ) %>%
  tidyr::pivot_wider(
    names_from = HDBSCAN_cluster,
    values_from = n_clicks,
    values_fill = 0,
    names_prefix = "cluster_"
  ) %>%
  mutate(
    # Compatibility with original make_pie_slice_data()
    # It expects a column named EventID.
    EventID = DriftID_clean
  )

drift_pies_excluding_noise <- drift_cluster_counts %>%
  filter(HDBSCAN_cluster != "0") %>%
  select(
    DriftID_clean,
    Latitude,
    Longitude,
    n_events,
    HDBSCAN_cluster,
    n_clicks
  ) %>%
  tidyr::pivot_wider(
    names_from = HDBSCAN_cluster,
    values_from = n_clicks,
    values_fill = 0,
    names_prefix = "cluster_"
  ) %>%
  mutate(
    EventID = DriftID_clean
  )

# ------------------------------------------------------------
# Pie radius
# Use fixed size so pies are comparable by composition,
# not by number of clicks.
# ------------------------------------------------------------

drift_fixed_pie_radius <- 0.45

drift_pies_including_noise <- drift_pies_including_noise %>%
  mutate(pie_radius = drift_fixed_pie_radius)

drift_pies_excluding_noise <- drift_pies_excluding_noise %>%
  mutate(pie_radius = drift_fixed_pie_radius)

# ------------------------------------------------------------
# Save drift pie tables
# ------------------------------------------------------------

write.csv(
  drift_pies_including_noise,
  file.path(map_table_dir, "drift_pie_map_data_including_noise.csv"),
  row.names = FALSE
)

write.csv(
  drift_pies_excluding_noise,
  file.path(map_table_dir, "drift_pie_map_data_excluding_noise.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# Drift pie map function
# Uses original manual pie-slice workflow
# ------------------------------------------------------------

make_drift_pie_map <- function(pie_data, include_noise = TRUE, output_name) {
  
  cluster_cols <- names(pie_data)[stringr::str_detect(names(pie_data), "^cluster_")]
  
  cluster_cols <- cluster_cols[
    order(as.numeric(stringr::str_remove(cluster_cols, "cluster_")))
  ]
  
  pie_slices <- make_pie_slice_data(
    pie_data = pie_data,
    cluster_cols = cluster_cols,
    radius_col = "pie_radius"
  )
  
  xlim <- range(pie_data$Longitude, na.rm = TRUE) + c(-1.5, 1.5)
  ylim <- range(pie_data$Latitude, na.rm = TRUE) + c(-1.5, 1.5)
  
  p <- ggplot() +
    geom_sf(
      data = world,
      fill = "gray92",
      color = "gray60"
    ) +
    geom_polygon(
      data = pie_slices,
      aes(
        x = Longitude,
        y = Latitude,
        group = piece_id,
        fill = cluster
      ),
      color = "black",
      linewidth = 0.12,
      alpha = 0.95
    ) +
    scale_fill_manual(
      values = cluster_colors_consistent,
      drop = FALSE
    ) +
    coord_sf(
      xlim = xlim,
      ylim = ylim,
      expand = FALSE
    ) +
    labs(
      title = ifelse(
        include_noise,
        "Drift-Level HDBSCAN Cluster Composition: Including Noise",
        "Drift-Level HDBSCAN Cluster Composition: Excluding Noise"
      ),
      subtitle = "One pie per drift",
      x = "Longitude",
      y = "Latitude",
      fill = "HDBSCAN cluster"
    ) +
    theme_bw(base_size = 13) +
    theme(
      plot.title = element_text(face = "bold")
    )
  
  ggsave(
    filename = file.path(map_plot_dir, output_name),
    plot = p,
    width = 8.5,
    height = 7,
    dpi = 300
  )
  
  return(p)
}

# ------------------------------------------------------------
# Make drift pie maps
# ------------------------------------------------------------

p_drift_map_including_noise <- make_drift_pie_map(
  pie_data = drift_pies_including_noise,
  include_noise = TRUE,
  output_name = "hdbscan_drift_pie_map_including_noise_final_colors.png"
)

p_drift_map_excluding_noise <- make_drift_pie_map(
  pie_data = drift_pies_excluding_noise,
  include_noise = FALSE,
  output_name = "hdbscan_drift_pie_map_excluding_noise_final_colors.png"
)

print(p_drift_map_including_noise)
print(p_drift_map_excluding_noise)

# ============================================================
# DRIFT-LEVEL STACKED COMPOSITION PLOTS
# Same style as original event-level stacked plots
# ============================================================

make_drift_stacked_barplot <- function(pie_data, include_noise = TRUE, output_name) {
  
  cluster_cols <- names(pie_data)[stringr::str_detect(names(pie_data), "^cluster_")]
  
  cluster_cols <- cluster_cols[
    order(as.numeric(stringr::str_remove(cluster_cols, "cluster_")))
  ]
  
  bar_data <- pie_data %>%
    arrange(Longitude) %>%
    mutate(
      Drift_label = paste0(DriftID_clean, "\n", round(abs(Longitude), 1), "°W"),
      Drift_label = factor(Drift_label, levels = Drift_label)
    ) %>%
    select(
      EventID,
      DriftID_clean,
      Drift_label,
      Longitude,
      Latitude,
      all_of(cluster_cols)
    ) %>%
    tidyr::pivot_longer(
      cols = all_of(cluster_cols),
      names_to = "cluster",
      values_to = "n_clicks"
    ) %>%
    mutate(
      cluster = stringr::str_remove(cluster, "cluster_")
    ) %>%
    group_by(EventID, DriftID_clean, Drift_label, Longitude, Latitude) %>%
    mutate(
      prop_clicks = n_clicks / sum(n_clicks)
    ) %>%
    ungroup()
  
  fill_values <- if (include_noise) {
    cluster_colors_consistent
  } else {
    cluster_colors_consistent[names(cluster_colors_consistent) != "0"]
  }
  
  p <- ggplot(
    bar_data,
    aes(
      x = Drift_label,
      y = prop_clicks,
      fill = cluster
    )
  ) +
    geom_col(
      width = 0.9,
      color = "black",
      linewidth = 0.1
    ) +
    scale_fill_manual(
      values = fill_values,
      drop = FALSE
    ) +
    scale_y_continuous(
      labels = scales::percent_format(accuracy = 1),
      expand = c(0, 0)
    ) +
    labs(
      title = ifelse(
        include_noise,
        "Drift-Level HDBSCAN Cluster Composition: Including Noise",
        "Drift-Level HDBSCAN Cluster Composition: Excluding Noise"
      ),
      subtitle = "Drifts ordered west-to-east by longitude",
      x = "Drift",
      y = "Percent of analyzed clicks",
      fill = "HDBSCAN cluster"
    ) +
    theme_bw(base_size = 13) +
    theme(
      plot.title = element_text(face = "bold"),
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "right"
    )
  
  ggsave(
    filename = file.path(map_plot_dir, output_name),
    plot = p,
    width = 11,
    height = 6,
    dpi = 300
  )
  
  return(p)
}

# ------------------------------------------------------------
# Make drift stacked plots
# ------------------------------------------------------------

p_drift_bar_including_noise <- make_drift_stacked_barplot(
  pie_data = drift_pies_including_noise,
  include_noise = TRUE,
  output_name = "hdbscan_drift_stacked_composition_including_noise_final_colors.png"
)

p_drift_bar_excluding_noise <- make_drift_stacked_barplot(
  pie_data = drift_pies_excluding_noise,
  include_noise = FALSE,
  output_name = "hdbscan_drift_stacked_composition_excluding_noise_final_colors.png"
)

print(p_drift_bar_including_noise)
print(p_drift_bar_excluding_noise)

cat("\nDONE: Drift-level pie maps and stacked plots saved to:\n")
cat(map_plot_dir, "\n")

```

## Plot Groups for Crowded Areas

```{r}

# ------------------------------------------------------------
# Define manual ADRIFT plotting groups using coordinate zones
#
# These groups affect plotting only.
# Original DriftID_clean and HDBSCAN assignments are unchanged.
# ------------------------------------------------------------

manual_grouped_drift_counts <- drift_cluster_counts %>%
  mutate(
    PlotGroupID = case_when(
      
      # ------------------------------------------------------
      # Northern nearshore ADRIFT group
      # Around ~41N, ~124W
      # ------------------------------------------------------
      stringr::str_detect(DriftID_clean, "^ADRIFT") &
        Latitude >= 40.3 &
        Latitude <= 41.5 &
        Longitude >= -125.0 &
        Longitude <= -124.0 ~ "ADRIFT_North_41N_124W",
      
      # ------------------------------------------------------
      # Central nearshore ADRIFT group
      # Around ~38N, ~124W
      # ------------------------------------------------------
      stringr::str_detect(DriftID_clean, "^ADRIFT") &
        Latitude >= 37.3 &
        Latitude <= 38.5 &
        Longitude >= -124.2 &
        Longitude <= -123.2 ~ "ADRIFT_Central_38N_124W",
      
      # ------------------------------------------------------
      # Southern nearshore ADRIFT group
      # Around ~36N, ~122W
      # ------------------------------------------------------
      stringr::str_detect(DriftID_clean, "^ADRIFT") &
        Latitude >= 35.2 &
        Latitude <= 36.3 &
        Longitude >= -122.6 &
        Longitude <= -121.2 ~ "ADRIFT_South_36N_122W",
      
      # ------------------------------------------------------
      # Everything else stays as one pie per drift
      # ------------------------------------------------------
      TRUE ~ DriftID_clean
    )
  )

# ------------------------------------------------------------
# Check exactly which drifts are assigned to each plot group
# ------------------------------------------------------------

manual_plot_group_membership <- manual_grouped_drift_counts %>%
  distinct(
    PlotGroupID,
    DriftID_clean,
    Latitude,
    Longitude
  ) %>%
  group_by(PlotGroupID) %>%
  summarise(
    n_drifts_in_group = n_distinct(DriftID_clean),
    drifts_in_group = paste(sort(unique(DriftID_clean)), collapse = "; "),
    group_latitude = mean(Latitude, na.rm = TRUE),
    group_longitude = mean(Longitude, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(group_latitude)

cat("\nManual ADRIFT plot group membership:\n")
print(manual_plot_group_membership, n = Inf)

write.csv(
  manual_plot_group_membership,
  file.path(map_table_dir, "manual_adrift_plot_group_membership_improved.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# Create ONE coordinate per plotting group
# ------------------------------------------------------------

manual_group_coordinates <- manual_grouped_drift_counts %>%
  group_by(PlotGroupID) %>%
  summarise(
    Latitude = mean(Latitude, na.rm = TRUE),
    Longitude = mean(Longitude, na.rm = TRUE),
    n_drifts_in_group = n_distinct(DriftID_clean),
    drifts_in_group = paste(sort(unique(DriftID_clean)), collapse = "; "),
    .groups = "drop"
  )

# ------------------------------------------------------------
# Collapse cluster counts to plotting groups
# IMPORTANT:
# coordinates are added AFTER cluster summarisation
# so every cluster slice shares the SAME pie center
# ------------------------------------------------------------

manual_grouped_cluster_counts <- manual_grouped_drift_counts %>%
  group_by(PlotGroupID, HDBSCAN_cluster) %>%
  summarise(
    n_clicks = sum(n_clicks, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(
    manual_group_coordinates,
    by = "PlotGroupID"
  )

cat("\nManual grouped cluster counts:\n")
print(manual_grouped_cluster_counts, n = Inf)

write.csv(
  manual_grouped_cluster_counts,
  file.path(map_table_dir, "manual_adrift_grouped_cluster_counts_improved.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# Wide pie tables: including noise
# ------------------------------------------------------------

manual_grouped_pies_including_noise <- manual_grouped_cluster_counts %>%
  select(
    PlotGroupID,
    Latitude,
    Longitude,
    n_drifts_in_group,
    drifts_in_group,
    HDBSCAN_cluster,
    n_clicks
  ) %>%
  tidyr::pivot_wider(
    names_from = HDBSCAN_cluster,
    values_from = n_clicks,
    values_fill = 0,
    names_prefix = "cluster_"
  ) %>%
  mutate(
    EventID = PlotGroupID,
    pie_radius = 0.45
  )

# ------------------------------------------------------------
# Wide pie tables: excluding noise
# ------------------------------------------------------------

manual_grouped_pies_excluding_noise <- manual_grouped_cluster_counts %>%
  filter(HDBSCAN_cluster != "0") %>%
  select(
    PlotGroupID,
    Latitude,
    Longitude,
    n_drifts_in_group,
    drifts_in_group,
    HDBSCAN_cluster,
    n_clicks
  ) %>%
  tidyr::pivot_wider(
    names_from = HDBSCAN_cluster,
    values_from = n_clicks,
    values_fill = 0,
    names_prefix = "cluster_"
  ) %>%
  mutate(
    EventID = PlotGroupID,
    pie_radius = 0.45
  )

write.csv(
  manual_grouped_pies_including_noise,
  file.path(map_table_dir, "manual_adrift_grouped_pie_data_including_noise_improved.csv"),
  row.names = FALSE
)

write.csv(
  manual_grouped_pies_excluding_noise,
  file.path(map_table_dir, "manual_adrift_grouped_pie_data_excluding_noise_improved.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# Map helper
# Same style as existing drift-level pie maps
# ------------------------------------------------------------

make_manual_grouped_pie_map <- function(pie_data, include_noise = TRUE, output_name) {
  
  cluster_cols <- names(pie_data)[stringr::str_detect(names(pie_data), "^cluster_")]
  
  cluster_cols <- cluster_cols[
    order(as.numeric(stringr::str_remove(cluster_cols, "cluster_")))
  ]
  
  pie_slices <- make_pie_slice_data(
    pie_data = pie_data,
    cluster_cols = cluster_cols,
    radius_col = "pie_radius"
  )
  
  xlim <- range(pie_data$Longitude, na.rm = TRUE) + c(-1.5, 1.5)
  ylim <- range(pie_data$Latitude, na.rm = TRUE) + c(-1.5, 1.5)
  
  p <- ggplot() +
    geom_sf(
      data = world,
      fill = "gray92",
      color = "gray60"
    ) +
    geom_polygon(
      data = pie_slices,
      aes(
        x = Longitude,
        y = Latitude,
        group = piece_id,
        fill = cluster
      ),
      color = "black",
      linewidth = 0.12,
      alpha = 0.95
    ) +
    scale_fill_manual(
      values = cluster_colors_consistent,
      drop = FALSE
    ) +
    coord_sf(
      xlim = xlim,
      ylim = ylim,
      expand = FALSE
    ) +
    labs(
      title = ifelse(
        include_noise,
        "Manual ADRIFT Grouped Drift Pies: Including Noise",
        "Manual ADRIFT Grouped Drift Pies: Excluding Noise"
      ),
      subtitle = "Crowded ADRIFT drifts grouped near ~41N, ~38N, and ~36N for visualization only",
      x = "Longitude",
      y = "Latitude",
      fill = "HDBSCAN cluster"
    ) +
    theme_bw(base_size = 13) +
    theme(
      plot.title = element_text(face = "bold")
    )
  
  ggsave(
    filename = file.path(map_plot_dir, output_name),
    plot = p,
    width = 8.5,
    height = 7,
    dpi = 300
  )
  
  return(p)
}

# ------------------------------------------------------------
# Make improved grouped pie maps
# ------------------------------------------------------------

p_manual_grouped_map_including_noise <- make_manual_grouped_pie_map(
  pie_data = manual_grouped_pies_including_noise,
  include_noise = TRUE,
  output_name = "hdbscan_manual_adrift_grouped_pie_map_including_noise_improved.png"
)

p_manual_grouped_map_excluding_noise <- make_manual_grouped_pie_map(
  pie_data = manual_grouped_pies_excluding_noise,
  include_noise = FALSE,
  output_name = "hdbscan_manual_adrift_grouped_pie_map_excluding_noise_improved.png"
)

print(p_manual_grouped_map_including_noise)
print(p_manual_grouped_map_excluding_noise)

cat("\nDONE: Improved manual ADRIFT grouped pie maps saved to:\n")
cat(map_plot_dir, "\n")

```

## Mean click spectra by HDBSCAN cluster with 95% CI ribbons

```{r}

library(dplyr)
library(ggplot2)

# ------------------------------------------------------------
# Output folder
# ------------------------------------------------------------

minpts50_plot_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/Main_minPts50_5PC_scaled_logICI/plots"

dir.create(
  minpts50_plot_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

# ------------------------------------------------------------
# Build 95% CI table from mean_spectra_minpts50
# ------------------------------------------------------------

mean_spectra_minpts50_ci <- mean_spectra_minpts50 %>%
  mutate(
    HDBSCAN_cluster = as.character(HDBSCAN_cluster),
    se_minmax_spectrum = sd_minmax_spectrum / sqrt(n_clicks),
    ci95_lower = mean_minmax_spectrum - 1.96 * se_minmax_spectrum,
    ci95_upper = mean_minmax_spectrum + 1.96 * se_minmax_spectrum,
    ci95_lower_plot = pmax(ci95_lower, 0),
    ci95_upper_plot = pmin(ci95_upper, 1)
  ) %>%
  filter(
    HDBSCAN_cluster %in% c("1", "2", "3", "4")
  )

# ------------------------------------------------------------
# Sanity check
# ------------------------------------------------------------

cat("\nClusters included in CI spectra figure:\n")

mean_spectra_minpts50_ci %>%
  distinct(HDBSCAN_cluster, n_clicks) %>%
  arrange(as.numeric(HDBSCAN_cluster)) %>%
  print()

# ------------------------------------------------------------
# Plot: mean spectra with 95% CI ribbons
# ------------------------------------------------------------

p_mean_spectra_minpts50_ci <- ggplot(
  mean_spectra_minpts50_ci,
  aes(
    x = Frequency_kHz,
    y = mean_minmax_spectrum
  )
) +
  geom_ribbon(
    aes(
      ymin = ci95_lower_plot,
      ymax = ci95_upper_plot
    ),
    fill = "gray70",
    alpha = 0.55
  ) +
  geom_line(
    linewidth = 1.1,
    color = "black"
  ) +
  facet_wrap(
    ~ HDBSCAN_cluster,
    ncol = 2
  ) +
  coord_cartesian(
    ylim = c(0, 1)
  ) +
  labs(
    title = "Mean Click Spectra by HDBSCAN Cluster",
    subtitle = "Mean ± 95% CI; 5PC + scaled logICI; minPts = 50; noise excluded",
    x = "Frequency (kHz)",
    y = "Min-max normalized amplitude"
  ) +
  theme_bw(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold"),
    strip.background = element_rect(fill = "gray85", color = "gray40"),
    strip.text = element_text(face = "bold")
  )

print(p_mean_spectra_minpts50_ci)

# ------------------------------------------------------------
# Save figure and table
# ------------------------------------------------------------

ggsave(
  filename = file.path(
    minpts50_plot_dir,
    "hdbscan_minPts50_mean_click_spectra_95CI_noise_excluded.png"
  ),
  plot = p_mean_spectra_minpts50_ci,
  width = 11,
  height = 8,
  dpi = 300
)

ggsave(
  filename = file.path(
    minpts50_plot_dir,
    "hdbscan_minPts50_mean_click_spectra_95CI_noise_excluded.pdf"
  ),
  plot = p_mean_spectra_minpts50_ci,
  width = 11,
  height = 8
)

write.csv(
  mean_spectra_minpts50_ci,
  file.path(
    minpts50_plot_dir,
    "hdbscan_minPts50_mean_click_spectra_95CI_noise_excluded.csv"
  ),
  row.names = FALSE
)

cat("\nDONE: Mean click spectra with 95% CI saved to:\n")
cat(minpts50_plot_dir, "\n")

```

## Mean spectra with Variability Ribbons for All Non-Noise Clusters Color-coded to match maps / stacked plots

```{r}

library(dplyr)
library(ggplot2)

# ------------------------------------------------------------
# Output folder
# ------------------------------------------------------------

minpts50_plot_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/Main_minPts50_5PC_scaled_logICI/plots"

dir.create(
  minpts50_plot_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

# ------------------------------------------------------------
# Cluster colors used across maps / stacked plots
# ------------------------------------------------------------

cluster_colors_consistent <- c(
  "0" = "gray60",
  "1" = "#F8766D",
  "2" = "#7CAE00",
  "3" = "#1FBFCF",
  "4" = "#C77CFF"
)

cluster_colors_no_noise <- cluster_colors_consistent[c("1", "2", "3", "4")]

# ------------------------------------------------------------
# Use non-noise HDBSCAN clusters from final minPts50 analysis
# ------------------------------------------------------------

cluster_spectra_ribbon_data <- mean_spectra_minpts50 %>%
  mutate(
    HDBSCAN_cluster = as.character(HDBSCAN_cluster)
  ) %>%
  filter(
    HDBSCAN_cluster %in% c("1", "2", "3", "4")
  ) %>%
  mutate(
    lower_sd = pmax(mean_minmax_spectrum - sd_minmax_spectrum, 0),
    upper_sd = pmin(mean_minmax_spectrum + sd_minmax_spectrum, 1)
  )

# ------------------------------------------------------------
# Sanity check
# ------------------------------------------------------------

cat("\nClusters included in SD spectra figure:\n")

cluster_spectra_ribbon_data %>%
  distinct(HDBSCAN_cluster, n_clicks) %>%
  arrange(as.numeric(HDBSCAN_cluster)) %>%
  print()

# ------------------------------------------------------------
# Faceted mean ± SD ribbons
# ------------------------------------------------------------

p_cluster_spectra_ribbons <- ggplot(
  cluster_spectra_ribbon_data,
  aes(
    x = Frequency_kHz,
    y = mean_minmax_spectrum
  )
) +
  geom_ribbon(
    aes(
      ymin = lower_sd,
      ymax = upper_sd,
      fill = HDBSCAN_cluster
    ),
    alpha = 0.25,
    color = NA
  ) +
  geom_line(
    aes(color = HDBSCAN_cluster),
    linewidth = 1
  ) +
  facet_wrap(
    ~ HDBSCAN_cluster,
    ncol = 2
  ) +
  scale_fill_manual(
    values = cluster_colors_no_noise,
    drop = FALSE
  ) +
  scale_color_manual(
    values = cluster_colors_no_noise,
    drop = FALSE
  ) +
  coord_cartesian(
    ylim = c(0, 1)
  ) +
  labs(
    title = "Mean Spectra with Variability by HDBSCAN Cluster",
    subtitle = "Mean ± 1 SD; 5PC + scaled logICI; minPts = 50; noise excluded",
    x = "Frequency (kHz)",
    y = "Min-max normalized amplitude",
    fill = "HDBSCAN cluster",
    color = "HDBSCAN cluster"
  ) +
  theme_bw(base_size = 13) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold"),
    strip.background = element_rect(fill = "gray85", color = "gray40"),
    strip.text = element_text(face = "bold")
  )

print(p_cluster_spectra_ribbons)

ggsave(
  filename = file.path(
    minpts50_plot_dir,
    "hdbscan_minPts50_mean_spectra_SD_colorcoded_noise_excluded.png"
  ),
  plot = p_cluster_spectra_ribbons,
  width = 10,
  height = 8,
  dpi = 300
)

# ------------------------------------------------------------
# Save ribbon data
# ------------------------------------------------------------

write.csv(
  cluster_spectra_ribbon_data,
  file = file.path(
    minpts50_plot_dir,
    "hdbscan_minPts50_mean_spectra_SD_colorcoded_noise_excluded.csv"
  ),
  row.names = FALSE
)

cat("\nDONE: Color-coded mean spectra with ±1 SD saved to:\n")
cat(minpts50_plot_dir, "\n")

```




























