---
title: "PCA_ICI_UMAP_HDBSCAN_90PercentPC"
author: "Sarah Moseley"
date: "2026-05-09"
output: html_document
---

## Setup

```{r setup}

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)

required_packages <- c(
  "dplyr",
  "ggplot2",
  "tidyr",
  "stringr",
  "readr",
  "purrr",
  "tibble",
  "RColorBrewer",
  "uwot",
  "dbscan",
  "scales"
)

missing_packages <- required_packages[
  !sapply(required_packages, requireNamespace, quietly = TRUE)
]

if (length(missing_packages) > 0) {
  stop(
    "Missing required packages: ",
    paste(missing_packages, collapse = ", "),
    "\nInstall them before running this workflow."
  )
}

library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
library(readr)
library(purrr)
library(tibble)
library(RColorBrewer)
library(uwot)
library(dbscan)
library(scales)

```

## User Settings

```{r user_settings}

# ------------------------------------------------------------
# Input files
# ------------------------------------------------------------

# Corrected combined Tukey object
combined_file <- file.choose()

# Click-level ICI table
# click_metadata_with_ici_NEW.rds
ici_file <- file.choose()

# ------------------------------------------------------------
# Base output directory
# ------------------------------------------------------------

base_output_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN"

dir.create(base_output_dir,
           recursive = TRUE,
           showWarnings = FALSE)

# ------------------------------------------------------------
# Spectral settings
# ------------------------------------------------------------

freq_min_khz <- 19
freq_max_khz <- 50

# ------------------------------------------------------------
# ICI filtering settings
# ------------------------------------------------------------

ici_min_s <- 0.01
ici_max_s <- 0.4

# ------------------------------------------------------------
# Sampling settings
# ------------------------------------------------------------

clicks_per_event <- 200
sampling_seed <- 123

# ------------------------------------------------------------
# PCA settings
# ------------------------------------------------------------

target_variance <- 0.90

# ------------------------------------------------------------
# ICI scaling experiments
# ------------------------------------------------------------

ici_scaling_values <- c(1.0, 0.75, 0.5, 0.25)

# ------------------------------------------------------------
# UMAP settings
# ------------------------------------------------------------

umap_neighbors <- 30
umap_min_dist <- 0.1
umap_metric <- "euclidean"
umap_seed <- 123

# ------------------------------------------------------------
# HDBSCAN settings
# ------------------------------------------------------------

hdbscan_minPts_values <- c(10, 30, 50)

# ------------------------------------------------------------
# Concatenated spectrogram settings
# ------------------------------------------------------------

max_clicks_per_cluster_spectrogram <- 300
cluster_spectrogram_seed <- 456

# ------------------------------------------------------------
# CONSISTENT CLUSTER COLORS
#
# IMPORTANT:
# Cluster 0 (noise) ALWAYS light grey.
# These colors NEVER change across plots.
# ------------------------------------------------------------

cluster_color_map <- c(
  "0" = "grey80",
  "1" = "#E41A1C",  # red
  "2" = "#FF7F00",  # orange
  "3" = "#FFD92F",  # yellow
  "4" = "#4DAF4A",  # green
  "5" = "#377EB8",  # blue
  "6" = "#984EA3",  # purple
  "7" = "#F781BF",  # pink
  "8" = "#A65628",  # brown
  "9" = "#999999"   # fallback grey
)

cat("Base output directory:\n")
cat(base_output_dir, "\n")

```

## Load Combined Tukey and ICI Tables

```{r load_data}

# ------------------------------------------------------------
# Load corrected combined Tukey object
# ------------------------------------------------------------

combined_tukey <- readRDS(combined_file)

cat("Loaded combined_tukey object.\n")
cat("Number of events:", length(combined_tukey), "\n\n")

# ------------------------------------------------------------
# Load click-level ICI table
# ------------------------------------------------------------

click_metadata_with_ici <- readRDS(ici_file)

cat("Loaded click_metadata_with_ici object.\n")
cat("Rows in ICI table:", nrow(click_metadata_with_ici), "\n\n")

# ------------------------------------------------------------
# Preview structures
# ------------------------------------------------------------

cat("Class of combined_tukey:\n")
print(class(combined_tukey))

cat("\nClass of first event object:\n")
print(class(combined_tukey[[1]]))

cat("\nNames in first event object:\n")
print(names(combined_tukey[[1]]))

cat("\nColumns in click_metadata_with_ici:\n")
print(names(click_metadata_with_ici))

# ------------------------------------------------------------
# Quick sanity preview
# ------------------------------------------------------------

cat("\nPreview of first event:\n")
print(combined_tukey[[1]]$EventID)

cat("\nPreview of ICI table:\n")
print(head(click_metadata_with_ici))

```

## Sanity Checks for Combined Tukey Structure

```{r sanity_check_combined_tukey}

event_structure_summary <- data.frame(
  
  EventIndex = seq_along(combined_tukey),
  
  EventID = sapply(combined_tukey, function(x) x$EventID),
  
  n_freq = sapply(combined_tukey, function(x) length(x$freq)),
  
  n_uid = sapply(combined_tukey, function(x) length(x$UID)),
  
  allSpec_rows = sapply(combined_tukey, function(x) dim(x$allSpec)[1]),
  
  allSpec_cols = sapply(combined_tukey, function(x) dim(x$allSpec)[2]),
  
  stringsAsFactors = FALSE
)

print(event_structure_summary)

# ------------------------------------------------------------
# Total click count
# ------------------------------------------------------------

total_clicks <- sum(event_structure_summary$allSpec_cols)

cat("\nTotal clicks across all events:", total_clicks, "\n")

# ------------------------------------------------------------
# Core dimension checks
# ------------------------------------------------------------

uid_matches_allSpec <- all(
  event_structure_summary$n_uid ==
    event_structure_summary$allSpec_cols
)

freq_matches_allSpec <- all(
  event_structure_summary$n_freq ==
    event_structure_summary$allSpec_rows
)

cat("\nUID count matches allSpec columns:\n")
print(uid_matches_allSpec)

cat("\nFrequency count matches allSpec rows:\n")
print(freq_matches_allSpec)

if (uid_matches_allSpec && freq_matches_allSpec) {
  
  cat("\nPASS: Core spectral dimensions are aligned.\n")
  
} else {
  
  stop("FAIL: Spectral dimensions are misaligned.")
}

# ------------------------------------------------------------
# Save structure summary temporarily to base directory
# Final 90PC-specific folders get created after PCA determines PC count.
# ------------------------------------------------------------

write.csv(
  event_structure_summary,
  file = file.path(base_output_dir, "event_structure_summary_pre_pca.csv"),
  row.names = FALSE
)

```

## Check for Duplicated EventIDs and UID Problems

```{r check_duplicate_eventids_uids}

# ------------------------------------------------------------
# Build click-level metadata from Tukey object
# ------------------------------------------------------------

click_metadata_from_tukey <- list()

for (i in seq_along(combined_tukey)) {
  
  event_obj <- combined_tukey[[i]]
  
  n_clicks <- ncol(event_obj$allSpec)
  
  click_metadata_from_tukey[[i]] <- data.frame(
    
    EventIndex = i,
    
    EventID = event_obj$EventID,
    
    UID = as.character(event_obj$UID),
    
    ClickIndexWithinEvent = seq_len(n_clicks),
    
    stringsAsFactors = FALSE
  )
}

click_metadata_from_tukey <- bind_rows(click_metadata_from_tukey)

cat("Rows in Tukey click metadata:",
    nrow(click_metadata_from_tukey), "\n")

# ------------------------------------------------------------
# Duplicate EventID check
# ------------------------------------------------------------

duplicate_event_ids <- click_metadata_from_tukey %>%
  distinct(EventIndex, EventID) %>%
  count(EventID, name = "n_rows") %>%
  filter(n_rows > 1)

cat("\nDuplicated EventIDs:\n")
print(duplicate_event_ids)

# ------------------------------------------------------------
# Duplicate UID within EventID
# ------------------------------------------------------------

duplicate_uid_within_event <- click_metadata_from_tukey %>%
  count(EventID, UID, name = "n_rows") %>%
  filter(n_rows > 1)

cat("\nDuplicated UID values within EventID:\n")
print(duplicate_uid_within_event)

# ------------------------------------------------------------
# PASS / FAIL checks
# ------------------------------------------------------------

if (nrow(duplicate_event_ids) == 0) {
  
  cat("\nPASS: No duplicated EventIDs.\n")
  
} else {
  
  stop("FAIL: Duplicated EventIDs detected.")
}

if (nrow(duplicate_uid_within_event) == 0) {
  
  cat("\nPASS: No duplicated UID values within events.\n")
  
} else {
  
  stop("FAIL: Duplicated UID values detected within events.")
}

```

## Inspect Frequency Grids Before Interpolation

```{r inspect_frequency_grids}

freq_grid_summary <- data.frame(
  EventIndex = seq_along(combined_tukey),
  EventID = sapply(combined_tukey, function(x) x$EventID),
  n_freq = sapply(combined_tukey, function(x) length(x$freq)),
  min_freq = sapply(combined_tukey, function(x) min(x$freq, na.rm = TRUE)),
  max_freq = sapply(combined_tukey, function(x) max(x$freq, na.rm = TRUE)),
  median_freq_step = sapply(combined_tukey, function(x) median(diff(x$freq), na.rm = TRUE)),
  stringsAsFactors = FALSE
)

print(freq_grid_summary)

# ------------------------------------------------------------
# Save frequency-grid summary to base folder for now
# PC-specific output folders are created after PCA determines n_pcs_90.
# ------------------------------------------------------------

write.csv(
  freq_grid_summary,
  file = file.path(base_output_dir, "frequency_grid_summary_before_interpolation.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# Check whether all frequency grids are already identical
# ------------------------------------------------------------

reference_freq_original <- combined_tukey[[1]]$freq

all_freq_grids_identical <- all(
  sapply(
    combined_tukey,
    function(x) identical(x$freq, reference_freq_original)
  )
)

cat("\nAre all original frequency grids identical?\n")
print(all_freq_grids_identical)

```

## Interpolate All Events to Shared Frequency Grid

```{r interpolate_to_shared_frequency_grid}

# ------------------------------------------------------------
# Use first event frequency grid as reference
# ------------------------------------------------------------

reference_freq <- combined_tukey[[1]]$freq

cat("Reference grid length:", length(reference_freq), "\n")
cat("Reference min frequency:", min(reference_freq), "\n")
cat("Reference max frequency:", max(reference_freq), "\n")
cat("Reference median step:", median(diff(reference_freq)), "\n")

# ------------------------------------------------------------
# Helper function: interpolate one matrix to reference grid
# ------------------------------------------------------------

interpolate_spectra_matrix <- function(original_freq, spectra_matrix, reference_freq) {
  
  interpolated_matrix <- apply(
    spectra_matrix,
    2,
    function(one_spectrum) {
      approx(
        x = original_freq,
        y = one_spectrum,
        xout = reference_freq,
        method = "linear",
        rule = 2
      )$y
    }
  )
  
  return(interpolated_matrix)
}

# ------------------------------------------------------------
# Interpolate each event
# ------------------------------------------------------------

interpolated_tukey <- vector("list", length(combined_tukey))

for (i in seq_along(combined_tukey)) {
  
  event_obj <- combined_tukey[[i]]
  
  cat("\nInterpolating event", i, "of", length(combined_tukey), ":",
      event_obj$EventID, "\n")
  
  interpolated_allSpec <- interpolate_spectra_matrix(
    original_freq = event_obj$freq,
    spectra_matrix = event_obj$allSpec,
    reference_freq = reference_freq
  )
  
  interpolated_tukey[[i]] <- event_obj
  interpolated_tukey[[i]]$freq <- reference_freq
  interpolated_tukey[[i]]$allSpec <- interpolated_allSpec
}

# ------------------------------------------------------------
# Sanity check: all interpolated grids identical
# ------------------------------------------------------------

all_interpolated_grids_identical <- all(
  sapply(
    interpolated_tukey,
    function(x) identical(x$freq, reference_freq)
  )
)

cat("\nAre all interpolated frequency grids identical?\n")
print(all_interpolated_grids_identical)

if (all_interpolated_grids_identical) {
  cat("\nPASS: All events now share the same frequency grid.\n")
} else {
  stop("FAIL: Frequency-grid interpolation did not align all events.")
}

```

## Restrict Spectra to 19-50 kHz

```{r restrict_spectra_19_50_khz}

# ------------------------------------------------------------
# Convert reference frequency grid to kHz if needed
# ------------------------------------------------------------

if (max(reference_freq, na.rm = TRUE) > 1000) {
  reference_freq_khz <- reference_freq / 1000
  cat("Detected frequency grid in Hz. Converted to kHz.\n")
} else {
  reference_freq_khz <- reference_freq
  cat("Detected frequency grid already in kHz.\n")
}

# ------------------------------------------------------------
# Identify frequency bins inside target range
# ------------------------------------------------------------

freq_keep <- reference_freq_khz >= freq_min_khz &
  reference_freq_khz <= freq_max_khz

freq_focus <- reference_freq_khz[freq_keep]

cat("Frequency range requested:", freq_min_khz, "to", freq_max_khz, "kHz\n")
cat("Number of frequency bins retained:", length(freq_focus), "\n")

if (length(freq_focus) > 0) {
  cat("Actual minimum retained frequency:", min(freq_focus), "kHz\n")
  cat("Actual maximum retained frequency:", max(freq_focus), "kHz\n")
}

if (length(freq_focus) < 2) {
  stop("Too few frequency bins retained. Check frequency units and freq_min_khz/freq_max_khz.")
}

# ------------------------------------------------------------
# Save retained frequency grid to base folder for now
# ------------------------------------------------------------

freq_focus_table <- data.frame(
  FreqBin = seq_along(freq_focus),
  Frequency_kHz = freq_focus,
  Frequency_OriginalUnits = reference_freq[freq_keep]
)

write.csv(
  freq_focus_table,
  file = file.path(base_output_dir, "frequency_grid_19_50_khz.csv"),
  row.names = FALSE
)

```

## Build Click-By-Frequency Spectral Matrix

```{r build_click_by_frequency_matrix}

# ------------------------------------------------------------
# Build one click-level spectral matrix and matching metadata table
# ------------------------------------------------------------

spectra_list <- vector("list", length(interpolated_tukey))
metadata_list <- vector("list", length(interpolated_tukey))

for (i in seq_along(interpolated_tukey)) {
  
  event_obj <- interpolated_tukey[[i]]
  
  # allSpec is frequency x click.
  # First restrict to 19–50 kHz frequency bins, then transpose:
  # rows = clicks, columns = frequency bins.
  spectra_focus_event <- t(event_obj$allSpec[freq_keep, , drop = FALSE])
  
  spectra_list[[i]] <- spectra_focus_event
  
  metadata_list[[i]] <- data.frame(
    EventIndex = i,
    EventID = event_obj$EventID,
    DriftID = sub("_Event_.*$", "", event_obj$EventID),
    UID = as.character(event_obj$UID),
    ClickIndexWithinEvent = seq_len(ncol(event_obj$allSpec)),
    stringsAsFactors = FALSE
  )
}

# ------------------------------------------------------------
# Combine all events into one matrix/table
# ------------------------------------------------------------

spectra_matrix <- do.call(rbind, spectra_list)
spectra_metadata <- bind_rows(metadata_list)

colnames(spectra_matrix) <- paste0(
  "freq_",
  sprintf("%.3f", freq_focus),
  "_kHz"
)

# ------------------------------------------------------------
# Sanity checks
# ------------------------------------------------------------

cat("Rows in spectra matrix:", nrow(spectra_matrix), "\n")
cat("Columns in spectra matrix:", ncol(spectra_matrix), "\n")
cat("Rows in spectra metadata:", nrow(spectra_metadata), "\n")

if (nrow(spectra_matrix) != nrow(spectra_metadata)) {
  stop("FAIL: Spectra matrix rows do not match metadata rows.")
} else {
  cat("\nPASS: Spectra matrix and metadata rows are aligned.\n")
}

if (ncol(spectra_matrix) != length(freq_focus)) {
  stop("FAIL: Spectra matrix columns do not match retained frequency bins.")
} else {
  cat("PASS: Spectra matrix columns match retained frequency bins.\n")
}

cat("\nPreview spectra metadata:\n")
print(head(spectra_metadata))

cat("\nSpectra matrix dimensions:\n")
print(dim(spectra_matrix))

```

## Sanity Check Spectra Values

```{r sanity_check_spectra_values}

# ------------------------------------------------------------
# Basic matrix checks
# ------------------------------------------------------------

cat("Spectra matrix dimensions:\n")
print(dim(spectra_matrix))

cat("\nNumber of NA values in spectra matrix:\n")
print(sum(is.na(spectra_matrix)))

cat("\nNumber of infinite / non-finite values in spectra matrix:\n")
print(sum(!is.finite(spectra_matrix)))

if (sum(is.na(spectra_matrix)) > 0 || sum(!is.finite(spectra_matrix)) > 0) {
  stop("FAIL: Spectra matrix contains NA, NaN, or infinite values.")
}

# ------------------------------------------------------------
# Check per-click min/max values
# ------------------------------------------------------------

spectra_row_min <- apply(spectra_matrix, 1, min, na.rm = TRUE)
spectra_row_max <- apply(spectra_matrix, 1, max, na.rm = TRUE)
spectra_row_range <- spectra_row_max - spectra_row_min

cat("\nSummary of per-click minimum values:\n")
print(summary(spectra_row_min))

cat("\nSummary of per-click maximum values:\n")
print(summary(spectra_row_max))

cat("\nSummary of per-click spectral ranges:\n")
print(summary(spectra_row_range))

# ------------------------------------------------------------
# Flag flat spectra, if any
# ------------------------------------------------------------

flat_spectra_rows <- which(spectra_row_range == 0)

cat("\nNumber of flat spectra rows:", length(flat_spectra_rows), "\n")

if (length(flat_spectra_rows) > 0) {
  
  cat("\nWARNING: Some spectra have zero range across retained frequencies.\n")
  cat("These may cause issues later if additional scaling is applied.\n")
  
  flat_spectra_table <- spectra_metadata[flat_spectra_rows, ]
  
  write.csv(
    flat_spectra_table,
    file = file.path(base_output_dir, "flat_spectra_rows_19_50_khz.csv"),
    row.names = FALSE
  )
}

# ------------------------------------------------------------
# Save row-level min/max diagnostics
# ------------------------------------------------------------

spectra_row_diagnostics <- spectra_metadata %>%
  mutate(
    row_min_19_50_khz = spectra_row_min,
    row_max_19_50_khz = spectra_row_max,
    row_range_19_50_khz = spectra_row_range
  )

write.csv(
  spectra_row_diagnostics,
  file = file.path(base_output_dir, "spectra_row_min_max_diagnostics_19_50_khz.csv"),
  row.names = FALSE
)

cat("\nPASS: Spectra matrix contains finite values only.\n")

```

## Min-Max Normalize Spectra Across 19-50 kHz

```{r minmax_normalize_spectra}

# ------------------------------------------------------------
# Min-max normalize each click spectrum across 19–50 kHz
#
# rows = clicks
# columns = frequency bins
#
# Formula:
# normalized = (value - row_min) / (row_max - row_min)
# ------------------------------------------------------------

spectra_matrix_minmax <- sweep(
  spectra_matrix,
  1,
  spectra_row_min,
  FUN = "-"
)

spectra_matrix_minmax <- sweep(
  spectra_matrix_minmax,
  1,
  spectra_row_range,
  FUN = "/"
)

# ------------------------------------------------------------
# Sanity checks
# ------------------------------------------------------------

minmax_row_min <- apply(spectra_matrix_minmax, 1, min, na.rm = TRUE)
minmax_row_max <- apply(spectra_matrix_minmax, 1, max, na.rm = TRUE)

cat("Min-max normalized spectra dimensions:\n")
print(dim(spectra_matrix_minmax))

cat("\nSummary of normalized per-click minimum values:\n")
print(summary(minmax_row_min))

cat("\nSummary of normalized per-click maximum values:\n")
print(summary(minmax_row_max))

cat("\nNumber of NA values in normalized spectra:\n")
print(sum(is.na(spectra_matrix_minmax)))

cat("\nNumber of non-finite values in normalized spectra:\n")
print(sum(!is.finite(spectra_matrix_minmax)))

if (sum(is.na(spectra_matrix_minmax)) > 0 || sum(!is.finite(spectra_matrix_minmax)) > 0) {
  stop("FAIL: Min-max normalized spectra contain NA, NaN, or infinite values.")
}

cat("\nPASS: Spectra are now min-max normalized across 19–50 kHz.\n")

# ------------------------------------------------------------
# Save diagnostics
# ------------------------------------------------------------

minmax_diagnostics <- spectra_metadata %>%
  mutate(
    raw_min_19_50_khz = spectra_row_min,
    raw_max_19_50_khz = spectra_row_max,
    raw_range_19_50_khz = spectra_row_range,
    minmax_min_19_50_khz = minmax_row_min,
    minmax_max_19_50_khz = minmax_row_max
  )

write.csv(
  minmax_diagnostics,
  file = file.path(base_output_dir, "minmax_normalization_diagnostics_19_50_khz.csv"),
  row.names = FALSE
)

```

## Join ICI to Spectra Metadata

```{r join_ici_to_spectra_metadata}

# ------------------------------------------------------------
# Prepare ICI table for joining
# ------------------------------------------------------------

ici_for_join <- click_metadata_with_ici %>%
  mutate(
    UID = as.character(UID)
  ) %>%
  select(
    EventID,
    UID,
    DriftID,
    Final_UTC_clean,
    Final_UTC_ms,
    ICI_to_next_ms,
    ICI_to_next_flag,
    ICI_from_previous_ms,
    ICI_from_previous_flag,
    IsFirstClickInEvent,
    IsLastClickInEvent
  )

# ------------------------------------------------------------
# Check ICI table uniqueness before join
# ------------------------------------------------------------

ici_duplicate_check <- ici_for_join %>%
  count(EventID, UID, name = "n_rows") %>%
  filter(n_rows > 1)

cat("Duplicated EventID + UID rows in ICI table:\n")
print(ici_duplicate_check)

if (nrow(ici_duplicate_check) > 0) {
  stop("FAIL: ICI table has duplicated EventID + UID rows.")
}

# ------------------------------------------------------------
# Join ICI table to spectra metadata
# ------------------------------------------------------------

spectra_metadata_with_ici <- spectra_metadata %>%
  left_join(
    ici_for_join,
    by = c("EventID", "UID"),
    suffix = c("_spectra", "_ici")
  )

cat("Rows before join:", nrow(spectra_metadata), "\n")
cat("Rows after join:", nrow(spectra_metadata_with_ici), "\n")

if (nrow(spectra_metadata_with_ici) != nrow(spectra_metadata)) {
  stop("FAIL: ICI join changed the number of rows.")
}

# ------------------------------------------------------------
# Create ICI in seconds
# ------------------------------------------------------------

spectra_metadata_with_ici <- spectra_metadata_with_ici %>%
  mutate(
    ICI_to_next_s = ICI_to_next_ms / 1000,
    log10_ICI_to_next_s = log10(ICI_to_next_s)
  )

# ------------------------------------------------------------
# Join coverage checks
# ------------------------------------------------------------

cat("\nRows missing ICI_to_next_ms:\n")
print(sum(is.na(spectra_metadata_with_ici$ICI_to_next_ms)))

cat("\nICI_to_next_flag counts:\n")
print(table(spectra_metadata_with_ici$ICI_to_next_flag, useNA = "ifany"))

cat("\nSummary of ICI_to_next_s:\n")
print(summary(spectra_metadata_with_ici$ICI_to_next_s))

# ------------------------------------------------------------
# Save joined metadata to base folder for now
# ------------------------------------------------------------

saveRDS(
  spectra_metadata_with_ici,
  file = file.path(base_output_dir, "spectra_metadata_with_ici_pre_pca.rds")
)

write.csv(
  spectra_metadata_with_ici,
  file = file.path(base_output_dir, "spectra_metadata_with_ici_pre_pca.csv"),
  row.names = FALSE
)

cat("\nPASS: ICI metadata joined to spectra metadata.\n")

```

## Inspect ICI Filtering Effect

```{r inspect_ici_filtering_effect}

# ------------------------------------------------------------
# Label clicks by ICI filter category
# ------------------------------------------------------------

spectra_metadata_with_ici <- spectra_metadata_with_ici %>%
  mutate(
    ICI_filter_category = case_when(
      is.na(ICI_to_next_s) ~ "missing_ici",
      ICI_to_next_s == 0 ~ "zero_ici",
      ICI_to_next_s < ici_min_s ~ "below_0.01_s_likely_buzz",
      ICI_to_next_s >= ici_min_s & ICI_to_next_s <= ici_max_s ~ "kept_0.01_to_0.4_s",
      ICI_to_next_s > ici_max_s ~ "above_0.4_s_long_gap",
      TRUE ~ "other"
    )
  )

# ------------------------------------------------------------
# Overall filter summary
# ------------------------------------------------------------

ici_filter_summary <- spectra_metadata_with_ici %>%
  count(ICI_filter_category, name = "n_clicks") %>%
  mutate(
    percent_clicks = 100 * n_clicks / sum(n_clicks)
  ) %>%
  arrange(desc(n_clicks))

cat("Overall ICI filter summary:\n")
print(ici_filter_summary)

write.csv(
  ici_filter_summary,
  file = file.path(base_output_dir, "ici_filter_summary_overall.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# Filter summary by event
# ------------------------------------------------------------

ici_filter_summary_by_event <- spectra_metadata_with_ici %>%
  group_by(DriftID_spectra, EventID, ICI_filter_category) %>%
  summarise(
    n_clicks = n(),
    .groups = "drop"
  ) %>%
  group_by(DriftID_spectra, EventID) %>%
  mutate(
    percent_within_event = 100 * n_clicks / sum(n_clicks)
  ) %>%
  ungroup() %>%
  arrange(DriftID_spectra, EventID, ICI_filter_category)

cat("\nICI filter summary by event:\n")
print(ici_filter_summary_by_event, n = Inf)

write.csv(
  ici_filter_summary_by_event,
  file = file.path(base_output_dir, "ici_filter_summary_by_event.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# Kept-click summary by event
# ------------------------------------------------------------

kept_summary_by_event <- spectra_metadata_with_ici %>%
  group_by(DriftID_spectra, EventID) %>%
  summarise(
    n_total_clicks = n(),
    n_kept_0.01_to_0.4_s = sum(ICI_filter_category == "kept_0.01_to_0.4_s"),
    percent_kept = 100 * n_kept_0.01_to_0.4_s / n_total_clicks,
    .groups = "drop"
  ) %>%
  arrange(DriftID_spectra, EventID)

cat("\nKept-click summary by event:\n")
print(kept_summary_by_event, n = Inf)

write.csv(
  kept_summary_by_event,
  file = file.path(base_output_dir, "kept_click_summary_by_event.csv"),
  row.names = FALSE
)

```

## Filter to 0.01-0.4 s and Sample Up To 200 Clicks Per Event

```{r filter_and_sample_clicks}

set.seed(sampling_seed)

eligible_clicks <- spectra_metadata_with_ici %>%
  mutate(
    RowID = row_number()
  ) %>%
  filter(
    ICI_filter_category == "kept_0.01_to_0.4_s"
  )

cat("Eligible clicks after ICI filtering:", nrow(eligible_clicks), "\n")

sampled_clicks <- eligible_clicks %>%
  group_by(EventID) %>%
  group_modify(~ {
    slice_sample(
      .x,
      n = min(nrow(.x), clicks_per_event),
      replace = FALSE
    )
  }) %>%
  ungroup()

cat("Sampled clicks:", nrow(sampled_clicks), "\n")
cat("Number of events represented:", n_distinct(sampled_clicks$EventID), "\n")

sampled_row_ids <- sampled_clicks$RowID

spectra_matrix_sampled_minmax <- spectra_matrix_minmax[
  sampled_row_ids,
  ,
  drop = FALSE
]

cat("Sampled spectra matrix dimensions:\n")
print(dim(spectra_matrix_sampled_minmax))

if (nrow(spectra_matrix_sampled_minmax) != nrow(sampled_clicks)) {
  stop("FAIL: Sampled spectra matrix rows do not match sampled metadata rows.")
}

sample_summary_by_event <- sampled_clicks %>%
  count(DriftID_spectra, EventID, name = "n_sampled_clicks") %>%
  arrange(DriftID_spectra, EventID)

cat("\nSampled clicks by event:\n")
print(sample_summary_by_event, n = Inf)

write.csv(
  sample_summary_by_event,
  file = file.path(base_output_dir, "sampled_clicks_by_event_pre_pca.csv"),
  row.names = FALSE
)

saveRDS(
  sampled_clicks,
  file = file.path(base_output_dir, "sampled_click_metadata_ici_filtered_pre_pca.rds")
)

saveRDS(
  spectra_matrix_sampled_minmax,
  file = file.path(base_output_dir, "sampled_minmax_spectra_matrix_19_50_khz_pre_pca.rds")
)

cat("\nPASS: ICI filtering and post-filter sampling complete.\n")

```

## Run PCA on Sampled Min-Max Spectra

```{r run_pca_90_percent_variance}

# ------------------------------------------------------------
# PCA on min-max normalized spectra only
#
# rows = sampled clicks
# columns = frequency bins from 19–50 kHz
# ------------------------------------------------------------

pca_spectra <- prcomp(
  spectra_matrix_sampled_minmax,
  center = TRUE,
  scale. = FALSE
)

# ------------------------------------------------------------
# Variance explained
# ------------------------------------------------------------

pca_variance <- pca_spectra$sdev^2
pca_variance_explained <- pca_variance / sum(pca_variance)
pca_cumulative_variance <- cumsum(pca_variance_explained)

pca_variance_table <- data.frame(
  PC = seq_along(pca_variance_explained),
  PC_label = paste0("PC", seq_along(pca_variance_explained)),
  variance_explained = pca_variance_explained,
  cumulative_variance_explained = pca_cumulative_variance
)

# ------------------------------------------------------------
# PCs needed for target variance
# ------------------------------------------------------------

n_pcs_target <- which(pca_cumulative_variance >= target_variance)[1]

pc_folder_name <- paste0(n_pcs_target, "PC")

output_dir <- file.path(base_output_dir, pc_folder_name)
plot_dir <- file.path(output_dir, "plots")
data_dir <- file.path(output_dir, "data")
table_dir <- file.path(output_dir, "tables")

dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(data_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(table_dir, recursive = TRUE, showWarnings = FALSE)

cat("PCA complete.\n")
cat("Number of PCs available:", ncol(pca_spectra$x), "\n")
cat("Target variance:", target_variance, "\n")
cat("Number of PCs needed:", n_pcs_target, "\n")
cat("Output directory:\n")
cat(output_dir, "\n\n")

cat("First 20 PCs variance explained:\n")
print(head(pca_variance_table, 20))

# ------------------------------------------------------------
# Save PCA objects and variance table
# ------------------------------------------------------------

saveRDS(
  pca_spectra,
  file = file.path(data_dir, paste0("pca_spectra_sampled_minmax_19_50_khz_", pc_folder_name, ".rds"))
)

write.csv(
  pca_variance_table,
  file = file.path(table_dir, paste0("pca_variance_explained_", pc_folder_name, ".csv")),
  row.names = FALSE
)

# ------------------------------------------------------------
# Move/save key pre-PCA objects into the final PC-specific folder
# ------------------------------------------------------------

saveRDS(
  sampled_clicks,
  file = file.path(data_dir, paste0("sampled_click_metadata_ici_filtered_", pc_folder_name, ".rds"))
)

saveRDS(
  spectra_matrix_sampled_minmax,
  file = file.path(data_dir, paste0("sampled_minmax_spectra_matrix_19_50_khz_", pc_folder_name, ".rds"))
)

write.csv(
  sample_summary_by_event,
  file = file.path(table_dir, paste0("sampled_clicks_by_event_", pc_folder_name, ".csv")),
  row.names = FALSE
)

write.csv(
  freq_focus_table,
  file = file.path(table_dir, "frequency_grid_19_50_khz.csv"),
  row.names = FALSE
)

cat("\nPASS: PCA completed and PC-specific output folders created.\n")

```

## Plot PCA Variance Explained

```{r plot_pca_variance_explained}

# ------------------------------------------------------------
# Scree plot: variance explained by each PC
# ------------------------------------------------------------

p_scree <- ggplot(
  pca_variance_table,
  aes(x = PC, y = variance_explained)
) +
  geom_col() +
  geom_vline(xintercept = n_pcs_target, linetype = "dotted") +
  labs(
    title = "PCA Scree Plot",
    subtitle = paste0(
      n_pcs_target,
      " PCs reach ",
      target_variance * 100,
      "% cumulative spectral variance"
    ),
    x = "Principal Component",
    y = "Proportion of variance explained"
  ) +
  theme_bw(base_size = 13)

print(p_scree)

ggsave(
  filename = file.path(plot_dir, paste0("pca_scree_plot_", pc_folder_name, ".png")),
  plot = p_scree,
  width = 8,
  height = 5,
  dpi = 300
)

# ------------------------------------------------------------
# Cumulative variance plot
# ------------------------------------------------------------

p_cumulative <- ggplot(
  pca_variance_table,
  aes(x = PC, y = cumulative_variance_explained)
) +
  geom_line() +
  geom_point(size = 1) +
  geom_hline(yintercept = target_variance, linetype = "dotted") +
  geom_vline(xintercept = n_pcs_target, linetype = "dotted") +
  labs(
    title = "PCA Cumulative Variance Explained",
    subtitle = paste0(
      n_pcs_target,
      " PCs reach ",
      target_variance * 100,
      "% cumulative spectral variance"
    ),
    x = "Number of principal components",
    y = "Cumulative variance explained"
  ) +
  theme_bw(base_size = 13)

print(p_cumulative)

ggsave(
  filename = file.path(plot_dir, paste0("pca_cumulative_variance_plot_", pc_folder_name, ".png")),
  plot = p_cumulative,
  width = 8,
  height = 5,
  dpi = 300
)

cat("Saved PCA variance plots to:\n")
cat(plot_dir, "\n")

```

## Build 18PC + Scaled logICI Feature Matrices

```{r build_feature_matrices}

# ------------------------------------------------------------
# Add PCA scores to sampled click metadata
# ------------------------------------------------------------

pc_scores <- as.data.frame(pca_spectra$x)

sampled_clicks_with_pcs <- bind_cols(
  sampled_clicks,
  pc_scores
)

# ------------------------------------------------------------
# Prepare log-transformed and scaled ICI
# ------------------------------------------------------------

sampled_clicks_with_pcs <- sampled_clicks_with_pcs %>%
  mutate(
    log10_ICI_to_next_s = log10(ICI_to_next_s),
    scaled_log10_ICI = as.numeric(scale(log10_ICI_to_next_s))
  )

cat("Summary of raw ICI seconds:\n")
print(summary(sampled_clicks_with_pcs$ICI_to_next_s))

cat("\nSummary of log10 ICI seconds:\n")
print(summary(sampled_clicks_with_pcs$log10_ICI_to_next_s))

cat("\nSummary of scaled log10 ICI:\n")
print(summary(sampled_clicks_with_pcs$scaled_log10_ICI))

# ------------------------------------------------------------
# Define 18-PC set
# ------------------------------------------------------------

pc_set_target <- paste0("PC", 1:n_pcs_target)

# ------------------------------------------------------------
# Build spectra-only matrix
# ------------------------------------------------------------

feature_matrix_spectra_only <- sampled_clicks_with_pcs %>%
  select(all_of(pc_set_target)) %>%
  as.matrix()

# ------------------------------------------------------------
# Build scaled-ICI feature matrices
# ------------------------------------------------------------

feature_matrix_list <- list()

for (ici_scale in ici_scaling_values) {
  
  scaled_ici_column <- sampled_clicks_with_pcs$scaled_log10_ICI * ici_scale
  
  feature_matrix <- sampled_clicks_with_pcs %>%
    select(all_of(pc_set_target)) %>%
    mutate(
      scaled_log10_ICI = scaled_ici_column
    ) %>%
    as.matrix()
  
  matrix_name <- paste0(
    pc_folder_name,
    "_scaledICI_",
    ici_scale
  )
  
  feature_matrix_list[[matrix_name]] <- feature_matrix
  
  cat("\nBuilt feature matrix:", matrix_name, "\n")
  cat("Dimensions:", dim(feature_matrix), "\n")
  
  saveRDS(
    feature_matrix,
    file = file.path(
      data_dir,
      paste0("feature_matrix_", matrix_name, ".rds")
    )
  )
}

# ------------------------------------------------------------
# Save spectra-only matrix
# ------------------------------------------------------------

saveRDS(
  feature_matrix_spectra_only,
  file = file.path(
    data_dir,
    paste0("feature_matrix_", pc_folder_name, "_spectra_only.rds")
  )
)

# ------------------------------------------------------------
# Save metadata with PCs
# ------------------------------------------------------------

saveRDS(
  sampled_clicks_with_pcs,
  file = file.path(
    data_dir,
    paste0("sampled_clicks_with_", pc_folder_name, "_and_scaled_logici.rds")
  )
)

cat("\nPASS: All feature matrices created and saved.\n")

```

## Run UMAP on 18PC Feature Spaces

```{r run_umap_analyses}

set.seed(umap_seed)

# ------------------------------------------------------------
# Helper function to run UMAP
# ------------------------------------------------------------

run_umap_analysis <- function(feature_matrix, analysis_name) {
  
  cat("\n============================\n")
  cat("Running UMAP:", analysis_name, "\n")
  cat("============================\n")
  
  umap_result <- uwot::umap(
    feature_matrix,
    n_neighbors = umap_neighbors,
    min_dist = umap_min_dist,
    metric = umap_metric,
    n_components = 2,
    ret_model = FALSE,
    verbose = TRUE,
    seed = umap_seed
  )
  
  umap_df <- as.data.frame(umap_result)
  
  names(umap_df) <- c("UMAP1", "UMAP2")
  
  umap_df$Analysis <- analysis_name
  
  return(umap_df)
}

# ------------------------------------------------------------
# Run spectra-only UMAP
# ------------------------------------------------------------

umap_results_list <- list()

spectra_only_name <- paste0(pc_folder_name, "_spectra_only")

umap_results_list[[spectra_only_name]] <- bind_cols(
  sampled_clicks_with_pcs,
  run_umap_analysis(
    feature_matrix_spectra_only,
    spectra_only_name
  )
)

# ------------------------------------------------------------
# Run scaled-ICI UMAPs
# ------------------------------------------------------------

for (matrix_name in names(feature_matrix_list)) {
  
  umap_results_list[[matrix_name]] <- bind_cols(
    sampled_clicks_with_pcs,
    run_umap_analysis(
      feature_matrix_list[[matrix_name]],
      matrix_name
    )
  )
}

# ------------------------------------------------------------
# Combine all UMAP results
# ------------------------------------------------------------

umap_results_all <- bind_rows(umap_results_list)

cat("\nRows in combined UMAP results:\n")
print(nrow(umap_results_all))

cat("\nUnique analyses:\n")
print(unique(umap_results_all$Analysis))

# ------------------------------------------------------------
# Save outputs
# ------------------------------------------------------------

saveRDS(
  umap_results_all,
  file = file.path(
    data_dir,
    paste0("umap_results_all_", pc_folder_name, ".rds")
  )
)

write.csv(
  umap_results_all,
  file = file.path(
    table_dir,
    paste0("umap_results_all_", pc_folder_name, ".csv")
  ),
  row.names = FALSE
)

cat("\nPASS: All UMAP analyses complete.\n")

```

## Plot UMAPs Colored by log10 ICI

```{r plot_umaps_by_logici}

plot_umap_logici <- function(df, analysis_name) {
  
  plot_data <- df %>%
    filter(Analysis == analysis_name)
  
  p <- ggplot(
    plot_data,
    aes(
      x = UMAP1,
      y = UMAP2,
      color = log10_ICI_to_next_s
    )
  ) +
    geom_point(
      size = 0.45,
      alpha = 0.55
    ) +
    scale_color_viridis_c(
      option = "plasma"
    ) +
    labs(
      title = paste0("UMAP: ", analysis_name),
      subtitle = "Colored by log10 ICI to next click",
      x = "UMAP1",
      y = "UMAP2",
      color = "log10 ICI"
    ) +
    theme_bw(base_size = 13)
  
  return(p)
}

for (analysis_name in unique(umap_results_all$Analysis)) {
  
  p <- plot_umap_logici(
    df = umap_results_all,
    analysis_name = analysis_name
  )
  
  print(p)
  
  ggsave(
    filename = file.path(
      plot_dir,
      paste0("umap_", analysis_name, "_colored_by_log10_ici.png")
    ),
    plot = p,
    width = 7,
    height = 6,
    dpi = 300
  )
}

cat("Saved UMAP log10 ICI plots to:\n")
cat(plot_dir, "\n")

```

## Create Per-ICI-Weight Output Folders

```{r create_per_weight_output_folders}

# ------------------------------------------------------------
# Create output folders for each 18PC scaled-ICI analysis
# ------------------------------------------------------------

analysis_names <- names(feature_matrix_list)

analysis_output_dirs <- list()

for (analysis_name in analysis_names) {
  
  analysis_dir <- file.path(output_dir, analysis_name)
  
  analysis_output_dirs[[analysis_name]] <- list(
    root = analysis_dir,
    plots = file.path(analysis_dir, "plots"),
    tables = file.path(analysis_dir, "tables"),
    data = file.path(analysis_dir, "data")
  )
  
  dir.create(analysis_output_dirs[[analysis_name]]$root,
             recursive = TRUE,
             showWarnings = FALSE)
  
  dir.create(analysis_output_dirs[[analysis_name]]$plots,
             recursive = TRUE,
             showWarnings = FALSE)
  
  dir.create(analysis_output_dirs[[analysis_name]]$tables,
             recursive = TRUE,
             showWarnings = FALSE)
  
  dir.create(analysis_output_dirs[[analysis_name]]$data,
             recursive = TRUE,
             showWarnings = FALSE)
}

cat("Created per-analysis output folders:\n")
print(names(analysis_output_dirs))

cat("\nFolder locations:\n")
print(lapply(analysis_output_dirs, function(x) x$root))

```

## Run HDBSCAN Across All ICI Weights and minPts Values

```{r run_hdbscan_all_conditions}

# ------------------------------------------------------------
# Helper function
# ------------------------------------------------------------

run_hdbscan_analysis <- function(df, analysis_name, minPts_value) {
  
  cat("\n=====================================\n")
  cat("Running HDBSCAN\n")
  cat("Analysis:", analysis_name, "\n")
  cat("minPts:", minPts_value, "\n")
  cat("=====================================\n")
  
  umap_subset <- df %>%
    filter(Analysis == analysis_name)
  
  umap_matrix <- umap_subset %>%
    select(UMAP1, UMAP2) %>%
    as.matrix()
  
  hdbscan_result <- dbscan::hdbscan(
    umap_matrix,
    minPts = minPts_value
  )
  
  output_df <- umap_subset %>%
    mutate(
      HDBSCAN_cluster = factor(hdbscan_result$cluster),
      HDBSCAN_membership_probability = hdbscan_result$membership_prob,
      HDBSCAN_minPts = minPts_value
    )
  
  cluster_summary <- output_df %>%
    count(HDBSCAN_cluster, name = "n_clicks") %>%
    mutate(
      percent_clicks = 100 * n_clicks / sum(n_clicks),
      cluster_type = if_else(
        HDBSCAN_cluster == "0",
        "noise_unassigned",
        "cluster"
      ),
      Analysis = analysis_name,
      minPts = minPts_value
    ) %>%
    arrange(desc(n_clicks))
  
  return(list(
    result_df = output_df,
    cluster_summary = cluster_summary
  ))
}

# ------------------------------------------------------------
# Run all analyses
# ------------------------------------------------------------

hdbscan_results_list <- list()
hdbscan_summary_list <- list()

for (analysis_name in names(feature_matrix_list)) {
  
  for (minPts_value in hdbscan_minPts_values) {
    
    result <- run_hdbscan_analysis(
      df = umap_results_all,
      analysis_name = analysis_name,
      minPts_value = minPts_value
    )
    
    result_name <- paste0(
      analysis_name,
      "_minPts",
      minPts_value
    )
    
    hdbscan_results_list[[result_name]] <- result$result_df
    
    hdbscan_summary_list[[result_name]] <- result$cluster_summary
    
    # ------------------------------------------------------------
    # Save per-analysis outputs
    # ------------------------------------------------------------
    
    saveRDS(
      result$result_df,
      file = file.path(
        analysis_output_dirs[[analysis_name]]$data,
        paste0("hdbscan_results_", result_name, ".rds")
      )
    )
    
    write.csv(
      result$result_df,
      file = file.path(
        analysis_output_dirs[[analysis_name]]$tables,
        paste0("hdbscan_results_", result_name, ".csv")
      ),
      row.names = FALSE
    )
    
    write.csv(
      result$cluster_summary,
      file = file.path(
        analysis_output_dirs[[analysis_name]]$tables,
        paste0("hdbscan_cluster_summary_", result_name, ".csv")
      ),
      row.names = FALSE
    )
    
    cat("\nCluster summary:\n")
    print(result$cluster_summary)
  }
}

# ------------------------------------------------------------
# Combine all summaries
# ------------------------------------------------------------

hdbscan_summary_all <- bind_rows(hdbscan_summary_list)

write.csv(
  hdbscan_summary_all,
  file = file.path(
    output_dir,
    "hdbscan_summary_all_conditions.csv"
  ),
  row.names = FALSE
)

cat("\nPASS: All HDBSCAN analyses completed.\n")


```

## Select Primary HDBSCAN Result and Plot Clusters

```{r plot_primary_hdbscan_clusters}

# ------------------------------------------------------------
# Primary candidate based on HDBSCAN summary review
# ------------------------------------------------------------

primary_analysis_name <- "18PC_scaledICI_1"
primary_minPts <- 50

primary_result_name <- paste0(
  primary_analysis_name,
  "_minPts",
  primary_minPts
)

# ------------------------------------------------------------
# Load/use primary HDBSCAN result
# ------------------------------------------------------------

primary_hdbscan <- hdbscan_results_list[[primary_result_name]]

if (is.null(primary_hdbscan)) {
  stop("Could not find primary HDBSCAN result: ", primary_result_name)
}

primary_hdbscan <- primary_hdbscan %>%
  mutate(
    HDBSCAN_cluster = factor(as.character(HDBSCAN_cluster))
  )

# ------------------------------------------------------------
# Make sure color map includes all observed clusters
# ------------------------------------------------------------

observed_clusters <- levels(primary_hdbscan$HDBSCAN_cluster)

missing_colors <- setdiff(observed_clusters, names(cluster_color_map))

if (length(missing_colors) > 0) {
  extra_colors <- hue_pal()(length(missing_colors))
  names(extra_colors) <- missing_colors
  
  cluster_color_map_extended <- c(cluster_color_map, extra_colors)
} else {
  cluster_color_map_extended <- cluster_color_map
}

# Force noise to stay light grey
cluster_color_map_extended["0"] <- "grey80"

# ------------------------------------------------------------
# UMAP with noise included
# ------------------------------------------------------------

p_primary_umap_with_noise <- ggplot(
  primary_hdbscan,
  aes(
    x = UMAP1,
    y = UMAP2,
    color = HDBSCAN_cluster
  )
) +
  geom_point(size = 0.55, alpha = 0.7) +
  scale_color_manual(
    values = cluster_color_map_extended,
    drop = FALSE
  ) +
  labs(
    title = paste0("HDBSCAN Clusters: ", primary_analysis_name),
    subtitle = paste0("minPts = ", primary_minPts, "; cluster 0 = noise/unassigned"),
    x = "UMAP1",
    y = "UMAP2",
    color = "HDBSCAN\ncluster"
  ) +
  theme_bw(base_size = 13)

print(p_primary_umap_with_noise)

ggsave(
  filename = file.path(
    analysis_output_dirs[[primary_analysis_name]]$plots,
    paste0("umap_hdbscan_", primary_result_name, "_with_noise.png")
  ),
  plot = p_primary_umap_with_noise,
  width = 8,
  height = 6,
  dpi = 300
)

# ------------------------------------------------------------
# UMAP with noise removed
# ------------------------------------------------------------

primary_hdbscan_no_noise <- primary_hdbscan %>%
  filter(HDBSCAN_cluster != "0")

p_primary_umap_no_noise <- ggplot(
  primary_hdbscan_no_noise,
  aes(
    x = UMAP1,
    y = UMAP2,
    color = HDBSCAN_cluster
  )
) +
  geom_point(size = 0.65, alpha = 0.8) +
  scale_color_manual(
    values = cluster_color_map_extended,
    drop = FALSE
  ) +
  labs(
    title = paste0("HDBSCAN Clusters: ", primary_analysis_name),
    subtitle = paste0("minPts = ", primary_minPts, "; noise removed"),
    x = "UMAP1",
    y = "UMAP2",
    color = "HDBSCAN\ncluster"
  ) +
  theme_bw(base_size = 13)

print(p_primary_umap_no_noise)

ggsave(
  filename = file.path(
    analysis_output_dirs[[primary_analysis_name]]$plots,
    paste0("umap_hdbscan_", primary_result_name, "_no_noise.png")
  ),
  plot = p_primary_umap_no_noise,
  width = 8,
  height = 6,
  dpi = 300
)

# ------------------------------------------------------------
# Save primary result table
# ------------------------------------------------------------

write.csv(
  primary_hdbscan,
  file = file.path(
    analysis_output_dirs[[primary_analysis_name]]$tables,
    paste0("PRIMARY_hdbscan_assignments_", primary_result_name, ".csv")
  ),
  row.names = FALSE
)

saveRDS(
  primary_hdbscan,
  file = file.path(
    analysis_output_dirs[[primary_analysis_name]]$data,
    paste0("PRIMARY_hdbscan_assignments_", primary_result_name, ".rds")
  )
)

cat("\nPASS: Primary HDBSCAN UMAP plots saved with and without noise.\n")

```

## Clean Mean Spectra Outputs for Primary HDBSCAN Result

```{r clean_primary_cluster_spectra_outputs}

# ------------------------------------------------------------
# Clean rainbow cluster colors
# Noise cluster 0 is always light grey.
# Non-noise clusters follow rainbow order.
# ------------------------------------------------------------

observed_clusters <- sort(as.numeric(as.character(unique(primary_hdbscan$HDBSCAN_cluster))))
observed_clusters <- as.character(observed_clusters)

non_noise_clusters <- observed_clusters[observed_clusters != "0"]

rainbow_cluster_colors <- c(
  "1" = "#E41A1C",  # red
  "2" = "#FF7F00",  # orange
  "3" = "#FFD92F",  # yellow
  "4" = "#4DAF4A",  # green
  "5" = "#00BFC4",  # cyan
  "6" = "#377EB8",  # blue
  "7" = "#984EA3",  # purple
  "8" = "#F781BF",  # pink
  "9" = "#A65628",  # brown
  "10" = "#999999",
  "11" = "#66C2A5",
  "12" = "#FC8D62"
)

cluster_colors_current <- c(
  "0" = "grey80",
  rainbow_cluster_colors[non_noise_clusters]
)

# ------------------------------------------------------------
# Pull spectra using original RowID
# ------------------------------------------------------------

primary_hdbscan_clean <- primary_hdbscan %>%
  mutate(
    HDBSCAN_cluster = factor(
      as.character(HDBSCAN_cluster),
      levels = observed_clusters
    ),
    RowID = as.integer(RowID)
  )

primary_spectra_matrix <- spectra_matrix_minmax[
  primary_hdbscan_clean$RowID,
  ,
  drop = FALSE
]

if (nrow(primary_spectra_matrix) != nrow(primary_hdbscan_clean)) {
  stop("FAIL: Spectra rows do not match HDBSCAN metadata rows.")
}

# ------------------------------------------------------------
# Compute mean and SD spectra by cluster
# ------------------------------------------------------------

cluster_spectra_summary <- list()

for (cl in observed_clusters) {
  
  cluster_rows <- which(primary_hdbscan_clean$HDBSCAN_cluster == cl)
  
  cluster_matrix <- primary_spectra_matrix[
    cluster_rows,
    ,
    drop = FALSE
  ]
  
  cluster_spectra_summary[[cl]] <- data.frame(
    HDBSCAN_cluster = cl,
    Frequency_kHz = freq_focus,
    mean_minmax_spectrum = colMeans(cluster_matrix, na.rm = TRUE),
    sd_minmax_spectrum = apply(cluster_matrix, 2, sd, na.rm = TRUE),
    n_clicks = length(cluster_rows)
  )
}

cluster_spectra_summary <- bind_rows(cluster_spectra_summary) %>%
  mutate(
    HDBSCAN_cluster = factor(HDBSCAN_cluster, levels = observed_clusters),
    lower_sd = pmax(mean_minmax_spectrum - sd_minmax_spectrum, 0),
    upper_sd = pmin(mean_minmax_spectrum + sd_minmax_spectrum, 1)
  )

write.csv(
  cluster_spectra_summary,
  file = file.path(
    analysis_output_dirs[[primary_analysis_name]]$tables,
    paste0("clean_cluster_mean_spectra_", primary_result_name, ".csv")
  ),
  row.names = FALSE
)

## Cleaner Mean Spectra Outputs: No SD on Combined, Taller Stacked Plot

# ============================================================
# OUTPUT 1: Combined mean spectra, all clusters, NO SD shading
# ============================================================

p_mean_all_clusters <- ggplot(
  cluster_spectra_summary,
  aes(
    x = Frequency_kHz,
    y = mean_minmax_spectrum,
    color = HDBSCAN_cluster,
    group = HDBSCAN_cluster
  )
) +
  geom_line(linewidth = 1.15) +
  scale_color_manual(
    values = cluster_colors_current,
    drop = FALSE
  ) +
  labs(
    title = "Mean Min-Max Normalized Spectra by HDBSCAN Cluster",
    subtitle = paste0(
      pc_folder_name,
      " + 1x scaled logICI; minPts = ",
      primary_minPts,
      "; cluster 0 = unassigned"
    ),
    x = "Frequency (kHz)",
    y = "Mean min-max normalized amplitude",
    color = "HDBSCAN\ncluster"
  ) +
  theme_bw(base_size = 13)

print(p_mean_all_clusters)

ggsave(
  filename = file.path(
    analysis_output_dirs[[primary_analysis_name]]$plots,
    paste0("clean_mean_spectra_all_clusters_NO_SD_", primary_result_name, ".png")
  ),
  plot = p_mean_all_clusters,
  width = 10,
  height = 6,
  dpi = 300
)

# ============================================================
# OUTPUT 2: Faceted mean ± SD, noise excluded
# ============================================================

cluster_spectra_summary_no_noise <- cluster_spectra_summary %>%
  filter(HDBSCAN_cluster != "0")

p_mean_sd_faceted_no_noise <- ggplot(
  cluster_spectra_summary_no_noise,
  aes(
    x = Frequency_kHz,
    y = mean_minmax_spectrum,
    color = HDBSCAN_cluster,
    fill = HDBSCAN_cluster
  )
) +
  geom_ribbon(
    aes(
      ymin = lower_sd,
      ymax = upper_sd
    ),
    alpha = 0.22,
    color = NA
  ) +
  geom_line(linewidth = 1.1) +
  facet_wrap(~ HDBSCAN_cluster, ncol = 3) +
  scale_color_manual(
    values = cluster_colors_current,
    drop = FALSE
  ) +
  scale_fill_manual(
    values = cluster_colors_current,
    drop = FALSE
  ) +
  labs(
    title = "Mean Spectra with Variability by HDBSCAN Cluster",
    subtitle = paste0(
      "Mean ± 1 SD; ",
      pc_folder_name,
      " + 1x scaled logICI; minPts = ",
      primary_minPts,
      "; noise excluded"
    ),
    x = "Frequency (kHz)",
    y = "Min-max normalized amplitude"
  ) +
  theme_bw(base_size = 13) +
  theme(
    legend.position = "none",
    strip.text = element_text(face = "bold")
  )

print(p_mean_sd_faceted_no_noise)

ggsave(
  filename = file.path(
    analysis_output_dirs[[primary_analysis_name]]$plots,
    paste0("clean_mean_spectra_SD_faceted_no_noise_", primary_result_name, ".png")
  ),
  plot = p_mean_sd_faceted_no_noise,
  width = 12,
  height = 9,
  dpi = 300
)

# ============================================================
# OUTPUT 3: Vertically amplified stacked mean spectra
# Noise excluded
# ============================================================

spectral_amplification <- 2.8
cluster_spacing <- 1.15

stacked_cluster_spectra <- cluster_spectra_summary_no_noise %>%
  group_by(HDBSCAN_cluster) %>%
  mutate(
    HDBSCAN_cluster_numeric = as.numeric(as.character(HDBSCAN_cluster)),
    spectrum_centered = mean_minmax_spectrum - mean(mean_minmax_spectrum, na.rm = TRUE),
    offset = HDBSCAN_cluster_numeric * cluster_spacing,
    stacked_spectrum = offset + spectrum_centered * spectral_amplification
  ) %>%
  ungroup()

p_stacked_spectra_amplified <- ggplot(
  stacked_cluster_spectra,
  aes(
    x = Frequency_kHz,
    y = stacked_spectrum,
    color = HDBSCAN_cluster,
    group = HDBSCAN_cluster
  )
) +
  geom_line(linewidth = 1.4) +
  scale_color_manual(
    values = cluster_colors_current,
    drop = FALSE
  ) +
  scale_y_continuous(
    breaks = sort(unique(stacked_cluster_spectra$offset)),
    labels = sort(unique(stacked_cluster_spectra$HDBSCAN_cluster))
  ) +
  coord_cartesian(xlim = c(19, 50), clip = "off") +
  labs(
    title = "Stacked Mean Spectra by HDBSCAN Cluster",
    subtitle = paste0(
      "Vertically amplified spectra; ",
      pc_folder_name,
      " + 1x scaled logICI; minPts = ",
      primary_minPts,
      "; noise excluded"
    ),
    x = "Frequency (kHz)",
    y = "HDBSCAN cluster"
  ) +
  theme_classic(base_size = 15) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold"),
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    plot.margin = margin(10, 10, 10, 10)
  )

print(p_stacked_spectra_amplified)

ggsave(
  filename = file.path(
    analysis_output_dirs[[primary_analysis_name]]$plots,
    paste0("clean_stacked_mean_spectra_VERTICAL_AMPLIFIED_no_noise_", primary_result_name, ".png")
  ),
  plot = p_stacked_spectra_amplified,
  width = 3.8,
  height = 12,
  dpi = 300
)

```

## Generate Full Output Set for All ICI Weights

```{r full_outputs_all_ici_weights}

# ------------------------------------------------------------
# Settings
# ------------------------------------------------------------

output_minPts <- 50

# Use all scaled ICI analyses
output_analysis_names <- names(feature_matrix_list)

# ------------------------------------------------------------
# Helper: build clean rainbow color map for observed clusters
# ------------------------------------------------------------

make_cluster_colors <- function(cluster_ids) {
  
  cluster_ids <- sort(as.numeric(as.character(cluster_ids)))
  cluster_ids <- as.character(cluster_ids)
  
  non_noise <- cluster_ids[cluster_ids != "0"]
  
  rainbow_cluster_colors <- c(
    "1" = "#E41A1C",  # red
    "2" = "#FF7F00",  # orange
    "3" = "#FFD92F",  # yellow
    "4" = "#4DAF4A",  # green
    "5" = "#00BFC4",  # cyan
    "6" = "#377EB8",  # blue
    "7" = "#984EA3",  # purple
    "8" = "#F781BF",  # pink
    "9" = "#A65628",  # brown
    "10" = "#999999",
    "11" = "#66C2A5",
    "12" = "#FC8D62",
    "13" = "#8DA0CB",
    "14" = "#E78AC3",
    "15" = "#A6D854"
  )
  
  out <- c(
    "0" = "grey80",
    rainbow_cluster_colors[non_noise]
  )
  
  return(out)
}

# ------------------------------------------------------------
# Main loop
# ------------------------------------------------------------

for (analysis_name in output_analysis_names) {
  
  cat("\n====================================================\n")
  cat("Generating outputs for:", analysis_name, "\n")
  cat("minPts:", output_minPts, "\n")
  cat("====================================================\n")
  
  result_name <- paste0(
    analysis_name,
    "_minPts",
    output_minPts
  )
  
  hdbscan_df <- hdbscan_results_list[[result_name]]
  
  if (is.null(hdbscan_df)) {
    warning("Skipping missing result: ", result_name)
    next
  }
  
  hdbscan_df <- hdbscan_df %>%
    mutate(
      HDBSCAN_cluster = factor(as.character(HDBSCAN_cluster)),
      RowID = as.integer(RowID)
    )
  
  observed_clusters <- sort(as.numeric(as.character(unique(hdbscan_df$HDBSCAN_cluster))))
  observed_clusters <- as.character(observed_clusters)
  
  hdbscan_df <- hdbscan_df %>%
    mutate(
      HDBSCAN_cluster = factor(
        as.character(HDBSCAN_cluster),
        levels = observed_clusters
      )
    )
  
  cluster_colors_current <- make_cluster_colors(observed_clusters)
  
  analysis_plot_dir <- analysis_output_dirs[[analysis_name]]$plots
  analysis_table_dir <- analysis_output_dirs[[analysis_name]]$tables
  analysis_data_dir <- analysis_output_dirs[[analysis_name]]$data
  
  # ------------------------------------------------------------
  # Cluster summary
  # ------------------------------------------------------------
  
  cluster_summary <- hdbscan_df %>%
    count(HDBSCAN_cluster, name = "n_clicks") %>%
    mutate(
      percent_clicks = 100 * n_clicks / sum(n_clicks),
      cluster_type = if_else(
        HDBSCAN_cluster == "0",
        "unassigned_low_density",
        "cluster"
      ),
      Analysis = analysis_name,
      minPts = output_minPts
    ) %>%
    arrange(as.numeric(as.character(HDBSCAN_cluster)))
  
  write.csv(
    cluster_summary,
    file = file.path(
      analysis_table_dir,
      paste0("FULL_cluster_summary_", result_name, ".csv")
    ),
    row.names = FALSE
  )
  
  # ============================================================
  # 1. UMAP with noise
  # ============================================================
  
  p_umap_with_noise <- ggplot(
    hdbscan_df,
    aes(
      x = UMAP1,
      y = UMAP2,
      color = HDBSCAN_cluster
    )
  ) +
    geom_point(size = 0.55, alpha = 0.7) +
    scale_color_manual(values = cluster_colors_current, drop = FALSE) +
    labs(
      title = paste0("HDBSCAN Clusters: ", analysis_name),
      subtitle = paste0("minPts = ", output_minPts, "; cluster 0 = unassigned low-density"),
      x = "UMAP1",
      y = "UMAP2",
      color = "HDBSCAN\ncluster"
    ) +
    theme_bw(base_size = 13)
  
  print(p_umap_with_noise)
  
  ggsave(
    file.path(
      analysis_plot_dir,
      paste0("FULL_umap_clusters_with_noise_", result_name, ".png")
    ),
    p_umap_with_noise,
    width = 8,
    height = 6,
    dpi = 300
  )
  
  # ============================================================
  # 2. UMAP without noise
  # ============================================================
  
  hdbscan_no_noise <- hdbscan_df %>%
    filter(HDBSCAN_cluster != "0")
  
  p_umap_no_noise <- ggplot(
    hdbscan_no_noise,
    aes(
      x = UMAP1,
      y = UMAP2,
      color = HDBSCAN_cluster
    )
  ) +
    geom_point(size = 0.65, alpha = 0.8) +
    scale_color_manual(values = cluster_colors_current, drop = FALSE) +
    labs(
      title = paste0("HDBSCAN Clusters: ", analysis_name),
      subtitle = paste0("minPts = ", output_minPts, "; unassigned removed"),
      x = "UMAP1",
      y = "UMAP2",
      color = "HDBSCAN\ncluster"
    ) +
    theme_bw(base_size = 13)
  
  print(p_umap_no_noise)
  
  ggsave(
    file.path(
      analysis_plot_dir,
      paste0("FULL_umap_clusters_no_noise_", result_name, ".png")
    ),
    p_umap_no_noise,
    width = 8,
    height = 6,
    dpi = 300
  )
  
  # ============================================================
  # Spectra summary
  # ============================================================
  
  cluster_spectra_summary <- list()
  
  spectra_for_result <- spectra_matrix_minmax[
    hdbscan_df$RowID,
    ,
    drop = FALSE
  ]
  
  for (cl in observed_clusters) {
    
    cluster_rows <- which(hdbscan_df$HDBSCAN_cluster == cl)
    
    cluster_matrix <- spectra_for_result[
      cluster_rows,
      ,
      drop = FALSE
    ]
    
    cluster_spectra_summary[[cl]] <- data.frame(
      HDBSCAN_cluster = cl,
      Frequency_kHz = freq_focus,
      mean_minmax_spectrum = colMeans(cluster_matrix, na.rm = TRUE),
      sd_minmax_spectrum = apply(cluster_matrix, 2, sd, na.rm = TRUE),
      n_clicks = length(cluster_rows)
    )
  }
  
  cluster_spectra_summary <- bind_rows(cluster_spectra_summary) %>%
    mutate(
      HDBSCAN_cluster = factor(HDBSCAN_cluster, levels = observed_clusters),
      lower_sd = pmax(mean_minmax_spectrum - sd_minmax_spectrum, 0),
      upper_sd = pmin(mean_minmax_spectrum + sd_minmax_spectrum, 1)
    )
  
  write.csv(
    cluster_spectra_summary,
    file = file.path(
      analysis_table_dir,
      paste0("FULL_cluster_mean_spectra_", result_name, ".csv")
    ),
    row.names = FALSE
  )
  
  # ============================================================
  # 3. Mean spectra with noise
  # ============================================================
  
  p_mean_with_noise <- ggplot(
    cluster_spectra_summary,
    aes(
      x = Frequency_kHz,
      y = mean_minmax_spectrum,
      color = HDBSCAN_cluster,
      group = HDBSCAN_cluster
    )
  ) +
    geom_line(linewidth = 1.1) +
    scale_color_manual(values = cluster_colors_current, drop = FALSE) +
    labs(
      title = "Mean Min-Max Normalized Spectra by HDBSCAN Cluster",
      subtitle = paste0(analysis_name, "; minPts = ", output_minPts, "; cluster 0 = unassigned"),
      x = "Frequency (kHz)",
      y = "Mean min-max normalized amplitude",
      color = "HDBSCAN\ncluster"
    ) +
    theme_bw(base_size = 13)
  
  print(p_mean_with_noise)
  
  ggsave(
    file.path(
      analysis_plot_dir,
      paste0("FULL_mean_spectra_with_noise_", result_name, ".png")
    ),
    p_mean_with_noise,
    width = 10,
    height = 6,
    dpi = 300
  )
  
  # ============================================================
  # 4. Mean spectra without noise
  # ============================================================
  
  cluster_spectra_no_noise <- cluster_spectra_summary %>%
    filter(HDBSCAN_cluster != "0")
  
  p_mean_no_noise <- ggplot(
    cluster_spectra_no_noise,
    aes(
      x = Frequency_kHz,
      y = mean_minmax_spectrum,
      color = HDBSCAN_cluster,
      group = HDBSCAN_cluster
    )
  ) +
    geom_line(linewidth = 1.1) +
    scale_color_manual(values = cluster_colors_current, drop = FALSE) +
    labs(
      title = "Mean Min-Max Normalized Spectra by HDBSCAN Cluster",
      subtitle = paste0(analysis_name, "; minPts = ", output_minPts, "; unassigned removed"),
      x = "Frequency (kHz)",
      y = "Mean min-max normalized amplitude",
      color = "HDBSCAN\ncluster"
    ) +
    theme_bw(base_size = 13)
  
  print(p_mean_no_noise)
  
  ggsave(
    file.path(
      analysis_plot_dir,
      paste0("FULL_mean_spectra_no_noise_", result_name, ".png")
    ),
    p_mean_no_noise,
    width = 10,
    height = 6,
    dpi = 300
  )
  
  # ============================================================
  # 5. Faceted mean ± SD without noise
  # ============================================================
  
  p_sd_faceted <- ggplot(
    cluster_spectra_no_noise,
    aes(
      x = Frequency_kHz,
      y = mean_minmax_spectrum,
      color = HDBSCAN_cluster,
      fill = HDBSCAN_cluster
    )
  ) +
    geom_ribbon(
      aes(ymin = lower_sd, ymax = upper_sd),
      alpha = 0.22,
      color = NA
    ) +
    geom_line(linewidth = 1.1) +
    facet_wrap(~ HDBSCAN_cluster, ncol = 3) +
    scale_color_manual(values = cluster_colors_current, drop = FALSE) +
    scale_fill_manual(values = cluster_colors_current, drop = FALSE) +
    labs(
      title = "Mean Spectra with Variability by HDBSCAN Cluster",
      subtitle = paste0("Mean ± 1 SD; ", analysis_name, "; minPts = ", output_minPts, "; unassigned removed"),
      x = "Frequency (kHz)",
      y = "Min-max normalized amplitude"
    ) +
    theme_bw(base_size = 13) +
    theme(
      legend.position = "none",
      strip.text = element_text(face = "bold")
    )
  
  print(p_sd_faceted)
  
  ggsave(
    file.path(
      analysis_plot_dir,
      paste0("FULL_mean_spectra_SD_faceted_no_noise_", result_name, ".png")
    ),
    p_sd_faceted,
    width = 12,
    height = 9,
    dpi = 300
  )
  
  # ============================================================
  # 6. Vertically amplified stacked spectra without noise
  # ============================================================
  
  spectral_amplification <- 2.8
  cluster_spacing <- 1.15
  
  stacked_cluster_spectra <- cluster_spectra_no_noise %>%
    group_by(HDBSCAN_cluster) %>%
    mutate(
      HDBSCAN_cluster_numeric = as.numeric(as.character(HDBSCAN_cluster)),
      spectrum_centered = mean_minmax_spectrum - mean(mean_minmax_spectrum, na.rm = TRUE),
      offset = HDBSCAN_cluster_numeric * cluster_spacing,
      stacked_spectrum = offset + spectrum_centered * spectral_amplification
    ) %>%
    ungroup()
  
  p_stacked <- ggplot(
    stacked_cluster_spectra,
    aes(
      x = Frequency_kHz,
      y = stacked_spectrum,
      color = HDBSCAN_cluster,
      group = HDBSCAN_cluster
    )
  ) +
    geom_line(linewidth = 1.4) +
    scale_color_manual(values = cluster_colors_current, drop = FALSE) +
    scale_y_continuous(
      breaks = sort(unique(stacked_cluster_spectra$offset)),
      labels = sort(unique(stacked_cluster_spectra$HDBSCAN_cluster))
    ) +
    coord_cartesian(xlim = c(19, 50), clip = "off") +
    labs(
      title = "Stacked Mean Spectra by HDBSCAN Cluster",
      subtitle = paste0("Vertically amplified spectra; ", analysis_name, "; minPts = ", output_minPts),
      x = "Frequency (kHz)",
      y = "HDBSCAN cluster"
    ) +
    theme_classic(base_size = 15) +
    theme(
      legend.position = "none",
      plot.title = element_text(face = "bold"),
      axis.title.y = element_text(margin = margin(r = 10))
    )
  
  print(p_stacked)
  
  ggsave(
    file.path(
      analysis_plot_dir,
      paste0("FULL_stacked_mean_spectra_VERTICAL_AMPLIFIED_no_noise_", result_name, ".png")
    ),
    p_stacked,
    width = 3.8,
    height = 12,
    dpi = 300
  )
  
  # ============================================================
  # 7. ICI summaries
  # ============================================================
  
  ici_summary <- hdbscan_df %>%
    group_by(HDBSCAN_cluster) %>%
    summarise(
      n_clicks = n(),
      median_ici_s = median(ICI_to_next_s, na.rm = TRUE),
      mean_ici_s = mean(ICI_to_next_s, na.rm = TRUE),
      sd_ici_s = sd(ICI_to_next_s, na.rm = TRUE),
      q25_ici_s = quantile(ICI_to_next_s, 0.25, na.rm = TRUE),
      q75_ici_s = quantile(ICI_to_next_s, 0.75, na.rm = TRUE),
      min_ici_s = min(ICI_to_next_s, na.rm = TRUE),
      max_ici_s = max(ICI_to_next_s, na.rm = TRUE),
      n_events = n_distinct(EventID),
      n_drifts = n_distinct(DriftID_spectra),
      .groups = "drop"
    ) %>%
    arrange(as.numeric(as.character(HDBSCAN_cluster)))
  
  write.csv(
    ici_summary,
    file = file.path(
      analysis_table_dir,
      paste0("FULL_ici_summary_by_cluster_", result_name, ".csv")
    ),
    row.names = FALSE
  )
  
  # ============================================================
  # 8. ICI boxplot with noise
  # ============================================================
  
  p_ici_with_noise <- ggplot(
    hdbscan_df,
    aes(
      x = HDBSCAN_cluster,
      y = ICI_to_next_s,
      fill = HDBSCAN_cluster
    )
  ) +
    geom_boxplot(outlier.size = 0.25) +
    scale_y_log10() +
    scale_fill_manual(values = cluster_colors_current, drop = FALSE) +
    labs(
      title = "ICI Distribution by HDBSCAN Cluster",
      subtitle = paste0(analysis_name, "; minPts = ", output_minPts, "; cluster 0 included"),
      x = "HDBSCAN cluster",
      y = "ICI to next click (seconds, log scale)"
    ) +
    theme_bw(base_size = 13) +
    theme(legend.position = "none")
  
  print(p_ici_with_noise)
  
  ggsave(
    file.path(
      analysis_plot_dir,
      paste0("FULL_ici_boxplot_with_noise_", result_name, ".png")
    ),
    p_ici_with_noise,
    width = 9,
    height = 6,
    dpi = 300
  )
  
  # ============================================================
  # 9. ICI boxplot without noise
  # ============================================================
  
  p_ici_no_noise <- ggplot(
    hdbscan_no_noise,
    aes(
      x = HDBSCAN_cluster,
      y = ICI_to_next_s,
      fill = HDBSCAN_cluster
    )
  ) +
    geom_boxplot(outlier.size = 0.25) +
    scale_y_log10() +
    scale_fill_manual(values = cluster_colors_current, drop = FALSE) +
    labs(
      title = "ICI Distribution by HDBSCAN Cluster",
      subtitle = paste0(analysis_name, "; minPts = ", output_minPts, "; unassigned removed"),
      x = "HDBSCAN cluster",
      y = "ICI to next click (seconds, log scale)"
    ) +
    theme_bw(base_size = 13) +
    theme(legend.position = "none")
  
  print(p_ici_no_noise)
  
  ggsave(
    file.path(
      analysis_plot_dir,
      paste0("FULL_ici_boxplot_no_noise_", result_name, ".png")
    ),
    p_ici_no_noise,
    width = 9,
    height = 6,
    dpi = 300
  )
  
  # ------------------------------------------------------------
  # Save full assignments again with clean naming
  # ------------------------------------------------------------
  
  write.csv(
    hdbscan_df,
    file = file.path(
      analysis_table_dir,
      paste0("FULL_hdbscan_assignments_", result_name, ".csv")
    ),
    row.names = FALSE
  )
  
  saveRDS(
    hdbscan_df,
    file = file.path(
      analysis_data_dir,
      paste0("FULL_hdbscan_assignments_", result_name, ".rds")
    )
  )
  
  cat("Finished:", analysis_name, "\n")
}

cat("\nPASS: Full output generation complete for all ICI weights.\n")

```

## Load Event Coordinates and Join to Primary HDBSCAN Result

```{r load_coordinates_join_primary_hdbscan}

# ------------------------------------------------------------
# Input: event latitude-longitude table
# This is the same location file used in the old mapping workflow.
# ------------------------------------------------------------

adrift_detection_file <- "E:/Logs/ADRIFT_GgDetections.csv"

if (!file.exists(adrift_detection_file)) {
  stop("Could not find coordinate file: ", adrift_detection_file)
}

adrift_detections <- readr::read_csv(
  adrift_detection_file,
  show_col_types = FALSE
)

cat("Columns in ADRIFT_GgDetections:\n")
print(names(adrift_detections))

# ------------------------------------------------------------
# Inspect coordinate file
# ------------------------------------------------------------

cat("\nRows in coordinate file:", nrow(adrift_detections), "\n")

cat("\nCoordinate file preview:\n")
print(
  adrift_detections %>%
    select(DriftName, EventID, Latitude, Longitude) %>%
    head(20)
)

# ------------------------------------------------------------
# Create event-level click counts from current primary HDBSCAN result
# ------------------------------------------------------------

hdbscan_event_click_counts <- mapping_hdbscan %>%
  mutate(
    EventID = as.character(EventID),
    EventID_clean = stringr::str_replace(EventID, "_Event_", "_"),
    DriftID_clean = stringr::str_remove(EventID_clean, "_[0-9]+$"),
    EventPrefix = stringr::str_extract(EventID, "^[^_]+")
  ) %>%
  count(
    EventPrefix,
    DriftID_clean,
    EventID,
    EventID_clean,
    name = "n_clicks_used"
  ) %>%
  arrange(EventPrefix, DriftID_clean, EventID_clean)

cat("\nTotal clicks in primary HDBSCAN object:", sum(hdbscan_event_click_counts$n_clicks_used), "\n")
cat("Unique events used:", nrow(hdbscan_event_click_counts), "\n")
cat("Unique drifts used:", n_distinct(hdbscan_event_click_counts$DriftID_clean), "\n")

# ------------------------------------------------------------
# Clean coordinate table
# ------------------------------------------------------------

event_locations <- adrift_detections %>%
  mutate(
    EventID_clean = as.character(EventID),
    Latitude = as.numeric(Latitude),
    Longitude = as.numeric(Longitude)
  ) %>%
  filter(
    !is.na(EventID_clean),
    !is.na(Latitude),
    !is.na(Longitude)
  ) %>%
  select(
    EventID_clean,
    DriftName,
    Latitude,
    Longitude
  ) %>%
  distinct()

# ------------------------------------------------------------
# Join event coordinates to HDBSCAN event table
# ------------------------------------------------------------

hdbscan_events_with_locations <- hdbscan_event_click_counts %>%
  left_join(
    event_locations,
    by = "EventID_clean"
  )

cat("\nHDBSCAN events:", nrow(hdbscan_event_click_counts), "\n")
cat("Matched location rows:", sum(!is.na(hdbscan_events_with_locations$Latitude)), "\n")
cat("Missing location rows:", sum(is.na(hdbscan_events_with_locations$Latitude)), "\n\n")

cat("Missing events, if any:\n")
print(
  hdbscan_events_with_locations %>%
    filter(is.na(Latitude) | is.na(Longitude))
)

cat("\nMatched event coordinate preview:\n")
print(hdbscan_events_with_locations, n = Inf)

# ------------------------------------------------------------
# Join coordinates back to click-level HDBSCAN result
# ------------------------------------------------------------

mapping_hdbscan_with_locations <- mapping_hdbscan %>%
  mutate(
    EventID = as.character(EventID),
    EventID_clean = stringr::str_replace(EventID, "_Event_", "_"),
    DriftID_clean = stringr::str_remove(EventID_clean, "_[0-9]+$")
  ) %>%
  left_join(
    hdbscan_events_with_locations %>%
      select(
        EventID,
        EventID_clean,
        DriftID_clean,
        DriftName,
        Latitude,
        Longitude
      ),
    by = c("EventID", "EventID_clean", "DriftID_clean")
  )

cat("\nRows in mapping_hdbscan_with_locations:", nrow(mapping_hdbscan_with_locations), "\n")
cat("Rows missing Latitude:", sum(is.na(mapping_hdbscan_with_locations$Latitude)), "\n")
cat("Rows missing Longitude:", sum(is.na(mapping_hdbscan_with_locations$Longitude)), "\n")

if (any(is.na(mapping_hdbscan_with_locations$Latitude)) ||
    any(is.na(mapping_hdbscan_with_locations$Longitude))) {
  stop("Some HDBSCAN rows still lack coordinates. Check missing events above.")
}

## Mapping Output Folder Setup

mapping_analysis_name <- "18PC_scaledICI_1"
mapping_minPts <- 50

mapping_result_name <- paste0(
  mapping_analysis_name,
  "_minPts",
  mapping_minPts
)

mapping_base_dir <- file.path(
  output_dir,
  mapping_analysis_name,
  "mapping_outputs"
)

event_map_dir <- file.path(mapping_base_dir, "by_event")
drift_map_dir <- file.path(mapping_base_dir, "by_drift")

event_map_plot_dir <- file.path(event_map_dir, "plots")
event_map_table_dir <- file.path(event_map_dir, "tables")

drift_map_plot_dir <- file.path(drift_map_dir, "plots")
drift_map_table_dir <- file.path(drift_map_dir, "tables")

dir.create(event_map_plot_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(event_map_table_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(drift_map_plot_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(drift_map_table_dir, recursive = TRUE, showWarnings = FALSE)

cat("Event mapping plots:\n", event_map_plot_dir, "\n\n")
cat("Event mapping tables:\n", event_map_table_dir, "\n\n")
cat("Drift mapping plots:\n", drift_map_plot_dir, "\n\n")
cat("Drift mapping tables:\n", drift_map_table_dir, "\n")

# ------------------------------------------------------------
# Save joined coordinate tables
# ------------------------------------------------------------

write.csv(
  hdbscan_events_with_locations,
  file.path(event_map_table_dir, "event_locations_matched_to_hdbscan.csv"),
  row.names = FALSE
)

write.csv(
  mapping_hdbscan_with_locations,
  file.path(event_map_table_dir, "click_level_hdbscan_with_event_locations.csv"),
  row.names = FALSE
)

cat("\nPASS: Coordinates joined to primary HDBSCAN result.\n")

```

## Build Event-Level Cluster Composition Tables

```{r build_event_cluster_composition_tables}

# ------------------------------------------------------------
# Event-level cluster counts WITH noise/unassigned
# ------------------------------------------------------------

event_cluster_counts_with_noise <- mapping_hdbscan_with_locations %>%
  count(
    EventID_clean,
    DriftID_clean,
    DriftName,
    Latitude,
    Longitude,
    HDBSCAN_cluster,
    name = "n_clicks"
  ) %>%
  group_by(EventID_clean) %>%
  mutate(
    total_clicks_event = sum(n_clicks),
    proportion = n_clicks / total_clicks_event,
    percent = 100 * proportion
  ) %>%
  ungroup() %>%
  arrange(EventID_clean, as.numeric(as.character(HDBSCAN_cluster)))

# ------------------------------------------------------------
# Event-level cluster counts WITHOUT noise/unassigned
# ------------------------------------------------------------

event_cluster_counts_no_noise <- mapping_hdbscan_with_locations %>%
  filter(HDBSCAN_cluster != "0") %>%
  count(
    EventID_clean,
    DriftID_clean,
    DriftName,
    Latitude,
    Longitude,
    HDBSCAN_cluster,
    name = "n_clicks"
  ) %>%
  group_by(EventID_clean) %>%
  mutate(
    total_clicks_event_nonnoise = sum(n_clicks),
    proportion = n_clicks / total_clicks_event_nonnoise,
    percent = 100 * proportion
  ) %>%
  ungroup() %>%
  arrange(EventID_clean, as.numeric(as.character(HDBSCAN_cluster)))

# ------------------------------------------------------------
# Event summary table
# ------------------------------------------------------------

event_cluster_summary <- mapping_hdbscan_with_locations %>%
  group_by(
    EventID_clean,
    DriftID_clean,
    DriftName,
    Latitude,
    Longitude
  ) %>%
  summarise(
    n_clicks_total = n(),
    n_clicks_unassigned = sum(HDBSCAN_cluster == "0"),
    n_clicks_assigned = sum(HDBSCAN_cluster != "0"),
    percent_unassigned = 100 * n_clicks_unassigned / n_clicks_total,
    percent_assigned = 100 * n_clicks_assigned / n_clicks_total,
    n_clusters_present_with_noise = n_distinct(HDBSCAN_cluster),
    n_clusters_present_no_noise = n_distinct(HDBSCAN_cluster[HDBSCAN_cluster != "0"]),
    .groups = "drop"
  ) %>%
  arrange(DriftID_clean, EventID_clean)

# ------------------------------------------------------------
# Save tables
# ------------------------------------------------------------

write.csv(
  event_cluster_counts_with_noise,
  file.path(
    event_map_table_dir,
    paste0("event_cluster_composition_WITH_noise_", mapping_result_name, ".csv")
  ),
  row.names = FALSE
)

write.csv(
  event_cluster_counts_no_noise,
  file.path(
    event_map_table_dir,
    paste0("event_cluster_composition_NO_noise_", mapping_result_name, ".csv")
  ),
  row.names = FALSE
)

write.csv(
  event_cluster_summary,
  file.path(
    event_map_table_dir,
    paste0("event_cluster_summary_", mapping_result_name, ".csv")
  ),
  row.names = FALSE
)

cat("Event cluster composition WITH noise:\n")
print(head(event_cluster_counts_with_noise, 20))

cat("\nEvent cluster summary:\n")
print(event_cluster_summary, n = Inf)

cat("\nPASS: Event-level cluster composition tables created.\n")

```

## Event Pie Maps Without scatterpie

```{r event_pie_maps_no_scatterpie}

# ------------------------------------------------------------
# Helper: create pie slice polygons
# ------------------------------------------------------------

make_pie_polygons <- function(df, group_col, lon_col, lat_col, cluster_col, prop_col, radius = 0.18) {
  
  df <- df %>%
    arrange(.data[[group_col]], as.numeric(as.character(.data[[cluster_col]]))) %>%
    group_by(.data[[group_col]]) %>%
    mutate(
      ymin = lag(cumsum(.data[[prop_col]]), default = 0),
      ymax = cumsum(.data[[prop_col]])
    ) %>%
    ungroup()
  
  pie_list <- list()
  counter <- 1
  
  for (i in seq_len(nrow(df))) {
    
    row <- df[i, ]
    
    angles <- seq(
      2 * pi * row$ymin,
      2 * pi * row$ymax,
      length.out = 40
    )
    
    polygon_df <- data.frame(
      x = c(
        row[[lon_col]],
        row[[lon_col]] + radius * cos(angles)
      ),
      y = c(
        row[[lat_col]],
        row[[lat_col]] + radius * sin(angles)
      ),
      HDBSCAN_cluster = as.character(row[[cluster_col]]),
      PieGroup = as.character(row[[group_col]]),
      stringsAsFactors = FALSE
    )
    
    pie_list[[counter]] <- polygon_df
    counter <- counter + 1
  }
  
  bind_rows(pie_list)
}

# ------------------------------------------------------------
# Consistent colors
# ------------------------------------------------------------

event_clusters_all <- sort(
  unique(as.character(mapping_hdbscan_with_locations$HDBSCAN_cluster))
)

cluster_colors_current <- make_cluster_colors(event_clusters_all)

# ------------------------------------------------------------
# World basemap
# ------------------------------------------------------------

world <- rnaturalearth::ne_countries(
  scale = "medium",
  returnclass = "sf"
)

# ------------------------------------------------------------
# Event pie polygons WITH noise
# ------------------------------------------------------------

event_pie_polygons_with_noise <- make_pie_polygons(
  df = event_cluster_counts_with_noise,
  group_col = "EventID_clean",
  lon_col = "Longitude",
  lat_col = "Latitude",
  cluster_col = "HDBSCAN_cluster",
  prop_col = "proportion",
  radius = 0.18
)

# ------------------------------------------------------------
# Event pie map WITH noise
# ------------------------------------------------------------

p_event_pies_with_noise <- ggplot() +
  geom_sf(
    data = world,
    fill = "grey95",
    color = "grey70",
    linewidth = 0.2
  ) +
  geom_polygon(
    data = event_pie_polygons_with_noise,
    aes(
      x = x,
      y = y,
      group = interaction(PieGroup, HDBSCAN_cluster),
      fill = HDBSCAN_cluster
    ),
    color = "black",
    linewidth = 0.15,
    alpha = 0.95
  ) +
  coord_sf(
    xlim = c(
      min(mapping_hdbscan_with_locations$Longitude) - 2,
      max(mapping_hdbscan_with_locations$Longitude) + 2
    ),
    ylim = c(
      min(mapping_hdbscan_with_locations$Latitude) - 2,
      max(mapping_hdbscan_with_locations$Latitude) + 2
    )
  ) +
  scale_fill_manual(
    values = cluster_colors_current,
    drop = FALSE
  ) +
  labs(
    title = "Event-Level HDBSCAN Cluster Composition",
    subtitle = paste0(
      mapping_analysis_name,
      "; minPts = ",
      mapping_minPts,
      "; including unassigned cluster 0"
    ),
    fill = "HDBSCAN cluster"
  ) +
  theme_bw(base_size = 13)

print(p_event_pies_with_noise)

ggsave(
  file.path(
    event_map_plot_dir,
    paste0("event_pie_map_WITH_noise_", mapping_result_name, ".png")
  ),
  p_event_pies_with_noise,
  width = 12,
  height = 8,
  dpi = 300
)

# ------------------------------------------------------------
# Event pie polygons WITHOUT noise
# ------------------------------------------------------------

event_pie_polygons_no_noise <- make_pie_polygons(
  df = event_cluster_counts_no_noise,
  group_col = "EventID_clean",
  lon_col = "Longitude",
  lat_col = "Latitude",
  cluster_col = "HDBSCAN_cluster",
  prop_col = "proportion",
  radius = 0.18
)

# ------------------------------------------------------------
# Event pie map WITHOUT noise
# ------------------------------------------------------------

p_event_pies_no_noise <- ggplot() +
  geom_sf(
    data = world,
    fill = "grey95",
    color = "grey70",
    linewidth = 0.2
  ) +
  geom_polygon(
    data = event_pie_polygons_no_noise,
    aes(
      x = x,
      y = y,
      group = interaction(PieGroup, HDBSCAN_cluster),
      fill = HDBSCAN_cluster
    ),
    color = "black",
    linewidth = 0.15,
    alpha = 0.95
  ) +
  coord_sf(
    xlim = c(
      min(mapping_hdbscan_with_locations$Longitude) - 2,
      max(mapping_hdbscan_with_locations$Longitude) + 2
    ),
    ylim = c(
      min(mapping_hdbscan_with_locations$Latitude) - 2,
      max(mapping_hdbscan_with_locations$Latitude) + 2
    )
  ) +
  scale_fill_manual(
    values = cluster_colors_current,
    drop = FALSE
  ) +
  labs(
    title = "Event-Level HDBSCAN Cluster Composition",
    subtitle = paste0(
      mapping_analysis_name,
      "; minPts = ",
      mapping_minPts,
      "; unassigned cluster removed"
    ),
    fill = "HDBSCAN cluster"
  ) +
  theme_bw(base_size = 13)

print(p_event_pies_no_noise)

ggsave(
  file.path(
    event_map_plot_dir,
    paste0("event_pie_map_NO_noise_", mapping_result_name, ".png")
  ),
  p_event_pies_no_noise,
  width = 12,
  height = 8,
  dpi = 300
)

cat("\nPASS: Event pie maps generated without scatterpie.\n")

```

## Event Pie Maps Without scatterpie — Bigger Pies

```{r event_pie_maps_no_scatterpie_bigger}

event_pie_radius <- 0.45

# ------------------------------------------------------------
# Helper: create pie slice polygons
# ------------------------------------------------------------

make_pie_polygons <- function(df, group_col, lon_col, lat_col, cluster_col, prop_col, radius = 0.45) {
  
  df <- df %>%
    arrange(.data[[group_col]], as.numeric(as.character(.data[[cluster_col]]))) %>%
    group_by(.data[[group_col]]) %>%
    mutate(
      ymin = lag(cumsum(.data[[prop_col]]), default = 0),
      ymax = cumsum(.data[[prop_col]])
    ) %>%
    ungroup()
  
  pie_list <- list()
  counter <- 1
  
  for (i in seq_len(nrow(df))) {
    
    row <- df[i, ]
    
    angles <- seq(
      2 * pi * row$ymin,
      2 * pi * row$ymax,
      length.out = 60
    )
    
    polygon_df <- data.frame(
      x = c(row[[lon_col]], row[[lon_col]] + radius * cos(angles)),
      y = c(row[[lat_col]], row[[lat_col]] + radius * sin(angles)),
      HDBSCAN_cluster = as.character(row[[cluster_col]]),
      PieGroup = as.character(row[[group_col]]),
      stringsAsFactors = FALSE
    )
    
    pie_list[[counter]] <- polygon_df
    counter <- counter + 1
  }
  
  bind_rows(pie_list)
}

event_pie_polygons_with_noise <- make_pie_polygons(
  df = event_cluster_counts_with_noise,
  group_col = "EventID_clean",
  lon_col = "Longitude",
  lat_col = "Latitude",
  cluster_col = "HDBSCAN_cluster",
  prop_col = "proportion",
  radius = event_pie_radius
)

event_pie_polygons_no_noise <- make_pie_polygons(
  df = event_cluster_counts_no_noise,
  group_col = "EventID_clean",
  lon_col = "Longitude",
  lat_col = "Latitude",
  cluster_col = "HDBSCAN_cluster",
  prop_col = "proportion",
  radius = event_pie_radius
)

p_event_pies_with_noise <- ggplot() +
  geom_sf(data = world, fill = "grey95", color = "grey70", linewidth = 0.2) +
  geom_polygon(
    data = event_pie_polygons_with_noise,
    aes(x = x, y = y, group = interaction(PieGroup, HDBSCAN_cluster), fill = HDBSCAN_cluster),
    color = "black",
    linewidth = 0.15,
    alpha = 0.95
  ) +
  coord_sf(
    xlim = c(min(mapping_hdbscan_with_locations$Longitude) - 2, max(mapping_hdbscan_with_locations$Longitude) + 2),
    ylim = c(min(mapping_hdbscan_with_locations$Latitude) - 2, max(mapping_hdbscan_with_locations$Latitude) + 2)
  ) +
  scale_fill_manual(values = cluster_colors_current, drop = FALSE) +
  labs(
    title = "Event-Level HDBSCAN Cluster Composition",
    subtitle = paste0(mapping_analysis_name, "; minPts = ", mapping_minPts, "; including unassigned cluster 0"),
    fill = "HDBSCAN cluster"
  ) +
  theme_bw(base_size = 13)

print(p_event_pies_with_noise)

ggsave(
  file.path(event_map_plot_dir, paste0("event_pie_map_WITH_noise_BIGGER_", mapping_result_name, ".png")),
  p_event_pies_with_noise,
  width = 12,
  height = 8,
  dpi = 300
)

p_event_pies_no_noise <- ggplot() +
  geom_sf(data = world, fill = "grey95", color = "grey70", linewidth = 0.2) +
  geom_polygon(
    data = event_pie_polygons_no_noise,
    aes(x = x, y = y, group = interaction(PieGroup, HDBSCAN_cluster), fill = HDBSCAN_cluster),
    color = "black",
    linewidth = 0.15,
    alpha = 0.95
  ) +
  coord_sf(
    xlim = c(min(mapping_hdbscan_with_locations$Longitude) - 2, max(mapping_hdbscan_with_locations$Longitude) + 2),
    ylim = c(min(mapping_hdbscan_with_locations$Latitude) - 2, max(mapping_hdbscan_with_locations$Latitude) + 2)
  ) +
  scale_fill_manual(values = cluster_colors_current, drop = FALSE) +
  labs(
    title = "Event-Level HDBSCAN Cluster Composition",
    subtitle = paste0(mapping_analysis_name, "; minPts = ", mapping_minPts, "; unassigned cluster removed"),
    fill = "HDBSCAN cluster"
  ) +
  theme_bw(base_size = 13)

print(p_event_pies_no_noise)

ggsave(
  file.path(event_map_plot_dir, paste0("event_pie_map_NO_noise_BIGGER_", mapping_result_name, ".png")),
  p_event_pies_no_noise,
  width = 12,
  height = 8,
  dpi = 300
)

cat("\nPASS: Bigger event pie maps generated.\n")

```

## Drift-Level Big Pie Maps

```{r drift_level_big_pie_maps}

# ------------------------------------------------------------
# Big drift pie radius
# ------------------------------------------------------------

drift_pie_radius <- 0.75

# ------------------------------------------------------------
# Drift-level cluster counts WITH noise
# ------------------------------------------------------------

drift_cluster_counts_with_noise <- mapping_hdbscan_with_locations %>%
  count(
    DriftID_clean,
    DriftName,
    HDBSCAN_cluster,
    name = "n_clicks"
  ) %>%
  group_by(DriftID_clean) %>%
  mutate(
    total_clicks_drift = sum(n_clicks),
    proportion = n_clicks / total_clicks_drift,
    percent = 100 * proportion
  ) %>%
  ungroup() %>%
  arrange(DriftID_clean, as.numeric(as.character(HDBSCAN_cluster)))

# ------------------------------------------------------------
# Drift-level cluster counts WITHOUT noise
# ------------------------------------------------------------

drift_cluster_counts_no_noise <- mapping_hdbscan_with_locations %>%
  filter(HDBSCAN_cluster != "0") %>%
  count(
    DriftID_clean,
    DriftName,
    HDBSCAN_cluster,
    name = "n_clicks"
  ) %>%
  group_by(DriftID_clean) %>%
  mutate(
    total_clicks_drift_nonnoise = sum(n_clicks),
    proportion = n_clicks / total_clicks_drift_nonnoise,
    percent = 100 * proportion
  ) %>%
  ungroup() %>%
  arrange(DriftID_clean, as.numeric(as.character(HDBSCAN_cluster)))

# ------------------------------------------------------------
# Drift-level locations
# Mean of event locations within each drift
# ------------------------------------------------------------

drift_locations <- mapping_hdbscan_with_locations %>%
  distinct(
    DriftID_clean,
    DriftName,
    EventID_clean,
    Latitude,
    Longitude
  ) %>%
  group_by(DriftID_clean, DriftName) %>%
  summarise(
    Latitude = mean(Latitude, na.rm = TRUE),
    Longitude = mean(Longitude, na.rm = TRUE),
    n_events_in_drift = n_distinct(EventID_clean),
    .groups = "drop"
  )

# ------------------------------------------------------------
# Join drift locations to drift cluster tables
# ------------------------------------------------------------

drift_cluster_counts_with_noise <- drift_cluster_counts_with_noise %>%
  left_join(
    drift_locations,
    by = c("DriftID_clean", "DriftName")
  )

drift_cluster_counts_no_noise <- drift_cluster_counts_no_noise %>%
  left_join(
    drift_locations,
    by = c("DriftID_clean", "DriftName")
  )

# ------------------------------------------------------------
# Save drift composition tables
# ------------------------------------------------------------

write.csv(
  drift_cluster_counts_with_noise,
  file.path(
    drift_map_table_dir,
    paste0("drift_cluster_composition_WITH_noise_", mapping_result_name, ".csv")
  ),
  row.names = FALSE
)

write.csv(
  drift_cluster_counts_no_noise,
  file.path(
    drift_map_table_dir,
    paste0("drift_cluster_composition_NO_noise_", mapping_result_name, ".csv")
  ),
  row.names = FALSE
)

write.csv(
  drift_locations,
  file.path(
    drift_map_table_dir,
    paste0("drift_mean_locations_", mapping_result_name, ".csv")
  ),
  row.names = FALSE
)

# ------------------------------------------------------------
# Consistent colors
# ------------------------------------------------------------

drift_clusters_all <- sort(
  unique(as.character(mapping_hdbscan_with_locations$HDBSCAN_cluster))
)

cluster_colors_current <- make_cluster_colors(drift_clusters_all)

# ------------------------------------------------------------
# Drift pie polygons WITH noise
# ------------------------------------------------------------

drift_pie_polygons_with_noise <- make_pie_polygons(
  df = drift_cluster_counts_with_noise,
  group_col = "DriftID_clean",
  lon_col = "Longitude",
  lat_col = "Latitude",
  cluster_col = "HDBSCAN_cluster",
  prop_col = "proportion",
  radius = drift_pie_radius
)

# ------------------------------------------------------------
# Drift pie map WITH noise
# ------------------------------------------------------------

p_drift_pies_with_noise <- ggplot() +
  geom_sf(
    data = world,
    fill = "grey95",
    color = "grey70",
    linewidth = 0.2
  ) +
  geom_polygon(
    data = drift_pie_polygons_with_noise,
    aes(
      x = x,
      y = y,
      group = interaction(PieGroup, HDBSCAN_cluster),
      fill = HDBSCAN_cluster
    ),
    color = "black",
    linewidth = 0.18,
    alpha = 0.95
  ) +
  coord_sf(
    xlim = c(
      min(drift_locations$Longitude) - 2,
      max(drift_locations$Longitude) + 2
    ),
    ylim = c(
      min(drift_locations$Latitude) - 2,
      max(drift_locations$Latitude) + 2
    )
  ) +
  scale_fill_manual(
    values = cluster_colors_current,
    drop = FALSE
  ) +
  labs(
    title = "Drift-Level HDBSCAN Cluster Composition",
    subtitle = paste0(
      mapping_analysis_name,
      "; minPts = ",
      mapping_minPts,
      "; including unassigned cluster 0"
    ),
    fill = "HDBSCAN cluster"
  ) +
  theme_bw(base_size = 13)

print(p_drift_pies_with_noise)

ggsave(
  file.path(
    drift_map_plot_dir,
    paste0("drift_pie_map_WITH_noise_BIG_", mapping_result_name, ".png")
  ),
  p_drift_pies_with_noise,
  width = 12,
  height = 8,
  dpi = 300
)

# ------------------------------------------------------------
# Drift pie polygons WITHOUT noise
# ------------------------------------------------------------

drift_pie_polygons_no_noise <- make_pie_polygons(
  df = drift_cluster_counts_no_noise,
  group_col = "DriftID_clean",
  lon_col = "Longitude",
  lat_col = "Latitude",
  cluster_col = "HDBSCAN_cluster",
  prop_col = "proportion",
  radius = drift_pie_radius
)

# ------------------------------------------------------------
# Drift pie map WITHOUT noise
# ------------------------------------------------------------

p_drift_pies_no_noise <- ggplot() +
  geom_sf(
    data = world,
    fill = "grey95",
    color = "grey70",
    linewidth = 0.2
  ) +
  geom_polygon(
    data = drift_pie_polygons_no_noise,
    aes(
      x = x,
      y = y,
      group = interaction(PieGroup, HDBSCAN_cluster),
      fill = HDBSCAN_cluster
    ),
    color = "black",
    linewidth = 0.18,
    alpha = 0.95
  ) +
  coord_sf(
    xlim = c(
      min(drift_locations$Longitude) - 2,
      max(drift_locations$Longitude) + 2
    ),
    ylim = c(
      min(drift_locations$Latitude) - 2,
      max(drift_locations$Latitude) + 2
    )
  ) +
  scale_fill_manual(
    values = cluster_colors_current,
    drop = FALSE
  ) +
  labs(
    title = "Drift-Level HDBSCAN Cluster Composition",
    subtitle = paste0(
      mapping_analysis_name,
      "; minPts = ",
      mapping_minPts,
      "; unassigned cluster removed"
    ),
    fill = "HDBSCAN cluster"
  ) +
  theme_bw(base_size = 13)

print(p_drift_pies_no_noise)

ggsave(
  file.path(
    drift_map_plot_dir,
    paste0("drift_pie_map_NO_noise_BIG_", mapping_result_name, ".png")
  ),
  p_drift_pies_no_noise,
  width = 12,
  height = 8,
  dpi = 300
)

cat("\nPASS: Big drift-level pie maps generated.\n")

```

## Drift-Level Pie Maps With Expanded Manual Coordinate Grouping

```{r drift_level_pie_maps_manual_coordinate_grouping_v2}

# ------------------------------------------------------------
# Pie size
# ------------------------------------------------------------

drift_pie_radius <- 0.45

# ------------------------------------------------------------
# Create manual display coordinates
# ------------------------------------------------------------

manual_display_locations <- drift_locations %>%
  mutate(
    DisplayGroup = case_when(
      # Around 45N, 124W
      Latitude >= 44.5 & Latitude <= 46.0 &
        Longitude >= -125.5 & Longitude <= -123.0 ~ "Group_45N_124W",

      # Around 41N, 124W
      Latitude >= 40.3 & Latitude <= 41.5 &
        Longitude >= -125.2 & Longitude <= -123.8 ~ "Group_41N_124W",

      # Around 37–38N, 122–123W / Point Reyes-ish overlap
      Latitude >= 37.5 & Latitude <= 38.5 &
        Longitude >= -124.0 & Longitude <= -122.5 ~ "Group_38N_123W",

      # Around 36N, 122W / Monterey-ish overlap
      Latitude >= 35.0 & Latitude <= 36.8 &
        Longitude >= -122.5 & Longitude <= -121.0 ~ "Group_36N_122W",

      TRUE ~ DriftID_clean
    )
  ) %>%
  group_by(DisplayGroup) %>%
  mutate(
    DisplayLatitude = mean(Latitude, na.rm = TRUE),
    DisplayLongitude = mean(Longitude, na.rm = TRUE)
  ) %>%
  ungroup()

cat("Manual display coordinate groups:\n")
print(
  manual_display_locations %>%
    arrange(DisplayGroup, DriftID_clean),
  n = Inf
)

write.csv(
  manual_display_locations,
  file.path(
    drift_map_table_dir,
    paste0("manual_display_locations_grouped_v2_", mapping_result_name, ".csv")
  ),
  row.names = FALSE
)

# ------------------------------------------------------------
# Join display coordinates onto drift cluster composition tables
# ------------------------------------------------------------

drift_cluster_counts_with_noise_display <- drift_cluster_counts_with_noise %>%
  select(-Latitude, -Longitude) %>%
  left_join(
    manual_display_locations %>%
      select(
        DriftID_clean,
        DisplayGroup,
        DisplayLatitude,
        DisplayLongitude
      ),
    by = "DriftID_clean"
  ) %>%
  mutate(
    Latitude = DisplayLatitude,
    Longitude = DisplayLongitude
  )

drift_cluster_counts_no_noise_display <- drift_cluster_counts_no_noise %>%
  select(-Latitude, -Longitude) %>%
  left_join(
    manual_display_locations %>%
      select(
        DriftID_clean,
        DisplayGroup,
        DisplayLatitude,
        DisplayLongitude
      ),
    by = "DriftID_clean"
  ) %>%
  mutate(
    Latitude = DisplayLatitude,
    Longitude = DisplayLongitude
  )

# ------------------------------------------------------------
# Collapse by DisplayGroup so nearby drifts become one pie
# ------------------------------------------------------------

drift_display_counts_with_noise <- drift_cluster_counts_with_noise_display %>%
  group_by(
    DisplayGroup,
    Latitude,
    Longitude,
    HDBSCAN_cluster
  ) %>%
  summarise(
    n_clicks = sum(n_clicks),
    .groups = "drop"
  ) %>%
  group_by(DisplayGroup) %>%
  mutate(
    total_clicks_display_group = sum(n_clicks),
    proportion = n_clicks / total_clicks_display_group,
    percent = 100 * proportion
  ) %>%
  ungroup()

drift_display_counts_no_noise <- drift_cluster_counts_no_noise_display %>%
  group_by(
    DisplayGroup,
    Latitude,
    Longitude,
    HDBSCAN_cluster
  ) %>%
  summarise(
    n_clicks = sum(n_clicks),
    .groups = "drop"
  ) %>%
  group_by(DisplayGroup) %>%
  mutate(
    total_clicks_display_group_nonnoise = sum(n_clicks),
    proportion = n_clicks / total_clicks_display_group_nonnoise,
    percent = 100 * proportion
  ) %>%
  ungroup()

write.csv(
  drift_display_counts_with_noise,
  file.path(
    drift_map_table_dir,
    paste0("drift_DISPLAY_GROUP_composition_WITH_noise_v2_", mapping_result_name, ".csv")
  ),
  row.names = FALSE
)

write.csv(
  drift_display_counts_no_noise,
  file.path(
    drift_map_table_dir,
    paste0("drift_DISPLAY_GROUP_composition_NO_noise_v2_", mapping_result_name, ".csv")
  ),
  row.names = FALSE
)

# ------------------------------------------------------------
# Build polygons WITH noise
# ------------------------------------------------------------

drift_display_pie_polygons_with_noise <- make_pie_polygons(
  df = drift_display_counts_with_noise,
  group_col = "DisplayGroup",
  lon_col = "Longitude",
  lat_col = "Latitude",
  cluster_col = "HDBSCAN_cluster",
  prop_col = "proportion",
  radius = drift_pie_radius
)

# ------------------------------------------------------------
# Map WITH noise
# ------------------------------------------------------------

p_drift_display_with_noise <- ggplot() +
  geom_sf(
    data = world,
    fill = "grey95",
    color = "grey70",
    linewidth = 0.2
  ) +
  geom_polygon(
    data = drift_display_pie_polygons_with_noise,
    aes(
      x = x,
      y = y,
      group = interaction(PieGroup, HDBSCAN_cluster),
      fill = HDBSCAN_cluster
    ),
    color = "black",
    linewidth = 0.18,
    alpha = 0.95
  ) +
  coord_sf(
    xlim = c(
      min(drift_display_counts_with_noise$Longitude) - 2,
      max(drift_display_counts_with_noise$Longitude) + 2
    ),
    ylim = c(
      min(drift_display_counts_with_noise$Latitude) - 2,
      max(drift_display_counts_with_noise$Latitude) + 2
    )
  ) +
  scale_fill_manual(
    values = cluster_colors_current,
    drop = FALSE
  ) +
  labs(
    title = "Drift-Level HDBSCAN Cluster Composition",
    subtitle = paste0(
      mapping_analysis_name,
      "; minPts = ",
      mapping_minPts,
      "; nearby drift coordinates grouped; including unassigned cluster 0"
    ),
    x = "Longitude (°W)",
    y = "Latitude (°N)",
    fill = "HDBSCAN cluster"
  ) +
  theme_bw(base_size = 13)

print(p_drift_display_with_noise)

ggsave(
  file.path(
    drift_map_plot_dir,
    paste0("drift_pie_map_WITH_noise_GROUPED_COORDS_radius0.45_v2_", mapping_result_name, ".png")
  ),
  p_drift_display_with_noise,
  width = 12,
  height = 8,
  dpi = 300
)

# ------------------------------------------------------------
# Build polygons WITHOUT noise
# ------------------------------------------------------------

drift_display_pie_polygons_no_noise <- make_pie_polygons(
  df = drift_display_counts_no_noise,
  group_col = "DisplayGroup",
  lon_col = "Longitude",
  lat_col = "Latitude",
  cluster_col = "HDBSCAN_cluster",
  prop_col = "proportion",
  radius = drift_pie_radius
)

# ------------------------------------------------------------
# Map WITHOUT noise
# ------------------------------------------------------------

p_drift_display_no_noise <- ggplot() +
  geom_sf(
    data = world,
    fill = "grey95",
    color = "grey70",
    linewidth = 0.2
  ) +
  geom_polygon(
    data = drift_display_pie_polygons_no_noise,
    aes(
      x = x,
      y = y,
      group = interaction(PieGroup, HDBSCAN_cluster),
      fill = HDBSCAN_cluster
    ),
    color = "black",
    linewidth = 0.18,
    alpha = 0.95
  ) +
  coord_sf(
    xlim = c(
      min(drift_display_counts_no_noise$Longitude) - 2,
      max(drift_display_counts_no_noise$Longitude) + 2
    ),
    ylim = c(
      min(drift_display_counts_no_noise$Latitude) - 2,
      max(drift_display_counts_no_noise$Latitude) + 2
    )
  ) +
  scale_fill_manual(
    values = cluster_colors_current,
    drop = FALSE
  ) +
  labs(
    title = "Drift-Level HDBSCAN Cluster Composition",
    subtitle = paste0(
      mapping_analysis_name,
      "; minPts = ",
      mapping_minPts,
      "; nearby drift coordinates grouped; unassigned cluster removed"
    ),
    x = "Longitude (°W)",
    y = "Latitude (°N)",
    fill = "HDBSCAN cluster"
  ) +
  theme_bw(base_size = 13)

print(p_drift_display_no_noise)

ggsave(
  file.path(
    drift_map_plot_dir,
    paste0("drift_pie_map_NO_noise_GROUPED_COORDS_radius0.45_v2_", mapping_result_name, ".png")
  ),
  p_drift_display_no_noise,
  width = 12,
  height = 8,
  dpi = 300
)

cat("\nPASS: Grouped-coordinate drift pie maps v2 generated.\n")

```

## Drift-Level Pie Maps With Expanded Manual Coordinate Grouping v3

```{r drift_level_pie_maps_manual_coordinate_grouping_v3}

# ------------------------------------------------------------
# Pie size
# ------------------------------------------------------------

drift_pie_radius <- 0.75

# ------------------------------------------------------------
# Create manual display coordinates
# ------------------------------------------------------------

manual_display_locations <- drift_locations %>%
  mutate(
    DisplayGroup = case_when(
      # Around 45N, 124W
      Latitude >= 44.5 & Latitude <= 46.0 &
        Longitude >= -125.5 & Longitude <= -123.0 ~ "Group_45N_124W",

      # Around 41N, 124W
      Latitude >= 40.3 & Latitude <= 41.5 &
        Longitude >= -125.2 & Longitude <= -123.8 ~ "Group_41N_124W",

      # Around 37–38N, 122–123W
      Latitude >= 37.5 & Latitude <= 38.5 &
        Longitude >= -124.0 & Longitude <= -122.5 ~ "Group_38N_123W",

      # Around 36N, 122W
      Latitude >= 35.0 & Latitude <= 36.8 &
        Longitude >= -122.5 & Longitude <= -121.0 ~ "Group_36N_122W",

      # Around 36N, 128W
      Latitude >= 35.5 & Latitude <= 37.0 &
        Longitude >= -129.0 & Longitude <= -127.0 ~ "Group_36N_128W",

      TRUE ~ DriftID_clean
    )
  ) %>%
  group_by(DisplayGroup) %>%
  mutate(
    DisplayLatitude = mean(Latitude, na.rm = TRUE),
    DisplayLongitude = mean(Longitude, na.rm = TRUE)
  ) %>%
  ungroup()

write.csv(
  manual_display_locations,
  file.path(
    drift_map_table_dir,
    paste0("manual_display_locations_grouped_v3_", mapping_result_name, ".csv")
  ),
  row.names = FALSE
)

print(manual_display_locations %>% arrange(DisplayGroup, DriftID_clean), n = Inf)

# ------------------------------------------------------------
# Join display coordinates
# ------------------------------------------------------------

drift_cluster_counts_with_noise_display <- drift_cluster_counts_with_noise %>%
  select(-Latitude, -Longitude) %>%
  left_join(
    manual_display_locations %>%
      select(DriftID_clean, DisplayGroup, DisplayLatitude, DisplayLongitude),
    by = "DriftID_clean"
  ) %>%
  mutate(
    Latitude = DisplayLatitude,
    Longitude = DisplayLongitude
  )

drift_cluster_counts_no_noise_display <- drift_cluster_counts_no_noise %>%
  select(-Latitude, -Longitude) %>%
  left_join(
    manual_display_locations %>%
      select(DriftID_clean, DisplayGroup, DisplayLatitude, DisplayLongitude),
    by = "DriftID_clean"
  ) %>%
  mutate(
    Latitude = DisplayLatitude,
    Longitude = DisplayLongitude
  )

# ------------------------------------------------------------
# Collapse by DisplayGroup
# ------------------------------------------------------------

drift_display_counts_with_noise <- drift_cluster_counts_with_noise_display %>%
  group_by(DisplayGroup, Latitude, Longitude, HDBSCAN_cluster) %>%
  summarise(n_clicks = sum(n_clicks), .groups = "drop") %>%
  group_by(DisplayGroup) %>%
  mutate(
    total_clicks_display_group = sum(n_clicks),
    proportion = n_clicks / total_clicks_display_group,
    percent = 100 * proportion
  ) %>%
  ungroup()

drift_display_counts_no_noise <- drift_cluster_counts_no_noise_display %>%
  group_by(DisplayGroup, Latitude, Longitude, HDBSCAN_cluster) %>%
  summarise(n_clicks = sum(n_clicks), .groups = "drop") %>%
  group_by(DisplayGroup) %>%
  mutate(
    total_clicks_display_group_nonnoise = sum(n_clicks),
    proportion = n_clicks / total_clicks_display_group_nonnoise,
    percent = 100 * proportion
  ) %>%
  ungroup()

write.csv(
  drift_display_counts_with_noise,
  file.path(
    drift_map_table_dir,
    paste0("drift_DISPLAY_GROUP_composition_WITH_noise_v3_", mapping_result_name, ".csv")
  ),
  row.names = FALSE
)

write.csv(
  drift_display_counts_no_noise,
  file.path(
    drift_map_table_dir,
    paste0("drift_DISPLAY_GROUP_composition_NO_noise_v3_", mapping_result_name, ".csv")
  ),
  row.names = FALSE
)

# ------------------------------------------------------------
# WITH noise map
# ------------------------------------------------------------

drift_display_pie_polygons_with_noise <- make_pie_polygons(
  df = drift_display_counts_with_noise,
  group_col = "DisplayGroup",
  lon_col = "Longitude",
  lat_col = "Latitude",
  cluster_col = "HDBSCAN_cluster",
  prop_col = "proportion",
  radius = drift_pie_radius
)

p_drift_display_with_noise <- ggplot() +
  geom_sf(
    data = world,
    fill = "grey95",
    color = "grey70",
    linewidth = 0.2
  ) +
  geom_polygon(
    data = drift_display_pie_polygons_with_noise,
    aes(
      x = x,
      y = y,
      group = interaction(PieGroup, HDBSCAN_cluster),
      fill = HDBSCAN_cluster
    ),
    color = "black",
    linewidth = 0.18,
    alpha = 0.95
  ) +
  coord_sf(
    xlim = c(
      min(drift_display_counts_with_noise$Longitude) - 2,
      max(drift_display_counts_with_noise$Longitude) + 2
    ),
    ylim = c(
      min(drift_display_counts_with_noise$Latitude) - 2,
      max(drift_display_counts_with_noise$Latitude) + 2
    )
  ) +
  scale_fill_manual(values = cluster_colors_current, drop = FALSE) +
  labs(
    title = "Drift-Level HDBSCAN Cluster Composition",
    subtitle = paste0(
      mapping_analysis_name,
      "; minPts = ",
      mapping_minPts,
      "; nearby drift coordinates grouped; including unassigned cluster 0"
    ),
    x = "Longitude (°W)",
    y = "Latitude (°N)",
    fill = "HDBSCAN cluster"
  ) +
  theme_bw(base_size = 13)

print(p_drift_display_with_noise)

ggsave(
  file.path(
    drift_map_plot_dir,
    paste0("drift_pie_map_WITH_noise_GROUPED_COORDS_radius0.75_v3_", mapping_result_name, ".png")
  ),
  p_drift_display_with_noise,
  width = 12,
  height = 8,
  dpi = 300
)

# ------------------------------------------------------------
# WITHOUT noise map
# ------------------------------------------------------------

drift_display_pie_polygons_no_noise <- make_pie_polygons(
  df = drift_display_counts_no_noise,
  group_col = "DisplayGroup",
  lon_col = "Longitude",
  lat_col = "Latitude",
  cluster_col = "HDBSCAN_cluster",
  prop_col = "proportion",
  radius = drift_pie_radius
)

p_drift_display_no_noise <- ggplot() +
  geom_sf(
    data = world,
    fill = "grey95",
    color = "grey70",
    linewidth = 0.2
  ) +
  geom_polygon(
    data = drift_display_pie_polygons_no_noise,
    aes(
      x = x,
      y = y,
      group = interaction(PieGroup, HDBSCAN_cluster),
      fill = HDBSCAN_cluster
    ),
    color = "black",
    linewidth = 0.18,
    alpha = 0.95
  ) +
  coord_sf(
    xlim = c(
      min(drift_display_counts_no_noise$Longitude) - 2,
      max(drift_display_counts_no_noise$Longitude) + 2
    ),
    ylim = c(
      min(drift_display_counts_no_noise$Latitude) - 2,
      max(drift_display_counts_no_noise$Latitude) + 2
    )
  ) +
  scale_fill_manual(values = cluster_colors_current, drop = FALSE) +
  labs(
    title = "Drift-Level HDBSCAN Cluster Composition",
    subtitle = paste0(
      mapping_analysis_name,
      "; minPts = ",
      mapping_minPts,
      "; nearby drift coordinates grouped; unassigned cluster removed"
    ),
    x = "Longitude (°W)",
    y = "Latitude (°N)",
    fill = "HDBSCAN cluster"
  ) +
  theme_bw(base_size = 13)

print(p_drift_display_no_noise)

ggsave(
  file.path(
    drift_map_plot_dir,
    paste0("drift_pie_map_NO_noise_GROUPED_COORDS_radius0.75_v3_", mapping_result_name, ".png")
  ),
  p_drift_display_no_noise,
  width = 12,
  height = 8,
  dpi = 300
)

cat("\nPASS: Grouped-coordinate drift pie maps v3 generated.\n")

```

## Save Drift and Cluster Percent Summary Tables

```{r save_drift_cluster_percent_summary_tables}

# ------------------------------------------------------------
# Drift total composition: percent of all sampled clicks by drift
# ------------------------------------------------------------

drift_total_click_percent_summary <- mapping_hdbscan_with_locations %>%
  count(
    DriftID_clean,
    DriftName,
    name = "n_clicks_in_drift"
  ) %>%
  mutate(
    total_sampled_clicks = sum(n_clicks_in_drift),
    percent_of_all_sampled_clicks = 100 * n_clicks_in_drift / total_sampled_clicks
  ) %>%
  arrange(desc(n_clicks_in_drift))

# ------------------------------------------------------------
# Cluster total composition: percent of all sampled clicks by cluster
# ------------------------------------------------------------

cluster_total_click_percent_summary <- mapping_hdbscan_with_locations %>%
  count(
    HDBSCAN_cluster,
    name = "n_clicks_in_cluster"
  ) %>%
  mutate(
    total_sampled_clicks = sum(n_clicks_in_cluster),
    percent_of_all_sampled_clicks = 100 * n_clicks_in_cluster / total_sampled_clicks,
    cluster_type = if_else(
      HDBSCAN_cluster == "0",
      "unassigned_low_density",
      "assigned_cluster"
    )
  ) %>%
  arrange(as.numeric(as.character(HDBSCAN_cluster)))

# ------------------------------------------------------------
# Drift-by-cluster composition
# ------------------------------------------------------------

drift_cluster_percent_summary <- mapping_hdbscan_with_locations %>%
  count(
    DriftID_clean,
    DriftName,
    HDBSCAN_cluster,
    name = "n_clicks"
  ) %>%
  group_by(DriftID_clean, DriftName) %>%
  mutate(
    n_clicks_in_drift = sum(n_clicks),
    percent_within_drift = 100 * n_clicks / n_clicks_in_drift
  ) %>%
  ungroup() %>%
  mutate(
    total_sampled_clicks = sum(n_clicks),
    percent_of_all_sampled_clicks = 100 * n_clicks / total_sampled_clicks,
    cluster_type = if_else(
      HDBSCAN_cluster == "0",
      "unassigned_low_density",
      "assigned_cluster"
    )
  ) %>%
  arrange(
    DriftID_clean,
    as.numeric(as.character(HDBSCAN_cluster))
  )

# ------------------------------------------------------------
# Save clearly named files
# ------------------------------------------------------------

write.csv(
  drift_total_click_percent_summary,
  file.path(
    drift_map_table_dir,
    paste0("CLEAR_SUMMARY_percent_of_clicks_by_drift_", mapping_result_name, ".csv")
  ),
  row.names = FALSE
)

write.csv(
  cluster_total_click_percent_summary,
  file.path(
    drift_map_table_dir,
    paste0("CLEAR_SUMMARY_percent_of_clicks_by_cluster_TOTAL_", mapping_result_name, ".csv")
  ),
  row.names = FALSE
)

write.csv(
  drift_cluster_percent_summary,
  file.path(
    drift_map_table_dir,
    paste0("CLEAR_SUMMARY_percent_of_clicks_by_drift_AND_cluster_", mapping_result_name, ".csv")
  ),
  row.names = FALSE
)

cat("Saved clear summary tables:\n\n")

cat(file.path(
  drift_map_table_dir,
  paste0("CLEAR_SUMMARY_percent_of_clicks_by_drift_", mapping_result_name, ".csv")
), "\n")

cat(file.path(
  drift_map_table_dir,
  paste0("CLEAR_SUMMARY_percent_of_clicks_by_cluster_TOTAL_", mapping_result_name, ".csv")
), "\n")

cat(file.path(
  drift_map_table_dir,
  paste0("CLEAR_SUMMARY_percent_of_clicks_by_drift_AND_cluster_", mapping_result_name, ".csv")
), "\n")

cat("\nDrift totals preview:\n")
print(drift_total_click_percent_summary, n = Inf)

cat("\nCluster totals preview:\n")
print(cluster_total_click_percent_summary, n = Inf)

cat("\nPASS: Clear drift/cluster percent summary files saved.\n")

```

## ICI boxplot only: 18PC + 1x scaled logICI; minPts = 50

```{r}

library(dplyr)
library(ggplot2)

analysis_name <- "18PC_scaledICI_1"
output_minPts <- 50

result_name <- paste0(analysis_name, "_minPts", output_minPts)

hdbscan_df <- hdbscan_results_list[[result_name]]

if (is.null(hdbscan_df)) {
  stop("Could not find result in hdbscan_results_list: ", result_name)
}

hdbscan_df <- hdbscan_df %>%
  mutate(
    HDBSCAN_cluster = factor(as.character(HDBSCAN_cluster)),
    RowID = as.integer(RowID)
  )

observed_clusters <- sort(as.numeric(as.character(unique(hdbscan_df$HDBSCAN_cluster))))
observed_clusters <- as.character(observed_clusters)

hdbscan_df <- hdbscan_df %>%
  mutate(
    HDBSCAN_cluster = factor(
      as.character(HDBSCAN_cluster),
      levels = observed_clusters
    )
  )

rainbow_cluster_colors <- c(
  "1" = "#E41A1C",
  "2" = "#FF7F00",
  "3" = "#FFD92F",
  "4" = "#4DAF4A",
  "5" = "#00BFC4",
  "6" = "#377EB8",
  "7" = "#984EA3",
  "8" = "#F781BF",
  "9" = "#A65628",
  "10" = "#999999",
  "11" = "#66C2A5",
  "12" = "#FC8D62",
  "13" = "#8DA0CB",
  "14" = "#E78AC3",
  "15" = "#A6D854"
)

cluster_colors_current <- rainbow_cluster_colors[observed_clusters[observed_clusters != "0"]]

hdbscan_no_noise <- hdbscan_df %>%
  filter(HDBSCAN_cluster != "0") %>%
  droplevels()

analysis_plot_dir <- analysis_output_dirs[[analysis_name]]$plots

p_ici_no_noise <- ggplot(
  hdbscan_no_noise,
  aes(
    x = HDBSCAN_cluster,
    y = ICI_to_next_s,
    fill = HDBSCAN_cluster
  )
) +
  geom_boxplot(outlier.size = 0.25) +
  scale_y_log10() +
  scale_fill_manual(values = cluster_colors_current, drop = FALSE) +
  labs(
    title = "ICI Distribution by HDBSCAN Cluster",
    subtitle = "18PC + 1x scaled logICI; minPts = 50; noise excluded",
    x = "HDBSCAN cluster",
    y = "ICI to next click (seconds, log scale)"
  ) +
  theme_bw(base_size = 13) +
  theme(legend.position = "none")

print(p_ici_no_noise)

ggsave(
  file.path(
    analysis_plot_dir,
    "FULL_ici_boxplot_no_noise_18PC_plus_1x_scaled_logICI_minPts50.png"
  ),
  p_ici_no_noise,
  width = 9,
  height = 6,
  dpi = 300
)

```

## Stacked bar charts of event-level cluster proportions by longitude
## 18PC + 1x scaled ICI; minPts = 50

```{r stacked_event_cluster_barplots_18pc_1x_scaled_ici}

# ------------------------------------------------------------
# Safety checks
# ------------------------------------------------------------

if (!exists("mapping_hdbscan_with_locations")) {
  stop("mapping_hdbscan_with_locations does not exist. Run the mapping coordinate-join chunk first.")
}

if (!exists("event_map_plot_dir")) {
  event_map_plot_dir <- file.path(output_dir, "18PC_scaledICI_1", "mapping_outputs", "by_event", "plots")
  dir.create(event_map_plot_dir, recursive = TRUE, showWarnings = FALSE)
}

if (!exists("event_map_table_dir")) {
  event_map_table_dir <- file.path(output_dir, "18PC_scaledICI_1", "mapping_outputs", "by_event", "tables")
  dir.create(event_map_table_dir, recursive = TRUE, showWarnings = FALSE)
}

# ------------------------------------------------------------
# Consistent colors
# ------------------------------------------------------------

event_clusters_all <- sort(
  unique(as.character(mapping_hdbscan_with_locations$HDBSCAN_cluster))
)

cluster_colors_current <- make_cluster_colors(event_clusters_all)

# ------------------------------------------------------------
# Build event-level cluster composition WITH noise
# ------------------------------------------------------------

event_cluster_counts_with_noise <- mapping_hdbscan_with_locations %>%
  mutate(
    HDBSCAN_cluster = as.character(HDBSCAN_cluster)
  ) %>%
  count(
    EventID_clean,
    DriftID_clean,
    DriftName,
    Latitude,
    Longitude,
    HDBSCAN_cluster,
    name = "n_clicks"
  ) %>%
  group_by(EventID_clean) %>%
  mutate(
    total_clicks_event = sum(n_clicks),
    proportion = n_clicks / total_clicks_event,
    percent = 100 * proportion
  ) %>%
  ungroup() %>%
  arrange(Longitude, EventID_clean, as.numeric(HDBSCAN_cluster))

# ------------------------------------------------------------
# Build event-level cluster composition WITHOUT noise
# ------------------------------------------------------------

event_cluster_counts_no_noise <- mapping_hdbscan_with_locations %>%
  mutate(
    HDBSCAN_cluster = as.character(HDBSCAN_cluster)
  ) %>%
  filter(HDBSCAN_cluster != "0") %>%
  count(
    EventID_clean,
    DriftID_clean,
    DriftName,
    Latitude,
    Longitude,
    HDBSCAN_cluster,
    name = "n_clicks"
  ) %>%
  group_by(EventID_clean) %>%
  mutate(
    total_clicks_event_nonnoise = sum(n_clicks),
    proportion = n_clicks / total_clicks_event_nonnoise,
    percent = 100 * proportion
  ) %>%
  ungroup() %>%
  arrange(Longitude, EventID_clean, as.numeric(HDBSCAN_cluster))

# ------------------------------------------------------------
# Save composition files
# ------------------------------------------------------------

write.csv(
  event_cluster_counts_with_noise,
  file.path(
    event_map_table_dir,
    "18PC_1x_scaled_ICI_event_cluster_composition_WITH_noise_minPts50.csv"
  ),
  row.names = FALSE
)

write.csv(
  event_cluster_counts_no_noise,
  file.path(
    event_map_table_dir,
    "18PC_1x_scaled_ICI_event_cluster_composition_NO_noise_minPts50.csv"
  ),
  row.names = FALSE
)

# ------------------------------------------------------------
# Helper function for stacked barplots
# ------------------------------------------------------------

make_event_stacked_barplot <- function(bar_data, include_noise = TRUE, output_name) {
  
  plot_data <- bar_data %>%
    mutate(
      Event_label = paste0(EventID_clean, "\n", round(abs(Longitude), 1), "°W")
    ) %>%
    arrange(Longitude, EventID_clean) %>%
    mutate(
      Event_label = factor(Event_label, levels = unique(Event_label)),
      
      # Reversed levels flip the stacked-bar order.
      # Higher-numbered clusters appear on top.
      HDBSCAN_cluster = factor(
        HDBSCAN_cluster,
        levels = rev(as.character(sort(as.numeric(unique(HDBSCAN_cluster)))))
      )
    )
  
  observed_clusters <- levels(plot_data$HDBSCAN_cluster)
  plot_colors <- cluster_colors_current[observed_clusters]
  
  p <- ggplot(
    plot_data,
    aes(
      x = Event_label,
      y = proportion,
      fill = HDBSCAN_cluster
    )
  ) +
    geom_col(width = 0.9, color = "black", linewidth = 0.1) +
    scale_fill_manual(values = plot_colors, drop = FALSE) +
    scale_y_continuous(
      labels = scales::percent_format(accuracy = 1),
      expand = c(0, 0)
    ) +
    labs(
      title = ifelse(
        include_noise,
        "Event-Level HDBSCAN Cluster Proportions: Including Noise",
        "Event-Level HDBSCAN Cluster Proportions: Excluding Noise"
      ),
      subtitle = "18PC + 1x scaled ICI; minPts = 50; events sorted west-to-east by longitude",
      x = "Event / Longitude",
      y = "Clicks per cluster (%)",
      fill = "HDBSCAN cluster"
    ) +
    theme_bw(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold"),
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 6),
      panel.grid.major.x = element_blank(),
      panel.grid.minor = element_blank(),
      legend.position = "right"
    )
  
  ggsave(
    filename = file.path(event_map_plot_dir, output_name),
    plot = p,
    width = 14,
    height = 6,
    dpi = 300
  )
  
  return(p)
}

# ------------------------------------------------------------
# Generate stacked plots
# ------------------------------------------------------------

p_bar_including_noise <- make_event_stacked_barplot(
  bar_data = event_cluster_counts_with_noise,
  include_noise = TRUE,
  output_name = "18PC_1x_scaled_ICI_event_stacked_barplot_WITH_noise_minPts50_reversed_stack.png"
)

p_bar_excluding_noise <- make_event_stacked_barplot(
  bar_data = event_cluster_counts_no_noise,
  include_noise = FALSE,
  output_name = "18PC_1x_scaled_ICI_event_stacked_barplot_NO_noise_minPts50_reversed_stack.png"
)

print(p_bar_including_noise)
print(p_bar_excluding_noise)

cat("Saved event cluster composition tables to:\n")
cat(event_map_table_dir, "\n\n")

cat("Saved stacked event barplots to:\n")
cat(event_map_plot_dir, "\n")

```

## Stacked bar charts of event-level cluster proportions by latitude
## 18PC + 1x scaled ICI; minPts = 50

```{r stacked_event_cluster_barplots_by_latitude_18pc_1x_scaled_ici}

# ------------------------------------------------------------
# Safety checks
# ------------------------------------------------------------

if (!exists("mapping_hdbscan_with_locations")) {
  stop("mapping_hdbscan_with_locations does not exist. Run the mapping coordinate-join chunk first.")
}

if (!exists("event_map_plot_dir")) {
  event_map_plot_dir <- file.path(output_dir, "18PC_scaledICI_1", "mapping_outputs", "by_event", "plots")
  dir.create(event_map_plot_dir, recursive = TRUE, showWarnings = FALSE)
}

if (!exists("event_map_table_dir")) {
  event_map_table_dir <- file.path(output_dir, "18PC_scaledICI_1", "mapping_outputs", "by_event", "tables")
  dir.create(event_map_table_dir, recursive = TRUE, showWarnings = FALSE)
}

# ------------------------------------------------------------
# Consistent colors
# ------------------------------------------------------------

event_clusters_all <- sort(
  unique(as.character(mapping_hdbscan_with_locations$HDBSCAN_cluster))
)

cluster_colors_current <- make_cluster_colors(event_clusters_all)

# ------------------------------------------------------------
# Build event-level cluster composition WITH noise
# ------------------------------------------------------------

event_cluster_counts_with_noise <- mapping_hdbscan_with_locations %>%
  mutate(
    HDBSCAN_cluster = as.character(HDBSCAN_cluster)
  ) %>%
  count(
    EventID_clean,
    DriftID_clean,
    DriftName,
    Latitude,
    Longitude,
    HDBSCAN_cluster,
    name = "n_clicks"
  ) %>%
  group_by(EventID_clean) %>%
  mutate(
    total_clicks_event = sum(n_clicks),
    proportion = n_clicks / total_clicks_event,
    percent = 100 * proportion
  ) %>%
  ungroup() %>%
  arrange(Latitude, EventID_clean, as.numeric(HDBSCAN_cluster))

# ------------------------------------------------------------
# Build event-level cluster composition WITHOUT noise
# ------------------------------------------------------------

event_cluster_counts_no_noise <- mapping_hdbscan_with_locations %>%
  mutate(
    HDBSCAN_cluster = as.character(HDBSCAN_cluster)
  ) %>%
  filter(HDBSCAN_cluster != "0") %>%
  count(
    EventID_clean,
    DriftID_clean,
    DriftName,
    Latitude,
    Longitude,
    HDBSCAN_cluster,
    name = "n_clicks"
  ) %>%
  group_by(EventID_clean) %>%
  mutate(
    total_clicks_event_nonnoise = sum(n_clicks),
    proportion = n_clicks / total_clicks_event_nonnoise,
    percent = 100 * proportion
  ) %>%
  ungroup() %>%
  arrange(Latitude, EventID_clean, as.numeric(HDBSCAN_cluster))

# ------------------------------------------------------------
# Save composition files
# ------------------------------------------------------------

write.csv(
  event_cluster_counts_with_noise,
  file.path(
    event_map_table_dir,
    "18PC_1x_scaled_ICI_event_cluster_composition_WITH_noise_minPts50_sorted_by_latitude.csv"
  ),
  row.names = FALSE
)

write.csv(
  event_cluster_counts_no_noise,
  file.path(
    event_map_table_dir,
    "18PC_1x_scaled_ICI_event_cluster_composition_NO_noise_minPts50_sorted_by_latitude.csv"
  ),
  row.names = FALSE
)

# ------------------------------------------------------------
# Helper function for stacked barplots
# ------------------------------------------------------------

make_event_stacked_barplot <- function(bar_data, include_noise = TRUE, output_name) {
  
  plot_data <- bar_data %>%
    mutate(
      Event_label = paste0(EventID_clean, "\n", round(Latitude, 1), "°N")
    ) %>%
    arrange(Latitude, EventID_clean) %>%
    mutate(
      Event_label = factor(Event_label, levels = unique(Event_label)),
      
      # Reversed levels flip the stacked-bar order.
      # Higher-numbered clusters appear on top.
      HDBSCAN_cluster = factor(
        HDBSCAN_cluster,
        levels = rev(as.character(sort(as.numeric(unique(HDBSCAN_cluster)))))
      )
    )
  
  observed_clusters <- levels(plot_data$HDBSCAN_cluster)
  plot_colors <- cluster_colors_current[observed_clusters]
  
  p <- ggplot(
    plot_data,
    aes(
      x = Event_label,
      y = proportion,
      fill = HDBSCAN_cluster
    )
  ) +
    geom_col(width = 0.9, color = "black", linewidth = 0.1) +
    scale_fill_manual(values = plot_colors, drop = FALSE) +
    scale_y_continuous(
      labels = scales::percent_format(accuracy = 1),
      expand = c(0, 0)
    ) +
    labs(
      title = ifelse(
        include_noise,
        "Event-Level HDBSCAN Cluster Proportions: Including Noise",
        "Event-Level HDBSCAN Cluster Proportions: Excluding Noise"
      ),
      subtitle = "18PC + 1x scaled ICI; minPts = 50; events sorted south-to-north by latitude",
      x = "Event / Latitude",
      y = "Clicks per cluster (%)",
      fill = "HDBSCAN cluster"
    ) +
    theme_bw(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold"),
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 6),
      panel.grid.major.x = element_blank(),
      panel.grid.minor = element_blank(),
      legend.position = "right"
    )
  
  ggsave(
    filename = file.path(event_map_plot_dir, output_name),
    plot = p,
    width = 14,
    height = 6,
    dpi = 300
  )
  
  return(p)
}

# ------------------------------------------------------------
# Generate stacked plots
# ------------------------------------------------------------

p_bar_including_noise_by_latitude <- make_event_stacked_barplot(
  bar_data = event_cluster_counts_with_noise,
  include_noise = TRUE,
  output_name = "18PC_1x_scaled_ICI_event_stacked_barplot_WITH_noise_minPts50_sorted_by_latitude.png"
)

p_bar_excluding_noise_by_latitude <- make_event_stacked_barplot(
  bar_data = event_cluster_counts_no_noise,
  include_noise = FALSE,
  output_name = "18PC_1x_scaled_ICI_event_stacked_barplot_NO_noise_minPts50_sorted_by_latitude.png"
)

print(p_bar_including_noise_by_latitude)
print(p_bar_excluding_noise_by_latitude)

cat("Saved event cluster composition tables to:\n")
cat(event_map_table_dir, "\n\n")

cat("Saved stacked event barplots to:\n")
cat(event_map_plot_dir, "\n")

```

## Pick Peaks from Saved Primary Cluster 9 Mean Spectrum

```{r pickpeaks_cluster_9_saved_primary}

# ------------------------------------------------------------
# Load packages
# ------------------------------------------------------------

library(dplyr)
library(ggplot2)
library(readr)
library(pracma)

# ------------------------------------------------------------
# Load the ORIGINAL saved primary spectra table
# ------------------------------------------------------------

primary_cluster_spectra_summary <- read_csv(
  "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/tables/clean_cluster_mean_spectra_18PC_scaledICI_1_minPts50.csv",
  show_col_types = FALSE
) %>%
  mutate(
    HDBSCAN_cluster = as.character(HDBSCAN_cluster)
  )

cat("Clusters found in saved primary spectra table:\n")
print(sort(unique(primary_cluster_spectra_summary$HDBSCAN_cluster)))

# ------------------------------------------------------------
# Pull cluster 9
# ------------------------------------------------------------

cluster_9_spectrum <- primary_cluster_spectra_summary %>%
  filter(HDBSCAN_cluster == "9") %>%
  arrange(Frequency_kHz)

if (nrow(cluster_9_spectrum) == 0) {
  stop("Cluster 9 was not found in the saved primary spectra table.")
}

cat("\nRows in cluster 9 spectrum:\n")
print(nrow(cluster_9_spectrum))

# ------------------------------------------------------------
# Pick peaks from cluster 9 mean spectrum
# ------------------------------------------------------------

peak_matrix <- pracma::findpeaks(
  cluster_9_spectrum$mean_minmax_spectrum,
  nups = 2,
  ndowns = 2,
  minpeakdistance = 3,
  minpeakheight = 0.15
)

if (is.null(peak_matrix)) {
  stop("No peaks detected. Try lowering minpeakheight or minpeakdistance.")
}

# ------------------------------------------------------------
# Build peak table
# ------------------------------------------------------------

cluster_9_peaks <- data.frame(
  peak_height = peak_matrix[, 1],
  peak_index = peak_matrix[, 2],
  peak_start_index = peak_matrix[, 3],
  peak_end_index = peak_matrix[, 4]
) %>%
  mutate(
    Frequency_kHz = cluster_9_spectrum$Frequency_kHz[peak_index]
  ) %>%
  arrange(Frequency_kHz)

cat("\nDetected peaks:\n")
print(cluster_9_peaks)

# ------------------------------------------------------------
# Save peak table
# ------------------------------------------------------------

write.csv(
  cluster_9_peaks,
  file = "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/tables/cluster_9_pickpeaks_mean_spectrum.csv",
  row.names = FALSE
)

# ------------------------------------------------------------
# Plot cluster 9 spectrum with peaks
# ------------------------------------------------------------

p_cluster_9_pickpeaks <- ggplot(
  cluster_9_spectrum,
  aes(
    x = Frequency_kHz,
    y = mean_minmax_spectrum
  )
) +
  geom_line(
    linewidth = 1.3,
    color = "#A65628"
  ) +
  geom_point(
    data = cluster_9_peaks,
    aes(
      x = Frequency_kHz,
      y = peak_height
    ),
    size = 3,
    color = "black"
  ) +
  geom_text(
    data = cluster_9_peaks,
    aes(
      x = Frequency_kHz,
      y = peak_height,
      label = paste0(round(Frequency_kHz, 1), " kHz")
    ),
    vjust = -0.9,
    size = 3.8
  ) +
  coord_cartesian(
    xlim = c(19, 50),
    ylim = c(0, 1.08)
  ) +
  labs(
    title = "Picked Peaks from Cluster 9 Mean Spectrum",
    subtitle = "18PC + 1x scaled logICI; minPts = 50",
    x = "Frequency (kHz)",
    y = "Mean min-max normalized amplitude"
  ) +
  theme_bw(base_size = 14)

print(p_cluster_9_pickpeaks)

# ------------------------------------------------------------
# Save peak plot
# ------------------------------------------------------------

ggsave(
  filename = "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/plots/cluster_9_pickpeaks_mean_spectrum.png",
  plot = p_cluster_9_pickpeaks,
  width = 9,
  height = 5.5,
  dpi = 300
)

cat("\nSaved:\n")
cat("- cluster_9_pickpeaks_mean_spectrum.csv\n")
cat("- cluster_9_pickpeaks_mean_spectrum.png\n")

```

## Pick Peaks and Notches for Every Cluster Mean Spectrum

```{r pickpeaks_notches_all_clusters_saved_primary}

library(dplyr)
library(ggplot2)
library(readr)
library(pracma)
library(purrr)

# ------------------------------------------------------------
# Input / output paths
# ------------------------------------------------------------

primary_spectra_file <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/tables/clean_cluster_mean_spectra_18PC_scaledICI_1_minPts50.csv"

feature_table_out <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/tables/all_clusters_pickpeaks_and_notches_mean_spectra.csv"

feature_plot_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/plots/peak_notch_picking_by_cluster"

dir.create(
  feature_plot_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

# ------------------------------------------------------------
# Cluster colors from earlier script
# Noise cluster 0 = grey80
# ------------------------------------------------------------

cluster_colors_for_features <- c(
  "0" = "grey60",
  "1" = "#E41A1C",  # red
  "2" = "#FF7F00",  # orange
  "3" = "#FFD92F",  # yellow
  "4" = "#4DAF4A",  # green
  "5" = "#00BFC4",  # cyan
  "6" = "#377EB8",  # blue
  "7" = "#984EA3",  # purple
  "8" = "#F781BF",  # pink
  "9" = "#A65628",  # brown
  "10" = "#999999",
  "11" = "#66C2A5",
  "12" = "#FC8D62"
)

# ------------------------------------------------------------
# Load saved primary spectra table
# ------------------------------------------------------------

primary_cluster_spectra_summary <- read_csv(
  primary_spectra_file,
  show_col_types = FALSE
) %>%
  mutate(
    HDBSCAN_cluster = as.character(HDBSCAN_cluster)
  )

available_clusters <- sort(
  as.numeric(unique(primary_cluster_spectra_summary$HDBSCAN_cluster))
)

available_clusters <- as.character(available_clusters)

cat("Clusters found:\n")
print(available_clusters)

# ------------------------------------------------------------
# Helper function: pick peaks and notches for one cluster
# ------------------------------------------------------------

pick_features_one_cluster <- function(cluster_id) {
  
  one_spectrum <- primary_cluster_spectra_summary %>%
    filter(HDBSCAN_cluster == cluster_id) %>%
    arrange(Frequency_kHz)
  
  if (nrow(one_spectrum) == 0) {
    warning("Skipping missing cluster: ", cluster_id)
    return(NULL)
  }
  
  # --------------------------
  # Peaks
  # --------------------------
  
  peak_matrix <- pracma::findpeaks(
    one_spectrum$mean_minmax_spectrum,
    nups = 2,
    ndowns = 2,
    minpeakdistance = 3,
    minpeakheight = 0.15
  )
  
  if (is.null(peak_matrix)) {
    
    peaks_df <- tibble()
    
  } else {
    
    peaks_df <- tibble(
      HDBSCAN_cluster = cluster_id,
      feature_type = "peak",
      feature_height = peak_matrix[, 1],
      feature_index = peak_matrix[, 2],
      feature_start_index = peak_matrix[, 3],
      feature_end_index = peak_matrix[, 4],
      Frequency_kHz = one_spectrum$Frequency_kHz[feature_index]
    )
  }
  
  # --------------------------
  # Notches
  # Notches are peaks in the inverted spectrum.
  # --------------------------
  
  notch_matrix <- pracma::findpeaks(
    -one_spectrum$mean_minmax_spectrum,
    nups = 2,
    ndowns = 2,
    minpeakdistance = 3,
    minpeakheight = -0.85
  )
  
  if (is.null(notch_matrix)) {
    
    notches_df <- tibble()
    
  } else {
    
    notches_df <- tibble(
      HDBSCAN_cluster = cluster_id,
      feature_type = "notch",
      feature_height = one_spectrum$mean_minmax_spectrum[notch_matrix[, 2]],
      feature_index = notch_matrix[, 2],
      feature_start_index = notch_matrix[, 3],
      feature_end_index = notch_matrix[, 4],
      Frequency_kHz = one_spectrum$Frequency_kHz[feature_index]
    )
  }
  
  features_df <- bind_rows(
    peaks_df,
    notches_df
  ) %>%
    arrange(Frequency_kHz, feature_type)
  
  return(features_df)
}

# ------------------------------------------------------------
# Run peak/notch picking for all clusters
# ------------------------------------------------------------

all_cluster_features <- map_dfr(
  available_clusters,
  pick_features_one_cluster
) %>%
  mutate(
    HDBSCAN_cluster = as.character(HDBSCAN_cluster),
    feature_type = factor(feature_type, levels = c("peak", "notch"))
  ) %>%
  arrange(
    as.numeric(HDBSCAN_cluster),
    Frequency_kHz,
    feature_type
  )

cat("\nDetected peak/notch features:\n")
print(all_cluster_features, n = Inf)

# ------------------------------------------------------------
# Save combined feature table
# ------------------------------------------------------------

write.csv(
  all_cluster_features,
  file = feature_table_out,
  row.names = FALSE
)

# ------------------------------------------------------------
# Plot and save one figure per cluster
# ------------------------------------------------------------

for (cluster_id in available_clusters) {
  
  one_spectrum <- primary_cluster_spectra_summary %>%
    filter(HDBSCAN_cluster == cluster_id) %>%
    arrange(Frequency_kHz)
  
  one_features <- all_cluster_features %>%
    filter(HDBSCAN_cluster == cluster_id)
  
  cluster_line_color <- cluster_colors_for_features[[cluster_id]]
  
  if (is.null(cluster_line_color) || is.na(cluster_line_color)) {
    cluster_line_color <- "black"
  }
  
  cluster_label <- ifelse(
    cluster_id == "0",
    "Noise / unassigned cluster 0",
    paste0("Cluster ", cluster_id)
  )
  
  p_one_cluster <- ggplot(
    one_spectrum,
    aes(
      x = Frequency_kHz,
      y = mean_minmax_spectrum
    )
  ) +
    geom_line(
      linewidth = 1.35,
      color = cluster_line_color
    ) +
    geom_point(
      data = one_features,
      aes(
        x = Frequency_kHz,
        y = feature_height,
        shape = feature_type
      ),
      size = 3.2,
      color = "black"
    ) +
    geom_text(
      data = one_features,
      aes(
        x = Frequency_kHz,
        y = feature_height,
        label = paste0(round(Frequency_kHz, 1), " kHz")
      ),
      vjust = ifelse(one_features$feature_type == "peak", -0.9, 1.5),
      size = 3.4
    ) +
    scale_shape_manual(
      values = c(
        "peak" = 17,
        "notch" = 25
      ),
      drop = FALSE
    ) +
    coord_cartesian(
      xlim = c(19, 50),
      ylim = c(0, 1.08)
    ) +
    labs(
      title = paste0("Picked Peaks and Notches: ", cluster_label),
      subtitle = "18PC + 1x scaled logICI; minPts = 50",
      x = "Frequency (kHz)",
      y = "Mean min-max normalized amplitude",
      shape = "Feature"
    ) +
    theme_bw(base_size = 14) +
    theme(
      plot.title = element_text(face = "bold")
    )
  
  print(p_one_cluster)
  
  ggsave(
    filename = file.path(
      feature_plot_dir,
      paste0("cluster_", cluster_id, "_picked_peaks_and_notches.png")
    ),
    plot = p_one_cluster,
    width = 10,
    height = 5.8,
    dpi = 300
  )
}

cat("\nSaved combined peak/notch table to:\n")
cat(feature_table_out, "\n\n")

cat("Saved per-cluster plots to:\n")
cat(feature_plot_dir, "\n")

```














