---
title: "ScalingTest"
author: "Sarah Shiers"
date: "2026-04-22"
output: html_document
---

## Load required package

```{r}

library(ggplot2)

```

## User Settings

```{r}

freq_min <- 19000
freq_max <- 50000
n_clicks_to_keep <- 5

```

## Load combined Tukey spectra

```{r}

combined_file <- file.choose()
combined_tukey <- readRDS(combined_file)

```

## Choose reference frequency grid

```{r}

reference_freq <- combined_tukey[[1]]$freq

```

## Helper function to interpolate spectra to the shared grid

```{r}

interpolate_spectrum <- function(old_freq, old_spec, new_freq) {
  approx(
    x = old_freq,
    y = old_spec,
    xout = new_freq,
    method = "linear",
    rule = 2
  )$y
}

```

## Interpolate all events to the shared frequency grid

```{r}

interpolated_tukey <- vector("list", length(combined_tukey))

for (i in seq_along(combined_tukey)) {
  
  tukey_spectra <- combined_tukey[[i]]
  old_freq <- tukey_spectra$freq
  old_allSpec <- tukey_spectra$allSpec
  
  interpolated_allSpec <- apply(
    old_allSpec,
    2,
    function(spec) interpolate_spectrum(old_freq, spec, reference_freq)
  )
  
  interpolated_allSpec <- as.matrix(interpolated_allSpec)
  
  tukey_spectra$allSpec <- interpolated_allSpec
  tukey_spectra$freq <- reference_freq
  
  interpolated_tukey[[i]] <- tukey_spectra
}

all_match_after_interp <- all(
  sapply(interpolated_tukey, function(x) identical(x$freq, reference_freq))
)

all_match_after_interp

```

## Combine all click spectra across events

```{r}

all_click_spectra <- list()

for (i in seq_along(interpolated_tukey)) {
  all_click_spectra[[i]] <- t(interpolated_tukey[[i]]$allSpec)
}

all_spec_matrix <- do.call(rbind, all_click_spectra)

```

## Keep only the first 5 clicks

```{r}

first5_matrix <- all_spec_matrix[1:n_clicks_to_keep, , drop = FALSE]

```

## Restrict to 19–50 kHz

```{r}

keepFreq <- which(reference_freq >= freq_min & reference_freq <= freq_max)

first5_focus <- first5_matrix[, keepFreq, drop = FALSE]
focus_freq <- reference_freq[keepFreq]

```

## Check dimensions

```{r}

dim(first5_focus)
focus_freq[1]
focus_freq[length(focus_freq)]

```

## Normalize each click to a zero baseline - Subtract the first retained frequency value from each spectrum

```{r}

first5_norm <- sweep(first5_focus, 1, first5_focus[, 1], FUN = "-")

```

## Whole-spectrum scaling - Scaling within each click / row

```{r}

first5_whole_scaled <- t(scale(t(first5_norm)))
first5_whole_scaled[is.na(first5_whole_scaled)] <- 0

```

## Frequency-value scaling - Scaling within each frequency bin / column across clicks)

```{r}

first5_freq_scaled <- scale(first5_norm)
first5_freq_scaled[is.na(first5_freq_scaled)] <- 0

```

## Sanity Checks

```{r}

## Check dimensions
dim(first5_norm)
dim(first5_whole_scaled)
dim(first5_freq_scaled)

## Confirm normalization worked:
## first value in each click should now be 0
first5_norm[, 1]

```

## Plot the first 5 clicks for each scaling approach

```{r}

## Save plots

output_dir <- "E:/Analysis/ScalingTest"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

click_colors <- 1:5

## 1. Normalized only
png(file.path(output_dir, "first5_normalized.png"), width = 1000, height = 700)

matplot(
  x = focus_freq,
  y = t(first5_norm),
  type = "l",
  lty = 1,
  lwd = 2,
  col = click_colors,
  xlab = "Frequency (Hz)",
  ylab = "Spectral value",
  main = "First 5 Clicks - Normalized Only"
)

legend(
  "topright",
  legend = paste0("Click_", 1:5),
  col = click_colors,
  lty = 1,
  lwd = 2
)

dev.off()


## 2. Whole-spectrum scaling
png(file.path(output_dir, "first5_whole_scaled.png"), width = 1000, height = 700)

matplot(
  x = focus_freq,
  y = t(first5_whole_scaled),
  type = "l",
  lty = 1,
  lwd = 2,
  col = click_colors,
  xlab = "Frequency (Hz)",
  ylab = "Scaled spectral value",
  main = "First 5 Clicks - Whole Spectrum Scaling"
)

legend(
  "topright",
  legend = paste0("Click_", 1:5),
  col = click_colors,
  lty = 1,
  lwd = 2
)

dev.off()


## 3. Frequency-value scaling
png(file.path(output_dir, "first5_freq_scaled.png"), width = 1000, height = 700)

matplot(
  x = focus_freq,
  y = t(first5_freq_scaled),
  type = "l",
  lty = 1,
  lwd = 2,
  col = click_colors,
  xlab = "Frequency (Hz)",
  ylab = "Scaled spectral value",
  main = "First 5 Clicks - Frequency Value Scaling"
)

legend(
  "topright",
  legend = paste0("Click_", 1:5),
  col = click_colors,
  lty = 1,
  lwd = 2
)

dev.off()

```

## Save each click as its own comparison plot

```{r}

comparison_colors <- c("black", "red", "blue")

for (i in 1:n_clicks_to_keep) {
  
  png(
    file.path(output_dir, paste0("click_", i, "_comparison.png")),
    width = 1000,
    height = 700
  )
  
  y_min <- min(
    first5_norm[i, ],
    first5_whole_scaled[i, ],
    first5_freq_scaled[i, ]
  )
  
  y_max <- max(
    first5_norm[i, ],
    first5_whole_scaled[i, ],
    first5_freq_scaled[i, ]
  )
  
  plot(
    focus_freq,
    first5_norm[i, ],
    type = "l",
    lwd = 2,
    col = comparison_colors[1],
    xlab = "Frequency (Hz)",
    ylab = "Spectral value",
    main = paste("Click", i, "- Scaling Comparison"),
    ylim = c(y_min, y_max)
  )
  
  lines(
    focus_freq,
    first5_whole_scaled[i, ],
    lwd = 2,
    col = comparison_colors[2]
  )
  
  lines(
    focus_freq,
    first5_freq_scaled[i, ],
    lwd = 2,
    col = comparison_colors[3]
  )
  
  legend(
    "topright",
    legend = c("Normalized only", "Whole spectrum", "Frequency value"),
    col = comparison_colors,
    lty = 1,
    lwd = 2
  )
  
  dev.off()
}

```

## Two-panel Comparison Plots

```{r}

scaled_colors <- c("red", "blue")

for (i in 1:n_clicks_to_keep) {
  
  png(
    file.path(output_dir, paste0("click_", i, "_two_panel_comparison.png")),
    width = 1400,
    height = 700
  )
  
  par(mfrow = c(1, 2), mar = c(4, 4, 3, 1))
  
  ## Left panel: normalized only
  plot(
    focus_freq,
    first5_norm[i, ],
    type = "l",
    lwd = 2,
    col = "black",
    xlab = "Frequency (Hz)",
    ylab = "Spectral value",
    main = paste("Click", i, "- Normalized Only")
  )
  
  ## Right panel: scaled versions only
  y_min_scaled <- min(first5_whole_scaled[i, ], first5_freq_scaled[i, ])
  y_max_scaled <- max(first5_whole_scaled[i, ], first5_freq_scaled[i, ])
  
  plot(
    focus_freq,
    first5_whole_scaled[i, ],
    type = "l",
    lwd = 2,
    col = scaled_colors[1],
    xlab = "Frequency (Hz)",
    ylab = "Scaled spectral value",
    main = paste("Click", i, "- Scaled Comparison"),
    ylim = c(y_min_scaled, y_max_scaled)
  )
  
  lines(
    focus_freq,
    first5_freq_scaled[i, ],
    lwd = 2,
    col = scaled_colors[2]
  )
  
  legend(
    "topright",
    legend = c("Whole spectrum", "Frequency value"),
    col = scaled_colors,
    lty = 1,
    lwd = 2
  )
  
  dev.off()
}

```























