---
title: "Soldevilla 2017 Cluster Peak and Notch Extraction"
author: "Sarah Moseley"
date: "2026-06-08"
output: html_document
---

## Setup

```{r}

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)

required_packages <- c(
  "dplyr",
  "ggplot2",
  "tidyr",
  "readr",
  "purrr",
  "tibble",
  "pracma",
  "R.matlab"
)

missing_packages <- required_packages[
  !sapply(required_packages, requireNamespace, quietly = TRUE)
]

if (length(missing_packages) > 0) {
  stop(
    "Missing required packages: ",
    paste(missing_packages, collapse = ", "),
    "\nInstall these before running the workflow."
  )
}

library(dplyr)
library(ggplot2)
library(tidyr)
library(readr)
library(purrr)
library(tibble)
library(pracma)
library(R.matlab)

```

## User Settings

```{r}

# ------------------------------------------------------------
# Input directory: Melissa Soldevilla figure files
# ------------------------------------------------------------

soldevilla_input_dir <- "C:/Users/sarah/OneDrive/Documents/GitHub/Rissos_Geographic_Variation/Data/Soldevilla Clusters"

# Mean spectra MATLAB figure
mean_spectra_fig <- file.path(
  soldevilla_input_dir,
  "FinalColorCodedClusterMeans_9k9best.fig"
)

# Concatenated spectrogram MATLAB figure
# We are NOT using this yet, but keeping the path here for reference.
spectrogram_fig <- file.path(
  soldevilla_input_dir,
  "FinalCMSClusterSpectrograms_9k9best.fig"
)

# ------------------------------------------------------------
# Output directory
# ------------------------------------------------------------

base_output_dir <- "E:/Analysis/Soldevilla Peaks and Notches"

dir.create(
  base_output_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

# ------------------------------------------------------------
# Basic frequency settings
# ------------------------------------------------------------

freq_min_khz <- 20
freq_max_khz <- 50

# ------------------------------------------------------------
# Sanity checks
# ------------------------------------------------------------

cat("Mean spectra figure exists:\n")
print(file.exists(mean_spectra_fig))

cat("\nSpectrogram figure exists:\n")
print(file.exists(spectrogram_fig))

cat("\nOutput directory exists:\n")
print(dir.exists(base_output_dir))

cat("\nMean spectra figure path:\n")
cat(mean_spectra_fig, "\n")

cat("\nOutput directory:\n")
cat(base_output_dir, "\n")

```

## Read MATLAB .fig file

```{r}

# ------------------------------------------------------------
# Load MATLAB .fig file containing the 9 mean spectra
# ------------------------------------------------------------

mean_spectra_mat <- R.matlab::readMat(mean_spectra_fig)

cat("Loaded MATLAB figure object.\n\n")

cat("Top-level object names:\n")
print(names(mean_spectra_mat))

```

## Inspect file structure

```{r}

# ------------------------------------------------------------
# Inspect the MATLAB figure structure
# ------------------------------------------------------------

cat("Class of loaded object:\n")
print(class(mean_spectra_mat))

cat("\nStructure preview:\n")
str(mean_spectra_mat, max.level = 3)

```

## Inspect the Axes Children

```{r}

# ------------------------------------------------------------
# Pull out figure, axes, and axes children using numeric positions
#
# From the structure:
# fig[[1]] = type
# fig[[2]] = handle
# fig[[3]] = properties
# fig[[4]] = children
# ------------------------------------------------------------

fig <- mean_spectra_mat[[1]]

fig_type <- fig[[1]]
fig_children <- fig[[4]]

cat("Figure type:\n")
print(fig_type)

cat("\nNumber of figure children:\n")
print(length(fig_children))

cat("\nStructure of figure children:\n")
str(fig_children, max.level = 2)

```

## Extract the 9 Line Objects

```{r}

# ------------------------------------------------------------
# Extract line objects from the axes children
#
# The axes children are stored in repeating groups of 5:
#   1 = object type
#   2 = handle
#   3 = properties
#   4 = children
#   5 = special
# ------------------------------------------------------------

axes_children <- fig_children[[4]]

n_object_fields <- 5

object_starts <- seq(
  from = 1,
  to = length(axes_children),
  by = n_object_fields
)

object_types <- sapply(object_starts, function(i) {
  as.character(axes_children[[i]][1, 1])
})

object_summary <- tibble(
  object_index = seq_along(object_starts),
  start_position = object_starts,
  object_type = object_types
)

cat("Objects found in axes children:\n")
print(object_summary)

# ------------------------------------------------------------
# Keep only the 9 plotted mean spectra line objects
# ------------------------------------------------------------

line_object_starts <- object_starts[
  object_types == "graph2d.lineseries"
]

cat("\nNumber of line objects found:\n")
print(length(line_object_starts))

if (length(line_object_starts) != 9) {
  stop("Expected 9 line objects, but found ", length(line_object_starts))
}

line_objects <- lapply(line_object_starts, function(i) {
  list(
    type = axes_children[[i]],
    handle = axes_children[[i + 1]],
    properties = axes_children[[i + 2]],
    children = axes_children[[i + 3]],
    special = axes_children[[i + 4]]
  )
})

cat("\nPASS: Extracted 9 line objects.\n")

```

## Inspect One Line's Object Properties

```{r}

# ------------------------------------------------------------
# Inspect properties for the first line object
# ------------------------------------------------------------

first_line <- line_objects[[1]]

cat("First line object type:\n")
print(first_line$type)

cat("\nFirst line handle:\n")
print(first_line$handle)

cat("\nFirst line property names:\n")
print(dimnames(first_line$properties)[[1]])

cat("\nFirst line property structure:\n")
str(first_line$properties, max.level = 2)

```

## Extract XData / YData From All 9 Clusters

```{r}

# ------------------------------------------------------------
# Helper function to extract XData and YData from one line object
# ------------------------------------------------------------

extract_line_xy <- function(line_obj, line_number) {
  
  props <- line_obj$properties
  
  prop_names <- dimnames(props)[[1]]
  
  x_index <- which(prop_names == "XData")
  y_index <- which(prop_names == "YData")
  
  if (length(x_index) != 1 || length(y_index) != 1) {
    stop("Could not uniquely find XData/YData for line ", line_number)
  }
  
  x_data <- as.numeric(props[[x_index]])
  y_data <- as.numeric(props[[y_index]])
  
  if (length(x_data) != length(y_data)) {
    stop("XData and YData have different lengths for line ", line_number)
  }
  
  tibble(
    original_line_number = line_number,
    Frequency_kHz = x_data,
    mean_spectrum_raw = y_data
  )
}

# ------------------------------------------------------------
# Extract all 9 spectra
# ------------------------------------------------------------

soldevilla_mean_spectra_raw <- map2_dfr(
  line_objects,
  seq_along(line_objects),
  extract_line_xy
)

cat("Extracted spectra table dimensions:\n")
print(dim(soldevilla_mean_spectra_raw))

cat("\nRows per original line:\n")
print(
  soldevilla_mean_spectra_raw %>%
    count(original_line_number)
)

cat("\nFrequency summary:\n")
print(summary(soldevilla_mean_spectra_raw$Frequency_kHz))

cat("\nPreview:\n")
print(head(soldevilla_mean_spectra_raw))

```

## Assign Soldevilla Cluster Numbers

```{r}

# ------------------------------------------------------------
# Assign Soldevilla cluster numbers based on vertical order
#
# User-confirmed figure layout:
#   top line    = Soldevilla cluster 1
#   bottom line = Soldevilla cluster 9
# ------------------------------------------------------------

line_vertical_order <- soldevilla_mean_spectra_raw %>%
  group_by(original_line_number) %>%
  summarise(
    mean_y_position = mean(mean_spectrum_raw, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(mean_y_position)) %>%
  mutate(
    Soldevilla_cluster = row_number()
  )

cat("Line-to-cluster mapping:\n")
print(line_vertical_order)

soldevilla_mean_spectra <- soldevilla_mean_spectra_raw %>%
  left_join(
    line_vertical_order %>%
      select(original_line_number, Soldevilla_cluster),
    by = "original_line_number"
  ) %>%
  arrange(Soldevilla_cluster, Frequency_kHz)

cat("\nRows per Soldevilla cluster:\n")
print(
  soldevilla_mean_spectra %>%
    count(Soldevilla_cluster)
)

cat("\nPreview labeled spectra:\n")
print(head(soldevilla_mean_spectra))

```

## Visualize Extracted Spectra

```{r}

# ------------------------------------------------------------
# Plot extracted spectra exactly as stored in the figure
# ------------------------------------------------------------

p_raw <- ggplot(
  soldevilla_mean_spectra,
  aes(
    x = Frequency_kHz,
    y = mean_spectrum_raw,
    color = factor(Soldevilla_cluster)
  )
) +
  geom_line(linewidth = 1) +
  labs(
    title = "Extracted Soldevilla Mean Spectra",
    subtitle = "Raw values from MATLAB figure",
    x = "Frequency (kHz)",
    y = "Amplitude",
    color = "Cluster"
  ) +
  theme_bw()

print(p_raw)

```

## Test Peak Picking on Cluster 1

```{r}

cluster_1_spectrum <- soldevilla_mean_spectra %>%
  filter(Soldevilla_cluster == 1) %>%
  arrange(Frequency_kHz) %>%
  mutate(
    spectrum_minmax =
      (mean_spectrum_raw - min(mean_spectrum_raw, na.rm = TRUE)) /
      (max(mean_spectrum_raw, na.rm = TRUE) - min(mean_spectrum_raw, na.rm = TRUE))
  )

peak_matrix <- pracma::findpeaks(
  cluster_1_spectrum$spectrum_minmax,
  nups = 2,
  ndowns = 2,
  minpeakdistance = 3,
  minpeakheight = 0.15
)

cluster_1_peaks <- tibble(
  peak_height = peak_matrix[, 1],
  peak_index = peak_matrix[, 2],
  peak_start_index = peak_matrix[, 3],
  peak_end_index = peak_matrix[, 4],
  peak_frequency_kHz = cluster_1_spectrum$Frequency_kHz[peak_index]
) %>%
  arrange(peak_frequency_kHz) %>%
  mutate(
    peak_label = paste0("Peak ", row_number())
  )

print(cluster_1_peaks)

p_cluster_1_peaks <- ggplot(
  cluster_1_spectrum,
  aes(x = Frequency_kHz, y = spectrum_minmax)
) +
  geom_line(linewidth = 1) +
  geom_point(
    data = cluster_1_peaks,
    aes(x = peak_frequency_kHz, y = peak_height),
    color = "red",
    size = 3
  ) +
  geom_text(
    data = cluster_1_peaks,
    aes(
      x = peak_frequency_kHz,
      y = peak_height,
      label = paste0(round(peak_frequency_kHz, 1), " kHz")
    ),
    vjust = -0.8,
    size = 3.5
  ) +
  labs(
    title = "Peak Picking Test: Soldevilla Cluster 1",
    subtitle = "Within-cluster min-max spectrum; peaks sorted by frequency",
    x = "Frequency (kHz)",
    y = "Min-max normalized amplitude"
  ) +
  theme_bw()

print(p_cluster_1_peaks)

```

## Detect Peaks and Notches for All 9 Clusters

```{r}

# ------------------------------------------------------------
# Normalize each Soldevilla cluster spectrum within cluster
# ------------------------------------------------------------

soldevilla_spectra_minmax <- soldevilla_mean_spectra %>%
  group_by(Soldevilla_cluster) %>%
  mutate(
    spectrum_minmax =
      (mean_spectrum_raw - min(mean_spectrum_raw, na.rm = TRUE)) /
      (max(mean_spectrum_raw, na.rm = TRUE) - min(mean_spectrum_raw, na.rm = TRUE))
  ) %>%
  ungroup()

# ------------------------------------------------------------
# Helper function: pick peaks and notches for one cluster
# ------------------------------------------------------------

pick_features_one_cluster <- function(cluster_df) {
  
  cluster_id <- unique(cluster_df$Soldevilla_cluster)
  
  cluster_df <- cluster_df %>%
    arrange(Frequency_kHz)
  
  # Peaks
  peak_matrix <- pracma::findpeaks(
    cluster_df$spectrum_minmax,
    nups = 2,
    ndowns = 2,
    minpeakdistance = 3,
    minpeakheight = 0.15
  )
  
  peak_table <- tibble()
  
  if (!is.null(peak_matrix)) {
    peak_table <- tibble(
      Soldevilla_cluster = cluster_id,
      feature_type = "peak",
      feature_height = peak_matrix[, 1],
      feature_index = peak_matrix[, 2],
      feature_start_index = peak_matrix[, 3],
      feature_end_index = peak_matrix[, 4],
      Frequency_kHz = cluster_df$Frequency_kHz[feature_index]
    )
  }
  
  # Notches = peaks in the inverted spectrum
  notch_matrix <- pracma::findpeaks(
    -cluster_df$spectrum_minmax,
    nups = 2,
    ndowns = 2,
    minpeakdistance = 3
  )
  
  notch_table <- tibble()
  
  if (!is.null(notch_matrix)) {
    notch_table <- tibble(
      Soldevilla_cluster = cluster_id,
      feature_type = "notch",
      feature_height = -notch_matrix[, 1],
      feature_index = notch_matrix[, 2],
      feature_start_index = notch_matrix[, 3],
      feature_end_index = notch_matrix[, 4],
      Frequency_kHz = cluster_df$Frequency_kHz[feature_index]
    )
  }
  
  bind_rows(peak_table, notch_table) %>%
    arrange(Frequency_kHz)
}

# ------------------------------------------------------------
# Apply to all clusters
# ------------------------------------------------------------

soldevilla_features_all <- soldevilla_spectra_minmax %>%
  group_by(Soldevilla_cluster) %>%
  group_split() %>%
  map_dfr(pick_features_one_cluster) %>%
  group_by(Soldevilla_cluster, feature_type) %>%
  arrange(Frequency_kHz, .by_group = TRUE) %>%
  mutate(
    feature_number = row_number(),
    feature_label = case_when(
      feature_type == "peak" ~ paste0("Peak_", feature_number),
      feature_type == "notch" ~ paste0("Notch_", feature_number)
    )
  ) %>%
  ungroup()

cat("Detected Soldevilla peaks and notches:\n")
print(soldevilla_features_all, n = Inf)

# ------------------------------------------------------------
# Save full feature table
# ------------------------------------------------------------

write.csv(
  soldevilla_features_all,
  file = file.path(base_output_dir, "Soldevilla_all_detected_peaks_and_notches.csv"),
  row.names = FALSE
)

# ------------------------------------------------------------
# Keep only first 4 peaks and first 3 notches per cluster
# "First" means ordered from low frequency to high frequency.
# ------------------------------------------------------------

soldevilla_features_selected <- soldevilla_features_all %>%
  group_by(Soldevilla_cluster, feature_type) %>%
  arrange(Frequency_kHz, .by_group = TRUE) %>%
  mutate(feature_number = row_number()) %>%
  filter(
    (feature_type == "peak" & feature_number <= 4) |
      (feature_type == "notch" & feature_number <= 3)
  ) %>%
  ungroup()

cat("Selected Soldevilla peaks/notches:\n")
print(soldevilla_features_selected, n = Inf)

write.csv(
  soldevilla_features_selected,
  file = file.path(base_output_dir, "Soldevilla_SELECTED_first4_peaks_first3_notches.csv"),
  row.names = FALSE
)


soldevilla_peak_notch_summary <- soldevilla_features_selected %>%
  mutate(
    feature_column = case_when(
      feature_type == "peak" ~ paste0("Peak_", feature_number, "_kHz"),
      feature_type == "notch" ~ paste0("Notch_", feature_number, "_kHz")
    )
  ) %>%
  select(
    Soldevilla_cluster,
    feature_column,
    Frequency_kHz
  ) %>%
  tidyr::pivot_wider(
    names_from = feature_column,
    values_from = Frequency_kHz
  ) %>%
  arrange(Soldevilla_cluster) %>%
  select(
    Soldevilla_cluster,
    Peak_1_kHz,
    Peak_2_kHz,
    Peak_3_kHz,
    Peak_4_kHz,
    Notch_1_kHz,
    Notch_2_kHz,
    Notch_3_kHz
  )

cat("Wide Soldevilla peak/notch summary table:\n")
print(soldevilla_peak_notch_summary)

write.csv(
  soldevilla_peak_notch_summary,
  file = file.path(base_output_dir, "Soldevilla_peak_notch_summary_first4_first3.csv"),
  row.names = FALSE
)

```

## Generate Visuals

```{r}

# ------------------------------------------------------------
# Output folder for full diagnostic peak/notch figures
# ------------------------------------------------------------

figure_dir_all <- file.path(
  base_output_dir,
  "Cluster Figures - All Detected Features"
)

dir.create(
  figure_dir_all,
  recursive = TRUE,
  showWarnings = FALSE
)

# ------------------------------------------------------------
# Helper function: plot one cluster with ALL detected features
# ------------------------------------------------------------

plot_all_features_one_cluster <- function(cluster_id) {
  
  cluster_spectrum <- soldevilla_spectra_minmax %>%
    filter(Soldevilla_cluster == cluster_id) %>%
    arrange(Frequency_kHz)
  
  cluster_features <- soldevilla_features_all %>%
    filter(Soldevilla_cluster == cluster_id) %>%
    mutate(
      feature_type = factor(
        feature_type,
        levels = c("peak", "notch")
      ),
      label_text = paste0(round(Frequency_kHz, 1), " kHz"),
      label_y = case_when(
        feature_type == "peak" ~ feature_height + 0.06,
        feature_type == "notch" ~ feature_height - 0.06
      ),
      label_vjust = case_when(
        feature_type == "peak" ~ 0,
        feature_type == "notch" ~ 1
      )
    )
  
  p <- ggplot(
    cluster_spectrum,
    aes(x = Frequency_kHz, y = spectrum_minmax)
  ) +
    geom_line(
      color = "grey55",
      linewidth = 1.3
    ) +
    geom_point(
      data = cluster_features,
      aes(
        x = Frequency_kHz,
        y = feature_height,
        shape = feature_type,
        fill = feature_type
      ),
      color = "black",
      size = 3.2,
      stroke = 0.8
    ) +
    geom_text(
      data = cluster_features,
      aes(
        x = Frequency_kHz,
        y = label_y,
        label = label_text,
        vjust = label_vjust
      ),
      size = 3.2,
      color = "black"
    ) +
    scale_shape_manual(
      values = c(
        "peak" = 24,
        "notch" = 25
      )
    ) +
    scale_fill_manual(
      values = c(
        "peak" = "black",
        "notch" = "white"
      )
    ) +
    coord_cartesian(
      ylim = c(-0.12, 1.12),
      xlim = c(freq_min_khz, freq_max_khz),
      clip = "off"
    ) +
    labs(
      title = paste0(
        "All Detected Peaks and Notches: Soldevilla Cluster ",
        cluster_id
      ),
      subtitle = "Within-cluster min-max normalized spectrum; all detected features labeled",
      x = "Frequency (kHz)",
      y = "Min-max normalized amplitude",
      shape = "Feature",
      fill = "Feature"
    ) +
    theme_bw(base_size = 13) +
    theme(
      legend.position = "right",
      plot.title = element_text(face = "bold"),
      plot.margin = margin(10, 30, 10, 10)
    )
  
  print(p)
  
  ggsave(
    filename = file.path(
      figure_dir_all,
      paste0(
        "Soldevilla_Cluster_",
        cluster_id,
        "_ALL_detected_peaks_notches.png"
      )
    ),
    plot = p,
    width = 9,
    height = 5.5,
    dpi = 300
  )
}

# ------------------------------------------------------------
# Generate one figure per Soldevilla cluster
# ------------------------------------------------------------

for (cluster_id in sort(unique(soldevilla_spectra_minmax$Soldevilla_cluster))) {
  
  plot_all_features_one_cluster(cluster_id)
}

cat("Saved all detected peak/notch figures to:\n")
cat(figure_dir_all, "\n")

```

## Try More Permissive Settings (For Cluster 2 Specifically)

```{r}

# ------------------------------------------------------------
# Alternative peak/notch detection: permissive settings
# Keeps the original soldevilla_features_all unchanged.
# ------------------------------------------------------------

soldevilla_spectra_minmax_permissive <- soldevilla_mean_spectra %>%
  group_by(Soldevilla_cluster) %>%
  mutate(
    spectrum_minmax =
      (mean_spectrum_raw - min(mean_spectrum_raw, na.rm = TRUE)) /
      (max(mean_spectrum_raw, na.rm = TRUE) - min(mean_spectrum_raw, na.rm = TRUE))
  ) %>%
  ungroup()

pick_features_one_cluster_permissive <- function(cluster_df) {
  
  cluster_id <- unique(cluster_df$Soldevilla_cluster)
  
  cluster_df <- cluster_df %>%
    arrange(Frequency_kHz)
  
  peak_matrix <- pracma::findpeaks(
    cluster_df$spectrum_minmax,
    nups = 1,
    ndowns = 1,
    minpeakdistance = 3,
    minpeakheight = 0.15
  )
  
  peak_table <- tibble()
  
  if (!is.null(peak_matrix)) {
    peak_table <- tibble(
      Soldevilla_cluster = cluster_id,
      feature_type = "peak",
      feature_height = peak_matrix[, 1],
      feature_index = peak_matrix[, 2],
      feature_start_index = peak_matrix[, 3],
      feature_end_index = peak_matrix[, 4],
      Frequency_kHz = cluster_df$Frequency_kHz[feature_index]
    )
  }
  
  notch_matrix <- pracma::findpeaks(
    -cluster_df$spectrum_minmax,
    nups = 1,
    ndowns = 1,
    minpeakdistance = 3
  )
  
  notch_table <- tibble()
  
  if (!is.null(notch_matrix)) {
    notch_table <- tibble(
      Soldevilla_cluster = cluster_id,
      feature_type = "notch",
      feature_height = -notch_matrix[, 1],
      feature_index = notch_matrix[, 2],
      feature_start_index = notch_matrix[, 3],
      feature_end_index = notch_matrix[, 4],
      Frequency_kHz = cluster_df$Frequency_kHz[feature_index]
    )
  }
  
  bind_rows(peak_table, notch_table) %>%
    arrange(Frequency_kHz)
}

soldevilla_features_all_permissive <- soldevilla_spectra_minmax_permissive %>%
  group_by(Soldevilla_cluster) %>%
  group_split() %>%
  map_dfr(pick_features_one_cluster_permissive) %>%
  group_by(Soldevilla_cluster, feature_type) %>%
  arrange(Frequency_kHz, .by_group = TRUE) %>%
  mutate(
    feature_number = row_number(),
    feature_label = case_when(
      feature_type == "peak" ~ paste0("Peak_", feature_number),
      feature_type == "notch" ~ paste0("Notch_", feature_number)
    )
  ) %>%
  ungroup()

write.csv(
  soldevilla_features_all_permissive,
  file = file.path(base_output_dir, "Soldevilla_all_detected_peaks_and_notches_PERMISSIVE.csv"),
  row.names = FALSE
)

cat("Saved permissive full feature table.\n")

```

## New Visuals

```{r}

# ------------------------------------------------------------
# Output folder for permissive diagnostic figures
# ------------------------------------------------------------

figure_dir_all_permissive <- file.path(
  base_output_dir,
  "Cluster Figures - All Detected Features - Permissive"
)

dir.create(
  figure_dir_all_permissive,
  recursive = TRUE,
  showWarnings = FALSE
)

# ------------------------------------------------------------
# Plot one cluster with ALL permissive detected features
# ------------------------------------------------------------

plot_all_features_one_cluster_permissive <- function(cluster_id) {
  
  cluster_spectrum <- soldevilla_spectra_minmax_permissive %>%
    filter(Soldevilla_cluster == cluster_id) %>%
    arrange(Frequency_kHz)
  
  cluster_features <- soldevilla_features_all_permissive %>%
    filter(Soldevilla_cluster == cluster_id) %>%
    mutate(
      feature_type = factor(feature_type, levels = c("peak", "notch")),
      label_text = paste0(round(Frequency_kHz, 1), " kHz"),
      label_y = case_when(
        feature_type == "peak" ~ feature_height + 0.06,
        feature_type == "notch" ~ feature_height - 0.06
      ),
      label_vjust = case_when(
        feature_type == "peak" ~ 0,
        feature_type == "notch" ~ 1
      )
    )
  
  p <- ggplot(
    cluster_spectrum,
    aes(x = Frequency_kHz, y = spectrum_minmax)
  ) +
    geom_line(color = "grey55", linewidth = 1.3) +
    geom_point(
      data = cluster_features,
      aes(
        x = Frequency_kHz,
        y = feature_height,
        shape = feature_type,
        fill = feature_type
      ),
      color = "black",
      size = 3.2,
      stroke = 0.8
    ) +
    geom_text(
      data = cluster_features,
      aes(
        x = Frequency_kHz,
        y = label_y,
        label = label_text,
        vjust = label_vjust
      ),
      size = 3.2,
      color = "black"
    ) +
    scale_shape_manual(values = c("peak" = 24, "notch" = 25)) +
    scale_fill_manual(values = c("peak" = "black", "notch" = "white")) +
    coord_cartesian(
      ylim = c(-0.12, 1.12),
      xlim = c(freq_min_khz, freq_max_khz),
      clip = "off"
    ) +
    labs(
      title = paste0(
        "Permissive Detected Peaks and Notches: Soldevilla Cluster ",
        cluster_id
      ),
      subtitle = "nups = 1, ndowns = 1; all detected features labeled",
      x = "Frequency (kHz)",
      y = "Min-max normalized amplitude",
      shape = "Feature",
      fill = "Feature"
    ) +
    theme_bw(base_size = 13) +
    theme(
      legend.position = "right",
      plot.title = element_text(face = "bold"),
      plot.margin = margin(10, 30, 10, 10)
    )
  
  print(p)
  
  ggsave(
    filename = file.path(
      figure_dir_all_permissive,
      paste0(
        "Soldevilla_Cluster_",
        cluster_id,
        "_ALL_detected_peaks_notches_PERMISSIVE.png"
      )
    ),
    plot = p,
    width = 9,
    height = 5.5,
    dpi = 300
  )
}

# ------------------------------------------------------------
# Generate figures for all clusters
# ------------------------------------------------------------

for (cluster_id in sort(unique(soldevilla_spectra_minmax_permissive$Soldevilla_cluster))) {
  plot_all_features_one_cluster_permissive(cluster_id)
}

cat("Saved permissive diagnostic figures to:\n")
cat(figure_dir_all_permissive, "\n")

```

## Exploring MY Cluster 6

```{r}

# ------------------------------------------------------------
# Load my mean spectra
# ------------------------------------------------------------

my_cluster_spectra <- read.csv(
  "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/tables/clean_cluster_mean_spectra_18PC_scaledICI_1_minPts50.csv"
)

# ------------------------------------------------------------
# Check column names
# ------------------------------------------------------------

names(my_cluster_spectra)

# ------------------------------------------------------------
# Extract my cluster 6
# ------------------------------------------------------------

my_cluster_6 <- my_cluster_spectra %>%
  filter(HDBSCAN_cluster == 6)

# ------------------------------------------------------------
# Extract Soldevilla cluster 4
# ------------------------------------------------------------

soldevilla_cluster_4 <- soldevilla_spectra_minmax %>%
  filter(Soldevilla_cluster == 4)

# ------------------------------------------------------------
# Plot my cluster 6
# ------------------------------------------------------------

ggplot(
  my_cluster_6,
  aes(
    x = Frequency_kHz,
    y = mean_minmax_spectrum
  )
) +
  geom_line(
    linewidth = 1.2,
    color = "black"
  ) +
  theme_bw() +
  labs(
    title = "My Cluster 6",
    x = "Frequency (kHz)",
    y = "Mean min-max normalized amplitude"
  )

# ------------------------------------------------------------
# Plot Soldevilla cluster 4
# ------------------------------------------------------------

ggplot(
  soldevilla_cluster_4,
  aes(
    x = Frequency_kHz,
    y = spectrum_minmax
  )
) +
  geom_line(
    linewidth = 1.2,
    color = "black"
  ) +
  theme_bw() +
  labs(
    title = "Soldevilla Cluster 4",
    x = "Frequency (kHz)",
    y = "Mean min-max normalized amplitude"
  )

```

## Continued

```{r}

# Try more sensitive peak/notch detection for My Cluster 6 only

my_c6_peak_matrix_sensitive <- pracma::findpeaks(
  my_cluster_6$mean_minmax_spectrum,
  nups = 1,
  ndowns = 1,
  minpeakdistance = 1,
  minpeakheight = 0.15
)

my_c6_notch_matrix_sensitive <- pracma::findpeaks(
  -my_cluster_6$mean_minmax_spectrum,
  nups = 1,
  ndowns = 1,
  minpeakdistance = 1
)

# ------------------------------------------------------------
# Extra-sensitive peak/notch picking for MY Cluster 6 only
# Goal: test possible peak near ~30 kHz and notch near ~31 kHz
# ------------------------------------------------------------

my_c6_peak_matrix_sensitive <- pracma::findpeaks(
  my_cluster_6$mean_minmax_spectrum,
  nups = 1,
  ndowns = 1,
  minpeakdistance = 1,
  minpeakheight = 0.15
)

my_c6_notch_matrix_sensitive <- pracma::findpeaks(
  -my_cluster_6$mean_minmax_spectrum,
  nups = 1,
  ndowns = 1,
  minpeakdistance = 1
)

my_c6_peaks_sensitive <- tibble()

if (!is.null(my_c6_peak_matrix_sensitive)) {
  my_c6_peaks_sensitive <- tibble(
    feature_type = "peak",
    feature_height = my_c6_peak_matrix_sensitive[, 1],
    feature_index = my_c6_peak_matrix_sensitive[, 2],
    Frequency_kHz = my_cluster_6$Frequency_kHz[feature_index]
  )
}

my_c6_notches_sensitive <- tibble()

if (!is.null(my_c6_notch_matrix_sensitive)) {
  my_c6_notches_sensitive <- tibble(
    feature_type = "notch",
    feature_height = -my_c6_notch_matrix_sensitive[, 1],
    feature_index = my_c6_notch_matrix_sensitive[, 2],
    Frequency_kHz = my_cluster_6$Frequency_kHz[feature_index]
  )
}

my_c6_features_sensitive <- bind_rows(
  my_c6_peaks_sensitive,
  my_c6_notches_sensitive
) %>%
  arrange(Frequency_kHz) %>%
  mutate(
    label_text = paste0(round(Frequency_kHz, 1), " kHz"),
    label_y = if_else(
      feature_type == "peak",
      feature_height + 0.07,
      feature_height - 0.07
    ),
    label_vjust = if_else(feature_type == "peak", 0, 1)
  )

print(my_c6_features_sensitive, n = Inf)

p_my_c6_sensitive <- ggplot(
  my_cluster_6,
  aes(x = Frequency_kHz, y = mean_minmax_spectrum)
) +
  geom_line(color = "steelblue", linewidth = 1.5) +
  geom_point(
    data = my_c6_features_sensitive,
    aes(
      x = Frequency_kHz,
      y = feature_height,
      shape = feature_type,
      fill = feature_type
    ),
    color = "black",
    size = 3.4,
    stroke = 0.8
  ) +
  geom_text(
    data = my_c6_features_sensitive,
    aes(
      x = Frequency_kHz,
      y = label_y,
      label = label_text,
      vjust = label_vjust
    ),
    size = 3.2
  ) +
  scale_shape_manual(values = c("peak" = 24, "notch" = 25)) +
  scale_fill_manual(values = c("peak" = "black", "notch" = "white")) +
  coord_cartesian(
    xlim = c(19, 50),
    ylim = c(-0.05, 1.05),
    clip = "off"
  ) +
  labs(
    title = "Extra-Sensitive Peak/Notch Picking: My Cluster 6",
    subtitle = "nups = 1, ndowns = 1, minpeakdistance = 1",
    x = "Frequency (kHz)",
    y = "Mean min-max normalized amplitude",
    shape = "Feature",
    fill = "Feature"
  ) +
  theme_bw(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "right"
  )

print(p_my_c6_sensitive)

```

## New peak picking

```{r}

# ------------------------------------------------------------
# Permissive peak/notch picking for all MY clusters
#
# Settings chosen after Soldevilla comparison:
#   nups = 1
#   ndowns = 1
#   minpeakdistance = 1
#   minpeakheight = 0.15
# ------------------------------------------------------------

my_permissive_output_dir <- file.path(
  "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/plots",
  "Peak_Notch_Picking_Permissive_After_Soldevilla_Comparison"
)

dir.create(
  my_permissive_output_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

# ------------------------------------------------------------
# Load my cluster mean spectra
# ------------------------------------------------------------

my_cluster_spectra <- read.csv(
  "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/tables/clean_cluster_mean_spectra_18PC_scaledICI_1_minPts50.csv",
  stringsAsFactors = FALSE
)

# ------------------------------------------------------------
# Helper function for one cluster
# ------------------------------------------------------------

plot_my_cluster_permissive <- function(cluster_id) {
  
  cluster_df <- my_cluster_spectra %>%
    mutate(HDBSCAN_cluster = as.character(HDBSCAN_cluster)) %>%
    filter(HDBSCAN_cluster == as.character(cluster_id)) %>%
    arrange(Frequency_kHz)
  
  if (nrow(cluster_df) == 0) {
    warning("No rows found for cluster ", cluster_id)
    return(NULL)
  }
  
  peak_matrix <- pracma::findpeaks(
    cluster_df$mean_minmax_spectrum,
    nups = 1,
    ndowns = 1,
    minpeakdistance = 1,
    minpeakheight = 0.15
  )
  
  notch_matrix <- pracma::findpeaks(
    -cluster_df$mean_minmax_spectrum,
    nups = 1,
    ndowns = 1,
    minpeakdistance = 1
  )
  
  peak_table <- tibble()
  
  if (!is.null(peak_matrix)) {
    peak_table <- tibble(
      HDBSCAN_cluster = cluster_id,
      feature_type = "peak",
      feature_height = peak_matrix[, 1],
      feature_index = peak_matrix[, 2],
      feature_start_index = peak_matrix[, 3],
      feature_end_index = peak_matrix[, 4],
      Frequency_kHz = cluster_df$Frequency_kHz[feature_index]
    )
  }
  
  notch_table <- tibble()
  
  if (!is.null(notch_matrix)) {
    notch_table <- tibble(
      HDBSCAN_cluster = cluster_id,
      feature_type = "notch",
      feature_height = -notch_matrix[, 1],
      feature_index = notch_matrix[, 2],
      feature_start_index = notch_matrix[, 3],
      feature_end_index = notch_matrix[, 4],
      Frequency_kHz = cluster_df$Frequency_kHz[feature_index]
    )
  }
  
  feature_table <- bind_rows(peak_table, notch_table) %>%
    arrange(Frequency_kHz) %>%
    group_by(feature_type) %>%
    mutate(
      feature_number = row_number(),
      feature_label = case_when(
        feature_type == "peak" ~ paste0("Peak_", feature_number),
        feature_type == "notch" ~ paste0("Notch_", feature_number)
      ),
      label_text = paste0(round(Frequency_kHz, 1), " kHz"),
      label_y = if_else(
        feature_type == "peak",
        feature_height + 0.06,
        feature_height - 0.06
      ),
      label_vjust = if_else(feature_type == "peak", 0, 1)
    ) %>%
    ungroup()
  
  p <- ggplot(
    cluster_df,
    aes(x = Frequency_kHz, y = mean_minmax_spectrum)
  ) +
    geom_line(color = "steelblue", linewidth = 1.5) +
    geom_point(
      data = feature_table,
      aes(
        x = Frequency_kHz,
        y = feature_height,
        shape = feature_type,
        fill = feature_type
      ),
      color = "black",
      size = 3.4,
      stroke = 0.8
    ) +
    geom_text(
      data = feature_table,
      aes(
        x = Frequency_kHz,
        y = label_y,
        label = label_text,
        vjust = label_vjust
      ),
      size = 3.2
    ) +
    scale_shape_manual(values = c("peak" = 24, "notch" = 25)) +
    scale_fill_manual(values = c("peak" = "black", "notch" = "white")) +
    coord_cartesian(
      xlim = c(19, 50),
      ylim = c(-0.05, 1.05),
      clip = "off"
    ) +
    labs(
      title = paste0("Permissive Picked Peaks and Notches: My Cluster ", cluster_id),
      subtitle = "After Soldevilla comparison; nups = 1, ndowns = 1, minpeakdistance = 1",
      x = "Frequency (kHz)",
      y = "Mean min-max normalized amplitude",
      shape = "Feature",
      fill = "Feature"
    ) +
    theme_bw(base_size = 13) +
    theme(
      plot.title = element_text(face = "bold"),
      legend.position = "right",
      plot.margin = margin(10, 35, 10, 10)
    )
  
  print(p)
  
  ggsave(
    filename = file.path(
      my_permissive_output_dir,
      paste0("My_Cluster_", cluster_id, "_permissive_peaks_notches_after_soldevilla.png")
    ),
    plot = p,
    width = 10,
    height = 6,
    dpi = 300
  )
  
  feature_table
}

# ------------------------------------------------------------
# Run for all observed non-noise clusters
# ------------------------------------------------------------

my_clusters_to_plot <- sort(
  as.numeric(
    unique(as.character(my_cluster_spectra$HDBSCAN_cluster))
  )
)

my_clusters_to_plot <- my_clusters_to_plot[my_clusters_to_plot != 0]

my_permissive_features_all <- purrr::map_dfr(
  my_clusters_to_plot,
  plot_my_cluster_permissive
)

# ------------------------------------------------------------
# Save full permissive feature table
# ------------------------------------------------------------

write.csv(
  my_permissive_features_all,
  file = file.path(
    my_permissive_output_dir,
    "My_clusters_all_permissive_peaks_notches_after_soldevilla.csv"
  ),
  row.names = FALSE
)

cat("Saved permissive cluster figures and table to:\n")
cat(my_permissive_output_dir, "\n")

# Save all detected permissive peaks/notches
write.csv(
  my_permissive_features_all,
  file = file.path(
    my_permissive_output_dir,
    "My_clusters_ALL_permissive_peaks_notches_after_soldevilla.csv"
  ),
  row.names = FALSE
)

# Keep first 4 peaks and first 3 notches per cluster
my_permissive_features_selected <- my_permissive_features_all %>%
  group_by(HDBSCAN_cluster, feature_type) %>%
  arrange(Frequency_kHz, .by_group = TRUE) %>%
  mutate(feature_number = row_number()) %>%
  filter(
    (feature_type == "peak" & feature_number <= 4) |
      (feature_type == "notch" & feature_number <= 3)
  ) %>%
  ungroup()

write.csv(
  my_permissive_features_selected,
  file = file.path(
    my_permissive_output_dir,
    "My_clusters_SELECTED_first4_peaks_first3_notches_after_soldevilla.csv"
  ),
  row.names = FALSE
)

# Optional wide summary table
my_permissive_peak_notch_summary <- my_permissive_features_selected %>%
  mutate(
    feature_column = case_when(
      feature_type == "peak" ~ paste0("Peak_", feature_number, "_kHz"),
      feature_type == "notch" ~ paste0("Notch_", feature_number, "_kHz")
    )
  ) %>%
  select(HDBSCAN_cluster, feature_column, Frequency_kHz) %>%
  tidyr::pivot_wider(
    names_from = feature_column,
    values_from = Frequency_kHz
  ) %>%
  arrange(as.numeric(HDBSCAN_cluster))

write.csv(
  my_permissive_peak_notch_summary,
  file = file.path(
    my_permissive_output_dir,
    "My_clusters_WIDE_first4_peaks_first3_notches_after_soldevilla.csv"
  ),
  row.names = FALSE
)

cat("Saved all, selected, and wide peak/notch CSVs to:\n")
cat(my_permissive_output_dir, "\n")

```



































