---
title: "Soldevilla_Style_Peak_Picking"
author: "Sarah Moseley"
date: "2026-05-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

This script follows a Soldevilla-style workflow for identifying consistent spectral peak bands in Risso's dolphin click spectra.

The workflow is:

1. Load combined Tukey spectra.
2. Check and interpolate spectra to a shared frequency grid.
3. Restrict spectra to 19–50 kHz.
4. Randomly sample up to 200 clicks per event.
5. Peak-pick individual click spectra within each event.
6. Build per-event histograms of detected peak frequencies.
7. Identify consistent event-level peak bands.

## Setup and User Settings

```{r}
library(ggplot2)
library(dplyr)

# -----------------------------
# Base Output Directory
# -----------------------------
base_output_dir <- "E:/Analysis/Peak_Picking_Soldevilla_Style"
dir.create(base_output_dir, recursive = TRUE, showWarnings = FALSE)

# -----------------------------
# Analysis Label
# -----------------------------
analysis_name <- "Soldevilla Style Peak Picking"

# -----------------------------
# Frequency Settings
# -----------------------------
freq_min <- 19000
freq_max <- 50000

# -----------------------------
# Sampling Settings
# -----------------------------
max_clicks_per_event <- 200
sampling_seed <- 123

# -----------------------------
# Histogram Settings
# -----------------------------
hist_bin_width <- 400
```

## Load Combined Tukey Spectra

```{r}

combined_file <- "E:/Analysis/TukeySpectra/Combined_Tukey_Spectra.rds"

combined_tukey <- readRDS(combined_file)

class(combined_tukey)
length(combined_tukey)

```

## Inspect Combined Tukey Structure

```{r}

names(combined_tukey[[1]])

combined_tukey[[1]]$EventID
length(combined_tukey[[1]]$freq)
dim(combined_tukey[[1]]$allSpec)
length(combined_tukey[[1]]$UID)

```

## Sanity Checks for Combined Tukey Structure

```{r}

event_structure_summary <- data.frame(
  EventIndex = seq_along(combined_tukey),
  EventID = sapply(combined_tukey, function(x) x$EventID),
  n_freq = sapply(combined_tukey, function(x) length(x$freq)),
  n_uid = sapply(combined_tukey, function(x) length(x$UID)),
  allSpec_rows = sapply(combined_tukey, function(x) dim(x$allSpec)[1]),
  allSpec_cols = sapply(combined_tukey, function(x) dim(x$allSpec)[2]),
  stringsAsFactors = FALSE
)

print(event_structure_summary)

uid_matches_allSpec <- all(
  event_structure_summary$n_uid == event_structure_summary$allSpec_cols
)

freq_matches_allSpec <- all(
  event_structure_summary$n_freq == event_structure_summary$allSpec_rows
)

cat("\nUID count matches allSpec columns:\n")
print(uid_matches_allSpec)

cat("\nFrequency count matches allSpec rows:\n")
print(freq_matches_allSpec)

if (uid_matches_allSpec && freq_matches_allSpec) {
  cat("\nPASS: Core spectral dimensions are aligned.\n")
} else {
  stop("FAIL: Spectral dimensions are misaligned.")
}

```

## Inspect Frequency Grids Before Interpolation

```{r}

freq_grid_summary <- data.frame(
  EventIndex = seq_along(combined_tukey),
  EventID = sapply(combined_tukey, function(x) x$EventID),
  n_freq = sapply(combined_tukey, function(x) length(x$freq)),
  min_freq = sapply(combined_tukey, function(x) min(x$freq, na.rm = TRUE)),
  max_freq = sapply(combined_tukey, function(x) max(x$freq, na.rm = TRUE)),
  median_freq_step = sapply(combined_tukey, function(x) median(diff(x$freq), na.rm = TRUE)),
  stringsAsFactors = FALSE
)

print(freq_grid_summary)

# ------------------------------------------------------------
# Check whether all frequency grids are identical
# ------------------------------------------------------------

reference_freq_original <- combined_tukey[[1]]$freq

all_freq_grids_identical <- all(
  sapply(
    combined_tukey,
    function(x) identical(x$freq, reference_freq_original)
  )
)

cat("\nAre all original frequency grids identical?\n")
print(all_freq_grids_identical)

```

## Interpolate Spectra to Shared Grid

```{r}

reference_freq <- combined_tukey[[1]]$freq

interpolate_spectrum <- function(old_freq, old_spec, new_freq) {
  approx(
    x = old_freq,
    y = old_spec,
    xout = new_freq,
    method = "linear",
    rule = 2
  )$y
}

interpolated_tukey <- vector("list", length(combined_tukey))

for (i in seq_along(combined_tukey)) {
  
  tukey_spectra <- combined_tukey[[i]]
  
  old_freq <- tukey_spectra$freq
  old_allSpec <- tukey_spectra$allSpec
  
  interpolated_allSpec <- apply(
    old_allSpec,
    2,
    function(spec) interpolate_spectrum(old_freq, spec, reference_freq)
  )
  
  tukey_spectra$allSpec <- as.matrix(interpolated_allSpec)
  tukey_spectra$freq <- reference_freq
  
  interpolated_tukey[[i]] <- tukey_spectra
}

all_match_after_interp <- all(
  sapply(interpolated_tukey, function(x) identical(x$freq, reference_freq))
)

all_match_after_interp

```

##Restrict Spectra to 19–50 kHz

```{r}

freq_min <- 19000
freq_max <- 50000

keepFreq <- which(reference_freq >= freq_min & reference_freq <= freq_max)

freq_focus <- reference_freq[keepFreq]

restricted_tukey <- vector("list", length(interpolated_tukey))

for (i in seq_along(interpolated_tukey)) {
  
  tukey_spectra <- interpolated_tukey[[i]]
  
  tukey_spectra$freq_focus <- freq_focus
  tukey_spectra$allSpec_focus <- tukey_spectra$allSpec[keepFreq, , drop = FALSE]
  
  restricted_tukey[[i]] <- tukey_spectra
}

cat("Number of retained frequency bins:", length(freq_focus), "\n")
cat("Frequency range:", min(freq_focus), "to", max(freq_focus), "Hz\n")

restricted_check <- data.frame(
  EventID = sapply(restricted_tukey, function(x) x$EventID),
  NumClicks = sapply(restricted_tukey, function(x) ncol(x$allSpec_focus)),
  NumFreqBins = sapply(restricted_tukey, function(x) nrow(x$allSpec_focus)),
  MinFreq = sapply(restricted_tukey, function(x) min(x$freq_focus)),
  MaxFreq = sapply(restricted_tukey, function(x) max(x$freq_focus)),
  stringsAsFactors = FALSE
)

print(restricted_check)

cat("\nAll events same number of retained bins:\n")
print(length(unique(restricted_check$NumFreqBins)) == 1)

cat("\nAll events same retained frequency range:\n")
print(
  length(unique(restricted_check$MinFreq)) == 1 &&
    length(unique(restricted_check$MaxFreq)) == 1
)

```

## Sample Up To 200 Clicks Per Event

```{r}

max_clicks_per_event <- 200
sampling_seed <- 123

if (!is.null(sampling_seed)) set.seed(sampling_seed)

sampled_tukey <- vector("list", length(restricted_tukey))

for (i in seq_along(restricted_tukey)) {
  
  tukey_spectra <- restricted_tukey[[i]]
  
  n_clicks <- ncol(tukey_spectra$allSpec_focus)
  
  sampled_idx <- sample(
    seq_len(n_clicks),
    size = min(n_clicks, max_clicks_per_event),
    replace = FALSE
  )
  
  tukey_spectra$sampled_click_index <- sampled_idx
  tukey_spectra$UID_sampled <- tukey_spectra$UID[sampled_idx]
  tukey_spectra$allSpec_sampled <- tukey_spectra$allSpec_focus[, sampled_idx, drop = FALSE]
  
  sampled_tukey[[i]] <- tukey_spectra
}

## Sanity Check Sampled Clicks

sample_check <- data.frame(
  EventID = sapply(sampled_tukey, function(x) x$EventID),
  OriginalClicks = sapply(sampled_tukey, function(x) ncol(x$allSpec_focus)),
  SampledClicks = sapply(sampled_tukey, function(x) ncol(x$allSpec_sampled)),
  NumFreqBins = sapply(sampled_tukey, function(x) nrow(x$allSpec_sampled)),
  stringsAsFactors = FALSE
)

print(sample_check)

cat("\nMax sampled clicks per event:", max(sample_check$SampledClicks), "\n")
cat("Total sampled clicks:", sum(sample_check$SampledClicks), "\n")
cat("All sampled events have 83 frequency bins:", all(sample_check$NumFreqBins == 83), "\n")

```

## Build Sampled Click Metadata Table

```{r}

sampled_click_metadata <- data.frame(
  EventIndex = integer(),
  EventID = character(),
  UID = character(),
  ClickIndexWithinEvent = integer(),
  SampleIndexWithinEvent = integer(),
  stringsAsFactors = FALSE
)

for (i in seq_along(sampled_tukey)) {
  
  tukey_spectra <- sampled_tukey[[i]]
  
  event_meta <- data.frame(
    EventIndex = i,
    EventID = tukey_spectra$EventID,
    UID = as.character(tukey_spectra$UID_sampled),
    ClickIndexWithinEvent = tukey_spectra$sampled_click_index,
    SampleIndexWithinEvent = seq_along(tukey_spectra$sampled_click_index),
    stringsAsFactors = FALSE
  )
  
  sampled_click_metadata <- rbind(sampled_click_metadata, event_meta)
}

head(sampled_click_metadata)
dim(sampled_click_metadata)

```

## Peak-Picking Function for One Click Spectrum

```{r}

## Peak-Picking Parameters

# Your spectra are already 5-point Tukey-smoothed, so we do not smooth again.

min_peak_separation_khz <- 1.5

# This is not exactly Soldevilla's derivative threshold,
# but it serves the same purpose: avoid tiny local wiggles.
min_prominence_db <- 2

# Keep this low / inactive for now because spectra are in negative dB.
peak_height_quantile <- NULL

```

## Peak-Picking Function

```{r}

pick_peaks_one_spectrum <- function(
    freq_khz,
    spec,
    min_peak_separation_khz = 1.5,
    min_prominence_db = 2
) {
  
  # ------------------------------------------------------------
  # Step 1: Find local maxima
  # ------------------------------------------------------------
  
  candidate_idx <- which(
    spec[2:(length(spec) - 1)] > spec[1:(length(spec) - 2)] &
    spec[2:(length(spec) - 1)] > spec[3:length(spec)]
  ) + 1
  
  if (length(candidate_idx) == 0) {
    return(data.frame())
  }
  
  # ------------------------------------------------------------
  # Step 2: Calculate local prominence
  # ------------------------------------------------------------
  
  prominence <- sapply(candidate_idx, function(idx) {
    
    left_diff  <- spec[idx] - spec[idx - 1]
    right_diff <- spec[idx] - spec[idx + 1]
    
    min(left_diff, right_diff)
  })
  
  keep_idx <- candidate_idx[prominence >= min_prominence_db]
  
  if (length(keep_idx) == 0) {
    return(data.frame())
  }
  
  # ------------------------------------------------------------
  # Step 3: Enforce minimum frequency separation
  # ------------------------------------------------------------
  
  keep_freq <- freq_khz[keep_idx]
  keep_amp  <- spec[keep_idx]
  
  peak_table <- data.frame(
    PeakIndex = keep_idx,
    PeakFrequency_kHz = keep_freq,
    PeakAmplitude = keep_amp
  )
  
  peak_table <- peak_table[order(-peak_table$PeakAmplitude), ]
  
  selected_peaks <- data.frame()
  
  while (nrow(peak_table) > 0) {
    
    best_peak <- peak_table[1, ]
    
    selected_peaks <- rbind(selected_peaks, best_peak)
    
    peak_table <- peak_table[
      abs(peak_table$PeakFrequency_kHz - best_peak$PeakFrequency_kHz) >=
        min_peak_separation_khz,
    ]
  }
  
  selected_peaks <- selected_peaks[
    order(selected_peaks$PeakFrequency_kHz),
  ]
  
  rownames(selected_peaks) <- NULL
  
  return(selected_peaks)
}

```

## ## Test Peak Picking

```{r}

test_peaks <- pick_peaks_one_spectrum(
  freq_khz = test_freq_khz,
  spec = test_spec,
  min_peak_separation_khz = 1.5,
  min_prominence_db = 0.5
)

test_peaks

plot(
  test_freq_khz,
  test_spec,
  type = "l",
  xlab = "Frequency (kHz)",
  ylab = "Spectral amplitude",
  main = paste("Peak picking:", test_event$EventID)
)

points(
  test_peaks$PeakFrequency_kHz,
  test_peaks$PeakAmplitude,
  col = "red",
  pch = 19
)

```































