---
title: "Temporal_ClickType_Visualization_SampledClicks"
author: "Sarah Moseley"
date: "2026-05-21"
output: html_document
---

## Overall Tempral Click-Type

```{r}

library(dplyr)
library(ggplot2)
library(readr)

# ------------------------------------------------------------
# Input: sampled click-level HDBSCAN assignments
# ------------------------------------------------------------

assignments_file <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/tables/Usable tables/FULL_hdbscan_assignments_18PC_scaledICI_1_minPts50.csv"

plot_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/plots/temporal_cluster_plots"

dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)

clicks <- read_csv(assignments_file, show_col_types = FALSE)

# ------------------------------------------------------------
# Cluster colors used throughout workflow
# ------------------------------------------------------------

cluster_colors <- c(
  "0" = "grey80",
  "1" = "#E41A1C",
  "2" = "#FF7F00",
  "3" = "#FFD92F",
  "4" = "#4DAF4A",
  "5" = "#00BFC4",
  "6" = "#377EB8",
  "7" = "#984EA3",
  "8" = "#F781BF",
  "9" = "#A65628"
)

# ------------------------------------------------------------
# Build relative time within each event
# ------------------------------------------------------------

clicks_temporal <- clicks %>%
  mutate(
    HDBSCAN_cluster = factor(
      as.character(HDBSCAN_cluster),
      levels = as.character(0:9)
    ),
    Final_UTC_ms = as.numeric(Final_UTC_ms)
  ) %>%
  group_by(EventID) %>%
  arrange(Final_UTC_ms, .by_group = TRUE) %>%
  mutate(
    seconds_since_event_start = (Final_UTC_ms - min(Final_UTC_ms, na.rm = TRUE)) / 1000
  ) %>%
  ungroup()

# ------------------------------------------------------------
# Plot: when each click type appears in each event
# ------------------------------------------------------------

p_temporal_clusters <- ggplot(
  clicks_temporal,
  aes(
    x = seconds_since_event_start,
    y = HDBSCAN_cluster,
    color = HDBSCAN_cluster
  )
) +
  geom_point(
    size = 0.8,
    alpha = 0.75,
    position = position_jitter(height = 0.12, width = 0)
  ) +
  facet_wrap(~ EventID, scales = "free_x") +
  scale_color_manual(values = cluster_colors, drop = FALSE) +
  labs(
    title = "Temporal Occurrence of HDBSCAN Click Types by Event",
    subtitle = "Sampled clicks only; 18PC + 1x scaled logICI; minPts = 50",
    x = "Seconds since event start",
    y = "HDBSCAN cluster",
    color = "Cluster"
  ) +
  theme_bw(base_size = 12) +
  theme(
    strip.text = element_text(size = 8, face = "bold"),
    legend.position = "bottom"
  )

print(p_temporal_clusters)

ggsave(
  filename = file.path(plot_dir, "temporal_click_types_by_event_sampled_clicks.png"),
  plot = p_temporal_clusters,
  width = 16,
  height = 12,
  dpi = 300
)

```

## Bearing source audit: targeted folder scan

```{r}

library(dplyr)
library(stringr)
library(readr)
library(purrr)
library(tibble)

search_roots <- c(
  "E:/Analysis/Useable Databases",
  "E:/Analysis/CleanAcousticStudies",
  "E:/Analysis/Binaries"
)

search_roots <- search_roots[dir.exists(search_roots)]

cat("Searching these folders:\n")
print(search_roots)

all_files <- unlist(
  lapply(search_roots, function(root) {
    list.files(root, recursive = TRUE, full.names = TRUE)
  })
)

file_inventory <- tibble(
  file = all_files,
  folder = dirname(all_files),
  filename = basename(all_files),
  extension = tolower(tools::file_ext(all_files)),
  size_mb = round(file.info(all_files)$size / 1024^2, 3),
  modified = file.info(all_files)$mtime
)

candidate_files <- file_inventory %>%
  filter(
    str_detect(
      filename,
      regex(
        "bearing|angle|local|loc|click|det|pam|target|motion|hydro|array|event|database|sqlite|db",
        ignore_case = TRUE
      )
    ) |
      extension %in% c("sqlite", "sqlite3", "db", "database", "csv", "txt", "tsv")
  ) %>%
  arrange(extension, filename)

print(candidate_files, n = Inf)

audit_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/temporal_visualizations/bearing_source_audit"

dir.create(audit_dir, recursive = TRUE, showWarnings = FALSE)

write_csv(
  candidate_files,
  file.path(audit_dir, "candidate_bearing_source_files.csv")
)

cat("\nSaved candidate file inventory to:\n")
cat(file.path(audit_dir, "candidate_bearing_source_files.csv"), "\n")

```

## Inspect SQLite database schemas for bearing / angle / localization fields

```{r}

library(dplyr)
library(purrr)
library(stringr)
library(tibble)
library(readr)
library(DBI)
library(RSQLite)

# ------------------------------------------------------------
# Search only the clean/useable databases first
# ------------------------------------------------------------

db_dir <- "E:/Analysis/Useable Databases"

db_files <- list.files(
  db_dir,
  pattern = "\\.sqlite3$",
  full.names = TRUE
)

cat("SQLite databases found:", length(db_files), "\n")
print(basename(db_files))

# ------------------------------------------------------------
# Patterns we care about
# ------------------------------------------------------------

bearing_pattern <- regex(
  "bearing|bear|angle|theta|azimuth|azi|deg|degree|degrees",
  ignore_case = TRUE
)

localization_pattern <- regex(
  "locali|loc|target|motion|hydro|array|heading|direction",
  ignore_case = TRUE
)

time_pattern <- regex(
  "time|utc|date|millis|millisecond|msec|timestamp",
  ignore_case = TRUE
)

id_pattern <- regex(
  "uid|event|click|databaseindex|id",
  ignore_case = TRUE
)

# ------------------------------------------------------------
# Helper: safely count rows in a table
# ------------------------------------------------------------

safe_count_rows <- function(con, table_name) {
  out <- tryCatch({
    qtbl <- DBI::dbQuoteIdentifier(con, table_name)
    DBI::dbGetQuery(con, paste0("SELECT COUNT(*) AS n FROM ", qtbl))$n[1]
  }, error = function(e) {
    NA_integer_
  })
  
  out
}

# ------------------------------------------------------------
# Helper: scan one database
# ------------------------------------------------------------

scan_one_db_schema <- function(db_path) {
  
  con <- DBI::dbConnect(RSQLite::SQLite(), db_path)
  on.exit(DBI::dbDisconnect(con), add = TRUE)
  
  tables <- DBI::dbListTables(con)
  
  map_dfr(tables, function(tbl) {
    
    fields <- tryCatch(
      DBI::dbListFields(con, tbl),
      error = function(e) character(0)
    )
    
    n_rows <- safe_count_rows(con, tbl)
    
    tibble(
      database_file = db_path,
      database_name = basename(db_path),
      table_name = tbl,
      n_rows = n_rows,
      column_name = fields
    )
  })
}

# ------------------------------------------------------------
# Scan all database schemas
# ------------------------------------------------------------

db_schema_inventory <- map_dfr(db_files, scan_one_db_schema) %>%
  mutate(
    bearing_match = str_detect(column_name, bearing_pattern),
    localization_match = str_detect(column_name, localization_pattern),
    time_match = str_detect(column_name, time_pattern),
    id_match = str_detect(column_name, id_pattern),
    any_relevant_match = bearing_match | localization_match | time_match | id_match
  )

# ------------------------------------------------------------
# Table-level summary: which tables look promising?
# ------------------------------------------------------------

promising_tables <- db_schema_inventory %>%
  group_by(database_name, database_file, table_name, n_rows) %>%
  summarise(
    bearing_columns = paste(column_name[bearing_match], collapse = " | "),
    localization_columns = paste(column_name[localization_match], collapse = " | "),
    time_columns = paste(column_name[time_match], collapse = " | "),
    id_columns = paste(column_name[id_match], collapse = " | "),
    n_bearing_cols = sum(bearing_match),
    n_localization_cols = sum(localization_match),
    n_time_cols = sum(time_match),
    n_id_cols = sum(id_match),
    all_columns = paste(column_name, collapse = " | "),
    .groups = "drop"
  ) %>%
  filter(
    n_bearing_cols > 0 |
      n_localization_cols > 0 |
      n_time_cols > 0 |
      n_id_cols > 0
  ) %>%
  arrange(
    desc(n_bearing_cols),
    desc(n_localization_cols),
    desc(n_time_cols),
    desc(n_id_cols),
    database_name,
    table_name
  )

cat("\nPromising tables:\n")
print(promising_tables, n = Inf)

# ------------------------------------------------------------
# Strong candidates: tables with possible bearing/angle columns
# ------------------------------------------------------------

bearing_candidate_tables <- promising_tables %>%
  filter(n_bearing_cols > 0 | n_localization_cols > 0)

cat("\nStrong bearing/localization candidate tables:\n")
print(bearing_candidate_tables, n = Inf)

# ------------------------------------------------------------
# Save outputs
# ------------------------------------------------------------

audit_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/temporal_visualizations/bearing_source_audit"

dir.create(audit_dir, recursive = TRUE, showWarnings = FALSE)

write_csv(
  db_schema_inventory,
  file.path(audit_dir, "sqlite_schema_inventory_all_columns.csv")
)

write_csv(
  promising_tables,
  file.path(audit_dir, "sqlite_promising_tables_uid_time_bearing.csv")
)

write_csv(
  bearing_candidate_tables,
  file.path(audit_dir, "sqlite_bearing_localization_candidate_tables.csv")
)

cat("\nSaved schema audit files to:\n")
cat(audit_dir, "\n")

```

## Join sampled HDBSCAN clicks to Click_Detector_OfflineClicks
## Goal: confirm database UID match and identify relevant PAMGuard binary files

```{r}

library(dplyr)
library(readr)
library(purrr)
library(stringr)
library(tibble)
library(DBI)
library(RSQLite)

# ------------------------------------------------------------
# Inputs
# ------------------------------------------------------------

assignments_file <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/tables/Usable tables/FULL_hdbscan_assignments_18PC_scaledICI_1_minPts50.csv"

db_dir <- "E:/Analysis/Useable Databases"

audit_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/temporal_visualizations/bearing_source_audit"

dir.create(audit_dir, recursive = TRUE, showWarnings = FALSE)

# ------------------------------------------------------------
# Load sampled labeled clicks
# ------------------------------------------------------------

sampled_clicks <- read_csv(assignments_file, show_col_types = FALSE) %>%
  mutate(
    UID = as.character(UID),
    DriftID_clean = DriftID_spectra
  )

cat("Sampled labeled clicks:", nrow(sampled_clicks), "\n")
cat("Unique drifts:", n_distinct(sampled_clicks$DriftID_clean), "\n\n")

# ------------------------------------------------------------
# Get database files
# Database names look like: ADRIFT_003 - Clean.sqlite3
# Drift IDs look like: ADRIFT_003
# ------------------------------------------------------------

db_files <- list.files(
  db_dir,
  pattern = "\\.sqlite3$",
  full.names = TRUE
)

db_lookup <- tibble(
  database_file = db_files,
  database_name = basename(db_files),
  DriftID_clean = str_remove(basename(db_files), " - Clean\\.sqlite3$")
)

print(db_lookup)

# ------------------------------------------------------------
# Helper: read Click_Detector_OfflineClicks from one database
# ------------------------------------------------------------

read_offline_clicks <- function(database_file, DriftID_clean) {
  
  con <- dbConnect(SQLite(), database_file)
  on.exit(dbDisconnect(con), add = TRUE)
  
  tables <- dbListTables(con)
  
  if (!"Click_Detector_OfflineClicks" %in% tables) {
    warning("No Click_Detector_OfflineClicks table in: ", basename(database_file))
    return(tibble())
  }
  
  available_cols <- dbListFields(con, "Click_Detector_OfflineClicks")
  
  wanted_cols <- c(
    "Id",
    "UID",
    "UTC",
    "UTCMilliseconds",
    "PCLocalTime",
    "PCTime",
    "parentID",
    "parentUID",
    "LongDataName",
    "BinaryFile",
    "EventId",
    "ClickNo",
    "Amplitude",
    "Channels"
  )
  
  cols_to_pull <- intersect(wanted_cols, available_cols)
  
  query <- paste0(
    "SELECT ",
    paste(DBI::dbQuoteIdentifier(con, cols_to_pull), collapse = ", "),
    " FROM ",
    DBI::dbQuoteIdentifier(con, "Click_Detector_OfflineClicks")
  )
  
  out <- dbGetQuery(con, query) %>%
    as_tibble() %>%
    mutate(
      UID = as.character(UID),
      DriftID_clean = DriftID_clean,
      database_name = basename(database_file),
      database_file = database_file
    )
  
  return(out)
}

# ------------------------------------------------------------
# Read offline click metadata from all usable databases
# ------------------------------------------------------------

offline_clicks_all <- pmap_dfr(
  db_lookup,
  function(database_file, database_name, DriftID_clean) {
    cat("Reading:", database_name, "\n")
    read_offline_clicks(database_file, DriftID_clean)
  }
)

cat("\nOffline click rows read:", nrow(offline_clicks_all), "\n")
cat("Unique UIDs in offline clicks:", n_distinct(offline_clicks_all$UID), "\n\n")

# ------------------------------------------------------------
# Join sampled labels to offline click metadata
# ------------------------------------------------------------

sampled_with_offline_clicks <- sampled_clicks %>%
  left_join(
    offline_clicks_all,
    by = c("DriftID_clean", "UID"),
    suffix = c("_sampled", "_db")
  )

join_summary <- sampled_with_offline_clicks %>%
  summarise(
    n_sampled_clicks = n(),
    n_joined_to_offline_clicks = sum(!is.na(BinaryFile)),
    n_missing_offline_click_match = sum(is.na(BinaryFile)),
    percent_joined = 100 * n_joined_to_offline_clicks / n_sampled_clicks
  )

cat("\nJoin summary:\n")
print(join_summary)

# ------------------------------------------------------------
# Check join by event
# ------------------------------------------------------------

join_summary_by_event <- sampled_with_offline_clicks %>%
  group_by(DriftID_clean, EventID) %>%
  summarise(
    n_sampled_clicks = n(),
    n_joined = sum(!is.na(BinaryFile)),
    n_missing = sum(is.na(BinaryFile)),
    percent_joined = 100 * n_joined / n_sampled_clicks,
    .groups = "drop"
  ) %>%
  arrange(percent_joined, DriftID_clean, EventID)

cat("\nJoin summary by event:\n")
print(join_summary_by_event, n = Inf)

# ------------------------------------------------------------
# Which binary files contain the sampled clicks?
# ------------------------------------------------------------

sampled_binary_files <- sampled_with_offline_clicks %>%
  filter(!is.na(BinaryFile)) %>%
  count(DriftID_clean, database_name, BinaryFile, name = "n_sampled_clicks") %>%
  arrange(DriftID_clean, desc(n_sampled_clicks))

cat("\nBinary files represented in sampled clicks:\n")
print(sampled_binary_files, n = Inf)

# ------------------------------------------------------------
# Save outputs
# ------------------------------------------------------------

write_csv(
  sampled_with_offline_clicks,
  file.path(audit_dir, "sampled_hdbscan_clicks_joined_to_offline_clicks.csv")
)

write_csv(
  join_summary,
  file.path(audit_dir, "sampled_to_offline_clicks_join_summary.csv")
)

write_csv(
  join_summary_by_event,
  file.path(audit_dir, "sampled_to_offline_clicks_join_summary_by_event.csv")
)

write_csv(
  sampled_binary_files,
  file.path(audit_dir, "sampled_clicks_relevant_binary_files.csv")
)

cat("\nSaved outputs to:\n")
cat(audit_dir, "\n")

```

## Locate relevant PAMGuard binary files on disk

```{r}

library(dplyr)
library(readr)
library(stringr)
library(tibble)
library(purrr)

audit_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/temporal_visualizations/bearing_source_audit"

binary_root <- "E:/Analysis/Binaries"

sampled_binary_files <- read_csv(
  file.path(audit_dir, "sampled_clicks_relevant_binary_files.csv"),
  show_col_types = FALSE
)

# ------------------------------------------------------------
# Find all binary-like files under the binary root
# ------------------------------------------------------------

all_binary_paths <- list.files(
  binary_root,
  recursive = TRUE,
  full.names = TRUE
)

binary_inventory <- tibble(
  full_path = all_binary_paths,
  filename = basename(all_binary_paths),
  extension = tolower(tools::file_ext(all_binary_paths)),
  size_mb = round(file.info(all_binary_paths)$size / 1024^2, 3),
  modified = file.info(all_binary_paths)$mtime
)

# ------------------------------------------------------------
# Match required .pgdf files
# ------------------------------------------------------------

matched_pgdf <- sampled_binary_files %>%
  left_join(
    binary_inventory,
    by = c("BinaryFile" = "filename")
  ) %>%
  mutate(
    pgdf_found = !is.na(full_path),
    expected_pgdx = str_replace(BinaryFile, "\\.pgdf$", ".pgdx")
  )

# ------------------------------------------------------------
# Match expected .pgdx sidecar files
# ------------------------------------------------------------

pgdx_lookup <- binary_inventory %>%
  filter(extension == "pgdx") %>%
  select(
    expected_pgdx = filename,
    pgdx_full_path = full_path,
    pgdx_size_mb = size_mb
  )

matched_binary_pairs <- matched_pgdf %>%
  left_join(
    pgdx_lookup,
    by = "expected_pgdx"
  ) %>%
  mutate(
    pgdx_found = !is.na(pgdx_full_path)
  )

# ------------------------------------------------------------
# Summaries
# ------------------------------------------------------------

binary_match_summary <- matched_binary_pairs %>%
  summarise(
    n_binary_file_rows = n(),
    n_unique_pgdf_needed = n_distinct(BinaryFile),
    n_pgdf_found = n_distinct(BinaryFile[pgdf_found]),
    n_pgdf_missing = n_distinct(BinaryFile[!pgdf_found]),
    n_pgdx_found = n_distinct(expected_pgdx[pgdx_found]),
    n_pgdx_missing = n_distinct(expected_pgdx[!pgdx_found])
  )

cat("\nBinary match summary:\n")
print(binary_match_summary)

cat("\nMissing .pgdf files:\n")
print(
  matched_binary_pairs %>%
    filter(!pgdf_found) %>%
    distinct(DriftID_clean, BinaryFile),
  n = Inf
)

cat("\nMissing .pgdx files:\n")
print(
  matched_binary_pairs %>%
    filter(!pgdx_found) %>%
    distinct(DriftID_clean, expected_pgdx),
  n = Inf
)

# ------------------------------------------------------------
# Save outputs
# ------------------------------------------------------------

write_csv(
  matched_binary_pairs,
  file.path(audit_dir, "sampled_clicks_relevant_binary_files_matched_to_disk.csv")
)

write_csv(
  binary_match_summary,
  file.path(audit_dir, "sampled_clicks_binary_file_match_summary.csv")
)

cat("\nSaved binary matching outputs to:\n")
cat(audit_dir, "\n")

```

## Create a compact manifest of relevant PAMGuard binary files

```{r}

library(dplyr)
library(readr)

audit_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/temporal_visualizations/bearing_source_audit"

matched_binary_pairs <- read_csv(
  file.path(audit_dir, "sampled_clicks_relevant_binary_files_matched_to_disk.csv"),
  show_col_types = FALSE
)

binary_manifest <- matched_binary_pairs %>%
  filter(pgdf_found, pgdx_found) %>%
  distinct(
    DriftID_clean,
    database_name,
    BinaryFile,
    full_path,
    expected_pgdx,
    pgdx_full_path
  ) %>%
  arrange(DriftID_clean, BinaryFile)

write_csv(
  binary_manifest,
  file.path(audit_dir, "PAMGuard_binary_manifest_for_bearing_export.csv")
)

print(binary_manifest, n = Inf)

cat("\nSaved manifest to:\n")
cat(file.path(audit_dir, "PAMGuard_binary_manifest_for_bearing_export.csv"), "\n")

---------------------------------------------

library(dplyr)
library(readr)

audit_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/temporal_visualizations/bearing_source_audit"

manifest <- read_csv(
  file.path(audit_dir, "PAMGuard_binary_manifest_for_bearing_export.csv"),
  show_col_types = FALSE
)

manifest %>%
  count(BinaryFile, name = "n_manifest_rows") %>%
  filter(n_manifest_rows > 1) %>%
  arrange(desc(n_manifest_rows)) %>%
  print(n = Inf)

manifest %>%
  group_by(BinaryFile) %>%
  filter(n() > 1) %>%
  select(DriftID_clean, BinaryFile, full_path, pgdx_full_path) %>%
  arrange(BinaryFile, DriftID_clean, full_path) %>%
  print(n = Inf)

manifest %>%
  group_by(BinaryFile) %>%
  filter(n() > 1) %>%
  select(
    DriftID_clean,
    BinaryFile,
    full_path,
    pgdx_full_path
  ) %>%
  arrange(BinaryFile, DriftID_clean, full_path, pgdx_full_path) %>%
  print(n = Inf)

```

## Clean PAMGuard binary manifest for bearing export
## Keep only rows where BOTH the .pgdf and .pgdx paths match the DriftID_clean folder

```{r}

library(dplyr)
library(readr)
library(stringr)

audit_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/temporal_visualizations/bearing_source_audit"

manifest <- read_csv(
  file.path(audit_dir, "PAMGuard_binary_manifest_for_bearing_export.csv"),
  show_col_types = FALSE
)

manifest_clean <- manifest %>%
  mutate(
    full_path_norm = str_replace_all(full_path, "\\\\", "/"),
    pgdx_full_path_norm = str_replace_all(pgdx_full_path, "\\\\", "/"),
    drift_folder_pattern = paste0("/", DriftID_clean, "/"),
    pgdf_path_matches_drift = str_detect(full_path_norm, fixed(drift_folder_pattern)),
    pgdx_path_matches_drift = str_detect(pgdx_full_path_norm, fixed(drift_folder_pattern))
  ) %>%
  filter(
    pgdf_path_matches_drift,
    pgdx_path_matches_drift
  ) %>%
  distinct(
    DriftID_clean,
    database_name,
    BinaryFile,
    full_path,
    expected_pgdx,
    pgdx_full_path
  ) %>%
  arrange(DriftID_clean, BinaryFile)

manifest_clean_summary <- manifest_clean %>%
  summarise(
    n_rows = n(),
    n_unique_drifts = n_distinct(DriftID_clean),
    n_unique_pgdf_files = n_distinct(BinaryFile),
    n_unique_pgdf_paths = n_distinct(full_path),
    n_unique_pgdx_paths = n_distinct(pgdx_full_path)
  )

print(manifest_clean_summary)

write_csv(
  manifest_clean,
  file.path(audit_dir, "PAMGuard_binary_manifest_for_bearing_export_CLEAN.csv")
)

cat("\nSaved cleaned manifest to:\n")
cat(file.path(audit_dir, "PAMGuard_binary_manifest_for_bearing_export_CLEAN.csv"), "\n")

```

## Smoke test: inspect one PAMGuard binary file for bearing/localization fields

```{r}

library(dplyr)
library(readr)
library(stringr)

# Install/load PamBinaries
# This reads PAMGuard .pgdf binary files into R.
if (!requireNamespace("PamBinaries", quietly = TRUE)) {
  install.packages("PamBinaries")
}

library(PamBinaries)

audit_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/temporal_visualizations/bearing_source_audit"

manifest_clean <- read_csv(
  file.path(audit_dir, "PAMGuard_binary_manifest_for_bearing_export_CLEAN.csv"),
  show_col_types = FALSE
)

# Pick one binary file to inspect
test_pgdf <- manifest_clean$full_path[1]

cat("Testing binary file:\n")
cat(test_pgdf, "\n\n")

# Load the PAMGuard binary
# skipLarge = TRUE avoids pulling huge waveform-like objects if possible.
pg <- PamBinaries::loadPamguardBinaryFile(
  fileName = test_pgdf,
  skipLarge = TRUE,
  convertDate = TRUE
)

# Convert to data frame
pg_df <- as.data.frame(pg)

cat("Rows read:", nrow(pg_df), "\n")
cat("Columns read:", ncol(pg_df), "\n\n")

cat("All column names:\n")
print(names(pg_df))

# Look for likely useful fields
candidate_cols <- names(pg_df)[
  str_detect(
    names(pg_df),
    regex("uid|time|utc|millis|binary|bear|angle|theta|azimuth|azi|loc|channel|amp", ignore_case = TRUE)
  )
]

cat("\nCandidate UID/time/bearing/localization columns:\n")
print(candidate_cols)

cat("\nPreview of candidate columns:\n")
print(
  pg_df %>%
    select(any_of(candidate_cols)) %>%
    head(20)
)

```

## Extract PAMGuard angles from relevant binaries and join to sampled HDBSCAN clicks

```{r}

library(dplyr)
library(readr)
library(purrr)
library(stringr)
library(tibble)
library(ggplot2)
library(PamBinaries)

# ------------------------------------------------------------
# Paths
# ------------------------------------------------------------

audit_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/temporal_visualizations/bearing_source_audit"

plot_dir <- "E:/Analysis/PCA_ICI_UMAP_HDBSCAN/18PC/18PC_scaledICI_1/plots/temporal_cluster_plots"

dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)

manifest_file <- file.path(
  audit_dir,
  "PAMGuard_binary_manifest_for_bearing_export_CLEAN.csv"
)

sampled_joined_file <- file.path(
  audit_dir,
  "sampled_hdbscan_clicks_joined_to_offline_clicks.csv"
)

# ------------------------------------------------------------
# Load sampled HDBSCAN clicks already joined to OfflineClicks
# This file has HDBSCAN_cluster + EventID + UID + BinaryFile
# ------------------------------------------------------------

sampled_clicks <- read_csv(sampled_joined_file, show_col_types = FALSE) %>%
  mutate(
    UID = as.character(UID),
    BinaryFile = as.character(BinaryFile),
    DriftID_clean = as.character(DriftID_clean),
    HDBSCAN_cluster = factor(
      as.character(HDBSCAN_cluster),
      levels = as.character(0:9)
    )
  )

needed_click_keys <- sampled_clicks %>%
  distinct(DriftID_clean, BinaryFile, UID)

cat("Sampled clicks needing bearing:", nrow(needed_click_keys), "\n")

# ------------------------------------------------------------
# Load clean binary manifest
# ------------------------------------------------------------

manifest_clean <- read_csv(manifest_file, show_col_types = FALSE) %>%
  mutate(
    DriftID_clean = as.character(DriftID_clean),
    BinaryFile = as.character(BinaryFile)
  )

cat("Relevant binary files:", nrow(manifest_clean), "\n")

# ------------------------------------------------------------
# Helper function: read one PAMGuard .pgdf and keep angle fields
# ------------------------------------------------------------

read_pgdf_angles <- function(pgdf_path, DriftID_clean, BinaryFile) {
  
  cat("Reading:", DriftID_clean, BinaryFile, "\n")
  
  out <- tryCatch({
    
    pg <- PamBinaries::loadPamguardBinaryFile(
      fileName = pgdf_path,
      skipLarge = TRUE,
      convertDate = TRUE
    )
    
    pg_df <- as.data.frame(pg) %>%
      as_tibble()
    
    # Keep only fields we need
    pg_df %>%
      transmute(
        DriftID_clean = as.character(DriftID_clean),
        BinaryFile = as.character(BinaryFile),
        UID = as.character(UID),
        binary_millis = as.numeric(millis),
        binary_date = as.character(date),
        time_delay_s = as.numeric(timeDelays),
        pamguard_angle_rad = as.numeric(angles),
        pamguard_angle_error_rad = as.numeric(angleErrors),
        pamguard_angle_deg = pamguard_angle_rad * 180 / pi,
        pamguard_angle_error_deg = pamguard_angle_error_rad * 180 / pi,
        maxAmplitude = as.numeric(maxAmplitude),
        channelMap = as.integer(channelMap),
        nChan = as.integer(nChan)
      )
    
  }, error = function(e) {
    
    warning("Failed to read: ", pgdf_path, "\n", e$message)
    
    tibble(
      DriftID_clean = character(),
      BinaryFile = character(),
      UID = character(),
      binary_millis = numeric(),
      binary_date = character(),
      time_delay_s = numeric(),
      pamguard_angle_rad = numeric(),
      pamguard_angle_error_rad = numeric(),
      pamguard_angle_deg = numeric(),
      pamguard_angle_error_deg = numeric(),
      maxAmplitude = numeric(),
      channelMap = integer(),
      nChan = integer()
    )
  })
  
  return(out)
}

# ------------------------------------------------------------
# Extract angles from all relevant binaries
# ------------------------------------------------------------
binary_angles_all <- pmap_dfr(
  manifest_clean %>%
    transmute(
      pgdf_path = full_path,
      DriftID_clean = DriftID_clean,
      BinaryFile = BinaryFile
    ),
  read_pgdf_angles
)

cat("\nRows read from binaries:", nrow(binary_angles_all), "\n")

# ------------------------------------------------------------
# Keep only the sampled clicks
# ------------------------------------------------------------

binary_angles_sampled <- binary_angles_all %>%
  semi_join(
    needed_click_keys,
    by = c("DriftID_clean", "BinaryFile", "UID")
  )

cat("Binary angle rows matching sampled clicks:", nrow(binary_angles_sampled), "\n")

# ------------------------------------------------------------
# Join angle fields back to sampled HDBSCAN table
# ------------------------------------------------------------

sampled_with_bearing <- sampled_clicks %>%
  left_join(
    binary_angles_sampled,
    by = c("DriftID_clean", "BinaryFile", "UID")
  )

bearing_join_summary <- sampled_with_bearing %>%
  summarise(
    n_sampled_clicks = n(),
    n_with_angle = sum(!is.na(pamguard_angle_deg)),
    n_missing_angle = sum(is.na(pamguard_angle_deg)),
    percent_with_angle = 100 * n_with_angle / n_sampled_clicks,
    median_angle_error_deg = median(pamguard_angle_error_deg, na.rm = TRUE),
    mean_angle_error_deg = mean(pamguard_angle_error_deg, na.rm = TRUE)
  )

print(bearing_join_summary)

# ------------------------------------------------------------
# Save joined bearing table
# ------------------------------------------------------------

write_csv(
  sampled_with_bearing,
  file.path(audit_dir, "sampled_hdbscan_clicks_with_pamguard_angles.csv")
)

write_csv(
  bearing_join_summary,
  file.path(audit_dir, "sampled_hdbscan_clicks_bearing_join_summary.csv")
)

cat("\nSaved joined table to:\n")
cat(file.path(audit_dir, "sampled_hdbscan_clicks_with_pamguard_angles.csv"), "\n")

sampled_with_bearing %>%
  summarise(
    n_rows = n(),
    n_unique_events = n_distinct(EventID),
    n_unique_drifts = n_distinct(DriftID_clean),
    n_unique_uids = n_distinct(paste(DriftID_clean, UID)),
    n_with_angle = sum(!is.na(pamguard_angle_deg)),
    n_missing_angle = sum(is.na(pamguard_angle_deg))
  )

sampled_with_bearing %>%
  summarise(
    min_angle_deg = min(pamguard_angle_deg, na.rm = TRUE),
    q1_angle_deg = quantile(pamguard_angle_deg, 0.25, na.rm = TRUE),
    median_angle_deg = median(pamguard_angle_deg, na.rm = TRUE),
    q3_angle_deg = quantile(pamguard_angle_deg, 0.75, na.rm = TRUE),
    max_angle_deg = max(pamguard_angle_deg, na.rm = TRUE),
    median_angle_error_deg = median(pamguard_angle_error_deg, na.rm = TRUE),
    max_angle_error_deg = max(pamguard_angle_error_deg, na.rm = TRUE)
  )

sampled_with_bearing %>%
  group_by(EventID) %>%
  summarise(
    n_clicks = n(),
    min_angle_deg = min(pamguard_angle_deg, na.rm = TRUE),
    median_angle_deg = median(pamguard_angle_deg, na.rm = TRUE),
    max_angle_deg = max(pamguard_angle_deg, na.rm = TRUE),
    median_angle_error_deg = median(pamguard_angle_error_deg, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(median_angle_deg) %>%
  print(n = Inf)

```

## Plotting bearing versus time

```{r}

cluster_colors <- c(
  "0" = "grey80",
  "1" = "#E41A1C",
  "2" = "#FF7F00",
  "3" = "#FFD92F",
  "4" = "#4DAF4A",
  "5" = "#00BFC4",
  "6" = "#377EB8",
  "7" = "#984EA3",
  "8" = "#F781BF",
  "9" = "#A65628"
)

bearing_plot_data <- sampled_with_bearing %>%
  filter(!is.na(pamguard_angle_deg)) %>%
  mutate(
    Final_UTC_ms = as.numeric(Final_UTC_ms),
    HDBSCAN_cluster = factor(
      as.character(HDBSCAN_cluster),
      levels = as.character(0:9)
    )
  ) %>%
  group_by(EventID) %>%
  arrange(Final_UTC_ms, .by_group = TRUE) %>%
  mutate(
    seconds_since_event_start = (Final_UTC_ms - min(Final_UTC_ms, na.rm = TRUE)) / 1000
  ) %>%
  ungroup()

p_bearing_by_event <- ggplot(
  bearing_plot_data,
  aes(
    x = seconds_since_event_start,
    y = pamguard_angle_deg,
    color = HDBSCAN_cluster
  )
) +
  geom_point(
    size = 0.8,
    alpha = 0.75
  ) +
  facet_wrap(~ EventID, scales = "free_x") +
  scale_color_manual(values = cluster_colors, drop = FALSE) +
  labs(
    title = "Temporal PAMGuard Click Angle by Event",
    subtitle = "Sampled clicks only; colored by HDBSCAN cluster",
    x = "Seconds since event start",
    y = "PAMGuard click angle (degrees)",
    color = "Cluster"
  ) +
  theme_bw(base_size = 12) +
  theme(
    strip.text = element_text(size = 8, face = "bold"),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

print(p_bearing_by_event)

ggsave(
  filename = file.path(
    plot_dir,
    "temporal_pamguard_angle_by_event_colored_by_cluster.png"
  ),
  plot = p_bearing_by_event,
  width = 16,
  height = 12,
  dpi = 300
)

```

## Plotting bearing versus time excluding noise

```{r}

bearing_plot_no_noise <- sampled_with_bearing %>%
  filter(
    !is.na(pamguard_angle_deg),
    as.character(HDBSCAN_cluster) != "0"
  ) %>%
  mutate(
    Final_UTC_ms = as.numeric(Final_UTC_ms),
    HDBSCAN_cluster = factor(
      as.character(HDBSCAN_cluster),
      levels = as.character(1:9)
    )
  ) %>%
  group_by(EventID) %>%
  arrange(Final_UTC_ms, .by_group = TRUE) %>%
  mutate(
    seconds_since_event_start = (Final_UTC_ms - min(Final_UTC_ms, na.rm = TRUE)) / 1000
  ) %>%
  ungroup()

cluster_colors_no_noise <- c(
  "1" = "#E41A1C",
  "2" = "#FF7F00",
  "3" = "#FFD92F",
  "4" = "#4DAF4A",
  "5" = "#00BFC4",
  "6" = "#377EB8",
  "7" = "#984EA3",
  "8" = "#F781BF",
  "9" = "#A65628"
)

p_bearing_by_event_no_noise <- ggplot(
  bearing_plot_no_noise,
  aes(
    x = seconds_since_event_start,
    y = pamguard_angle_deg,
    color = HDBSCAN_cluster
  )
) +
  geom_point(
    size = 0.8,
    alpha = 0.75
  ) +
  facet_wrap(~ EventID, scales = "free_x") +
  scale_color_manual(values = cluster_colors_no_noise, drop = FALSE) +
  labs(
    title = "Temporal PAMGuard Click Angle by Event (Noise Excluded)",
    subtitle = "Sampled clicks only; HDBSCAN cluster 0 removed",
    x = "Seconds since event start",
    y = "PAMGuard click angle (degrees)",
    color = "Cluster"
  ) +
  theme_bw(base_size = 12) +
  theme(
    strip.text = element_text(size = 8, face = "bold"),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

print(p_bearing_by_event_no_noise)

ggsave(
  filename = file.path(
    plot_dir,
    "temporal_pamguard_angle_by_event_colored_by_cluster_NO_NOISE.png"
  ),
  plot = p_bearing_by_event_no_noise,
  width = 16,
  height = 12,
  dpi = 300
)

```



























